### 2017-03-20 mlppi-comsis17/paper.tex
experiments done

\documentclass[runningheads]{comsis2}

\usepackage[ruled,linesnumbered]{algorithm2e}
%\usepackage{algorithm}
%\usepackage{amssymb,amsthm,amsmath}
\usepackage{amssymb}
\usepackage{balance}
\usepackage{subfigure}
\usepackage{float}
%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{lemma}[theorem]{Lemma}
\usepackage{latexsym}
\usepackage{multirow}
\usepackage{color}
\usepackage{url}
\usepackage{comment}
\usepackage[dvips]{graphicx}          % graphicx package for including ps files

%% Necessary definitions for the running heads
\def\journalissue{Computer Science and Information Systems ?(?):??--??}
\def\paperidnum{DOI: N/A}
\setcounter{page}{1}

\title{MLPPI Wizard: An Automated Multi-level Partitioning Tool on Analytical Workloads}
%\titlenote{An earlier version of this article was presented as a short paper at ACM CIKM'12.}
%% Use this if the title is too long for the running heads
%\titlerunning{An Automated Multi-level Partitioning Tool on Analytical Workloads}

\author{Young-Kyoon Suh\inst{1} \and Alain Crolotte\inst{2} \and Pekka Kostamaa\inst{2}}
%Correspondence to Young-Kyoon Suh, now at Korea Institute of Science and Technology Information
%% Use this the list of authors is too long for the running heads
%\authorrunning{Correspondence to Young-Kyoon Suh, now at Korea Institute of Science and Technology Information}

\institute{Department of Computer Science, The University of Arizona, AZ 85721, USA\footnote{Currently, he is at Korea Institute of Science and Technology Information (KISTI). This work was conducted while he interned at Teradata during his PhD in the University of Arizona.}\\ 
  %\email{yksuh@cs.arizona.edu}
  \email{yksuh@cs.arizona.edu}
  \and
  Teradata Corporation, El Segundo, CA 90245, USA\\
  \email{\{alain.crolotte,pekka.kostamaa\}@teradata.com}
}

\def\form#1{$\langle{#1}\rangle$}
\def\range#1{$[{#1}]$}
\def\openrange#1{$({#1})$}
\def\lopenrange#1{$({#1}]$}
\def\ropenrange#1{$[{#1})$}
\def\product#1{$(#1)$}
\definecolor{grey}{RGB}{200,200,200}
\newcommand{\hilite}[1]{\colorbox{grey}{#1}}
\newcommand{\hilitey}[1]{\colorbox{yellow}{#1}}
\newcommand{\hiliting}[1]{\colorbox{grey}{#1}}
\long\def\todo#1{\hilitey{{\bf TODO:} {\em #1}}}
\long\def\shorten#1{}
\long\def\comment#1{}
\def\MAX{\mbox{\sl MAX}}
\def\LIMIT{\mbox{\sl LIMIT}}
\def\getQueryCost{\mbox{\sl getQueryCost}}
\def\getScanCost{\mbox{\sl getScanCost}}
\def\BigO{\mbox{\sl O}}
\def\NULL{\mbox{\sl NULL}}
\def\SolA{\mbox{\sl A}}
\def\SolB{\mbox{\sl B}}
\def\SolC{\mbox{\sl C}}
\def\SolS{\mbox{\sl S}}
\def\Qone{\mbox{\sl q1}}
\def\Qtwo{\mbox{\sl q2}}
\def\Sone{\mbox{\sl s1}}
\def\Stwo{\mbox{\sl s2}}
\def\SolSone{\mbox{\sl S1}}
\def\SolStwo{\mbox{\sl S2}}
\def\SolSthree{\mbox{\sl S3}}
\def\SolSfour{\mbox{\sl S4}}
\def\SolSfive{\mbox{\sl S5}}

\begin{document}

\maketitle

\begin{abstract} 
Typically, it is a daunting task for a database administrator (DBA) to figure out how to partition a huge fact table accessed by query workloads for better performance. To relieve such a burden, we introduce an intelligent partitioning tool to recommend an optimized partitioning on the fact table. This tool uses a greedy algorithm for search space enumeration. This space is driven by predicates of a given query workload. The tool takes advantage of the cost model of a query optimizer to prune the search space. The tool resides completely on a client and interacts with the optimizer via APIs. Thus, there is no overhead to instrument the optimizer code. Furthermore, the predicate-driven method can be applied to any clustering or partitioning scheme. We show that the tool's recommendation outperforms a human expert's solution. We also demonstrate that the recommendation scale very well with increasing workload and growing fact table.

\vspace{6pt}\textbf{Keywords:} Star Schema, Fact Table, Multi-Level Partitioning.
\end{abstract}

\section{Introduction}
\label{sec:intro}

Over the past decades much attention has been paid 
to improve the performance on analytical workloads in a data warehousing environment. 
Typically, a relational database management system (\hbox{DBMS}) 
has been exploited to process such analytical queries faster 
and assist its customers to make a timely business-decision in 
rapidly changing enterprise data warehousing.
Teradata DBMS has been a leading database product in this data warehousing marketplace 
that has been increasingly more competitive than ever.

%\subsection{Background}
%\label{sec:background}

\paragraph{Background:} 
To optimize the performance of analytical processing, 
tables and materialized views in the Teradata DBMS are hash-distributed based 
on a user-specified column or set of columns called {\em primary index}. 
Each virtual processor, a unit of parallelism called {\em AMP} in 
the \hbox{Teradata} \hbox{DBMS}, 
receives a subset of the data and stores it in hash order. 
Users typically choose primary index fields in the DBMS, 
so that the data is evenly distributed among the AMPs. 
The primary index fields are also chosen to reflect join fields 
in workloads to accomplish cheaper local joins that do not require data shuffling.

Partitioned Primary Index (PPI)~\cite{sinclair:ppi} is an optional horizontal 
partitioning scheme applied locally on the data belonging to each AMP. 
In some database products, this type of partitioning is termed {\em clustering}, 
but hereafter we call the term \hbox{{\em partitioning}} only to avoid confusion.
Database \hbox{administrators} (DBAs) usually choose the columns for PPI, 
based on join fields and single table \hbox{predicates} to optimize the queries included 
in the workload. 
The PPI columns are used to physically group data with the same values 
together in contiguous data blocks. 
This grouping enables ``partition elimination'' in scans and joins for performance enhancement.
PPI can be specified as single or multiple levels. 
This type of partitioning scheme is known as 
\hbox{Multi-Level} PPI ({\em MLPPI)}~\cite{klindt09mlppi}.

By allowing non-qualified partitions to be eliminated, MLPPI can considerably 
reduce the amount of data to be scanned to answer a query. 
But a large number of partitions can create significant overhead, 
particularly in joins and table maintenance operations 
such as inserts and deletes, so that the selection of partitions can be usually a balancing act.

Figure~\ref{fig:exam} exemplifies an MLPPI table with a defined partitioning option. 
This example is a subset of the {\tt LINEORDER} table from the Star Schema Benchmark (SSB)~\cite{oneil:ssb}. 
(Note that the full table with modified data types was actually used in our experiments.)

\begin{figure}[t]
\begin{center}
{\small
\begin{tabular}{l}
{\tt CREATE TABLE LINEORDER(} \\
\hspace{0.1in}{\tt LO\_ORDERKEY INTEGER,} \\
\hspace{0.1in}{\tt LO\_QUANTITY INTEGER,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT INTEGER} \\
{\tt )} \\
{\tt PRIMARY INDEX ( LO\_ORDERKEY )} \\
{\tt PARTITION BY (} \\
\hspace{0.1in}{\tt CASE\_N(} \\
\hspace{0.2in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.2in}{\tt NO CASE OR UNKNOWN),} \\
\hspace{0.1in}{\tt CASE\_N(}\\
\hspace{0.2in}{\tt LO\_QUANTITY < 25,} \\
\hspace{0.2in}{\tt LO\_QUANTITY >= 25 AND LO\_QUANTITY <= 30,} \\
\hspace{0.2in}{\tt LO\_QUANTITY > 30 AND LO\_QUANTITY <= 35,} \\
\hspace{0.2in}{\tt NO CASE OR UNKNOWN)} \\
{\tt );} \\
\end{tabular}
}
\end{center}
\vspace{-0.2in}
\caption{An Example of an MLPPI Table in the Teradata DBMS\label{fig:exam}}
\end{figure}

In the definition of {\tt LINEORDER} in the figure, the primary index 
is on {\tt LO\_ORDERKEY}. 
The \hbox{primary} index \hbox{dictates} the AMP on which a row will be located 
while the partitioning of data will be dictated by the values and 
ranges associated with {\tt LO\_DISCOUNT} and {\tt LO\_QUANTITY}. 
For example, {\tt LO\_DISCOUNT} has two ranges: (i)~one for \hbox{values} $\geq$ 7 
and (ii)~another for all the other values (or {\tt NO CASE OR UNKNOWN}). 
{\tt LO\_QUANTITY} has four ranges. 
Hence, the relation will have a total of 2$\times$4 = 8 partitions as follows.

\begin{center}
{\footnotesize
\begin{tabular}{|c|l|}\hline 
Partition Number & Condition \\ \hline
1		& {{\tt LO\_DISCOUNT >= 7 \&\& LO\_QUANTITY < 25}} \\ \hline
2		& {{\tt LO\_DISCOUNT >= 7 \&\& 25 <= LO\_QUANTITY <= 30}} \\ \hline
3		& {{\tt LO\_DISCOUNT >= 7 \&\& 30 < LO\_QUANTITY <= 35}} \\ \hline
4		& {{\tt LO\_DISCOUNT >= 7 \&\& LO\_QUANTITY no case}}\\ \hline
5       & {{\tt LO\_DISCOUNT no case \&\& LO\_QUANTITY < 25}} \\ \hline 
6		& {{\tt LO\_DISCOUNT no case \&\& 25 <= LO\_QUANTITY < 30}} \\\hline								  	        
7		& {{\tt LO\_DISCOUNT no case \&\& 30 < LO\_QUANTITY <= 35}} \\ \hline
8		& {{\tt LO\_DISCOUNT no case \&\& LO\_QUANTITY no case}} \\ \hline
\end{tabular}
}
\end{center}

Whatever partition is chosen, 
the selection must be semantically correct; namely, 
the mapping of rows to partitions must be an injection. 
In other words, the constraints must form a covering of 
the entire range, so that a row will belong to exactly one and one partition. 
The Teradata optimizer then applies \hbox{partition} elimination for queries 
that specify conditions on {\tt LO\_DISCOUNT} and/or {\tt LO\_QUANTITY}. 
For \hbox{example}, only partitions~1 and~2 are needed for the query 
``{\tt SELECT * FROM LINEORDER WHERE} \linebreak {\tt LO\_DISCOUNT >= 7 AND} {\tt LO\_QUANTITY <= 30}.'' 
Similarly, partitions~1 and~5 are sufficient to answer the query 
``{\tt SELECT * FROM LINEORDER WHERE} \linebreak {\tt LO\_QUANTITY < 25}.'' 
The respective sizes of the partitions are a factor of the data distribution. 
To see MLPPI in greater detail, check our references~\cite{klindt09mlppi,sinclair:ppi}.

%\vspace{-.1in}

%Next, using {\tt LINEORDER} defined in the SSB we provide examples of partitioning options that 
%a DBA may define based on a small query set. 
\paragraph{Challenge:}
Consider the below query set $Q$ having two queries $q$1 and $q$2 on the SSB.
\vspace{-.05in}
\begin{center}
{\small
\begin{tabular}{rl}
$q$1:	& {\tt SELECT SUM(l.LO$\_$EXTENDEDPRICE*l.LO\_DISCOUNT)} \\ 
		& {\tt FROM LINEORDER l, DDATE d} \\
		& {\tt WHERE l.LO\_ORDERDATE = d.D\_DATEKEY} \\
		& {\tt AND d.D\_YEAR = `1993'} \\
		& {\tt AND l.LO\_DISCOUNT IN (1, 4, 5)} \\
        & {\tt AND l.LO\_QUANTITY <= 30} \\ 
$q$2:	& {\tt SELECT c.C\_NATION, SUM(l.LO\_REVENUE)} \\ 
		& {\tt FROM CUSTOMER c, LINEORDER l} \\
		& {\tt WHERE l.LO\_CUSTKEY = c.LO\_CUSTKEY} \\
		& {\tt AND c.C\_REGION=`EUROPE'} \\
		& {\tt AND l.LO\_DISCOUNT >= 7} \\
		& {\tt AND l.LO\_QUANTITY >= 25 AND l.LO\_QUANTITY <= 35} \\
		& {\tt GROUP BY c.C\_NATION} \\
		& {\tt ORDER BY revenue desc} \\
\end{tabular}
}
\vspace{-.01in}
\end{center}

A DBA may focus on the {\tt LINEORDER} fact table and 
the predicates involving \linebreak {\tt LINEORDER} fields only. 
He can figure out the following five constraints, identified by the query number 
and the sequence number of predicates in each query: 
\vspace{-.05in}
\begin{center}
\begin{tabular}{cl}
$q$1.1: & {\tt LO\_DISCOUNT IN (1,4,5)} \\ 
$q$1.2: & {\tt LO\_QUANTITY <= 30} \\ 
$q$2.1: & {\tt LO\_DISCOUNT >= 7} \\ 
$q$2.2: & {\tt LO\_QUANTITY >= 25}\\ 
$q$2.3: & {\tt LO\_QUANTITY <= 35}. \\ 
\end{tabular}
\end{center}
\vspace{-.05in}
At this point the DBA needs to take into account options 
based only on two fields and the five constraints. 
There are many other possibilities from a {\em fine-grained} partition set 
to a {\em coarser-grained} definition.

Considering for the time being the column {\tt LO\_DISCOUNT}, 
one solution is to identify each value 
for the {\tt IN} predicate in $q$1.1 
and use $q$2.1 as is. This yields the following partitioning expression 
for {\tt LO\_DISCOUNT}:
\vspace{-.05in}
\begin{center}
{\footnotesize
\begin{tabular}{l}
{\tt CASE\_N(} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 1,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 4,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT = 5,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.1in}{\tt NO CASE OR UNKNOWN} \\
{)}. \\
\end{tabular}
}
\end{center}
This expression minimizes the size of the partitions 
by focusing exactly on the values required to satisfy the constraints 
but creates a large number of small partitions.

Another possibility is to look at the maximum and minimum 
values in the {\tt IN} set and to build an {\tt AND} clause equivalent to 
a between clause yielding the following:
\vspace{-.1in}
\begin{center}
\begin{tabular}{l}
{\tt CASE\_N(} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 1 AND LO\_DISCOUNT <= 5,} \\
\hspace{0.1in}{\tt LO\_DISCOUNT >= 7,} \\
\hspace{0.1in}{\tt NO CASE OR UNKNOWN} \\
{)}. \\
\end{tabular}
\vspace{-.1in}
\end{center}

\noindent The above partitioning solution decreases the number of partitions, 
compared to the previous solution but still focuses sharply on the constraints 
with still relatively small partitions.

%(Another possibility would be to use the partitioning associated with {\tt LO\_DISCOUNT} shown 
%in the beginning of this section with only two partitions.)

Similar considerations apply to the partitioning associated with {\tt LO\_QUANTITY}. 
The resulting partitioning definition for the table includes both {\tt LO\_DISCOUNT}
and \linebreak {\tt LO\_QUANTITY}, and thus the number of possible combinations is the product of 
the potential combinations for each field.

Furthermore, given two queries one partitioning scheme may be favorable for one query but not for the other, or vice versa. 
As a result, the DBA becomes faced with a daunting combinatorial search problem 
and no clear basis to decide on which combination is the best. 
This state of affairs begs for a {\em tool} for the DBA.

\paragraph{Solution:} 
To ease a burden of partitioning decision, 
we introduce a novel physical database design tool, called {\em MLPPI wizard}, 
designed to recommend an effective partitioning solution for a given workload.
%Indeed, the \hbox{wizard} is the first physical database design tool 
%developed at Teradata.
This tool uses a greedy algorithm for search space enumeration 
and is based on a general framework allowing general expressions, 
ranges and case expressions for partition definitions. 
%is particularly well-suited for an MLPPI definition. 
We emphasize that the \hbox{predicate-driven} algorithm used by the tool can be applied 
to {\em any} clustering or partitioning scheme based on simple fields and expressions or complex SQL predicates 
in an {\em arbitrary} DBMS. 
The wizard also borrows the optimizer's cost model 
to prune the search space and reach a final solution.

\begin{figure*}[htp!]
\centering
\includegraphics[scale=0.45]{architecture}
\vspace{-0.1in}
\caption{The MLPPI wizard architecture\label{fig:arch}}
\end{figure*}

Figure~\ref{fig:arch} illustrates how our tool recommends 
the final MLPPI customized for a given workload consisting of 
a set of queries and corresponding weights. 
Until yielding a final partitioning recommendation, 
the wizard passes through a series of phases: {\em Preprocessing}, 
{\em Initial}, and {\em Optimized} Phases. 
%In the pre-processing phase the tool produces the most granular partitioning (MGP) based on the predicates collected from every query.
%In the initial phase the wizard refines the passed MGP 
%by merging a pair of partitions with the least (scan) cost 
%and then yields an initial (feasible) MLPPI based on the refined partitions.
%In the \hbox{optimization} phase, the tool makes a further attempt to optimize the initial MLPPI. 
We elaborate in greater detail on each of the three phases, 
using the running query set $Q$ consisting of $q$1 and $q2$ in the rest of the article. 
As illustrated in Figure~\ref{fig:arch}, the wizard resides completely on the client side as opposed to 
several previous tools~\cite{agrawal04:integrating,Lightstone04:db2auto,nehme2011automated} 
that require optimizer code instrumentation. 
Our tool simply invokes the server's existing APIs used to simplify the queries, capture fact table 
predicates, and estimate processing costs. 
%This makes the wizard extensible and portable to different releases 
%of the Teradata DBMS server. 
%This \hbox{loosely-coupled} interaction 
%between both sides makes the wizard extensible and portable to different releases 
%of the database server. 
To the best of our knowledge, our wizard  
is the very first physical database design tool 
to address such a {\em \hbox{multi-level} partitioning} problem.

\paragraph{Contribution:} This article provides 
the following contributions.

\begin{itemize}
\item We propose a physical database design tool, termed MLPPI wizard, to recommend a multi-level partitioning definition 
for reducing the processing cost of a given query workload on a star schema benchmark. 
\item This tool is totally outside the database server, thereby incurring no overhead of instrumenting query optimizer's code.
\item We empirically show that a partitioning recommendation of our tool outperforms that of human expert 
over increasing workload size and growing fact table.

\end{itemize}

The article is a substantial extension of prior work~\cite{Suh12}. 
The rest of this article is \hbox{organized} in the following way. 
In the following section we propose a \hbox{multi-level} partitioning algorithm---consisting of a series of phases---used by our MLPPI wizard, 
which is applicable to any clustering or partitioning scheme in an arbitrary DBMS. 
In turn, we conduct a detailed analysis of the complexity of the algorithms. 
Next, we report the performance evaluation results. 
We then elaborate on how our wizard is distinguished from several existing tools proposed by other DBMS vendors. 
Finally, we conclude this article by summarizing our discussion.

\section{The MLPPI wizard}
\label{sec:algorithms}

As illustrated in Figure~\ref{fig:arch}, 
the MLPPI wizard has the three phases for a final partitioning recommendation. 
In this section we delve into each phase. For better exposition we use the running 
query set $Q$ \hbox{provided} in Section~\ref{sec:intro}.

\subsection{The Preprocessing Phase}
\label{sec:pre_phase}

The first phase is called {\em Preprocessing} 
to extract query predicates and check any constraint in partitioning. 
This phase consists of a total of six steps, as shown in Algorithm~\ref{algo:pre_phase}. 
%We elaborate on each step in the subsequent section.

%\vspace{-.1in}

\begin{algorithm}	
\caption{The Preprocessing Phase}
\label{algo:pre_phase}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{func}{Function}
\SetKwBlock{begin}{}{end}
{
\Input{$Q$ (input query set) }
\Output{$R$ (non-overlapping range set), $M$ (query-to-range-set map)}
\hspace{.05in} Query Simplification (on $Q$)\\
\hspace{.05in} Range Extraction (from $Q$) \\
\hspace{.05in} Non-Overlapping Range ($R$) Construction \\
\hspace{.05in} Field Count Limit Check (on $R$) \\
\hspace{.05in} Query-to-Range-Set Map ($M$) Construction \\
\hspace{.05in} Partition Count Limit Check (on $R$) \\
}
\end{algorithm}

\subsubsection{Query Simplification}

The first step is to simplify the predicates of queries in a given workload.
In other words, we remove redundant conditions among the queries. 
For instance, if a query predicate has the condition of 
``{\tt LO\_DISCOUNT IN (1, 4, 5) AND LO\_DISCOUNT IN (2, 3, 4, 5)}'', 
then the predicate can be simplified as \linebreak ``\hbox{{\tt LO\_DISCOUNT}} {\tt IN (4, 5)}.'' 
This can be done via an API call to the database server. 
This simplification is not required for the running example.

%\vspace{-.3in}

\subsubsection{Range Extraction}

In the subsequent step we gather the simplified predicates and 
then extract ranges from the collected predicates. 
This task can also be performed through an API call to the server. 
The query predicate is of the form 
\begin{center}
\begin{tabular}{l} 
\form{\tt variable} \form{\tt op} \form{\tt constant},
\end{tabular}
\end{center}
where \form{\tt variable} is a field (column) from {\tt LINEORDER}, 
\form{\tt op} is in 
\begin{center}
\begin{tabular}{l} 
\{{\tt $ = $ }, {\tt $ < $}, {\tt $ <= $}, {\tt $ >= $}, {\tt $ > $}, 
{\tt IN}\},
\end{tabular}
\end{center}
and \form{\tt constant} represents a constant value(s). 
All \form{\tt op}s are self-explanatory. 
In particular, `{\tt IN}' is a list predicate implying `{\tt OR}' operator.
In the running example, 
we can collect a set of predicates $P$ = \{$p$1,$p$2,$p$3,$p$4\} from $Q$, where

%\vspace{-.15in}

\begin{center}
\begin{tabular}{l} 
$p$1: {\tt l.LO\_DISCOUNT IN (1, 4, 5)} \\ 
$p$2: {\tt l.LO\_DISCOUNT >= 7} \\ 
$p$3: {\tt l.LO\_QUANTITY <= 30} \\ 
$p$4: {\tt l.LO\_QUANTITY >= 25} {\tt AND} {\tt l.LO\_QUANTITY <= 35}. \\ 
\end{tabular}
\end{center}

%\vspace{-.15in}

Next, we construct a bi-directional {\em map} between a query and 
associated predicates, so that for each query the corresponding predicate(s) 
can be found directly in the map, and vice versa. 
In the example, we get a map, called $M$1, built between $Q$ and $P$. 
$M$1 gets filled with the following entries:

$M$1:
\begin{center}
\begin{tabular}{ll} 
\form{q1, \{p1, p3\}} \\
\form{q2, \{p2, p4\}}.
\end{tabular}
\end{center}

Now we build a set of distinct ranges for each field referenced by the predicates. 
Specifically, we gather fields referenced by predicates in $P$, extract 
ranges from the predicates, and group the ranges of each of the referenced fields. 
Suppose $I$ to be a list of fact table fields referenced by $P$. 
In the workload $Q$ we add to $I$ two fields, {\tt LO\_DISCOUNT} and {\tt LO\_QUANTITY}, used by $P$. 
The whole range of the gathered predicates can be represented by a kind of two-dimensional array, called $R$.

Each entry under a field in $R$ represents a pair of (start and end) values forming the range. 
When it comes to range representation for an entry, 
``['' and ``]'' are used for closed ranges while ``('' and ``)'' for open ranges. 
Also, infinity ($\infty$) and -infinity (-$\infty$) can be used for unbounded ranges. 
In particular, {\tt IN} predicate is represented by multiple ranges. 
For instance, the above $p$1 can be represented by the following three ranges 
\range{1, 1}, \range{4, 4}, \range{5, 5}; however, the last two consecutive 
ranges can be consolidated as \range{4, 5}. 
In the continuing example the field-wise distinct range set ($R$) is formed by the entries below.

$R$:
\begin{center}
\begin{tabular}{c|c} 
 {\tt LO\_DISCOUNT} & {\tt LO\_QUANTITY} \\ \hline
\range{1, 1} & \lopenrange{\infty, 30}\\
\range{4, 5} & \range{25, 35} \\
\ropenrange{7, \infty} &  \\
\end{tabular}
\end{center}

In turn, we build another map, called $M$2, between $P$ and $R$. 
That is, we associate each predicate in $P$ with a corresponding range in $R$. 
Let $R$[$i$, $j$] denote the interval value for the $i$-th field and $j$-th range in $R$. 
$R$[1, 3], for instance, indicates range \ropenrange{7, \infty} under {\tt LO\_DISCOUNT}. 
In the running example, $M$2 is constructed as follows:

$M$2:
\begin{center}
\begin{tabular}{l}
$\langle${$p$1}, \{$R$[1, 1],$R$[1, 2]\}$\rangle$ \\
$\langle${$p$2}, \{$R$[1 ,3]\}$\rangle$ \\ 
$\langle${$p$3}, \{$R$[2, 1]\}$\rangle$ \\
$\langle${$p$4}, \{$R$[2, 2]\}$\rangle$. \\
\end{tabular}
\end{center}

\subsubsection{Non-Overlapping Range Construction}

Since there may be multiple predicates on the same field across queries, 
extracted ranges for the field may overlap each other. 
The overlapping ranges must be broken without any common portion, in order 
that we can consider a pair of consecutive, \hbox{non-overlapping} ranges for 
a merge in subsequent phases.

\begin{algorithm}[t]
\caption{Splitting Overlapping Ranges}
\label{algo:split_overlap}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{func}{Function}
\SetKwBlock{begin}{}{end}
{
\Input{$R$ having overlapping ranges}
\Output{$R$ with consecutive, non-overlapping ranges}
	\ForEach{field $i$ $\in$ $R$}{	
		$L$ $\leftarrow$ Get the range set of $i$ from $R$ and sort by start value \\
		$j$ = 0 \;
		\While{$j$ $<$ $|L|$-1}{
			$r_{j}$, $r_{j+1}$ $\leftarrow$ Adjacent ranges in $L$ \\
			\If{$r_{j}$.$end$ $\geq$ $r_{j+1}$.$start$}{
				Make a split by modifying values of $r_{j}$ and $r_{j+1}$. \\
				Insert into $L$ an intermediate range if any. \\
				Re-sort ranges in $L$ and reset $j$ to 0. \\
			}\lElse{$j$++;}
		}
		Update $R$ with the final $L$. 
	} 
}
\end{algorithm}

Algorithm~\ref{algo:split_overlap} describes the way of breaking 
the \hbox{overlap} of ranges. 
For each field, the algorithm 
gets associated ranges of each field and sorts them by start value. 
We then check whether or not adjacent ranges overlap each other.
If that is the case, we make a split between the ranges. 
In the example, the split is applied on the 
overlapping ranges, \lopenrange{\infty, 30} and \range{25, 35} in {\tt LO\_QUANTITY}. 
Subsequently, a common (or intermediate) range (\range{25, 30}) is created and inserted into the range set 
belonging to the {\tt LO\_QUANTITY} field, in order to fill the gap. 
$R$ is shown in the following.

{\it $R$}:
\begin{center}
\begin{tabular}{c|c} 
 {\tt LO\_DISCOUNT} & {\tt LO\_QUANTITY} \\ \hline
\range{1, 1}      & \openrange{\infty, 25} \\
\range{4, 5}      & \range{25, 30} \\
\ropenrange{7, \infty} & \range{31, 35} \\
\end{tabular}
\end{center}

The update of $R$ affects the existing map $M$2. 
In the running example, $p$3 and $p$4 influenced by the split get mapped to 
new range sets, \{$R$[2, 1], $R$[2, 2]\} and \{$R$[2, 2], $R$[2, 3]\}, respectively. 
Of course, the range sets corresponding to $p$1 and $p$2 referencing {\tt LO\_DISCOUNT} remain unchanged. 
As a result, in the running example we obtain the updated $M$2 in the following.

\pagebreak
$M$2:
\begin{center}
\begin{tabular}{l}
$\langle${$p$1}, \{$R$[1, 1],$R$[1, 2]\}$\rangle$ \\
$\langle${$p$2}, \{$R$[1, 3]\}$\rangle$ \\ 
$\langle${$p$3}, \{$R$[2, 1], $R$[2, 2]\}$\rangle$ \\
$\langle${$p$4}, \{$R$[2, 2], $R$[2, 3]\}$\rangle$. \\
\end{tabular}
\end{center}

\subsubsection{Field Count Limit Check}

One question that occurs is what if a DBMS has 
a limit of fields that can be used for a partitioning definition. 
(In the case of Teradata DBMS, up to 64 fields can be allowed for 
an MLPPI definition.)
If the number of fields present in $R$ exceeds the field count limit, 
we determine which field(s) should be thrown away to satisfy the limit. 
(The star schema fact table we use has much fewer fields than the limit, 
and thus, this step will not be executed.)
To choose victim fields, our tool computes the weighted sum of 
{\em query cost}, to be discussed in Section~\ref{sec:opt_phase}, 
of queries regarding each field. 
The sum can be obtained by adding up every query cost on an MLPPI using 
only the ranges under the field. 
We incrementally discard a field with the largest sum of query cost 
until the limit is reached. 
Correspondingly, we can update the existing $M$2. 
One might say that such an MLPPI may not be exploited if too many ranges 
over a partition count limit (65,536 in Teradata DBMS) are found in one field.
But we assume that such an extreme case is not expected. Even if 
such a case happens, the merges of consecutive ranges can make the MLPPI feasible.

\subsubsection{Query-to-Range-Set Map Construction}

Using the existing maps $M$1 and $M$2, 
the tool can create the final bi-directional {\em query-to-range-set} map, say $M$, 
between $Q$ and $R$. 
To construct $M$, we leverage a transitive property from $M$1 to $M$2.
Once we compute $M$, there is no need to keep 
the intermediate maps ($M$1 and $M$2) for the subsequent phases. 
In the running example, we can build $M$ in the following.

\pagebreak

{\it $M$}:
\begin{center}
\begin{tabular}{ll} 
$\langle${$q$1}, \{$R$[1, 1], $R$[1, 2], $R$[2, 1], $R$[2, 2]\}$\rangle$ \\
$\langle${$q$2}, \{$R$[1, 3], $R$[2, 2], $R$[2, 3]\}$\rangle$.
\end{tabular}
\end{center} 
\vspace{-.1in}
\noindent In the subsequent phases $M$ gets used for computing costs and updated 
along with a merge of ranges, and $R$ is refined for an MLPPI recommendation.

\subsubsection{Partition Count Limit Check}

The range set $R$ can be used to define an MLPPI using each field as one level. 
However, if the current number of partitions by $R$ is greater than 
the aforementioned partition count limit, 
then it is not possible to make a feasible MLPPI based on the ranges in $R$.

The actual partition count limit is ample enough to pass the running example, 
but to continue our discussion, assume that in our discussion the limit is {\it 15}.
From $R$ in the running example, we obtain a total of (number of ranges in {\tt LO\_DISCOUNT}) 
$\cdot$(number of ranges in {\tt LO\_QUANTITY}) = (3+1) $\cdot$ (3+1) = {\it 16}
partitions. Note here that one reserved range covering the rest but the identified 
is added by default to each field in the calculation, 
and the default range is equivalent to the ``{\tt NO CASE OR UNKNOWN)}'' case in Figure~\ref{fig:exam}. 
Since the total partitions surpass the limit, 
the tool should go through the initial phase in which 
we make fewer partitions than the limit. 
(Otherwise, the wizard can bypass the initial phase and then immediately proceed to the optimized phase.)

The tool is now ready to proceed to the next phases, 
with $R$ (input range set, or MGP) and the prepared $M$ (query-to-range-set map). 
%In the following subsections, we describe how the MLPPI solution for $Q$ 
%can be eventually made using $M$ and $R$.

\subsection{The Initial Phase}
\label{sec:init_phase}

This phase incrementally merges a range pair in $R$ to reduce partitions. 
The merge continues until the number of ongoing partitions 
drops below the partition count limit. 
After finishing the merge, the tool can derive a feasible MLPPI with the remaining partitions.

Overall, I/O cost may be increased by the merge, 
in that the full scan on a partition is performed to retrieve all the rows potentially matching a given query. 
In the case of a merged partition, we may read the non-qualifying rows 
that would not be seen before the merge, thereby paying more I/O to answer 
the query. To minimize the merge overhead, we pick up the range pair incurring 
the least I/O increase in a heuristic fashion. 
To choose a range pair for a merge, 
we calculate the {\em scan cost} of a query on the merged range pair.

Algorithm~\ref{algo:comp_scan_cost} represents how 
the scan cost can be computed on the range pair that \hbox{influences} a given query. 
The scan cost represents the I/O cost to answer the query when the range pair is merged. 
It is defined as the {\em number of blocks} that are to be read 
for the given query. To compute the scan cost of a query $q$, we write 
a corresponding scan cost query ($s$), which can be constructed as follows: 
\vspace{-.1in}
\begin{center}
$s$: {\tt SELECT * FROM $F$ WHERE} {\it CP},
\end{center} 
\vspace{-.1in}
where $F$ is a fact table, and 
{\it CP} indicates a predicate(s) reconstructed from 
a range(s) mapped to $q$ in $M$.

\begin{algorithm}[t]
\caption{Scan Cost Computation}
\label{algo:comp_scan_cost}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{func}{Function}
\SetKwBlock{begin}{}{end}
{
\Input{$q$ (query), $rp$ (range pair), $M$ (query-to-range-set map)}
\Output{Scan cost for answering $q$ when considering a merge of $rp$}
	$r_{m}$ $\leftarrow$ Consolidate $rp$ \;
	$L$ $\leftarrow$ Copy the range set mapped to $q$ from $M$\; 
	Delete ranges in $rp$ from and insert $r_{m}$ into $L$. \\
	$s$ $\leftarrow$ ``{\tt SELECT * FROM FACT$\_$TABLE WHERE }'' \;
	\ForEach{range $r$ $\in$ $L$}{	
		Reconstruct a predicate $p$ from $r$. \\
		Add $p$ to {\tt WHERE} clause of $s$.
	}
	Make an API call with $s$ to the server. \\
	Extract the spool size (in bytes) from the result. \\
	$blc$ $\leftarrow$ Calculate the block counts from the spool size. \\
	Update $q$'s scan cost to $blc$. \\
	{\bf return} $blc$ \;
}
\end{algorithm}

If $q$ turns out being affected by the merge of a range pair, 
then the range set associated with $q$ in $M$ is temporarily updated 
by removing the parent ranges and adding the merged range. 
The altered range set is remapped to $q$ in $M$ and then 
translated to the equivalent predicates. Then {\it CP} for $s$ can be built based on the predicates.
Unless the merge range pair influences $q$, 
we can simply derive {\it CP} from the existing ranges mapped to $q$.

To see in the Algorithm~\ref{algo:comp_scan_cost} how a scan cost query is built, 
consider a range pair ($rp$) of \hbox{$R$[2, 2]} and $R$[2, 3] in the running example. 
$rp$ produces the merged range, \range{25, 35}, \hbox{under} {\tt LO\_QUANTITY}. 
The merge by $rp$ affects both $q$1 and $q$2 in the workload. 
Thus, the range sets of $q$1 and $q$2 on {\tt LO\_QUANTITY} are 
correspondingly altered to \{\openrange{\infty, 25}, \range{25, 35}\} and \{\range{25, 35}\}, 
respectively. Of course, no change is made to the existing range sets 
of the queries on {\tt LO\_DISCOUNT}. 
Therefore, the corresponding scan cost queries for $q$1 and $q$2 
can be built as shown below:
\vspace{-.05in}
\begin{center}
\begin{tabular}{rl}
$\Sone$:	& {\tt SELECT * FROM LINEORDER l} \\
		& {\tt WHERE ((l.LO\_DISCOUNT = 1) OR}\\
		& {\tt (l.LO\_DISCOUNT >= 4 AND l.LO\_DISCOUNT <= 5))} \\
		& {\tt AND l.LO\_QUANTITY <= 35}  \\ 
$\Stwo$:	& {\tt SELECT * FROM LINEORDER l} \\
        & {\tt WHERE l.LO\_DISCOUNT >= 7} \\
        & {\tt AND (l.LO\_QUANTITY >= 25 AND l.LO\_QUANTITY <= 35)}.
\end{tabular}
\end{center} 
\vspace{-.05in}
\noindent Since the altered ranges (\openrange{\infty, 25} and \range{25, 35}) of $q$1 
are consecutive, we can simply build the merged, single predicate ({\tt l.LO\_QUANTITY <= 35}) for $\Sone$.

In turn, the wizard sends each scan cost query to the server through an API call. 
The tool extracts from the server's response the {\em spool size} (in bytes), which is 
the result size of the sent scan cost query, and then 
calculates as scan cost the {\em block counts} ($blc$) using the spool size. 
If the merge of a range pair does not impact on any query in the workload, 
the existing (previously computed) scan cost of queries is reused for the range pair. 
In other words, the wizard only recomputes the scan cost of a query {\em only if} 
the query is affected by the merge of two consecutive ranges. 
The weighted sum of scan cost of queries, $T_{s}$ can be computed 
for each range pair. 
$T_{s}$ can be defined as follows:
\begin{center}
\begin{tabular}{c} 
$T_{s}$ = $\sum_{i=1}^{n}$($sc_{i}${$\cdot$}{$w_{i}$}),
\end{tabular}
\end{center}
where $n$ is the number of queries in $Q$, $sc_{i}$ is the scan cost of a query $q_{i}$ in $Q$, and 
$w_{i}$ denotes the (\hbox{non-negative}) weight associated with $q_{i}$.

Once all range pairs are examined, the tool chooses the range pair 
with the least $T_{s}$ for a merge. 
If several range pairs end up with the same least $T_{s}$, 
then the tool applies the heuristic of favoring the range pair that produces 
fewer number of partitions when merged, to make 
it faster to reach the partition count limit.

In the example, suppose that the running range pair $rp$ produces the least $T_{s}$, 
and thus, the tool merges $rp$. 
Thus we can update both $R$ and $M$ as follows.

$R$:
\begin{center}
\begin{tabular}{c|c} 
{\tt LO\_DISCOUNT} & {\tt LO\_QUANTITY} \\ \hline
\range{1, 1} & \openrange{\infty, 25} \\
\range{4, 5} & \range{25, 35}		  \\
\ropenrange{7, \infty} &  \\
\end{tabular}
\end{center}

\vspace{-0.1in}
$M$:
\begin{center}
\begin{tabular}{ll} 
$\langle${$q$1}, \{$R$[1, 1], $R$[1, 2], $R$[2, 1], $R$[2, 2]\}$\rangle$ \\
$\langle${$q$2}, \{$R$[1, 3], $R$[2, 2]\}$\rangle$
\end{tabular}
\end{center}

Note that we see that $q$2 may have the most customized partition based on the updated $M$. 
Only the single partition formed by ($R$[1, 3]$\cap${$R$[2, 2]}) is sufficient 
to retrieve all the qualifying 
rows for $q$2; thus, the other partitions can be simply eliminated. 
In the meantime, to answer $q$1, we need to read the four partitions formed by 
the top two ranges of each field. 
As the partitions formed by (({$R$[1, 1]{$\cup$}$R$[1, 2]})$\cap$($R$[2, 2]) 
contain the \hbox{non-qualifying} rows for $q$1, unfortunately, 
$R$ cannot provide $q$1 with as much benefit as $q$2. 
$R$, however, can be a good compromise to satisfy both queries, in that 
the MLPPI derived by $R$ can potentially minimize the total execution cost of $Q$.

Algorithm~\ref{algo:init_phase} represents the initial phase 
that has been described so far. 
Given a \linebreak \hbox{query-to-range} set map ($M$) and an input range set ($R$), 
the algorithm determines which range pair to yield the least total execution cost ($T_{s}$) 
and updates $M$ and $R$ using the chosen range pair, 
until the number of the partitions by $R$ falls below a pre-defined partition limit count.
Finally, the algorithm produces the final recommendation based on $R$.

\begin{algorithm}[t]
\caption{The Initial Phase}
\label{algo:init_phase}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{func}{Function}
\SetKwBlock{begin}{}{end}
{
\Input{$M$ (query-to-range-set map), $R$ (input range set)}
\Output{$R$ with \# partitions $\leq$ partition limit}
	$W$ $\leftarrow$ Query weights \\
	\While{(\# partitions by $R$ $>$ partition limit)}
	{
		\ForEach{a range pair ($rp$) in $R$}
		{	 
			$T_{s}$ $\leftarrow$ 0 \;
			\ForEach{$q$ in $M$}
			{	
				$w$ $\leftarrow$ Get $q$'s weight from $W$ \\
				$L$ $\leftarrow$ Get the range set mapped to $q$ in $M$ \\
				\uIf(\tcp*[f]{Affected}) {(($rp$ $\cap$ $L$) $\neq$ $\varnothing$)}
				{
					$T_{s}$ += $w{\cdot}$({\it getSC}($q$,$rp$,$M$)) \tcp*[r]{See Algo.~\ref{algo:comp_scan_cost} regarding {\it getSC}()}
				}
				%\lElse{ $T_{s}$ += $w{\cdot}$($q$'s existing scan cost)}
			}
			Associate $T_{s}$ with $rp$.
		}
		Find $rp$(s) with the least $T_{s}$. \\
		If a tie happens, select the $rp$ to make fewer partitions when consolidated. \\
		Update $M$ and $R$ by the chosen $rp$.
	}
	{\bf return} $R$\;
}
\end{algorithm}

In the running example the total partition counts ($4${$\times$}$3$=$12$) by $R$ is under 
the assumed limit ($15$), which meets the partitioning limit. 
Hence, the wizard can enter the optimized phase with $R$, 
without repeating the initial phase.

\subsection{The Optimized Phase}
\label{sec:opt_phase}

%We verified the effectiveness of the tool by running it on a workload based
After passing the initial phase, 
we now have an initial MLPPI recommendation based on $R$, for the {\tt LINEORDER} fact table.
The initial recommendation can be used sufficiently.

But we emphasize that additional merges may make fewer partitions, 
which actually can enhance the overall workload performance for the following reasons. 
First, multiple file contexts by many partitions can incur 
huge overhead that impacts on the query optimizer. 
There also exists an operational threshold 
that the optimizer can handle the maximum number of the partitions at a time.
Therefore, further reducing partitions greatly helps the optimizer to manage these partitions.

A similar algorithm, as suggested in the initial phase, 
can be applied. 
In this phase, we use {\em query cost}, analogous to scan cost, 
which can be obtained via a feasible MLPPI. 
Specifically, the query cost is modeled as the estimated processing time 
of a query on a {\it faked} fact table applying the MLPPI definition that reflects the merge of a range pair. 
This cost estimation technique is similar 
to the ``\hbox{what-if}" approach~\cite{chaudhuri1998autoadmin}.

\begin{algorithm}[t]
\caption{Query Cost Computation}
\label{algo:comp_query_cost}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{func}{Function}
\SetKwBlock{begin}{}{end}
{
\Input{$q$ (query), $rp$ (range pair), $M$ (query-to-range-set map), $R$ (input range set)}
\Output{Query cost to answer $q$ when considering a merge of $rp$}
	Create an empty shadow table $H$ having the same definition as the fact table $F$, 
	including indexes and all constraints (check and referential integrity constraints). \\
	Propagate all (field and index) statistics of $F$ to $H$. \\
	\If{$F$ has materialized views (MVs)}{
		Create the equivalent MVs on $H$. \\
		Propagate the statistics of the MVs on $F$ to those of $H$. \\
	}
	$r_{m}$ $\leftarrow$ Merge $rp$ \;
	$L$ $\leftarrow$ Copy ranges associated with $q$ from $M$\; 
	Remove ranges in $rp$ from and add $r_{m}$ to $L$. \\
	$R'$ $\leftarrow$ Update $R$ with $L$ \\
	Alter $H$ using a fictitious MLPPI with $R'$. \\
	Construct and send to the server an $H$-based shadow query $h$ equivalent to $q$. \\	
	$ept$ $\leftarrow$ Estimated processing time of $h$ extracted from the server response \;
	Update $q$'s query cost to $ept$. \\
	{\bf return $ept$} 
}
\end{algorithm}

\begin{algorithm}[t]
\caption{The Optimized Phase}
\label{algo:opt_phase}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{func}{Function}
\SetKwBlock{begin}{}{end}
{
\Input{$M$ (query-to-range-set map), $R$ (input range set)}
\Output{The MLPPI on which the total query cost is minimal}
	$W$ $\leftarrow$ Query weights \\
	$T_{p}$ $\leftarrow$ Total query cost sum on an initial MLPPI by $R$ \\
	\While{(Pre-defined iterations or $R$ $\neq$ $\varnothing$)}{
		\ForEach{range pair $rp$ in $R$}{	 
			$T_{q}$ $\leftarrow$ 0 \;
			\ForEach{$q$ in $M$}{	
				$w$ $\leftarrow$ Get $q$'s weight from $W$ \\
				$L$ $\leftarrow$ Get the range set mapped to $q$ in $M$ \\
				\uIf(\tcp*[f]{Affected}) 
				{(($rp$ $\cap$ $L$) $\neq$ $\varnothing$)}{
					$T_{q}$ += $w${$\cdot$}({\it getQC}($q$,$rp$,$M$,$R$)) 	
					\tcp*[r]{See Algo.~\ref{algo:comp_query_cost} regarding {\it getQC}()}				
				}
				\lElse{ $T_{q}$ += $w${$\cdot$}($q$'s existing query cost)}
			}
			Map $T_{q}$ to $rp$.
		}
		$T$ $\leftarrow$ The least $T_{q}$ that has been so far seen.  \\
		\lIf{$T$ $>$ $T_{p}$}{ {\bf break} \;}
		\Else{
			$T_{p}$ $\leftarrow$ $T$ \tcp*[r]{Query Cost Sum Update} 
			Find $rp$(s) with $T$. \\
			If a tie happens, select the $rp$ to make fewer partitions when consolidated. \\
			Update $M$ and $R$ by the chosen $rp$. \\
		}
	}
	{\bf return} an MLPPI by $R$ \;
}
\end{algorithm}

Algorithm~\ref{algo:comp_query_cost} illustrates the steps for computing the query cost of a query $q$. 
Given a range pair ($rp$) affecting $q$ when merged, 
we first create an empty shadow table $H$ with the same definition as $F$, 
including all indexes and constraints such as check and referential integrity 
ones. Next, we propagate all statistics of $F$ to $H$. 
This covers field and index statistics. 
If $F$ has materialized views (MVs), then we create the equivalent MVs on $H$ and subsequently, 
propagate the statistics of the MVs on $F$ to the new MVs on $H$. 
After that, we alter $H$ by an MLPPI with the updated $R$ ($R$'), applying the merge of $rp$. 
The algorithm then builds a {\em shadow} query $h$, equivalent to the input query $q$,
which is built by replacing $F$ in $q$ with $H$. 
The algorithm in turn makes an API call to the {\tt EXPLAIN} tool with $h$. 
Eventually, the algorithm extracts from the result and returns the estimated elapsed time 
of $h$ as the query cost of $q$. 
As done in the initial phase, the query cost of a query is recomputed only if the merge affects the query.

Algorithm~\ref{algo:opt_phase} proposes the algorithm of the optimized phase. 
$W$ indicates a set of weights assigned to individual queries in a given workload. 
$T_{p}$ keeps track of the existing least query cost sum, and initially is computed on an initial MLPPI. 
For each range pair, the algorithm 
computes the respective weighted sum ($T_{q}$) of query costs of the queries in a given workload. 
($T_{q}$ can be similarly defined as $T_{s}$, described in Section~\ref{sec:init_phase}, 
and thus the definition of $T_{q}$ is omitted here.)
Let the current least $T_{q}$ be $T$. 
If $T$ $<=$ $T_{p}$ (existing least), then for the next iteration $T_{p}$ is updated to $T$. 
We perform the merge on the range pair that produces that $T$ and update $M$ and $R$ along with the merge. 
The algorithm repeats the process until (1)~no more ranges in $R$ are left, or (2)~predefined iterations are reached. 
If $T$ $\geq$ $T_{p}$, then the optimized phase is terminated.

Note that to break a tie between range pairs, 
we apply the same heuristic to favor one of them that produces fewer partitions. 
The order of the different partitioning levels may have some impact on 
the execution time. A range condition on lower levels like the query 
in Section~\ref{sec:intro} 
``{\tt SELECT * FROM LINEORDER WHERE LO\_QUANTITY < 25}'' 
requires scanning non-consecutive partitions. 
This may incur some overhead at run time. 
To reduce this effect, we sort the partition levels based on the number of 
partitions in descending order before making the final recommendation. 
The solution in Figure~\ref{fig:exam} follows this heuristic 
making the two-partition case in the first level 
followed by the four-partition case in the second level. 
The order in Figure~\ref{fig:mlppi} can go \hbox{either} way 
since both levels have the same number of \hbox{partitions}.

In the running example, suppose that in the first round, ranges $R$[1, 1] and $R$[1, 2] are chosen for a merge. 
Then, the wizard produces a merged range \range{1, 5} \hbox{under} {\tt LO\_DISCOUNT} and proceeds to the next round. 
If a range pair selected for the subsequent merge fails to improve $T_{p}$, 
then, the tool exits the optimized phase and \hbox{produces} the final MLPPI recommendation for the given workload $Q$ 
as shown in Figure~\ref{fig:mlppi}.

\vspace{-.25in}
\begin{figure}[h!]
\begin{center}
\begin{tabular}{|l|} \hline
{\tt ALTER TABLE LINEORDER} \\
{\tt MODIFY PRIMARY INDEX} \\
{\tt PARTITION BY(} \\
\hspace{.1in}{\tt CASE\_N(} \\
	\hspace{.2in}{\tt LO\_DISCOUNT $\geq$ 1 AND LO\_DISCOUNT $\leq$ 5,} \\
	\hspace{.2in}{\tt LO\_DISCOUNT $\geq$ 7,} \\ 
	\hspace{.2in}{\tt NO CASE OR UNKNOWN),} \\
\hspace{.1in}{\tt CASE\_N(} \\
	\hspace{.2in}{\tt LO\_QUANTITY $<$ 25,} \\
	\hspace{.2in}{\tt LO\_QUANTITY $\geq$ 25 AND LO\_QUANTITY $\leq$ 35,} \\
	\hspace{.2in}{\tt NO CASE OR UNKNOWN)} \\ 
{\tt );} \\ \hline
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{The Final MLPPI Recommendation for the Given Workload $Q$\label{fig:mlppi}}
\vspace{-.2in}
\end{figure}

\section{Analysis}
\label{sec:analysis}

In this section we analyze the time complexity of 
the proposed algorithms: preprocessing, initial, and optimized phases. 
The metric of this complexity analysis on each phase is the logical running time of the wizard, 
equivalent to the number of API calls made to the database server during that phase.

\subsection{The Preprocessing Phase}

\begin{lemma}
\label{lemma:pre_phase}  
The running time complexity for building a query-to-range map is 
$\BigO$($N${$\cdot$}$C${$\cdot$}$V$), where $N$ is the number of queries 
in a workload, $C$ is the total number of fields in the fact table, 
and $V$ is the max number of values in a field. 
\end{lemma}

{\bf Proof}. Consider the worst case such that given a workload of 
$N$ queries, each query references all the fields of the fact table, and 
each field is associated with at most $V$ values by the predicates of the queries. 
Each query may have a single-value range. 
A range consisting of only one value per field can be mapped to each query. 
Thus, the running time complexity for the map construction 
is $\BigO$($N${$\cdot$}$C${$\cdot$}$V$). \qed

But note that in our experiments the bound of $\BigO$($N${$\cdot$}$C${$\cdot$}$V$) 
was overly pessimistic.

\subsection{The Initial and Optimized Phases}

\begin{lemma}
\label{lemma:init_phase}  
The running time complexity for the initial (or optimized) phase is 
$\BigO$($M^{2}${$\cdot$}$N$), 
where $M$ is the number of range pairs, and $N$ is the number of queries in a workload. 
\end{lemma}

{\bf Proof}. Suppose that the merge of every range pair influences all the queries. 
Every iteration, either the scan or query costs need to be recomputed for each query. 
In the first round we pay the \hbox{re-computation} cost of $M${$\cdot$}$N$ 
and merge a victim range pair. In the subsequent round, $M$-1 range pairs remain, 
so the cost amounts to ($M$-1){$\cdot$}$N$. 
In the worst case we end up consolidating all range pairs, thus having no partition on 
the target table. The total calls made to the server can be added up to

\begin{center}
$M${$\cdot$}$N$+($M$-1){$\cdot$}$N$+$\cdots$+$N$ = ($\sum_{i=1}^M i$){$\cdot$}$N$ = $\frac{M(M-1)}{2}${$\cdot$}$N$. 
\end{center}

\noindent Hence, the running time complexity is $\BigO$($M^{2}${$\cdot$}$N$). $\qed$

%\vspace{0.1in}

Again, our experiments showed that $\BigO$($M^{2}${$\cdot$}$N$), the running time complexity, 
was overly pessimistic. Specifically, 
the number of calls made in our experiments was by far fewer than that bound. 
That was because it was very rare for a range pair to involve all queries in a given workload 
(as seen in Table~\ref{tab:workload_stat}).

%In the following we evaluate the performance of our wizard using the proposed algorithms.

\section{Experiment}
\label{sec:experiment}

In this section we evaluate the performance of our wizard using the proposed algorithms.
We first describe our environment settings. 
We then report the statistics measured during our experiments. 
Finally, we compare the performance of the recommendation 
by the MLPPI wizard (WIZARD) with (1)~that of no partitioning (NO PPI)
and (2)~the partitioning of a human expert (EXP) under different setups.

\subsection{Environment Settings}
\label{sec:env_set}

\paragraph{Development:} 
We implemented our MLPPI wizard as a prototype on top of the \hbox{Teradata} DBMS server. 
The wizard was written in Java. 
We evaluated the performance of the \hbox{wizard} 
on the Teradata DBMS server machine running Unix.

\paragraph{Workload:} When it comes to workload generation, we took advantage of 
our simple star schema query \hbox{generator}. 
The generator assumes that (i)~available operators, (ii)~fields, 
and (iii)~the minimum and \hbox{maximum} values of the fields are already known. 
For each query, the generator first randomly selects the number of single-table 
predicates to create. 
The generator then arbitrarily chooses a field and a specific operator 
for each predicate. 
If {\tt IN} operator on a chosen field is picked up, the query generator 
determines the number of values to add to the {\tt IN} list and then chooses random values within the min-max value 
range of the field. 
In this way the generator makes a query involving the generated predicates.

A generated query is based on a template that joins {\tt LINEORDER} and 
{\tt DDATE} with constraints defined by the predicates. 
This template is common in customer cases like reports and form 
templates. 
Using the query generator we generated two workloads consisting of 10 queries~(10Q) and 20 queries~(20Q).

For the query workloads we populated the fact table with a scale factor of 1TByte~(1TB) and 3TBytes~(3TB).

\begin{table*}[t]
\caption{The statistics obtained on the used workloads\label{tab:workload_stat}}
\begin{center}
{\scriptsize %
\begin{tabular}{|c|l|l|l|l|l|} \hline
\multicolumn{2}{|c|}{\multirow{2}{*}{Item}} & \multicolumn{2}{|c|}{The Initial Phase} & \multicolumn{2}{|c|}{The Optimized Phase} \\ \cline{3-6}
 \multicolumn{2}{|c|}{} & 10Q & 20Q & 10Q & 20Q \\ \hline
1 & {\# of input partitions}  	 & 1,008,000 & 110,739,200 & 51,840 & 59,520 \\ \hline
2 & {\# of input range pairs} 	 & 441 & 4,180 & 24 & 48\\ \hline
 % {\# of total iterations}   & 14 & 55 & 1 & 1 \\ \hline
3 & {\# of total API calls (\# of worst calls by Lemma~\ref{lemma:init_phase}})   & 1,305 (1,944,810) & 31,915 (349,448,000) & 78 (5,760) & 388 (46,080) \\ \hline
4 & {\# of average API calls per range pair} 	& 3 & 8 & 3 & 8 \\ \hline
5 & {\# of average range pairs compared per iteration} 	   & 32 & 76 & 24 & 48 \\ \hline
6 & {The maximum \# of range pairs compared per an iteration} & 38 & 103 & 24 & 48 \\ \hline
7 &{The minimum \# of range pairs compared per an iteration} & 25 & 49 & 24 & 48\\ \hline
%{\# max. API Calls made per range pair} 		   & 4.92 & 16.14 & & \\ \hline
%{\# min. API Calls made per range pair} 		   & 1.92 & & & \\ \hline
8 & {\# of average API calls made per iteration}		   & 93 & 580 & 78 & 388 \\ \hline
9 & {The maximum \# of API calls made per an iteration} 	   & 123 & 791 & 78 & 388 \\ \hline
10 & {The minimum \# of API calls made per an iteration} 	   & 73 & 381 & 78 & 388 \\ \hline
\end{tabular}
}
\end{center}
\vspace{-0.3in}
\end{table*}

\subsection{Statistics Report}
\label{sec:stat}

Table~\ref{tab:workload_stat} summarizes the execution statistics of 
which were obtained while the recommendations for the \hbox{workloads} were produced. 
The statistics includes input partitions, 
range pairs, number of API calls made, and total iterations 
observed in each phase.

About 1 million partitions were initially derived for 10Q, and roughly 110 million partitions for 20Q (Item \#1). 
Although the workload size doubled from 10Q to 20Q, the total number of partitions increased exponentially. 
Much more predicates with different bindings produced about 110x more ranges in 20Q than those of 10Q. 
While a huge number of input partitions were made, the number of range pairs collected per field was relatively small 
(Item \#2).

The total number of iterations in the initial phase 
tended to be proportional to the input partitions 
generated from the 10Q and 20Q workloads. 
Note that the optimized phase ran only once for those two workloads.
Since the partitions produced by the initial phase were customized enough,
no more iterations were necessary for further optimization.

As mentioned in Section~\ref{sec:analysis}, for each workload the number of the total calls 
to the server were extremely small, compared to that of the theoretical worst calls shown in the parentheses (Item \#3). 
The worst call counts were calculated regarding the number of queries and range pairs observed in each phase. 
Furthermore, the average number of calls per range pair (was limited within 10 across the workloads and phases (Item \#4).

Finally,  the per-iteration numbers of the average, minimum, and maximum range pairs proportionally 
increased along with the growing workload size (Items \#5$\sim$\#7), while 
the per-iteration numbers of the average, minimum, and maximum API calls 
increased more than the scale factor (of 2x) (Items \#8$\sim$\#10). 
But empirically the number of invoked API calls were much fewer than that of the theoretical bound.

\begin{figure*}[t]
\vspace{-.2in}
\centering
\subfigure[1TB]{
	\includegraphics[scale=0.59]{1T_total_time}
	\label{fig:1T_eval_res}
}
\subfigure[3TB]{
	\includegraphics[scale=0.59]{3T_total_time}
	\label{fig:3T_eval_res}
}
\vspace{-0.1in}
\caption{Total Elapsed Time\label{fig:total_elapsed_time}}
\end{figure*}

\begin{figure*}[t]
\vspace{-.2in}
\centering
\subfigure[1TB]{
	\includegraphics[scale=0.59]{1T_quality}
	\label{fig:1T_quality}
}
\subfigure[3TB]{
	\includegraphics[scale=0.59]{3T_quality}
	\label{fig:3T_quality}
}
\vspace{-0.1in}
\caption{Recommendation Quality Comparison\label{fig:recomm_qual}}
\vspace{-.1in}
\end{figure*}

\subsection{Performance Evaluation}

The main focus of our experiments is to see the performance (quality) 
of MLPPI recommendations made by the MLPPI wizard (WIZARD), 
compared with the performance of NO PPI (no partitioning) and EXP (partitioning 
by a human expert).

%Our evalution results are exhibited in Figures ~\ref{fig:eval_res} 
%and ~\ref{fig:improvement}.
%Figure~\ref{fig:eval_res} shows the times (in seconds) taken 
%to run the 10Q or 20Q workloads over the 1TB or 3TB fact tables. 
%Based on the run results, Figure~\ref{fig:improvement} displays 
%how much improvement could be achieved by the EXP or the WIZARD 
%recommendations against the NO PPI one.

Figures~\ref{fig:total_elapsed_time} and \ref{fig:recomm_qual} exhibit our evaluation results.
Figure~\ref{fig:total_elapsed_time} shows the times (in seconds) taken 
to run the 10Q and 20Q workloads over the 1TB and 3TB fact tables that we populated 
into the database server.
Based on the run results, Figure~\ref{fig:recomm_qual} shows 
how much improvement was gained by the EXP and WIZARD solutions, compared with the NO PPI solution.

Overall, the WIZARD recommendations were very effective at partitioning 
the fact tables by leveraging the predicates in the workloads. 
On average, the \hbox{total} execution time of the workloads on the WIZARD solutions was 
about 3.5 and 2 times faster than those of NO PPI and EXP, respectively. 
Also, the quality of the \hbox{WIZARD} solutions outperformed that of the EXP solutions; 
the WIZARD solutions yielded an average of about 77\% improvement on the NO PPI solutions.

Furthermore, the WIZARD recommendations 
scaled very well with growing workload size and table size, 
as illustrated in Figure~\ref{fig:recomm_qual}. 
Although we doubled the workload scale from 10Q to 20Q, 
the performance improvement of the WIZARD solution stayed the same in Figure~\ref{fig:1T_quality} 
and got almost negligibly decreased in Figure~\ref{fig:3T_quality}. 
This result showed that our WIZARD \hbox{solution} was scalable over increasing workload size.

We also increased the table scale from 1TByte to 3TBbytes. 
For 10Q the performance improvement (about 77\%) was almost the same over the increasing scales, 
and for 20Q it was decreased from 76\% to 74\%, which was \hbox{almost} negligible.
The WIZARD partitioning \hbox{recommendation} scaled well with the increasing table size.

In sum, our experiment results attest the effectiveness of the proposed algorithms for WIZARD 
and show the superiority of the recommendations of WIZARD.

\shorten{Regarding the scale of 1TB in Figure~\ref{fig:1T_quality}, 
although we doubled the workload scale (from 10Q to 20Q) 
the performance improvement of the WIZARD solution compared with that of NO PPI was not decrease although we doubled the workload scale (from 10Q to 20Q).
In Figure~\ref{fig:3T_quality}, for the 3TB scale the improvement of WIZARD was slightly decreased, but it was negligible.
In sum, the performance of the WIZARD solution was scalable over increasing workload and table scale.

%Also, we scaled up the table from 1TB to 3TB on the fixed workload, but 
%the \hbox{recommendation} quality was not degraded at all. 
These results attest the effectiveness of our algorithms, 
and show the superiority of our recommendations.}

\section{Related Work}
\label{sec:related_work}

%Physical database design~\cite{finkelstein1988physical,labio1997physical,rozen1991framework,Zilio1998} 
Physical database design~\cite{finkelstein1988physical,labio1997physical,rozen1991framework} 
has been discussed in \hbox{academic} research and industrial sectors in the past years.
The major DBMS vendors (e.g., IBM, MS, and Oracle) have driven much of the work. 
Their specific interests have been in automating the physical design for table 
partitioning~\cite{agrawal04:integrating,Lightstone04:db2auto,nehme2011automated,sheet2009oracle,saa2007oracle,rao2002automating}, indexes/materialized views~\cite{Agrawal04:sqlserver,agrawal2000automated,dash2011cophy,Kimura11,itwsqlserver,valentin2000db2,zilio2004recommending}, 
and integration~\cite{Zilio04:db2design}.

%Lightstone's work~\cite{Lightstone04:db2auto} introduces the MDC (Multidimensional 
%Clustering) Advisor that automates the selection of MDC 
%keys customized for a specific workload on DB2 Universal Databases to use MDC structure. 
%DB2 Advisor~\cite{Zilio04:db2design} is an integrated commercial 
%tool that captures a variety of aspects of physical database design 
%by automation in a comprehensive way. 
%Oracle Partitioning Advisor~\cite{sheet2009oracle} for recommending 
%table partitionings is  integrated into 
%the SQL Access Advisor~\cite{saa2007oracle}.

%Since data clustering significantly influences on the performance of 
%multidimensional queries, the MDC advisor helps DBAs that may have 
%nontrivial troubles in finding proper clustering dimensions and 
%granularity. 
%To enumerate and determine possible clustering dimensions, 
%his work relies on {\it what-if} query cost modeling, data sampling and a search algorithm.

%DB2 Advisor~\cite{Zilio04:db2design} is an integrated commercial 
%tool that captures a variety of aspects of physical database design 
%by automation in a comprehensive way. 
%It makes an attempt to cover a variety of design choices-partitioning, 
%MDC, materialized views and indexes. 
%DB2 Advisor leverages the DB2 query optimizer for yielding a recommendation, and 
%takes a hybrid searching approach for considering different features.

%Oracle has a tool, called Partitioning Advisor~\cite{sheet2009oracle}, 
%to make a partitioning recommendation for tables. 
%The tool is now integrated into 
%the SQL Access Advisor~\cite{saa2007oracle} that recommends indexes and materialized 
%views. 
%However, its technical details are not yet released in public. 
%
%Also, MS SQL Server provides some similar, automated tools~\cite{agrawal2000automated,agrawal04:integrating,
%itwsqlserver} for table partitioning, materialized views or indexes fitting to a given workload.
%The optimizer cost model is exploited to acquire estimates for 
%candidate materialized views or indexes. 
%In particular, Agrawal's work~\cite{agrawal04:integrating} presents 
%an approach of integrating vertical and horizontal partitioning schemes 
%into their physical database design, considering scalability.

Some of the tools in IBM DB2~\cite{Lightstone04:db2auto}, 
Oracle~\cite{sheet2009oracle}, and MS SQL Server~\cite{agrawal04:integrating} 
appear to be similar to our MLPPI wizard. 
However, in light of problem scope and approach, the wizard 
is fundamentally different from the existing tools excluding Oracle Partitioning \hbox{Advisor}~\cite{sheet2009oracle} 
providing no published technical details.

DB2 MDC Advisor~\cite{Lightstone04:db2auto} actually tackles 
a different problem of automatically recommending the most well-suited MDC 
keys for a given workload. 
Agrawal's work~\cite{agrawal04:integrating} 
discusses another problem of merging \hbox{single-level} range partitionings on objects such as tables and indexes. 
This article addresses the multi-level \hbox{partitioning} problem. 
Hence, the existing solutions cannot be directly applied to our MLPPI wizard.

Regarding the approach, DB2 MDC Advisor~\cite{Lightstone04:db2auto} uses 
the search space driven by {\em fields}. In contrast, our search 
space is driven by query-predicates, which is superior because 
predicates are more customized and specific to a workload than fields.

Agrawal's horizontal partitioning scheme~\cite{agrawal04:integrating} 
also uses the search space driven by simple range predicates, but 
his technique has a shortcoming as follows. 
His work produces a solution for each individual query and attempt 
to merge the solutions. But this approach cannot reach 
an optimized solution in a global perspective.
Our wizard generates the whole search space upfront and in turn merges partitions, leading to a \hbox{globally-optimized} solution. 
While his work considers a single column only, we deal with multiple fields.

Implementations of previous tools~\cite{agrawal04:integrating,Lightstone04:db2auto,nehme2011automated} \hbox{required} instrumentation for optimizer code.
These instrumentations are needed to facilitate the required 
information for the physical design tools API calls. 
The \hbox{instrumentation} code need to be enhanced and tested for new database 
releases that add complexity and additional cost for software upgrades. 
But our algorithm is based existing APIs supported by an optimizer (like Teradata), 
which requires no code change in the optimizer.
%The Teradata optimizer with a rich set of existing APIs originally coded 
%for system and workload management tools. 
%The APIs are sufficient to avoid the costly optimizer code change.

Also, Nehme's work~\cite{nehme2011automated}, deeply-integrated with 
optimizer, reveals a concern about the quality of the recommendations 
made by some tools, shallowly integrated with optimizer. 
But we observed in our experiments that the quality of 
the wizard's solutions was much superior to that of the base solutions. 
In addition, some might argue that in our loosely-coupled approach 
the cost to invoke the optimizer might be significant. 
But the measured call counts were much fewer than 
the theoretical bound, since most calls were made only when queries were affected by a merge.  
%If the remote communication cost is expensive, the tool can also be run, 
%{\em co-located} with the server.

%The existing work requires their database engine to be extended 
%to support their features. 
%This strong coupling may impose a nontrivial burden on their servers whenever 
%a new feature of the existing tools is desired. 
%As illustrated in Figure~\ref{fig:arch}, our tool is completely outside 
%the DBMS server. Thus, we are free of the constraint. 
%A literature~\cite{nehme2011automated} claims that 
%such loosely-coupled approaches might suffer from the reduced performance of a tool 
%and the poor quality of the partitioning recommendations. 
%The powerful Teradata APIs, however, help the wizard receive from the server 
%everything needed for yielding a MLPPI recommendation. 
%Moreover, as demonstrated in Figure~\ref{fig:perf_eval}, the quality of the 
%partitioning solutions by the wizard was also much superior to that of the base solutions.
%Some might argue that the cost of invoking the optimizer could be significant as well. 
%Again, the actual calls are mainly made only when queries are affected by a merge. 
%As a result, they were much fewer than 
%the theoretical bound as shown in Table~\ref{tab:workload_stat}.  
%To further save the remote communication cost, the tool can also be run, 
%{\em co-located} with the server. 
%Therefore, we believe that the shallowly-integrated strategy perfectly matches 
%with the wizard.

Lastly, there has been some previous work~\cite{nehme2011automated,rao2002automating,tatarowicz2012lookup} 
regarding table partitioning in multi-node \hbox{systems}, but 
our problem is discussed in the context of a single node system, as in the existing work~\cite{agrawal04:integrating}. 
Database cracking~\cite{idreoskm07}
%, self-reorganizing column values in accordance with query workload, 
assumes a single node environment, but 
it does not address our multi-level partitioning problem.

\section{Summary}
\label{sec:conclusion}
%We verified the effectiveness of the tool by running it on a workload based

Given workloads, it is difficult for DBAs to select appropriate fields 
in partitioning the fact table due to large search space. 
DBAs cannot easily determine how granular partitions should be made 
for the workloads.

To address this concern, we presented the MLPPI wizard to recommend a fact-table partitioning for a given star schema workload. 
Since \hbox{query-predicates} are exploited 
to capture necessary ranges for fields and define partitions, 
the MLPPI \hbox{recommendation} can be very optimized, customized to the workload.

We proposed the wizard's algorithms consisting of the three phases. 
The wizard incrementally reduced its search space by merging 
the range pair with the least scan or query cost, and eventually reached an MLPPI recommendation. 
In addition, we analyzed the running complexity for the initial and optimized phases. 
Despite the theoretically high bound, in practice, the wizard made much fewer calls in the phases.

We measured the performance of the recommendation by the wizard using 
our workloads.  
We demonstrated that the produced MLPPI solutions by 
the wizard could reduce the total elapsed time by more than a factor of two, 
compared with those of no partitioning approach and partitioning 
done by a human expert.

%\section*{Acknowledgments} 
%This work was carried out while the first author visited Teradata. 
%The first author is now at Korea Institute of Science and Technology Information (yksuh@kisti.re.kr).

{\footnotesize
\bibliographystyle{splncs03}
\bibliography{paper}
}

\end{document}

### 2017-03-27 mlppi-comsis17/cover-letter.tex
ComSIS paper finalized

% Cover letter using letter.sty
\documentclass{letter} % Uses 10pt
%Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
% the following commands control the margins:
\topmargin=-1in    % Make letterhead start about 1 inch from top of page 
\textheight=8in  % text height can be bigger for a longer letter
\oddsidemargin=0pt % leftmargin is 1 inch
\textwidth=6.5in   % textwidth of 6.5in leaves 1 inch for right margin

\begin{document}

\signature{Young-Kyoon Suh}           % name for signature 
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

\begin{letter}{
%Editors-in-Chief of ACM Transactions on Computer Systems (TOCS): \\
%\begin{tabular}{ll}
%\vspace{.1in}
%
%Prof. Todd C. Mowry\\
%Computer Science Department\\
%Carnegie Mellon University\\
%5000 Forbes Avenue\\
%Pittsburgh, PA 15213 \\
%USA\\ 
}

\begin{flushleft}
{\large\bf Young-Kyoon Suh}
\end{flushleft}
\medskip\hrule height 1pt
%\begin{flushright}
%\hfill Department of Computer Science, The University of Arizona, Tucson, AZ 85721 USA\\
%\hfill Tel: +1-520-370-1865
%(Please understand we cannot reveal our affiliation, as the manuscript should be double-blind.)
%\end{flushright} 
\begin{flushright}
\hfill Supercomputing R\&D Center, 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Korea Institute of Science and Technology Information (KISTI), 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Republic of Korea, 34141, 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Tel: +82-42-869-0725
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Email: yksuh@kisti.re.kr
\end{flushright} 
%\vfill % forces letterhead to top of page

%\vspace{-0.5in}

\opening{Dear Prof. Mirjana Ivanovi\'c (Editor-in-Chief):}

\noindent An earlier version of this submission was published in Proceedings of the 21st ACM \hbox{International} Conference on Information and Knowledge Management (CIKM). 
This version that I now submit to Computer Science and Information Systems (ComSIS) includes the following substantial extensions.

\begin{itemize}

\item Expanded algorithms: we substantially expand the description of the pre-processing, initial and optimized phase algorithms shortened in the earlier version. Furthermore, this submission presents three additional algorithms as follows:

\begin{itemize}

\item Splitting overlapping ranges in Algorithm~2 on page~8,
\item Checking the allowed field count limit on the split ranges on page~8,
\item Computing scan cost in Algorithm~3 on page~10, and
\item Computing query cost in Algorithm~5 on page~13.

\end{itemize}

\item Complexity analysis: we conduct the asymptotic running time analysis on each phase in Section~3 on pages~14-15.

\item Experiments: we provide additional runtime statistics measured from our experiments, as exhibited in Table~1 on page~16. 
In addition, we compare the recommendation quality of our tool with that of a human expert, as shown in Figure~5 on page~18.

\item More thorough literature review: Section~5 on page~18 discusses in greater detail the similarities and differences between existing tools 
and ours.

\end{itemize}

\closing{Sincerely yours,}

\end{letter}

\end{document}

### 2017-03-27 mlppi-comsis17/paper.tex
ComSIS paper finalized

@@
 
-\section{Analysis}
+\section{Complexity Analysis}
 \label{sec:analysis}
@@
 \subfigure[1TB]{
-	\includegraphics[scale=0.59]{1T_total_time}
+	\includegraphics[scale=0.58]{1T_total_time}
 	\label{fig:1T_eval_res}
@@
 \subfigure[3TB]{
-	\includegraphics[scale=0.59]{3T_total_time}
+	\includegraphics[scale=0.58]{3T_total_time}
 	\label{fig:3T_eval_res}
@@
 \caption{Total Elapsed Time\label{fig:total_elapsed_time}}
+\vspace{-0.2in}
 \end{figure*}
@@
 \subfigure[1TB]{
-	\includegraphics[scale=0.59]{1T_quality}
+	\includegraphics[scale=0.58]{1T_quality}
 	\label{fig:1T_quality}
@@
 \subfigure[3TB]{
-	\includegraphics[scale=0.59]{3T_quality}
+	\includegraphics[scale=0.58]{3T_quality}
 	\label{fig:3T_quality}
@@
 
-%Our evalution results are exhibited in Figures ~\ref{fig:eval_res} 
-%and ~\ref{fig:improvement}.
-%Figure~\ref{fig:eval_res} shows the times (in seconds) taken 
-%to run the 10Q or 20Q workloads over the 1TB or 3TB fact tables. 
-%Based on the run results, Figure~\ref{fig:improvement} displays 
-%how much improvement could be achieved by the EXP or the WIZARD 
-%recommendations against the NO PPI one.
-
 Figures~\ref{fig:total_elapsed_time} and \ref{fig:recomm_qual} exhibit our evaluation results.
-Figure~\ref{fig:total_elapsed_time} shows the times (in seconds) taken 
+Figure~\ref{fig:total_elapsed_time} shows the times taken 
 to run the 10Q and 20Q workloads over the 1TB and 3TB fact tables that we populated 
@@
 Based on the run results, Figure~\ref{fig:recomm_qual} shows 
-how much improvement was gained by the EXP and WIZARD solutions, compared with the NO PPI solution.
+how much improvement was gained by the EXP and WIZARD solutions, compared with that of NO~PPI.
 
 Overall, the WIZARD recommendations were very effective at partitioning 
-the fact tables by leveraging the predicates in the workloads. 
-On average, the \hbox{total} execution time of the workloads on the WIZARD solutions was 
-about 3.5 and 2 times faster than those of NO PPI and EXP, respectively. 
-Also, the quality of the \hbox{WIZARD} solutions outperformed that of the EXP solutions; 
+the fact table via the predicates in the given workloads. 
+As shown in Figure~\ref{fig:total_elapsed_time}, the 
+total execution times of the 10Q and 20Q on the WIZARD recommendations was 
+on average about 3.5 and 2 times faster than those of NO PPI and EXP, respectively. 
+The quality of the \hbox{WIZARD} solutions also outperformed that of the EXP solutions; 
 the WIZARD solutions yielded an average of about 77\% improvement on the NO PPI solutions.
 
-Furthermore, the WIZARD recommendations 
-scaled very well with growing workload size and table size, 
+In addition, the WIZARD recommendations 
+scaled very well with the combination of growing workload/table size, 
 as illustrated in Figure~\ref{fig:recomm_qual}. 
-Although we doubled the workload scale from 10Q to 20Q, 
-the performance improvement of the WIZARD solution stayed the same in Figure~\ref{fig:1T_quality} 
-and got almost negligibly decreased in Figure~\ref{fig:3T_quality}. 
+Even though we doubly increased the workload size from 10Q to 20Q, 
+the performance improvement of our solution (\hbox{WIZARD}) retained the same in Figure~\ref{fig:1T_quality}. 
+Also, for the 3T table the performance gain of WIZARD was almost unchanged (at around 80\%) as shown in Figure~\ref{fig:3T_quality}. 
 This result showed that our WIZARD \hbox{solution} was scalable over increasing workload size.
@@
 We also increased the table scale from 1TByte to 3TBbytes. 
-For 10Q the performance improvement (about 77\%) was almost the same over the increasing scales, 
-and for 20Q it was decreased from 76\% to 74\%, which was \hbox{almost} negligible.
-The WIZARD partitioning \hbox{recommendation} scaled well with the increasing table size.
+For 10Q the performance improvement (about 77\%) was almost the same over the increasing scales.
+For 20Q the gain was reduced from 76\% to 74\%, but it was almost negligible.
+Hence, the WIZARD partitioning \hbox{recommendation} scaled well with the increasing table size.
 
 In sum, our experiment results attest the effectiveness of the proposed algorithms for WIZARD 
-and show the superiority of the recommendations of WIZARD.
+and demonstrate the superiority of the recommendations of WIZARD.
 
@@
 
-%Also, we scaled up the table from 1TB to 3TB on the fixed workload, but 
-%the \hbox{recommendation} quality was not degraded at all. 
 These results attest the effectiveness of our algorithms, 
@@
 
-%Physical database design~\cite{finkelstein1988physical,labio1997physical,rozen1991framework,Zilio1998} 
 Physical database design~\cite{finkelstein1988physical,labio1997physical,rozen1991framework} 
-has been discussed in \hbox{academic} research and industrial sectors in the past years.
-The major DBMS vendors (e.g., IBM, MS, and Oracle) have driven much of the work. 
-Their specific interests have been in automating the physical design for table 
-partitioning~\cite{agrawal04:integrating,Lightstone04:db2auto,nehme2011automated,sheet2009oracle,saa2007oracle,rao2002automating}, indexes/materialized views~\cite{Agrawal04:sqlserver,agrawal2000automated,dash2011cophy,Kimura11,itwsqlserver,valentin2000db2,zilio2004recommending},
-and integration~\cite{Zilio04:db2design}.
-
-%Lightstone's work~\cite{Lightstone04:db2auto} introduces the MDC (Multidimensional 
-%Clustering) Advisor that automates the selection of MDC 
-%keys customized for a specific workload on DB2 Universal Databases to use MDC structure. 
-%DB2 Advisor~\cite{Zilio04:db2design} is an integrated commercial 
-%tool that captures a variety of aspects of physical database design 
-%by automation in a comprehensive way. 
-%Oracle Partitioning Advisor~\cite{sheet2009oracle} for recommending 
-%table partitionings is  integrated into 
-%the SQL Access Advisor~\cite{saa2007oracle}.
-
-%Since data clustering significantly influences on the performance of 
-%multidimensional queries, the MDC advisor helps DBAs that may have 
-%nontrivial troubles in finding proper clustering dimensions and 
-%granularity. 
-%To enumerate and determine possible clustering dimensions, 
-%his work relies on {\it what-if} query cost modeling, data sampling and a search algorithm.
-
-%DB2 Advisor~\cite{Zilio04:db2design} is an integrated commercial 
-%tool that captures a variety of aspects of physical database design 
-%by automation in a comprehensive way. 
-%It makes an attempt to cover a variety of design choices-partitioning, 
-%MDC, materialized views and indexes. 
-%DB2 Advisor leverages the DB2 query optimizer for yielding a recommendation, and 
-%takes a hybrid searching approach for considering different features.
-
-%Oracle has a tool, called Partitioning Advisor~\cite{sheet2009oracle}, 
-%to make a partitioning recommendation for tables. 
-%The tool is now integrated into 
-%the SQL Access Advisor~\cite{saa2007oracle} that recommends indexes and materialized 
-%views. 
-%However, its technical details are not yet released in public. 
-%
-%Also, MS SQL Server provides some similar, automated tools~\cite{agrawal2000automated,agrawal04:integrating,
-%itwsqlserver} for table partitioning, materialized views or indexes fitting to a given workload.
-%The optimizer cost model is exploited to acquire estimates for 
-%candidate materialized views or indexes. 
-%In particular, Agrawal's work~\cite{agrawal04:integrating} presents 
-%an approach of integrating vertical and horizontal partitioning schemes 
-%into their physical database design, considering scalability.
+has been discussed in \hbox{academic} research and industrial communities in the past decades.
+The major DBMS vendors have led much of the work. 
+Their specific interests have been around automating the physical design for table 
+partitioning~\cite{agrawal04:integrating,Lightstone04:db2auto,nehme2011automated,sheet2009oracle,saa2007oracle,rao2002automating}, indexes/materialized views~\cite{Agrawal04:sqlserver,agrawal2000automated,dash2011cophy,Kimura11,itwsqlserver,valentin2000db2,zilio2004recommending}, and integration~\cite{Zilio04:db2design}.
 
@@
 appear to be similar to our MLPPI wizard. 
-However, in light of problem scope and approach, the wizard 
-is fundamentally different from the existing tools excluding Oracle Partitioning \hbox{Advisor}~\cite{sheet2009oracle} 
-providing no published technical details.
-
-DB2 MDC Advisor~\cite{Lightstone04:db2auto} actually tackles 
+But with respect to problem scope and approach the wizard 
+is {\em fundamentally different} from the existing tools, 
+except Oracle Partitioning \hbox{Advisor}~\cite{sheet2009oracle} that provides no published technical details.
+
+First, DB2 MDC Advisor~\cite{Lightstone04:db2auto} actually tackles 
 a different problem of automatically recommending the most well-suited MDC 
@@
 discusses another problem of merging \hbox{single-level} range partitionings on objects such as tables and indexes. 
-This article addresses the multi-level \hbox{partitioning} problem. 
+On the other hand, we address the multi-level \hbox{partitioning} problem. 
 Hence, the existing solutions cannot be directly applied to our MLPPI wizard.
@@
 Regarding the approach, DB2 MDC Advisor~\cite{Lightstone04:db2auto} uses 
-the search space driven by {\em fields}. In contrast, our search 
-space is driven by query-predicates, which is superior because 
-predicates are more customized and specific to a workload than fields.
+the search space driven by fields. In contrast, our search 
+space is driven by {\em query-predicates}, of which the use is superior to 
+that of the fields, as utilizing predicates can be more customized and specific to a given workload.
 
@@
 also uses the search space driven by simple range predicates, but 
-his technique has a shortcoming as follows. 
-His work produces a solution for each individual query and attempt 
+his technique has a shortcoming. 
+Specifically, his work produces a solution for each individual query and attempt 
 to merge the solutions. But this approach cannot reach 
 an optimized solution in a global perspective.
-Our wizard generates the whole search space upfront and in turn merges partitions, leading to a \hbox{globally-optimized} solution. 
-While his work considers a single column only, we deal with multiple fields.
+We generate the whole search space upfront and in turn merges partitions, 
+leading to a \hbox{globally-optimized} solution. 
+While his work considers a single column only, our wizard deals with {\em multiple} fields.
 
@@
 
-\section{Summary}
+\section{Conclusion}
 \label{sec:conclusion}
@@
 the range pair with the least scan or query cost, and eventually reached an MLPPI recommendation. 
-In addition, we analyzed the running complexity for the initial and optimized phases. 
-Despite the theoretically high bound, in practice, the wizard made much fewer calls in the phases.
-
-We measured the performance of the recommendation by the wizard using 
-our workloads.  
+We also analyzed the running complexity for the initial and optimized phases. 
+Despite the theoretically high bound, the wizard in practice invoked much fewer calls in these phases.
+
+Using synthetic workloads, we evaluated the performance of the recommendation produced by the wizard.  
 We demonstrated that the produced MLPPI solutions by 
-the wizard could reduce the total elapsed time by more than a factor of two, 
+the wizard reduced the total elapsed time by more than a factor of two, 
 compared with those of no partitioning approach and partitioning 
 done by a human expert. 
-
+%We believe our algorithm can be applied to any clustering and partitioning algorithm.
 %\section*{Acknowledgments} 

### 2017-03-29 pt_characterization/rep_check/paper.tex
Characterizing drafts

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
 \usepackage{auto-pst-pdf}
\usepackage{graphicx}
%\usepackage{arydshln}
\usepackage{ifpdf}
\ifpdf
  \usepackage{epstopdf}
\fi
\usepackage{multirow}
\usepackage{epsfig}
\usepackage{float}
\usepackage{url}
\usepackage{color}
\usepackage{subfigure}

%\newcommand\solidrule[1][1cm]{\rule[0.5ex]{#1}{.4pt}}
%\newcommand\dashedrule{\mbox{%
%  \solidrule[2mm]\hspace{2mm}\solidrule[2mm]\hspace{2mm}\solidrule[2mm]}}

%\usepackage{hyperref}

\begin{document}
\title{Repeatability Test}

\author{
Young-Kyoon Suh\\
}
\maketitle

\section{Description}
This document presents two sets of histograms based on the two runs 
of a program under test, called {\em INC}, with a set of increasing task lengths 
from 1 second to 4096 seconds.
We would like to see if both of the sets have the same shape for the same task length.
If so then, we can say that run repeatability is satisfied in our experiment setting.

\section{Experiment Notes}
Table~\ref{tab:exp_notes} provides a short description of our experimental runs, 
on which the following histograms are based.

%% done
\begin{table}[h]
\begin{center}
\begin{tabular}{|p{2cm}|p{3cm}|p{5cm}|p{5cm}|} \hline
Machine & Task Length (sec) & Description & Experiment Period\\ \hline
{\tt sodb9} &  INC1$\sim$INC64 & Two runs with 1000 samples & 2017-03-02 $\sim$ 2017-03-?? / 2017-03-?? $\sim$ 2017-03-??\\ \hline
{\tt sodb9} &  INC128$\sim$INC1024 & Two runs with 300 samples & 2017-03-02 $\sim$ 2017-03-?? / 2017-03-?? $\sim$ 2017-03-??\\ \hline
{\tt sodb10} & INC2048 & Two runs of 300 samples & 2017-03-02 $\sim$ 2017-03-09 / 2017-03-13 $\sim$ 2017-03-20\\ \hline
{\tt sodb12} & INC4096 & Two runs of 300 samples & 2017-02-13 $\sim$ 2017-02-27 / 2017-03-02 $\sim$ 2017-03-17\\ \hline
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{Notes on experiment runs used for histograms\label{tab:exp_notes}}
\end{table}

Now we show histograms of elapsed time (ET) and process time (PT) of INC.

\pagebreak

\input{res1.tex}

\pagebreak
\newpage

\input{res2.tex}

\end{document}

### 2017-03-29 pt_characterization/rep_check/res1.tex
Characterizing drafts

\section{Histograms on the First Run~\label{sec:sodb9_r1_hist}} 
This section exhibits histograms on the first run of 
INC with its task length increasing from 1 second to 4096 seconds, via EMPv5. 
The detailed description of the base data is from Table~\ref{tab:exp_notes}.

\subsection{ET}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC1]{
		\includegraphics[scale=0.43]{repet_data1/1_sec_et_hist_v5.eps}
		\label{fig:inc1_r1_et_hist_v5}
	}
	\subfigure[ET frequency on INC2]{
		\includegraphics[scale=0.43]{repet_data1/2_sec_et_hist_v5.eps}
		\label{fig:inc2_r1_et_hist_v5}
	}
	\subfigure[ET frequency on INC4]{
		\includegraphics[scale=0.43]{repet_data1/4_sec_et_hist_v5.eps}
		\label{fig:inc4_r1_et_hist_v5}
	}
	\subfigure[ET frequency on INC8]{
		\includegraphics[scale=0.43]{repet_data1/8_sec_et_hist_v5.eps}
		\label{fig:inc8_r1_et_hist_v5}
	}
	\caption{ET Histograms of INC1 ... INC8~\label{fig:s9_r1_et_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC16]{
		\includegraphics[scale=0.43]{repet_data1/16_sec_et_hist_v5.eps}
		\label{fig:inc16_r1_et_hist_v5}
	}
	\subfigure[ET frequency on INC32]{
		\includegraphics[scale=0.43]{repet_data1/32_sec_et_hist_v5.eps}
		\label{fig:inc32_r1_et_hist_v5}
	}
	\subfigure[ET frequency on INC64]{
		\includegraphics[scale=0.43]{repet_data1/64_sec_et_hist_v5.eps}
		\label{fig:inc64_r1_et_hist_v5}
	}
	\caption{ET Histograms of INC16 ... INC64~\label{fig:s9_r1_et_hist2}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC128]{
		\includegraphics[scale=0.43]{repet_data1/128_sec_et_hist_v5.eps}
		\label{fig:inc128_r1_et_hist_v5}
	}
	\subfigure[ET frequency on INC256]{
		\includegraphics[scale=0.43]{repet_data1/256_sec_et_hist_v5.eps}
		\label{fig:inc256_r1_et_hist_v5}
	}
	\subfigure[ET frequency on INC512]{
		\includegraphics[scale=0.43]{repet_data1/512_sec_et_hist_v5.eps}
		\label{fig:inc512_r1_et_hist_v5}
	}
	\subfigure[ET frequency on INC1024]{
		\includegraphics[scale=0.43]{repet_data1/1024_sec_et_hist_v5.eps}
		\label{fig:inc1024_r1_et_hist_v5}
	}
	\caption{ET Histograms of INC128 ... INC1024~\label{fig:s9_r1_et_hist3}}
\end{figure}

\pagebreak

\begin{figure}[t]
	\centering
	\subfigure[ET frequency on INC2048]{
		\includegraphics[scale=0.43]{repet_data1/2048_sec_et_hist_v5.eps}
		\label{fig:inc2048_r1_et_hist_v5}
	}
	\subfigure[ET frequency on INC4096]{
		\includegraphics[scale=0.43]{repet_data1/4096_sec_et_hist_v5.eps}
		\label{fig:inc4096_r1_et_hist_v5}
	}
	\caption{ET Histograms of INC2048 and INC4096~\label{fig:s9_r1_et_hist4}}
\end{figure}

\vspace\fill
\clearpage

\subsection{PT}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC1]{
		\includegraphics[scale=0.43]{repet_data1/1_sec_pt_hist_v5.eps}
		\label{fig:inc1_r1_hist_v5}
	}
	\subfigure[PT frequency on INC2]{
		\includegraphics[scale=0.43]{repet_data1/2_sec_pt_hist_v5.eps}
		\label{fig:inc2_r1_hist_v5}
	}
	\subfigure[PT frequency on INC4]{
		\includegraphics[scale=0.43]{repet_data1/4_sec_pt_hist_v5.eps}
		\label{fig:inc4_r1_hist_v5}
	}
	\subfigure[PT frequency on INC8]{
		\includegraphics[scale=0.43]{repet_data1/8_sec_pt_hist_v5.eps}
		\label{fig:inc8_r1_hist_v5}
	}
	\caption{PT Histograms of INC1 ... INC8~\label{fig:s9_r1_pt_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC16]{
		\includegraphics[scale=0.43]{repet_data1/16_sec_pt_hist_v5.eps}
		\label{fig:inc16_r1_hist_v5}
	}
	\subfigure[PT frequency on INC32]{
		\includegraphics[scale=0.43]{repet_data1/32_sec_pt_hist_v5.eps}
		\label{fig:inc32_r1_hist_v5}
	}
	\subfigure[PT frequency on INC64]{
		\includegraphics[scale=0.43]{repet_data1/64_sec_pt_hist_v5.eps}
		\label{fig:inc64_r1_hist_v5}
	}
	\caption{PT Histograms of INC16 ... INC64\label{fig:s9_r1_pt_hist2}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC128]{
		\includegraphics[scale=0.43]{repet_data1/128_sec_pt_hist_v5.eps}
		\label{fig:inc128_r1_hist_v5}
	}
	\subfigure[PT frequency on INC256]{
		\includegraphics[scale=0.43]{repet_data1/256_sec_pt_hist_v5.eps}
		\label{fig:inc256_r1_hist_v5}
	}
	\subfigure[PT frequency on INC512]{
		\includegraphics[scale=0.43]{repet_data1/512_sec_pt_hist_v5.eps}
		\label{fig:inc512_r1_hist_v5}
	}
	\subfigure[PT frequency on INC1024]{
		\includegraphics[scale=0.43]{repet_data1/1024_sec_pt_hist_v5.eps}
		\label{fig:inc1024_r1_hist_v5}
	}
	\caption{PT Histograms of INC256 ... INC1024~\label{fig:s9_r1_pt_hist3}}
\end{figure}

\begin{figure}[t]
	\centering
	\subfigure[PT frequency on INC2048]{
		\includegraphics[scale=0.43]{repet_data1/2048_sec_pt_hist_v5.eps}
		\label{fig:inc2048_r1_hist_v5}
	}
	\subfigure[PT frequency on INC4096]{
		\includegraphics[scale=0.43]{repet_data1/4096_sec_pt_hist_v5.eps}
		\label{fig:inc4096_r1_hist_v5}
	}
	\caption{PT Histograms of INC2048 and INC4096~\label{fig:s9_r1_pt_hist4}}
\end{figure}

\clearpage

### 2017-03-29 pt_characterization/rep_check/res2.tex
Characterizing drafts

\section{Histograms on the Second Run~\label{sec:sodb9_r2_hist}} 
This section exhibits histograms on the second run of 
INC with its task length increasing from 1 second to 4096 seconds, via EMPv5. 
The detailed description of the base data is from Table~\ref{tab:exp_notes}.

\subsection{ET}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC1]{
		\includegraphics[scale=0.43]{repet_data1/1_sec_et_hist_v5.eps}
		\label{fig:inc1_r2_et_hist_v5}
	}
	\subfigure[ET frequency on INC2]{
		\includegraphics[scale=0.43]{repet_data2/2_sec_et_hist_v5.eps}
		\label{fig:inc2_r2_et_hist_v5}
	}
	\subfigure[ET frequency on INC4]{
		\includegraphics[scale=0.43]{repet_data2/4_sec_et_hist_v5.eps}
		\label{fig:inc4_r2_et_hist_v5}
	}
	\subfigure[ET frequency on INC8]{
		\includegraphics[scale=0.43]{repet_data2/8_sec_et_hist_v5.eps}
		\label{fig:inc8_r2_et_hist_v5}
	}
	\caption{ET Histograms of INC1 ... INC8~\label{fig:s9_r2_et_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC16]{
		\includegraphics[scale=0.43]{repet_data2/16_sec_et_hist_v5.eps}
		\label{fig:inc16_r2_et_hist_v5}
	}
	\subfigure[ET frequency on INC32]{
		\includegraphics[scale=0.43]{repet_data2/32_sec_et_hist_v5.eps}
		\label{fig:inc32_r2_et_hist_v5}
	}
	\subfigure[ET frequency on INC64]{
		\includegraphics[scale=0.43]{repet_data2/64_sec_et_hist_v5.eps}
		\label{fig:inc64_r2_et_hist_v5}
	}
	\caption{ET Histograms of INC16 ... INC64~\label{fig:s9_r2_et_hist2}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC128]{
		\includegraphics[scale=0.43]{repet_data2/128_sec_et_hist_v5.eps}
		\label{fig:inc128_r2_et_hist_v5}
	}
	\subfigure[ET frequency on INC256]{
		\includegraphics[scale=0.43]{repet_data2/256_sec_et_hist_v5.eps}
		\label{fig:inc256_r2_et_hist_v5}
	}
	\subfigure[ET frequency on INC512]{
		\includegraphics[scale=0.43]{repet_data2/512_sec_et_hist_v5.eps}
		\label{fig:inc512_r2_et_hist_v5}
	}
	\subfigure[ET frequency on INC1024]{
		\includegraphics[scale=0.43]{repet_data2/1024_sec_et_hist_v5.eps}
		\label{fig:inc1024_r2_et_hist_v5}
	}
	\caption{ET Histograms of INC128 ... INC1024~\label{fig:s9_r2_et_hist3}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC2048]{
		\includegraphics[scale=0.43]{repet_data2/2048_sec_et_hist_v5.eps}
		\label{fig:inc2048_r2_et_hist_v5}
	}
	\subfigure[ET frequency on INC4096]{
		\includegraphics[scale=0.43]{repet_data2/4096_sec_et_hist_v5.eps}
		\label{fig:inc4096_r2_et_hist_v5}
	}
	\caption{ET Histograms of INC2048 and INC4096~\label{fig:s9_r2_et_hist4}}
\end{figure}

\vspace\fill
\clearpage

\subsection{PT}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC1]{
		\includegraphics[scale=0.43]{repet_data2/1_sec_pt_hist_v5.eps}
		\label{fig:inc1_r2_hist_v5}
	}
	\subfigure[PT frequency on INC2]{
		\includegraphics[scale=0.43]{repet_data2/2_sec_pt_hist_v5.eps}
		\label{fig:inc2_r2_hist_v5}
	}
	\subfigure[PT frequency on INC4]{
		\includegraphics[scale=0.43]{repet_data2/4_sec_pt_hist_v5.eps}
		\label{fig:inc4_r2_hist_v5}
	}
	\subfigure[PT frequency on INC8]{
		\includegraphics[scale=0.43]{repet_data2/8_sec_pt_hist_v5.eps}
		\label{fig:inc8_r2_hist_v5}
	}
	\caption{PT Histograms of INC1 ... INC8~\label{fig:s9_r2_pt_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC16]{
		\includegraphics[scale=0.43]{repet_data2/16_sec_pt_hist_v5.eps}
		\label{fig:inc16_r2_hist_v5}
	}
	\subfigure[PT frequency on INC32]{
		\includegraphics[scale=0.43]{repet_data2/32_sec_pt_hist_v5.eps}
		\label{fig:inc32_r2_hist_v5}
	}
	\subfigure[PT frequency on INC64]{
		\includegraphics[scale=0.43]{repet_data2/64_sec_pt_hist_v5.eps}
		\label{fig:inc64_r2_hist_v5}
	}
	\caption{PT Histograms of INC16 ... INC64\label{fig:s9_r2_pt_hist2}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC128]{
		\includegraphics[scale=0.43]{repet_data2/128_sec_pt_hist_v5.eps}
		\label{fig:inc128_r2_hist_v5}
	}
	\subfigure[PT frequency on INC256]{
		\includegraphics[scale=0.43]{repet_data2/256_sec_pt_hist_v5.eps}
		\label{fig:inc256_r2_hist_v5}
	}
	\subfigure[PT frequency on INC512]{
		\includegraphics[scale=0.43]{repet_data2/512_sec_pt_hist_v5.eps}
		\label{fig:inc512_r2_hist_v5}
	}
	\subfigure[PT frequency on INC1024]{
		\includegraphics[scale=0.43]{repet_data2/1024_sec_pt_hist_v5.eps}
		\label{fig:inc1024_r2_hist_v5}
	}
	\caption{PT Histograms of INC256 ... INC1024~\label{fig:s9_r2_pt_hist3}}
\end{figure}

\begin{figure}[t]
	\centering
	\subfigure[PT frequency on INC2048]{
		\includegraphics[scale=0.43]{repet_data2/2048_sec_pt_hist_v5.eps}
		\label{fig:inc2048_r2_hist_v5}
	}
	\subfigure[PT frequency on INC4096]{
		\includegraphics[scale=0.43]{repet_data2/4096_sec_pt_hist_v5.eps}
		\label{fig:inc4096_r2_hist_v5}
	}
	\caption{PT Histograms of INC2048 and INC4096~\label{fig:s9_r2_pt_hist4}}
\end{figure}

### 2017-03-29 pt_characterization/rep_check/repet_data1/et_hist_sodb9.r

@@ -1,0 +1,279 @@
+### sodb9
+numSamples <- 1000
+x = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("1_sec_et_hist_v5.eps")
+binsize=1
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
+h = hist(x$METIME, right=F, breaks=nbins,plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+xmin <-min(x$METIME)
+xmax <-max(x$METIME)
+plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC1', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='ET (ms)', ylab=expression('Frequency'))
+xaxtl <- seq(xmin-1,xmax+1,by=1)
+axis(side=1, at=xaxtl, labels=xaxtl)
+dev.off()
+
+x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("2_sec_et_hist_v5.eps")
+binsize=1
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
+h = hist(x$METIME, right=F, breaks=nbins,plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+xmin <-min(x$METIME)
+xmax <-max(x$METIME)
+plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC2', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
[…]

### 2017-03-29 pt_characterization/rep_check/repet_data1/pt_hist_sodb9.r

@@ -1,0 +1,255 @@
+### sodb9
+numSamples <- 1000
+x = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("1_sec_pt_hist_v5.eps")
+binsize=1
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+nbins <- ceiling((max(x$PRTIME)-min(x$PRTIME)) / binsize)
+h = hist(x$PRTIME, right=F, breaks=nbins,plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="green", main='PT frequency on INC1', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+xaxtl <- seq(xmin-1,xmax+1,by=1)
+axis(side=1, at=xaxtl, labels=xaxtl)
+dev.off()
+
+x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("2_sec_pt_hist_v5.eps")
+binsize=1
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+nbins <- ceiling((max(x$PRTIME)-min(x$PRTIME)) / binsize)
+h = hist(x$PRTIME, right=F, breaks=nbins,plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+#ymax <- 400
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="green", main='PT frequency on INC2', 
[…]

### 2017-03-29 pt_characterization/rep_check/repet_data2/et_hist_sodb9.r

@@ -1,0 +1,280 @@
+### sodb9
+numSamples <- 1000
+x = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("1_sec_et_hist_v5.eps")
+binsize=1
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
+h = hist(x$METIME, right=F, breaks=nbins,plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+xmin <-min(x$METIME)
+xmax <-max(x$METIME)
+plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC1', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='ET (ms)', ylab=expression('Frequency'))
+xaxtl <- seq(xmin-1,xmax+1,by=1)
+axis(side=1, at=xaxtl, labels=xaxtl)
+dev.off()
+
+x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("2_sec_et_hist_v5.eps")
+binsize=1
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
+h = hist(x$METIME, right=F, breaks=nbins,plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+xmin <-min(x$METIME)
+xmax <-max(x$METIME)
+plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC2', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
[…]

### 2017-03-29 pt_characterization/rep_check/repet_data2/pt_hist_sodb9.r

@@ -1,0 +1,256 @@
+### sodb9
+numSamples <- 1000
+x = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("1_sec_pt_hist_v5.eps")
+binsize=1
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+nbins <- ceiling((max(x$PRTIME)-min(x$PRTIME)) / binsize)
+h = hist(x$PRTIME, right=F, breaks=nbins,plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="green", main='PT frequency on INC1', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+xaxtl <- seq(xmin-1,xmax+1,by=1)
+axis(side=1, at=xaxtl, labels=xaxtl)
+dev.off()
+
+x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("2_sec_pt_hist_v5.eps")
+binsize=1
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+nbins <- ceiling((max(x$PRTIME)-min(x$PRTIME)) / binsize)
+h = hist(x$PRTIME, right=F, breaks=nbins,plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+#ymax <- 250
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="green", main='PT frequency on INC2', 
[…]

### 2017-03-31 sedona-ieice16/final/reviews/review_response.tex
SEDONA minor revision

\documentclass[10pt,letterpaper]{article}

\usepackage{graphicx}
\usepackage{longtable}
\usepackage{algorithmic}
\usepackage{alltt}
\usepackage{array}
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
%\usepackage[caption=false]{caption}
\usepackage[dvipsnames]{color}
%\usepackage{colortbl}
%\usepackage{enumitem}
%\usepackage{eqparbox}
%\usepackage{fixltx2e}
%\usepackage{float}
%\usepackage{floatflt}
%\usepackage{mdwmath}
%\usepackage{mdwtab}
%\usepackage{multirow}
\usepackage{stfloats}
%\usepackage[font=footnotesize]{subfig}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\usepackage[tight,normalsize,sf,SF]{subfigure}
%\usepackage{subfigure} 
\usepackage{times}
\usepackage{url}
\usepackage{verbatim} 
\usepackage{wrapfig}
\usepackage{xspace}
\usepackage{color}

\definecolor{grey}{RGB}{200,200,200}
\newcommand{\hilite}[1]{\colorbox{yellow}{#1}}
\newcommand{\hiliting}[1]{\colorbox{grey}{#1}}
\long\def\todo#1{\hilite{{\bf TODO:} {\em #1}}}

\setlength{\oddsidemargin}{0in}
\setlength{\textwidth}{6.5in}

% new commands: from file -- to permit re-use
%\input{newcommands.tex}

%Hyphenation problems
\hyphenation{name-space}
\hyphenation{schema-location}
\hyphenation{xml-lint}
\hyphenation{time-stamp}
\hyphenation{time-stamps}
\hyphenation{time-stamped}

\urlstyle{rm}
\graphicspath{{./figures/}}

\definecolor{gray}{rgb}{0.5,0.5,0.5}
%\newcommand{\rev}[1]{\textcolor{blue}{#1}}
%\newcommand{\quo}[1]{\textcolor{gray}{#1}}
%\newcommand{\rev}[1]{\vspace{3mm} \noindent {\em {\color{blue}{#1}}}}

\newenvironment{myindentpar}[1]%
{\begin{list}{}
         {\vspace{10pt}
					\setlength{\leftmargin}{#1}}
          \item[]
}
{\end{list}}
\newcommand{\rev}[1]{\begin{myindentpar}{.25in} {\em {\color{blue}{#1}}}\end{myindentpar}}

\newenvironment{myindentparQ}[1]%
{\begin{list}{}
         {\setlength{\leftmargin}{#1}}
          \item[]
}
{\end{list}}
\newcommand{\quo}[1]{\begin{myindentparQ}{.25in} {{\color{gray}{#1}}}\end{myindentparQ}}

\begin{document}

\title{Author's Reply}
\author{}
\maketitle

\section*{Overview}\label{sec:overview}
I appreciate your review and address your remaining question below. 
Note that my response to your question is 
incorporated into the revision and marked in {\color{blue}blue}.

\newcounter{RP}

\clearpage
\section*{Reviewer}\label{sec:rev1}

% 1
\rev{
$<<$ Reviewer's comments to the author(s) $>>$

The manuscript is well revised.

However, I have a question how to estimate each daemon's influence as time.
The estimation is important for accuracy of your proposed method.

For acceptance, please clarify the estimation method and its validity.

}

If I understand your question correctly, 
the proposed {\em cutoff} method addresses your question. 
We elaborate on deriving the cutoffs for each infrequent, long-running daemon 
from Steps~2 to~9 in Figure~3, which is marked in blue 
with a provided one-line comment above Step~2, for better clarification.

Here is the summarized procedure of the cutoff estimation. 
We first collect all the daemon processes 
while running PUT128 800 times enough to fit in one day.

For more details, please see the prose in Section~2 marked in blue.

\end{document}

### 2017-03-31 sedona-ieice16/final/paper.tex
SEDONA minor revision

\documentclass[letter]{ieice}
\usepackage{balance}
\usepackage{flushend}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{amssymb,amsthm,amsmath}
\usepackage{balance}
\usepackage{graphicx}          % graphicx package for including ps files  
\usepackage{epsfig}
\usepackage{url}
\usepackage{subfigure}
\usepackage{algorithmic}
\usepackage{float}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\usepackage{latexsym}
\usepackage{multirow}
\usepackage{color}
\usepackage{url}
\usepackage{comment}
%\usepackage{pstricks}
\usepackage{array}
\usepackage{epstopdf}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\setcounter{page}{1}
%\breakauthorline{}% breaks lines after the n-th author

\field{D}
\title{SEDONA: A Novel Protocol for Identifying Infrequent, Long-Running Daemons on a Linux System}
%\titlenote{This work was in part supported by the EDISON program (NRF-2011-0020576).}
\authorlist{
 \authorentry[yksuh@kisti.re.kr]{Young-Kyoon SUH}{m}{labelA}
 %\authorentry[rts@cs.arizona.edu]{Richard T. Snodgrass}{n}{labelB}}
\affiliate[labelA]{Dept. of Scientific Platform Development, Korea Inst. of Sci. and Tech. Info. (KISTI), Daejeon, 34141, Korea}
%\affiliate[labelB]{Department of Computer Science, University of Arizona, Tucson, AZ 85721, USA}
}
\received{2017}{4}{7}

\definecolor{grey}{RGB}{200,200,200}
\newcommand{\hilite}[1]{\colorbox{grey}{#1}}
\newcommand{\hilitey}[1]{\colorbox{yellow}{#1}}
\newcommand{\hiliting}[1]{\colorbox{grey}{#1}}
\long\def\todo#1{\hilitey{{\bf TODO:} {\em #1}}}
\long\def\shorten#1{}
\long\def\comment#1{}

\begin{document}

\maketitle

\begin{summary}
Measuring program execution time is 
a \hbox{much-used} technique for performance evaluation in computer science. 
Without proper care, however, timed results may vary a lot, 
thus making it hard to trust their validity. 
In this paper we propose a novel timing protocol to significantly 
reduce such variability.

\end{summary}
\begin{keywords}
Infrequent Long-running Daemon, Execution-Time Measurement
\end{keywords}

\section{Introduction}
\label{sec:intro}

Measuring program execution time is a much-used
technique for performance evaluation in computer science. 
Despite the importance of accurate and precise execution-time measurement, 
how to achieve {\em better} timing has not been well addressed. 
Surprisingly, there is considerable variability in the measured time.
The goal of this paper is to propose
a better timing protocol that can significantly reduce such variability 
and thus enable better timing results without a doubt of their validity.

Consider a simple compute-bound program, what we term {\em PUT} (program-under-test), 
as shown in Fig.~\ref{alg:put}. 
This program runs a nested for-loop with a specified task length ($tl$) (in seconds). 
The $tl$ value is used to compute 
the number of iterations ($t$) for which that for-loop is performed 
to reach the specified task length.

\vspace{-.2in}
\begin{figure}[h]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} PerformManyIncrements($tl$): \\
\STATE $t$ = $tl$ * {CONSTANT} \\
\FOR{$k$ = 1 to $t$ by 1}
	\FOR{$i$ = 1 to {UINT\_MAX}-1 by 1}
		\STATE $j$ += 1 \\
	\ENDFOR 
\ENDFOR 
\end{algorithmic}
\end{center}
\caption{Computation by a Program-Under-Test (PUT)\label{alg:put}}
\vspace{-.2in}
\end{figure}

We used 128~seconds as the task length and ran the PUT program (termed PUT128) 
800~times. 
Note that PUT128 was run on a machine, with 
a specification illustrated in Table~\ref{tab:machine_config}, 
in the same timing environment as used in our prior work~\cite{Currim}, 
with (i)~non-critical daemons deactivated, 
(ii)~the Network Timing Protocol daemon switched on,  
(iii)~particular CPU features~\cite{intel15,intelSpeed15} turned off, 
and (iv)~an up-to-date kernel installed.

Fig.~\ref{fig:meas_comp} shows two plots of the timing results on PUT128.
Most of the 800 executions measured in elapsed time, 
via  {\tt System.currentTimeMillis()} in Java, form the central band 
of \hbox{128,000--129,000 ms,} as can be seen in Fig.~\ref{fig:hist_before_et}. 
Despite the well-controlled timing environment, there exist over a dozen of executions above that band, or {\em outliers}. Indeed, these outliers {\em critically} impact the overall measurement quality: a standard deviation of 1,730 ms and a relative error of 0.01. This result suggests adequately eliminating these outliers while retaining most of the executions.

Our protocol, which we characterize shortly, on the contrary, 
exactly identified and removed those outliers. 
In effect, the protocol considerably reduced such variability, 
as illustrated in Fig.~\ref{fig:hist_after_pt}. 
Specifically, regarding standard deviation and relative error 
the protocol presented a significant performance gain up to by 
three orders of magnitude for the same PUT128 results.

{\bf Characteristics.} There are several important specifics of our protocol. 
First, we utilize {\em process time} (PT) rather than on elapsed time (ET). 
The use of PT is preferred, as it takes into account 
the time taken for only a ``process'' (or program) of interest.
PT is defined as the sum of ticks (where one tick is equal to 10 msec)
in user and system mode. 
These tick measures are obtainable via {\tt taskstats} C struct, 
provided by the Linux NetLink facility~\cite{Netlink}. 
Second, the protocol throws out some executions that involve
{\em infrequent, \hbox{long-running} daemon processes}. 
We observe that such daemon processes substantially impact 
the execution time of the process.  
By eliminating those executions, the protocol greatly improves
the overall measurement quality, which would otherwise be biased by them.

{\bf Related Work.} 
McGeoch introduced
two basic methods of measuring program time: elapsed time and CPU time~\cite{Mcgeoch12}. 
Bryant and O'Hallaron~\cite{Randal03} 
presented two timing schemes of using clock-cycle and interval counters. 
They proposed a measurement protocol, called {\em minimum-of-k}, 
that for observed elapsed ticks the minimum is chosen as the most accurate one. 
Odom et al.'s work~\cite{Odom05} focused on timing long-running programs 
in a simulation framework via dynamic sampling of trace snippets during program execution. 
But none of these prior works considers the variability in 
timing and the impact of daemons that may disturb the timing a lot.

\begin{figure}[t]
	\centering
	\subfigure[Initial Elapsed Time Measurement]{
		\includegraphics[width=0.224\textwidth]{128_sec_et_old.eps}
		\label{fig:hist_before_et}
	}
	\subfigure[Program Time After Our Protocol]{
		\includegraphics[width=0.224\textwidth]{128_sec_pt_new.eps}
		\label{fig:hist_after_pt}
	}
	\caption{An Example of Measurement Results on a \linebreak \hbox{Compute-bound} 
	Program-Under-Test~\label{fig:meas_comp}}
	\vspace{-.18in}
\end{figure}

Commercial software tools measure execution time~\cite{VTune,TimeSys,WindView}. 
Since the tools' source code is not disclosed, there is no way of figuring out 
whether they can prevent such a daemon from timing.

\vspace{0.05in}
{\bf Contribution.} 
Our contributions are following.
\vspace{-0.05in}
\begin{itemize}

\item We show empirical evidence that 
measuring execution time 
can be seriously affected by daemons.

\shorten{
\item We propose a novel timing protocol that
identifies infrequent, long-running daemons that impact the timing results for that program.

\item We evaluate the performance of the protocol with rigorous experiments 
using real-world programs.

\item The experimental results show a support for the effectiveness of our scheme\shorten{SEDONA: 
early eliminating dirty executions infected with such a daemon 
and eventually contributing to enhancing the overall measurement-quality}. 
}

{\color{blue}
%\item We present an algorithm to identify infrequent, long-running daemons impacting 
%the timing of a compute-bound program.
%
%\item We propose a more accurate and precise protocol based on the identification variance 
%variance for that program.

\item We present an algorithm to identify and eliminate such infrequent and long-running 
daemons.

\item We propose a novel timing protocol that can considerably 
reduce variance via the elimination.

\item The evaluation results on real-world programs 
show a support for the effectiveness of the protocol. 
}

\end{itemize}
\vspace{-0.2in}

\noindent
The rest is organized as follows. 
We next elaborate on the proposed timing protocol. In turn we evaluate the performance of the scheme using real workloads.

\section{Proposed Scheme}
\label{sec:prop_appach}

In this section we propose a novel execution time measurement scheme, 
called {\em SEDONA} (Selective Elimination through Detection of infrequent, lOng-ruNning dAemons). 
Our scheme (1) utilizes PT (process time) to avoid absorbing timing noise from other co-running processes and (2)~identifies and eliminates executions including daemon processes that are infrequent and \hbox{long-running} via a {\em cutoff} measure.

Our SEDONA protocol consists of a total of ten steps, as described in Fig.~\ref{alg:find}.
Note that this protocol (algorithm) is applicable to any arbitrary Linux system.

\begin{figure}[h]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} The SEDONA Timing Protocol: \\
\STATE Step 1. Set up the timing environment.

{\color{blue}
\tcc{{\scriptsize Steps 2---9 produce cutoffs for each identified daemon.}}
\STATE Step 2. Perform a single PUT run (specifically, PUT128) for many samples (specifically, 800).
\STATE Step 3. Consider each pair of elapsed time measurements to be a dual-PUT measurement 
and examine a scatter-plot to see if it it displays an $L$-shape.
\STATE Step 4. Zoom into the central cluster to ensure that it is symmetric (roughly circular).
\STATE Step 5. Compute the maximum and standard deviation of the process time 
for each daemon encountered within the central cluster samples.
\STATE Step 6. Identify for each sample in the $L$-shape infrequent, long-running daemon executions. 
\STATE Step 7. Determine potentially periodic daemons based on the $L$-executions 
and for each daemon compute the minimum process time from those executions identified. 
\STATE Step 8. Perform Steps 1--6 above for a single run 
consisting of a small number of executions (specifically, 40) of PUT16384.  
\STATE Step 9. Compute the cutoffs for each identified daemon. 
}
\STATE Step 10. Discard an execution including a daemon of which process time is greater than the respective cutoff time.
\end{algorithmic}
\end{center}
\caption{Summary of the SEDONA Timing Protocol\label{alg:find}}
\vspace{-0.25in}
\end{figure}

{\bf A Running Example.} 
First, we set up the same measurement environment as described earlier in \hbox{Section}~\ref{sec:intro}, 
to eliminate known timing factors (Step~1).

{\color{blue}
We then run a program-under-test (called PUT) of 
Fig.~\ref{alg:put} with a task length of 128~sec, 
termed {\em PUT128}, 800 times (Step 2). 
(We render Fig.~\ref{fig:hist_before_et} using the 800 samples from this run.)
We use 128 seconds because that is long enough to perhaps experience an infrequent daemon. 
We run it 800 times to capture infrequent daemons that perhaps run every few hours or even 
once a day. 
Note that we collect all daemon processes as well as the PUT and their measures 
through the Netlink interface from the kernel before and after each 
timing.

%Fig.~\ref{fig:reg_put128} (= Fig.~\ref{fig:hist_before_et}) 
%plots all the 800 elapsed times (ETs) of PUT128.
%The plot clearly shows three rows; that is, 
%the top and middle rows represent over a dozen of outliers far from 
%the rest of the samples clustered in the bottom row. 
%We will now drill down into these outliers to 
%show how to reliably eliminate the indirect influence of 
%some ``infrequent, \hbox{long-running} daemons'' on PT (\hbox{process} time) of the PUT.
Fig.~\ref{fig:reg_put128} plots all the 800 elapsed times (ETs) of the run of PUT128.
The plot clearly shows three rows; that is, 
the top and middle rows represent over a dozen of outliers far from 
the rest of the samples clustered in the bottom row. 
We will now drill down into these outliers to 
show how to reliably eliminate the indirect influence of 
some ``infrequent, \hbox{long-running} daemons'' on PT (\hbox{process} time) of the PUT.

To identify such daemons, we use a novel \hbox{scatter} plot: 
those of {\em pairs of successive samples}. 
So samples~\#{1} and~\#{2} form the first pair and samples~\#{3} and~\#{4} 
form the second pair.
Such a pair is termed ``\hbox{{\it dual-PUT}}.'' 
A sample of \hbox{dual-PUT256}, for instance, 
consists of two consecutive (odd and even) 
samples of PUT128: 
sample~\#{1} of dual-PUT256 is equivalent 
to samples~\#{1} and~\#{2} of PUT128.

Fig.~\ref{fig:raw_put128} presents such a scatter plot 
of 400 samples of a run of \hbox{dual-PUT256} constructed 
from a run of the 800 PUT128 samples (Step~3). 
The $xy$-plane corresponds to the dual-PUT samples. 
For example, \hbox{sample}~\#{2} of a run of dual-PUT256 
is plotted as a point with samples~\#{3} and~\#{4} of a run of PUT128 
on the respective $x$ and $y$ axes. In that sense 
the $x$ ($y$) axis is named as ``Odd (Even) Sample's ET.''
There are two quite obvious outliers with ETs 
of 163,913 ms (rightmost) and 161,785 ms (uppermost), respectively.

\begin{figure}[ht]
	\centering
	\subfigure[PUT128 with 800 Samples]{
		\includegraphics[width=0.224\textwidth]{put128_empv4_new.eps}
		\label{fig:reg_put128}
	}
	\subfigure[Dual-PUT256 Samples]{
		\includegraphics[width=0.224\textwidth]{dual_et_put256_empv4.eps}
		\label{fig:raw_put128}
	}\vspace{-0.15in}
	\subfigure[Zooming in on the $L$-shape]{
		\includegraphics[width=0.224\textwidth]{zooming_in_dual_et_put256_empv4.eps}
		\label{fig:dual_put256_liters}
	}
	\subfigure[The Central Cluster]{
		\includegraphics[width=0.224\textwidth]{dual_et_clump_put256_empv4.eps}
		\label{fig:dual_et_put256_cluster}
        }
    \caption{Successive Scatter Plots of PUT128 with 800 samples 
    (equivalent to Dual-PUT256 with 400 samples) in Steps~2---4~\label{fig:put128_plot}} 
    \vspace{-0.3in}     
\end{figure}

We informally term this phenomenon of a scatter plot of a dual-PUT run an
``$L$-shape,'' and attribute it to the
presence of infrequent long-running daemons.

Figure~\ref{fig:dual_put256_liters} zooms into the lower left region,
focusing on the
tight cluster of samples.  Interestingly, this plot continues to exhibit an
$L$-shape, with perhaps a dozen or more $L$-samples in the left and bottom arms of
the ``$L$'', and again no samples in the upper right portion of the
scatter plot.

We continue zooming until we get to Figure~\ref{fig:dual_et_put256_cluster},
which shows a central cluster (Step 4). 
We confirm the symmetry of the ET measurements in the central cluster: there
is no $L$-shape, and thus no $L$-samples, and thus no obvious infrequent
long-running daemons.

We then perform Step~5, which computes 
the maximum process time and standard deviation 
of PT (process time, note the switch in emphasis from ET to PT) of the daemon processes
(i.e. {\tt flush-9:0}) 
observed in the central cluster samples in Fig.~\ref{fig:dual_et_put256_cluster}.

In Step~6 we identify, for each daemon in the \hbox{$L$-samples}, 
those that are actual long-running daemon
executions.  We define such executions as those whose PT 
is over two standard deviations above the maximum PT for
that daemon in the central cluster samples. 
In the running example  {\tt flush-9:0}, 
{\tt jbd2/md0-8}, and {\tt md0\_raid1} 
are determined as infrequent, \hbox{long-running}.
We also identify ``extra'' infrequent daemons: 
{\tt bash}, {\tt grep}, {\tt rhn\_check}, {\tt rhnsd}, {\tt rhsmcertd}, 
{\tt rhsmcertd-worke}, and {\tt sshd}, those 
found only in the $L$-samples but not in the central cluster.

For each of the infrequent daemons 
we use a heuristic to determine the daemon's periodicity:
the daemon must occur regularly in a sequence of samples.
For instance, {\tt rhn\_check} appears 
roughly every 112 samples (or almost every four hours). 
Four others ({\tt flush-9:0}, {\tt jbd2/md0-8}, {\tt md0\_raid1}, and {\tt rhn\_check}) 
all occur together and have a periodicity of about every 559 samples 
(5x longer, or just about 20 hours).

Next, we can compute for each so-identified infrequent, long-running daemon its
minimum time in the \hbox{$L$-samples} (Step 7). 
This computation provides a rough, initial distinction of a ``long-running''
daemon, or the valley between the maximum PT from the central cluster 
and the minimum PT from the $L$-samples, to differentiate ``short-running'' 
from ``\hbox{long-running}'' executions of the daemon. 
For those daemons (i.e. {\tt grep}) never appearing in the central cluster,
this initial \hbox{analysis} concludes only that they are infrequent.

In Step 8 we repeat Steps 1--6, but instead with the much-longer running
PUT16384 (4.5 hours per \hbox{sample} versus 2 minutes), to see 
if any of our identified \hbox{infrequent} daemons are actually frequent 
at that much longer PUT execution time. 
We find some frequent daemon processes appearing in both of the 
clusters of \hbox{dual-PUT256} and \hbox{dual-PUT32768}, each consisting of 
pairs of two successive samples of PUT128 and PUT16384.
That said, the central cluster also contains other
  processes not seen in the \hbox{dual-PUT256} central cluster: 
  {\tt grep}, {\tt rhn\_check}, {\tt rhnsd}, {\tt rhsmcertd}, 
  {\tt rhsmcertd-worke}, and {\tt sshd}. 
  But these daemons were categorized in the \hbox{dual-PUT256} analysis as {\em
  infrequent}, several having \hbox{periodicities} estimated at four or twenty hours.
When PUT128 had a ``short'' 
program time (or, two minutes), daemons with a periodicity of
hours are \hbox{infrequent}. But with PUT16384 
with a ``long'' program time (or, 4.5 hours),
some of those daemons are now frequent, and appear in the central cluster.

In Step~9 we compute 
the cutoff for each of those infrequent, long-running daemons so
identified, based on the runs of PUT128 and 
PUT16384 as collected in Tab.~\ref{tab:final_infrequent_cutoff}. 
Here is how to compute the cutoff. For the cutoff of such a daemon with PUT128, 
we take the \hbox{midpoint} between the maximum of that daemon's PTs in the 
central cluster (or 0, if absent) and the \hbox{minimum} of those in the $L$-samples. 
For the cutoff of such a daemon with PUT16384, we do the same. 
We then compute a ``task time'' as 5\% of the inferred periodicity. 
This 5\% ensures that such infrequent daemons will impact 
only a small percentage of the shorter PUTs, while presumably being 
associated with much larger cutoffs for the very long PUTs. 
We also include daemons that (a)~were identified as infrequent and 
\hbox{long-running} from PUT128 and (b)~were not identified as so in the PUT16384 
\hbox{$L$-samples}, but may have in the \hbox{dual-PUT32768} central cluster.
We then take the {\em maximum} of the two cutoffs for the final cutoff PT 
(the last column of Tab.~\ref{tab:final_infrequent_cutoff}). 
}

\vspace{-0.2in}
\begin{table}[ht]
\centering
{\tiny
\begin{tabular}{|p{1.2cm}|c|c|p{1cm}|p{1.45cm}|} \hline
{\tiny Process}  & {\tiny Cutoff PT} & {\tiny Cutoff PT} & {\tiny Task}  & {\tiny{\bf Final }} \\
{\tiny Name} & {\tiny on PUT128}  & {\tiny on PUT16K} & {\tiny Time} & {\tiny {\bf Cutoff PT}} \\\hline
{\tt bash} & 1 msec & --- & --- & {\bf 1 msec} \\ \hline
{\tt flush-9:0} & 64 msec & --- & $< 1$ hour & {\bf 64 msec} \\
                & ---     & 48 msec & $\geq 1$ hour & {\bf 48 msec} \\ \hline
{\tt grep }     & 1 msec & 12 msec & --- & {\bf 12 msec}\\ \hline
{\tt jbd2/md0-8} & 4 msec & --- & $< 1$ hour & {\bf 4 msec} \\
                & ---     & 11 msec & $\geq 1$ hour & {\bf 11 msec} \\ \hline
{\tt md0\_raid1} & 35 msec & ---     & $< 1$ hour  & {\bf 35 msec}\\
                 & ---     & 51 msec & $\geq 1$ hour & {\bf 51 msec} \\ \hline
{\tt rhn\_check}  & 281 msec & --- & $< 12$ min & {\bf 281 msec} \\
                 &  --- & 12,828 msec & $\geq 12$ min & {\bf 12,828 msec}\\ \hline
{\tt rhnsd} & 2 msec & --- & $< 12$ min & {\bf 2 msec} \\
            & ---    & 12 msec &$\geq 12$ min & {\bf 12 msec} \\ \hline
{\tt rhsmcertd}  & 1 msec & 1 msec & --- & {\bf 1 msec} \\  \hline
{\tt rhsmcertd}  & 57 msec & --- & $< 12$ min & {\bf 57 msec} \\
{\tt -worke}           &  --- & 119 msec  & $\geq 12$ min & {\bf 119 msec}\\ \hline
{\tt sshd} & 2 msec & 23 msec & --- & {\bf 23 msec}\\ \hline
\end{tabular}
}
\caption{Collected Infrequent, Long-running Daemons and Their Final Cutoff
  Process Time \hbox{(Step~9)}\label{tab:final_infrequent_cutoff}}
  \vspace{-0.3in}
\end{table}

Based on Tab.~\ref{tab:final_infrequent_cutoff}, 
we discard any \hbox{sample} \hbox{containing} an infrequent, long-running daemon execution
over the corresponding cutoff. We thus end up dropping just fifteen of the 800
PUT128 samples and only two of the forty PUT16384 samples. 
As a result, the \hbox{overall} \hbox{measurement} quality---the standard deviation and \hbox{relative} error 
---for PUT128 was improved 
by three orders of magnitude (as shown in Fig.~\ref{fig:meas_comp}) 
and by about two \hbox{orders} of magnitude for PUT16384.

\section{Evaluation}
\label{sec:eval}
\vspace{-0.07in}
We now evaluate the \hbox{performance} of SEDONA 
compared to that of the original timing scheme, termed ORG, based on 
elapsed time.
Our experiments were conducted on a \hbox{machine}
described in Table~\ref{tab:machine_config}. 
\begin{table}[h]
\vspace{-0.2in}
\begin{center}
{\tiny
\begin{tabular}{|l|p{7cm}|}\hline
OS & Red Hat Ent. Linux (RHEL) 6.4 with a kernel of 2.6.32 \\ \hline
CPU & Intel Core i7-870 Lynnfield 2.93GHz quad-core \hbox{processor}\\ \hline
RAM & 4GB of DDR3 1333 dual-channel memory\\ \hline
HDD & Western Digital Caviar Black 1TB 7200rpm SATA Drive\\ \hline
\end{tabular}
}
\end{center}
\caption{Machine Configurations\label{tab:machine_config}}
\vspace{-0.35in}
\end{table}

First, we compared the performance of the two approaches 
using real-world programs such as insertion sort, termed SORT and 
matrix multiplication (in column major), termed MM. 
In this comparison SEDONA outperformed ORG, as can be seen in Fig.~\ref{fig:synprog_test}. 
As workload size increased, the performance gap between the two 
were increasingly widened, reaching up to by 6x.

\vspace{-.2in}
\begin{figure}[h]
	\centering
	\subfigure[SORT - Standard Deviation]{
		\includegraphics[scale=0.31]{sort_std.eps}
		\label{fig:sort_std}
	}
	\subfigure[MM - Standard Deviation]{
		\includegraphics[scale=0.31]{matc_std.eps}
		\label{fig:matc_std}
	}
	\vspace{-.1in}
	\caption{Performance Comparison on Real-world Programs~\label{fig:synprog_test}}
\vspace{-.25in}
\end{figure}

In turn, our evaluation proceeded with SPEC CPU2006 
benchmarks~\cite{specCpu2006}, providing various \linebreak 
\hbox{compute-bound} real applications. The results are \hbox{provided} in Table~\ref{tab:spec_real}. 
Note that in the table the results for {\tt 481} and {\tt 483} \hbox{benchmarks} 
are omitted because of some runtime error and incurred I/O, respectively.

Table~\ref{tab:spec_real} shows that 
SEDONA bettered ORG,
on the standard deviation and relative error across the very different SPEC benchmarks. 
To be more specific, all the benchmarks revealed a smaller standard deviation 
from SEDONA as compared to that of ORG. 
Our timing protocol quite effectively filtered out infrequent daemon executions 
in the industrial workloads. The relative error of SEDONA was also 
lower than that of ORG for almost every benchmark.
For example, about a 10x \hbox{margin} between the two resulted from {\tt 434}.

SEDONA also scaled well for the SPEC workloads, 
with regard to growth of relative error as the execution time lengthened.
For the short benchmarks 
(e.g. {\tt 400}, {\tt 403}, {\tt 410}, 
{\tt 434}, {\tt 445}, and {\tt 999}: those taking \hbox{under} 100 sec), 
our scheme outperformed the ORG method by about 3.5x, on average. 
The SEDONA protocol continued its dominance against the conventional technique 
for the medium-length benchmarks (e.g. {\tt 447}, {\tt 456}, {\tt 470}, and {\tt 473}).
Even for the long-running benchmarks (e.g. {\tt 416}, {\tt 436}, and {\tt 454}, both $>$1,000 sec), 
the relative error of \hbox{SEDONA} was still lower than that of ORG.

To summarize, our proposed SEDONA protocol can achieve {\em better} accuracy,
precision, and scalability in measuring the execution time of real \hbox{compute-bound} programs 
than the ET-based existing method. 
The experimental results demonstrate 
the general applicability of SEDONA to timing any CPU-bound program.

%\vspace{-0.in}
\begin{table}[t]
\centering
{
\tiny
\begin{tabular}{|p{0.1cm}|R{1.1cm}|R{1.1cm}|R{0.6cm}|R{0.6cm}|R{0.92cm}|R{0.92cm}|} \hline
 		  & \multicolumn{2}{c|}{{\tiny Execution Time (ms)}} & \multicolumn{2}{c|}{{\tiny Standard}}   & \multicolumn{2}{c|}{\multirow{2}{*}{{\tiny Relative Error}}}\\ \cline{2-3}
 & \multicolumn{1}{c|}{{\tiny{ORG}}}  & \multicolumn{1}{c|}{{\tiny{SED.}}} & \multicolumn{2}{c|}{{\tiny Deviation (ms)}}  & \multicolumn{2}{c|}{{\tiny}}\\ \cline{4-7}	
 & \multicolumn{1}{c|}{{\tiny{(in ET)}}} & \multicolumn{1}{c|}{{\tiny{(in PT)}}} & {\tiny{ORG}} & {\tiny{SED.}} & \multicolumn{1}{c|}{{\tiny{ORG}}}& \multicolumn{1}{c|}{{\tiny{SED.}}}\\ \hline
{{\tt 400}} & 454 & 445 & {3} & {2} & {6$\times$10$^{-3}$} & {3$\times$10$^{-3}$}\\
{{\tt 401}} & 536,639 & 528,517 & {1,185} & {1,161} & {2$\times$10$^{-3}$} & {2$\times$10$^{-3}$}\\
{{\tt 403}} & 26,109	& 25,695 & {138} & {96} & {5$\times$10$^{-3}$} & {4$\times$10$^{-3}$}\\
{{\tt 410}} & 7,938 & 7,801 & {46} & {19} & {6$\times$10$^{-3}$} & {2$\times$10$^{-3}$}\\
{{\tt 416}} & 1,015,342 & 999,846 & {965} & {876} & {1$\times$10$^{-3}$} & {9$\times$10$^{-4}$}\\%% V
{{\tt 429}} & 235,811 & 232,209 & {623} & {600} & {3$\times$10$^{-3}$} & {3$\times$10$^{-3}$}\\
{{\tt 433}} & 480,586 & 473,256  & {743} & {725} & {2$\times$10$^{-3}$} & {2$\times$10$^{-3}$}\\ %% V
{{\tt 434}} & 16,495  & 16,242  & {75} & {7} & {5$\times$10$^{-3}$} & {5$\times$10$^{-4}$}\\
{{\tt 435}} & 990,575 & 975,445  & {947} & {900} & {1$\times$10$^{-3}$} & {9$\times$10$^{-4}$}\\
{{\tt 436}} & 1,160,742 & 1,143,078   & {3,914} & {3,843} & {3$\times$10$^{-3}$} & {3$\times$10$^{-3}$}\\
{{\tt 437}} & 581,635 & 572,775 & {1,492} & {1,475}  & {3$\times$10$^{-3}$} & {3$\times$10$^{-3}$}\\
{{\tt 444}} & 591,201 & 582,229 & {294} & {281}  & {5$\times$10$^{-4}$} & {5$\times$10$^{-4}$}\\
{{\tt 445}} & 84,435  & 83,164  & {91} & {28}  & {1$\times$10$^{-3}$} & {3$\times$10$^{-4}$}\\
{{\tt 447}} & 521,846 & 513,493  & {150} & {108}  & {3$\times$10$^{-4}$} & {2$\times$10$^{-4}$}\\
{{\tt 450}} & 341,030 & 335,848  & {99} & {91}  & {3$\times$10$^{-4}$} & {2$\times$10$^{-4}$}\\
{{\tt 453}} & 258,797 & 254,496  & {623} & {582}  & {2$\times$10$^{-3}$} & {2$\times$10$^{-3}$}\\
{{\tt 454}} & 1,721,804 & 1,695,613  & {678} & {627}  & {4$\times$10$^{-4}$} & {4$\times$10$^{-4}$}\\
{{\tt 456}} & 410,533 & 404,328  & {85} & {50}  & {2$\times$10$^{-4}$} & {1$\times$10$^{-4}$}\\
{{\tt 458}} & 589,541 & 580,591  & {542} & {513} &  {9$\times$10$^{-4}$} & {9$\times$10$^{-4}$}\\
{{\tt 459}} & 798,917 & 786,726  & {2,143} & {2,132}  & {3$\times$10$^{-3}$} & {3$\times$10$^{-3}$}\\
{{\tt 462}} & 595,188 & 586,120  & {3,326} & {3,274}  & {6$\times$10$^{-3}$} & {6$\times$10$^{-3}$}\\
{{\tt 464}} & 649,838 & 639,939  & {601} & {563}  & {9$\times$10$^{-4}$} & {9$\times$10$^{-4}$}\\
{{\tt 465}} & 895,754 & 882,106  & {883} & {797}  & {1$\times$10$^{-3}$} & {1$\times$10$^{-3}$}\\
{{\tt 470}}	& 349,830 & 344,510 & {143} & {94} & {4$\times$10$^{-4}$} & {3$\times$10$^{-4}$}\\
{{\tt 471}} & 367,589 & 361,959  & {2,114} & {2,072} & {6$\times$10$^{-4}$} & {6$\times$10$^{-4}$}\\
{{\tt 473}} & 362,587 & 357,090  & {359} & {317} &  {1$\times$10$^{-3}$} & {9$\times$10$^{-4}$}\\
%{{\tt 481}} & \multicolumn{4}{c}{Runtime Error}\\
{{\tt 482}} & 654,208	 & 644,223  & {2,436} & {2,392} &   {4$\times$10$^{-3}$} & {4$\times$10$^{-3}$}\\ %% V
%{{\tt 483}} & \multicolumn{4}{c}{Out-of-Scope}\\
{{\tt 998}} & 128	 & 127  & {0.6} & {0.6} &  {4$\times$10$^{-3}$} & {4$\times$10$^{-3}$}\\
{{\tt 999}} & 128 &  127 & {0.8} & {0.6} &  {6$\times$10$^{-3}$} & {5$\times$10$^{-3}$}\\ \hline 
\multicolumn{3}{|r|}{Averages} & 852 & 815 &  {3$\times$10$^{-3}$} & {2$\times$10$^{-3}$}\\ \hline
\end{tabular}
}
\caption{Performance Evaluation on the SPEC Benchmarks\label{tab:spec_real}}
\vspace{-0.25in}
\end{table}

\section{Conclusion}
\label{sec:conclusion}
\vspace{-0.07in}
We proposed a novel timing protocol called {\em SEDONA}, and 
performed an empirical evaluation to show that 
it is more precise and accurate than the extant method.

\balance

\vspace{-0.03in}
{\tiny 
\bibliographystyle{ieicetr} 
\bibliography{paper}
%\vspace{-0.2in}
}

\vspace\fill
\end{document}

### 2017-03-31 sedona-ieice16/final/paper_old.tex
SEDONA minor revision

--- 2017-03-31 sedona-ieice16/final/paper.tex
@@
 \title{SEDONA: A Novel Protocol for Identifying Infrequent, Long-Running Daemons on a Linux System}
-%\titlenote{This work was in part supported by the EDISON program (NRF-2011-0020576).}
+\titlenote{This work was in part supported by the EDISON program (NRF-2011-0020576).}
 \authorlist{
@@
  %\authorentry[rts@cs.arizona.edu]{Richard T. Snodgrass}{n}{labelB}}
-\affiliate[labelA]{Dept. of Scientific Platform Development, Korea Inst. of Sci. and Tech. Info. (KISTI), Daejeon, 34141, Korea}
+\affiliate[labelA]{Supercomputing R\&{D} Center, KISTI, 245 Daehangno, Yuseong-gu, Daejeon, 34141, Republic of Korea}
 %\affiliate[labelB]{Department of Computer Science, University of Arizona, Tucson, AZ 85721, USA}
 }
-\received{2017}{4}{7}
+\received{2017}{2}{10}
 
@@
 \begin{summary}
+{\color{blue}
 Measuring program execution time is 
@@
 reduce such variability.
+}
 
@@
 Surprisingly, there is considerable variability in the measured time.
+{\color{blue}
 The goal of this paper is to propose
@@
 and thus enable better timing results without a doubt of their validity.
+}
 
@@
 This program runs a nested for-loop with a specified task length ($tl$) (in seconds). 
+{\color{blue}
 The $tl$ value is used to compute 
 the number of iterations ($t$) for which that for-loop is performed 
-to reach the specified task length.
+to reach the specified task length.}
 
@@
 800~times. 
+{\color{blue}
 Note that PUT128 was run on a machine, with 
@@
 the overall measurement quality, which would otherwise be biased by them.
+}
 
@@
 	\centering
-	\subfigure[Initial Elapsed Time Measurement]{
+	\subfigure[{\color{blue}Initial Elapsed Time Measurement}]{
 		\includegraphics[width=0.224\textwidth]{128_sec_et_old.eps}
@@
 	}
-	\subfigure[Program Time After Our Protocol]{
+	\subfigure[{\color{blue}Program Time After Our Protocol}]{
 		\includegraphics[width=0.224\textwidth]{128_sec_pt_new.eps}
@@
 	}
-	\caption{An Example of Measurement Results on a \linebreak \hbox{Compute-bound} 
-	Program-Under-Test~\label{fig:meas_comp}}
+	\caption{{\color{blue}An Example of Measurement Results on a \linebreak \hbox{Compute-bound} 
+	Program-Under-Test}~\label{fig:meas_comp}}
 	\vspace{-.18in}
@@
 
-\shorten{
 \item We propose a novel timing protocol that
@@
 
+{\color{blue}
 \item We evaluate the performance of the protocol with rigorous experiments 
 using real-world programs.
+}
 
@@
 and eventually contributing to enhancing the overall measurement-quality}.
-}
-
-{\color{blue}
-%\item We present an algorithm to identify infrequent, long-running daemons impacting 
-%the timing of a compute-bound program.
-%
-%\item We propose a more accurate and precise protocol based on the identification variance 
-%variance for that program.
-
-\item We present an algorithm to identify and eliminate such infrequent and long-running 
-daemons.
-
-\item We propose a novel timing protocol that can considerably 
-reduce variance via the elimination.
-
-\item The evaluation results on real-world programs 
-show a support for the effectiveness of the protocol. 
-}
 
 \end{itemize}
-\vspace{-0.2in}
+\vspace{-0.05in}
 
 \noindent
+{\color{blue}
 The rest is organized as follows. 
 We next elaborate on the proposed timing protocol. In turn we evaluate the performance of the scheme using real workloads. 
+}
 
@@
 called {\em SEDONA} (Selective Elimination through Detection of infrequent, lOng-ruNning dAemons). 
+{\color{blue}
 Our scheme (1) utilizes PT (process time) to avoid absorbing timing noise from other co-running processes and (2)~identifies and eliminates executions including daemon processes that are infrequent and \hbox{long-running} via a {\em cutoff} measure.
+}
 
 Our SEDONA protocol consists of a total of ten steps, as described in Fig.~\ref{alg:find}.
-Note that this protocol (algorithm) is applicable to any arbitrary Linux system.
+{\color{blue} 
+Note that this protocol (algorithm) is applicable to any arbitrary Linux system.}
 
@@
 \STATE Step 1. Set up the timing environment.
-
-{\color{blue}
-\tcc{{\scriptsize Steps 2---9 produce cutoffs for each identified daemon.}}
 \STATE Step 2. Perform a single PUT run (specifically, PUT128) for many samples (specifically, 800).
@@
 \STATE Step 9. Compute the cutoffs for each identified daemon. 
-}
 \STATE Step 10. Discard an execution including a daemon of which process time is greater than the respective cutoff time.
@@
 {\bf A Running Example.} 
+{\color{blue}
 First, we set up the same measurement environment as described earlier in \hbox{Section}~\ref{sec:intro}, 
 to eliminate known timing factors (Step~1).
+}
 
-{\color{blue}
 We then run a program-under-test (called PUT) of 
@@
 
+{\color{blue}
 To identify such daemons, we use a novel \hbox{scatter} plot: 
@@
 to samples~\#{1} and~\#{2} of PUT128.
+}
 
@@
 from a run of the 800 PUT128 samples (Step~3). 
+{\color{blue}
 The $xy$-plane corresponds to the dual-PUT samples. 
@@
 the $x$ ($y$) axis is named as ``Odd (Even) Sample's ET.''
+}
 There are two quite obvious outliers with ETs 
@@
 are determined as infrequent, \hbox{long-running}.
+{\color{blue}
 We also identify ``extra'' infrequent daemons: 
@@
 (5x longer, or just about 20 hours). 
+}
 
@@
 We find some frequent daemon processes appearing in both of the 
-clusters of \hbox{dual-PUT256} and \hbox{dual-PUT32768}, each consisting of 
-pairs of two successive samples of PUT128 and PUT16384.
+clusters of \hbox{dual-PUT256} and \hbox{dual-PUT32768}{\color{blue}, each consisting of 
+pairs of two successive samples of PUT128 and PUT16384.}
 That said, the central cluster also contains other
@@
   infrequent}, several having \hbox{periodicities} estimated at four or twenty hours.
+{\color{blue}
 When PUT128 had a ``short'' 
@@
 hours are \hbox{infrequent}. But with PUT16384 
-with a ``long'' program time (or, 4.5 hours),
+with a ``long'' program time (or, 4.5 hours),} 
 some of those daemons are now frequent, and appear in the central cluster.
@@
 the cutoff for each of those infrequent, long-running daemons so
-identified, based on the runs of PUT128 and 
-PUT16384 as collected in Tab.~\ref{tab:final_infrequent_cutoff}. 
+identified, based on {\color{blue}the runs of PUT128 and 
+PUT16384} as collected in Tab.~\ref{tab:final_infrequent_cutoff}. 
 Here is how to compute the cutoff. For the cutoff of such a daemon with PUT128, 
@@
 (the last column of Tab.~\ref{tab:final_infrequent_cutoff}).
-}
 
@@
 
+{\color{blue}
 First, we compared the performance of the two approaches 
@@
 were increasingly widened, reaching up to by 6x.
+}
 
@@
 	\centering
-	\subfigure[SORT - Standard Deviation]{
+	\subfigure[{\color{blue}SORT - Standard Deviation}]{
 		\includegraphics[scale=0.31]{sort_std.eps}
@@
 	}
-	\subfigure[MM - Standard Deviation]{
+	\subfigure[{\color{blue}MM - Standard Deviation}]{
 		\includegraphics[scale=0.31]{matc_std.eps}
@@
 
+{\color{blue}
 Table~\ref{tab:spec_real} shows that 
@@
 For example, about a 10x \hbox{margin} between the two resulted from {\tt 434}.
+}
 
@@
 for the medium-length benchmarks (e.g. {\tt 447}, {\tt 456}, {\tt 470}, and {\tt 473}).
+{\color{blue}
 Even for the long-running benchmarks (e.g. {\tt 416}, {\tt 436}, and {\tt 454}, both $>$1,000 sec), 
 the relative error of \hbox{SEDONA} was still lower than that of ORG.
+}
 
@@
 than the ET-based existing method. 
+{\color{blue}
 The experimental results demonstrate 
-the general applicability of SEDONA to timing any CPU-bound program.
+the general applicability of SEDONA to timing any CPU-bound program.}
 
@@
 {{\tt 433}} & 480,586 & 473,256  & {743} & {725} & {2$\times$10$^{-3}$} & {2$\times$10$^{-3}$}\\ %% V
-{{\tt 434}} & 16,495  & 16,242  & {75} & {7} & {5$\times$10$^{-3}$} & {5$\times$10$^{-4}$}\\
+{{\tt 434}} & 16,495  & 16,242  & {\color{blue}{75}} & {\color{blue}{7}} & {{\color{blue}5$\times$10$^{-3}$}} & {{\color{blue}5$\times$10$^{-4}$}}\\
 {{\tt 435}} & 990,575 & 975,445  & {947} & {900} & {1$\times$10$^{-3}$} & {9$\times$10$^{-4}$}\\
@@
 
+{\color{blue}
 \section{Conclusion}
@@
 performed an empirical evaluation to show that 
-it is more precise and accurate than the extant method.
+it is more precise and accurate than the extant method.}
 

### 2017-04-02 sedona-ieice16/final/reviews/review_response.tex
Sedona revision

@@
 
-If I understand your question correctly, 
-the proposed {\em cutoff} method addresses your question. 
-We elaborate on deriving the cutoffs for each infrequent, long-running daemon 
-from Steps~2 to~9 in Figure~3, which is marked in blue 
-with a provided one-line comment above Step~2, for better clarification.
+If I interpret your question correctly,
+(1)~the identification process of infrequent, long-running daemons 
+and (2)~the proposed {\em cutoff} method address your question. 
+The cutoff is used as a criterion to determine whether to discard a sample, to improve measurement quality. 
+Both are elaborated from Steps~2 through~9 in Figure~3, augmented with a one-line comment (above Step~2), 
+for better clarification.
 
-Here is the summarized procedure of the cutoff estimation. 
-We first collect all the daemon processes 
-while running PUT128 800 times enough to fit in one day.
+%Here is the summary of how to estimate each daemon's influence as time via the cutoffs. 
+To estimate each daemon's influence, we do the following. 
+\begin{itemize}
+\item We run a program-under-test with 128 seconds, termed PUT128, 800 times 
+enough to experience an infrequent, long-running daemon 
+that may run every few hours or even once a day (Step~2).
+
+\item From the 800 samples, we render a scatter-plot of pairs of successive samples of elapsed time (ET) of PUT128
+as illustrated in Fig.~4~(b) and see if that scatter-plot has an $L$-shape (Step~3).
+
+\item We check if the central cluster emerges (Step~4) by zooming into the scatter-plot (Step~4).
+
+\item For each daemon in the central cluster, we compute the maximum and standard deviation 
+of the process time (PT) of that daemon (Step~5). (As specified in the Characteristics paragraph, PT is defined as 
+the sum of user and system time, supplied from the kernel via the Netlink interface, before and after each run.)
+
+\item (1)~We identify, for each daemon in the samples in the $L$-shape (or $L$-sample), 
+which daemon has its PT greater than two standard deviations above the maximum PT 
+of that daemon. (2)~We also collect some extra processes (e.g. {\tt bash}, {\tt rhn\_check}) 
+that are not in the cluster but only in the $L$-sample (Step~6). 
+The identified processes are treated as infrequent, long-running.
+
+\item For each so-identified infrequent, long-running daemon, we compute the minimum time 
+in the $L$-samples (Step~7). (This computation provides the valley between the maximum PT from 
+the central cluster and the minimum PT from the $L$-shape.)
+
+\item We repeat Steps 2---6 with the much-longer running of PUT16384, to confirm if 
+any of the identified daemons is actually not infrequent at the much larger task length of 16384 seconds (Step~8).
+This step is required to prevent the false positive case in our identification process, 
+as some daemons with a periodicity of hours are infrequent for a ``short'' program time like 128 seconds.
+
+\item We now compute the {\em cutoff} for each of those identified infrequent, long-running daemons, 
+based on the PUT128 and PUT16384 runs. 
+\end{itemize}
+
+%(The user- and system-time measures of captured processes are supplied by the kernel, via the Netlink interface.)
 

### 2017-04-02 sedona-ieice16/final/reviews/review_response.tex
Updates on SEDONA

--- 2017-03-31 sedona-ieice16/final/reviews/review_response.tex
@@
 
-If I understand your question correctly, 
-the proposed {\em cutoff} method addresses your question. 
-We elaborate on deriving the cutoffs for each infrequent, long-running daemon 
-from Steps~2 to~9 in Figure~3, which is marked in blue 
-with a provided one-line comment above Step~2, for better clarification.
+If I understand your question right, 
+we actually *measure* each daemon's process time (PT) for each run of a program-under-test, 
+rather than estimate its influence as time. 
+PT is computed based on the measures of user and system time received from the kernel via 
+the Netlink socket.
 
-Here is the summarized procedure of the cutoff estimation. 
-We first collect all the daemon processes 
-while running PUT128 800 times enough to fit in one day.
-
+To see whether the ``influence" (in your term) was significant on the timing, 
+using the PT data we first identify infrequent, long-running daemons and then compute the *cutoff* of each so-identified daemon. 
+As exhibited in Table~1, we obtain the detailed cutoff information for our identified infrequent, long-running daemon. 
+The computed cutoffs are used to determine whether to discard a sample involving 
+a daemon whose PT is above the respective cutoff, meaning that such a daemon's significant influence was there. 
+The identification and cutoff computation are described from Steps~2 through~9, marked in blue, in Figure~3. 
 For more details, please see the prose in Section~2 marked in blue.
 
+The validity of our protocol using the cutoff was assessed on real-world programs and the SPEC CPU benchmark. 
+As shown in Figure~5 and Table~3, we could obtain better measurement quality for those real workloads, compared to 
+the existing timing scheme using elapsed time.
+
+Note that to address your question, 
+I slightly modified the Contribution paragraph for better clarification, and provided 
+a single-line comment for better understanding about the cutoff computation.
+
+If I didn't catch your question correctly, please kindly let me know so that I will re-address it.
+
 \end{document}

### 2017-04-03 sedona-ieice16/final/reviews/review_response.tex
Some changes

--- 2017-03-31 sedona-ieice16/final/reviews/review_response.tex
@@
 
-If I understand your question correctly, 
-the proposed {\em cutoff} method addresses your question. 
-We elaborate on deriving the cutoffs for each infrequent, long-running daemon 
-from Steps~2 to~9 in Figure~3, which is marked in blue 
-with a provided one-line comment above Step~2, for better clarification.
+If I understand your question right, 
+we actually *measure* each daemon's process time (PT) for each timing (sample) of a program-under-test, 
+rather than estimate the influence of that daemon as time. 
+Remind that PT is computed based on the measures of user and system time received from the kernel via 
+the Netlink socket. To check whether the influence (in your terminology) was significant on each sample, 
+we came up with an idea of the {\em cutoff} method by 
+which we could eliminate a certain sample containing a daemon that ran over the respective cutoff time.
 
-Here is the summarized procedure of the cutoff estimation. 
-We first collect all the daemon processes 
-while running PUT128 800 times enough to fit in one day.
+To determine the cutoff associated with each daemon, we first identify infrequent, long-running daemons (\hbox{Steps~2---8}). 
+We then compute the cutoff of each so-identified daemon (Step~9). 
+The detailed cutoff computation is specified in the second paragraph in the right column on page~3. 
+The final cutoff data that we obtained for our identified daemon are exhibited in Table~1. 
+For more details about the identification and cutoff calculation, please revisit the blue text in Section~2.
 
-For more details, please see the prose in Section~2 marked in blue.
+We applied the table to discard a sample involving 
+a daemon whose PT is above the respective cutoff. Dropping a certain sample means that there happened such a daemon's significant influence on that sample. 
+This indicates that we can't trust and include that sample in the base data to calculate the execution time of the given program.
+
+The validity of our protocol using the cutoff was assessed on real-world programs (sort and matrix multiplication) and the SPEC CPU2006 benchmark. 
+As shown in Figure~5 and Table~3, our protocol obtained better measurement quality 
+for those real workloads, compared with that of the existing timing scheme using elapsed time.
+
+For better clarification (to resolve your question), I slightly modified the Contribution paragraph, 
+provided a single-line comment about the cutoff computation, and finally 
+added the last sentence in parentheses in the prose of Step~9. 
+These changes are marked in green. (Due to a space limit, more text could not be inserted.)
+
+I hope my response to your concern resolved your question. But if I misunderstood you, please kindly let me know. I will re-address.
 

### 2017-04-03 sedona-ieice16/final/paper.tex
Sedona revision
Updates on SEDONA
Some changes

@@
 
-{\color{blue}
+{\color{green}
 %\item We present an algorithm to identify infrequent, long-running daemons impacting 
@@
 
-\item We present an algorithm to identify and eliminate such infrequent and long-running 
-daemons.
+\item We present an algorithm to identify and eliminate such daemons that are infrequent, long-running.
 
@@
 {\color{blue}
-\tcc{{\scriptsize Steps 2---9 produce cutoffs for each identified daemon.}}
+\tcc{{\color{green}{\scriptsize Daemon-cutoff times are derived from Steps~2 through~9.}}}
 \STATE Step 2. Perform a single PUT run (specifically, PUT128) for many samples (specifically, 800).
@@
 to eliminate known timing factors (Step~1).
-
 {\color{blue}
@@
 (the last column of Tab.~\ref{tab:final_infrequent_cutoff}). 
+({\color{green}Note that different cutoff data from Tab.~\ref{tab:final_infrequent_cutoff} may be yielded on a different kernel.})
 }
 
-\vspace{-0.2in}
+\vspace{-0.3in}
 \begin{table}[ht]
@@
 
-%\vspace{-0.in}
+\vspace{-0.in}
 \begin{table}[t]
@@
 
-\vspace{-0.03in}
 {\tiny 
+%\vspace{-0.05in}
 \bibliographystyle{ieicetr} 

### 2017-04-03 sedona-ieice16/final/reviews/review_response.tex
Sedona finalization

@@
 \section*{Overview}\label{sec:overview}
-I appreciate your review and address your remaining question below. 
-Note that my response to your question is 
-incorporated into the revision and marked in {\color{blue}blue}.
+I appreciate your review and address the last question below. 
+Note that my response to that question is 
+reflected into the revision and marked in {\color{blue}blue}.
 
@@
 The validity of our protocol using the cutoff was assessed on real-world programs (sort and matrix multiplication) and the SPEC CPU2006 benchmark. 
-As shown in Figure~5 and Table~3, our protocol obtained better measurement quality 
+As shown in Fig.~5 and Tab.~3, our protocol obtained better measurement quality 
 for those real workloads, compared with that of the existing timing scheme using elapsed time.  
+For more details, please refer to the blue text in Section~3.
 
-For better clarification (to resolve your question), I slightly modified the Contribution paragraph, 
+For better clarification, I slightly modified the Contribution paragraph, 
 provided a single-line comment about the cutoff computation, and finally 
 added the last sentence in parentheses in the prose of Step~9. 
-These changes are marked in green. (Due to a space limit, more text could not be inserted.)
+These changes are marked in blue and incorporated into the revision.
 
-I hope my response to your concern resolved your question. But if I misunderstood you, please kindly let me know. I will re-address.
+I hope my response to your concern resolved your question. But if I misunderstood you, please kindly let me know. Many thanks for your great comments.
 

### 2017-04-03 sedona-ieice16/final/paper.tex
Sedona finalization

@@
 
-{\color{green}
+{\color{blue}
 %\item We present an algorithm to identify infrequent, long-running daemons impacting 
@@
 {\color{blue}
-\tcc{{\color{green}{\scriptsize Daemon-cutoff times are derived from Steps~2 through~9.}}}
+\tcc{{\color{blue}{\scriptsize Daemon-cutoff times are derived from Steps~2 through~9.}}}
 \STATE Step 2. Perform a single PUT run (specifically, PUT128) for many samples (specifically, 800).
@@
 (the last column of Tab.~\ref{tab:final_infrequent_cutoff}). 
-({\color{green}Note that different cutoff data from Tab.~\ref{tab:final_infrequent_cutoff} may be yielded on a different kernel.})
+({\color{blue}A different Linux distribution may yield different cutoff data than Tab.~\ref{tab:final_infrequent_cutoff}.})
 }
 
-\vspace{-0.3in}
+\vspace{-0.2in}
 \begin{table}[ht]
@@
 
+{\color{blue}
 Based on Tab.~\ref{tab:final_infrequent_cutoff}, 
-we discard any \hbox{sample} \hbox{containing} an infrequent, long-running daemon execution
-over the corresponding cutoff. We thus end up dropping just fifteen of the 800
+we discarded any \hbox{sample} \hbox{containing} an infrequent, long-running daemon execution
+over the corresponding cutoff.} We thus ended up dropping just fifteen of the 800
 PUT128 samples and only two of the forty PUT16384 samples. 
@@
 matrix multiplication (in column major), termed MM. 
-In this comparison SEDONA outperformed ORG, as can be seen in Fig.~\ref{fig:synprog_test}. 
+{\color{blue}
+In this comparison SEDONA bettered ORG, as can be seen in Fig.~\ref{fig:synprog_test}. 
 As workload size increased, the performance gap between the two 
 were increasingly widened, reaching up to by 6x.
+}
 
@@
 
+{\color{blue}
 Table~\ref{tab:spec_real} shows that 
-SEDONA bettered ORG,
+SEDONA outperformed ORG,
 on the standard deviation and relative error across the very different SPEC benchmarks.
+}
 To be more specific, all the benchmarks revealed a smaller standard deviation 
@@
 
+{\color{blue}
 SEDONA also scaled well for the SPEC workloads, 
 with regard to growth of relative error as the execution time lengthened.
+}
 For the short benchmarks 

### 2017-04-04 sedona-ieice16/final/paper.tex
SEDONA new version

--- 2017-03-31 sedona-ieice16/final/paper.tex
@@
 \title{SEDONA: A Novel Protocol for Identifying Infrequent, Long-Running Daemons on a Linux System}
-%\titlenote{This work was in part supported by the EDISON program (NRF-2011-0020576).}
+\titlenote{This work was in part supported by the EDISON program (NRF-2011-0020576). The author acknowledges Prof. Rick Snodgrass for his suggestions to improve this article.}
 \authorlist{
- \authorentry[yksuh@kisti.re.kr]{Young-Kyoon SUH}{m}{labelA}
+ \authorentry[yksuh@kisti.re.kr (Corresponding author)]{Young-Kyoon SUH}{m}{labelA}
  %\authorentry[rts@cs.arizona.edu]{Richard T. Snodgrass}{n}{labelB}}
@@
 
-Consider a simple compute-bound program, what we term {\em PUT} (program-under-test), 
-as shown in Fig.~\ref{alg:put}. 
-This program runs a nested for-loop with a specified task length ($tl$) (in seconds). 
-The $tl$ value is used to compute 
-the number of iterations ($t$) for which that for-loop is performed 
-to reach the specified task length.
+Fig.~\ref{fig:synprog_test} compared the performance of 
+the existing timing scheme, termed {\em ORG} using \hbox{end-to-end} (elapsed) time 
+and our timing protocol, termed {\em \hbox{SEDONA}} 
+(\hbox{Selective} Elimination through Detection of infrequent, \hbox{lOng-ruNning} dAemons), 
+to be proposed shortly. 
+In this comparison we used 
+two kinds of real-world programs such as insertion sort, termed SORT and 
+matrix multiplication (in column major), termed MM. 
+In this comparison SEDONA outperformed ORG in terms of measurement quality 
+gauged via standard deviation. 
+To be more specific, the performance gap between the two schemes over growing workload
+were increasingly widened, reaching up to by 6x.
 
-\vspace{-.2in}
-\begin{figure}[h]
-\begin{center}
-\begin{algorithmic}
-{\bf Algorithm} PerformManyIncrements($tl$): \\
-\STATE $t$ = $tl$ * {CONSTANT} \\
-\FOR{$k$ = 1 to $t$ by 1}
-	\FOR{$i$ = 1 to {UINT\_MAX}-1 by 1}
-		\STATE $j$ += 1 \\
-	\ENDFOR 
-\ENDFOR 
-\end{algorithmic}
-\end{center}
-\caption{Computation by a Program-Under-Test (PUT)\label{alg:put}}
-\vspace{-.2in}
+\begin{figure}[t]
+	\centering
+	\subfigure[SORT - Standard Deviation]{
+		\includegraphics[scale=0.31]{sort_std.eps}
+		\label{fig:sort_std}
+	}
+	\subfigure[MM - Standard Deviation]{
+		\includegraphics[scale=0.31]{matc_std.eps}
+		\label{fig:matc_std}
+	}
+	%\vspace{-.1in}
+	\caption{Performance Comparison on Real-world Programs~\label{fig:synprog_test}}
+\vspace{-.1in}
 \end{figure}
-
-We used 128~seconds as the task length and ran the PUT program (termed PUT128) 
-800~times. 
-Note that PUT128 was run on a machine, with 
-a specification illustrated in Table~\ref{tab:machine_config}, 
-in the same timing environment as used in our prior work~\cite{Currim}, 
-with (i)~non-critical daemons deactivated, 
-(ii)~the Network Timing Protocol daemon switched on,  
-(iii)~particular CPU features~\cite{intel15,intelSpeed15} turned off, 
-and (iv)~an up-to-date kernel installed.
-
-Fig.~\ref{fig:meas_comp} shows two plots of the timing results on PUT128.
-Most of the 800 executions measured in elapsed time, 
-via  {\tt System.currentTimeMillis()} in Java, form the central band 
-of \hbox{128,000--129,000 ms,} as can be seen in Fig.~\ref{fig:hist_before_et}. 
-Despite the well-controlled timing environment, there exist over a dozen of executions above that band, or {\em outliers}.
-Indeed, these outliers {\em critically} impact the overall measurement quality: a standard deviation of 1,730 ms and a relative error of 0.01.
-This result suggests adequately eliminating these outliers while retaining most of the executions.
-
-Our protocol, which we characterize shortly, on the contrary, 
-exactly identified and removed those outliers. 
-In effect, the protocol considerably reduced such variability, 
-as illustrated in Fig.~\ref{fig:hist_after_pt}. 
-Specifically, regarding standard deviation and relative error 
-the protocol presented a significant performance gain up to by 
-three orders of magnitude for the same PUT128 results.
 
@@
 
-\begin{figure}[t]
-	\centering
-	\subfigure[Initial Elapsed Time Measurement]{
-		\includegraphics[width=0.224\textwidth]{128_sec_et_old.eps}
-		\label{fig:hist_before_et}
-	}
-	\subfigure[Program Time After Our Protocol]{
-		\includegraphics[width=0.224\textwidth]{128_sec_pt_new.eps}
-		\label{fig:hist_after_pt}
-	}
-	\caption{An Example of Measurement Results on a \linebreak \hbox{Compute-bound} 
-	Program-Under-Test~\label{fig:meas_comp}}
-	\vspace{-.18in}
-\end{figure}
-
 Commercial software tools measure execution time~\cite{VTune,TimeSys,WindView}. 
@@
 
-\vspace{0.05in}
+%\vspace{0.05in}
 {\bf Contribution.} 
 Our contributions are following.
-\vspace{-0.05in}
+%\vspace{-0.05in}
 \begin{itemize}
@@
 
-\shorten{
-\item We propose a novel timing protocol that
-identifies infrequent, long-running daemons that impact the timing results for that program.
-
-\item We evaluate the performance of the protocol with rigorous experiments 
-using real-world programs.
-
-\item The experimental results show a support for the effectiveness of our scheme\shorten{SEDONA: 
-early eliminating dirty executions infected with such a daemon 
-and eventually contributing to enhancing the overall measurement-quality}. 
-}
-
-{\color{blue}
 %\item We present an algorithm to identify infrequent, long-running daemons impacting 
 %the timing of a compute-bound program.
-%
-%\item We propose a more accurate and precise protocol based on the identification variance 
-%variance for that program.
 
-\item We present an algorithm to identify and eliminate such infrequent and long-running 
-daemons.
+\item We present an algorithm to identify and eliminate such daemons that are infrequent, \hbox{long-running} 
+and thus impact the timing of a given program.
 
@@
 show a support for the effectiveness of the protocol.
-}
 
 \end{itemize}
-\vspace{-0.2in}
+%\vspace{-0.2in}
 
@@
 
-In this section we propose a novel execution time measurement scheme, 
-called {\em SEDONA} (Selective Elimination through Detection of infrequent, lOng-ruNning dAemons). 
-Our scheme (1) utilizes PT (process time) to avoid absorbing timing noise from other co-running processes and (2)~identifies and eliminates executions including daemon processes that are infrequent and \hbox{long-running} via a {\em cutoff} measure.
+In this section we propose our SEDONA timing protocol. 
+SEDONA (1) utilizes PT (process time) to avoid absorbing timing noise from other co-running processes and (2)~identifies and eliminates executions including daemon processes that are infrequent and \hbox{long-running} via a {\em cutoff} measure.
 
@@
 \caption{Summary of the SEDONA Timing Protocol\label{alg:find}}
-\vspace{-0.25in}
 \end{figure}
@@
 {\bf A Running Example.} 
-First, we set up the same measurement environment as described earlier in \hbox{Section}~\ref{sec:intro}, 
-to eliminate known timing factors (Step~1).
+In Step~1 we configure the same timing environment as used in our prior work~\cite{Currim}, 
+to eliminate known timing factors. 
+The environment requires (i)~deactivating non-critical daemons, 
+(ii)~switching on the Network Timing Protocol daemon,  
+(iii)~turning off particular CPU features~\cite{intel15,intelSpeed15}, 
+and (iv)~getting an up-to-date kernel installed.
 
-{\color{blue}
-We then run a program-under-test (called PUT) of 
-Fig.~\ref{alg:put} with a task length of 128~sec, 
-termed {\em PUT128}, 800 times (Step 2). 
-(We render Fig.~\ref{fig:hist_before_et} using the 800 samples from this run.)
+We then run a simple program-under-test (called PUT) many times, as shown in Fig.~\ref{alg:put} (Step~2). 
+PUT runs a nested for-loop with a specified task length ($tl$) (in seconds). 
+The $tl$ value is used to compute 
+the number of iterations ($t$) for which that for-loop is performed 
+to reach the specified task length.
+
+\vspace{-.2in}
+\begin{figure}[h]
+\begin{center}
+\begin{algorithmic}
+{\bf Algorithm} PerformManyIncrements($tl$): \\
+\STATE $t$ = $tl$ * {CONSTANT} \\
+\FOR{$k$ = 1 to $t$ by 1}
+	\FOR{$i$ = 1 to {UINT\_MAX}-1 by 1}
+		\STATE $j$ += 1 \\
+	\ENDFOR 
+\ENDFOR 
+\end{algorithmic}
+\end{center}
+\caption{Computation by a Program-Under-Test (PUT)\label{alg:put}}
+\vspace{-.2in}
+\end{figure}
+
+The run of Step~2 assigns a task length of 128~sec to PUT, termed {\em PUT128}, and repeats executing PUT128 800~times.
 We use 128 seconds because that is long enough to perhaps experience an infrequent daemon. 
@@
 
-%Fig.~\ref{fig:reg_put128} (= Fig.~\ref{fig:hist_before_et}) 
-%plots all the 800 elapsed times (ETs) of PUT128.
-%The plot clearly shows three rows; that is, 
-%the top and middle rows represent over a dozen of outliers far from 
-%the rest of the samples clustered in the bottom row. 
-%We will now drill down into these outliers to 
-%show how to reliably eliminate the indirect influence of 
-%some ``infrequent, \hbox{long-running} daemons'' on PT (\hbox{process} time) of the PUT.
 Fig.~\ref{fig:reg_put128} plots all the 800 elapsed times (ETs) of the run of PUT128.
@@
 some ``infrequent, \hbox{long-running} daemons'' on PT (\hbox{process} time) of the PUT.
+
+\begin{figure}[t]
+	\centering
+	\subfigure[PUT128 with 800 Samples]{
+		\includegraphics[width=0.224\textwidth]{put128_empv4_new.eps}
+		\label{fig:reg_put128}
+	}
+	\subfigure[Dual-PUT256 Samples]{
+		\includegraphics[width=0.224\textwidth]{dual_et_put256_empv4.eps}
+		\label{fig:raw_put128}
+	}\vspace{-0.15in}
+	\subfigure[Zooming in on the $L$-shape]{
+		\includegraphics[width=0.224\textwidth]{zooming_in_dual_et_put256_empv4.eps}
+		\label{fig:dual_put256_liters}
+	}
+	\subfigure[The Central Cluster]{
+		\includegraphics[width=0.224\textwidth]{dual_et_clump_put256_empv4.eps}
+		\label{fig:dual_et_put256_cluster}
+        }
+    \caption{Successive Scatter Plots of PUT128 with 800 samples 
+    (equivalent to Dual-PUT256 with 400 samples) in Steps~2---4~\label{fig:put128_plot}} 
+    %\vspace{-0.3in}     
+\end{figure}
 
@@
 of 163,913 ms (rightmost) and 161,785 ms (uppermost), respectively.
-
-\begin{figure}[ht]
-	\centering
-	\subfigure[PUT128 with 800 Samples]{
-		\includegraphics[width=0.224\textwidth]{put128_empv4_new.eps}
-		\label{fig:reg_put128}
-	}
-	\subfigure[Dual-PUT256 Samples]{
-		\includegraphics[width=0.224\textwidth]{dual_et_put256_empv4.eps}
-		\label{fig:raw_put128}
-	}\vspace{-0.15in}
-	\subfigure[Zooming in on the $L$-shape]{
-		\includegraphics[width=0.224\textwidth]{zooming_in_dual_et_put256_empv4.eps}
-		\label{fig:dual_put256_liters}
-	}
-	\subfigure[The Central Cluster]{
-		\includegraphics[width=0.224\textwidth]{dual_et_clump_put256_empv4.eps}
-		\label{fig:dual_et_put256_cluster}
-        }
-    \caption{Successive Scatter Plots of PUT128 with 800 samples 
-    (equivalent to Dual-PUT256 with 400 samples) in Steps~2---4~\label{fig:put128_plot}} 
-    \vspace{-0.3in}     
-\end{figure}
 
@@
 (the last column of Tab.~\ref{tab:final_infrequent_cutoff}). 
-}
+(Note that a different distribution than RedHat Enterprise Linux may yield different cutoff data from those of Tab.~\ref{tab:final_infrequent_cutoff}, but this cutoff idea is applicable to any Linux distribution and arbitrary program.)
 
-\vspace{-0.2in}
-\begin{table}[ht]
+\vspace{-0.1in}
+\begin{table}[h]
 \centering
-{\tiny
-\begin{tabular}{|p{1.2cm}|c|c|p{1cm}|p{1.45cm}|} \hline
+{\scriptsize
+\begin{tabular}{|p{1.2cm}|p{1.2cm}|p{1.4cm}|p{1cm}|p{1.45cm}|} \hline
 {\tiny Process}  & {\tiny Cutoff PT} & {\tiny Cutoff PT} & {\tiny Task}  & {\tiny{\bf Final }} \\
@@
   Process Time \hbox{(Step~9)}\label{tab:final_infrequent_cutoff}}
-  \vspace{-0.3in}
+  \vspace{-0.15in}
 \end{table}
@@
 Based on Tab.~\ref{tab:final_infrequent_cutoff}, 
-we discard any \hbox{sample} \hbox{containing} an infrequent, long-running daemon execution
-over the corresponding cutoff. We thus end up dropping just fifteen of the 800
-PUT128 samples and only two of the forty PUT16384 samples. 
-As a result, the \hbox{overall} \hbox{measurement} quality---the standard deviation and \hbox{relative} error 
----for PUT128 was improved 
-by three orders of magnitude (as shown in Fig.~\ref{fig:meas_comp}) 
-and by about two \hbox{orders} of magnitude for PUT16384.
+SEDONA discards any \hbox{sample} \hbox{containing} an infrequent, long-running daemon execution
+over the corresponding cutoff.
+
+The following section presents the evaluation results of SEDONA via a popular compute-bound benchmark suite.
 
@@
 \label{sec:eval}
-\vspace{-0.07in}
+%\vspace{-0.07in}
 We now evaluate the \hbox{performance} of SEDONA 
@@
 \begin{table}[h]
-\vspace{-0.2in}
+\vspace{-0.1in}
 \begin{center}
-{\tiny
+{%\tiny
 \begin{tabular}{|l|p{7cm}|}\hline
@@
 \caption{Machine Configurations\label{tab:machine_config}}
-\vspace{-0.35in}
+\vspace{-0.2in}
 \end{table}
 
-First, we compared the performance of the two approaches 
-using real-world programs such as insertion sort, termed SORT and 
-matrix multiplication (in column major), termed MM. 
-In this comparison SEDONA outperformed ORG, as can be seen in Fig.~\ref{fig:synprog_test}. 
-As workload size increased, the performance gap between the two 
-were increasingly widened, reaching up to by 6x.
-
-\vspace{-.2in}
-\begin{figure}[h]
-	\centering
-	\subfigure[SORT - Standard Deviation]{
-		\includegraphics[scale=0.31]{sort_std.eps}
-		\label{fig:sort_std}
-	}
-	\subfigure[MM - Standard Deviation]{
-		\includegraphics[scale=0.31]{matc_std.eps}
-		\label{fig:matc_std}
-	}
-	\vspace{-.1in}
-	\caption{Performance Comparison on Real-world Programs~\label{fig:synprog_test}}
-\vspace{-.25in}
-\end{figure}
-
-In turn, our evaluation proceeded with SPEC CPU2006 
-benchmarks~\cite{specCpu2006}, providing various \linebreak 
-\hbox{compute-bound} real applications. The results are \hbox{provided} in Table~\ref{tab:spec_real}. 
+The validity of our protocol was evaluated with SPEC CPU2006 
+benchmarks~\cite{specCpu2006}, providing various 
+\hbox{compute-bound} real applications (as well as the real-world programs 
+as shown in Fig.~\ref{fig:synprog_test}).
+The results are \hbox{provided} in Table~\ref{tab:spec_real}. 
 Note that in the table the results for {\tt 481} and {\tt 483} \hbox{benchmarks} 
@@
 Table~\ref{tab:spec_real} shows that 
-SEDONA bettered ORG,
+SEDONA outperformed ORG,
 on the standard deviation and relative error across the very different SPEC benchmarks.
@@
 
-%\vspace{-0.in}
 \begin{table}[t]
 \centering
-{
-\tiny
+{\tiny
 \begin{tabular}{|p{0.1cm}|R{1.1cm}|R{1.1cm}|R{0.6cm}|R{0.6cm}|R{0.92cm}|R{0.92cm}|} \hline
@@
 \label{sec:conclusion}
-\vspace{-0.07in}
+%\vspace{-0.07in}
 We proposed a novel timing protocol called {\em SEDONA}, and 
@@
 
-\balance
+%\balance
 
-\vspace{-0.03in}
-{\tiny 
+%{\tiny 
+%\vspace{-0.05in}
 \bibliographystyle{ieicetr} 
@@
 %\vspace{-0.2in}
-}
+%}
 

### 2017-04-05 sedona-ieice16/final/paper.tex
SEDONA updates

@@
 
+{\color{blue}
 Fig.~\ref{fig:synprog_test} compared the performance of 
@@
 to be proposed shortly. 
-In this comparison we used 
+For this comparison we used 
 two kinds of real-world programs such as insertion sort, termed SORT and 
 matrix multiplication (in column major), termed MM. 
-In this comparison SEDONA outperformed ORG in terms of measurement quality 
-gauged via standard deviation. 
-To be more specific, the performance gap between the two schemes over growing workload
-were increasingly widened, reaching up to by 6x.
+As illustrated in Fig.~\ref{fig:synprog_test}, 
+SEDONA outperformed ORG with regard to measurement quality by standard deviation. 
+To be more specific, we observed that as workload size increased the performance gap between the two methods 
+was increasingly bigger, reaching up to by six~times. 
+}
 
@@
 	\caption{Performance Comparison on Real-world Programs~\label{fig:synprog_test}}
-\vspace{-.1in}
+\vspace{-.3in}
 \end{figure}
 
+\vspace{-0.15in}
 {\bf Characteristics.} There are several important specifics of our protocol. 
@@
 {\em infrequent, \hbox{long-running} daemon processes}. 
-We observe that such daemon processes substantially impact 
+We witness that such daemon processes substantially impact 
 the execution time of the process.  
@@
 
-%\vspace{0.05in}
+%\vspace{0.02in}
 {\bf Contribution.} 
 Our contributions are following.
-%\vspace{-0.05in}
+\vspace{-0.1in}
 \begin{itemize}
@@
 
+{\color{blue}
 %\item We present an algorithm to identify infrequent, long-running daemons impacting 
@@
 show a support for the effectiveness of the protocol. 
+}
 
 \end{itemize}
-%\vspace{-0.2in}
+\vspace{-0.2in}
 
@@
 
-In this section we propose our SEDONA timing protocol. 
-SEDONA (1) utilizes PT (process time) to avoid absorbing timing noise from other co-running processes and (2)~identifies and eliminates executions including daemon processes that are infrequent and \hbox{long-running} via a {\em cutoff} measure.
+{\color{blue}
+In this section we propose our timing protocol. 
+The protocol (1) utilizes PT (process time) to avoid absorbing timing noise from other co-running processes and (2)~identifies and eliminates executions including daemon processes that are infrequent and \hbox{long-running} via a {\em cutoff} measure.
 
-Our SEDONA protocol consists of a total of ten steps, as described in Fig.~\ref{alg:find}.
-Note that this protocol (algorithm) is applicable to any arbitrary Linux system.
+Our SEDONA protocol consists of the three major steps, as described in Fig.~\ref{alg:find}. 
+Note that this protocol (algorithm) is applicable to any arbitrary Linux system and to any compute-bound program.
+}
 
+\vspace{-0.2in}
 \begin{figure}[h]
@@
 {\bf Algorithm} The SEDONA Timing Protocol: \\
-\STATE Step 1. Set up the timing environment.
-\STATE Step 2. Perform a single PUT run (specifically, PUT128) for many samples (specifically, 800).
-\STATE Step 3. Consider each pair of elapsed time measurements to be a dual-PUT measurement 
-and examine a scatter-plot to see if it it displays an $L$-shape.
-\STATE Step 4. Zoom into the central cluster to ensure that it is symmetric (roughly circular).
-\STATE Step 5. Compute the maximum and standard deviation of the process time 
-for each daemon encountered within the central cluster samples.
-\STATE Step 6. Identify for each sample in the $L$-shape infrequent, long-running daemon executions. 
-\STATE Step 7. Determine potentially periodic daemons based on the $L$-executions 
-and for each daemon compute the minimum process time from those executions identified. 
-\STATE Step 8. Perform Steps 1--6 above for a single run 
-consisting of a small number of executions (specifically, 40) of PUT16384.  
-\STATE Step 9. Compute the cutoffs for each identified daemon. 
-\STATE Step 10. Discard an execution including a daemon of which process time is greater than the respective cutoff time.
+{\color{blue}
+\STATE {\bf Step~A. Set up the timing environment.}
+\STATE {\bf Step~B. Determine the cutoffs.}
+\STATE  \ B-1. Perform a single run of a simple program-under-test (PUT) (specifically, PUT128) for many samples (specifically, 800).
+\STATE  \ B-2. Consider each pair of elapsed time measurements to be a dual-PUT measurement and examine a scatter-plot to see if it it displays an $L$-shape.
+\STATE  \ B-3. Zoom into the central cluster to ensure that it is symmetric (roughly circular).
+\STATE  \ B-4. Compute the maximum and standard deviation of the process time for each daemon encountered within the central cluster samples.
+\STATE  \ B-5. Identify for each sample in the $L$-shape infrequent, long-running daemon executions. 
+\STATE  \ B-6. Determine potentially periodic daemons based on the $L$-executions and for each daemon compute the minimum process time from those executions identified. 
+\STATE  \ B-7. Perform Steps 1--6 above for a single run consisting of a small number of executions (specifically, 40) of PUT16384.  
+\STATE  \ B-8. Compute the cutoffs for each identified daemon. 
+\STATE {\bf Step~C. Time an arbitrary compute-bound program.}
+\STATE  \ C-1. Run the program ten times, collecting all the process times (PTs). 
+\STATE  \ C-2. Discard any run involving daemon executions over the cutoffs.
+\STATE  \ C-3. Compute the average to get the PT of that program.
+}
 \end{algorithmic}
 \end{center}
-\caption{Summary of the SEDONA Timing Protocol\label{alg:find}}
+\caption{The SEDONA Timing Protocol\label{alg:find}}
 \end{figure}
 
-{\bf A Running Example.} 
-In Step~1 we configure the same timing environment as used in our prior work~\cite{Currim}, 
+%\vspace{-0.2in}
+%{\color{blue}
+%{\bf Determining the Cutoffs.} 
+In Step~A we configure the same timing environment as used in our prior work~\cite{Currim}, 
 to eliminate known timing factors. 
@@
 
-We then run a simple program-under-test (called PUT) many times, as shown in Fig.~\ref{alg:put} (Step~2). 
+Step~B consists of eight sub-steps, resulting in determining 
+the cutoffs for each identified infrequent, long-running daemon. 
+In Step~B-1, we run a simple program-under-test (called PUT) many times, as shown in Fig.~\ref{alg:put}. 
 PUT runs a nested for-loop with a specified task length ($tl$) (in seconds). 
@@
 
-\vspace{-.2in}
 \begin{figure}[h]
@@
 \caption{Computation by a Program-Under-Test (PUT)\label{alg:put}}
-\vspace{-.2in}
+%\vspace{-.2in}
 \end{figure}
@@
     (equivalent to Dual-PUT256 with 400 samples) in Steps~2---4~\label{fig:put128_plot}} 
-    %\vspace{-0.3in}     
+    \vspace{-0.3in}     
 \end{figure}
@@
 %\vspace{-0.07in}
-We now evaluate the \hbox{performance} of SEDONA 
-compared to that of the original timing scheme, termed ORG, based on 
-elapsed time.
-Our experiments were conducted on a \hbox{machine}
+{\color{blue}
+We now evaluate the \hbox{performance} of SEDONA, 
+compared to that of the original timing scheme (ORG) based on elapsed time.
+Our experiments were conducted on a commodity \hbox{machine}
 described in Table~\ref{tab:machine_config}. 
 \begin{table}[h]
-\vspace{-0.1in}
+\vspace{-0.2in}
 \begin{center}
@@
 
-The validity of our protocol was evaluated with SPEC CPU2006 
-benchmarks~\cite{specCpu2006}, providing various 
-\hbox{compute-bound} real applications (as well as the real-world programs 
-as shown in Fig.~\ref{fig:synprog_test}).
-The results are \hbox{provided} in Table~\ref{tab:spec_real}. 
-Note that in the table the results for {\tt 481} and {\tt 483} \hbox{benchmarks} 
-are omitted because of some runtime error and incurred I/O, respectively.
+As already mentioned in Section~\ref{sec:intro}, we first 
+confirmed the validity of our protocol with two 
+real \hbox{CPU-bound} programs---insertion sort and \hbox{column-major} matrix multiplication---with heavy memory references.
 
-Table~\ref{tab:spec_real} shows that 
-SEDONA outperformed ORG,
+To further assess the performance of the protocol, 
+we proceeded with SPEC CPU2006 benchmarks~\cite{specCpu2006}, providing various 
+\hbox{compute-bound} real applications.
+%The results are \hbox{provided} in Table~\ref{tab:spec_real}. 
+%Note that in the table the results for {\tt 481} and {\tt 483} \hbox{benchmarks} 
+%are omitted because of some runtime error and incurred I/O, respectively.
+Note that in this evaluation we could not obtain the results for {\tt 481} and {\tt 483} \hbox{benchmarks}. 
+{\tt 481} threw some runtime error that we could not resolve, and {\tt 483} incurred I/O, which was out of scope of this article.
+Hence, those two results are omitted in Table~\ref{tab:spec_real}.
+
+We reaffirmed from Table~\ref{tab:spec_real} that SEDONA outperformed ORG,
 on the standard deviation and relative error across the very different SPEC benchmarks.
-To be more specific, all the benchmarks revealed a smaller standard deviation 
+Specifically, all the benchmarks revealed a smaller standard deviation 
 from SEDONA as compared to that of ORG.
@@
 lower than that of ORG for almost every benchmark.
-For example, about a 10x \hbox{margin} between the two resulted from {\tt 434}.
+Roughly a 10x margin between the two schemes resulted from the {\tt 434} benchmark, for instance.
+}
 
@@
 \caption{Performance Evaluation on the SPEC Benchmarks\label{tab:spec_real}}
-\vspace{-0.25in}
+\vspace{-0.3in}
 \end{table}
+
+%\vspace\fill
 
@@
 \label{sec:conclusion}
-%\vspace{-0.07in}
+\vspace{-0.1in}
+
 We proposed a novel timing protocol called {\em SEDONA}, and 
@@
 it is more precise and accurate than the extant method. 
+This protocol is generic, in that {\em any} arbitrary \hbox{CPU-bound} program 
+can be timed with reduced variance on {\em any} Linux distribution, via the protocol.
 
-%\balance
+\balance
 
-%{\tiny 
-%\vspace{-0.05in}
+{\scriptsize
+%\vspace{-0.2in}
 \bibliographystyle{ieicetr} 
 \bibliography{paper}
-%\vspace{-0.2in}
-%}
+%\vspace{-0.1in}
+}
 

### 2017-04-06 sedona-ieice16/final/reviews/review_response.tex
SEDONA done

--- 2017-04-02 sedona-ieice16/final/reviews/review_response.tex
@@
 \section*{Overview}\label{sec:overview}
-I appreciate your review and address your remaining question below. 
-Note that my response to your question is 
-incorporated into the revision and marked in {\color{blue}blue}.
+I appreciate your review and address the last question below. 
+Note that I mark all the changes made to this revision to respond to your concern are marked in {\color{blue}blue}. 
+I hope that these changes helped you better read the flow of the paper.
 
@@
 
-If I understand your question right, 
-we actually *measure* each daemon's process time (PT) for each run of a program-under-test, 
-rather than estimate its influence as time. 
-PT is computed based on the measures of user and system time received from the kernel via 
-the Netlink socket.
+If I understood your question right, 
+each daemon's influence is translated to its respective cutoff, which I proposed and derived in this paper. 
+For the detailed cutoffs, please refer to Tab.~1 and for computation process (estimation method in your term), the second paragraph in the right column on page~3.
+As mentioned in the paper, the cutoffs are used to drop any execution (sample) involving daemon executions 
+over their respective cutoffs. This drop indicates that such an execution was *significantly influenced* by infrequent, long-running daemons and thus we cannot include 
+that execution in the computation of the execution time of a given program.
 
-To see whether the ``influence" (in your term) was significant on the timing, 
-using the PT data we first identify infrequent, long-running daemons and then compute the *cutoff* of each so-identified daemon. 
-As exhibited in Table~1, we obtain the detailed cutoff information for our identified infrequent, long-running daemon. 
-The computed cutoffs are used to determine whether to discard a sample involving 
-a daemon whose PT is above the respective cutoff, meaning that such a daemon's significant influence was there. 
-The identification and cutoff computation are described from Steps~2 through~9, marked in blue, in Figure~3. 
-For more details, please see the prose in Section~2 marked in blue.
+To assess the validity of my protocol using the cutoffs, I used the SPEC CPU2006 benchmark suite and observed that the protocol 
+outperformed the extant method using elapsed time, as exhibited in Tab.~3 on page~4 (as already indicated by the previous revision). 
+Note also that I took out Fig.~1 and instead moved Fig.~5 in the former revision (now Fig.~1) upfront (on page~1) in this revision, 
+to early draw the attention of the protocol's effectiveness on two real programs (insertion sort and matrix multiplication) with increasing workloads.
 
-The validity of our protocol using the cutoff was assessed on real-world programs and the SPEC CPU benchmark. 
-As shown in Figure~5 and Table~3, we could obtain better measurement quality for those real workloads, compared to 
-the existing timing scheme using elapsed time.
+Last but not least, there seems a confusion from the used program-under-test (PUT) (e.g., PUT128 and PUT16384) 
+for not only defining but also evaluating the protocol. (In fact, the PUT is also an arbitrary \hbox{compute-bound} program, 
+so there's no problem with applying the protocol to that. But I acknowledge that the dual use of the PUT potentially confuses readers including you.)
+To unravel the confusion, I have redesigned the protocol, divided into the three major steps, as shown in Fig.~2 on page~2. 
+In the new design, I talk about the PUT, only in the context of protocol definition (Step~B), but do not use it for evaluation. 
+As described above, the evaluation was performed on those two real-world programs and the SPEC CPU benchmarks.
 
-Note that to address your question, 
-I slightly modified the Contribution paragraph for better clarification, and provided 
-a single-line comment for better understanding about the cutoff computation.
+%If I understand your question right, 
+%we actually *measure* each daemon's process time (PT) for each timing (sample) of a program-under-test, 
+%rather than estimate the influence of that daemon as time. 
+%Remind that PT is computed based on the measures of user and system time received from the kernel via 
+%the Netlink socket. To check whether the influence (in your terminology) was significant on each sample, 
+%we came up with an idea of the {\em cutoff} method by 
+%which we could eliminate a certain sample containing a daemon that ran over the respective cutoff time. 
+%
+%To determine the cutoff associated with each daemon, we first identify infrequent, long-running daemons (\hbox{Steps~2---8}). 
+%We then compute the cutoff of each so-identified daemon (Step~9). 
+%The detailed cutoff computation is specified in the second paragraph in the right column on page~3. 
+%The final cutoff data that we obtained for our identified daemon are exhibited in Table~1. 
+%For more details about the identification and cutoff calculation, please revisit the blue text in Section~2. 
+%
+%We applied the table to discard a sample involving 
+%a daemon whose PT is above the respective cutoff. Dropping a certain sample means that there happened such a daemon's significant influence on that sample. 
+%This indicates that we can't trust and include that sample in the base data to calculate the execution time of the given program. 
+%
+%The validity of our protocol using the cutoff was assessed on real-world programs (sort and matrix multiplication) and the SPEC CPU2006 benchmark. 
+%As shown in Fig.~5 and Tab.~3, our protocol obtained better measurement quality 
+%for those real workloads, compared with that of the existing timing scheme using elapsed time.  
+%For more details, please refer to the blue text in Section~3.
+%
+%For better clarification, I slightly modified the Contribution paragraph, 
+%provided a single-line comment about the cutoff computation, and finally 
+%added the last sentence in parentheses in the prose of Step~9. 
+%These changes are marked in blue and incorporated into the revision. 
+%
 
-If I didn't catch your question correctly, please kindly let me know so that I will re-address it.
+If I'm misunderstood, please kindly let me know. I'll re-address. Many thanks for your great comments.
 

### 2017-04-06 sedona-ieice16/final/paper.tex
SEDONA done

@@
 \STATE  \ B-4. Compute the maximum and standard deviation of the process time for each daemon encountered within the central cluster samples.
-\STATE  \ B-5. Identify for each sample in the $L$-shape infrequent, long-running daemon executions. 
-\STATE  \ B-6. Determine potentially periodic daemons based on the $L$-executions and for each daemon compute the minimum process time from those executions identified. 
-\STATE  \ B-7. Perform Steps 1--6 above for a single run consisting of a small number of executions (specifically, 40) of PUT16384.  
+\STATE  \ B-5. Identify for each sample in the $L$-shape infrequent, \hbox{long-running} daemon executions. 
+\STATE  \ B-6. Determine potentially periodic daemons based on the \hbox{$L$-executions} and for each daemon compute the minimum process time from those executions identified. 
+\STATE  \ B-7. Perform Steps 1--B-6 above for a single run consisting of a small number of executions (specifically, 40) of PUT16384.  
 \STATE  \ B-8. Compute the cutoffs for each identified daemon. 
@@
 \end{center}
+\vspace{-0.05in}
 \caption{The SEDONA Timing Protocol\label{alg:find}}
+\vspace{-0.2in}
 \end{figure}
@@
 %\vspace{-0.2in}
-%{\color{blue}
-%{\bf Determining the Cutoffs.} 
-In Step~A we configure the same timing environment as used in our prior work~\cite{Currim}, 
+{\color{blue}
+{\bf Timing configuration.} 
+Step~A sets up the same timing environment as used in our prior work~\cite{Currim}, 
 to eliminate known timing factors. 
@@
 
+{\bf Determining the Cutoffs}
 Step~B consists of eight sub-steps, resulting in determining 
 the cutoffs for each identified infrequent, long-running daemon.
+
 In Step~B-1, we run a simple program-under-test (called PUT) many times, as shown in Fig.~\ref{alg:put}. 
@@
 to reach the specified task length.
+}
 
@@
 \caption{Computation by a Program-Under-Test (PUT)\label{alg:put}}
-%\vspace{-.2in}
+\vspace{-0.25in}
 \end{figure}
 
-The run of Step~2 assigns a task length of 128~sec to PUT, termed {\em PUT128}, and repeats executing PUT128 800~times.
-We use 128 seconds because that is long enough to perhaps experience an infrequent daemon. 
-We run it 800 times to capture infrequent daemons that perhaps run every few hours or even 
-once a day. 
+{\color{blue}
+We assign a task length of 128~sec to PUT, termed {\em PUT128} and run PUT128 800~times.
+We use 128~seconds because that is long enough to perhaps experience an infrequent daemon. 
+We execute PUT128 800~times to capture infrequent daemons that perhaps run every few hours or even 
+once a day.} 
 Note that we collect all daemon processes as well as the PUT and their measures 
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
+	\vspace{-0.3in} 
 	\centering
@@
     \caption{Successive Scatter Plots of PUT128 with 800 samples 
-    (equivalent to Dual-PUT256 with 400 samples) in Steps~2---4~\label{fig:put128_plot}} 
-    \vspace{-0.3in}     
+    (equivalent to Dual-PUT256 with 400 samples) in {\color{blue}Steps~B-1--3}~\label{fig:put128_plot}} 
+    \vspace{-0.25in}     
 \end{figure}
@@
 of 400 samples of a run of \hbox{dual-PUT256} constructed 
-from a run of the 800 PUT128 samples (Step~3). 
+from a run of the 800 PUT128 samples ({\color{blue}Step~B-2}). 
 The $xy$-plane corresponds to the dual-PUT samples. 
@@
 We continue zooming until we get to Figure~\ref{fig:dual_et_put256_cluster},
-which shows a central cluster (Step 4). 
+which shows a central cluster ({\color{blue}Step~B-3}). 
 We confirm the symmetry of the ET measurements in the central cluster: there
@@
 
-We then perform Step~5, which computes 
+We then perform {\color{blue}Step~B-4}, which computes 
 the maximum process time and standard deviation 
@@
 
-In Step~6 we identify, for each daemon in the \hbox{$L$-samples}, 
+{\color{blue}Step~B-5} identifies, for each daemon in the \hbox{$L$-samples}, 
 those that are actual long-running daemon
@@
 Next, we can compute for each so-identified infrequent, long-running daemon its
-minimum time in the \hbox{$L$-samples} (Step 7). 
+minimum time in the \hbox{$L$-samples} ({\color{blue}Step~B-6}). 
 This computation provides a rough, initial distinction of a ``long-running''
@@
 
-In Step 8 we repeat Steps 1--6, but instead with the much-longer running
+In {\color{blue}Step~B-7} we repeat {\color{blue}Steps~1--B-6}, but instead with the much-longer running
 PUT16384 (4.5 hours per \hbox{sample} versus 2 minutes), to see 
@@
 
-In Step~9 we compute 
+In {\color{blue}Step~B-8} we compute 
 the cutoff for each of those infrequent, long-running daemons so
@@
 (the last column of Tab.~\ref{tab:final_infrequent_cutoff}). 
-(Note that a different distribution than RedHat Enterprise Linux may yield different cutoff data from those of Tab.~\ref{tab:final_infrequent_cutoff}, but this cutoff idea is applicable to any Linux distribution and arbitrary program.)
+%(Note that a different distribution than RedHat Enterprise Linux may yield different cutoff data from those of Tab.~\ref{tab:final_infrequent_cutoff}, but this cutoff idea is applicable to any Linux distribution and arbitrary program.)
 
-\vspace{-0.1in}
+\vspace{-0.2in}
 \begin{table}[h]
@@
 \caption{Collected Infrequent, Long-running Daemons and Their Final Cutoff
-  Process Time \hbox{(Step~9)}\label{tab:final_infrequent_cutoff}}
-  \vspace{-0.15in}
+  Process Time \hbox{({\color{blue}Step~B-8})}\label{tab:final_infrequent_cutoff}}
+  \vspace{-0.25in}
 \end{table}
 
-Based on Tab.~\ref{tab:final_infrequent_cutoff}, 
-SEDONA discards any \hbox{sample} \hbox{containing} an infrequent, long-running daemon execution
-over the corresponding cutoff.
+{\color{blue}
+{\bf Time an arbitrary compute-bound program.} 
+Step~C measures the execution time of a given program. 
+In Step~C-1, we execute that program ten times~\cite{Currim}, collecting all the process times (PTs).
+We then drop any execution involving daemons of which PTs are over 
+their respective cutoffs in Tab.~\ref{tab:final_infrequent_cutoff} (Step~C-2).
+Finally, \hbox{Step~C-3} calculates the PT of that program as the average among the retained executions' PTs.
 
-The following section presents the evaluation results of SEDONA via a popular compute-bound benchmark suite.
+%Now, we report the evaluation results of SEDONA. 
+}
 
@@
 \label{sec:eval}
-%\vspace{-0.07in}
+\vspace{-0.07in}
 {\color{blue}
@@
 described in Table~\ref{tab:machine_config}. 
+\vspace{-0.3in}
 \begin{table}[h]
-\vspace{-0.2in}
 \begin{center}
-{%\tiny
-\begin{tabular}{|l|p{7cm}|}\hline
+{\tiny
+\begin{tabular}{|l|p{6.8cm}|}\hline
 OS & Red Hat Ent. Linux (RHEL) 6.4 with a kernel of 2.6.32 \\ \hline
@@
 \caption{Machine Configurations\label{tab:machine_config}}
-\vspace{-0.2in}
+\vspace{-0.3in}
 \end{table}
@@
 
-We reaffirmed from Table~\ref{tab:spec_real} that SEDONA outperformed ORG,
+Consequently, we reaffirmed from Table~\ref{tab:spec_real} that \hbox{SEDONA} bettered ORG,
 on the standard deviation and relative error across the very different SPEC benchmarks.
@@
 \label{sec:conclusion}
-\vspace{-0.1in}
+\vspace{-0.05in}
 

### 2017-04-07 sedona-ieice16/final/paper.tex
SEDONA finalization

@@
 thus making it hard to trust their validity. 
-In this paper we propose a novel timing protocol to significantly 
-reduce such variability.
+{\color{blue}We propose a novel timing protocol to significantly 
+reduce such variability by eliminating executions involving infrequent, \hbox{long-running} daemons.}
 
@@
 
+\vspace{-0.05in}
+
 \section{Introduction}
 \label{sec:intro}
+\vspace{-0.05in}
 
@@
 {\color{blue}
-Fig.~\ref{fig:synprog_test} compared the performance of 
-the existing timing scheme, termed {\em ORG} using \hbox{end-to-end} (elapsed) time 
-and our timing protocol, termed {\em \hbox{SEDONA}} 
-(\hbox{Selective} Elimination through Detection of infrequent, \hbox{lOng-ruNning} dAemons), 
+Fig.~\ref{fig:synprog_test} compares the performance of 
+the existing timing scheme, termed {\em ORG}, using elapsed time and our timing protocol, termed {\em \hbox{SEDONA}} 
+(\hbox{Selective} \hbox{Elimination} through Detection of infrequent, \hbox{lOng-ruNning} dAemons), 
 to be proposed shortly. 
 For this comparison we used 
-two kinds of real-world programs such as insertion sort, termed SORT and 
+two kinds of real-world programs, specifically insertion sort, termed SORT and 
 matrix multiplication (in column major), termed MM. 
@@
 SEDONA outperformed ORG with regard to measurement quality by standard deviation. 
-To be more specific, we observed that as workload size increased the performance gap between the two methods 
-was increasingly bigger, reaching up to by six~times. 
+To be more specific, we observed that as workload size increased the performance gap between the two methods also increased, 
+reaching up to by six~times. 
 }
@@
 	}
-	%\vspace{-.1in}
+	\vspace{-.15in}
 	\caption{Performance Comparison on Real-world Programs~\label{fig:synprog_test}}
-\vspace{-.3in}
+\vspace{-.25in}
 \end{figure}
 
-\vspace{-0.15in}
+%\vspace{-0.15in}
 {\bf Characteristics.} There are several important specifics of our protocol. 
@@
 Our contributions are following.
-\vspace{-0.1in}
+\vspace{-0.05in}
 \begin{itemize}
@@
 
-\item The evaluation results on real-world programs 
-show a support for the effectiveness of the protocol. 
+\item An evaluation on real-world programs 
+shows a support for the effectiveness of the protocol. 
 }
@@
 \end{itemize}
-\vspace{-0.2in}
+\vspace{-0.05in}
 
@@
 The rest is organized as follows. 
-We next elaborate on the proposed timing protocol. In turn we evaluate the performance of the scheme using real workloads.
+We next elaborate on the proposed timing protocol. 
+{\color{blue}In turn we assess the performance of the protocol using a popular industrial benchmark suite. 
+%Finally, we conclude this letter by summarizing our discussion on the proposed protocol.}
+Finally, we summarize our discussion.}
 
@@
 
-{\bf Determining the Cutoffs}
+{\bf Determining the Cutoffs.}
 Step~B consists of eight sub-steps, resulting in determining 
@@
 
-As already mentioned in Section~\ref{sec:intro}, we first 
+As already seen in Figure~\ref{fig:synprog_test}, we 
 confirmed the validity of our protocol with two 
@@
 
-To further assess the performance of the protocol, 
+To assess the performance of the protocol, 
 we proceeded with SPEC CPU2006 benchmarks~\cite{specCpu2006}, providing various 
@@
 
-Consequently, we reaffirmed from Table~\ref{tab:spec_real} that \hbox{SEDONA} bettered ORG,
+Consequently, we affirmed from Table~\ref{tab:spec_real} that \hbox{SEDONA} bettered ORG,
 on the standard deviation and relative error across the very different SPEC benchmarks.

### 2017-04-07 sedona-ieice16/final/paper_black.tex
SEDONA finalization

--- 2017-04-07 sedona-ieice16/final/paper.tex
@@
 thus making it hard to trust their validity. 
-{\color{blue}We propose a novel timing protocol to significantly 
-reduce such variability by eliminating executions involving infrequent, \hbox{long-running} daemons.}
+We propose a novel timing protocol to significantly 
+reduce such variability by eliminating executions involving infrequent, \hbox{long-running} daemons.
 
@@
 
-{\color{blue}
 Fig.~\ref{fig:synprog_test} compares the performance of 
@@
 reaching up to by six~times.
-}
 
@@
 
-%\vspace{0.02in}
 {\bf Contribution.} 
@@
 
-{\color{blue}
-%\item We present an algorithm to identify infrequent, long-running daemons impacting 
-%the timing of a compute-bound program.
-
 \item We present an algorithm to identify and eliminate such daemons that are infrequent, \hbox{long-running} 
@@
 shows a support for the effectiveness of the protocol.
-}
 
@@
 We next elaborate on the proposed timing protocol. 
-{\color{blue}In turn we assess the performance of the protocol using a popular industrial benchmark suite. 
-%Finally, we conclude this letter by summarizing our discussion on the proposed protocol.}
-Finally, we summarize our discussion.}
+In turn we assess the performance of the protocol using a popular industrial benchmark suite. 
+Finally, we summarize our discussion.
 
@@
 
-{\color{blue}
 In this section we propose our timing protocol. 
@@
 Note that this protocol (algorithm) is applicable to any arbitrary Linux system and to any compute-bound program.
-}
 
@@
 {\bf Algorithm} The SEDONA Timing Protocol: \\
-{\color{blue}
 \STATE {\bf Step~A. Set up the timing environment.}
@@
 \STATE  \ C-3. Compute the average to get the PT of that program.
-}
 \end{algorithmic}
@@
 
-%\vspace{-0.2in}
-{\color{blue}
 {\bf Timing configuration.} 
@@
 to reach the specified task length.
-}
 
@@
 
-{\color{blue}
 We assign a task length of 128~sec to PUT, termed {\em PUT128} and run PUT128 800~times.
@@
 We execute PUT128 800~times to capture infrequent daemons that perhaps run every few hours or even 
-once a day.} 
+once a day.
 Note that we collect all daemon processes as well as the PUT and their measures 
@@
     \caption{Successive Scatter Plots of PUT128 with 800 samples 
-    (equivalent to Dual-PUT256 with 400 samples) in {\color{blue}Steps~B-1--3}~\label{fig:put128_plot}} 
+    (equivalent to Dual-PUT256 with 400 samples) in Steps~B-1--3~\label{fig:put128_plot}} 
     \vspace{-0.25in}     
@@
 of 400 samples of a run of \hbox{dual-PUT256} constructed 
-from a run of the 800 PUT128 samples ({\color{blue}Step~B-2}). 
+from a run of the 800 PUT128 samples (Step~B-2). 
 The $xy$-plane corresponds to the dual-PUT samples. 
@@
 We continue zooming until we get to Figure~\ref{fig:dual_et_put256_cluster},
-which shows a central cluster ({\color{blue}Step~B-3}). 
+which shows a central cluster (Step~B-3). 
 We confirm the symmetry of the ET measurements in the central cluster: there
@@
 
-We then perform {\color{blue}Step~B-4}, which computes 
+We then perform Step~B-4, which computes 
 the maximum process time and standard deviation 
@@
 
-{\color{blue}Step~B-5} identifies, for each daemon in the \hbox{$L$-samples}, 
+Step~B-5 identifies, for each daemon in the \hbox{$L$-samples}, 
 those that are actual long-running daemon
@@
 Next, we can compute for each so-identified infrequent, long-running daemon its
-minimum time in the \hbox{$L$-samples} ({\color{blue}Step~B-6}). 
+minimum time in the \hbox{$L$-samples} (Step~B-6). 
 This computation provides a rough, initial distinction of a ``long-running''
@@
 
-In {\color{blue}Step~B-7} we repeat {\color{blue}Steps~1--B-6}, but instead with the much-longer running
+In Step~B-7 we repeat Steps~1--B-6, but instead with the much-longer running
 PUT16384 (4.5 hours per \hbox{sample} versus 2 minutes), to see 
@@
 
-In {\color{blue}Step~B-8} we compute 
+In Step~B-8 we compute 
 the cutoff for each of those infrequent, long-running daemons so
@@
 \caption{Collected Infrequent, Long-running Daemons and Their Final Cutoff
-  Process Time \hbox{({\color{blue}Step~B-8})}\label{tab:final_infrequent_cutoff}}
+  Process Time \hbox{(Step~B-8)}\label{tab:final_infrequent_cutoff}}
   \vspace{-0.25in}
@@
 
-{\color{blue}
 {\bf Time an arbitrary compute-bound program.} 
@@
 
-%Now, we report the evaluation results of SEDONA. 
-}
-
 \section{Evaluation}
@@
 \vspace{-0.07in}
-{\color{blue}
 We now evaluate the \hbox{performance} of SEDONA, 
@@
 Roughly a 10x margin between the two schemes resulted from the {\tt 434} benchmark, for instance.
-}
 

### 2017-04-12 pt_characterization/rep_check/paper.tex
repeatability test draft

@@
 If so then, we can say that run repeatability is satisfied in our experiment setting. 
+In this study the used protocol is equivalent to SEDONA, eliminating samples involving infrequent, long-running daemons, which is the first step of EMPv5.
+In other words, the protocol skips the second step
+of EMPv5 that removes samples with process times over/below two standard deviations 
+from the average after performing the first step.
 
-\section{Experiment Notes}
-Table~\ref{tab:exp_notes} provides a short description of our experimental runs, 
-on which the following histograms are based.
-
-%% done
-\begin{table}[h]
-\begin{center}
-\begin{tabular}{|p{2cm}|p{3cm}|p{5cm}|p{5cm}|} \hline
-Machine & Task Length (sec) & Description & Experiment Period\\ \hline
-{\tt sodb9} &  INC1$\sim$INC64 & Two runs with 1000 samples & 2017-03-02 $\sim$ 2017-03-?? / 2017-03-?? $\sim$ 2017-03-??\\ \hline
-{\tt sodb9} &  INC128$\sim$INC1024 & Two runs with 300 samples & 2017-03-02 $\sim$ 2017-03-?? / 2017-03-?? $\sim$ 2017-03-??\\ \hline
-{\tt sodb10} & INC2048 & Two runs of 300 samples & 2017-03-02 $\sim$ 2017-03-09 / 2017-03-13 $\sim$ 2017-03-20\\ \hline
-{\tt sodb12} & INC4096 & Two runs of 300 samples & 2017-02-13 $\sim$ 2017-02-27 / 2017-03-02 $\sim$ 2017-03-17\\ \hline
-\end{tabular}
-\end{center}
-\vspace{-.2in}
-\caption{Notes on experiment runs used for histograms\label{tab:exp_notes}}
-\end{table}
-
-Now we show histograms of elapsed time (ET) and process time (PT) of INC.
+Now we show histograms of elapsed time (ET) and process time (PT) of INC via the SEDONA protocol.
 

### 2017-04-12 pt_characterization/rep_check/res1.tex
repeatability test draft

@@
 INC with its task length increasing from 1 second to 4096 seconds, via EMPv5. 
-The detailed description of the base data is from Table~\ref{tab:exp_notes}.
+The detailed description of the base data is from Table~\ref{tab:exp_notes1}.
+
+\section{Experiment Notes}
+Table~\ref{tab:exp_notes1} provides a short description of our experimental runs, 
+on which the following histograms are based.
+
+%% done
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{2cm}|p{3cm}|p{3cm}|p{4cm}|p{3cm}|} \hline
+Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
+{\tt sodb9} &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-02 $\sim$ 2017-03-04 & Figs.~\ref{fig:s9_r1_et_hist1},~\ref{fig:s9_r1_et_hist2},~\ref{fig:s9_r1_pt_hist1}, and~\ref{fig:s9_r1_pt_hist2}\\ \hline
+{\tt sodb9} &  INC128$\sim$INC1024 & 300 samples, each & 2017-03-04 $\sim$ 2017-03-11 & 
+Figs.~\ref{fig:s9_r1_et_hist3} and~\ref{fig:s9_r1_pt_hist3}\\ \hline
+{\tt sodb10} & INC2048 & 300 samples & 2017-03-02 $\sim$ 2017-03-09 & Figs.~\ref{fig:inc2048_r1_et_hist_v5} and~\ref{fig:inc2048_r1_hist_v5}\\ \hline
+{\tt sodb12} & INC4096 & 300 samples & 2017-02-13 $\sim$ 2017-02-27 & Figs.~\ref{fig:inc4096_r1_et_hist_v5} and~\ref{fig:inc4096_r1_hist_v5}\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Notes on experiment runs used for histograms\label{tab:exp_notes1}}
+\end{table}
+
+\pagebreak
 

### 2017-04-12 pt_characterization/rep_check/res2.tex
repeatability test draft

@@
 INC with its task length increasing from 1 second to 4096 seconds, via EMPv5. 
-The detailed description of the base data is from Table~\ref{tab:exp_notes}.
+The detailed description of the base data is from Table~\ref{tab:exp_notes2}.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{2cm}|p{3cm}|p{3cm}|p{4cm}|p{3.5cm}|} \hline
+Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
+{\tt sodb9} &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-13 $\sim$ 2017-03-14 & Figs.~\ref{fig:s9_r2_et_hist1},~\ref{fig:s9_r2_et_hist2},~\ref{fig:s9_r2_pt_hist1}, and~\ref{fig:s9_r2_pt_hist2}\\ \hline
+{\tt sodb9} &  INC128$\sim$INC1024 & 300 samples, each & 2017-03-14 $\sim$ 2017-03-21 & 
+Figs.~\ref{fig:s9_r2_et_hist3} and~\ref{fig:s9_r2_pt_hist3}\\ \hline
+{\tt sodb10} & INC2048 & 300 samples & 2017-03-13 $\sim$ 2017-03-20 & Figs.~\ref{fig:inc2048_r2_et_hist_v5} and~\ref{fig:inc2048_r2_hist_v5}\\ \hline
+{\tt sodb12} & INC4096 & 300 samples & 2017-03-02 $\sim$ 2017-03-17 & Figs.~\ref{fig:inc4096_r2_et_hist_v5} and~\ref{fig:inc4096_r2_hist_v5}\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Notes on experiment runs used for histograms\label{tab:exp_notes2}}
+\end{table}
+
+\pagebreak
 
@@
 \end{figure}
+
+\clearpage
+\pagebreak
+\subsubsection{Analysis}
+In this section we look into what happened inside the peaks observed in a certain histogram. 
+We consider Figure~\ref{fig:inc8_r2_hist_v5} for this study. 
+In the figure, we see the peaks at 8015 msec, 8020 msec, and 8021 msec.
+
+Table~\ref{tab:inc8_daemons} shows captured daemons and their runtime statistics per bin 
+of figure. Note that bin is at the unit of PT. 
+It appears that the peaks are definitely correlated with (1) appearances of some daemons 
+and (2) times that those daemons co-ran with INC8.
+
+\begin{table}[h]
+\begin{center}
+{\scriptsize
+\begin{tabular}{l|l|l|l|l|l|l|l} \hline
+TASK\_LEN  & BIN (PT) &  DAEMON   & MIN\_PT  &  MAX\_PT  &  AVG\_PT  &   STD\_PT & Counts \\ \hline
+INC8     & 8013  & jbd2/md0-8    & 1     & 1     & 1     & 0     & 1\\ \hline
+INC8     & 8013  & kslowd000     & 1     & 1     & 1     & 0     & 1\\ \hline
+INC8     & 8013  & md0\_raid1     & 1     & 1     & 1     & 0     & 17\\ \hline
+INC8     & 8013  & proc\_monitor  & 196   & 200   & 197.72        & 1.07  & 18\\ \hline \hline
+
+INC8     & 8014  & jbd2/md0-8    & 1     & 1     & 1     & 0     & 5\\ \hline
+INC8     & 8014  & kslowd000     & 1     & 1     & 1     & 0     & 35\\ \hline
+INC8     & 8014  & kslowd001     & 1     & 1     & 1     & 0     & 26\\ \hline
+INC8     & 8014  & md0\_raid1     & 1     & 1     & 1     & 0     & 58\\ \hline
+INC8     & 8014  & proc\_monitor  & 196   & 200   & 197.31        & 1.06  & 95\\ \hline \hline
+
+INC8     & {\bf 8015}  & java  & 2     & 7     & 4.5   & 3.54  & 2\\ \hline
+INC8     & {\bf 8015}  & jbd2/md0-8    & 1     & 1     & 1     & 0     & 2\\ \hline
+INC8     & {\bf 8015}  & kslowd000     & 1     & 1     & 1     & 0     & 86\\ \hline
+INC8     & {\bf 8015}  & kslowd001     & 1     & 1     & 1     & 0     & 89\\ \hline
+INC8     & {\bf 8015}  & md0\_raid1     & 1     & 1     & 1     & 0     & 18\\ \hline
+INC8     & {\bf 8015}  & proc\_monitor  & 196   & 200   & 197.28        & 1.01  & 194\\ \hline\hline
+
+INC8     & 8016  & kslowd000     & 1     & 1     & 1     & 0     & 36\\ \hline
+INC8     & 8016  & kslowd001     & 1     & 1     & 1     & 0     & 40\\ \hline
+INC8     & 8016  & md0\_raid1     & 1     & 1     & 1     & 0     & 8\\ \hline
+INC8     & 8016  & proc\_monitor  & 196   & 200   & 196.45        & .95   & 78\\ \hline\hline
+
+INC8     & 8017  & kslowd000     & 1     & 1     & 1     & 0     & 11\\ \hline
+INC8     & 8017  & kslowd001     & 1     & 1     & 1     & 0     & 10\\ \hline
+INC8     & 8017  & md0\_raid1     & 1     & 1     & 1     & 0     & 3\\ \hline
+INC8     & 8017  & proc\_monitor  & 196   & 200   & 197.15        & 1.16  & 26\\ \hline\hline
+
+INC8     & 8018  & kslowd000     & 1     & 1     & 1     & 0     & 13\\ \hline
+INC8     & 8018  & kslowd001     & 1     & 1     & 1     & 0     & 9\\ \hline
+INC8     & 8018  & md0\_raid1     & 1     & 1     & 1     & 0     & 6\\ \hline
+INC8     & 8018  & proc\_monitor  & 196   & 200   & 197.24        & 1.27  & 29\\ \hline\hline
+
+INC8     & 8019  & jbd2/md0-8    & 1     & 1     & 1     & 0     & 3\\ \hline
+INC8     & 8019  & kslowd000     & 1     & 1     & 1     & 0     & 9\\ \hline
+INC8     & 8019  & kslowd001     & 1     & 2     & 1.06  & .24   & 18\\ \hline
+INC8     & 8019  & md0\_raid1     & 1     & 1     & 1     & 0     & 27\\ \hline
+INC8     & 8019  & proc\_monitor  & 196   & 200   & 197.1         & 1.18  & 52\\ \hline\hline
+
+INC8     & {\bf 8020}  & jbd2/md0-8    & 1     & 1     & 1     & 0     & 8\\ \hline
+INC8     & {\bf 8020}  & kslowd000     & 1     & 1     & 1     & 0     & 52\\ \hline
+INC8     & {\bf 8020}  & kslowd001     & 1     & 1     & 1     & 0     & 57\\ \hline
+INC8     & {\bf 8020}  & md0\_raid1     & 1     & 1     & 1     & 0     & 91\\ \hline
+INC8     & {\bf 8020}  & proc\_monitor  & 196   & 200   & 197.03        & 1.02  & 180\\ \hline\hline
+
+INC8     & {\bf 8021}  & cifsd         & 1     & 1     & 1     & 0     & 1\\ \hline
+INC8     & {\bf 8021}  & java  & 2     & 37    & 19.5  & 24.75         & 2\\ \hline
+INC8     & {\bf 8021}  & kslowd000     & 1     & 1     & 1     & 0     & 146\\ \hline
+INC8     & {\bf 8021}  & kslowd001     & 1     & 1     & 1     & 0     & 143\\ \hline
+INC8     & {\bf 8021}  & md0\_raid1     & 1     & 1     & 1     & 0     & 11\\ \hline
+INC8     & {\bf 8021}  & proc\_monitor  & 196   & 198   & 197.15        & .98   & 299\\ \hline\hline
+
+INC8     & 8022  & kslowd000     & 1     & 1     & 1     & 0     & 20\\ \hline
+INC8     & 8022  & kslowd001     & 1     & 1     & 1     & 0     & 9\\ \hline
+INC8     & 8022  & proc\_monitor  & 196   & 198   & 196.07        & .37   & 29\\ \hline\hline
+\end{tabular}
+}
+\end{center}
+\caption{Daemons observed from the INC8 run~\label{tab:inc8_daemons}}
+\end{table}

### 2017-04-12 pt_characterization/recent_data/sodb10_10k/hist.r

@@ -1,0 +1,41 @@
+### sodb10
+numSamples <- 10000
+
+x = read.csv(file="16_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("16_sec_et_hist_v5.eps")
+#x <- subset(x, x$ITERNUM != 686 & x$ITERNUM != 700)
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
+h = hist(x$METIME, right=F, breaks=nbins,plot=F)
+xmin <-min(x$METIME)
+xmax <-max(x$METIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+#ymax <- 100
+plot(h, freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-3,xmax+2),col="blue", main='ET frequency on INC16', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='ET (ms)', ylab=expression('Frequency'))
+dev.off()
+
+x = read.csv(file="16_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("16_sec_pt_hist_v5.eps")
+#x <- subset(x, x$ITERNUM != 686 & x$ITERNUM != 700)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+h = hist(x$PRTIME, right=F,breaks=max(x$PRTIME)-min(x$PRTIME)+1,plot=F)
+binsize=1
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+#ymax <- 100
+plot(h, ylim=c(0,ymax), axes = TRUE,freq=TRUE,xlim=c(min(x$PRTIME)-1,max(x$PRTIME)+3),col="green", main='PT frequency on INC16', 
[…]

### 2017-04-12 pt_characterization/rep_check/repet_data1/et_hist_sodb9.r

@@ -7,18 +7,22 @@ postscript("1_sec_et_hist_v5.eps")
 binsize=1
 x_up = mean(x$METIME) + 2*sd(x$METIME)
 x_dn = mean(x$METIME) - 2*sd(x$METIME)
-x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
-nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
-h = hist(x$METIME, right=F, breaks=nbins,plot=F)
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+nbins <- ceiling((max(x$METIME)-min(x$METIME)+1) / binsize)
+h = hist(x$METIME, right=F, breaks=nbins+1,plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
 xmin <-min(x$METIME)
 xmax <-max(x$METIME)
-plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC1', 
-sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
-xlab='ET (ms)', ylab=expression('Frequency'))
-xaxtl <- seq(xmin-1,xmax+1,by=1)
-axis(side=1, at=xaxtl, labels=xaxtl)
+plot(h, xaxt="n",freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC1', sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), xlab='ET (ms)', ylab=expression('Frequency'))
+#xaxtl <- seq(xmin-1,xmax+1,by=1)
+#axis(side=1, at=xaxtl, labels=xaxtl)
+#xaxtl <- seq(floor(min(h$mids)),ceiling(max(h$mids)-1),by=binsize)
+#xaxtl <- seq(floor(min(h$mids)),ceiling(max(h$mids)),by=binsize)
+#axis(side=1, at=h$breaks, labels=xaxtl)
+#yaxtl <- seq(0,ymax,by=50)
+#axis(side=2, at=yaxtl, labels=yaxtl)
+axis(side=1,at=h$mids,labels=seq(xmin,xmax-1))
 dev.off()
 
 x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
[…]

### 2017-04-12 pt_characterization/rep_check/repet_data1/non-centered/et_hist_sodb9.r

@@ -1,0 +1,279 @@
+### sodb9
+numSamples <- 1000
+x = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("1_sec_et_hist_v5.eps")
+binsize=1
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
+h = hist(x$METIME, right=F, breaks=nbins,plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+xmin <-min(x$METIME)
+xmax <-max(x$METIME)
+plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC1', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='ET (ms)', ylab=expression('Frequency'))
+xaxtl <- seq(xmin-1,xmax+1,by=1)
+axis(side=1, at=xaxtl, labels=xaxtl)
+dev.off()
+
+x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("2_sec_et_hist_v5.eps")
+binsize=1
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
+h = hist(x$METIME, right=F, breaks=nbins,plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+xmin <-min(x$METIME)
+xmax <-max(x$METIME)
+plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC2', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
[…]

### 2017-04-12 pt_characterization/rep_check/repet_data2/et_hist_sodb9.r

@@ -7,18 +7,18 @@ postscript("1_sec_et_hist_v5.eps")
 binsize=1
 x_up = mean(x$METIME) + 2*sd(x$METIME)
 x_dn = mean(x$METIME) - 2*sd(x$METIME)
-x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
 nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
 h = hist(x$METIME, right=F, breaks=nbins,plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
 xmin <-min(x$METIME)
 xmax <-max(x$METIME)
-plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC1', 
+plot(h, xaxt="n",freq=TRUE,ylim=c(0,ymax), col="blue", main='ET frequency on INC1', 
 sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
 xlab='ET (ms)', ylab=expression('Frequency'))
 xaxtl <- seq(xmin-1,xmax+1,by=1)
-axis(side=1, at=xaxtl, labels=xaxtl)
+axis(side=1,at=h$mids,labels=seq(xmin,xmax-1))
 dev.off()
 
 x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
@@ -28,18 +28,18 @@ postscript("2_sec_et_hist_v5.eps")
 binsize=1
 x_up = mean(x$METIME) + 2*sd(x$METIME)
 x_dn = mean(x$METIME) - 2*sd(x$METIME)
-x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
 nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
 h = hist(x$METIME, right=F, breaks=nbins,plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
 xmin <-min(x$METIME)
 xmax <-max(x$METIME)
-plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC2', 
[…]

### 2017-04-12 pt_characterization/rep_check/repet_data2/non-centered/et_hist_sodb9.r

@@ -1,0 +1,281 @@
+### sodb9
+numSamples <- 1000
+x = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("1_sec_et_hist_v5.eps")
+binsize=1
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
+h = hist(x$METIME, right=F, breaks=nbins,plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+xmin <-min(x$METIME)
+xmax <-max(x$METIME)
+plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC1', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='ET (ms)', ylab=expression('Frequency'))
+xaxtl <- seq(xmin-1,xmax+1,by=1)
+axis(side=1, at=xaxtl, labels=xaxtl)
+dev.off()
+
+x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("2_sec_et_hist_v5.eps")
+binsize=1
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
+h = hist(x$METIME, right=F, breaks=nbins,plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+xmin <-min(x$METIME)
+xmax <-max(x$METIME)
+plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC2', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
[…]

### 2017-04-12 pt_characterization/rep_check/repet_data2/pt_hist_sodb9.r

@@ -7,18 +7,17 @@ postscript("1_sec_pt_hist_v5.eps")
 binsize=1
 x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
 x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
-x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
 nbins <- ceiling((max(x$PRTIME)-min(x$PRTIME)) / binsize)
 h = hist(x$PRTIME, right=F, breaks=nbins,plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
 xmin <-min(x$PRTIME)
 xmax <-max(x$PRTIME)
-plot(h, axes = TRUE,freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="green", main='PT frequency on INC1', 
+plot(h, xaxt="n",freq=TRUE,ylim=c(0,ymax), col="green", main='PT frequency on INC1', 
 sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
 xlab='PT (ms)', ylab=expression('Frequency'))
-xaxtl <- seq(xmin-1,xmax+1,by=1)
-axis(side=1, at=xaxtl, labels=xaxtl)
+axis(side=1,at=h$mids,labels=seq(xmin,xmax-1))
 dev.off()
 
 x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
@@ -28,7 +27,7 @@ postscript("2_sec_pt_hist_v5.eps")
 binsize=1
 x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
 x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
-x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
 nbins <- ceiling((max(x$PRTIME)-min(x$PRTIME)) / binsize)
 h = hist(x$PRTIME, right=F, breaks=nbins,plot=F)
 ymax <- max(h$counts)
@@ -36,11 +35,11 @@ ymax <- ceiling(ymax/100)*100
 #ymax <- 250
 xmin <-min(x$PRTIME)
 xmax <-max(x$PRTIME)
[…]

### 2017-04-21 pt_characterization/rep_check/10k_run.tex
new draft for repeatability

\section{Histograms with 10,000 samples~\label{sec:sodb9_10k_hist}} 
This section exhibits histograms on two runs of 
INC, each with 8 and 16 seconds as its task length, having 10,000 repetitions.
The detailed description of the base data is from Table~\ref{tab:exp_notes3}.

\begin{table}[h]
\begin{center}
\begin{tabular}{|p{2cm}|p{3cm}|p{3cm}|p{4cm}|p{3.5cm}|} \hline
Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
{\tt sodb9} &  INC8 & 10000 samples & 2017-03-29 $\sim$ 2017-03-30 & Figs.~\ref{fig:inc8_10k_et_hist_v5} and~\ref{fig:inc8_10k_pt_hist_v5}\\ \hline
{\tt sodb10} &  INC16 & 10000 samples & 2017-03-29 $\sim$ 2017-03-31 & 
Figs.~\ref{fig:inc16_10k_et_hist_v5} and~\ref{fig:inc16_10k_pt_hist_v5}\\ \hline
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{Notes on experiment runs used for histograms\label{tab:exp_notes3}}
\end{table}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC8]{
		\includegraphics[scale=0.43]{10k_run/8_sec_et_hist_v5.eps}
		\label{fig:inc8_10k_et_hist_v5}
	}
	\subfigure[PT frequency on INC8]{
		\includegraphics[scale=0.43]{10k_run/8_sec_pt_hist_v5.eps}
		\label{fig:inc8_10k_pt_hist_v5}
	}
	\subfigure[ET frequency on INC16]{
		\includegraphics[scale=0.43]{10k_run/16_sec_et_hist_v5.eps}
		\label{fig:inc16_10k_et_hist_v5}
	}
	\subfigure[PT frequency on INC16]{
		\includegraphics[scale=0.43]{10k_run/16_sec_pt_hist_v5.eps}
		\label{fig:inc16_10k_pt_hist_v5}
	}
	\caption{Histograms of INC8 and INC16 with 10,000 samples~\label{fig:10k_run_hist}}
\end{figure}

### 2017-04-21 pt_characterization/rep_check/res1.tex
new draft for repeatability

@@
 This section exhibits histograms on the first run of 
-INC with its task length increasing from 1 second to 4096 seconds, via EMPv5. 
+INC with its task length increasing from 1 second to 4096 seconds, via SEDONA. 
 The detailed description of the base data is from Table~\ref{tab:exp_notes1}.

### 2017-04-21 pt_characterization/rep_check/res2.tex
new draft for repeatability

@@
 This section exhibits histograms on the second run of 
-INC with its task length increasing from 1 second to 4096 seconds, via EMPv5. 
+INC with its task length increasing from 1 second to 4096 seconds, via SEDONA. 
 The detailed description of the base data is from Table~\ref{tab:exp_notes2}.
@@
 	\subfigure[ET frequency on INC1]{
-		\includegraphics[scale=0.43]{repet_data1/1_sec_et_hist_v5.eps}
+		\includegraphics[scale=0.43]{repet_data2/1_sec_et_hist_v5.eps}
 		\label{fig:inc1_r2_et_hist_v5}

### 2017-04-21 pt_characterization/rep_check/u_s_time.tex
new draft for repeatability

\section{User and System Time Histograms on the Second Run~\label{sec:u_s_time_hist}} 
This section exhibits user and system time histograms on the second run of 
INC with its task length increasing from 1 second to 4096 seconds, via SEDONA. 
The detailed description of the base data is from Table~\ref{tab:exp_notes2}.

\subsection{User Time}

\begin{figure}[hp!]
	\centering
	\subfigure[User time frequency on INC1]{
		\includegraphics[scale=0.43]{u_s_time/1_sec_ut_hist.eps}
		\label{fig:inc1_ut_hist}
	}
	\subfigure[User time frequency on INC2]{
		\includegraphics[scale=0.43]{u_s_time/2_sec_ut_hist.eps}
		\label{fig:inc2_ut_hist}
	}
	\subfigure[User time frequency on INC4]{
		\includegraphics[scale=0.43]{u_s_time/4_sec_ut_hist.eps}
		\label{fig:inc4_ut_hist}
	}
	\subfigure[User time frequency on INC8]{
		\includegraphics[scale=0.43]{u_s_time/8_sec_ut_hist.eps}
		\label{fig:inc8_ut_hist}
	}
	\caption{User Time Histograms of INC1 ... INC8~\label{fig:ut_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[User time frequency on INC16]{
		\includegraphics[scale=0.43]{u_s_time/16_sec_ut_hist.eps}
		\label{fig:inc16_ut_hist}
	}
	\subfigure[User time frequency on INC32]{
		\includegraphics[scale=0.43]{u_s_time/32_sec_ut_hist.eps}
		\label{fig:inc32_ut_hist}
	}
	\subfigure[User time frequency on INC64]{
		\includegraphics[scale=0.43]{u_s_time/64_sec_ut_hist.eps}
		\label{fig:inc64_ut_hist}
	}
	\caption{User Time Histograms of INC16 ... INC64~\label{fig:ut_hist2}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[User time frequency on INC128]{
		\includegraphics[scale=0.43]{u_s_time/128_sec_ut_hist.eps}
		\label{fig:inc128_ut_hist}
	}
	\subfigure[User time frequency on INC256]{
		\includegraphics[scale=0.43]{u_s_time/256_sec_ut_hist.eps}
		\label{fig:inc256_ut_hist}
	}
	\subfigure[User time frequency on INC512]{
		\includegraphics[scale=0.43]{u_s_time/512_sec_ut_hist.eps}
		\label{fig:inc512_ut_hist}
	}
	\subfigure[User time frequency on INC1024]{
		\includegraphics[scale=0.43]{u_s_time/1024_sec_ut_hist.eps}
		\label{fig:inc1024_ut_hist}
	}
	\caption{User Time Histograms of INC128 ... INC1024~\label{fig:ut_hist3}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[User time frequency on INC2048]{
		\includegraphics[scale=0.43]{u_s_time/2048_sec_ut_hist.eps}
		\label{fig:inc2048_ut_hist}
	}
	\subfigure[User time frequency on INC4096]{
		\includegraphics[scale=0.43]{u_s_time/4096_sec_ut_hist.eps}
		\label{fig:inc4096_ut_hist}
	}
	\caption{User Time Histograms of INC2048 and INC4096~\label{fig:ut_hist4}}
\end{figure}

\vspace\fill
\clearpage

\subsection{System Time}

\begin{figure}[hp!]
	\centering
	\subfigure[System time frequency on INC1]{
		\includegraphics[scale=0.43]{u_s_time/1_sec_st_hist.eps}
		\label{fig:inc1_hist_v5}
	}
	\subfigure[System time frequency on INC2]{
		\includegraphics[scale=0.43]{u_s_time/2_sec_st_hist.eps}
		\label{fig:inc2_hist_st}
	}
	\subfigure[System time frequency on INC4]{
		\includegraphics[scale=0.43]{u_s_time/4_sec_st_hist.eps}
		\label{fig:inc4_hist_st}
	}
	\subfigure[System time frequency on INC8]{
		\includegraphics[scale=0.43]{u_s_time/8_sec_st_hist.eps}
		\label{fig:inc8_hist_st}
	}
	\caption{System Time Histograms of INC1 ... INC8~\label{fig:st_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[System time frequency on INC16]{
		\includegraphics[scale=0.43]{u_s_time/16_sec_st_hist.eps}
		\label{fig:inc16_hist_st}
	}
	\subfigure[System time frequency on INC32]{
		\includegraphics[scale=0.43]{u_s_time/32_sec_st_hist.eps}
		\label{fig:inc32_hist_st}
	}
	\subfigure[System time frequency on INC64]{
		\includegraphics[scale=0.43]{u_s_time/64_sec_st_hist.eps}
		\label{fig:inc64_hist_st}
	}
	\caption{System Time Histograms of INC16 ... INC64\label{fig:st_hist2}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[System time frequency on INC128]{
		\includegraphics[scale=0.43]{u_s_time/128_sec_st_hist.eps}
		\label{fig:inc128_hist_st}
	}
	\subfigure[System time frequency on INC256]{
		\includegraphics[scale=0.43]{u_s_time/256_sec_st_hist.eps}
		\label{fig:inc256_hist_st}
	}
	\subfigure[System time frequency on INC512]{
		\includegraphics[scale=0.43]{u_s_time/512_sec_st_hist.eps}
		\label{fig:inc512_hist_st}
	}
	\subfigure[System time frequency on INC1024]{
		\includegraphics[scale=0.43]{u_s_time/1024_sec_st_hist.eps}
		\label{fig:inc1024_hist_st}
	}
	\caption{System Time Histograms of INC256 ... INC1024~\label{fig:st_hist3}}
\end{figure}

\begin{figure}[t]
	\centering
	\subfigure[System time frequency on INC2048]{
		\includegraphics[scale=0.43]{u_s_time/2048_sec_st_hist.eps}
		\label{fig:inc2048_hist_st}
	}
	\subfigure[System time frequency on INC4096]{
		\includegraphics[scale=0.43]{u_s_time/4096_sec_st_hist.eps}
		\label{fig:inc4096_hist_st}
	}
	\caption{System Time Histograms of INC2048 and INC4096~\label{fig:st_hist4}}
\end{figure}

### 2017-04-21 pt_characterization/recent_data/sodb9_10k/hist.r

@@ -1,41 +1,0 @@
-### sodb9
-numSamples <- 10000
-
-x = read.csv(file="8_sec.dat",head=TRUE,sep="\t")
-x <- subset(x, x$ITERNUM <= numSamples)
-setEPS()
-postscript("8_sec_et_hist_v5.eps")
-binsize=1
-x <- subset(x, x$ITERNUM != 27)
-#x_up = mean(x$METIME) + 2*sd(x$METIME)
-#x_dn = mean(x$METIME) - 2*sd(x$METIME)
-#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
-nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
-h = hist(x$METIME, right=F, breaks=nbins,plot=F)
-xmin <-min(x$METIME)
-xmax <-max(x$METIME)
-ymax <- max(h$counts)
-ymax <- ceiling(ymax/100)*100
-#ymax <- 150
-plot(h, freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-binsize-5,xmax+binsize+1),col="blue", main='ET frequency on INC8 with 10K samples', 
-sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
-xlab='ET (ms)', ylab=expression('Frequency'))
-dev.off()
-
-x = read.csv(file="8_sec.dat",head=TRUE,sep="\t")
-x <- subset(x, x$ITERNUM <= numSamples)
-setEPS()
-postscript("8_sec_pt_hist_v5.eps")
-x <- subset(x, x$ITERNUM != 27)
-#x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
-#x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
-x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
-h = hist(x$PRTIME, right=F,breaks=max(x$PRTIME)-min(x$PRTIME)+1,plot=F)
-ymax <- max(h$counts)
-ymax <- ceiling(ymax/100)*100
-#ymax <- 100
-binsize=1
-plot(h, ylim=c(0,ymax), axes = TRUE,freq=TRUE,xlim=c(min(x$PRTIME)-4,max(x$PRTIME)+4),col="green", main='PT frequency on INC8 with 10K samples', 
[…]

### 2017-04-21 pt_characterization/rep_check/10k_run/hist.r

@@ -1,0 +1,95 @@
+### sodb9
+numSamples <- 10000
+x = read.csv(file="8_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("8_sec_et_hist_v5.eps")
+binsize=1
+x <- subset(x, x$ITERNUM != 9394)
+#x_up = mean(x$METIME) + 2*sd(x$METIME)
+#x_dn = mean(x$METIME) - 2*sd(x$METIME)
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+xmin <-min(x$METIME)
+xmax <-max(x$METIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$METIME)
+xmax <-max(x$METIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 4000
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='ET frequency on INC8 with 10K samples', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='ET (ms)', ylab=expression('Frequency'))
+axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+dev.off()
+
+numSamples <- 10000
+x = read.csv(file="16_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("16_sec_et_hist_v5.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+#x_up = mean(x$METIME) + 2*sd(x$METIME)
+#x_dn = mean(x$METIME) - 2*sd(x$METIME)
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+xmin <-min(x$METIME)
+xmax <-max(x$METIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
[…]

### 2017-04-21 pt_characterization/rep_check/repet_data1/et_hist_sodb9.r

@@ -8,21 +8,16 @@ binsize=1
 x_up = mean(x$METIME) + 2*sd(x$METIME)
 x_dn = mean(x$METIME) - 2*sd(x$METIME)
 #x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
-nbins <- ceiling((max(x$METIME)-min(x$METIME)+1) / binsize)
-h = hist(x$METIME, right=F, breaks=nbins+1,plot=F)
+xmin <-min(x$METIME)
+xmin
+xmax <-max(x$METIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
-xmin <-min(x$METIME)
-xmax <-max(x$METIME)
-plot(h, xaxt="n",freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC1', sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), xlab='ET (ms)', ylab=expression('Frequency'))
-#xaxtl <- seq(xmin-1,xmax+1,by=1)
-#axis(side=1, at=xaxtl, labels=xaxtl)
-#xaxtl <- seq(floor(min(h$mids)),ceiling(max(h$mids)-1),by=binsize)
-#xaxtl <- seq(floor(min(h$mids)),ceiling(max(h$mids)),by=binsize)
-#axis(side=1, at=h$breaks, labels=xaxtl)
-#yaxtl <- seq(0,ymax,by=50)
-#axis(side=2, at=yaxtl, labels=yaxtl)
-axis(side=1,at=h$mids,labels=seq(xmin,xmax-1))
+plot(h, xaxt="n",freq=TRUE,ylim=c(0,ymax), col="blue", main='ET frequency on INC1', sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), xlab='ET (ms)', ylab=expression('Frequency'))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
 x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
@@ -33,8 +28,12 @@ binsize=1
 x_up = mean(x$METIME) + 2*sd(x$METIME)
[…]

### 2017-04-21 pt_characterization/rep_check/repet_data1/pt_hist_sodb9.r

@@ -8,16 +8,18 @@ binsize=1
 x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
 x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
 #x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
-nbins <- ceiling((max(x$PRTIME)-min(x$PRTIME)) / binsize)
-h = hist(x$PRTIME, right=F, breaks=nbins,plot=F)
+xmin <-min(x$PRTIME)
+xmin
+xmax <-max(x$PRTIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
-xmin <-min(x$PRTIME)
-xmax <-max(x$PRTIME)
 plot(h, xaxt="n",axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green", main='PT frequency on INC1', 
 sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
 xlab='PT (ms)', ylab=expression('Frequency'))
-axis(side=1,at=h$mids,labels=seq(xmin,xmax-1))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
 x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
@@ -28,17 +30,19 @@ binsize=1
 x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
 x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
 #x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
-nbins <- ceiling((max(x$PRTIME)-min(x$PRTIME)) / binsize)
-h = hist(x$PRTIME, right=F, breaks=nbins,plot=F)
+xmin <-min(x$PRTIME)
+xmin
+xmax <-max(x$PRTIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
 #ymax <- 400
-xmin <-min(x$PRTIME)
[…]

### 2017-04-21 pt_characterization/rep_check/repet_data2/et_hist_sodb9.r

@@ -8,8 +8,12 @@ binsize=1
 x_up = mean(x$METIME) + 2*sd(x$METIME)
 x_dn = mean(x$METIME) - 2*sd(x$METIME)
 #x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
-nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
-h = hist(x$METIME, right=F, breaks=nbins,plot=F)
+xmin <-min(x$METIME)
+xmin
+xmax <-max(x$METIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
 xmin <-min(x$METIME)
@@ -17,8 +21,7 @@ xmax <-max(x$METIME)
 plot(h, xaxt="n",freq=TRUE,ylim=c(0,ymax), col="blue", main='ET frequency on INC1', 
 sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
 xlab='ET (ms)', ylab=expression('Frequency'))
-xaxtl <- seq(xmin-1,xmax+1,by=1)
-axis(side=1,at=h$mids,labels=seq(xmin,xmax-1))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
 x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
@@ -29,8 +32,12 @@ binsize=1
 x_up = mean(x$METIME) + 2*sd(x$METIME)
 x_dn = mean(x$METIME) - 2*sd(x$METIME)
 #x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
-nbins <- ceiling((max(x$METIME)-min(x$METIME)) / binsize)
-h = hist(x$METIME, right=F, breaks=nbins,plot=F)
+xmin <-min(x$METIME)
+xmin
+xmax <-max(x$METIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
 xmin <-min(x$METIME)
[…]

### 2017-04-21 pt_characterization/rep_check/repet_data2/pt_hist_sodb9.r

@@ -8,16 +8,18 @@ binsize=1
 x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
 x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
 #x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
-nbins <- ceiling((max(x$PRTIME)-min(x$PRTIME)) / binsize)
-h = hist(x$PRTIME, right=F, breaks=nbins,plot=F)
+xmin <-min(x$PRTIME)
+xmin
+xmax <-max(x$PRTIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
-xmin <-min(x$PRTIME)
-xmax <-max(x$PRTIME)
 plot(h, xaxt="n",freq=TRUE,ylim=c(0,ymax), col="green", main='PT frequency on INC1', 
 sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
 xlab='PT (ms)', ylab=expression('Frequency'))
-axis(side=1,at=h$mids,labels=seq(xmin,xmax-1))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
 x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
@@ -28,18 +30,19 @@ binsize=1
 x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
 x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
 #x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
-nbins <- ceiling((max(x$PRTIME)-min(x$PRTIME)) / binsize)
-h = hist(x$PRTIME, right=F, breaks=nbins,plot=F)
+xmin <-min(x$PRTIME)
+xmin
+xmax <-max(x$PRTIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
 #ymax <- 250
-xmin <-min(x$PRTIME)
-xmax <-max(x$PRTIME)
[…]

### 2017-04-21 pt_characterization/rep_check/s_time_hist/s_time_hist.r

@@ -1,280 +1,0 @@
-### sodb9
-numSamples <- 1000
-x = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
-x <- subset(x, x$ITERNUM <= numSamples)
-setEPS()
-postscript("1_sec_st_hist.eps")
-binsize=1
-xmin <-min(x$S_TIME)
-xmax <-max(x$S_TIME)
-#x = subset(x, x$S_TIME >= x_dn & x$S_TIME <= x_up)
-nbins <- ceiling((max(x$S_TIME)-min(x$S_TIME)) / binsize)
-h = hist(x$S_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
-ymax <- max(h$counts)
-ymax <- ceiling(ymax/100)*100
-xmin <-min(x$S_TIME)
-xmax <-max(x$S_TIME)
-plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green", main='System time frequency on INC1', 
-sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
-xlab='System Time (ms)', ylab=expression('Frequency'))
-axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
-dev.off()
-
-x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
-x <- subset(x, x$ITERNUM <= numSamples)
-setEPS()
-postscript("2_sec_st_hist.eps")
-binsize=1
-xmin <-min(x$S_TIME)
-xmax <-max(x$S_TIME)
-#x = subset(x, x$S_TIME >= x_dn & x$S_TIME <= x_up)
-nbins <- ceiling((max(x$S_TIME)-min(x$S_TIME)) / binsize)
-h = hist(x$S_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
-ymax <- max(h$counts)
-ymax <- ceiling(ymax/100)*100
-xmin <-min(x$S_TIME)
-xmax <-max(x$S_TIME)
-plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green", main='System time frequency on INC2', 
-sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
-xlab='System Time (ms)', ylab=expression('Frequency'))
-axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
-dev.off()
-
[…]

### 2017-04-21 pt_characterization/rep_check/s_time_hist/u_time_hist.r

@@ -1,266 +1,0 @@
-### sodb9
-numSamples <- 1000
-x = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
-x <- subset(x, x$ITERNUM <= numSamples)
-setEPS()
-postscript("1_sec_ut_hist.eps")
-binsize=1
-#x = subset(x, x$U_TIME >= x_dn & x$U_TIME <= x_up)
-xmin <-min(x$U_TIME)
-xmax <-max(x$U_TIME)
-nbins <- ceiling((xmax-xmin) / binsize)
-h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
-ymax <- max(h$counts)
-ymax <- ceiling(ymax/100)*100
-plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green", main='User time frequency on INC1', 
-sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
-xlab='User time (ms)', ylab=expression('Frequency'))
-axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
-dev.off()
-
-x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
-x <- subset(x, x$ITERNUM <= numSamples)
-setEPS()
-postscript("2_sec_ut_hist.eps")
-binsize=1
-#x = subset(x, x$U_TIME >= x_dn & x$U_TIME <= x_up)
-xmin <-min(x$U_TIME)
-xmax <-max(x$U_TIME)
-nbins <- ceiling((xmax-xmin) / binsize)
-h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
-ymax <- max(h$counts)
-ymax <- ceiling(ymax/100)*100
-plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green", main='User time frequency on INC2', 
-sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
-xlab='User time (ms)', ylab=expression('Frequency'))
-axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
-dev.off()
-
-x = read.csv(file="4_sec.dat",head=TRUE,sep="\t")
-x <- subset(x, x$ITERNUM <= numSamples)
-setEPS()
[…]

### 2017-04-21 pt_characterization/rep_check/u_s_time/s_time_hist.r

@@ -1,0 +1,280 @@
+### sodb9
+numSamples <- 1000
+x = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("1_sec_st_hist.eps")
+binsize=1
+xmin <-min(x$S_TIME)
+xmax <-max(x$S_TIME)
+#x = subset(x, x$S_TIME >= x_dn & x$S_TIME <= x_up)
+nbins <- ceiling((max(x$S_TIME)-min(x$S_TIME)) / binsize)
+h = hist(x$S_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+xmin <-min(x$S_TIME)
+xmax <-max(x$S_TIME)
+plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="yellow",  main='System time frequency on INC1', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='System Time (ms)', ylab=expression('Frequency'))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("2_sec_st_hist.eps")
+binsize=1
+xmin <-min(x$S_TIME)
+xmax <-max(x$S_TIME)
+#x = subset(x, x$S_TIME >= x_dn & x$S_TIME <= x_up)
+nbins <- ceiling((max(x$S_TIME)-min(x$S_TIME)) / binsize)
+h = hist(x$S_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+xmin <-min(x$S_TIME)
+xmax <-max(x$S_TIME)
+plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="yellow",  main='System time frequency on INC2', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='System Time (ms)', ylab=expression('Frequency'))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
[…]

### 2017-04-21 pt_characterization/rep_check/u_s_time/u_time_hist.r

@@ -1,0 +1,266 @@
+### sodb9
+numSamples <- 1000
+x = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("1_sec_ut_hist.eps")
+binsize=1
+#x = subset(x, x$U_TIME >= x_dn & x$U_TIME <= x_up)
+xmin <-min(x$U_TIME)
+xmax <-max(x$U_TIME)
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="orange",  main='User time frequency on INC1', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='User time (ms)', ylab=expression('Frequency'))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("2_sec_ut_hist.eps")
+binsize=1
+#x = subset(x, x$U_TIME >= x_dn & x$U_TIME <= x_up)
+xmin <-min(x$U_TIME)
+xmax <-max(x$U_TIME)
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="orange",  main='User time frequency on INC2', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='User time (ms)', ylab=expression('Frequency'))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+x = read.csv(file="4_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
[…]

### 2017-04-26 pt_characterization/rep_check/1024_run.tex
April 27 draft

\section{Histograms on Consecutive INC1024 Runs~\label{sec:s9_1024_runs}} 
This section exhibits histograms on (three) consecutive runs of INC1024 via SEDONA. 
The detailed description of the base data is from Table~\ref{tab:exp_notes2}.

\begin{table}[h]
\begin{center}
\begin{tabular}{|p{2cm}|p{3cm}|p{3cm}|p{4cm}|p{3.5cm}|} \hline
Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
{\tt sodb9} &  INC1024 & 300 samples, each & 2017-04-12 $\sim$ 2017-04-23 & 
Figs.~\ref{fig:s9_inc1024_et} and~\ref{fig:s9_inc1024_pt}\\ \hline
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{Notes on experiment runs used for histograms\label{tab:inc1024_run_notes}}
\end{table}

\clearpage
\newpage

\subsection{ET}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC1024-Run1]{
		\includegraphics[scale=0.43]{1024_run/1024_sec_et_hist1.eps}
		\label{fig:inc1024_run1_et}
	}
	\subfigure[ET frequency on INC1024-Run2]{
		\includegraphics[scale=0.43]{1024_run/1024_sec_et_hist2.eps}
		\label{fig:inc1024_run2_et}
	}
	\subfigure[ET frequency on INC1024-Run3]{
		\includegraphics[scale=0.43]{1024_run/1024_sec_et_hist3.eps}
		\label{fig:inc1024_run3_et}
	}
	\caption{ET Histograms of Three Consecutive INC4096 Runs\label{fig:s9_inc1024_et}}
\end{figure}

\vspace\fill
\clearpage

\subsection{PT}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC1024-Run1]{
		\includegraphics[scale=0.43]{1024_run/1024_sec_pt_hist1.eps}
		\label{fig:inc1024_run1_pt}
	}
	\subfigure[PT frequency on INC1024-Run2]{
		\includegraphics[scale=0.43]{1024_run/1024_sec_pt_hist2.eps}
		\label{fig:inc1024_run2_pt}
	}
	\subfigure[PT frequency on INC1024-Run3]{
		\includegraphics[scale=0.43]{1024_run/1024_sec_pt_hist3.eps}
		\label{fig:inc1024_run3_pt}
	}
	\caption{PT Histograms of Three Consecutive INC4096 Runs\label{fig:s9_inc1024_pt}}
\end{figure}

### 2017-04-26 pt_characterization/rep_check/u_s_time.tex
April 27 draft

@@
-\section{User and System Time Histograms on the Second Run~\label{sec:u_s_time_hist}} 
+\section{Further invstigation of the Second Run~\label{sec:u_s_time_hist}} 
 This section exhibits user and system time histograms on the second run of 
@@
 \end{figure}
+
+\clearpage
+\pagebreak
+
+\subsection{Correlation}
+
+%% done
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|c|c|c|} \hline
+%Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
+%{\tt sodb9} &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-02 $\sim$ 2017-03-04 & Figs.~\ref{fig:s9_r1_et_hist1},~\ref{fig:s9_r1_et_hist2},~\ref{fig:s9_r1_pt_hist1}, and~\ref{fig:s9_r1_pt_hist2}\\ \hline
+%{\tt sodb9} &  INC128$\sim$INC1024 & 300 samples, each & 2017-03-04 $\sim$ 2017-03-11 & 
+%Figs.~\ref{fig:s9_r1_et_hist3} and~\ref{fig:s9_r1_pt_hist3}\\ \hline
+%{\tt sodb10} & INC2048 & 300 samples & 2017-03-02 $\sim$ 2017-03-09 & Figs.~\ref{fig:inc2048_r1_et_hist_v5} and~\ref{fig:inc2048_r1_hist_v5}\\ \hline
+%{\tt sodb12} & INC4096 & 300 samples & 2017-02-13 $\sim$ 2017-02-27 & Figs.~\ref{fig:inc4096_r1_et_hist_v5} and~\ref{fig:inc4096_r1_hist_v5}\\ \hline
+ 						& w/ u time & w/ s time\\ \hline
+daemon u time & 0.1 & {\bf 0.3} \\ \hline
+daemon s time & -0.09 & {\bf 0.19} \\ \hline
+daemon minor faults & {\bf 0.11} & {\bf 0.32} \\ \hline
+%read bytes & N/A & N/A
+daemon read char & 0.1 & {\bf 0.32} \\ \hline
+daemon read sys calls & 0.11 & {\bf 0.32} \\ \hline
+daemon write bytes & 0 & {\bf 0.26} \\ \hline
+daemon write char & 0.11 & {\bf 0.32}\\ \hline
+%read sys calls & N/A & N/A
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Correlation of user and system time of INC4096 with some daemon measures\label{tab:corr_table}}
+\end{table}
+
+\clearpage
+\newpage
+
+\subsection{Scatter Plots on Some Significant Correlations}
+The following scatter plots correspond to the correlations bold in Table~\ref{tab:corr_table}.
+
+\begin{figure}[htp!]
+	\centering
+	\subfigure[User time vs. All daemon system time]{
+		\includegraphics[scale=0.43]{u_s_time/corr_u_as_time.eps}
+		\label{fig:corr_u_as_time}
+	}
+	\subfigure[System time vs. All daemon system time]{
+		\includegraphics[scale=0.43]{u_s_time/corr_s_as_time.eps}
+		\label{fig:corr_s_as_time}
+	}
+	\subfigure[User time vs. All daemon minor faults]{
+		\includegraphics[scale=0.43]{u_s_time/corr_mnf_u_time.eps}
+		\label{fig:corr_u_mnf}
+	}
+	\subfigure[System time vs. All daemon minor faults]{
+		\includegraphics[scale=0.43]{u_s_time/corr_mnf_s_time.eps}
+		\label{fig:corr_s_mnf}
+	}
+	\caption{Scatter plots between measures on INC4096~\label{fig:corr1}}
+\end{figure}
+
+\begin{figure}[htp!]
+	\centering
+	\subfigure[System time vs. All daemon read char]{
+		\includegraphics[scale=0.43]{u_s_time/corr_s_rchar.eps}
+		\label{fig:corr_s_rchar}
+	}
+	\subfigure[System time vs. All daemon read syscalls]{
+		\includegraphics[scale=0.43]{u_s_time/corr_s_rsysc.eps}
+		\label{fig:corr_s_rsysc}
+	}
+	\subfigure[System time vs. All daemon write bytes]{
+		\includegraphics[scale=0.43]{u_s_time/corr_s_wbytes.eps}
+		\label{fig:corr_s_wbytes}
+	}
+	\subfigure[System time vs. All daemon write char]{
+		\includegraphics[scale=0.43]{u_s_time/corr_s_wchar.eps}
+		\label{fig:corr_s_wchar}
+	}
+	\caption{Scatter plots between measures on INC4096~\label{fig:corr2}}
+\end{figure}

### 2017-04-26 pt_characterization/rep_check/1024_run/et_hist_sodb9.r

@@ -1,0 +1,74 @@
+### sodb9 
+### 1st run
+### INC1024 with 300 samples: taking 3.5 days.
+x = read.csv(file="1024_sec1.dat",head=TRUE,sep="\t")
+setEPS()
+postscript("1024_sec_et_hist1.eps")
+x <- subset(x, x$ITERNUM != 40 & x$ITERNUM != 124 & x$ITERNUM != 207 & x$ITERNUM != 292)
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+binsize=2
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+xmin <-min(x$METIME)
+xmin
+xmax <-max(x$METIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- max(h$counts)
+ymax <- 50
+plot(h, xaxt="n",freq=TRUE,ylim=c(0,ymax), col="blue", main='', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='ET (ms)', ylab=expression('Frequency'))
+axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+dev.off()
+
+x = read.csv(file="1024_sec2.dat",head=TRUE,sep="\t")
+setEPS()
+postscript("1024_sec_et_hist2.eps")
+x <- subset(x, x$ITERNUM != 75 & x$ITERNUM != 159 & x$ITERNUM != 243)
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+binsize=2
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+xmin <-min(x$METIME)
+xmin
+xmax <-max(x$METIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+ymax <- max(h$counts)
[…]

### 2017-04-26 pt_characterization/rep_check/1024_run/pt_hist_sodb9.r

@@ -1,0 +1,74 @@
+### sodb9
+### 1st run
+### INC1024 with 300 samples: taking 3.5 days.
+numSamples <- 300
+x = read.csv(file="1024_sec1.dat",head=TRUE,sep="\t")
+setEPS()
+postscript("1024_sec_pt_hist1.eps")
+x <- subset(x, x$ITERNUM != 40 & x$ITERNUM != 124 & x$ITERNUM != 207 & x$ITERNUM != 292)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+binsize=5
+xmin <-min(x$PRTIME)
+xmin
+xmax <-max(x$PRTIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 100
+plot(h, ylim=c(0,ymax), xaxt="n",freq=TRUE,col="green", main='', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+dev.off()
+
+numSamples <- 300
+x = read.csv(file="1024_sec2.dat",head=TRUE,sep="\t")
+setEPS()
+postscript("1024_sec_pt_hist2.eps")
+x <- subset(x, x$ITERNUM != 75 & x$ITERNUM != 159 & x$ITERNUM != 243)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+binsize=5
+xmin <-min(x$PRTIME)
+xmin
+xmax <-max(x$PRTIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
[…]

### 2017-04-26 pt_characterization/rep_check/u_s_time/scatterplot.r

@@ -1,0 +1,126 @@
+### sodb9
+numSamples <- 300
+x = read.csv(file="per_iter_detailed.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+min_stime <- min(x$S_TIME)
+min_stime <- floor(min_stime/10)*10
+max_stime <- max(x$S_TIME)
+max_stime <- ceiling(max_stime/10)*10
+min_mnflt <- min(x$ALL_MIN_FLT)
+min_mnflt <- floor(min_mnflt/1000)*1000
+min_mnflt <-0
+max_mnflt <- max(x$ALL_MIN_FLT)
+max_mnflt <- ceiling(max_mnflt/1000)*1000
+setEPS()
+postscript("corr_mnf_s_time.eps")
+plot(x$S_TIME, x$ALL_MIN_FLT,xlim=c(min_stime,max_stime), 
+ylim=c(min_mnflt,max_mnflt),
+pch=1, main='', xlab='System time (ms)', ylab='All daemon minor faults')
+dev.off()
+
+min_utime <- min(x$U_TIME)
+min_utime <- floor(min_utime/10)*10
+max_utime <- max(x$U_TIME)
+max_utime <- ceiling(max_utime/10)*10
+min_mnflt <- min(x$ALL_MIN_FLT)
+min_mnflt <- floor(min_mnflt/1000)*1000
+min_mnflt <-0
+max_mnflt <- max(x$ALL_MIN_FLT)
+max_mnflt <- ceiling(max_mnflt/1000)*1000
+setEPS()
+postscript("corr_mnf_u_time.eps")
+plot(x$U_TIME, x$ALL_MIN_FLT,xlim=c(min_utime,max_utime), 
+ylim=c(min_mnflt,max_mnflt),
+pch=1, main='', xlab='User time (ms)', ylab='All daemon minor faults')
+dev.off()
+
+min_utime <- min(x$U_TIME)
+min_utime <- floor(min_utime/10)*10
+max_utime <- max(x$U_TIME)
+max_utime <- ceiling(max_utime/10)*10
+min_astime <- min(x$ALL_S_TIME)
+min_astime <- floor(min_astime/10)*10
+max_astime <- max(x$ALL_S_TIME)
+max_astime <- ceiling(max_astime/10)*10
+setEPS()
+postscript("corr_u_as_time.eps")
+plot(x$U_TIME, x$ALL_S_TIME,xlim=c(min_utime,max_utime), 
[…]

### 2017-05-24 pt_characterization/rep_check/1024_run.tex
New draft for concrete complexity

@@
+\clearpage
+
 \section{Histograms on Consecutive INC1024 Runs~\label{sec:s9_1024_runs}} 
@@
 	}
-	\caption{ET Histograms of Three Consecutive INC4096 Runs\label{fig:s9_inc1024_et}}
+	\caption{ET Histograms of Three Consecutive INC1024 Runs\label{fig:s9_inc1024_et}}
 \end{figure}
@@
 	}
-	\caption{PT Histograms of Three Consecutive INC4096 Runs\label{fig:s9_inc1024_pt}}
+	\caption{PT Histograms of Three Consecutive INC1024 Runs\label{fig:s9_inc1024_pt}}
 \end{figure}

### 2017-05-24 pt_characterization/rep_check/paper.tex
new draft for repeatability
April 27 draft
New draft for concrete complexity

@@
 \usepackage{graphicx}
+\usepackage{footnote}
 %\usepackage{arydshln}
@@
 If so then, we can say that run repeatability is satisfied in our experiment setting. 
-In this study the used protocol is equivalent to SEDONA, eliminating samples involving infrequent, long-running daemons, which is the first step of EMPv5.
-In other words, the protocol skips the second step
-of EMPv5 that removes samples with process times over/below two standard deviations 
-from the average after performing the first step.
+In this study the used protocol is equivalent to SEDONA~\cite{Sedona}, 
+eliminating samples involving infrequent, long-running daemons, which is the first step of EMPv5.
+
+In the SEDONA protocol, we use taskstats C struct to get measures of a captured process. 
+The taskstat's data is delivered via a netlink socket from the kernel space. 
+The receive buffer for the socket is not robust for many observed processes~\cite{Metrology}. 
+Fortunately, there is an average of 95 processes per iteration of a run, 
+which turns out to be fine with the struct. 
+For a much more number of processes, 
+the use of  {\tt /proc/}[pid]{\tt{/stat}} is preferred, 
+as (i) there are equivalent measures available in the {\tt /proc} filesystem, 
+and (ii) there's little constraint on the use as opposed to taskstats.
 
@@
 
+\pagebreak
+\newpage
+
+\input{10k_run.tex}
+
+\pagebreak
+\newpage
+
+\input{u_s_time.tex}
+
+\pagebreak
+\newpage
+
+\input{1024_run.tex}
+
+\bibliographystyle{abbrv}
+\newcommand{\etalchar}[1]{$^{#1}$}
+\begin{thebibliography}{99}
+\vspace{0.1em}
+\bibitem
+%{Avnur:2000:ECA:342009.335420}
+{Sedona}
+Young-Kyoon Suh, ``SEDONA: A Novel Protocol for Identifying Infrequent, Long-running Daemons on a Linux System'', in {\em IEICE Transactions on Information Systems}, Vol.
+100D, No.
+9, pp.~??--??, 2017.
+\vspace{0.1em}
+\bibitem
+%{Avnur:2000:ECA:342009.335420}
+{Metrology}
+Sabah Currim,Richard T. Snodgrass, Young-Kyoon Suh, and Rui Zhang, ``DBMS Metrology: Measuring Query Time'', in {\em ACM Transactions on Database Systems}, 42(1):3:1--42(+8), 2017.
+\end{thebibliography}
+
 \end{document}

### 2017-05-24 pt_characterization/rep_check/res2.tex
New draft for concrete complexity

@@
 {\tt sodb12} & INC4096 & 300 samples & 2017-03-02 $\sim$ 2017-03-17 & Figs.~\ref{fig:inc4096_r2_et_hist_v5} and~\ref{fig:inc4096_r2_hist_v5}\\ \hline
+{\tt sodb10} & INC8192& 261 samples & 2017-04-27 $\sim$ 2017-05-21 & Figs.~\ref{fig:inc8192_r2_et_hist_v5} and~\ref{fig:inc8192_r2_hist_v5}\\ \hline
+{\tt sodb12} & INC16384& 130 samples & 2017-04-27 $\sim$ 2017-05-21 & Figs.~\ref{fig:inc16384_r2_et_hist_v5} and~\ref{fig:inc16384_r2_hist_v5}\\ \hline
 \end{tabular}
@@
 \end{table}
+
+INC8192/INC16384 unfortunately stopped in the middle of their runs due to a frozen vnc problem. So couldn't finish 300 samples.
 
@@
 	}
-	\caption{ET Histograms of INC2048 and INC4096~\label{fig:s9_r2_et_hist4}}
+	\subfigure[ET frequency on INC8192]{
+		\includegraphics[scale=0.43]{repet_data2/8192_sec_et_hist.eps}
+		\label{fig:inc8192_r2_et_hist_v5}
+	}
+	\subfigure[ET frequency on INC16384]{
+		\includegraphics[scale=0.43]{repet_data2/16384_sec_et_hist.eps}
+		\label{fig:inc16384_r2_et_hist_v5}
+	}
+	\caption{ET Histograms of INC2048 ... INC16384~\label{fig:s9_r2_et_hist4}}
 \end{figure}
@@
 	}
-	\caption{PT Histograms of INC2048 and INC4096~\label{fig:s9_r2_pt_hist4}}
+	\subfigure[PT frequency on INC8192]{
+		\includegraphics[scale=0.43]{repet_data2/8192_sec_pt_hist.eps}
+		\label{fig:inc8192_r2_hist_v5}
+	}
+	\subfigure[PT frequency on INC16384]{
+		\includegraphics[scale=0.43]{repet_data2/16384_sec_pt_hist.eps}
+		\label{fig:inc16384_r2_hist_v5}
+	}
+	\caption{PT Histograms of INC2048 ... INC16384~\label{fig:s9_r2_pt_hist4}}
 \end{figure}

### 2017-05-24 pt_characterization/rep_check/u_s_time.tex
New draft for concrete complexity

@@
 
-%% done
 \begin{table}[h]
@@
 \begin{tabular}{|c|c|c|} \hline
-%Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
-%{\tt sodb9} &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-02 $\sim$ 2017-03-04 & Figs.~\ref{fig:s9_r1_et_hist1},~\ref{fig:s9_r1_et_hist2},~\ref{fig:s9_r1_pt_hist1}, and~\ref{fig:s9_r1_pt_hist2}\\ \hline
-%{\tt sodb9} &  INC128$\sim$INC1024 & 300 samples, each & 2017-03-04 $\sim$ 2017-03-11 & 
-%Figs.~\ref{fig:s9_r1_et_hist3} and~\ref{fig:s9_r1_pt_hist3}\\ \hline
-%{\tt sodb10} & INC2048 & 300 samples & 2017-03-02 $\sim$ 2017-03-09 & Figs.~\ref{fig:inc2048_r1_et_hist_v5} and~\ref{fig:inc2048_r1_hist_v5}\\ \hline
-%{\tt sodb12} & INC4096 & 300 samples & 2017-02-13 $\sim$ 2017-02-27 & Figs.~\ref{fig:inc4096_r1_et_hist_v5} and~\ref{fig:inc4096_r1_hist_v5}\\ \hline
- 						& w/ u time & w/ s time\\ \hline
+ 						   & INC2048's u time & INC2048's s time\\ \hline
+INC2048's u time  &	 -  &  0.05\\ \hline					
+daemon u time & 0.56 & 0.61\\ \hline
+daemon s time & 0.48 & 0.57\\ \hline
+daemon minor faults &  0.53 & 0.62\\ \hline
+daemon major faults & 0.55 & 0.59 \\ \hline
+daemon read bytes & 0.55 & 0.59 \\ \hline
+daemon read char & 0.56 & 0.61\\ \hline
+daemon read sys calls & 0.57 & 0.63\\ \hline
+daemon write bytes & 0.57 &  0.64\\ \hline
+daemon write char & 0.53 & 0.62\\ \hline
+%write sys calls & N/A & N/A\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Correlation of user and system time of INC2048 with some daemon measures\label{tab:corr_table2}}
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|c|c|c|} \hline
+ 						   & INC4096's u time & INC4096's s time\\ \hline
+INC4096's u time  &	 -  & -0.30 \\ \hline					
 daemon u time & 0.1 & {\bf 0.3} \\ \hline
@@
 daemon write char & 0.11 & {\bf 0.32}\\ \hline
-%read sys calls & N/A & N/A
+%write sys calls & N/A & N/A
 \end{tabular}
@@
 	\centering
-	\subfigure[User time vs. All daemon system time]{
+	\subfigure[User time vs. Sum of all daemon system time]{
 		\includegraphics[scale=0.43]{u_s_time/corr_u_as_time.eps}
@@
 	}
-	\subfigure[System time vs. All daemon system time]{
+	\subfigure[System time vs. Sum of all  daemon system time]{
 		\includegraphics[scale=0.43]{u_s_time/corr_s_as_time.eps}
@@
 	}
-	\subfigure[User time vs. All daemon minor faults]{
+	\subfigure[User time vs. Sum of all  daemon minor faults]{
 		\includegraphics[scale=0.43]{u_s_time/corr_mnf_u_time.eps}
@@
 	}
-	\subfigure[System time vs. All daemon minor faults]{
+	\subfigure[System time vs. Sum of all  daemon minor faults]{
 		\includegraphics[scale=0.43]{u_s_time/corr_mnf_s_time.eps}
@@
 	\centering
-	\subfigure[System time vs. All daemon read char]{
+	\subfigure[System time vs. Sum of all  daemon read char]{
 		\includegraphics[scale=0.43]{u_s_time/corr_s_rchar.eps}
@@
 	}
-	\subfigure[System time vs. All daemon read syscalls]{
+	\subfigure[System time vs. Sum of all  daemon read syscalls]{
 		\includegraphics[scale=0.43]{u_s_time/corr_s_rsysc.eps}
@@
 	}
-	\subfigure[System time vs. All daemon write bytes]{
+	\subfigure[System time vs. Sum of all  daemon write bytes]{
 		\includegraphics[scale=0.43]{u_s_time/corr_s_wbytes.eps}
@@
 	}
-	\subfigure[System time vs. All daemon write char]{
+	\subfigure[System time vs. Sum of all daemon write char]{
 		\includegraphics[scale=0.43]{u_s_time/corr_s_wchar.eps}
@@
 \end{figure}
+
+\clearpage
+
+\subsection{Further Investigation of Some Samples}
+
+\subsubsection{Samples in Fig.~\ref{fig:corr_u_mnf}}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|} \hline
+proc name (id) & u time & s time & min flt & maj flt & r bytes & r char & r sysc & w bytes & w char & w sysc\\ \hline
+INC4096 (3559) & 4103920 & 36 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 \\ \hline
+proc\_monitor (25917) & 194 & 6 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+md0\_raid1 (484) & 0 &  3 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+java (3549) & 2 & 1 & 1093 & 0 & 0 & 11480 & 20 & 0 & 0 & 0 \\ \hline
+cifsd (1927) & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+kblockd/0 (16) & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+ntpd (28232) & 0 & 1 & 1 & 0 & 0 & 0 & 42 & 4096 & 7 & 0 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Observed values of measures of processes on the leftmost sample in Fig.~\ref{fig:corr_u_mnf}~\label{tab:breakdown}}
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|} \hline
+proc name (id) & u time & s time & min flt & maj flt & r bytes & r char & r sysc & w bytes & w char & w sysc\\ \hline
+INC4096 (3559) & 4103984 & 19 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 \\ \hline
+proc\_monitor (25917) & 190 & 6 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+java (3549) & 2 & 1 & 1093 & 0 & 0 & 11480 & 20 & 0 & 0 & 0 \\ \hline
+md0\_raid1 (484) & 0 & 3 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+ntpd (28232) & 0 & 0 & 1 & 0 & 0 & 0 & 39 & 4096 & 7 & 0 \\ \hline
+java (4108) & 0 & 0  & 20 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Observed values of measures of processes on the second rightmost sample in Fig.~\ref{fig:corr_u_mnf}~\label{tab:breakdown2}}
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|} \hline
+proc name (id) & u time & s time & min flt & maj flt & r bytes & r char & r sysc & w bytes & w char & w sysc\\ \hline
+INC4096 (3559) & 4103988 & 25 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 \\ \hline
+proc\_monitor  (25917)  & 194 & 2 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+sshd (3609) & 8 & 4 & 1382 & 0 & 0 & 512357 & 400 & 0 & 20881 & 0 \\ \hline
+bash (3611)  & 4 & 1 & 835 & 0 & 0 & 283911 & 155 & 0 & 136 & 0 \\ \hline
+md0\_raid1 (484) & 0 & 3 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+java (3549) & 2 & 1 & 1093 & 0 & 0 & 11480 & 20 & 0 & 0 & 0 \\ \hline
+cifsd (1927) & 0 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+java (3606) & 0 & 1 & 22 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+jbd2/md0-8 (497) & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 450560 & 0 & 0 \\ \hline
+kblockd/0 (16) & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+grep (3617) & 1 & 0 & 311 & 0 & 0 & 5417 & 11 & 0 & 0 & 0 \\ \hline
+bash (3612) & 0 & 0 & 158 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+consoletype (3613) & 0 & 0 & 127 & 0 & 0 & 1956 & 6 & 0 & 7 & 0 \\ \hline
+bash (3614) & 0 & 0 & 174 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+uname (3615) & 0 & 0 & 189 & 0 & 0 & 1956  & 6 & 0 & 7 & 0 \\ \hline
+sshd (3610) & 0 & 0 & 425 & 0 & 0 & 22656 & 29 & 0 & 4630 & 0 \\ \hline
+sshd (2105) & 0 & 0 & 14 & 0 & 0 & 0 & 0 & 0 & 594 & 0 \\ \hline
+bash (3618) & 0 & 0 & 170 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+id (3619) & 0 & 0 & 225 & 0 & 0 & 4352 & 12 & 0 & 2 & 0 \\ \hline
+ntpd (28232) & 0 & 0 & 1 & 0 & 0 & 0 & 42 & 4096 & 7 & 0 \\ \hline
+bash (3616) & 0 & 0 & 131 & 0 & 0 & 0 & 0 & 0 & 61 & 0 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Observed values of measures of processes on the rightmost sample in Fig.~\ref{fig:corr_u_mnf}~\label{tab:breakdown1}}
+\end{table}
+
+\clearpage
+\newpage
+
+\subsubsection{Samples in Fig.~\ref{fig:corr_s_wbytes}}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|} \hline
+proc name (id) & u time & s time & min flt & maj flt & r bytes & r char & r sysc & w bytes & w char & w sysc\\ \hline
+INC4096 (3559) & 4103981 & 14 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 \\ \hline
+proc\_monitor (25917) & 194 & 2 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+md0\_raid1 (484) & 0 & 5 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+java (3549) & 2 & 1 & 1093 & 0 & 0 & 11480 & 20 & 0 & 0 & 0 \\ \hline
+flush-9:0 (3548) & 0 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+ntpd (28232) & 0 & 0 & 1 & 0 & 0 & 0 & 42 & 4096 & 7 & 0 \\ \hline
+java (4585) & 0 & 0 & 20 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Observed values of measures of processes on the leftmost sample in Fig.~\ref{fig:corr_s_wbytes}~\label{tab:breakdown4}}
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|} \hline
+proc name (id) & u time & s time & min flt & maj flt & r bytes & r char & r sysc & w bytes & w char & w sysc\\ \hline
+INC4096 (3559) & 4103958 & 44 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 \\ \hline
+proc\_monitor (25917) & 194 & 2 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+sshd (4877) & 10 & 4 & 1382 & 0 & 0 & 519568 & 392 & 0 & 20868 & 0 \\ \hline
+sshd (4886) & 10 & 3 & 1383 & 0 & 0 & 515456 & 393 & 0 & 20868 & 0 \\ \hline
+grep (4888) & 6 & 1 & 994 & 0 & 0 & 287034 & 154 & 0 & 136 & 0 \\ \hline
+grep (4879) & 3 & 2 & 991 & 0 & 0 & 286801 & 159 & 0 & 136 & 0 \\ \hline
+java (3549) & 2 & 1 & 1093 & 0 & 0 & 11480 & 20 & 0 & 0 & 0 \\ \hline
+md0\_raid1 (484) & 0 & 3 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+sshd (4878) & 2 & 1 & 424 & 0 & 0 & 22403 & 26 & 0 & 4268 & 0 \\ \hline
+jbd2/md0-8 (497) & 0 & 2 & 0 & 0 & 0 & 0 & 0 & 696320 & 0 & 0 \\ \hline
+flush-9:0 (3548) & 0 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+cifsd (1927) & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+kblockd/0 (16) & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+grep (4892) & 0 & 1 & 309 & 0 & 0 & 5417 & 11 & 0 & 0 & 0 \\ \hline
+grep (4883) & 1 & 0 & 310 & 0 & 0 & 5417 & 11 & 0 & 0 & 0 \\ \hline
+id (4894) & 0 & 0 & 226 & 0 & 0 & 4352 & 12 & 0 & 2 & 0 \\ \hline
+bash (4884) & 0 & 0 & 164 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+id (4885) & 0 & 0 & 226 & 0 & 0 & 0 & 12 & 0 & 2 & 0 \\ \hline
+sshd (2105) & 0 & 0 & 28 & 0 & 0 & 0 & 0 & 0 & 1188 & 0 \\ \hline
+sshd (4887) & 0 & 0 & 425 & 0 & 0 & 22403 &26 & 0 & 4268 & 0 \\ \hline
+ntpd (20232) & 0 & 0 & 1 & 0 & 0 & 0 & 42 & 4096 & 7 & 0 \\ \hline
+bash (4880) & 0 & 0 & 165 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+uname (4890)  & 0 & 0 & 190 & 0 & 0 & 1956  & 6 & 0 & 7 & 0 \\ \hline
+bash (4891) & 0 & 0 & 128 & 0 & 0 & 0 & 0 & 0 & 61 & 0 \\ \hline
+java  (4874) & 0 & 0 & 20 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+bash (4893) & 0 & 0 & 164 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+uname (4881)  & 0 & 0 & 191 & 0 & 0 & 1956  & 6 & 0 & 7 & 0 \\ \hline
+bash (4889) & 0 & 0 & 165 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Observed values of measures of processes on the second rightmost sample in Fig.~\ref{fig:corr_s_wbytes}~\label{tab:breakdown5}}
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|} \hline
+proc name (id) & u time & s time & min flt & maj flt & r bytes & r char & r sysc & w bytes & w char & w sysc\\ \hline
+INC4096 (3559) & 4103948 & 45 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 \\ \hline
+proc\_monitor (25917) & 190 & 10 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+md0\_raid1 (484) & 0 & 5 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+jbd2/md0-8 (497) & 0 & 3 & 0 & 0 & 0 & 0 & 0 & 471040 & 0 & 0 \\ \hline
+java (3549) & 2 & 1 & 1093 & 0 & 0 & 11480 & 20 & 0 & 0 & 0 \\ \hline
+cifsd (1927) & 0 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+flush-9:0 (3548) & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+java (5022) & 1 & 0 & 21 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\ \hline
+ntpd (28232) & 0 & 0 & 1 & 0 & 0 & 0 & 42 & 4096 & 7 & 0 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Observed values of measures of processes on the rightmost sample in Fig.~\ref{fig:corr_s_wbytes}~\label{tab:breakdown6}}
+\end{table}

### 2017-05-24 pt_characterization/rep_check/repet_data2/et_hist_sodb9.r

@@ -313,3 +313,52 @@ xlab='ET (ms)', ylab=expression('Frequency'))
 axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
 dev.off()
 
+x = read.csv(file="8192_sec.dat",head=TRUE,sep="\t")
+setEPS()
+postscript("8192_sec_et_hist.eps")
+#x <- subset(x, x$ITERNUM != 40 & x$ITERNUM != 124 & x$ITERNUM != 207 & x$ITERNUM != 292)
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+binsize=5
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+xmin <-min(x$METIME)
+xmin <- xmin-2
+xmax <-max(x$METIME)
+xmax <- xmax+2
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- max(h$counts)
+ymax <- 40
+plot(h, xaxt="n",freq=TRUE,ylim=c(0,ymax), col="blue", main='', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='ET (ms)', ylab=expression('Frequency'))
+axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+dev.off()
+
+x = read.csv(file="16384_sec.dat",head=TRUE,sep="\t")
+setEPS()
+postscript("16384_sec_et_hist.eps")
+#x <- subset(x, x$ITERNUM != 40 & x$ITERNUM != 124 & x$ITERNUM != 207 & x$ITERNUM != 292)
+x_up = mean(x$METIME) + 2*sd(x$METIME)
+x_dn = mean(x$METIME) - 2*sd(x$METIME)
+binsize=10
+#x = subset(x, x$METIME >= x_dn & x$METIME <= x_up)
+xmin <-min(x$METIME)
+xmin <- xmin-6
[…]

### 2017-05-24 pt_characterization/rep_check/repet_data2/pt_hist_sodb9.r

@@ -306,3 +306,47 @@ sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""),
 xlab='PT (ms)', ylab=expression('Frequency'))
 axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
 dev.off()
+
+x = read.csv(file="8192_sec.dat",head=TRUE,sep="\t")
+setEPS()
+postscript("8192_sec_pt_hist.eps")
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+binsize=5
+xmin <-min(x$PRTIME)
+xmin <- xmin-2
+xmax <-max(x$PRTIME)
+xmax <- xmax+1
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 40
+plot(h, ylim=c(0,ymax), xaxt="n",freq=TRUE,col="green", main='', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+dev.off()
+
+x = read.csv(file="16384_sec.dat",head=TRUE,sep="\t")
+setEPS()
+postscript("16384_sec_pt_hist.eps")
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+binsize=10
+xmin <-min(x$PRTIME)
+xmin <- xmin-7
+xmax <-max(x$PRTIME)
+xmax <- xmax+5
+nbins <- ceiling((xmax-xmin) / binsize)
[…]

### 2017-05-24 pt_characterization/rep_check/u_s_time/scatterplot.r

@@ -1,3 +1,7 @@
+numSamples <- 300
+x = read.csv(file="per_iter_detailed.dat",head=TRUE,sep="\t")
+corr(x$U_TIME, x$S_TIME)
+
 ### sodb9
 numSamples <- 300
 x = read.csv(file="per_iter_detailed.dat",head=TRUE,sep="\t")

### 2017-05-24 pt_characterization/rep_check/u_s_time/scatterplot_INC2048.r

@@ -1,0 +1,8 @@
+numSamples <- 300
+x = read.csv(file="INC2048_per_iter_detailed.dat",head=TRUE,sep="\t")
+corr(x$U_TIME, x$S_TIME)
+
+numSamples <- 300
+x = read.csv(file="per_iter_detailed.dat",head=TRUE,sep="\t")
+min(x$S_TIME)
+corr(x$U_TIME, x$S_TIME)

### 2017-05-25 pt_characterization/rep_check/paper.tex
Changed tex

@@
 If so then, we can say that run repeatability is satisfied in our experiment setting. 
-In this study the used protocol is equivalent to SEDONA~\cite{Sedona}, 
-eliminating samples involving infrequent, long-running daemons, which is the first step of EMPv5.
+%In this study the used protocol is equivalent to SEDONA~\cite{Sedona}, 
+%eliminating samples involving infrequent, long-running daemons, which is the first step of EMPv5.
+In this study the used protocol eliminates samples involving infrequent, long-running daemons, 
+which is the first step of EMPv5.
 
-In the SEDONA protocol, we use taskstats C struct to get measures of a captured process. 
+In the protocol, we use taskstats C struct to get measures of a captured process. 
 The taskstat's data is delivered via a netlink socket from the kernel space. 
@@
 %{Avnur:2000:ECA:342009.335420}
-{Sedona}
-Young-Kyoon Suh, ``SEDONA: A Novel Protocol for Identifying Infrequent, Long-running Daemons on a Linux System'', in {\em IEICE Transactions on Information Systems}, Vol.
+%{Sedona}
+%Young-Kyoon Suh, ``SEDONA: A Novel Protocol for Identifying Infrequent, Long-running Daemons on a Linux System'', in {\em IEICE Transactions on Information Systems}, Vol.
 100D, No.

### 2017-05-31 pt_characterization/rep_check/u_s_time.tex
DRAFT for May 31

@@
 	\centering
+	\subfigure[User time frequency on INC96]{
+		\includegraphics[scale=0.43]{u_s_time/96_sec_ut_hist.eps}
+		\label{fig:inc96_ut_hist}
+	}
 	\subfigure[User time frequency on INC128]{
-		\includegraphics[scale=0.43]{u_s_time/128_sec_ut_hist.eps}
-		\label{fig:inc128_ut_hist}
+		\includegraphics[scale=0.43]{u_s_time/128_sec_ut_hist2.eps}
+		\label{fig:inc128_ut_hist2}
 	}
+	\subfigure[User time frequency on INC160]{
+		\includegraphics[scale=0.43]{u_s_time/160_sec_ut_hist.eps}
+		\label{fig:inc160_ut_hist}
+	}
+	\subfigure[User time frequency on INC192]{
+		\includegraphics[scale=0.43]{u_s_time/192_sec_ut_hist.eps}
+		\label{fig:inc192_ut_hist}
+	}
+	\caption{User Time Histograms of INC96 ... INC192~\label{fig:ut_hist3}}
+\end{figure}
+
+\begin{figure}[hp!]
+	\centering
 	\subfigure[User time frequency on INC256]{
@@
 	}
-	\subfigure[User time frequency on INC512]{
+	\subfigure[User time frequency on INC512 (with one extreme outlier of 513,259 msec removed)]{
 		\includegraphics[scale=0.43]{u_s_time/512_sec_ut_hist.eps}
@@
 	}
-	\caption{User Time Histograms of INC128 ... INC1024~\label{fig:ut_hist3}}
+	\caption{User Time Histograms of INC256 ... INC1024~\label{fig:ut_hist4}}
 \end{figure}
 
-\begin{figure}[hp!]
-	\centering
-	\subfigure[User time frequency on INC2048]{
-		\includegraphics[scale=0.43]{u_s_time/2048_sec_ut_hist.eps}
-		\label{fig:inc2048_ut_hist}
-	}
-	\subfigure[User time frequency on INC4096]{
-		\includegraphics[scale=0.43]{u_s_time/4096_sec_ut_hist.eps}
-		\label{fig:inc4096_ut_hist}
-	}
-	\caption{User Time Histograms of INC2048 and INC4096~\label{fig:ut_hist4}}
-\end{figure}
+%\begin{figure}[hp!]
+%	\centering
+%	\subfigure[User time frequency on INC2048]{
+%		\includegraphics[scale=0.43]{u_s_time/2048_sec_ut_hist.eps}
+%		\label{fig:inc2048_ut_hist}
+%	}
+%	\subfigure[User time frequency on INC4096]{
+%		\includegraphics[scale=0.43]{u_s_time/4096_sec_ut_hist.eps}
+%		\label{fig:inc4096_ut_hist}
+%	}
+%	\caption{User Time Histograms of INC2048 and INC4096~\label{fig:ut_hist5}}
+%\end{figure}
 

### 2017-05-31 pt_characterization/rep_check/u_s_time/u_time_hist.r

@@ -132,6 +132,25 @@ xlab='User time (ms)', ylab=expression('Frequency'))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
+x = read.csv(file="96_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("96_sec_ut_hist.eps")
+binsize=1
+#x = subset(x, x$U_TIME >= x_dn & x$U_TIME <= x_up)
+x <- subset(x, x$ITERNUM != 474)
+xmin <-min(x$U_TIME)
+xmax <-max(x$U_TIME)
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="orange",  main='User time frequency on INC96', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='User time (ms)', ylab=expression('Frequency'))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
 x = read.csv(file="128_sec.dat",head=TRUE,sep="\t")
 x <- subset(x, x$ITERNUM <= numSamples)
 setEPS()
@@ -149,6 +168,72 @@ ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
 ymax <- 80
 plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="orange",  main='User time frequency on INC128', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='User time (ms)', ylab=expression('Frequency'))
+axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+dev.off()
+
+x = read.csv(file="128_sec2.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
[…]

### 2017-06-07 pt_characterization/rep_check/res2.tex
June 7 Draft

@@
 {\tt sodb9} &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-13 $\sim$ 2017-03-14 & Figs.~\ref{fig:s9_r2_et_hist1},~\ref{fig:s9_r2_et_hist2},~\ref{fig:s9_r2_pt_hist1}, and~\ref{fig:s9_r2_pt_hist2}\\ \hline
-{\tt sodb9} &  INC128$\sim$INC1024 & 300 samples, each & 2017-03-14 $\sim$ 2017-03-21 & 
+%{\tt sodb9} &  INC128$\sim$INC1024 & 300 samples, each & 2017-03-14 $\sim$ 2017-03-21 & 
+%Figs.~\ref{fig:s9_r2_et_hist3} and~\ref{fig:s9_r2_pt_hist3}\\ \hline
+{\tt sodb9} &  INC512$\sim$INC1024 & 300 samples, each & 2017-03-17 $\sim$ 2017-03-21 & 
 Figs.~\ref{fig:s9_r2_et_hist3} and~\ref{fig:s9_r2_pt_hist3}\\ \hline

### 2017-06-07 pt_characterization/rep_check/u_s_time.tex
June 7 Draft

@@
 INC with its task length increasing from 1 second to 4096 seconds, via SEDONA. 
-The detailed description of the base data is from Table~\ref{tab:exp_notes2}.
+The detailed description of the base data is from Table~\ref{tab:exp_notes2}, 
+and for some additional data, the corresponding description is given in Table~\ref{tab:exp_notes3}.
 
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{2cm}|p{3cm}|p{3cm}|p{4cm}|p{3.5cm}|} \hline
+Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
+{\tt sodb9} &  INC96$\sim$INC256 & 1000 samples, each & 2017-05-24 $\sim$ 2017-06-06 & 
+Figs.~\ref{fig:ut_hist3}, \ref{fig:inc224_ut_hist}, and~\ref{fig:inc256_ut_hist}\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Notes on experiment runs used for histograms\label{tab:exp_notes3}}
+\end{table}
+
+\newpage
 \subsection{User Time}
@@
 	\centering
+	\subfigure[User time frequency on INC224]{
+		\includegraphics[scale=0.43]{u_s_time/224_sec_ut_hist.eps}
+		\label{fig:inc224_ut_hist}
+	}
 	\subfigure[User time frequency on INC256]{
-		\includegraphics[scale=0.43]{u_s_time/256_sec_ut_hist.eps}
+		\includegraphics[scale=0.43]{u_s_time/256_sec_ut_hist2.eps}
 		\label{fig:inc256_ut_hist}
@@
 	}
-	\caption{User Time Histograms of INC256 ... INC1024~\label{fig:ut_hist4}}
+	\caption{User Time Histograms of INC224 ... INC256~\label{fig:ut_hist4}}
 \end{figure}
 
-%\begin{figure}[hp!]
-%	\centering
-%	\subfigure[User time frequency on INC2048]{
-%		\includegraphics[scale=0.43]{u_s_time/2048_sec_ut_hist.eps}
-%		\label{fig:inc2048_ut_hist}
-%	}
-%	\subfigure[User time frequency on INC4096]{
-%		\includegraphics[scale=0.43]{u_s_time/4096_sec_ut_hist.eps}
-%		\label{fig:inc4096_ut_hist}
-%	}
-%	\caption{User Time Histograms of INC2048 and INC4096~\label{fig:ut_hist5}}
-%\end{figure}
+\begin{figure}[hp!]
+	\centering
+	\subfigure[User time frequency on INC2048]{
+		\includegraphics[scale=0.43]{u_s_time/2048_sec_ut_hist.eps}
+		\label{fig:inc2048_ut_hist}
+	}
+	\subfigure[User time frequency on INC4096]{
+		\includegraphics[scale=0.43]{u_s_time/4096_sec_ut_hist.eps}
+		\label{fig:inc4096_ut_hist}
+	}
+	\caption{User Time Histograms of INC2048 and INC4096~\label{fig:ut_hist5}}
+\end{figure}
 

### 2017-06-07 pt_characterization/rep_check/u_s_time/u_time_hist.r

@@ -223,7 +223,29 @@ setEPS()
 postscript("192_sec_ut_hist.eps")
 binsize=2
 #x = subset(x, x$U_TIME >= x_dn & x$U_TIME <= x_up)
-x <- subset(x, x$ITERNUM != 26 & x$ITERNUM != 472)
+x <- subset(x, x$ITERNUM != 26 & x$ITERNUM != 472 & x$ITERNUM != 922)
+xmin <-min(x$U_TIME)
+xmax <-max(x$U_TIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 250
+plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="orange",  main='User time frequency on INC192', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='User time (ms)', ylab=expression('Frequency'))
+axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+dev.off()
+
+x = read.csv(file="224_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("224_sec_ut_hist.eps")
+binsize=2
+#x = subset(x, x$U_TIME >= x_dn & x$U_TIME <= x_up)
+x <- subset(x, x$ITERNUM != 321 & x$ITERNUM !=704)
 xmin <-min(x$U_TIME)
 xmax <-max(x$U_TIME)
 xmin
@@ -233,7 +255,7 @@ h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
 ymax <- 200
-plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="orange",  main='User time frequency on INC192', 
+plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="orange",  main='User time frequency on INC224', 
[…]

### 2017-06-14 pt_characterization/rep_check/u_s_time.tex
June 14 Draft

@@
 This section exhibits user and system time histograms on the second run of 
-INC with its task length increasing from 1 second to 4096 seconds, via SEDONA. 
+INC with its task length increasing from 1 second to 4096 seconds. 
 The detailed description of the base data is from Table~\ref{tab:exp_notes2}, 
@@
 Figs.~\ref{fig:ut_hist3}, \ref{fig:inc224_ut_hist}, and~\ref{fig:inc256_ut_hist}\\ \hline
+					&  INC3, 6, 12, 24, 48, 72, 80, 88, 104, 112 & 1000 samples, each & 2017-06-07 $\sim$ 2017-06-14 & Figs.~\ref{fig:inc3_ut_hist}, ~\ref{fig:inc6_ut_hist}, ~\ref{fig:inc12_ut_hist}, ~\ref{fig:inc24_ut_hist},
+~\ref{fig:inc48_ut_hist}, ~\ref{fig:inc72_ut_hist}, ~\ref{fig:inc80_ut_hist}, ~\ref{fig:inc88_ut_hist}, ~\ref{fig:inc104_ut_hist}, and ~\ref{fig:inc112_ut_hist}\\ \hline
 \end{tabular}
@@
 \vspace{-.2in}
-\caption{Notes on experiment runs used for histograms\label{tab:exp_notes3}}
+\caption{Notes on experiment runs used for histograms\label{tab:exp_notes4}}
 \end{table}
@@
 	}
+	\subfigure[User time frequency on INC3]{
+		\includegraphics[scale=0.43]{u_s_time/3_sec_ut_hist.eps}
+		\label{fig:inc3_ut_hist}
+	}
 	\subfigure[User time frequency on INC4]{
@@
 		\label{fig:inc4_ut_hist}
+	}
+	\caption{User Time Histograms of INC1 ... INC4~\label{fig:ut_hist1}}
+\end{figure}
+
+\begin{figure}[hp!]
+	\centering
+	\subfigure[User time frequency on INC6]{
+		\includegraphics[scale=0.43]{u_s_time/6_sec_ut_hist.eps}
+		\label{fig:inc6_ut_hist}
 	}
@@
 	}
-	\caption{User Time Histograms of INC1 ... INC8~\label{fig:ut_hist1}}
+	\subfigure[User time frequency on INC12]{
+		\includegraphics[scale=0.43]{u_s_time/12_sec_ut_hist.eps}
+		\label{fig:inc12_ut_hist}
+	}
+	\subfigure[User time frequency on INC16]{
+		\includegraphics[scale=0.43]{u_s_time/16_sec_ut_hist.eps}
+		\label{fig:inc16_ut_hist}
+	}
+	\caption{User Time Histograms of INC6 ... INC16~\label{fig:ut_hist2}}
 \end{figure}
@@
 	\centering
-	\subfigure[User time frequency on INC16]{
-		\includegraphics[scale=0.43]{u_s_time/16_sec_ut_hist.eps}
-		\label{fig:inc16_ut_hist}
+	\subfigure[User time frequency on INC24]{
+		\includegraphics[scale=0.43]{u_s_time/24_sec_ut_hist.eps}
+		\label{fig:inc24_ut_hist}
 	}
@@
 	}
+	\subfigure[User time frequency on INC48]{
+		\includegraphics[scale=0.43]{u_s_time/48_sec_ut_hist.eps}
+		\label{fig:inc48_ut_hist}
+	}
 	\subfigure[User time frequency on INC64]{
@@
 	}
-	\caption{User Time Histograms of INC16 ... INC64~\label{fig:ut_hist2}}
+	\caption{User Time Histograms of INC24 ... INC64~\label{fig:ut_hist3}}
 \end{figure}
@@
 	\centering
+	\subfigure[User time frequency on INC72]{
+		\includegraphics[scale=0.43]{u_s_time/72_sec_ut_hist.eps}
+		\label{fig:inc72_ut_hist}
+	}
+	\subfigure[User time frequency on INC80]{
+		\includegraphics[scale=0.43]{u_s_time/80_sec_ut_hist.eps}
+		\label{fig:inc80_ut_hist}
+	}
+	\subfigure[User time frequency on INC88]{
+		\includegraphics[scale=0.43]{u_s_time/88_sec_ut_hist.eps}
+		\label{fig:inc88_ut_hist}
+	}
 	\subfigure[User time frequency on INC96]{
@@
 	}
+	\caption{User Time Histograms of INC72 ... INC96~\label{fig:ut_hist4}}
+\end{figure}
+
+\begin{figure}[hp!]
+	\centering
+	\subfigure[User time frequency on INC104]{
+		\includegraphics[scale=0.43]{u_s_time/104_sec_ut_hist.eps}
+		\label{fig:inc104_ut_hist}
+	}
+	\subfigure[User time frequency on INC112]{
+		\includegraphics[scale=0.43]{u_s_time/112_sec_ut_hist.eps}
+		\label{fig:inc112_ut_hist}
+	}
+	\caption{User Time Histograms of INC104 and INC112~\label{fig:ut_hist5}}
+\end{figure}
+
+\begin{figure}[hp!]
+	\centering
 	\subfigure[User time frequency on INC128]{
@@
 	}
-	\caption{User Time Histograms of INC96 ... INC192~\label{fig:ut_hist3}}
+	\caption{User Time Histograms of INC128 ... INC192~\label{fig:ut_hist6}}
 \end{figure}
@@
 	}
-	\caption{User Time Histograms of INC224 ... INC256~\label{fig:ut_hist4}}
+	\caption{User Time Histograms of INC224 ... INC256~\label{fig:ut_hist7}}
 \end{figure}
@@
 	}
-	\caption{User Time Histograms of INC2048 and INC4096~\label{fig:ut_hist5}}
+	\caption{User Time Histograms of INC2048 and INC4096~\label{fig:ut_hist8}}
 \end{figure}

### 2017-06-14 pt_characterization/rep_check/u_s_time/u_time_hist.r

@@ -36,6 +36,24 @@ xlab='User time (ms)', ylab=expression('Frequency'))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
+x = read.csv(file="3_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("3_sec_ut_hist.eps")
+binsize=1
+#x = subset(x, x$U_TIME >= x_dn & x$U_TIME <= x_up)
+xmin <-min(x$U_TIME)
+xmax <-max(x$U_TIME)
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="orange",  main='User time frequency on INC3', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='User time (ms)', ylab=expression('Frequency'))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
 x = read.csv(file="4_sec.dat",head=TRUE,sep="\t")
 x <- subset(x, x$ITERNUM <= numSamples)
 setEPS()
@@ -49,6 +67,24 @@ h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
 plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="orange",  main='User time frequency on INC4', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='User time (ms)', ylab=expression('Frequency'))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+x = read.csv(file="6_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("6_sec_ut_hist.eps")
+binsize=1
+#x = subset(x, x$U_TIME >= x_dn & x$U_TIME <= x_up)
[…]

### 2017-06-16 pt_characterization/rep_check/u_s_time.tex
June 16 Draft

@@
 Figs.~\ref{fig:ut_hist3}, \ref{fig:inc224_ut_hist}, and~\ref{fig:inc256_ut_hist}\\ \hline
-					&  INC3, 6, 12, 24, 48, 72, 80, 88, 104, 112 & 1000 samples, each & 2017-06-07 $\sim$ 2017-06-14 & Figs.~\ref{fig:inc3_ut_hist}, ~\ref{fig:inc6_ut_hist}, ~\ref{fig:inc12_ut_hist}, ~\ref{fig:inc24_ut_hist},
-~\ref{fig:inc48_ut_hist}, ~\ref{fig:inc72_ut_hist}, ~\ref{fig:inc80_ut_hist}, ~\ref{fig:inc88_ut_hist}, ~\ref{fig:inc104_ut_hist}, and ~\ref{fig:inc112_ut_hist}\\ \hline
+					&  INC3, 6, 12, 24, 48, 72, 80, 88, 104, 112, and 120 & 1000 samples, each & 2017-06-07 $\sim$ 2017-06-16 & Figs.~\ref{fig:inc3_ut_hist}, ~\ref{fig:inc6_ut_hist}, ~\ref{fig:inc12_ut_hist}, ~\ref{fig:inc24_ut_hist},
+~\ref{fig:inc48_ut_hist}, ~\ref{fig:inc72_ut_hist}, ~\ref{fig:inc80_ut_hist}, ~\ref{fig:inc88_ut_hist}, ~\ref{fig:inc104_ut_hist},~\ref{fig:inc112_ut_hist}, and~\ref{fig:inc120_ut_hist}\\ \hline
 \end{tabular}
@@
 	}
-	\caption{User Time Histograms of INC104 and INC112~\label{fig:ut_hist5}}
+	\subfigure[User time frequency on INC120]{
+		\includegraphics[scale=0.43]{u_s_time/120_sec_ut_hist.eps}
+		\label{fig:inc120_ut_hist}
+	}
+	\subfigure[User time frequency on INC128]{
+		\includegraphics[scale=0.43]{u_s_time/128_sec_ut_hist2.eps}
+		\label{fig:inc128_ut_hist2}
+	}
+	\caption{User Time Histograms of INC104 and INC128~\label{fig:ut_hist5}}
 \end{figure}
@@
 	\centering
-	\subfigure[User time frequency on INC128]{
-		\includegraphics[scale=0.43]{u_s_time/128_sec_ut_hist2.eps}
-		\label{fig:inc128_ut_hist2}
-	}
 	\subfigure[User time frequency on INC160]{
@@
 	}
-	\caption{User Time Histograms of INC128 ... INC192~\label{fig:ut_hist6}}
-\end{figure}
-
-\begin{figure}[hp!]
-	\centering
 	\subfigure[User time frequency on INC224]{
@@
 	}
+	\caption{User Time Histograms of INC160 ... INC256~\label{fig:ut_hist6}}
+\end{figure}
+
+\begin{figure}[hp!]
+	\centering
 	\subfigure[User time frequency on INC512 (with one extreme outlier of 513,259 msec removed)]{
@@
 	}
-	\caption{User Time Histograms of INC224 ... INC256~\label{fig:ut_hist7}}
-\end{figure}
-
-\begin{figure}[hp!]
-	\centering
 	\subfigure[User time frequency on INC2048]{
@@
 	}
-	\caption{User Time Histograms of INC2048 and INC4096~\label{fig:ut_hist8}}
+	\caption{User Time Histograms of INC512 ... INC4096~\label{fig:ut_hist7}}
 \end{figure}

### 2017-06-16 pt_characterization/rep_check/u_s_time/u_time_hist.r

@@ -337,6 +337,47 @@ xlab='User time (ms)', ylab=expression('Frequency'))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
+numSamples <- 1000
+x = read.csv(file="120_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("120_sec_ut_hist1.eps")
+binsize=1
+#x = subset(x, x$U_TIME >= x_dn & x$U_TIME <= x_up)
+x <- subset(x, x$ITERNUM != 80 & x$ITERNUM != 797)
+xmin <-min(x$U_TIME)
+xmax <-max(x$U_TIME)
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
+ymax <- max(h$counts)
+ymax <- 150
+plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="orange",  main='User time frequency on INC120', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='User time (ms)', ylab=expression('Frequency'))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+x = read.csv(file="120_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("120_sec_ut_hist.eps")
+binsize=2
+x <- subset(x, x$ITERNUM != 80 & x$ITERNUM != 797)
+xmin <-min(x$U_TIME)
+xmax <-max(x$U_TIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 300
+plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="orange",  main='User time frequency on INC120', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='User time (ms)', ylab=expression('Frequency'))
[…]

### 2017-06-16 pt_characterization/rep_check/u_s_time.tex
Updated Draft for June 16

@@
 	}
-	\subfigure[User time frequency on INC120]{
+	\subfigure[User time frequency on INC120 with bin size = 1 ms]{
+		\includegraphics[scale=0.43]{u_s_time/120_sec_ut_hist1.eps}
+		\label{fig:inc120_ut_hist1}
+	}
+	\subfigure[User time frequency on INC120 with bin size = 2 ms]{
 		\includegraphics[scale=0.43]{u_s_time/120_sec_ut_hist.eps}
@@
 	}
+	\caption{User Time Histograms of INC104 and INC120~\label{fig:ut_hist5}}
+\end{figure}
+
+\begin{figure}[hp!]
+	\centering
 	\subfigure[User time frequency on INC128]{
@@
 	}
-	\caption{User Time Histograms of INC104 and INC128~\label{fig:ut_hist5}}
-\end{figure}
-
-\begin{figure}[hp!]
-	\centering
 	\subfigure[User time frequency on INC160]{
@@
 	}
+	\caption{User Time Histograms of INC160 ... INC224~\label{fig:ut_hist6}}
+\end{figure}
+
+\begin{figure}[hp!]
+	\centering
 	\subfigure[User time frequency on INC256]{
@@
 	}
-	\caption{User Time Histograms of INC160 ... INC256~\label{fig:ut_hist6}}
-\end{figure}
-
-\begin{figure}[hp!]
-	\centering
 	\subfigure[User time frequency on INC512 (with one extreme outlier of 513,259 msec removed)]{
@@
 	}
+	\caption{User Time Histograms of INC256 ... INC1024~\label{fig:ut_hist7}}
+\end{figure}
+
+\begin{figure}[hp!]
+	\centering
 	\subfigure[User time frequency on INC2048]{
@@
 	}
-	\caption{User Time Histograms of INC512 ... INC4096~\label{fig:ut_hist7}}
+	\caption{User Time Histograms of INC512 ... INC4096~\label{fig:ut_hist8}}
 \end{figure}

### 2017-07-04 pt_characterization/rep_check/u_s_time.tex
3d plot generation

@@
-\section{Further invstigation of the Second Run~\label{sec:u_s_time_hist}} 
+\section{Further investigation of the Second Run~\label{sec:u_s_time_hist}} 
 This section exhibits user and system time histograms on the second run of 

### 2017-07-04 pt_characterization/rep_check/repet_data1/3d_plot.r

@@ -1,0 +1,47 @@
+### sodb9
+#install.packages("devtools")
+#library(devtools)
+#install.packages("rgl")
+#library(rgl)
+require(rgl)
+
+install.packages("misc3d")
+install.packages("plot3D", repos="http://R-Forge.R-project.org")
+require(plot3D)
+library(plot3D)
+### 1 sec
+x1 = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
+binsize=1
+xmin <-min(x1$PRTIME)
+xmax <-max(x1$PRTIME)
+nbins <- ceiling((xmax-xmin) / binsize)
+h1 = hist(x1$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+z1 <- h1$mids
+z1[z1!=0] <- 1
+### 2 sec
+x2 = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
+x2 <- subset(x2, x2$ITERNUM <= numSamples)
+xmin <-min(x2$PRTIME)
+xmax <-max(x2$PRTIME)
+nbins <- ceiling((xmax-xmin) / binsize)
+h2 = hist(x2$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+z2 <- h2$mids
+z2[z2!=0] <- 1
+### 4 sec
+x4 = read.csv(file="4_sec.dat",head=TRUE,sep="\t")
+x4 <- subset(x4, x4$ITERNUM <= numSamples)
+xmin <-min(x4$PRTIME)
+xmax <-max(x4$PRTIME)
+nbins <- ceiling((xmax-xmin) / binsize)
+h4 = hist(x4$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+z4 <- h4$mids
+z4[z4!=0] <- 1
+
+x <- c(h1$mids,h2$mids,h4$mids)
+y <- c(h1$counts,h2$counts,h4$counts)
+z <- c(z1, z2, z4)
+
+
+hist3D(x,y,z, zlim=c(0,50), theta=40, phi=40, axes=TRUE,label=TRUE, nticks=5, 
+        ticktype="detailed", space=0.5, lighting=TRUE, light="diffuse", shade=0.5)
+

### 2017-07-07 pt_characterization/rep_check/res2.tex
Draft being ready for July 10

@@
 This section exhibits histograms on the second run of 
-INC with its task length increasing from 1 second to 4096 seconds, via SEDONA. 
+INC with its task length increasing from 1 second to 4096 seconds, via EMPv5 without Step 2. 
 The detailed description of the base data is from Table~\ref{tab:exp_notes2}.

### 2017-07-07 pt_characterization/rep_check/res3.tex
Draft being ready for July 10

\section{A Characterization Study on the First Run~\label{sec:sodb9_r3_hist}} 
This section further discusses the first run of 
INC with its task length increasing from 1 second to 4096 seconds via EMPv5 with no Step 2. 
The detailed description of the base data is from Table~\ref{tab:exp_notes3}.

%\begin{table}[h]
%\begin{center}
%\begin{tabular}{|p{2cm}|p{3cm}|p{3cm}|p{4cm}|p{3.5cm}|} \hline
%Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
%{\tt sodb9} &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-13 $\sim$ 2017-03-14 & %Figs.~\ref{fig:s9_r2_et_hist1},~\ref{fig:s9_r2_et_hist2},~\ref{fig:s9_r2_pt_hist1}, and~\ref{fig:s9_r2_pt_hist2}\\ \hline
%{\tt sodb9} &  INC512$\sim$INC1024 & 300 samples, each & 2017-03-17 $\sim$ 2017-03-21 & 
%%Figs.~\ref{fig:s9_r2_et_hist3} and~\ref{fig:s9_r2_pt_hist3}\\ \hline
%{\tt sodb10} & INC2048 & 300 samples & 2017-03-13 $\sim$ 2017-03-20 & 
%%Figs.~\ref{fig:inc2048_r2_et_hist_v5} and~\ref{fig:inc2048_r2_hist_v5}\\ \hline
%{\tt sodb12} & INC4096 & 300 samples & 2017-03-02 $\sim$ 2017-03-17 & 
%%Figs.~\ref{fig:inc4096_r2_et_hist_v5} and~\ref{fig:inc4096_r2_hist_v5}\\ \hline
%{\tt sodb10} & INC8192& 261 samples & 2017-04-27 $\sim$ 2017-05-21 & 
%%Figs.~\ref{fig:inc8192_r2_et_hist_v5} and~\ref{fig:inc8192_r2_hist_v5}\\ \hline
%{\tt sodb12} & INC16384& 130 samples & 2017-04-27 $\sim$ 2017-05-21 & 
%%Figs.~\ref{fig:inc16384_r2_et_hist_v5} and~\ref{fig:inc16384_r2_hist_v5}\\ \hline
%\end{tabular}
%\end{center}
%\vspace{-.2in}
%\caption{Notes on experiment runs used for histograms\label{tab:exp_notes3}}
%\end{table}
%
%INC8192/INC16384 unfortunately stopped in the middle of their runs due to a frozen vnc problem. So couldn't finish 300 samples.
%
%\pagebreak

\subsection{Summary of Histograms}

\begin{figure}[hp!]
\vspace{-.3in}
	\centering
	\includegraphics[scale=0.5,angle=270]{u_s_time/3d_plot.eps}\label{fig:3d_plot}
	\caption{3D histograms on INC~\label{fig:hist3d}}
\end{figure}

\subsection{Absolute and Relative Variance over Increasing Task Length}

\begin{figure}[hp!]
	\centering
	\subfigure[Standard deviation (absolute)]{
		\includegraphics[scale=0.6]{u_s_time/overall_std.eps}
		\label{fig:overall_std}
	}
	\subfigure[Coefficient of variation (relative)]{
		\includegraphics[scale=0.6]{u_s_time/overall_re.eps}
		\label{fig:overall_re}
	}
	\caption{Absolute and relative variances~\label{fig:cv_inc}}
\end{figure}

\pagebreak
\newpage

\subsection{ET}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC1]{
		\includegraphics[scale=0.43]{u_s_time/1_sec_et_hist.eps}
		\label{fig:inc1_r3_et_hist_v5}
	}
	\subfigure[ET frequency on INC2]{
		\includegraphics[scale=0.43]{u_s_time/2_sec_et_hist.eps}
		\label{fig:inc2_r3_et_hist_v5}
	}
	\subfigure[ET frequency on INC4]{
		\includegraphics[scale=0.43]{u_s_time/4_sec_et_hist.eps}
		\label{fig:inc4_r3_et_hist_v5}
	}
	\subfigure[ET frequency on INC8]{
		\includegraphics[scale=0.43]{u_s_time/8_sec_et_hist.eps}
		\label{fig:inc8_r3_et_hist_v5}
	}
	\caption{ET Histograms of INC1 ... INC8~\label{fig:s9_r3_et_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC16]{
		\includegraphics[scale=0.43]{u_s_time/16_sec_et_hist.eps}
		\label{fig:inc16_r3_et_hist_v5}
	}
	\subfigure[ET frequency on INC32]{
		\includegraphics[scale=0.43]{u_s_time/32_sec_et_hist.eps}
		\label{fig:inc32_r3_et_hist_v5}
	}
	\subfigure[ET frequency on INC64]{
		\includegraphics[scale=0.43]{u_s_time/64_sec_et_hist.eps}
		\label{fig:inc64_r3_et_hist_v5}
	}
	\caption{ET Histograms of INC16 ... INC64~\label{fig:s9_r3_et_hist2}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC128]{
		\includegraphics[scale=0.43]{u_s_time/128_sec_et_hist.eps}
		\label{fig:inc128_r3_et_hist_v5}
	}
	\subfigure[ET frequency on INC256]{
		\includegraphics[scale=0.43]{u_s_time/256_sec_et_hist.eps}
		\label{fig:inc256_r3_et_hist_v5}
	}
	\subfigure[ET frequency on INC512]{
		\includegraphics[scale=0.43]{u_s_time/512_sec_et_hist.eps}
		\label{fig:inc512_r3_et_hist_v5}
	}
	\subfigure[ET frequency on INC1024]{
		\includegraphics[scale=0.43]{u_s_time/1024_sec_et_hist.eps}
		\label{fig:inc1024_r3_et_hist_v5}
	}
	\caption{ET Histograms of INC128 ... INC1024~\label{fig:s9_r3_et_hist3}}
\end{figure}

%\begin{figure}[hp!]
%	\centering
%	\subfigure[ET frequency on INC2048]{
%		\includegraphics[scale=0.43]{u_s_time/2048_sec_et_hist.eps}
%		\label{fig:inc2048_r3_et_hist_v5}
%	}
%	\subfigure[ET frequency on INC4096]{
%		\includegraphics[scale=0.43]{u_s_time/4096_sec_et_hist.eps}
%		\label{fig:inc4096_r3_et_hist_v5}
%	}
%	\caption{ET Histograms of INC2048 ... INC4096~\label{fig:s9_r3_et_hist4}}
%\end{figure}

\vspace\fill
\clearpage

\subsection{PT}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC1]{
		\includegraphics[scale=0.43]{u_s_time/1_sec_pt_hist.eps}
		\label{fig:inc1_r3_hist_v5}
	}
	\subfigure[PT frequency on INC2]{
		\includegraphics[scale=0.43]{u_s_time/2_sec_pt_hist.eps}
		\label{fig:inc2_r3_hist_v5}
	}
	\subfigure[PT frequency on INC4]{
		\includegraphics[scale=0.43]{u_s_time/4_sec_pt_hist.eps}
		\label{fig:inc4_r3_hist_v5}
	}
	\subfigure[PT frequency on INC8]{
		\includegraphics[scale=0.43]{u_s_time/8_sec_pt_hist.eps}
		\label{fig:inc8_r3_hist_v5}
	}
	\caption{PT Histograms of INC1 ... INC8~\label{fig:s9_r3_pt_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC16]{
		\includegraphics[scale=0.43]{u_s_time/16_sec_pt_hist.eps}
		\label{fig:inc16_r3_hist_v5}
	}
	\subfigure[PT frequency on INC32]{
		\includegraphics[scale=0.43]{u_s_time/32_sec_pt_hist.eps}
		\label{fig:inc32_r3_hist_v5}
	}
	\subfigure[PT frequency on INC64]{
		\includegraphics[scale=0.43]{u_s_time/64_sec_pt_hist.eps}
		\label{fig:inc64_r3_hist_v5}
	}
	\caption{PT Histograms of INC16 ... INC64\label{fig:s9_r3_pt_hist2}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC128]{
		\includegraphics[scale=0.43]{u_s_time/128_sec_pt_hist.eps}
		\label{fig:inc128_r3_hist_v5}
	}
	\subfigure[PT frequency on INC256]{
		\includegraphics[scale=0.43]{u_s_time/256_sec_pt_hist.eps}
		\label{fig:inc256_r3_hist_v5}
	}
	\subfigure[PT frequency on INC512]{
		\includegraphics[scale=0.43]{u_s_time/512_sec_pt_hist.eps}
		\label{fig:inc512_r3_hist_v5}
	}
	\subfigure[PT frequency on INC1024]{
		\includegraphics[scale=0.43]{u_s_time/1024_sec_pt_hist.eps}
		\label{fig:inc1024_r3_hist_v5}
	}
	\caption{PT Histograms of INC256 ... INC1024~\label{fig:s9_r3_pt_hist3}}
\end{figure}

%\begin{figure}[t]
%	\centering
%	\subfigure[PT frequency on INC2048]{
%		\includegraphics[scale=0.43]{u_s_time/2048_sec_pt_hist.eps}
%		\label{fig:inc2048_r3_hist_v5}
%	}
%	\subfigure[PT frequency on INC4096]{
%		\includegraphics[scale=0.43]{u_s_time/4096_sec_pt_hist.eps}
%		\label{fig:inc4096_r3_hist_v5}
%	}
%%	\subfigure[PT frequency on INC8192]{
%%		\includegraphics[scale=0.43]{u_s_time/8192_sec_pt_hist.eps}
%%		\label{fig:inc8192_r3_hist_v5}
%%	}
%%	\subfigure[PT frequency on INC16384]{
%%		\includegraphics[scale=0.43]{u_s_time/16384_sec_pt_hist.eps}
%%		\label{fig:inc16384_r3_hist_v5}
%%	}
%	\caption{PT Histograms of INC2048 ... INC16384~\label{fig:s9_r3_pt_hist4}}
%\end{figure}

\clearpage

### 2017-07-07 pt_characterization/rep_check/u_s_time.tex
Draft being ready for July 10

--- 2017-06-16 pt_characterization/rep_check/u_s_time.tex
@@
-\section{Further invstigation of the Second Run~\label{sec:u_s_time_hist}} 
+\section{Further Investigation of the First Run and Some Additional Runs~\label{sec:u_s_time_hist}} 
 This section exhibits user and system time histograms on the second run of 
-INC with its task length increasing from 1 second to 4096 seconds. 
-The detailed description of the base data is from Table~\ref{tab:exp_notes2}, 
+INC with its task length increasing from 1 second to 4096 seconds via EMPv5 with no Step 2.
+The detailed description of the base data is from Table~\ref{tab:exp_notes1}, 
 and for some additional data, the corresponding description is given in Table~\ref{tab:exp_notes3}.
@@
 ~\ref{fig:inc48_ut_hist}, ~\ref{fig:inc72_ut_hist}, ~\ref{fig:inc80_ut_hist}, ~\ref{fig:inc88_ut_hist}, ~\ref{fig:inc104_ut_hist},~\ref{fig:inc112_ut_hist}, and~\ref{fig:inc120_ut_hist}\\ \hline
+%%%%% INC2048 and INC4096 come from the second run
 \end{tabular}
@@
 
+\subsection{Summary of Histograms}
+
+\begin{figure}[htp!]
+\vspace{-.3in}
+	\centering
+	\includegraphics[scale=0.5,angle=270]{u_s_time/3d_plot.eps}\label{fig:3d_plot}
+	\caption{3D histograms on INC~\label{fig:hist3d}}
+\end{figure}
+
+\clearpage
+\pagebreak
+
+\subsection{Absolute and Relative Variance over Increasing Task Length}
+
+\begin{figure}[htp!]
+	\centering
+	\subfigure[Standard deviation (absolute)]{
+		\includegraphics[scale=0.6]{u_s_time/overall_std.eps}
+		\label{fig:overall_std}
+	}
+	\subfigure[Coefficient of variation (relative)]{
+		\includegraphics[scale=0.6]{u_s_time/overall_re.eps}
+		\label{fig:overall_re}
+	}
+	\caption{Absolute and relative variances~\label{fig:cv_inc}}
+\end{figure}
+
+\pagebreak
 \newpage
+
+\subsection{Additional Histograms}
+
+\subsection{ET}
+
+%\begin{figure}[hp!]
+%	\centering
+%	\subfigure[User time frequency on INC3]{
+%		\includegraphics[scale=0.43]{u_s_time/3_sec_et_hist.eps}
+%		\label{fig:inc3_ut_hist}
+%	}
+%	\caption{User Time Histograms of INC1 ... INC4~\label{fig:ut_hist1}}
+%\end{figure}
+
+\subsection{PT}
+
+%\begin{figure}[hp!]
+%	\centering
+%	\subfigure[User time frequency on INC3]{
+%		\includegraphics[scale=0.43]{u_s_time/3_sec_et_hist.eps}
+%		\label{fig:inc3_ut_hist}
+%	}
+%	\caption{User Time Histograms of INC1 ... INC4~\label{fig:ut_hist1}}
+%\end{figure}
+
+\pagebreak
+\newpage
+
 \subsection{User Time}

### 2017-07-07 pt_characterization/rep_check/u_s_time/pt_hist.r

@@ -3,599 +3,266 @@ numSamples <- 1000
 x = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
 x <- subset(x, x$ITERNUM <= numSamples)
 setEPS()
-postscript("1_sec_ut_hist.eps")
+postscript("1_sec_pt_hist.eps")
 binsize=1
-#x = subset(x, x$U_TIME >= x_dn & x$U_TIME <= x_up)
-xmin <-min(x$U_TIME)
-xmax <-max(x$U_TIME)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
-h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
-plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="orange",  main='User time frequency on INC1', 
+plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green",  main='PT frequency on INC1', 
 sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
-xlab='User time (ms)', ylab=expression('Frequency'))
+xlab='PT (ms)', ylab=expression('Frequency'))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
 x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
 x <- subset(x, x$ITERNUM <= numSamples)
 setEPS()
-postscript("2_sec_ut_hist.eps")
+postscript("2_sec_pt_hist.eps")
 binsize=1
-#x = subset(x, x$U_TIME >= x_dn & x$U_TIME <= x_up)
-xmin <-min(x$U_TIME)
-xmax <-max(x$U_TIME)
+#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
-h = hist(x$U_TIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
[…]

### 2017-07-13 pt_characterization/rep_check/u_s_time.tex
July 12 Draft

@@
 The detailed description of the base data is from Table~\ref{tab:exp_notes1}, 
-and for some additional data, the corresponding description is given in Table~\ref{tab:exp_notes3}.
+and for some additional data, the corresponding description is given in Table~\ref{tab:exp_notes4}.
 
@@
 	\centering
-	\includegraphics[scale=0.5,angle=270]{u_s_time/3d_plot.eps}\label{fig:3d_plot}
+	\includegraphics[scale=0.6,angle=270]{u_s_time/3d_plot.eps}\label{fig:3d_plot}
 	\caption{3D histograms on INC~\label{fig:hist3d}}
@@
 
-\subsection{ET}
+\subsubsection{ET}
+Not available at this point, due to the labshelf server's unavailability for the time being.
 
-%\begin{figure}[hp!]
-%	\centering
-%	\subfigure[User time frequency on INC3]{
-%		\includegraphics[scale=0.43]{u_s_time/3_sec_et_hist.eps}
-%		\label{fig:inc3_ut_hist}
-%	}
-%	\caption{User Time Histograms of INC1 ... INC4~\label{fig:ut_hist1}}
-%\end{figure}
+\subsubsection{PT}
+The histograms of INC1 through INC4096 are 
+the same as those of Figures~\ref{fig:s9_r1_pt_hist1},~\ref{fig:s9_r1_pt_hist2},~\ref{fig:s9_r1_pt_hist3}, and~\ref{fig:s9_r1_pt_hist4}. 
+The following are the histograms of INC with some intermediate task lengths.
 
-\subsection{PT}
+\begin{figure}[hp!]
+	\centering
+	\subfigure[PT frequency on INC3]{
+		\includegraphics[scale=0.43]{u_s_time/3_sec_pt_hist.eps}
+		\label{fig:inc3_pt_hist}
+	}
+	\subfigure[PT frequency on INC6]{
+		\includegraphics[scale=0.43]{u_s_time/6_sec_pt_hist.eps}
+		\label{fig:inc6_pt_hist}
+	}
+	\subfigure[PT frequency on INC12]{
+		\includegraphics[scale=0.43]{u_s_time/12_sec_pt_hist.eps}
+		\label{fig:inc12_pt_hist}
+	}
+	\subfigure[PT frequency on INC24]{
+		\includegraphics[scale=0.43]{u_s_time/24_sec_pt_hist.eps}
+		\label{fig:inc24_pt_hist}
+	}
+	\caption{PT Histograms of INC3 ... INC24~\label{fig:new_pt_hist1}}
+\end{figure}
 
-%\begin{figure}[hp!]
-%	\centering
-%	\subfigure[User time frequency on INC3]{
-%		\includegraphics[scale=0.43]{u_s_time/3_sec_et_hist.eps}
-%		\label{fig:inc3_ut_hist}
-%	}
-%	\caption{User Time Histograms of INC1 ... INC4~\label{fig:ut_hist1}}
-%\end{figure}
+\pagebreak
+\newpage
+
+\begin{figure}[hp!]
+	\centering
+	\subfigure[PT frequency on INC48]{
+		\includegraphics[scale=0.43]{u_s_time/48_sec_pt_hist.eps}
+		\label{fig:inc48_pt_hist}
+	}
+	\subfigure[PT frequency on INC72]{
+		\includegraphics[scale=0.43]{u_s_time/72_sec_pt_hist.eps}
+		\label{fig:inc72_pt_hist}
+	}
+	\subfigure[PT frequency on INC80]{
+		\includegraphics[scale=0.43]{u_s_time/80_sec_pt_hist.eps}
+		\label{fig:inc80_pt_hist}
+	}
+	\subfigure[PT frequency on INC88]{
+		\includegraphics[scale=0.43]{u_s_time/88_sec_pt_hist.eps}
+		\label{fig:inc88_pt_hist}
+	}
+	\caption{PT Histograms of INC48 ... INC88~\label{fig:new_pt_hist2}}
+\end{figure}
+
+\pagebreak
+\newpage
+
+\begin{figure}[hp!]
+	\centering
+	\subfigure[PT frequency on INC96]{
+		\includegraphics[scale=0.43]{u_s_time/96_sec_pt_hist.eps}
+		\label{fig:inc96_pt_hist}
+	}
+	\subfigure[PT frequency on INC104]{
+		\includegraphics[scale=0.43]{u_s_time/104_sec_pt_hist.eps}
+		\label{fig:inc104_pt_hist}
+	}
+	\subfigure[PT frequency on INC112]{
+		\includegraphics[scale=0.43]{u_s_time/112_sec_pt_hist.eps}
+		\label{fig:inc112_pt_hist}
+	}
+	\subfigure[PT frequency on INC120]{
+		\includegraphics[scale=0.43]{u_s_time/120_sec_pt_hist.eps}
+		\label{fig:inc120_pt_hist}
+	}
+	\caption{PT Histograms of INC96 ... INC120~\label{fig:new_pt_hist3}}
+\end{figure}
+
+\pagebreak
+\newpage
+
+\begin{figure}[hp!]
+	\centering
+	\subfigure[PT frequency on INC160]{
+		\includegraphics[scale=0.43]{u_s_time/160_sec_pt_hist.eps}
+		\label{fig:inc160_pt_hist}
+	}
+	\subfigure[PT frequency on INC192]{
+		\includegraphics[scale=0.43]{u_s_time/192_sec_pt_hist.eps}
+		\label{fig:inc192_pt_hist}
+	}
+	\subfigure[PT frequency on INC224]{
+		\includegraphics[scale=0.43]{u_s_time/224_sec_pt_hist.eps}
+		\label{fig:inc224_pt_hist}
+	}
+	\caption{PT Histograms of INC169 ... INC224~\label{fig:new_pt_hist4}}
+\end{figure}
 

### 2017-07-13 pt_characterization/rep_check/u_s_time/3d_plot.r

@@ -1,13 +1,5 @@
-### sodb9
-#install.packages("devtools")
-#library(devtools)
-#install.packages("rgl")
-#library(rgl)
+#### An R script for 3D plot
 require(rgl)
-#install.packages("misc3d")
-#install.packages("plot3D", repos="http://R-Forge.R-project.org")
-require(plot3D)
-require(axis3D)
 library(plot3D)
 numSamples <- 1000
 ### 1 sec
@@ -17,11 +9,13 @@ xmin <-min(x1$PRTIME)
 xmax <-max(x1$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h1 = hist(x1$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
-#x1_pt <- (h1$mids-1000)
-x1_pt <- h1$mids-min(h1$mids)
-#x1_pt <- ((h1$mids-min(h1$mids))/(max(h1$mids)-min(h1$mids)))
-y1_pt <- 1
-z1_pt <- h1$density
+x1_pt <- (h1$mids-1000)/1
+y1_raw_pt <- 1
+y1_pt <- 0
+z1_pt <- h1$counts/numSamples
+z1_raw_pt <- h1$counts
+inc1 <- rep(1,length(h1$mids))
+ss_inc1 <- rep(numSamples,length(h1$mids)) 
 ### 2 sec
 x2 = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
 x2 <- subset(x2, x2$ITERNUM <= numSamples)
@@ -29,11 +23,13 @@ xmin <-min(x2$PRTIME)
 xmax <-max(x2$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h2 = hist(x2$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
-#x2_pt <- (h2$mids-2000)
-x2_pt <- (h2$mids-min(h2$mids))
-#x2_pt <- ((h2$mids-min(h2$mids))/(max(h2$mids)-min(h2$mids)))
-y2_pt <- 2
-z2_pt <- h2$density
+x2_pt <- (h2$mids-2000)/2
+y2_raw_pt <- 2
+y2_pt <- 1
+z2_pt <- h2$counts/numSamples
+z2_raw_pt <- h2$counts
+inc2 <- rep(2,length(h2$mids))
[…]

### 2017-07-13 pt_characterization/rep_check/u_s_time/3d_plot_bak.r

@@ -1,0 +1,651 @@
+#### An R script for 3D plot
+require(rgl)
+library(plot3D)
+numSamples <- 1000
+### 1 sec
+x1 = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
+binsize=1
+xmin <-min(x1$PRTIME)
+xmax <-max(x1$PRTIME)
+nbins <- ceiling((xmax-xmin) / binsize)
+h1 = hist(x1$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+x1_pt <- (h1$mids-1000)/1
+y1_raw_pt <- 1
+y1_pt <- 0
+z1_pt <- h1$counts/numSamples
+z1_raw_pt <- h1$counts
+inc1 <- rep(1,length(h1$mids))
+ss_inc1 <- rep(numSamples,length(h1$mids)) 
+### 2 sec
+x2 = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
+x2 <- subset(x2, x2$ITERNUM <= numSamples)
+xmin <-min(x2$PRTIME)
+xmax <-max(x2$PRTIME)
+nbins <- ceiling((xmax-xmin) / binsize)
+h2 = hist(x2$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+x2_pt <- (h2$mids-2000)/2
+y2_raw_pt <- 2
+y2_pt <- 1
+z2_pt <- h2$counts/numSamples
+z2_raw_pt <- h2$counts
+inc2 <- rep(2,length(h2$mids))
+ss_inc2 <- rep(numSamples,length(h2$mids)) 
+### 4 sec
+x4 = read.csv(file="4_sec.dat",head=TRUE,sep="\t")
+x4 <- subset(x4, x4$ITERNUM <= numSamples)
+xmin <-min(x4$PRTIME)
+xmax <-max(x4$PRTIME)
+nbins <- ceiling((xmax-xmin) / binsize)
+h4 = hist(x4$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+x4_pt <- (h4$mids-4000)/4
+y4_raw_pt <- 4
+y4_pt <- 2
+z4_pt <- h4$counts/numSamples
+z4_raw_pt <- h4$counts
+inc4 <- rep(4,length(h4$mids))
+ss_inc4 <- rep(numSamples,length(h4$mids)) 
+### 8 sec
[…]

### 2017-07-13 pt_characterization/rep_check/u_s_time/pt_hist.r

@@ -308,6 +308,7 @@ xlab='PT (ms)', ylab=expression('Frequency'))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
+numSamples <- 1000
 x = read.csv(file="6_sec.dat",head=TRUE,sep="\t")
 x <- subset(x, x$ITERNUM <= numSamples)
 setEPS()
@@ -319,7 +320,7 @@ xmax <-max(x$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
 ymax <- max(h$counts)
-ymax <- 250
+ymax <- ceiling(ymax/100)*100
 plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green",  main='PT frequency on INC6', 
 sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
 xlab='PT (ms)', ylab=expression('Frequency'))
@@ -337,7 +338,25 @@ xmax <-max(x$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
 ymax <- max(h$counts)
-ymax <- 250
+ymax <- ceiling(ymax/100)*100
+plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green",  main='PT frequency on INC12', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+x = read.csv(file="12_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("12_sec_pt_hist.eps")
+binsize=1
+#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
[…]

### 2017-07-14 pt_characterization/rep_check/u_s_time/3d_plot.r
July 19 Draft

@@ -199,6 +199,16 @@ ss_inc4096 <- rep(numSamples,length(h4096$mids))
 numSamples <- 260
 x8192 = read.csv(file="8192_sec.dat",head=TRUE,sep="\t")
 x8192 <- subset(x8192, x8192$ITERNUM <= numSamples)
+#x8192_up = mean(x8192$PRTIME) + 2*sd(x8192$PRTIME)
+#x8192_dn = mean(x8192$PRTIME) - 2*sd(x8192$PRTIME)
+#x8192 = subset(x8192, x8192$PRTIME >= x8192_dn & x8192$PRTIME <= x8192_up)
+#sd(x8192$PRTIME)
+#sd(x8192$PRTIME)/mean(x8192$PRTIME)
+#x8192_up = mean(x8192$METIME) + 2*sd(x8192$METIME)
+#x8192_dn = mean(x8192$METIME) - 2*sd(x8192$METIME)
+#x8192 = subset(x8192, x8192$METIME >= x8192_dn & x8192$METIME <= x8192_up)
+#sd(x8192$METIME)
+#sd(x8192$METIME)/mean(x8192$METIME)
 xmin <-min(x8192$PRTIME)
 xmax <-max(x8192$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
@@ -216,7 +226,14 @@ x16384 = read.csv(file="16384_sec.dat",head=TRUE,sep="\t")
 x16384 <- subset(x16384, x16384$ITERNUM <= numSamples)
 x16384_up = mean(x16384$PRTIME) + 2*sd(x16384$PRTIME)
 x16384_dn = mean(x16384$PRTIME) - 2*sd(x16384$PRTIME)
-#x16384 = subset(x16384, x16384$PRTIME >= x16384_dn & x16384$PRTIME <= x16384_up)
+x16384 = subset(x16384, x16384$PRTIME >= x16384_dn & x16384$PRTIME <= x16384_up)
+sd(x16384$PRTIME)
+sd(x16384$PRTIME)/mean(x16384$PRTIME)
+x16384_up = mean(x16384$METIME) + 2*sd(x16384$METIME)
+x16384_dn = mean(x16384$METIME) - 2*sd(x16384$METIME)
+x16384 = subset(x16384, x16384$METIME >= x16384_dn & x16384$METIME <= x16384_up)
+sd(x16384$METIME)
[…]

### 2017-07-16 pt_characterization/rep_check/1024_run.tex
Refactoring files for Characterizing paper

@@
 \section{Histograms on Consecutive INC1024 Runs~\label{sec:s9_1024_runs}} 
-This section exhibits histograms on (three) consecutive runs of INC1024 via SEDONA. 
+This section exhibits histograms on (three) consecutive runs of INC1024. 
 The detailed description of the base data is from Table~\ref{tab:exp_notes2}.

### 2017-07-16 pt_characterization/rep_check/paper.tex
Change the name of the protocol
DRAFT for May 31
Draft being ready for July 10
Refactoring files for Characterizing paper

@@
 \section{Description}
-This document presents two sets of histograms based on the two runs 
-of a program under test, called {\em INC}, with a set of increasing task lengths 
-from 1 second to 4096 seconds.
-We would like to see if both of the sets have the same shape for the same task length.
-If so then, we can say that run repeatability is satisfied in our experiment setting. 
-%In this study the used protocol is equivalent to SEDONA~\cite{Sedona}, 
-%eliminating samples involving infrequent, long-running daemons, which is the first step of EMPv5.
-In this study the used protocol eliminates samples involving infrequent, long-running daemons, 
-which is the first step of EMPv5.
+This document characterizes execution times 
+measured on a simple program in pure-computation mode, called {\em INC}, 
+with increasing task lengths (up to 16,384 seconds from 1 second). 
+For the characterization, we present various histograms of INC throughout this document.
+The main goal of examining these histograms is to make sure if 
+two (or more) histograms of INC with the same task length 
+have the same shape or not. If that's the case, then we can say that such an INC run is so called {\em repeatable}; 
+in other words, {\em repeatability} is satisfied in our experimental settings. 
+Another goal is to uncover and explicate several interesting structures behind the histograms. 
+The third goal is to build a statistical distribution (or model) fitting in the histograms, so that later we 
+are capable of predicting a concrete execution time via 
+the model on an arbitrary algorithm with a given input on a real execution environment.
 
+In our experiments, we used EMPv5~\cite{EMP}. 
+(That said, the second step of EMPv5 was on purpose omitted, just to obtain better histograms by retaining more samples.) 
 In the protocol, we use taskstats C struct to get measures of a captured process. 
@@
 
-Now we show histograms of elapsed time (ET) and process time (PT) of INC via the SEDONA protocol.
+Now we show histograms of elapsed time (ET) and process time (PT) of INC via the EMPv5 protocol.
 
@@
 
+%\pagebreak
+%\newpage
+%
+%\input{further.tex}
+
 \pagebreak
@@
 \vspace{0.1em}
-\bibitem
+%\bibitem
 %{Avnur:2000:ECA:342009.335420}
@@
 9, pp.~??--??, 2017.
-\vspace{0.1em}
+%\vspace{0.1em}
+\bibitem
+%{Avnur:2000:ECA:342009.335420}
+{EMP}
+Young-Kyoon Suh, Richard T.
+Snodgrass, John Kececioglu, Peter J.
+Downey, Rob S.
+Maier, and Cheng Yi, ``EMP: Execution Time Measurement Protocol for Compute-Bound Programs'', in {\em Software: Practice and Experience}, 47(4):559--597, 2017.
 \bibitem

### 2017-07-16 pt_characterization/rep_check/res1.tex
Refactoring files for Characterizing paper

--- 2017-04-12 pt_characterization/rep_check/res1.tex
@@
 This section exhibits histograms on the first run of 
-INC with its task length increasing from 1 second to 4096 seconds, via EMPv5. 
+INC with its task length increasing from 1 second to 4096 seconds. 
 The detailed description of the base data is from Table~\ref{tab:exp_notes1}.

### 2017-07-16 pt_characterization/rep_check/temp/further.tex
Refactoring files for Characterizing paper

--- 2017-07-07 pt_characterization/rep_check/res3.tex
@@
 INC with its task length increasing from 1 second to 4096 seconds via EMPv5 with no Step 2. 
-The detailed description of the base data is from Table~\ref{tab:exp_notes3}.
-
-%\begin{table}[h]
-%\begin{center}
-%\begin{tabular}{|p{2cm}|p{3cm}|p{3cm}|p{4cm}|p{3.5cm}|} \hline
-%Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
-%{\tt sodb9} &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-13 $\sim$ 2017-03-14 & %Figs.~\ref{fig:s9_r2_et_hist1},~\ref{fig:s9_r2_et_hist2},~\ref{fig:s9_r2_pt_hist1}, and~\ref{fig:s9_r2_pt_hist2}\\ \hline
-%{\tt sodb9} &  INC512$\sim$INC1024 & 300 samples, each & 2017-03-17 $\sim$ 2017-03-21 & 
-%%Figs.~\ref{fig:s9_r2_et_hist3} and~\ref{fig:s9_r2_pt_hist3}\\ \hline
-%{\tt sodb10} & INC2048 & 300 samples & 2017-03-13 $\sim$ 2017-03-20 & 
-%%Figs.~\ref{fig:inc2048_r2_et_hist_v5} and~\ref{fig:inc2048_r2_hist_v5}\\ \hline
-%{\tt sodb12} & INC4096 & 300 samples & 2017-03-02 $\sim$ 2017-03-17 & 
-%%Figs.~\ref{fig:inc4096_r2_et_hist_v5} and~\ref{fig:inc4096_r2_hist_v5}\\ \hline
-%{\tt sodb10} & INC8192& 261 samples & 2017-04-27 $\sim$ 2017-05-21 & 
-%%Figs.~\ref{fig:inc8192_r2_et_hist_v5} and~\ref{fig:inc8192_r2_hist_v5}\\ \hline
-%{\tt sodb12} & INC16384& 130 samples & 2017-04-27 $\sim$ 2017-05-21 & 
-%%Figs.~\ref{fig:inc16384_r2_et_hist_v5} and~\ref{fig:inc16384_r2_hist_v5}\\ \hline
-%\end{tabular}
-%\end{center}
-%\vspace{-.2in}
-%\caption{Notes on experiment runs used for histograms\label{tab:exp_notes3}}
-%\end{table}
-%
-%INC8192/INC16384 unfortunately stopped in the middle of their runs due to a frozen vnc problem. So couldn't finish 300 samples.
-%
-%\pagebreak
+The detailed description of the base data is from Table~\ref{tab:exp_notes1}.
 
@@
 \newpage
-
-\subsection{ET}
-
-\begin{figure}[hp!]
-	\centering
-	\subfigure[ET frequency on INC1]{
-		\includegraphics[scale=0.43]{u_s_time/1_sec_et_hist.eps}
-		\label{fig:inc1_r3_et_hist_v5}
-	}
-	\subfigure[ET frequency on INC2]{
-		\includegraphics[scale=0.43]{u_s_time/2_sec_et_hist.eps}
-		\label{fig:inc2_r3_et_hist_v5}
-	}
-	\subfigure[ET frequency on INC4]{
-		\includegraphics[scale=0.43]{u_s_time/4_sec_et_hist.eps}
-		\label{fig:inc4_r3_et_hist_v5}
-	}
-	\subfigure[ET frequency on INC8]{
-		\includegraphics[scale=0.43]{u_s_time/8_sec_et_hist.eps}
-		\label{fig:inc8_r3_et_hist_v5}
-	}
-	\caption{ET Histograms of INC1 ... INC8~\label{fig:s9_r3_et_hist1}}
-\end{figure}
-
-\begin{figure}[hp!]
-	\centering
-	\subfigure[ET frequency on INC16]{
-		\includegraphics[scale=0.43]{u_s_time/16_sec_et_hist.eps}
-		\label{fig:inc16_r3_et_hist_v5}
-	}
-	\subfigure[ET frequency on INC32]{
-		\includegraphics[scale=0.43]{u_s_time/32_sec_et_hist.eps}
-		\label{fig:inc32_r3_et_hist_v5}
-	}
-	\subfigure[ET frequency on INC64]{
-		\includegraphics[scale=0.43]{u_s_time/64_sec_et_hist.eps}
-		\label{fig:inc64_r3_et_hist_v5}
-	}
-	\caption{ET Histograms of INC16 ... INC64~\label{fig:s9_r3_et_hist2}}
-\end{figure}
-
-\begin{figure}[hp!]
-	\centering
-	\subfigure[ET frequency on INC128]{
-		\includegraphics[scale=0.43]{u_s_time/128_sec_et_hist.eps}
-		\label{fig:inc128_r3_et_hist_v5}
-	}
-	\subfigure[ET frequency on INC256]{
-		\includegraphics[scale=0.43]{u_s_time/256_sec_et_hist.eps}
-		\label{fig:inc256_r3_et_hist_v5}
-	}
-	\subfigure[ET frequency on INC512]{
-		\includegraphics[scale=0.43]{u_s_time/512_sec_et_hist.eps}
-		\label{fig:inc512_r3_et_hist_v5}
-	}
-	\subfigure[ET frequency on INC1024]{
-		\includegraphics[scale=0.43]{u_s_time/1024_sec_et_hist.eps}
-		\label{fig:inc1024_r3_et_hist_v5}
-	}
-	\caption{ET Histograms of INC128 ... INC1024~\label{fig:s9_r3_et_hist3}}
-\end{figure}
-
-%\begin{figure}[hp!]
-%	\centering
-%	\subfigure[ET frequency on INC2048]{
-%		\includegraphics[scale=0.43]{u_s_time/2048_sec_et_hist.eps}
-%		\label{fig:inc2048_r3_et_hist_v5}
-%	}
-%	\subfigure[ET frequency on INC4096]{
-%		\includegraphics[scale=0.43]{u_s_time/4096_sec_et_hist.eps}
-%		\label{fig:inc4096_r3_et_hist_v5}
-%	}
-%	\caption{ET Histograms of INC2048 ... INC4096~\label{fig:s9_r3_et_hist4}}
-%\end{figure}
-
-\vspace\fill
-\clearpage
-
-\subsection{PT}
-
-\begin{figure}[hp!]
-	\centering
-	\subfigure[PT frequency on INC1]{
-		\includegraphics[scale=0.43]{u_s_time/1_sec_pt_hist.eps}
-		\label{fig:inc1_r3_hist_v5}
-	}
-	\subfigure[PT frequency on INC2]{
-		\includegraphics[scale=0.43]{u_s_time/2_sec_pt_hist.eps}
-		\label{fig:inc2_r3_hist_v5}
-	}
-	\subfigure[PT frequency on INC4]{
-		\includegraphics[scale=0.43]{u_s_time/4_sec_pt_hist.eps}
-		\label{fig:inc4_r3_hist_v5}
-	}
-	\subfigure[PT frequency on INC8]{
-		\includegraphics[scale=0.43]{u_s_time/8_sec_pt_hist.eps}
-		\label{fig:inc8_r3_hist_v5}
-	}
-	\caption{PT Histograms of INC1 ... INC8~\label{fig:s9_r3_pt_hist1}}
-\end{figure}
-
-\begin{figure}[hp!]
-	\centering
-	\subfigure[PT frequency on INC16]{
-		\includegraphics[scale=0.43]{u_s_time/16_sec_pt_hist.eps}
-		\label{fig:inc16_r3_hist_v5}
-	}
-	\subfigure[PT frequency on INC32]{
-		\includegraphics[scale=0.43]{u_s_time/32_sec_pt_hist.eps}
-		\label{fig:inc32_r3_hist_v5}
-	}
-	\subfigure[PT frequency on INC64]{
-		\includegraphics[scale=0.43]{u_s_time/64_sec_pt_hist.eps}
-		\label{fig:inc64_r3_hist_v5}
-	}
-	\caption{PT Histograms of INC16 ... INC64\label{fig:s9_r3_pt_hist2}}
-\end{figure}
-
-\begin{figure}[hp!]
-	\centering
-	\subfigure[PT frequency on INC128]{
-		\includegraphics[scale=0.43]{u_s_time/128_sec_pt_hist.eps}
-		\label{fig:inc128_r3_hist_v5}
-	}
-	\subfigure[PT frequency on INC256]{
-		\includegraphics[scale=0.43]{u_s_time/256_sec_pt_hist.eps}
-		\label{fig:inc256_r3_hist_v5}
-	}
-	\subfigure[PT frequency on INC512]{
-		\includegraphics[scale=0.43]{u_s_time/512_sec_pt_hist.eps}
-		\label{fig:inc512_r3_hist_v5}
-	}
-	\subfigure[PT frequency on INC1024]{
-		\includegraphics[scale=0.43]{u_s_time/1024_sec_pt_hist.eps}
-		\label{fig:inc1024_r3_hist_v5}
-	}
-	\caption{PT Histograms of INC256 ... INC1024~\label{fig:s9_r3_pt_hist3}}
-\end{figure}
-
-%\begin{figure}[t]
-%	\centering
-%	\subfigure[PT frequency on INC2048]{
-%		\includegraphics[scale=0.43]{u_s_time/2048_sec_pt_hist.eps}
-%		\label{fig:inc2048_r3_hist_v5}
-%	}
-%	\subfigure[PT frequency on INC4096]{
-%		\includegraphics[scale=0.43]{u_s_time/4096_sec_pt_hist.eps}
-%		\label{fig:inc4096_r3_hist_v5}
-%	}
-%%	\subfigure[PT frequency on INC8192]{
-%%		\includegraphics[scale=0.43]{u_s_time/8192_sec_pt_hist.eps}
-%%		\label{fig:inc8192_r3_hist_v5}
-%%	}
-%%	\subfigure[PT frequency on INC16384]{
-%%		\includegraphics[scale=0.43]{u_s_time/16384_sec_pt_hist.eps}
-%%		\label{fig:inc16384_r3_hist_v5}
-%%	}
-%	\caption{PT Histograms of INC2048 ... INC16384~\label{fig:s9_r3_pt_hist4}}
-%\end{figure}
-
-\clearpage

### 2017-07-18 mlppi-ksii17/cover-letter.tex
MLPPI TIIS draft

--- 2017-03-27 mlppi-comsis17/cover-letter.tex
@@
 \documentclass{letter} % Uses 10pt
+\usepackage{color}
 %Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
@@
 \begin{flushright}
-\hfill Supercomputing R\&D Center, 
+\hfill Center of Computational Science and Engineering, 
 \end{flushright}
@@
 
-\opening{Dear Prof. Mirjana Ivanovi\'c (Editor-in-Chief):}
+\opening{Dear Dr. Minho Jo (Editor-in-Chief):}
 
-\noindent An earlier version of this submission was published in Proceedings of the 21st ACM \hbox{International} Conference on Information and Knowledge Management (CIKM). 
-This version that I now submit to Computer Science and Information Systems (ComSIS) includes the following substantial extensions.
+\noindent An earlier version of this submission appeared as a short paper in the Proceedings of the 21st ACM \hbox{International} Conference on Information and Knowledge Management (CIKM). 
+This version that I now submit to KSII Transactions on Internet and Information Systems (TIIS) includes the following substantial extensions that are definitely greater than 30\%.
 
@@
 
-\item Expanded algorithms: we substantially expand the description of the pre-processing, initial and optimized phase algorithms shortened in the earlier version.
-Furthermore, this submission presents three additional algorithms as follows:
+\item Expanded algorithms: we present a substantial expansion of the pre-processing, initial and optimized phase algorithms shortened in the earlier version. Moreover, this submission includes 
+several additional algorithms in the following:
 
@@
 
-\item Splitting overlapping ranges in Algorithm~2 on page~8,
-\item Checking the allowed field count limit on the split ranges on page~8,
-\item Computing scan cost in Algorithm~3 on page~10, and
-\item Computing query cost in Algorithm~5 on page~13.
+\item Splitting overlapping ranges in {\color{blue}Alg.~2} on page~8,
+\item Checking the pre-defined field count limit on the split ranges on page~8,
+\item Computing scan cost in {\color{blue}Alg.~3} on page~10, and
+\item Computing query cost in {\color{blue}Alg.~5} on page~13.
 
@@
 
-\item Complexity analysis: we conduct the asymptotic running time analysis on each phase in Section~3 on pages~14-15.
+\item Complexity analysis: we conduct the asymptotic running time analysis on the proposed algorithms in Section~3 on pages~14--15.
 
-\item Experiments: we provide additional runtime statistics measured from our experiments, as exhibited in Table~1 on page~16. 
-In addition, we compare the recommendation quality of our tool with that of a human expert, as shown in Figure~5 on page~18.
+\item Experiments: we provide additional runtime statistics measured from our experiments, as exhibited in {\color{blue}Table~1} on page~16.
+In addition, we compare the recommendation quality of our tool with those of a human expert and a vanilla approach (no partitioning), as shown in {\color{blue}Fig.~5} on page~17.
 
-\item More thorough literature review: Section~5 on page~18 discusses in greater detail the similarities and differences between existing tools 
-and ours.
+\item More thorough literature review: Section~5 on page~18 discusses in greater detail the similarities and differences between existing works and our work.
 

### 2017-07-21 mlppi-ksii17/cover-letter.tex
Submission

@@
 \item Experiments: we provide additional runtime statistics measured from our experiments, as exhibited in {\color{blue}Table~1} on page~16.
-In addition, we compare the recommendation quality of our tool with those of a human expert and a vanilla approach (no partitioning), as shown in {\color{blue}Fig.~5} on page~17.
+In addition, we compare the recommendation quality of our tool with those of a human expert and a vanilla approach (no partitioning), as shown in {\color{blue}Fig.~5~(b)} on page~17.
 

### 2017-07-26 subopt-tods16/cover-letter.tex
Cover letter added

% Cover letter using letter.sty
\documentclass{letter} % Uses 10pt
%Use \documentstyle[newcent]{letter} for New Century Schoolbook postscript font
% the following commands control the margins:
\topmargin=-1in    % Make letterhead start about 1 inch from top of page 
\textheight=8in  % text height can be bigger for a longer letter
\oddsidemargin=0pt % leftmargin is 1 inch
\textwidth=6.5in   % textwidth of 6.5in leaves 1 inch for right margin

\begin{document}

\signature{Young-Kyoon Suh}           % name for signature 
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

\begin{letter}{
%Editors-in-Chief of ACM Transactions on Computer Systems (TOCS): \\
%\begin{tabular}{ll}
%\vspace{.1in}
%
%Prof. Todd C. Mowry\\
%Computer Science Department\\
%Carnegie Mellon University\\
%5000 Forbes Avenue\\
%Pittsburgh, PA 15213 \\
%USA\\ 
}

\begin{flushleft}
{\large\bf Young-Kyoon Suh}
\end{flushleft}
\medskip\hrule height 1pt
%\begin{flushright}
%\hfill Department of Computer Science, The University of Arizona, Tucson, AZ 85721 USA\\
%\hfill Tel: +1-520-370-1865
%(Please understand we cannot reveal our affiliation, as the manuscript should be double-blind.)
%\end{flushright} 
\begin{flushright}
\hfill National Institute of Supercomputing and Networking, 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Korea Institute of Science and Technology Information (KISTI), 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Republic of Korea, 34141, 
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Tel: +82-42-869-0725
\end{flushright}
\vspace{-.15in}
\begin{flushright}
\hfill Email: yksuh@kisti.re.kr
\end{flushright} 
%\vfill % forces letterhead to top of page

%\vspace{-0.5in}

\opening{Dear Prof. Christian S. Jensen:}

We are submitting this manuscript, entitled ``A Causal Model of DBMS Suboptimality,'' to ACM {\em \hbox{Transactions} on Database Systems} (TODS).

In this cover letter we disclose the following information.

\begin{itemize}

\item No version of this manuscript has been previously published nor be currently under consideration for publication elsewhere. 
We have anonymized the current submission as directed by the double-blind
review guidelines on the TODS web site. Note that Section 9 (on pages 36--37) of
this manuscript summarizes the contributions in an attempt to draw 
a causal explanation of suboptimality and to uncover a fundamental limit 
on the number of operators that a DBMS may be able to support.

\item The authors of this manuscript are: Richard T.~Snodgrass (Department of
  Computer Science, University of Arizona), Sabah Currim (Alumni
  Association, University of Arizona), and Young-Kyoon Suh (KISTI). 
\item No anonymous citations appear in this manuscript.
\item Conflict of interests associated with the listed five categories:
  Richard T.~Snodgrass's and Young-Kyoon Suh's have a number of people to list. (See
  below for details.) For the other author, her collaborators are all at
  the University of Arizona.
\end{itemize}

Below is Richard T. Snodgrass's list of people in conflicts of interests.

Luca Anselma (University of Piedmonte, Italy), Michael B\"{o}hlen
(University of Zurich), Alessio Bottrighi (University of Piedmonte, Italy),
Glenn Browne (Texas Tech), James Cordy (Queen's University), Faiz \hbox{Currim} (University of Arizona), Sabah
Currim (University of Arizona), Curtis Dyreson (Utah State University),
Dengfeng Gao (IBM), Johann Gamper (Free University of Bozen-Bolzano), Christian S. Jensen (Aalborg
University), Shailesh Joshi (Expedia), Vijay Khatri (Indiana University),
Soumyadeb Mitra (Data Domain), Gianpaolo Molino (Torino, Italy), Bongki Moon (Seoul National University), Clayton
Morrison (University of Arizona), Jeffrey Parsons (Memorial University of
Newfoundland), Kyriacos Pavlou (University of Illinois), Sudha Ram
(University of Arizona), Young-Kyoon Suh (KISTI), Paolo Terenziani (University of Piedmonte, Italy),
Stephen Thomas (Queen's University), Bedirhan Urgun (Turkcell), Marianne
Winslett (University of Illinois), and Rui Zhang (Dataware Ventures).\\

Michael B\"{o}hlen
(University of Zurich) was a postdoctoral student. Richard Snodgrass has
advised the following doctoral students, Ilsoo Ahn (Samsung), Curtis
E.~Dyreson (Utah State University), Dengfeng Gao (IBM), Daniel Hoffman
(University of Victoria), Nick Kline (Google), Edwin E.~McKenzie, Kyri
Pavlou (University of Illinois), Karen P.~Shannon, Michael D.~Soo (Reliant
Steel), Rui Zhang (Dataware), and Young-Kyoon Suh (KISTI), along with eleven MS thesis students.\\

Richard Snodgrass's dissertation committee was chaired by William Wulf and
included Arthur Evans, Anita Jones, Joseph Newcomer, and Zary Segall.

Below is Young-Kyoon Suh' list of people in conflicts of interests.

Alain Crolotte (Teradata), 
Sabah Currim (University of Arizona), 
Ahmad Ghazal (Oracle),
John \hbox{Kececioglu} (University of Arizona), 
Jong Wook Kim (Sangmyung University),
Myoung Ho Kim (KAIST),
Pekka Kostamaa (Teradata),
Ki Yong Lee (Sookmyung Women's University),
Sang-Won Lee (Sungkyunkwan University),
Rob Maier (University of Arizona), 
Bongki Moon (Seoul National University),
Kyriacos Pavlou (University of Illinois),
Sudha Ram (University of Arizona),  
Richard Snodgrass (University of Arizona), 
Cheng Yi (Google), and 
Rui Zhang (Dataware Ventures).\\

Young-Kyoon Suh's dissertation committee was chaired by Richard T. Snodgrass and
included John \hbox{Kececioglu}, Peter Downey, and Sabah Currim.

\closing{Sincerely yours,}

\end{letter}

\end{document}

### 2017-07-30 pt_characterization/rep_check/u_s_time.tex
July 19 Draft
PT draft

@@
-\section{Further Investigation of the First Run and Some Additional Runs~\label{sec:u_s_time_hist}} 
-This section exhibits user and system time histograms on the second run of 
-INC with its task length increasing from 1 second to 4096 seconds via EMPv5 with no Step 2.
+\section{Further Investigation of the Existing Runs and Some Additional Runs~\label{sec:u_s_time_hist}} 
+This section exhibits user and system time histograms on the first run of 
+INC with its task length increasing from 1 second to 1024 seconds 
+and the second run of INC from 2048 seconds to 16384 seconds via EMPv5 with no Step 2.
 The detailed description of the base data is from Table~\ref{tab:exp_notes1}, 
@@
 	}
-	\subfigure[Coefficient of variation (relative)]{
+	\subfigure[Standard deviation (absolute) in log scale]{
+		\includegraphics[scale=0.6]{u_s_time/overall_std_log.eps}
+		\label{fig:overall_std_log}
+	}
+	\subfigure[Coefficient of variation in log scale (relative)]{
 		\includegraphics[scale=0.6]{u_s_time/overall_re.eps}

### 2017-07-30 pt_characterization/rep_check/repet_data2/pt_hist_sodb9.r

@@ -118,7 +118,7 @@ axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
 x = read.csv(file="32_sec.dat",head=TRUE,sep="\t")
-x <- subset(x, x$ITERNUM <= numSamples)
+#x <- subset(x, x$ITERNUM <= numSamples)
 setEPS()
 postscript("32_sec_pt_hist_v5.eps")
 #x <- subset(x, x$ITERNUM != 433)

### 2017-08-08 subopt-tods17/OLDtex/cover.tex
subopt tods submission

\documentclass{sig-alternate}
%\documentclass{vldb}
% \usepackage{fullpage}
% \usepackage{times}
\usepackage{url}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{epsfig}
% \usepackage[total={6.5in,9in}, top=1.25in,left=0.9in]{geometry}
%\usepackage[text={6.5in,9in},centerpage,includefoot]{geometry}
% \usepackage{setspace}
\usepackage{color}
\usepackage{url}
\usepackage{balance}

\definecolor{grey}{RGB}{200,200,200}
\newcommand{\hilite}[1]{\colorbox{grey}{#1}}
\newcommand{\hilitey}[1]{\colorbox{yellow}{#1}}
\newcommand{\hiliting}[1]{\colorbox{grey}{#1}}
\long\def\todo#1{\hilitey{{\bf TODO:} {\em #1}}}
\long\def\shorten#1{}

%For last edit: revise and resubmit for SIGMOD'14
\long\def\shortenF#1#2{}%First is letter indicating what has been removed

%For double blind, keep the second argument. To turn off doubleblind, keep
%the first
\long\def\double#1#2{#2}
\def\azdb{\double{\hbox{\sc AZDBLab}}{\hbox{\sc DBLab}}}
\def\QatC{Q{@}C}
\long\def\comment#1{}
%c2j: Conference to Journal: first parameter is conference, second is journal
% For Conferences: c2j#1#2{#1}
% For Journals: c2j#1#2{2}
\long\def\c2j#1#2{#1}

\begin{document}

{\centering\Large\bf Comments to the Reviewers}

We thank the reviewers for their very clear statement of the concerns they
had, enabling us to work over the last few weeks to articulate better what is
surprising and what is actionable.

\vspace{2ex}\noindent{\bf Meta-review}

\begin{itemize}
\item ``the reviewers felt that the derived conclusions are (1) non
  surprising, i.e., Well known, and (2) non actionable, i.e., Nothing one
  can do about it.''

The last three pages (Sections 7--9) have been substantially rewritten to list explicitly the
surprising results and to indicate the specific things that can (as well as
one that need not) be done, deriving from the results of this research. To
make space for this longer discussion, we removed about 1/2 page of
methodological details that are less important.

We know of no other paper that has presented those results or has been able
to provide the rather specific guidance across DBMSes to database
researchers and developers of places to improve and places where the problem
is not present.

\item ``Furthermore, in the discussion the relationship to Haritsas work was
  pointed out.''
We now explain in Section 7 how our results go beyond Haritsa's novel work. (We
mention here but not in the paper that Haritsa et al. did not look at all at
query suboptimality.)
\end{itemize}

In the following, we respond to just the high-level concerns expressed in
the detailed reviews, though emphasize that we have made small changes
throughout the paper in response to this feedback.

\vspace{2ex}\noindent{\bf Reviewer 1}
\begin{itemize}
\item ``...the results don't specify how the results can be used to improve
  the space of solutions in query optimization.''

Our paper is fundamentally about articulating and testing a causal model
that will thus provide novel guidance to us and to the rest of the community
about how to improve query optimization. This model identifies more
precisely than before where more work could substantially add to the space
of solutions in query optimization.

\item `` There isn't any discussion on
  how the results be integrated to improve the present DBMSes or tune the
  schema or queries better.''

We now discuss in some detail which research questions need more attention
and which do not. We feel that the ultimate goal is to have the DBMS handle
query optimization, and so do not consider manual schema or query tuning.

\item "complex sub-queries ... a discussion on how the model might extend to
  such cases is missing.''

There are several paragraphs in Section 9, modified and expanded, which discuss extension
of the model.

\end{itemize}

\newpage%\vspace{2ex}
\noindent{\bf Reviewer 2}
\begin{itemize}
\item ``There is nothing surprising in the findings and although the
  methodology is interesting, it does not seem to lead to any new insight.''

See the discussion above for the changes we have made.

\item ``It is a bit annoying to read these conclusions without any reference
  to the literature where such problems are addressed.''

We have added some additional references.

What our research does is help focus attention on more specifically where
the problems still reside in query optimization. For example, even when the
statistics are entirely accurate, existing query optimizers have challenges
even on simple queries, a finding that this paper contributes.

\item ``since no attempt is made to put the results in perspective over what
  has been done in the past''

We have not made such an attempt because we don't know (can't know) which
research results from the past have been adequately incorporated  into
existing proprietary DBMSes. Our audience comprises both researchers, who
can develop new solutions, and developers, who can incorporate existing and
new solutions into subsequent versions of their products.

\item ``this could be as simple as mentioning how these problems have been
  addressed in the past and saying that the paper now qualitatively confirms
  what was informally known before.''

This is an excellent suggestion. We mention that in the fourth paragraph of Section
4 and the third bullet of Section 8. We also repeatedly discuss where {\em
  more} research is needed, so as to not imply that research hasn't been
done in the past.

\item ``Put the result in context to what is known already and state if
  there are any conclusions that are different from the state-of-the-art in
  terms of query plans and optimizations.''

We have tried to be as clear and directed as we can, given the constraints
mentioned above.
\end{itemize}

\vspace{2ex}\noindent{\bf Reviewer 3}
\begin{itemize}
\item ``W1: It is not clear what is actionable about the hypotheses.''

See above about what we have added.

\item ``Sure, number of ``effective plans'', number of joins, etc will
  result in more complex plans, and more points where things are suboptimal,
  but most of these are not factors that a query optimizer can control.''\

No disagreement here. These factors do have implications on challenges that
query optimizers much contend with.

\newpage
\item ``In direct contrast, reasons such as errors in statistics,
  independence assumptions, which can be ameliorated by better/more
  expensive statistics are not covered at all.''

We discussed that in our investigation, the statistics are accurate and the
independence assumptions are also accurate (there is not skew in our data),
and so those are not sources for the query suboptimality observed in our
study. Sections 6.7 and 7 discuss the challenges to be addressed and possible
engineering approaches that might help.

\item ``W2: $\ldots$''

This paragraph was removed.

\item ``W3: The paper oversells in parts, eg the claim about manual
  optimization of queries in the introduction is meaningless since most
  queries are canned queries which are executed repeatedly.''

In our experience, canned queries are often those that are most problematic
and require the most manual tuning. In any case, after clarifying that
paragraph, we removed it due to space constraints.

\item ``W4: The paper only looks at one manifestation of suboptimality,
  which respect to plans that are optimal for nearby regions, i.e., only
  checking if the cutover point is chosen properly.''

Actually, we are not considering the specific case that the cutover point is
chosen properly or improperly. Rather, change points are used in our 
operationalization of suboptimality, especially for proprietary systems. So
change points are a fundamental notion in our methodology.

But the reviewer is correct that our definition of suboptimality is
conservative, in that we might miss some suboptimal plans. We mention that
near the end of Section 5.3.

\item ``W5: the experiments are not on any standard benchmark.''

We don't see how that invalidates any of the findings or conclusions we
report.

\item ``Why have you not explained what are the queries you used.''

The queries are generated randomly using the process detailed in Section 5.2.
\end{itemize}

\vspace{2ex}\noindent{\bf Postscript}

Our sincere thanks for the clear feedback, which has substantially benefited
the paper.
\end{document}

% LocalWords:  PVLDB

### 2017-08-08 subopt-tods17/OLDtex/dataanalysis.tex
subopt tods submission

\section{Data Analysis}
\subsection{Testing the Causal Model}
\todo{Give algorithm for the onpass scenario.}

\c2j{}{\subsection{Experimental Setup}\label{sec:setup}}
As mentioned, we used \azdb\ to run the experiments. This infrastructure,
now comprising about 45K lines of code, handles all the details of
generating the sample data, generating the actual queries, collecting the
various plans generated by the optimizer, running the queries, and storing
the results in a ``lab notebook,'' a set of tables managed by
a separate DBMS. This infrastructure also performs the many steps
required to ensure repeatability.

We have assembled a hardware lab of dedicated
machines, one \c2j{for each DBMS}{each for the three DBMSes} and one to run the DBMS used
to store the lab notebooks. Having dedicated hardware allows us to worry
less about other processes running on the machines that could dirty the
results (these machines are {\em only} for experiments), and allows us to
run extensive experiments involving days or weeks of computation. We have
used this software and hardware lab to perform a number of experiments
examining the phenomenon of suboptimality.

\azdb\ utilizes experiments defined in XML. It supports {\em experimental
  scenarios}, which are short (perhaps 100 lines of code) Java subclasses
that provide details of how to run kinds of experiments.

In the following, we describe in detail the data used by our experiments and
the experimental \c2j{scenario we utilized.}{scenarios we defined.}

\subsection{Datasets}\label{sec:datasets}
We generate our experiment datasets randomly in the experiments. However, we
use seeds to control the random data generator so that it can produce
repeatable data as required.

Our experiment datasets consist of relational tables. There are two types of
tables. The first is a ``fixed table'' that, once created and
populated, will never be modified in the future.  In contrast, the second
type is a ``variable table.'' We alter the cardinality\c2j{ statistics of
  this table}{, either physically or
virtually, of such tables} as the experiments are being performed. We will
elaborate the detailed discussion on modifying table cardinality
\c2j{statistics }{}when we
examine our experiment \c2j{scenario}{scenarios}.

\c2j{We used a}{We organized three configurations for the datasets. The first configuration
is a small} dataset with four tables, each with four integer-typed
columns. Each table is populated with one million rows. The size of each
table is roughly 16MByte ( =4 (bytes)$\times 4 \times$1,000,000).\c2j{}{ Second, to
perform queries with more relations, we create eight tables. In this case,
each table has the same configuration as in the first dataset. Third, to
test queries on large tables, we configure each table with 64 integer-typed
columns. In this dataset, we create four tables, but the size of each one is
512MByte.}

\shorten{\begin{figure*}[t]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} generateQueries({\em numQueries}, %\\
%\quad
{\em maxNumCorrNames}, {\em maxNumColumns}, {\em whereAttributes}, {\em tableMetaData}):
\FOR{$i$ $\leftarrow$ 1 {\bf to} {\em numQueries}}
\STATE    {\em numCorrNames}       $\leftarrow$ random(1, {\em maxNumCorrNames})
\STATE    {\em numColumns}      $\leftarrow$ random(1, {\em maxNumColumns})
\STATE    {\em corrNames}	$\leftarrow$ getCorrNames({\em numCorrNames})
	\FOR{$j$ $\leftarrow$ 1 {\bf to} {\em numCorrNames}}
	\STATE	{\em tableNames}[$j$]	$\leftarrow$ {\em tableMetaData}[random(1, %\\
%\hspace*{2in}
{\em tableMetaData}.numTables)].tableName
	\ENDFOR
	\FOR{$j$ $\leftarrow$ 1 {\bf to} {\em numColumns}}
	\STATE	{\em columnNames}[$j$]	$\leftarrow$ pickColumn({\em corrNames}, %\\
%\hspace*{1.5in} 
{\em tableMetaData})
	\ENDFOR
\STATE    {\em selectClause}   $\leftarrow$ generateSelectClause({\em columnNames})
\STATE    {\em fromClause}     $\leftarrow$ generateFromClause({\em tableNames}, %\\
%\hspace*{1.5in}
{\em corrNames})
\STATE    {\em whereClause}    $\leftarrow$ generateWhereClause({\em tableNames}, %\\
%\hspace*{1.5in}
{\em corrNames}, {\em whereAttributes})
\STATE    {\em queries}[$i$]      $\leftarrow$ {\em selectClause} + {\em fromClause} + {\em whereClause}
\ENDFOR
\STATE {\bf return} {\em queries}
\end{algorithmic}
\caption{Algorithm for Random Query Generation\label{alg:querygen}}
\end{center}
\end{figure*}

The XML specification of the \c2j{dataset}{first dataset (the others are similarly
specified)} is shown in Figure~\ref{fig:expspec}. Note that we do not specify any
constraints such as primary keys and foreign keys, nor did we specify
primary or secondary indexes. This will have the impact of eliminating
several operators which might be used in plan generation, such as index
scan. Interestingly, even though the data and query complexity are both
quite restricted (though we can vary query complexity \c2j{somewhat}{through
  our three datasets}), we still observed suboptimality\c2j{}{ in all three of our subject DBMSes}.

The experimental specification also configures the randomly generated
queries. The query specification stated that 500 queries were to be
generated\c2j{}{; this set of queries were generated for the first experimental
subject and then used with the other two experimental subjects}. Each query
is a simple select-project-join (SPJ) query, with a few attributes in the
SELECT clause, a few tables referenced in the FROM clause, and a few
equality predicates in the WHERE clause. As such some of the complexities
mentioned by Kabra and DeWitt~\cite{kabra98}, such as
user-defined data types, methods, and operators, are not considered.  The queries were generated
by a simple algorithm, the pseudo-code for which is shown in
Figure~\ref{alg:querygen}.

The query generator has three main components, which are {\tt
  generateSelectClause()}, {\tt generateFromClause()}, and {\tt
  generateWhereClause()}. The select clause will contain from one to four
columns\c2j{}{, hence, an average of 2.5 columns}. The number of correlation names
in the FROM clause varied from one to four, with duplication of tables
allowed (duplicate table names within a from clause implies a self-join).
In the queries that were generated, from one to four unique tables were
mentioned in the FROM clause. Somewhat fewer tables were mentioned than the
number of correlation names, as the presence of self-joins reduced the
number of unique tables named by the queries.

In {\tt generateWhereClause()}, in which all the joins will appear, we make
sure that Cartesian products are
eliminated\c2j{.}{(\verb.cartesianPossible="false".)}  We do this by
connecting all the correlation names that appear in the FROM statement via
comparisons on random columns. In this case, the comparisons are all
equality operators. To ensure that the queries are as simple as possible, we
do not include any additional predicates in the WHERE clause.  \c2j{}{This
  is realized by setting the attributes {\tt maxIsAbsolute} to {\tt true}
  and {\tt complexUsePercentage} to {\tt 100}.  Basically, ``complex''
  predicates eliminate the Cartesian product, and by setting complex
  predicates as ``absolute,'' no additional predicates are included except
  for those which are necessary for eliminating Cartesian product.} Also for
simplicity, we do not include disjunctions nor negations.

\c2j{}{The settings of the query parameters are related to the definition of the test
tables. For example, the experiment involving eight tables would expect the
queries to be generated also have the appearance of the eight tables (at most).
}
}

\todo{ Add discussion of exhaustive, onepass}

\subsection{Data Analysis}\label{sec:dataanalysis}
\todo{Include details of how many queries, plans were selected and how many runs across how many DBMSes to illustrate complexity. }
\todo{Discuss statistics in more detail. }
Through many weeks of testing, we were able to eliminate the caching and
buffering artifacts via five general steps.  First, we cleared the DBMS
buffer via the provided API, to discard cached content residing inside
the DBMS. \c2j{}{(In some cases, there were multiple caches to be cleared, each
with a different kind of command. As mentioned above, these caches were
often not clearly documented.) }Second, we flushed the OS-related
caches.\c2j{}{ Under our experimental system, which is Linux, we used the {\tt
  drop\_cache} facility to discard cached clean pages from
cache. Considering that in query plan generation, no dirty pages would
necessarily be involved, clearing the clean page is sufficient.} Third, we
filled the on-disk cache, which is resident on the hard-drive, by reading a
data file of size 64MB. The size of the data file was chosen such that it is
bigger than the hard-drive buffer, which is 32MB. Fourth, we disabled,
through a variety of means, over a
dozen daemons that were not needed on a dedicated server. The final step
disabled, again, in various ways, some of the daemons running within the DBMS.
Through all of these
efforts in concert, we were able to obtain quite stable and repeatable
measurements of the time each plan for each query executed. Skipping any
one of these steps adds significant variability to the timing of a query.

There was still some residual variability in the timing of the execution
plans for each query. We have evidence that this variability is due to DBMS
daemons that occasionally wake up and expend system resources for a second
or two, sometimes more. (While we were able to disable or stop some of the
DBMS daemons, others seemed to be required by the DBMS for correct execution or
could not be disabled or delayed.) To get a handle on this variability, we ran some
experiments several times and compared the times.  Out of 100 queries that
were run, with multiple plans for each, each set run eleven times
independently, we found that the times were tightly clustered around the
median time, with some outliers, either where one or more daemon(s) happened
not to run during that query or where one or more daemons ran more often or
longer than normal during that query.

To understand the variability (and thus the repeatability) of these timings,
we plot in Figure~\ref{fig:stddev} the distribution of the four measured
times around the median time for that plan, eliminating the three lowest and
highest times (hence, we are considering the two times immediately below and
above the median). The $x$-axis is in terms of percentage of the
median. Each plan took around 5 seconds (the fastest plan took 0.3 seconds,
the slowest, 527 seconds, for different queries...). A standard deviation of
2.73\% indicates that 95\% of the plans were within (above or below) an
average of 140msec of the median.

\begin{figure}[bth]\centering
\includegraphics[width=0.50\textwidth]{figures/variance_distribution.eps}
\caption{Distribution of plan timings around the median\label{fig:stddev}}
\end{figure}

Our experimental timing methodology emerged from this initial data. We ran each
plan of each query five times, clearing all the caches before each
execution. We then selected the median time out of these five runs. We
judged a plan~$A$ to be faster than a plan $B$, for the same query, if the
measured execution time of plan 
$A$, $t_A < 0.9454 \cdot t_B$, that is, if it is more than two standard deviations faster. If the query optimizer picked plan $B$ and
there existed such a faster plan~$A$, we marked the query as suboptimal. As an
additional check of the methodology, we randomly selected five sets of five
runs from the set of eleven runs of 100 queries discussed above. For each of
the five sets, we identified the suboptimal queries in each run. The number
of such queries differed by a total of only two queries across all five sets.

\comment{, perhaps
ten executions were extreme outliers; these plans were omitted from further
analysis. We then examined the remaining plans. We believe that the smallest
values is the accurate one since intuitively the execution of that
particular run suffered the least from other system activities. We then
normalized each run to its minimum value and computed the standard deviation
from that value to arrive at a relative threshold that we could use when
testing the hypotheses, as described in later sections.

For DBMS {\bf A}, the relative threshold was 0.169. For DBMS {\bf B}, that
value was 0.138. It was reassuring that these were quite similar, as the
underlying casual factors are likely to be similar. We did not have
sufficient data to compute the threshold for DBMS {\bf C}, and so used the
higher value, 0.169. We will use this threshold later to help
reduce the false positive instances in determining suboptimality.}

\begin{figure}[t]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} executeSpecificExperiment($experimentSpec$):
\STATE initializeLabNotebookContent()
\STATE initializeExperimentTables()
\FOR{$phase$ $\leftarrow$ 1 {\bf to} getNumPhases()}
\STATE stepA($phase$, $varTables$)
  \FOR{$q$ $\in$ getExperimentRunQueries()}
    \IF{$q$ has result}
      \STATE getResultFromLabNotebook($q$)
    \ELSIF{checkToBePaused()}
      \RETURN false
    \ELSE
      \STATE storeQueryResult(analyzeQuery($phase$, $q$, \\
      \hspace{47.0mm}$experimentSpec$))
    \ENDIF
  \ENDFOR
\ENDFOR
\RETURN true
\end{algorithmic}
\caption{Algorithm for Executing a Specific Experiment\label{alg:exeSpecExp}}
\end{center}
\end{figure}

\begin{figure}[t]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} analyzeQuery($phase$, $q$, $experimentSpec$):
  RecordedQueryRuns
\IF{$phase$ = 1}
  \STATE $plan$ $\leftarrow$ getQueryPlan($q.sql$)
  \STATE storeOptimalQueryRun($plan$, stepB($phase$, $plan$))
\ENDIF
\RETURN recordAllQueryRuns(\\
\hspace{18.0mm}findQueryRuns($phase$, $q$, $experimentSpec$))
\end{algorithmic}
\caption{Algorithm for Analyzing a Query\label{alg:analyzeQuery}}
\end{center}
\end{figure}

\begin{figure}[t]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} findQueryRuns($phase$, $q$, $experimentSpec$):
  QueryRunResult
\STATE $queryRunResult$ $\leftarrow$ $\Phi$
\FOR{$cardinality$ $\in$ \\
\hspace{5.0mm}generateCandidateCardinality($experimentSpec$)}
\IF{$\lnot$ stepC($phase$, $varTables$, $cardinality$)}
  \RETURN queryRunResult
\ENDIF
\STATE $newPlan$ $\leftarrow$ getQueryPlan($q.sql$)
\IF{($newPlan$ is subopt plan in the first phase $\wedge$ \\
\hspace{8.0mm} $newPlan$ has {\em NOT} been executed) \\
\hspace{3.0mm} $\vee$ $newPlan$ is different from the previous plan}
  \STATE $queryRunResult$ $\bigcup$ stepD($phase$, $q.sql$, \\
  \hspace{40.0mm} $cardinality$))
\ENDIF
\ENDFOR
\RETURN $queryRunResult$
\end{algorithmic}
\caption{Algorithm for Finding Query Runs\label{alg:findQueryRuns}}
\end{center}
\end{figure}

\c2j{}{\subsection{Experimental Scenarios}}
\c2j{Experiment scenarios use the {\em Template Pattern}~\cite{gamma95} to
  control their variability.}{As discussed in Section~\ref{sec:subopt}, we used two approaches to
determine which queries were suboptimal. These were expressed as two
experimental scenarios named ``VaryStats'' and ``Upper.'' We used the
{\em Template Pattern}~\cite{gamma95} to ensure that the commonality between
these two approaches was explicit.} Specifically, we implemented
an abstract class with several methods \c2j{and, for the purpose of this
  study, one concrete subclass, defining a specific scenario}{and concrete
  subclasses, each defining one of the specific scenarios} through the definition of five
short methods, namely {\tt stepA()}, {\tt stepB()}, {\tt stepC()},
{\tt stepD()}, and {\tt getNumPhases()}. The abstract class provides the
{\tt execute\-SpecificExperiment()}
method (Figure~\ref{alg:exeSpecExp}) along with two helper methods,
{\tt analyze\-Query()} (Figure~\ref{alg:analyzeQuery}) and {\tt findQuery\-Runs()}
(Figure~\ref{alg:findQueryRuns}). \c2j{Each scenario iterates}{Both scenarios iterate} through the
queries, analyzing each one. (Through \azdb, a long-running
experi\-ment, which may take a week or more, can be paused and
restarted.) Most of the work is done in {\tt findQuery\-Runs()}, with the
execution tailored by the individual steps provided by the concrete scenarios.

The ``VaryStats'' scenario produces different plans by changing the
statistics of the variable table.  However, the {\em actual} cardinality of
the varying table remains at 1M rows. Each plan is thus run on the identical
set of tables.

In this scenario, sub-optimality is defined by the existence
of a plan, produced at a non-actual table cardinality, that runs with a
shorter elapsed-time than the ``optimal'' plan chosen and executed at the
(same) actual cardinality. {\tt stepA} populates the variable table. {\tt
  stepB} times the execution plan for the actual table size and returns
statistics on that execution. {\tt stepC} changes the cardinality statistics
of the variable table and always returns true. {\tt stepD} times the plan
for the specified cardinality, which was set by the previous step.

\c2j{}{The ``Upper'' scenario is used when a DBMS does not allow the statistics to
be changed. This scenario simply executes the query by successively changing
the cardinality of the varying table, by starting at the maximum and
deleting tuples. Periodically (every 10K tuples deleted), the query is
optimized, and run whenever the plan changes. In this scenario, a more
conservative definition of suboptimality is used. Suboptimality is defined
by the existence of a plan, produced with a larger varying table table
cardinality, that still runs with a shorter elapsed-time than the
``optimal'' plan chosen and executed at the actual cardinality.

For this scenario, {\tt stepA} again populates the variable table and {\tt
  stepB} times the execution plan for the actual table size and returns
statistics on that execution. {\tt stepC} actually changes the variable
table by deleting rows; it returns false if the requested cardinality is
less than the actual cardinality (hence the name of the scenario: Upper).
As before, {\tt stepD} times the plan
for the specified cardinality, which was set by the previous step.
Note that this scenario avoids some of the complexities mentioned
by Kabra and DeWitt~\cite{kabra98}, as the statistics on the underlying tables are
completely accurate and the amount of available resources are known (only
the one query is run at a time) and the values of host language variables
are also known (the system configuration remains fixed as the queries are
optimized and then immediately evaluated.

\todo{Show same query in upper with changepoints. }

The idea is that the faster query on the larger table would run even faster
if executed on the table with the actual cardinality. We formalize this
assumption as follows.

\noindent
{\bf Definition:} {\em Monotonicity}: given a query $Q$ and an actual
cardinality $a$,

\quad\quad\quad$\forall p \in \hbox{\em plans} (p) \;\forall c > a \;( \hbox{\em time} (p,
c) \geq \hbox{\em time} (p,a) )$

\noindent
where {\em plans}($p$) is the set of plans generated by the optimizer and
{\em time}($p$, $c$) is the execution time of plan $p$ on the data set with
the varying table at cardinality $c$.Given the variability discussed in
Section~\ref{sec:dataanalysis}, we augment this definition with a threshold
$\sigma$,

\quad\quad\quad$\forall p \in \hbox{\em plans} (p) \;\forall c > a \;( \hbox{\em time} (p,
c) \geq \hbox{\em time} (p,a) )$

\hfill$\blacksquare$

To test this assumption, we counted the number of queries, out of the total
of 1000 queries, applying the above definition of a monotonic query. We
examined every case where the same plan was run at least two times, at
different varying cardinalities. Note that for some queries, multiple plans
were each run more than one. \todo{true?} We
expect that there should be no non-monotonic queries. The results of this
analysis is in Table~\ref{tab:monotonic}.

\begin{table}
\caption{\label{tab:monotonic}}
\end{table}

\todo{discussion}

Given this analysis, we can now use the Upper scenario as a more
conservative proxy for the suboptimality observed directly in the
VaryStats scenario. In the remainder, we will presents results for the
Upper scenario for DBMSes {\bf A}, {\bf B}, and {\bf C}. When appropriate,
we will also present results for the VaryStats scenario for DBMS {\bf A}.

\todo{some general discussion about the presence of suboptimal, and give the
graph of the time difference between the fastest plan and ``optimal'' plan.
\comment{\begin{figure}[t]
\centering
%originally 30pc
\epsfig{figure=vldb_figures/figure10/figure10.eps,width=20pc}
\caption{Difference in Execution Time between Chosen Plan and Fastest Plan\label{fig:diff}}
\end{figure}
}

Several initial conclusions can be drawn from this experiment, which was the
initial result of several semesters of programming effort and many days of
experimental runs. First, {\em every} DBMS exhibits suboptimality. Thus, this
phenomenon is likely to be a fundamental aspect of either the algorithm
(cost-based optimization) or the creator of the algorithm (human information
processing). Our theory includes both effects.}

\subsection{Validity of Upper Scenario}
Between the two categories of suboptimality, the {\em VaryStats}
scenario directly shows whether the plan selected for the actual cardinality
is optimal.  However, the {\em Upper} scenario relies on the monotonicity
assumption, which we showed holds. Therefore, we expect that the set of
queries observed with suboptimality in the {\em Upper} scenario is a subset
of the set of suboptimal queries in the {\em VaryStats} scenario. In
practice, due to the measurement variation in experiments, there exist
exception instances. Nevertheless, we determine for each scenario a proper
variation threshold by which most exceptions or false positive
suboptimalities can be eliminated. Thus, we expect that the suboptimality
set in ``Upper'' scenario is a very {\em close} subset of the one in
``Suboptimal'' scenario.

Figure~\ref{fig:mono_check} {\todo Draw figure}shows the percentage of suboptimality with various
variation threshold for both ``Upper'' and ``VaryStats'' scenarios. Note that
we choose for each scenario the threshold value at the obvious turning position
of the curves in that we believe that is where most the false positive 
suboptimality is filtered out. Table~\ref{tab:mono_check} demonstrates the
actual queries in the two suboptimality sets. {\todo what can be seen? }
}

### 2017-08-08 subopt-tods17/OLDtex/detecting.tex
subopt tods submission

\subsection{Detecting Suboptimality}
\todo{Get adjacent from operationalization.
 assumptions: time query repeatable, timing monotonic}

\todo{Rick has modify token: merge these two sections.}
\todo{ original adjacent scenario, change points, monotonicity}

How might such behavior be observed? We have developed a system, \azdb, that
allows us to perform experiments to study this phenomenon of suboptimality.
\azdb\ can externally modify data
dictionary statistics for tables and columns, thereby allowing the DBMS to
generate different plans for a given query.
\hbox{\azdb} then presents queries to the DBMS while varying
the statistics, in particular the cardinality of one of the tables,
requesting in each case the evaluation plan chosen by the DBMS.
This is done using the
{\tt EXPLAIN} SQL facility available in modern DBMSes. We can thus compare
the performance (execution time) of various plans for the query, to identify those
situations when the wrong plan was chosen, when in fact there was a
different plan that was semantically equivalent to the chosen plan (that is,
yielded the identical result) but which ran faster.

\begin{figure*}[tb]
\centering
%originally 30pc
\epsfig{figure=figures/model.eps,width=30pc}
\caption{Predictive Model of Suboptimality\label{fig:model}}
\end{figure*}

\comment{\azdb\ modifies the cardinality statistics to produce multiple execution
plans for a given query. For one of the DBMSes, we can modify the stored
table statistics directly. For the other DBMSes, we had to do so indirectly,
by varying the size of the table and running the optimizer on tables of
different size. As \azdb\ varies the cardinality statistics of tables, it
collects all the plans that the optimizer felt were appropriate for that
query.  After \azdb\ gathers the plans, it then executes each plan on the
original table(s), to determine which plan was actually the best, of the
plans generated by the optimizer. 
}
It is important to note that all of this manipulation can be done {\em
  outside} the DBMS. For proprietary systems we do not have access to the
internal code. We don't know ({\em can't} know) all the plans that were
considered, nor the details of how the plans are selected. But such access
is not needed; indeed, to be able to study a phenomenon across many DBMSes,
such access is not practical. But by designing the experiment to examine
the plans that the DBMS actually produces, and thus to examine phenomena
that can be externally visible, we can obtain valuable insights into general
classes of computational artifacts. Our ultimate goal is to make statements
that hold across cost-based optimizers in general, thereby moving up the
$y$-axis of Figure~\ref{fig:empirical}.

### 2017-08-08 subopt-tods17/OLDtex/introduction.tex
subopt tods submission

\section{Introduction}
\todo{Rick:motivate suboptimality. } 
\shorten{Put back flutter for journal version of paper}
DBMSes
underlie all information systems and hence optimizing their performance is of
critical importance. One could argue that optimizing a query's performance
could be done by the software engineer. However, the time taken to examine,
identify and correct inefficiencies for every query in the DBMS would result
in a huge expenditure of time and effort. What needs to be done instead is
to systematically examine the factors influencing the number of suboptimal
queries. There could be multiple causes of the suboptimality: One possible
cause could be some peculiarity within the tens of thousands of lines of
code of that query optimizer. Another possible cause could be the query
complexity. A third possible reason could be some fundamental limitation
{\em within the general architecture of cost-based optimization} that will
always render a good number of queries suboptimal. Prior research shows that
increasing the complexity negatively influences
performance~\cite{campbell88,moody98}.

To better understand the impact of different factors on suboptimality of
query performance and the interaction between operators, especially in a
dynamic environment, an experimental approach is needed. This is where our
research makes its contribution.  Our research introduces a novel approach
to better understand the factors influencing query performance in
DBMSes. Based on existing research and general knowledge of DBMSes, we
developed a simple predictive model to predict performance, in particular
suboptimality of query performance. We use an experimental methodology with
the DBMS as a subject to test our hypotheses with respect to factors
influencing suboptimality. Our research falls within creative development of
new evaluation methods and metrics for artifacts, which were identified as
important design-science contributions~\cite{hevner04}.

The key contributions of our research are as follows.
\begin{itemize}
\item It builds and tests a predictive model for DBMSes to better understand the factors influencing performance.

\item It uses an innovative methodology that treats DBMSes as experimental subjects.

\item The predictive model has some unsettling practical implications.
\end{itemize}

This paper takes a scientifically rigorous approach to an
area previously dominated by the engineering perspective, that of database
query optimization.  Our goal is to understand cost-based query optimizers
as a {\em general} class of computational artifacts and to come up with
insights and ultimately with predictive theories about how such optimizers,
again, as a general class, behave.  These theories can be used to further
improve DBMSes through engineering efforts that benefit from the fundamental
understanding that the scientific perspective can provide.

We focus here on one aspect, that of the effectiveness of query
optimization. The query optimization phase within a DBMS ostensibly finds the fastest
query execution plan from a potentially large set of enumerated plans, all of
which correctly compute the specified query. Occasionally the ``optimizer''
selects the wrong plan, for a variety of reasons. We define {\em
  suboptimality} as the existence of a query plan that performs more
efficiently than the plan chosen by the query optimizer. From the engineering
perspective, it is of critical importance to understand the prevalence
of suboptimality as well as how it arises.

We quantitatively study these aspects \c2j{}{across DBMSes }to identify the
underlying causes. We have developed an initial theory that identifies two
factors that may play a role in suboptimality. The ultimate goal is to {\em
  understand} a component of a DBMS, its cost-based optimizer, through the
articulation and empirical testing of a general scientific theory.

In Section~\ref{sec:related} we briefly summarize the vast amount of related
work in query optimization to establish the technical basis for our study.
The following section introduces the methodology we will follow, that of
empirical generalization. We then present a predictive, causal model of
suboptimality and empirically test several hypotheses derived from that
model\c2j{. These are the first scientific results that we are aware of that
  test a causal model of DBMSes in general.}{ on three DBMSes.
  These are the first scientific results that we are aware of that apply
  {\em across} DBMSes, rather than on a single, specific DBMS or on a
  specific algorithm.}

### 2017-08-08 subopt-tods17/OLDtex/measuring.tex
subopt tods submission

\subsection{Measuring Query Execution Time}\label{sec:subopt}
\todo{Young has modify token}
\todo{phantom...}
In order to quantitatively examine the effects of suboptimality, we need to
measure the execution time of each plan of each query.

In this scenario, sub-optimality is defined by the existence
of a plan, produced at a non-actual table cardinality, that runs with a
shorter elapsed-time than the ``optimal'' plan chosen and executed at the
(same) actual cardinality. {\tt stepA} populates the variable table. {\tt
  stepB} times the execution plan for the actual table size and returns
statistics on that execution. {\tt stepC} changes the cardinality statistics
of the variable table and always returns true. {\tt stepD} times the plan
for the specified cardinality, which was set by the previous step.

\todo{Rick: Could you please read through this paragraph. }
\todo{Extract text about change points and modify for adjacent. }
\todo{Talk about adjacent and why we measured it and how. } 
The ``optimal'' plan is the plan chosen at the actual cardinality. It may 
be chosen at other cardinalities too. 
The ``Adjacent'' scenario is used when a DBMS does not allow the statistics to
be changed. This scenario simply executes the query by successively changing
the cardinality of the varying table, by starting at the maximum and
deleting tuples. Periodically (every 10K tuples deleted), the query is
optimized, and run whenever the plan changes. In this scenario, a more
conservative definition of suboptimality is used. After the identification of the occurrences of change points, i.e., when the optimized plan chosen changes, 
the running time for each optimized plan is compared to the ``optimal'' plan that is 
closest to it at a lower cardinality. Suboptimality is defined
by the existence of a plan, produced with a larger varying table table
cardinality, that still runs with a shorter elapsed-time than the
``optimal'' plan chosen and executed.

We worked very hard to ensure the correctness and repeatability of the
execution time measurements. In developing \azdb, we encountered many sources
of DBMS variance for a given DBMS, a given query, and a given plan. Hence,
much of our effort was devoted to determining why a particular measurement
wasn't repeatable, and then eliminating that variance. We were hampered by
the fact that DBMS documentation is generally quite incomplete on this
subject\c2j{}{, and by the fact that our solutions were quite idiosyncratic to the
DBMS involved}. We were also challenged by the fact that we had to be able to
solve the problem only through the facilities available to JDBC: we could
not make any internal changes to the DBMS.

The presence of flutter, sometimes prevalent, indicates that the optimizer
appears to not be making good choices as the cardinality estimate is less
and less accurate. We saw that graphically in Figure~\ref{fig:query769}, as
the query plan fluttered between Plans 2 and 4 for a wide range of
cardinality estimation error for Query~769.

Perhaps the optimizer is not even making a good choice when the estimate is
entirely accurate. Such was the case for Query~769: that plan was not the
fastest of the five plans, having an execution time that was almost the same 
as the worst plan, Plan~2.

The presence of flutter indicates the optimizer is having a hard time
picking a plan. It seems possible that in such situations the optimizer may
pick the wrong plan.

Because we can execute the various plans generated by the optimizer on the
actual cardinality, we can see whether the optimizer in fact picked the best
plan, out of the options in the effective plan space. We ran the same 600 queries 
with timing enabled on DBMS {\bf A}.

It is interesting to see how far off the optimizer
was in such cases. Figure~\ref{fig:diff} shows the execution time difference by percentage
between the chosen plan and the fastest plan, shown ordered by
percentage. Our results show that most of the sub-optimal queries have chosen plans 
that are several times slower that the fastest plan.

\c2j{We obtained different plans for each query by varying the cardinality
  statistics of one of the tables; these statistics are used by the optimizer.}{We obtained different plans for each query by varying the cardinality
of one experiment table, which we called the ``variable table'' in our
experiments. We developed two approaches to perform the tests.

In the first approach, we fixed the actual size of the variable table. However,
we altered the table statistics stored in the DBMS system catalogs; these
statistics are used by the optimizer.In practice, 
only one DBMS out of the three allowed us to manipulate the 
table statistics through the JDBC API. Thus we present results on 
query plan execution time using such approach only for DBMS~{\bf A}.}
The purpose of this approach is to find out whether there exists a plan
that is better than the plan the DBMS chose as the ``optimal one'' with
accurate cardinality, given that these plans are run on the same set of
tables.

\c2j{Utilizing JDBC to manipulate independent variables from outside the
  DBMS allows us to empirically generalize by moving up the $y$ axis of
  Figure~\ref{fig:empirical}, from one system to several systems and then to
  a general theory. However, as we discuss in detail shortly, obtaining {\em
    repeatable} and {\em accurate} timings of database queries was
  unexpectedly difficult. So in this paper we report experiments over just
  one DBMS. We plan to generalize the results and the tests over a range of
  DBMSes, once we are able to obtain accurate, repeatable timing results for
  each. We don't reveal the identity of the DBMS we studied, for two reasons. First, commercial DBMSes
include in their user agreements requirements not to release 
performance data. This is detrimental to science, but we have no choice but
to live with that restriction. However, in some sense the specific DBMS
doesn't matter, as we are studying phenomena about cost-based optimizers
{\em in general}, and so are interested in making statements that apply to
all the experimental subjects in our study.}{In the second approach, we varied the cardinality of the variable
  table by actually deleting tuples from it. To ensure the correctness of
  the plan generation and thus time measurements, we performed an ``ANALYZE
  TABLE`` function to force the DBMS to update the table's new statistics to
  be accurate. The purpose of this approach is find whether there exists a
  plan that runs faster with a larger table than a different plan that is
  executed on a smaller table.}

\shorten{\begin{figure*}[htb]
\begin{small}
\begin{verbatim}
<experiment name="X" scenario="veryCardinalityStats" dbms="...">

    <dataDefinition name="ft">
      <table name="HT1" cardinality="1000000">
        <column name="id1" dataType="number" dataLength="10" dataGenerationType="sequential"
                distributionMinimum="0" distributionMaximum="1000000" inPrimaryKey="false"/>
        <column name="id2" dataType="number" dataLength="10" dataGenerationType="random"
                distributionMinimum="0" distributionMaximum="1000000" inPrimaryKey="false"/>
        <column name="id3" dataType="number" dataLength="10" dataGenerationType="random"
                distributionMinimum="0" distributionMaximum="1000000" inPrimaryKey="false"/>
        <column name="id4" dataType="number" dataLength="10" dataGenerationType="random"
                distributionMinimum="0" distributionMaximum="1000000" inPrimaryKey="false"/>
      </table>
      <table name="HT2" cardinality="1000000"> (same as above) </table>
      <table name="HT3" cardinality="1000000"> (same as above) </table>
      <table name="HT4" cardinality="1000000"> (same as above) </table>
    </dataDefinition>

    <tableConfiguration>
      <variableTableSet searchMethod="linear" searchGranularity="10000"> <table name="HT1" seed="1999">
          <cardinality hypotheticalMinimum="1" hypotheticalMaximum="2000000"/>
        </table>
      </variableTableSet>
      <fixedTableSet>
        <table name="HT2" seed="2999"> <cardinality hypothetical="actual"/> </table>
        <table name="HT3" seed="3999"> <cardinality hypothetical="actual"/> </table>
        <table name="HT4" seed="4999"> <cardinality hypothetical="actual"/> </table>
      </fixedTableSet>
    </tableConfiguration>

    <queryDefinition numberQueries="500" ...>
      <grammar>
        <select maxColumns="4"/>
        <from maxCorrNames="4" useDuplicates="true"/>
        <where cartesianPossible="false" maxIsAbsolute="true" maxPredicates="0" complexUsePercentage="100">
          <binaryOperators> <operator symbol="="/> </binaryOperators>
          <binaryLogicalOperators> <operator symbol="AND"/> </binaryLogicalOperators>
          <unaryLogicalOperators> <operator symbol="NOT" usePercentage="0"/> </unaryLogicalOperators>
        </where>
      </grammar>
    </queryDefinition>

</experiment>
\end{verbatim}
\end{small}
\vspace*{-4ex}\caption{The XML specification of the experiment\label{fig:expspec}}
\end{figure*}
}

\input{timing.tex}

### 2017-08-08 subopt-tods17/OLDtex/measuring_time.tex
subopt tods submission

\subsection{Time Measurement on A Single Query Execution}\label{sec:measure_time}

\begin{figure}[th]\centering
\includegraphics[width=0.50\textwidth]{figures/time_measurement.eps}
\caption{Processes considered when timing a query\label{fig:timing}}
\end{figure}

In this section we describe how the time on a single query execution 
is measured. 
There could be in practice quite a few non-DBMS processes simultaneously 
at the time of executing a query. 
Hence, a careful measurement is necessary so that we can later derive 
the time only spent on a DBMS process(es) involving in the single 
query execution (as described in the next subsection).

First, $GetProcs$() is called to capture all running processes, 
as illustrated in Figure~\ref{fig:timing}. 
To obtain each of the process information, 
the method either iterates {\tt /proc/\{pid\}/stat} or 
takes advantage of {\tt PsList}~\cite{russinovich10}, depending on Unix/Linux 
or Windows operating systems, respectively. 
Once all the information associated with a process 
is extracted, a record of \{process id (pid), command (process name), 
minimum and major faults, cpu time and system time\} is built and mapped 
to the pid to identify the process. 
Subsequently, $GetProcStat$() is invoked to obtain all detailed CPU ticks and 
the max pid at the time of the call. 
The method accesses {\tt /proc/stat} that maintains kernel and system statistics 
(not available on Windows), and it extracts the current 
CPU ticks by user mode, system mode, idle task, wait mode for I/O completion, 
servicing interrupt mode, etc., and lastly the max pid.
The tick information is needed to compute the correct single execution time, 
and the max pid is used to detect the presence of a 
{\em phatom} process(es) in the single query execution (as will be discussed shortly). 
Next, an individual call is made to $GetTime$() right before and 
after executing the query $Q$, thereby measuring the elapsed 
(probably not accurate) time by difference between the start and end times. 
$GetProcStat$() and $GetProcs$() are in turn invoked again after the execution. 
In particular, it is important not to reverse the order of calling 
$GetProcs$() and $GetProcStat$(), in that a phantom process(es), otherwise, 
would be more likely to happen while $GetProcs$() extracts all process information. 
The prevalence of a phantom process(es) included in a single query execution 
could spoil our analysis. 
We will discuss the phantom process in greater detail shortly.

Now let us consider how the captured processes may influence 
the single query execution. 
As illustrated in Figure~\ref{fig:timing}, 
$P$1, $P$2 and $P$3 are recorded into a map, $M_{1}$ when the first $GetProcs$() 
gets called. 
$P$4 might also appear in $M_{1}$ if $GetProcs$() captures it. 
Another map, $M_{2}$, built by the second $GetProcs$(), 
records $P$3, $P$6 and $P$7, and it might also have $P$8 
for the same reason as $P$4. 
$P$5 is called a phantom process, since it appears and completes 
its task before the query execution is finished. 
$P$5, thus, cannot be captured by $GetProcs$(), but at least we can tell 
whether it would have been present in the query execution 
by comparing the max pids each other extracted by $GetProcStat$(). 
If no difference is made, the query execution data is said to be {\em clean} 
without any phantom process. 
Otherwise, the phantom must have happened during the query execution, so 
the data gets dirty, and thus, it is not considered in our analysis.

There could be other situations where the execution data cannot be 
used. 
The comparison result with $M_{1}$ and $M_{2}$ reveals that 
$P$1 and $P$2 are not found in $M_{2}$. 
These processes are called {\em stopped} ones. 
It is evident that $P$2 involves the query execution, but it is not easy 
to determine that $P$1 is the same case as $P$2, only by the comparison. 
Therefore, our analysis do not make use of the execution data 
if any stopped processes are detected. 
If $P$4 is captured, it would be counted as a stopped process, and thus, 
the data gets thrown too. 
However, the data would not be unclean although $P$4 is not captured, 
for it does not get involved in the query execution.

$P$6 and $P$7 are located in $M_{2}$ but not in $M_{1}$. 
Thus, they are said to be {\em started} processes, which could involve 
the query execution. 
Because it is not impossible to get their start times, 
we do not throw the data having started processes. 
Instead, their times will be subtracted from the measured, elapsed time.
To minimize the subtraction as well, 
``{\em infrequently-running} processes'' followed by 
``{\em frequently-running} processes'' are checked in order 
at the first call of $GetProcs$(), and they are checked 
in the other way around at the second call of $GetProcs$() 
when the maps $M_{1}$ and $M_{2}$ are built individually.

The frequently-running processes mean ones found to be run during 
the execution by the comparison of $M_{1}$ and $M_{2}$. 
For instance, $P$3 could be in the frequently-running process list should it 
show different CPU ticks, thereby proving that it would run during the execution. 
The experiment starts by loading the existing frequently-running process list 
with a small (tunable) number of processes (e.g., 10). 
As the experiment goes, a process could be automatically added to the list 
as long as it did not run last execution but did run this time. 
By adjusting the order of examining processes in $GetProcs$(), 
the tick increase of the running processes could be minimized, and accordingly, 
more accurate query execution time could be computed, minimally 
incurring the subtraction caused by the increase ticks. 
Finally, $P$8 could be classified as a started process once detected. 
It would otherwise be fine since it does not be involved in the query 
execution, either.

Figure~\ref{alg:exeSpecExp} represents an algorithm of measuring 
the single query execution time as described above. 
Input arguments, $\iterNum$, $\query$, $\prevPlan$ and $\card$, 
denote the current iteration number for repeatability check, 
a given query, the previously detected query plan 
and the current cardinality at which the query gets executed, 
respectively. 
First, the algorithm checks if the given plan, $\prevPlan$ 
is identical to a current plan obtained 
at the time of executing the $\query$ via $GetQueryPlan$(). 
If both plans are inconsistent, then an error is reported. 
Next, to remove cache effects, all the caches from disk through 
operating system to DBMS are flushed.
Before executing $\query$, through $GetProcs$(), $GetProcStat$() and 
$GetTime$() we get $\procMap_{1}$, which 
records all existing process information, 
$cpuStat_{1}$, which stores current CPU ticks and max process ID, and 
$time_{1}$, which gets the current time in milliseconds, in order. 
After finishing the query execution, we again obtain 
$time_{2}$, $cpuStat_{2}$ and $\procMap_{2}$ by calling the above 
methods in reverse order. 
We then get $\mTime$, the measured time on this single query execution, 
and through $\diff$() we receive 
the $\procDiff$ string, which contains executed, started and stopped 
processes during the query execution and their detailed information (increased 
cpu and system times and min and major faults) and the increased CPU ticks 
per entry (such as user and system modes, etc.), if any. 
Lastly, $RecordQueryExecution$() stores all the results, 
$\mTime$ and $\procDiff$ with $\iterNum$, $\query$ and $\card$.

\begin{figure}[t]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} timeSingleQueryExecution($\iterNum$, $\query$, \\
						\hspace{44.0mm}$\prevPlan$, $\card$):
\STATE $\plan$ $\leftarrow$ $GetQueryPlan$($\query$)
\IF{$\plan$ $\neq$ $\prevPlan$}
	\STATE Error: two different plans are observed \\ 
	at the same cardinality. \\
\ENDIF
\STATE Flush disk drive, OS and DBMS caches. \\
\STATE $\procMap_{1}$ $\leftarrow$ $GetProcs$() \\
\STATE $cpuStat_{1}$ $\leftarrow$ $GetProcStat$() \\
\STATE $time_{1}$ $\leftarrow$ $GetTime$() \\
\STATE $executeQuery$($q$) \\ 
\STATE $time_{2}$ $\leftarrow$ $GetTime$() \\	
\STATE $cpuStat_{2}$ $\leftarrow$ $GetProcStat$() \\
\STATE $\procMap_{2}$ $\leftarrow$ $GetProcs$() \\
\STATE $\mTime$ $\leftarrow$ $time_{2}$ - $time_{1}$ \\	
\STATE $\procDiff$ $\leftarrow$ $\diff$($cpuStat_{1}$, $cpuStat_{2}$, \\
\hspace{22.0mm}							$\procMap_{1}$, $\procMap_{2}$) \\	
\STATE $RecordQueryExecution$($\iterNum$, $\query$, $\card$, \\ 
			\hspace{35.0mm}$\mTime$, $\procDiff$) \\
\end{algorithmic}
\caption{Algorithm for The Single Query Execution Time Measurement\label{alg:exeSpecExp}}
\end{center}
\end{figure}

\subsubsection{Query Execution Time Computation}\label{sec:qe_time}

Figure~\ref{fig:dbms_time} shows the interleaved processes, 
$P_{\other}$ and $P_{\dbms}$, running on a single core. 
$P_{\dbms}$ represents a DBMS-related process, and $P_{\other}$ 
denotes a non-DBMS process. 
To derive the query execution time ($T$) only taken by $P_{\dbms}$, 
the following equation is used:

\begin{center}
$T$ (ms) = (($U_{\dbms}$+$S_{\dbms}$+$\IO_{\total}$+$\Irq_{\total}$+
$\SoftIrq_{\total}$) - ($\sum_{i=1}^{n}{S_{i}}$)) (ticks) $\times$ 10, 
\end{center}
where $U_{\dbms}$ and $S_{\dbms}$ denote ticks on user and system 
modes of $P_{\dbms}$, and $\IO_{\total}$ and $\Irq_{\total}$ and 
$\SoftIrq_{\total}$ mean ticks by wait mode for IO completion 
and servicing interrupts and soft interrupts, respectively, 
and $S_{i}$ denotes the ticks subtracted due to a started process $i$.

$U_{\dbms}$ and $S_{\dbms}$ are obtained from a DBMS process entry 
and $\IO_{\total}$, $\Irq_{\total}$ and $\SoftIrq_{\total}$ from 
the detailed CPU tick information that are contained 
in the $\procDiff$ string by the $\diff$ method in Figure~\ref{alg:exeSpecExp}. 
Since a tick is typically one hundredth of a second (according to man page), 
it is required to multiply the total ticks by ten in the equation. 
In Figure~\ref{fig:dbms_time}, $T$ can be derived by summing up 
the times only taken by DBMS as follows: 
$T$ = ($t_{5}$-$t_{0}$)+($t_{7}$-$t_{6}$)

In particular, $\mTime$ in Figure~\ref{alg:exeSpecExp} is 
ensured to be clean by the following sanity check for consistency with 
the increased tick information: 
\begin{center}
$\vert$($U_{\total}$+$S_{\total}$+$\Idle_{\total}$+$\IO_{\total}$+$\Irq_{\total}$+$\SoftIrq_{\total}$) 
- ($\mTime$*0.1)$\vert$ <= 1 
\end{center}

\begin{figure}[t]\centering
\includegraphics[width=0.48\textwidth]{figures/dbms_time.eps}
\caption{Two interleaved, running processes\label{fig:dbms_time}}
\end{figure}

### 2017-08-08 subopt-tods17/OLDtex/methodology.tex
subopt tods submission

\section{Methodology}

\input{model.tex}

\input{detecting.tex}

\input{measuring.tex}

\input{repeatability.tex}

%\input{monotonicity.tex}

\subsection{Summary of Sub-Optimality}
\shorten{
In this paper we utilize a new methodology to ask new
questions about a mature topic, query optimization, in \hbox{order} to develop and
test a novel model and theory about when suboptimality arises, to
arrive at a new understanding of query optimization. This may later enable
new engineering approaches as well as a better understanding of the limits
of existing approaches.

\begin{figure}[bth]\centering
\includegraphics[width=0.50\textwidth]{figures/understanding.eps}
\caption{Empirical generalization ([3, page~5])\label{fig:empirical}}
\end{figure}

This paper utilizes {\em empirical
  generalization}, in which understanding proceeds along two conceptual
dimensions, as shown in Figure~\ref{fig:empirical}. As Cohen~\c2j{\cite{cohenbook}}{[1995]} has
articulated, science (in general, and this paper specifically) progresses
(in the figure, on the 
$y$-axis) from studies of a particular system (e.g, a particular database
query optimization algorithm in the context of a particular database
management system), through studies of a class of systems (e.g.,~a study
{\em across} several disparate DBMSs), to statements about computational
tools in general (e.g.,~a theory that holds for rule-based optimizers,
whether in DBMSs, AI systems, or compilers). This progression increases the
domain of applicability, and thus the generality, of a theory. Science also
progresses (in the figure, on the 
$x$-axis) from description of the phenomenon, to prediction, and eventually
through causal explanation, within an articulated, thoroughly-tested theory.
We seek the rather ambitious outcome of the articulation of a general theory
of suboptimality that applies to the entire class of DBMSes and which
provides causal explanation of suboptimality: when it will arise and why.

Note that this line of study is very different from conventional
approaches summarized above from the literature. We are trying to understand
cost-based query optimizers as a {\em general} class of computational artifacts,
to come up with insights and ultimately with predictive theories about how such
optimizers, again, as a general class, behave~\cite{davies}.}

### 2017-08-08 subopt-tods17/OLDtex/model.tex
subopt tods submission

\subsection{A Model of Suboptimality}
\todo{Sabah has the modify token}
The model suggests how some factors may impact the prevalence of
suboptimality. (In preliminary investigation, we have seen that
different queries are suboptimal in different DBMSes, and so it is
more productive to focus on the prevalence of suboptimality than on
the harder, and perhaps less illuminative, task of predicting which
individual queries produce suboptimal plans.) We present our model in
a hierarchical manner such that the dependencies among the components
can be directly revealed. We partition the factors in our model into
two categories, namely external and internal factors. ``Query
complexity'' is thus an external factor, in that they are not part of
the query optimizer. We can manipulate the external factors in the
experiment. Query complexity includes such factors as the number of
correlation names in the from clause, the presence of aggregates, the
number of self joins. The number of correlation names in the from
clause and the presence of aggregates will weakly influence
suboptimality directly. Instead, all the Query complexity
instantiations will influence suboptimality indirectly by influencing
plan space complexity. The number of correlation names in the from
clause will exponentially influence plan space complexity. The
presence of aggregates will linearly influence plan space complexity.
The number of self joins will negatively correlate with plan space
complexity. Structure of the tables used by the queries is another
external factor called ``Schema complexity''. Schema complexity
include factors like presence of primary keys and indexes.

[Is the sentence below accurate?]
Presence of primary keys, or indexes if present on columns involved in
the join condition between tables involved in a query, will reduce the
plan space complexity and thus inderectly reduce suboptimality present
in the query. [What is the function of the relationship between
presence of indexes and CEPS / Suboptimality]

Concerning the internal factors, some are independent variables, which
we can control, and some are {\em latent variables}, which are derived
from independent variables or which are internal to the optimizer. The
number of aggregates is an independent variable.

One of the central causal factors of this model is the optimizer
complexity. The model states that if the optimizer is very complex,
i.e., if the optimizer has to contend with a large number
of available operators, that will also increase the number of
interactions, and thus the
prevalence of suboptimal queries.

As we cannot observe the plan space complexity of the optimizer
directly, we include in the model other components, which may, from
different angles, indirectly help us understand the effects of the
complexity of the optimizer. Given a particular query, its measurable
complexity will impact the total number of candidate plans considered
by the optimizer. However, this latter factor is not directly
observable, again, especially for propriety systems. However, we {\em
can} measure the number of plans actually generated by the optimizer
when presented with different
cardinalities of the underlying tables. We term this set of plans
actually generated the
``effective plan space'' and the number of such plans, the
``cardinality of the effective plan
space,'' or {\em CEPS}. The model states a correlation between the
latent variable of the number
of plans considered and the measurable dependent variable of CEPS. The
model also states a
positive correlation between the number of operators available in the
DBMS and plan space
complexity. Another feature of the model is the positive correlation
between the number of
operators available in the DBMS and the number of operators observed
for each functionality. Only a subset of the physical operators
available will actually be considered depending on the
presence of a particular functionality (such as join) in the query.

We have just {\em described} how suboptimality might arise, through a
theory and its elaborated
model. We now discuss {\em prediction}.
How might we test such a theory? We hypothesize from our model that as
the optimizer becomes increasingly complex, more occurrences of
suboptimality will be observed, because of such unintended
interactions. Similarly, with a more simplistic optimizer, we
hypothesize that there will be fewer unanticipated interactions, and
so less suboptimality would be present. We predict that one source of
the suboptimality is the designer's focus on correctness and
performance, making it difficult to predict subtle interactions
between components of the optimizer. Later we explore the
ramifications of this hypothesis, finding support for our proposed
model.

\todo{Questions:}
1. I presume the following should go in section 1/ 4, since they are
related to how we
operationalized / tested hypotheses
- heuristic nature of some DBMSes
- Daemons and how we controlled for them

2. Do we still think that the different components of query
complexity, schema complexity and
optimizer complexity will influence suboptimality through CEPS? If so,
why didn't we see the
relationship last time? Or is the relationship between the independent
variables through cost
model accuracy (or some other factor) that we haven't operationalized yet?

3. How does primary keys / indexes impact suboptimality?

\todo{ operationalization def, high level}
Once a particular query is determined to be suboptimal, the obvious
next step would be to look at the code of the query optimizer (if
indeed the code is even available) and alter it in some way so that it
to reduce suboptimality. What we have done instead is to approach this
phenomenon from a scientific perspective, attempting to move to the
right on the $x$ axis of Figure~\ref{fig:empirical}, by developing a
simple predictive model of suboptimality, shown in
Figure~\ref{fig:model}.

The purpose of query optimization is to generate optimal plans.  So
why would suboptimality occur in the first place? Query optimizers are
highly complex, comprised of tens of thousands of lines of code. There
are several reasons for this complexity. First, an optimizer must
contend with the richness of the query language, SQL, whose definition
requires about 2000 pages~\cite{SQL2008}, with a multiple of
linguistic features. Second, an optimizer must content with the
richness of the physical operators available to it. DBMSes have a
range of algorithms available to evaluate each of many algebraic
operators. Third, the optimizer must contend with an exponential
number of query evaluation plans. Kabra and DeWitt~\cite{kabra98}
identify several other sources of complexity: inaccurate statistics on
the underlying tables and insufficient information about the runtime
system: ``amount of available resources (especially memory), the load
on the system, and the values of host language
variables.''~(page~106). They also mention user-defined data types,
methods, and operators allowed by the newer object-relational
systems~\cite{Melton03}. Thus, the task of optimization is very
complex, with the result that the optimizers themselves consist of a
collection of ``components,'' that is, the rules or heuristics that it
uses during optimization, with each of these components being itself
complex.

Our theory, which we will investigate in some detail and which is the
basis for our predictive model, is that suboptimality is due in part
to the complexity of the optimizer and concomitant unanticipated
interactions between the different components used within the 
optimizer.  We argue that with the proliferation of concerns and
variability that an optimizer must contend with, it is extremely
difficult to ensure that for an arbitrary query that the optimizer
will {\em always} pick the right plan. The components of the optimizer
are all interacting with each other; our theory implicate these {\em
interactions} as one source of suboptimality.

The model of Figure~\ref{fig:model} suggests how some factors may
impact the prevalence of suboptimality. (In preliminary investigation,
we have seen that different queries are suboptimal in different
DBMSes, and so it is more productive to focus on the prevalence of
suboptimality than on the harder, and perhaps less illuminative, task
of predicting which individual queries produce suboptimal plans.) We
present our model in a hierarchical manner such that the dependencies
among the components can be directly revealed. We partition the
factors in our model into two categories, namely external and internal
factors. ``Query complexity'' is thus an external factor, in that it is 
not part of the query optimizer. We can manipulate the external 
factors in the experiment. Query complexity includes such factors as
the number of correlation names in the from clause, the presence of
aggregates, the number of self joins. The number of correlation names
in the from clause and the presence of aggregates will weakly
influence suboptimality directly. Instead, all the Query complexity
instantiations will influence suboptimality indirectly by influencing
plan space complexity. The number of correlation names in the from
clause will exponentially influence plan space complexity. The
presence of aggregates will linearly influence plan space complexity.
The number of self joins will negatively correlate with plan space
complexity.
Therefore, our hypotheses with respect to the impact of Query Complexity are
Hypothesis 1: The number of correlation names will exponentially influence
Plan space complexity.
Hypothesis 2: The presence of aggregates will linearly influence plan space
complexity.
Hypothesis 3: The number of self joins will negatively correlate with
plan space complexity.

Concerning the internal factors, some are independent variables, which
we can control, and some are {\em latent variables}, which are derived
from independent variables or which are internal to the optimizer. The
number of aggregates is an independent variable.

One of the central causal factors of this model is the optimizer
complexity. The model states that if the optimizer is very complex,
i.e., if the optimizer has to contend with a large number of available
operators, that will also increase the number of interactions, and
thus the prevalence of suboptimal queries.
Hypothesis 4: As the number of operators increase, the prevalence of suboptimality will increase.

As we cannot observe the plan space complexity of the optimizer
directly, we include in the model other components, which may, from
different angles, indirectly help us understand the effects of the
complexity of the optimizer. Given a particular query, its measurable
complexity will impact the total number of candidate plans considered
by the optimizer. However, this latter factor is not directly
observable, again, especially for propriety systems. However, we {\em
can} measure the number of plans actually generated by the optimizer
when presented with different cardinalities of the underlying tables.
We term this set of plans actually generated the ``effective plan
space'' and the number of such plans, the ``cardinality of the
effective plan space,'' or {\em CEPS}. The model states a correlation
between the latent variable of the number of plans considered and the
measurable dependent variable of CEPS. The model also states a
positive correlation between the number of operators available in the
DBMS and plan space complexity. Another feature of the model is the
positive correlation between the number of operators available in the
DBMS and the number of operators observed for each functionality. Only
a subset of the physical operators available will actually be
considered depending on the presence of a particular functionality
(such as join) in the query.
Hypothesis 4: Plan space complexity (CEPS) linearly influences suboptimality.

\todo{Add a few sentences about number of operators observed across functionality. }
The types of functionality involved vary based on the type of SQL query. 
For each functionality, the DBMS may include multiple different physical operators
to implement the functionality. The main
types of functionality included in this study involve joins, and aggregates. 
Within joins, we also looked at the impact of self joins. Optimizers have
different operators for each functionality. For example, if the query did 
not involve joins, none of the join operators would be considered and thus
they would not contribute to the complexity for that query. Although we cannot
extract the operators that were considered for proprietary DBMSes. However, 
we can approximate this by observing the operators that were selected 
across different queries sharing functionality. And, that is what we did
to measure operators.

In the characterization of the $x$ axis of Figure~\ref{fig:empirical}
of empirical generalization, we have just {\em described} how
suboptimality might arise, through a theory and its elaborated model.
We now need to move right on that axis, to {\em prediction}.

How might we test such a theory? We hypothesize from our model that as
the optimizer becomes increasingly complex, more occurrences of
suboptimality will be observed, because of such unintended
interactions. Similarly, with a more simplistic optimizer, we
hypothesize that there will be fewer unanticipated interactions, and
so less suboptimality would be present. We predict that one source of
the suboptimality is the designer's focus on correctness and
performance, making it difficult to predict subtle interactions
between components of the optimizer. Later we explore the
ramifications of this hypothesis, finding support for our proposed
model.
\todo{Also add details why our model is necessary before examining data
skew, indexes, etc. }

### 2017-08-08 subopt-tods17/OLDtex/motivation.tex
subopt tods submission

\section{Motivation}
\todo{Rick has modify token}
Consider a simple select-project-join (SPJ) query, with a few attributes in
the SELECT clause, a few tables referenced in the FROM clause,
and a few equality predicates in the WHERE clause. This might be an
excerpt from a more complex query, with the tables being intermediate
results from other computations. We'll use as an example Query 769 from the
study to be described in Section~\ref{sec:detecting}. This query was one of 1000
queries that were randomly generated.

\vspace*{-2ex}
{\small\begin{verbatim}
        SELECT t2.id3 
        FROM ft_HT1 t3, ft_HT3 t2, ft_HT1 t0, ft_HT1 t1 
        WHERE (t3.id2=t2.id1 AND t2.id1=t0.id4 AND t0.id4=t1.id4)
\end{verbatim}
}

\vspace*{-2ex}
\noindent
This query utilizes four correlation names in the FROM clause, three of
which were associated with the same table, {\tt ft\_HT1}, whose cardinality
statistics were varied. Hence this query is on two tables. Due to predicates
in the WHERE clause, the joins between {\tt t0}, {\tt t1}, and {\tt t3} are
in fact self-joins, and with {\tt t2}, an equi-join. The optimizer generates
different plans for this query as the estimated cardinality of the {\tt
  ft\_HT1} table varies.
\shorten{
Assume initially that the cardinalities (actual cardinalities for the simple
query as an isolated query, estimated cardinalities when the query is an
excerpt) are entirely accurate. In this case, the cardinalities of the two
tables named in the FROM clause were 1,000,000 tuples. With this information,
the cost-based query optimizer presumably generates the best plan for this
query. We'll call this the {\em actual plan}, or simply Plan~0.

However, perhaps the cardinality estimates are inaccurate. Perhaps they do
not reflect the number of tuples in a base or intermediate table. With
this inaccurate information, the cost-based query optimizer generates
perhaps the same plan (that would be nice!) or, more likely, a somewhat less
efficient plan when evaluated on the {\em actual} base or intermediate
table. In this particular experiment, all the cardinalities are accurate
except for one of the tables in the from clause, \verb.ft_HT1.. Let's vary the
inaccuracy from exact to increasingly inaccurate by overriding the
cardinality statistics for \verb.ft_HT1. without changing the physical table. We
then examine the sequence of increasingly poorer plans that result.

One would expect that this sequence of plans is stable. Specifically, for
minor inaccuracy, one would expect that the same plan, Plan~0, to be
produced. And indeed, as the cardinality estimate becomes less and less accurate,
ranging from 570,000 tuples to 1,230,000 tuples, the optimizer continues with
Plan~0. At some point, the
inaccuracy passes a threshold and the optimizer chooses a different plan,
Plan~1 at 1,240,000 tuples and Plan~2 at 560,000 tuples.
Later we hit another threshold (here, at 480,000 tuples), and enter into
a new regime, where Plan~3 is chosen.

For some queries, we see an interesting phenomenon. For these queries
(generally different queries for each DBMS), as we increase the inaccuracy,
the optimizer returns to an {\em earlier} plan! Sometimes, in fact, the
optimizer starts oscillating between two plans, sometimes even switching
back and forth when the cardinality estimate changes by a small percentage.
We had no idea that such strange behavior would occur until we came across
it while testing our experiment manager. We call this phenomenon, in which
the query optimizer returns to a previous plan as the inaccuracy increases,
``query optimizer flutter,'' or simply ``flutter.''}  In fact, for this
query, the query optimizer returns to a previous plan as the inaccuracy
increases, which we call ``query optimizer flutter,'' or simply ``flutter.''

\begin{figure*}[tb]
\centering
%originally 30pc
\epsfig{figure=figures/plan769.eps,width=36pc}
\caption{Suboptimality and Fluttering in Query 769\label{fig:query769}}
\end{figure*}

\shorten{This behavior occurs in Query~769, as shown in}In Figure~\ref{fig:query769},
\shorten{which we term a {\em flutter diagram}. Here,} the \hbox{$x$-axis} depicts the
estimated cardinality and the $y$-axis the execution time of the plan on the
actual cardinality, which is 1,000,000 tuples, denoted by $x=10$. So for example Plan~3 is
selected when the optimizer {\em thought} that the cardinality of table {\tt
  ft\_HT1} was 400K rows; when that plan was run on a table with 1M rows, it
was twice as fast as Plan~0, the plan the optimizer selected when the
optimizer had an accurate cardinality estimate of 1M rows.

\shorten{In this flutter diagram, w}We see that as the cardinality is {\em
  over-}estimated, the optimizer switches to Plan~1 (which in fact is better than Plan~0) 
and stays reliably with that plan. As the cardinality 
is {\em under-}estimated, the optimizer switches to a better one, 
Plan~2 (as this plan is in fact faster than Plan~0, when run on a table with
1,000,000 tuples), and then to an even better one, Plan~3, and eventually,
reaches a sub-optimal plan, Plan~4. 
It is as if the optimizer can't make up its mind. This
behavior is somewhat unnerving.

As we will see shortly, flutter and suboptimality are all around us: {\em
  every} \hbox{DBMS} that we have examined, including several open source
\hbox{DBMSes} and several proprietary \shorten{commercial} \hbox{DBMSes},
covering much of the installed base worldwide, exhibit these phenomena, even
when optimizing very simple queries. Hence, such phenomena are not dependent
on a particular implementation of cost-based optimization. Rather, they seem
to be common to {\em any} cost-based optimizer, independent of the specific
cardinality estimation or plan costing or plan enumeration algorithm or
implementation.

This is where the shift to a new methodology comes in. We want to understand
cost-based optimization deeply, in this case to determine if a fundamental
limitation exists, which means that we need to study multiple instances of
that optimization architecture. To do so, we will articulate a model for how
and in what circumstances suboptimality arises and will provide compelling
evidence via hypothesis testing that this model can predict the behavior of
these several \hbox{DBMSes}. This is what we mean by ``empirical
generalization'' and why it is needed to answer such questions.\shorten{ In
  the long history of research in database query optimization, or even of
  databases in general, our model and its hypothesis tests are the first
  predictive results that we are aware of that apply {\em across}
  \hbox{DBMSes}, rather than on a single, specific \hbox{DBMS} or on a
  specific algorithm. This is due to the adoption of a new research
  methodology, that of empirical generalization.}

### 2017-08-08 subopt-tods17/OLDtex/onepass.tex
subopt tods submission

\subsection{{\tt OnePass} Scenario}\label{sec:onepass}

In {\tt OnePass} Scenario, a query is studied for suboptimality by keeping track 
of the {\em states} on the variable-table. 
In other words, we locate different query plans at adjacent cardinalities 
defined by the {\em even} or {\em odd} states, 
and query executions are made at both of the states simultaneously, 
while decreasing the variable-table cardinality in one pass.

Indeed, the adjacent cardinalities are not really called even and odd, 
but they are next to each other. 
They, thus, are simply separated by whether a given cardinality over 
granularity is even or odd. 
For instance, the even state will have 2M, 1.98M, 1.96M, $\cdots$, 1M cardinalities 
in order, while the odd state will take 1.99M, 1.97M, $\cdots$, 1.01M cardinalities 
(decremented by granularity (10K) from the even state's cardinality).

Each state initializes its table(s) same as the variable-table(s). 
In addition, it has the {\em pre-compiled} query plan at its own cardinality 
so that the given query can be executed by the {\em prepared} plan at the cardinality 
without any compiling overhead.

While stepping down from the max to the min cardinality, 
if different query plans are observed at the even and odd states, 
then the prepared query execution at both states is made at the same time 
unless the query has been executed at the cardinality before. 
In this regard, for repeatability we time the same query up to 10 times. 
The timing results at each state are inserted into the labshelf for 
suboptimality analysis. 
For more details about timing single query execution, 
please refer to Section~\ref{sec:timing}.

\begin{figure}[t]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} analyzeQuery($q$, $varTables$):
\STATE $evenState$.$initialize$($varTables$) \\
\STATE $oddState$.$initialize$($varTables$) \\
\FOR{$card$ = 2M to 1M by 10K}
	\IF{$\dfrac{card}{10K}$ is $even$}
		\STATE $evenState$.$card$ $\leftarrow$ $card$ \\
    	\STATE $evenState$.$plan$ $\leftarrow$ $evenState$.$getPreparedQueryPlan$($q$) \\
    \ELSE
 	    \STATE $oddState$.$card$ $\leftarrow$ $card$ \\
	    \STATE $oddState$.$plan$ $\leftarrow$ $oddState$.$getPreparedQueryPlan$($q$) \\
    \ENDIF
    \IF{($evenState$.$plan$ $\neq$ $oddState$.$plan$)}
		\IF{$\neg${$evenState$}.$isExecuted$()}
			\FOR{$i$ $\leftarrow$ 1 {\bf to} 10}
				\STATE $timeSingleQueryExecution$($i$, $q$,  \\
						\hspace{14.0mm}$evenState$.$plan$, $evenState$.$card$) \\
			\ENDFOR
		\ENDIF
		\IF{$\neg${$oddState$}.$isExecuted$()}
			\FOR{$i$ $\leftarrow$ 1 {\bf to} 10}
				\STATE $timeSingleQueryExecution$($i$, $q$,  \\
						\hspace{14.0mm}$oddState$.$plan$, $oddState$.$card$) \\	
			\ENDFOR
		\ENDIF
	\ENDIF
\ENDFOR 
\end{algorithmic}
\caption{Algorithm for Analyzing a Query by {\tt OnePass} Scenario\label{alg:onepass}}
\end{center}
\end{figure}

### 2017-08-08 subopt-tods17/OLDtex/operationalization.tex
subopt tods submission

\section{Operationalization}
\todo{discuss why discontinuity only at lower cardinalities}
\todo{describe experimental setup: number of queries, dbmses, etc}
In this model, we consider plan-operator discontinuity as an independent 
variable in that the occurrences of sub-optimality is not necessarily dependent
on the presence of discontinuous plan-operators, such as a hash-join
operator. However, we predict that when discontinuous plan operators
selected, they may introduce complexity to query optimization especially
at the point where discontinuity is observed. This is because the performance
of such operators is sensitive to the input size. An inaccurate estimation
in the plan statistics may lead the optimizer to keep a plan that
starts to require additional passes over the input data, which could be
significantly less efficient than an alternative plan. Therefore,
we consider such discontinuous plan operators to have positive effects
on the sub-optimality phenomenon.

To study the presence of discontinuous plan operators, more importantly,
to quantify their effects for the later-performed linear regression,
we designed a simple discontinuity-identification technique. We present
an example to illustrate how discontinuity, which is observed in the hash-join
operators for a particular query, is identified.

\begin{figure}[th]\centering
\includegraphics[width=0.40\textwidth]{figures/discontinuity.eps}
\caption{An Example of Discontinuous Plan Operators (Hash-Join)
\label{fig:discontinuity}}
\end{figure}

Figure~\ref{fig:discontinuity} presents for a single query
\shorten{({\em query4}, exhaustive-10q-1, oracle)}, as the cardinality
increase, the {\em cost} of all the hash-join operators utilized in each
plan at a particular cardinality. Each distinct color depicts an individual
hash-join operator.
Note that we ensure the plan-operator equality by examining first the
equality of the entire plans. In other words, if two plans are identical,
then the hash-join operator that appears at the same position in these two
plans are inherently identical. As shown by this figure, the hash-join shown
in cyan has two discontinuous points, both appearing below cardinality
200K. The hash-join in yellow has one discontinuous point.

To systematically identify such discontinuous instances, we employ a simple
slope-based measurement. According to the monotonic characteristic
of a plan operator, it is expected that as the input size increases,
the cost of executing the operator will steadily increase. In the case
of hash-join operator, since its performance is linear to the input size,
we expect that the increase in the execution cost can be characterized by
some slopes that are stable and consistent.
Therefore, by identifying a sudden change in the slope, we can effectively
spot the cardinality at which an operator becomes discontinuous.

Formally, we compute the slope between each pair of consecutive cardinalities
as the following.\\
$S_{card} = (C_{card + 10K} - C_{card}) / 10K$\\

\todo{Sabah will change the following discussion to distribution analysis}
As a result, a series of slope values are computed.
We then compute the standard deviation of all the slope values. Although
there are briefly four classes of operator cost, as indicated by the cyan
hash-join operator in Figure~\ref{fig:discontinuity}, (corresponding to four
groups of slope values), the major variance is introduced by the discontinuous
points. Hence, by identifying the slope values that are greater than one
standard deviation over the average value, the discontinuous can be effectively
discovered. For this particular example, we assign a value {\em 3} to this
variable (number of gaps) since there are three discontinuous points.

\todo{Rick: please read}\\
We operationalize discontinuity in our model by counting and summing the
number of discontinuous operators. Specifically, in the example shown in
Figure~\ref{fig:discontinuity}, two distinct hash-join operators, represented
as the cyan and yellow curves, respectively, are discontinuous. Hence the
value of discontinuity of this query is two. We choose such a
operationalization metric based on the following considerations.

\begin{itemize}
\item{} Due to the sensitivity of the performance of discontinuous plan
operator w.r.t. the operator-cost estimation, such operators can be a major
factor to the presence of suboptimality. Hence, the more discontinuous plan
operators are employed by a query, the higher chance suboptimality will be
observed for this query.

\item{} We do not particularly weight the series of cardinalities, with which
query plans are collected, differently. One might argue that discontinuous
operators that appear at higher cardinalities should have more significant
performance impact than those appearing at lower cardinalities.
We consider an operator to be discontinuous for all the cardinalities, where
this operator is observed, if it is discontinuous for at least once.
Hence the actual values of the cardinalities are not relevant to this metric.

\item{} Furthermore, we do not count the total number of instances of
discontinuity of each plan operator. For instance, the operator represented by
the cyan curve in Figure~\ref{fig:discontinuity} has two discontinuous
instances. The reason is that the frequency of discontinuity for a particular
operator is dependent on a number of factors. For instance, the position
of an operator, whether it is high-up in the plan tree or at the bottom,
may result in more or less occurrences of discontinuity for this operator.
In our study, we consider suboptimality at query plan level rather than within
a plan. In addition, extremely frequent occurrence of discontinuity is not
observable since discontinuity is determined largely by the input-size
estimation of a particular operator. Hence given the gradual change of the
cardinalities in our experiments, no more than three instances for an
individual operator are observed. Therefore, we exclude the in-plan
discontinuity-frequency measurement and focus on the occurrences of
discontinuous operators themselves.
\end{itemize}

\todo{later we will investigate non-linear operators..?}

\todo{The rest of the section below  should go before description of 
discontinuous operators}
An experiment was used to test the model. We selected five relational
DBMSes that are representative of the relational DBMS the market.

The experimental specification configures 600 randomly generated
queries. Each query
is a select-project-join-aggregate query, with a few attributes in the
SELECT clause, a few tables referenced in the FROM clause, a few
equality predicates in the WHERE clause, and zero or more aggregate
functions. As such some of the complexities
mentioned by Kabra and DeWitt~\cite{kabra98}, such as
user-defined data types, methods, and operators, are not considered.
The queries were generated
by a simple algorithm, the pseudo-code for which is shown in
Figure~\ref{alg:querygen}.

The query generator has \todo{is it four? Rui/Matt, could you please put
in the function names for Aggregates} 
main components, which are {\tt
  generateSelectClause()}, {\tt generateFromClause()}, {\tt
  generateWhereClause()} and {\tt generateGroupByClause()}. 
\todo{Q: does primary key have its own function?}
The select clause will contain from one to four
columns\c2j{}{, hence, an average of 2.5 columns}. The number of
correlation names
in the FROM clause varied from one to four, with duplication of tables
allowed (duplicate table names within a from clause implies a self-join).
In the queries that were generated, from one to four unique tables were
mentioned in the FROM clause. Somewhat fewer tables were mentioned than the
number of correlation names, as the presence of self-joins reduced the
number of unique tables named by the queries.

In {\tt generateWhereClause()}, in which all the joins will appear, we make
sure that Cartesian products are
eliminated\c2j{.}{(\verb.cartesianPossible="false".)}  We do this by
connecting all the correlation names that appear in the FROM statement via
comparisons on random columns. In this case, the comparisons are all
equality operators. To ensure that the queries are as simple as possible, we
do not include any additional predicates in the WHERE clause.  \c2j{}{This
  is realized by setting the attributes {\tt maxIsAbsolute} to {\tt true}
  and {\tt complexUsePercentage} to {\tt 100}.  Basically, ``complex''
  predicates eliminate the Cartesian product, and by setting complex
  predicates as ``absolute,'' no additional predicates are included except
  for those which are necessary for eliminating Cartesian product.} Also for
simplicity, we do not include disjunctions nor negations.

In the \todo{Correct name: generateGroupByClause}, where the aggregate
functions and
grouping occurs, the presence or absence of a group by clause is
recorded. We expect the
complexity of the query to increase when the first aggregate function
is introduced. However,
additional aggregate functions in the same SELECT clause will not add
significant complexity since
they share the same grouping function.

We are interested in predicting the behavior of DBMSes through our
model. Therefore, we vary the cardinality from 2M (maximum) to
100K (minimum), in increments of 10K. For one of the DBMSes that was
slower than the others and was timing out for the majority of the
queries
when run between 100K and 2M, we reduced the size of the tables and
varied the cardinality from 200K (maximum) to 10K (minimum), in
increments of 1K.

We identify the cardinalities where the plan selected changed (change
point) in a single iteration increment.
At the cardinality before the change (higher) and after (lower), we
run and time the query executions 10 times each.
We reduce the runtime by timing of other processes (identified by
psdiff) and remove any executions that had phantoms or stopped processes.
We then compare the average runtimes
and standard deviations at the cardinality just before the change
point (n-1) and at the change point (n). The query is said to be suboptimal
$if avg_{n-1} + 0.5 * stddev_{n-1} <= avg_n - 0.5 * stddev_n$.

\todo{Rick, please check}Suboptimality is coded as follows:
a) Not suboptimal, i.e., $0 IF avg_{n-1} + 0.5 * stddev_{n-1} > avg_n -
0.5 * stddev_n ;$
b) Low suboptimality, i.e., $1 IF avg_{n-1} + 0.5 * stddev_{n-1} <=
avg_n - 0.5 * stddev_n$, and
$avg_{n-1} + 1.0 * stddev_{n-1} > avg_n - 1.0 * stddev_n$;
c) Moderate suboptimality, i.e., $2 IF avg_{n-1} + 1.0 * stddev_{n-1}
<= avg_n - 1.0 * stddev_n$, and
$avg_{n-1} + 1.5 * stddev_{n-1} > avg_n - 1.5 * stddev_n$;
b) High suboptimality, i.e., $3 IF avg_{n-1} + 1.5 * stddev_{n-1} <=
avg_n - 1.5 * stddev_n$

One of the constructs that
our model predicts ``number of operators in the DBMS'' is a factor
that will impact suboptimality. By number of operators,
we mean the number of operators available for selection, projection,
join and aggregate functions. As the number
of operators available increases, the complexity of the optimizer
increases because it has to choose between more
operators, and hence the probability of suboptimality increases.

Not all operators available in a DBMS are considered by the optimizer.
For example, for joins, nested loop is rarely considered
if there are other operators available for joins. However, in
proprietary DBMSes, we cannot capture the latent variable number of
operators
considered. Therefore, we try to approximate this by counting ``Number
of operators observed'' chosen across plans for each query.

The overall effect of the primary key moderator is to reduce the strength 
of the relationship between the number of correlation names in from to 
CEPS and subopt and from presence of aggregates to CEPS/supopt.

More specifically, if primary key is not on 
a) any of the joining columns for tables (or there is only one table 
involved in the query), then presence of primary keys will NOT have any 
impact on the relationship of ``\# of correlation names in FROM'' (OR from 
``\# of self joins'') to any other variable (No moderating effect)

b) any of the Grouping columns in aggregate function and is not on any 
columns involved in the joins, then presence of primary keys will NOT 
have any impact on the relationship from "Presence of aggregates" to 
any other variable.

c) any of the Grouping columns in aggregate function but is on the 
columns involved on the joins then the presence of primary key will 
not have any moderating effect on the relationships from presence of 
aggregates to CEPS/supopt

If primary key is on 
d) one side of a join condition, for a subset of the joins, the 
primary key will have less of an impact on the relationship between 
"\# of correlation names in FROM" and Suboptimality compared to when 
all of the columns had a primary index defined on them.

e) one side of a join condition, for all the joins, the primary key 
will have more of an impact on the relationship between "\# of correlation 
names in FROM" compared to when some of the columns had a primary index 
defined on them (on one side of the join condition).

f) both sides of a join condition, for a subset of the joins, the primary 
key will have more of an impact on the relationship between "\# of 
correlation names in FROM" and Suboptimality compared to when none of 
the columns had a primary index defined on them and more compared to d) 
primary key is on one side of the join condition, for a subset of the 
joins (assuming that d) was similar in number of total joins, and had a 
similar number of joins involving primary keys).

g) both side of a join condition, for all of the joins, \# of correlation 
names will not affect plan space complexity / suboptimality.

h) If primary key is on a subset of SELECT /GROUP BY columns in aggregate, 
then the presence of primary keys will have more of an impact on the 
relationship "Presence of aggregates" to Suboptimality vs. no primary 
key on any of the columns in GROUP BY clause.

i) If primary key is on all the columns of SELECT /GROUP BY columns in 
aggregate, then presence of primary keys will have a very strong 
moderating effect: "Presence of aggregates" will have no impact on 
Suboptimality.

The aggregate is generally done after all the joins are completed; 
therefore, the presence of primary keys on non-GROUP BY columns should 
not impact the relationship between presence of aggregates and suboptimality.

\shorten{To estimate the error a DBMS makes in estimating the cost of a plan,
we use a variable called ``Flutter'' that counts the number of times a
plan repeats
in non contiguous blocks \todo{Rick, please check}
Flutter per query = (Max(Count(planID involved in change point)) - 1}

### 2017-08-08 subopt-tods17/OLDtex/prevalence.tex
subopt tods submission

\section{Prevalence of Suboptimality}
\todo{Sabah has modify token: rewrite according to the model}
\todo{path analysis}
As explained in Section~\ref{sec:datasets}, we looked
for query suboptimality among 500 randomly generated, quite simple queries,
for which the actual cardinality was known (the experiment scenario updated
the statistics before each plan was optimized). We observed that 266 of
these queries, or 53.2\%, were suboptimal.  This percentage comes with several
provisos. First, there were no foreign key constraints nor indexes, which in
some ways renders the optimizer's job harder.  Perhaps with
queries over data with more comprehensive schema available, the percentage
would be less. Second, for half of the suboptimal queries, the optimal plan
not chosen was within 30\% of the optimal plan (see
Figure~\ref{fig:time}). That said, that also means that there were a good
number of queries for which the selected plan was quite a bit off (with a
few running twice as long as the best plan for that query).

\begin{figure}[bth]\centering
\includegraphics[width=0.50\textwidth]{figures/runtime_diff.eps}
\caption{Cumulative percentage of queries differing from the optimal plan by
  the indicated percentage of time\label{fig:time}}
\end{figure}

It is also interesting to look at the {\em width} of suboptimality: the
range of cardinality for which the optimizer {\em didn't} choose the optimal
plan. This width is defined to be the interval
of suboptimality starting from the cardinality immediately at the first
faster plan, extending across the actual cardinality of 1M tuples, to the
end of the cardinality of the last faster plan.

If the query optimizer is well behaved, one would might expect that the
optimal plan is relatively close, cardinality-wise, from the plan chosen by
the query optimizer, and this is indeed the case. Figure~\ref{fig:width} shows
the cumulative percentage of suboptimal queries exhibiting a width greater
than or equal to the width shown on the $x$-axis, expressed in cardinality
(number of tuples, in steps of 100K tuples), as specified by the {\tt
  searchGranularity} attribute in the {\tt variable\-TableSet} element of the
XML specification in Figure~\ref{fig:expspec}. The assumption of a
relatively close faster plan implies a concave cumulative distribution, as
was observed.

There is a big jump at 990,000, with this width accounting for about a third
of the queries. This occurs when there are two plans that are structurally
similar (as discussed later in Section~\ref{sec:ceps}) but have different
execution times, when measured at cardinalities of 1M and
2M (that is, at the {\em actual} and {\em maximum} cardinalities, for which
the experiment always obtains an execution time). By definition, this
particular circumstance translates to a width of $1\hbox{M} - 10\hbox{K} =
990\hbox{K}$.

\begin{figure}[bth]\centering
\includegraphics[width=0.50\textwidth]{figures/width_cumulative.eps}
\caption{Cumulative percentage of queries by
  the indicated width\label{fig:width}}
\end{figure}

\c2j{}{

\begin{table}[t]
\begin{center}
\begin{tabular}{c|c}
{\em DBMS} & {\em Percentage of Suboptimality}\\
\hline
{\bf A} & X\%\\
{\bf B} & X\%\\
{\bf C} & X\%\\
{\bf D} & X\%
\end{tabular}
\end{center}
\caption{Percentage of Suboptimality by DBMS\label{tab:dbmsprev}}
\end{table}

Second, suboptimality is highly variable, ranging from {\todo: X} to {\todo: Y}
queries (out of 100) for DBMS~{\todo: X}.

Third, even for very simple queries, suboptimality is quite prevalent.
Across all DBMSes, the prevalence is about {\todo: X}\%: {\todo: Y} of all
simple queries exhibit suboptimality. So it is likely that most extant
database applications include one or more queries that suboptimality.

\begin{table}
\begin{center}
\begin{tabular}{c|c|c}
{\em Number of} & {\em Number of} &{\em Percentage of}\\
{\em DBMS(es)} & {\em Suboptimal Queries} & {\em Suboptimality} \\
\hline
4 & X & X\% \\
3 & X & X\% \\
2 & X & X\% \\
1 & X & X\% \\
\end{tabular}
\end{center}
\caption{Prevalence of Suboptimality Across DBMS(es)\label{tab:supoptacross}}
\end{table}

Fourth, suboptimality is observed on different queries for different DBMSes.
Table~\ref{tab:supoptacross} 
summarizes the common suboptimal queries across multiple \hbox{DBMSes}.
We notice that {\todo: some phenomenon} 
all four \hbox{DBMSes}.
}

\section{Testing the Predictive Model}
\section{Sabah, change this also}
Our theory, that unanticipated interactions within the optimizer cause
suboptimality, coupled with the associated model in Figure~\ref{fig:model}, can be used
to derive specific predictions.  Simply put, we predict that as optimizer
complexity increases, so will suboptimality prevalence. However, as that is
a latent variable, not subject to direct control nor measurement,
we need to operationalize the independent variable of
optimizer complexity. Here we examine two such operationalizations of optimizer
complexity that we {\em could} control for, to render optimizer complexity
an independent variable.

\subsection{Impact on Query Parameters}
We first examine the relationship between
the prevalence of suboptimality and various aspects of the queries. For this
purpose, we considered the query parameters that are under our control: the
number of columns in the SELECT list, the number of correlation names in the
FROM list, and the number of self-joins in the WHERE clause. Our model
relates the query complexity to the number of plans. Of these independent
variables, only the number of correlation names in the FROM list will impact
the number of plans.

Our theory predicts that the higher the complexity of the optimization, that
is, the more correlation names in the FROM clause,
the higher the chance that the query will be suboptimal.

\begin{quote}
{\bf Hypothesis H1}: Suboptimality prevalence will correlate positively
with the number of correlation \hbox{tuple} variables used in the query.
\end{quote}

\begin{table}[t]
\begin{center}
\begin{tabular}{c|c|c|r}
{\em Number of} & {\em Number of} & {\em Number of} & \multicolumn{1}{|c}{\em Suboptimality}\\
{\em Corr Names}&{\em Queries}&{\em Suboptimal}&{\em Prevalence}\\
{\em in} FROM&&Queries&\\
\hline
1 & 113 & 5 & 4\%\\
2 & 136 & 57 & 42\%\\
3 & 117 & 91 & 78\%\\
4 & 134 & 113 & 84\%
\end{tabular}
\end{center}
\caption{Prevalence of Suboptimality by Number of Correlation Names in
FROM\label{tab:numcnfrm}}
\end{table}

As indicated in Table~\ref{tab:numcnfrm}, our experiment shows that the percentage of suboptimality grows
monotonically as the number of correlation names in the FROM clause
increases. \c2j{}{When looking at each individual experimental DBMS separately,
this trend is still present (see Figure~\ref{fig:numcnfrm}). (Note that the
number of columns were uniformly chosen to be 1 to 4, which implies a mean
of about 2.5 in the second column. For DBMS~{\bf C}, the number of columns
was fixed at 2, 4, 6, 8, and 10, for a mean of about 6.)} \c2j{A two-sample
  \hbox{t-test} assuming unequal variances shows that the mean
  number of correlation names for optimal
  queries, 1.8, as compared with that for suboptimal queries, 3.1, rejects
  the null hypothesis that they are drawn from the same distribution, with a
  {\em p-value} of $p<0.001$.}{The t-test results
in Table~\ref{tab:ttests}(c) reject the null hypothesis in all cases, 
as predicted by our theory. (We note that the query generation algorithm 
ensures that the number of predicates is the same as the number of 
correlation names in FROM clause. This might have played a role.)

\begin{figure}[t]
\begin{center}
\comment{\epsfig{file=figures/subopt_from/subopt_from.eps, width = 20pc}}
\end{center}
\caption{Prevalence of Suboptimality by the Number of Correlation Names in
FROM, by DBMS\label{fig:numcnfrm}}
\end{figure}

\begin{table}[t]
\begin{center}
\begin{tabular}{c|c|c|r}
{\em Number of} & {\em Number of} & {\em Number of} & \multicolumn{1}{|c}{\em Suboptimality}\\
{\em Corr Names}&{\em Queries}&{\em Suboptimality}&{\em Prevalence}\\
{\em in} FROM&&&\\
\hline
1 & 21 & 0 & 0\%\\
2 & 27 & 0 & 0\%\\
3 & 26 & 6 & 23\%\\
4 & 26 & 17 & 65\%
\end{tabular}

\vspace*{1em}

{\bf (b)} DBMS {\bf A} VaryStats Scenario

\vspace*{1em}

\begin{tabular}{c|c|c|r}
{\em Number of} & {\em Number of} & {\em Number of} & \multicolumn{1}{|c}{\em Suboptimality}\\
{\em Corr Names}&{\em Queries}&{\em Suboptimality}&{\em Prevalence}\\
{\em in} FROM&&&\\
\hline
1 & 21 & 0 & 0\%\\
2 & 27 & 5 & 19\%\\
3 & 26 & 7 & 27\%\\
4 & 26 & 3 & 12\%
\end{tabular}

\vspace*{1em}

{\bf (b)} DBMS {\bf A} Upper Scenario

\vspace*{1em}

\end{center}
\caption{Prevalence of Suboptimality by Number of Correlation Names in
FROM\label{tab:numcnfrm}}
\end{table}

\begin{table}[t]
\begin{center}
\begin{tabular}{c|c|c|c|l}
{\em DBMS}/ & {\em Mean} & {\em Mean} & {\em Statistic} &
{\em P value} \\
{\em Scenario}&{\em Optimality} & {\em Suboptimality} &\\
\hline
{\bf A} VaryStats&2.22&3.74&$p < 0.001$\\
{\bf A} Upper&2.52&2.87&$p< 0.001$\\
{\bf B}&X&X&X\\
{\bf C}&X&X&X&X\\
Overall&X&X&X&X\\
\end{tabular}
\end{center}
\caption{t-test Results for the Number of Correlation Names in the FROM Clause\label{tab:ttests2}}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{c|c|c|r}
{\em Number of} & {\em Number of} & {\em Number of} & \multicolumn{1}{|c}{\em Suboptimality}\\
{\em Columns}&{\em Queries}&{\em Suboptimality}&{\em Prevalence}\\
{\em in} Select&&&\\
\hline
1 & X & X & X\%\\
2 & X & X & X\%\\
3 & X & X & X\%\\
4 & X & X & X\%
\end{tabular}
\end{center}
\caption{Prevalence of Suboptimality by the Number of Columns in
SELECT\label{tab:numcnsel}}
\end{table}

The impact of the number of self-joins on the query optimization complexity, 
and thus, on the suboptimality prevalence as predicted by our theory,
is unclear. Hence, our theory is silent on the effect. {\todo: verification is
needed.}

As indicated by Table~\ref{tab:numselfjoin}, although as the number of
self-joins grows from zero to two, {\todo: complete with new observation}
%the percentage of flutters grows, it
%drops when number of self-join is three. In our 1000-queries data set,
%there are only six (=24/4, as there are four identical 1000-query sets) 
%queries with three self-joins, thus there isn't a large
%enough sample space to draw conclusions in this case. We do note that for
%all DBMSes the null hypothesis for number of self-joins is rejected. 
%It is unclear whether this is due to an increase in query optimization complexity, 
%or is due to a confounding increase in the number of correlation names (if so, 
%our theory would predict an increase in flutter), or is due to some other factor.

\begin{figure}
\begin{center}
\comment{\epsfig{file=figures/subopt_selfjoin/subopt_selfjoin.eps, width = 20pc}}
\end{center}
\caption{Prevalence of Suboptimality by Number of Self-Joins, by
DBMS\label{fig:numselfjoin}}
\end{figure}

\begin{table}
\begin{center}
\begin{tabular}{c|c|c|r}
{\em Number of} & {\em Number of} & {\em Number of} & \multicolumn{1}{|c}{\em Suboptimality}\\
{\em Self-Joins}&{\em Queries}&{\em Suboptimality}&{\em Prevalence}\\
\hline
0 & X & X & X\%\\
1 & X & X & X\%\\
2 & X & X & X\%\\
3 & X & X & X\%

\end{tabular}
\end{center}
\caption{Prevalence of Suboptimality by Number of
Self-Joins\label{tab:numselfjoin}}
\end{table}

As discussed in Section~\ref{sec:setup}, many of the queries included
self-joins, and about a third included two or more correlation names over the
independent variable table. Does varying the size of a self-joined table
induce more or less occurrences of suboptimality? Again, our theory is silent
on this. We performed a chi-squared test to see if the distribution of
suboptimality in the presence of a self-join on the varying
table was independent from the distribution of suboptimality in the presence
of a self-join not on the varying table. We found that the null hypothesis was 
rejected in this case ($\chi^2(1)=X,\,p<Y$). Interestingly, 
as shown by Table~\ref{tab:corSelfVar}, the prevalence of suboptimality
decreased when there was a self-join on the variable table.
{\todo: verification}

\begin{table}
\begin{center}
\begin{tabular}{c|c|c|c}
{\em Self-Join} & {\em Optimality} & {\em Suboptimality} & {\em Suboptimality Prevalence}\\
\hline
{\em Non-Variable} & X & X & X\% \\
{\em Variable} & X & X & X\%

\end{tabular}
\end{center}
\caption{Correlation Between Existence of Self-Joins on the Variable Table and Prevalence of Suboptimality\label{tab:corSelfVar}}
\end{table}
}

\c2j{}{
\subsection{Number of Operators}
One measure of optimizer complexity is the number of operators used by a
DBMS. The thinking here is that optimizers that have a lot of alternatives
at their disposal for common algebraic operators (scans and joins) will have
more inherent complexity in choosing among those operators.

\begin{quote}
{\bf Hypothesis H3}: Suboptimality prevalence will correlate positively
with the number of operators generated by the DBMS.
\end{quote}

We analyzed all the plans and accumulated the operators used by each DBMS;
the result is shown in Table~\ref{tab:ops}. Here the correlation is quite
striking. In fact, it appears that suboptimality percentage might be 
correlated with the square of number of operators, which makes some sense,
because optimizer interactions might be related to inter-operator
combinations.{\todo: verification}

\begin{table}
\begin{center}
\begin{tabular}{c|c|c|l}
{\em DBMS} & {\em Number of} &{\em Suboptimality} & {\em Operators}\\
& {\em Operators} && \\
\hline
{\bf C} & X & X\% &NestedLoop\\
{\bf A} & X & X\% &SeqScan,HashJoin\\
{\bf D} & X & X\% &SeqScan,HashJoin,\\
&&&Sort,MergeJoin\\
{\bf B} & X & X\% & SeqScan,\\
&&&Hash,HashJoin,\\
&&&Sort,MergeJoin,\\
&&&NestedLoop\\
\end{tabular}
\end{center}
\caption{Prevalence of Suboptimality by Number of Operators\label{tab:ops}}
\end{table}
}

\subsection{Effective Plan Space}\label{sec:ceps}
\c2j{One latent variable that is linked to optimizer complexity is the
  size of the plan space considered by the optimizer for each specific query.}{The number of operators a DBMS uses is a quite coarse independent variable
for optimizer complexity. Perhaps the size of the plan space considered by
the optimizer for each specific query is a better measure.} An optimizer that
can enumerate many plans (from a collection of many available operators) for
a query would expose that query to more complex interactions between
optimizer rules than a query for which the optimizer that enumerated just a
few plans, using a small set of operators.

While we did not have access to the number of enumerated plans (DBMSes do
not report such information), we do know how many unique plans were
generated for each query as we varied the estimated cardinality of one of
the tables. We consider two plans to be
different if they have different {\em structure}, that is, a different join
order, a different tree structure, or use a different physical operator
somewhere in the plan. We term the set of plans that were actually generated the {\em
  effective plan space}\c2j{ and}{; for Query~{\bf X} in
Figure~\ref{fig:subopt_instance}, that set is 
$\{$Plan 0, Plan 1, Plan 2, Plan 3, Plan 4$\}$.  We} term the size of this set
the {\em cardinality of the effective plan space}, or {\em CEPS}. Our
refined hypothesis predicts that queries that utilize more of the query
optimizer, that is, have a higher CEPS, will exhibit more suboptimality.
(We note here the pioneering work in the Picasso project,
which graphically depicted the plans chosen for
different cardinalities~\cite{harish07,reddy05},
thereby providing a visual representation of CEPS for a query.)

\begin{quote}
{\bf Hypothesis \c2j{H2}{H4}}: Suboptimality prevalence will correlate positively
with the cardinality of the effective plan space.
\end{quote}

\begin{table}[t]
\begin{center}
\begin{tabular}{c|c|c|r}
{\em CEPS} & {\em Number of} & {\em Number of} & \multicolumn{1}{|c}{\em Suboptimality}\\
&{\em Queries} & {\em Suboptimal} &  \multicolumn{1}{|c}{\em Prevalence}\\
&&{\em Queries} & \\
\hline
1 & 157 & 34 & 22\%\\
2 & 117 & 45 & 38\%\\
3 & 74 &58 & 78\%\\
4 & 73 & 62 & 85\%\\
5 & 50& 41 & 82\%\\
6 & 22& 20 & 91\%\\
7 & 5& 5 & 100\%\\
8 & 2& 1 & 50\%\\
\end{tabular}
\end{center}
\caption{Prevalence of Suboptimality by CEPS\label{tab:ceps}}
\end{table}

\c2j{}{\begin{table}[t]
\begin{center}
\begin{tabular}{c|c|c|c|r}
{\em DBMS} & {\em CEPS} & {\em Number of} & {\em Number of} & \multicolumn{1}{|c}{\em Suboptimality}\\
&&{\em Queries} & {\em Suboptimality} &  \multicolumn{1}{|c}{\em Prevalence}\\
\hline
{\bf A} & X & X & X & X\%\\
{\bf A} & X & X & X & X\%\\

{\bf B} & X & X & X & X\%\\
{\bf B} & X & X & X & X\%\\

{\bf C} & X & X & X & X\%\\
{\bf C} & X & X & X & X\%\\

{\bf D} & X & X & X & X\%\\
{\bf D} & X & X & X & X\%\\
\end{tabular}
\end{center}
\caption{Prevalence of Suboptimality by CEPS, per DBMS\label{tab:ceps}}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{c|c|c|c|r}
{\em Number of} &{\em CEPS} & {\em Number of} & {\em Number of} & \multicolumn{1}{|c}{\em Suboptmality}\\
{\em Tables}&&{\em Queries} & {\em Suboptmality} &  \multicolumn{1}{|c}{\em Prevalence}\\
\hline
2 & X &X & X & X\%\\
2 & X &X & X & X\%\\

4 & X &X & X & X\%\\
4 & X &X & X & X\%\\

6 & X &X & X & X\%\\
6 & X &X & X & X\%\\

8 & X &X & X & X\%\\
8 & X &X & X & X\%\\

10 & X &X & X & X\%\\
10 & X &X & X & X\%\\
\end{tabular}
\end{center}
\caption{Prevalence of Suboptimality by CEPS, for DBMS~{\bf C}\label{tab:cepsC}}
\end{table}
}
We clustered the queries by their CEPS and computed the amount of
suboptimality. Table~\ref{tab:ceps} shows the results. There is a row for
each value of CEPS. The first row is for CEPS${}=1$, that is, the situation
where the
optimizer returned the same structure query evaluation plan across all
cardinalities. (We note that plans can differ by other factors not indicated
by {\tt EXPLAIN PLAN}; these plans can thus exhibit different execution
times. That is why even with CEPS${}=1$ there are suboptimal queries:
two plans that are structurally equivalent but differ in other ways had
different times, which are included in those plans that we saw at the top
right of Figure~\ref{fig:width}.)

This table indicates a quite striking positive correlation
between suboptimality and CEPS. The prevalence increases rather
smoothly except for a very few outliers.

We compared the mean of the CEPS for the
optimal queries with the mean for the suboptimal queries, using a
two-sample \hbox{t-test} assuming unequal variances; \c2j{the analysis shows
  that the mean CEPS for optimal queries, 1.8, as compared with that for
  suboptimal queries, 3.4, again rejects
  the null hypothesis that they are drawn from the same distribution, with 
  $p<0.001$.}{the results are shown in
Table~\ref{tab:ttests}. We looked at all DBMSes, as well as the union of 
the data for these three DBMSes, termed
``Overall'' in the results. For the third row  of Table~\ref{tab:ttests}, 
we don't report results from the original data set for 
DBMS~{\bf C} as that includes only two suboptimal queries; instead,
in that row we report results from the data described
in Table~\ref{tab:cepsC}, to be described shortly.  The null hypothesis, that the CEPS for
optimal queries does not differ from the CEPS for suboptimal queries,
is convincingly rejected for all four analysis. {\todo: verification}}

\c2j{}{\begin{table}[t]
\begin{center}
\begin{tabular}{c|c|c|c|l}
{\em DBMS} & {\em Mean} & {\em Mean} & {\em Statistic} &
{\em P value} \\
&{\em Optimal} & {\em Suboptimal} &\\
\hline
{\bf A} & X & X & X & X\\
{\bf B} & X & X & X & X\\
{\bf C} & X & X & X & X\\
{\bf D} & X & X & X & X\\
Overall & X & X & X & X\\
\end{tabular}

\end{center}
\caption{t-test Results for CEPS\label{tab:ttests}}
\end{table}

Let's examine Table~\ref{tab:ceps} more closely. Suboptimality obviously
doesn't occur for CEPS=1, because one plan can't be suboptimal. Suboptimality
appears for CEPS=2 only in DBMS~{\bf B}, and then, at low prevalence.
{\todo: verification}

The prevalence generally increases rather smoothly for all DBMSes
except for a few outliers. DBMS {\bf A} and {\bf D} both demonstrate increasing trend as
CEPS grows from two to six and five, respectively. The prevalence of 
suboptimality then decreases slightly. We consider this decreasing trend is due to first, 
small sample size, e.g.,~only five queries in the result from DBMS {\bf A} 
have eight plans involved. Second, when there are more plans generated, the 
chance of getting these plans to repeat is lower. To check that conjecture, we tried
just that DBMS on 10,000 queries, each with exactly four correlation names in the from
clause. The result, shown in Figure~\ref{fig:ceps10K}, demonstrates a general upward trend.
}
\c2j{}{
The relationship between suboptimality and CEPS may explain the very low 
prevalence of suboptimality for\linebreak DBMS~{\bf C}. From
the previous section, we see that this DBMS used very few operators. Thus
the plans varied only in terms of the join order. For queries with one
correlation name, there is necessarily only one plan; for queries with two
correlation names there are only two possible plans. There were only two
queries for which this DBMS was able to even come up with four plans, so the
opportunity for suboptimal was minimal.
 {\todo: rewrite}

{\todo: no time}
\begin{figure}
\begin{center}
\comment{\epsfig{figure=vldb_figures/figure4/figure4.eps,width=20pc}}
\end{center}
\caption{Prevalence of Flutter by CEPS, for DBMS~{\bf A}, for 10,000
  Queries, For Exactly Four Correlation Names\label{fig:ceps10K}}
\end{figure}

To explore this further, we ran a series of experiments, each with 1000
queries on DBMS~{\bf C}, on the data set described in Section~\ref{sec:setup}, over four tables at four and fixing the
number of correlation names to
2, 4, 6, 8, and 10, with the results shown in Table~\ref{tab:cepsC}. Here
we see a slightly smoother transition, though it seems to jump very quickly
to 100\% as CEPS increases.

\begin{quote}
{\bf Hypothesis H5} Increasing the number of operators used in a plan will
increase the {\tt width} of sub-optimality.
\end{quote}
}

It is important to note the different roles that the number of correlation
names and CEPS plays in the causal model in Figure~\ref{fig:model}.
This model contains two sources of suboptimality: optimizer complexity and
inadequacies of the cost model. While the independent variable of query
complexity (here, the number of correlation names) affects the impact of
both sources, CEPS only impacts the effect of optimizer complexity on
suboptimality. That hypothesis {\bf H2} is supported indicates that
optimizer complexity plays at least some role, independent of cost model, in
suboptimality.

\subsection{Width of Suboptimality}
\todo{Do we even want to include this?}
How should the width of suboptimality be affected by query complexity and optimizer
complexity? Our causal model predicts that as the query becomes more
complex, the number of components and plans involved will be greater,
thereby having more unanticipated interactions, with the result that a
faster plan may occur closer to the actual plan, thereby reducing the width
of suboptimality.

\begin{quote}
{\bf Hypothesis \c2j{H3}{H6}} Increasing the query complexity will {\bf decrease}
the width of sub-optimality.
\end{quote}

Table~\ref{tab:widthfrom} shows such a negative correlation.

\begin{table}[t]
\begin{center}
\begin{tabular}{c|c|c}
{\em Number of} & {\em Number of} & {\em Average}\\
{\em Corr Names} & {\em Suboptimal} &{\em Width}\\
{\em in FROM} &{\em Queries} & \\
\hline
1 & 5 & 990,000\\
2 & 57 & 930,000\\
3 & 91 & 480,000\\
4 & 113 & 440,000\\
\end{tabular}
\end{center}
\caption{Average Width of Suboptimality by Query Complexity\label{tab:widthfrom}}
\end{table}

Considering optimizer complexity, we have a proxy in CEPS. The causal
model predicts that as the CEPS increases, there will be more
unanticipated interactions, and thus the faster plan will again occur closer
to the actual plan, reducing the width.

\begin{quote}
{\bf Hypothesis \c2j{H4}{H7}} Increasing the size of the effective plan space will
{\bf decrease} the width of sub-optimality.
\end{quote}

As Table~\ref{tab:widthceps} shows, while there is generally a negative
correlation as CEPS increases, there is a slight tick at CEPS${} = 4$ and
anomalies for a CEPS of 7 and 8, though admittedly there is insufficient
data at these last two points to draw any conclusions about average width at
high CEPS.

\begin{table}[t]
\begin{center}
\begin{tabular}{c|c|c}
{\em CEPS} & {\em Number of} & {\em Average}\\
& {\em Suboptimal} & {\em Width}\\
&{\em Queries} & \\
\hline
1 & 34 & 990,000\\
2 & 45 & 810,000\\
3 & 58 & 440,000\\
4 & 62 & 460,000\\
5 & 41 & 410,000\\
6 & 20 & 350,000\\
7 & 5 &  650,000\\
8 & 1 & 750,000\\
\end{tabular}
\end{center}
\caption{Average Width of Suboptimality by CEPS\label{tab:widthceps}}
\end{table}

The model predicts that these hypotheses will hold for different DBMSes and
for a wider range of query and data complexity. It also suggests other
hypotheses concerning the impact on suboptimality of non-linearity of the
cost formulas and of error in the estimated cost of a produced plan, which
can be empirically investigated across DBMSes, query complexity, and data
complexity.

%\section{Prevalence of Sub-optimality}
%When sub-optimal plan occurs, one immediate question is how bad is it?
%Is the sub-optimal plan, indeed, unacceptable?
%It is obvious to first of all study how bad the sub-optimal situation is.

%1. The {\bf difference in percentage} between optimal execution time and the
%actual.
%2. The {\bf width} of sub-optimal plan
%Figure

%\begin{figure}[htb]
%\begin{center}
%\epsfig{figure=vldb_figures/figure5/figure5.eps,width=20pc}
%\end{center}
%\caption{Flutter Length Distribution for DBMSes~{\bf A} and {\bf D}\label{fig:length}}
%\end{figure}

%\begin{figure}[htb]
%\begin{center}
%\epsfig{figure=vldb_figures/figure5/figure5b.eps,width=20pc}
%\end{center}
%\caption{Flutter Length Distribution for DBMS~{\bf B}\label{fig:lengthb}}
%\end{figure}

%As a final characterization, we look at {\em when} flutter first appears, as
%the estimated cardinality is varied. In Figure~\ref{fig:start}, we show the
%distribution of the initial cardinality for each plan, by DBMS.
%This data, from DBMS~{\bf A} and DBMS~{\bf D}, is somewhat symmetric. We
%emphasize that these are two large and complex query optimizers written by
%different teams of programmers. It may be that there is something important
%going on here, though more data is needed before a coherent,
%predictive explanation is forthcoming.

%\begin{figure}[t]
%\begin{center}
%\epsfig{figure=vldb_figures/figure6/figure6.eps,width=20pc}
%\end{center}
%\caption{Distribution of the Start of Flutter for DBMSes~{\bf A} and {\bf D}\label{fig:start}}
%\end{figure}

%\begin{figure}[t]
%\begin{center}
%\epsfig{figure=vldb_figures/figure6/figure6b.eps,width=20pc}
%\end{center}
%\caption{Distribution of the Start of Flutter for DBMS~{\bf B}\label{fig:startb}}
%\end{figure}

%For DBMS~{\bf B}, as shown in Figure~\ref{fig:startb}, 22.8\% of the flutters start at the beginning 
%(10,000 tuples), with the percentage dropping as the cardinality increases. All the
%flutters start before the actual optimal cardinality, which is 1,000,000. This
%unusual behavior is related to the high flutter length (cf. Figure~\ref{fig:start}): the minimum number
%of length of a fluttered plan is 19 while the maximum is 60. 
%We believe that as we reduce the sampling granularity, a higher number
%of re-occurrences will be observed.

%\subsection{Optimizer Stability}
%When optimal situations were observed, we wanted to study
%the stability of optimizer. That is, whether the optimizer would change to a
%different (slower) plan immediately after a small change in the cardinality
%estimation, or the optimal plan would last for a while. In order to study this
%issue more intuitively, we defined the following four types of behavior
%(shape), as shown in figure~\ref{_figure}. Moreover, to quantify the
%stability of the optimizers, we defined {\tt CEI (Cardinality Estimation
%Instability)} formula to compute the score for the shapes.
%$score = \Sigma(1-|d|)*y$
%The idea behind the computation was that we want to focus on the area near
%the actual cardinality by assign higher weight values to the area, according
%to the distance to the actual cardinality. So if the optimal plan stays for
%a while on both sides of the actual cardinality, the score would be lower
%than the shape in which slower plans suddenly occurs around actual
%cardinality, therefore, not as stable as the shape with lower {\tt CEI} score.

### 2017-08-08 subopt-tods17/OLDtex/related.tex
subopt tods submission

\section{Related Work}\label{sec:related}

\todo{Rick has modify token}

SQL has emerged as the {\em de facto} and {\em de jure} standard relational
database query and update language. It emerged as an ANSI and ISO standard
in 1986--7, with a major extension and refinement as SQL-92~\cite{Melton93} 
and an even larger extension as SQL:1999~\cite{Melton03} and refinement as
SQL:2008~\cite{SQL2008}.

Query evaluation 
proceeds in several steps~\cite{Ramakrishnan03}. First, the SQL query is 
translated into alternative query evaluation plans based on the 
relational algebra via {\em query enumeration}. The cost
of each plan is then estimated and the plan with the lowest estimated cost
is chosen. These steps comprise query optimization, specifically {\em
cost-based query optimization}~\cite{Selinger}. The selected query plan is
then evaluated by the query execution engine which implements a set of
physical operators, often several for each logical operator such as
join~\cite{Graefe93}.

There has been extensive work in query optimization over the last 35
years. An influential survey~\cite{Chaudhuri98} does a superb job of
capturing the major themes that have pursued in the hundreds of articles
published on this general topic. Chaudhuri reviews the many techniques and
approaches that have been developed to represent the query plans, to
enumerate equivalent query plans, to handle some of the more complex lexical
constructs of SQL, to statistically summarize the base data, and to compute the
cost of evaluation plans. He also mentions some of the theoretical work
(which is much less prevalent) to understand the computational complexity of
these algorithms. Most of this work may be classified as adopting an
engineering perspective: how can we architect a query optimizer ``where 
(1)~the search space includes plans that have {\em low cost} (2)~the costing
technique is {\em accurate} (3)~the enumeration algorithm is {\em
  efficient}. Each of these three tasks is nontrivial and that is why
building a good optimizer is an enormous undertaking.'' \cite[page~35]{Chaudhuri98}

To determine the best query access plan, the cost model estimates the
execution time of each plan. There is a vibrant literature on this
subject~\hbox{\cite{Ioannidis03,Mannino88}}, including proposals for
histograms, sampling, and parametric methods. Again, these papers almost
always are engineering studies, providing new techniques that improve on the
state-of-the-art through increased accuracy or performance. There have also
been a few mathematical results, such as ``the task of estimating distinct
values is {\em provably} error prone, i.e., for any estimation scheme, there
exists a database where the error is significant''~\cite{Chaudhuri98}.

An optimizer for a language like SQL must content with a huge search space of
complex queries. Its first objective must be {\em correctness}: that the
resulting query evaluation plan produce the correct result for the
query. A secondary but clearly very important objective is {\em efficiency}; after
all, that is the raison d'etre for this phase. As is well known, the name
for this phase is an exaggeration, as existing optimizers do not produce
provably optimal plans. That said, the
query optimizers of prominent DBMSes generally do a superb job of producing
the best query evaluation plan for most queries. This performance is the
result of a fruitful collaboration between the research community and
developers.

However, even with great effort over decades, optimizers as a general class
are still poorly understood. As has been observed, ``query optimization has
acquired the dubious reputation of being something of a black
art''~\cite{Babcock05}. DeWitt has gone farther, stating that ``query
optimizers [do] a terrible job of producing reliable, good plans [for
  complex queries] without a lot of hand
tuning''~\cite[page~59]{winslett02}. And as we'll see, suboptimality is
possible even when considering only simple queries.

### 2017-08-08 subopt-tods17/OLDtex/repeatability.tex
subopt tods submission

\section{Repeatability}
We ensure the validity of our experiment results examining the
{\em repeatability} of our experiments. 
The repeatability study focuses on three factors, namely
{\em data repeatability}, {\em within-run repeatability},
and {\em across-run repeatability}, respectively.

First, repeatability is examined within data. Although the data 
that is populated for experiment tables are randomly generated, we ensure the
same numbers are generated each time.
Second, the query plans produced at the same cardinality for the same query
should be identical. Finally, as the cardinality of the variable-table
reduces, change points, where a different plan is detected, are identified. 
The experiment is repeatable if the same set of change points, which appears 
at the same cardinalities across multiple experiment runs, can be identified.

\shorten{
We use the following query $Q$ as an example to explain repeatability.

\begin{center}
\begin{tabular}{rl}
$Q$:	& {\tt SELECT t2.id4, t3.id1} \\
		& {\tt FROM HT2 t3, HT1 t0, HT4 t2, HT3 t1} \\
		& {\tt WHERE (t3.id4=t0.id1 AND }  \\
		& {\tt t0.id1=t2.id1 AND t2.id1=t1.id4)} \\
\end{tabular}
\end{center}
}
\shorten{
We first obtain a plan at 1M rows of {\tt HT1} (1M plan). 
Then, we find different plans, deleting the end of 10K rows 
from {\tt HT1} initially loaded with 2M rows (max) 
until the minimum cardinality which is 1M rows. 
All change points are also recorded with the plans.

In the second pass, $Q$ gets executed at all change points including 
the first 1M rows. 
We double-check whether the plan found at the change point is equal to one 
obtained at the time of execution. 
The executions for $Q$ are later used for telling the suboptimality of $Q$ 
by comparing the execution time of the 1M plan with that of other plans.

Interestingly, what we found was that in the first pass we happened to 
observe two different plans at 1M rows of {\tt HT1}. 
That is, the plans at 1M rows were inconsistent with each other 
even though all the involved tables' cardinalities were the same. 
}

\shorten{
The repeatability issue is initially raised from an experimental scenario, 
termed {\em Adjacent}, which studies query sub-optimality. 
(For more details about query suboptimality, please refer to 
Section~\ref{suboptimality}.)
The scenario includes data definition, query definition, and 
query plan detection and execution strategy. 
By data definition, we populate a few tables with fixed cardinality, 
saying 1M rows, and use a variable-table to get cardinality changing. 
Queries can be generated or pre-defined by query definition. 
For instance, let us consider the following query $Q$.}

This query involves four distinct tables, namely {\tt HT1}, {\tt HT2},
{\tt HT3}, and {\tt HT4}, which are all configured with the same sets of
columns, namely ({\tt id1}$\sim${\tt id4}). These four tables participate
in three joins, each joining a distinct pair of tables.
Tables {\tt HT2}, {\tt HT3}, and {\tt HT4} have the same fixed cardinality,
which is 1M rows. We term these tables the {\em fixed}-tables.
On the other hand, {\tt HT1} is termed a {\em variable}-table in that we
vary its cardinality from 2M rows down to 10K rows, at a step of 10K rows
each time.

\subsection{Data Repeatability}
To populate these tables, we utilize a random number generator. The values
for column {\tt id1} across the four tables are sequentially assigned.
The values for the other columns are populated by the generator. We control
the seed for the number generation such that the produced sequence of
``random'' numbers are identical each time. To change the cardinality of
the variable-table, we populate this table to the maximal cardinality,
which is 2M rows, and then {\tt DELETE} from the end of the table 10K rows
at a time.

\subsection{Within-Run Repeatability}
To study suboptimality in query evaluation, we define an experiment scenario
to carry out the experiment in an automated and structured fashion.
The scenario contains two execution stages.
In the first stage, we initially populate the variable table to have 1M rows,
which is the same to the fixed tables. We utilize the {\tt EXPLAIN} command
to produce the execution plan for $Q$ and then execute the query to measure
the running time. This measurement is used as a reference point for further
study.
We then re-populate the variable table to the maximal of 2M rows, and we delete
10K rows from the end of the table each time and call {\tt EXPLAIN} to
collect the plans. As the cardinality is reduced, if the newly generated plan
is different from the previous one, we note the current cardinality as a
{\em change point} and we store the cardinality as well as the plan. At the
end of the first stage, we usually collect a list of change points.

In the second stage, we re-populate the variable table to the maximal
cardinality again. According to the list of change points, we can directly
delete a number of rows from the variable table to achieve the target
cardinality at which a change point was found in the previous stage.
We run the query at the point to measure the running time. To ensure
that the current query execution utilizes the same time as the corresponding
change point, we compare the two plans.

An interesting observation is that sometimes in the two stages, a change
point is identified at cardinality 1M. The occurring plan at this change point
is different from the the reference plan initially generated. The reason
is that by deleting rows from a table, the deleted rows are simply flagged
without being physically removed. This leads to the issue that the number of
occupied pages, which is a factor considered by the cost model employed by
the optimizers, does not tightly correlate with the number of rows in the
table.

\shorten{
One of the reasons we speculated was that a way of setting the cardinality 
for the variable table might cause the optimizer to choose different 
plans in spite of the same cardinalities. 
To vary the cardinality of {\tt HT1}, a table with the same definition 
as {\tt HT1}, called {\tt max$\_$HT1}, is populated to the max cardinality, 
2M rows. 
To get the first 1M plan, the 1M rows of {\tt HT1} were copied from {\tt max$\_$HT1}. 
On the contrary, the second 1M plan was acquired due to the 10K-row-deletion.

To see whether the plan difference, or unrepeatability reappeared even after 
making cardinality by a consistent method, 
we also tried other variety of scenarios: {\it deleteTop10K}, {\it doubleTriple}, 
{\it copy1M}. 
In {\it deleteTop10K} scenario, {\tt HT1} is first loaded with 2M rows 
by the max table copy, and 10K rows from {\tt HT1} are incrementally 
deleted until 1M rows. 
A 1M plan is eventually acquired at the minimum cardinality as before. 
We repeat these steps so that both of the 1M plans can be compared 
for repeatability check.

{\it DoubleTriple} scenario basically extends {\it deleteTop10K}. 
Four plans are obtained by running {\tt deleteTop10K} scenario twice. 
Thereafter, instead of the incremental deletion, we get a 1M plan right after 
copying 1M rows into {\tt HT1} from the max table as before. 
By repeating this, the total six plans are obtained. 
We ensure repeatability by checking whether all of the six 1M plans are identical. 
}

To address the issue that the number of pages and the number of rows are not
correlated, we designed a new scenario, essentially with a different
cardinality-varying mechanism. We term this new scenario {\em copy1M}.
Unlike deleting rows, the alternative mechanism creates a new table and
utilizing the {\tt SELECT ... INTO ...} command to populate the newly created
table to the target cardinality. This approach effectively ensures that
the cardinality of the table is coupled with the number of pages the table
occupies.

\subsection{Across-Run Repeatability}
While the within-run repeatability is ensured, we run each experiment multiple
times to collect running time samples for statistical analysis. A change point
can be uniquely identified by the query ID and the cardinality value within
a run. Hence by grouping the change points with the same query ID and
cardinality value across multiple runs, should the samples be correctly
collected for each change point.

Nonetheless, we observed that for the same query, different sets of change
points are found for various runs. We utilized one DBMS to conduct a simple
test, that is to invoke the {\tt EXPLAIN} facility with the same query
at different times, which was often one minute apart or hours apart. The
underlying tables were of course not changed whatsoever. We found that
the produced plans at various times could actually be different even though
the tables stay unchanged. We concluded that this plan inconsistency issue
was due to the randomness introduced by the heuristics adopted by
DBMSes to compute the table stats.

Given that it is difficult to ensure the across-run repeatability,
we altered the approach to collecting multiple running-time samples.
Instead of execute multiple runs for each experiment, each plan, once
identified, is executed multiple times such that the samples are guaranteed
to be collected on the same plan.

In fact,\todo{is this true? I don't have facts to back this up...} the
within-run repeatability was sometimes violated due to a similar
reason. In the adjacent scenario, two stages (passes) are made to identify
and execute query plans, respectively. However, the plan produced in the
second pass could be different from the first pass.

To address the randomness caused by the optimizer, we restrict that
the optimizer only produces plan once for the same query at the same
cardinality. We term this new approach {\em onePass}. The resulting
new scenario is termed accordingly. In this scenario, instead of making
two passes of varying the cardinality of the variable table, the cardinality
is only varied once such that once a change point is identified, the generated
plan is executed multiple times. This scenario finally resolved the
repeatability violations in our experiments.
\todo{Do we talk about the two-table settings here?}

\input{onepass.tex}

\subsection{Summary}
In summary, we learned various tricky situations where repeatability
could be violated during experiments and designed the one-pass scenario
to address these issues. Moreover, we experienced great variance in
running-time measurements due to system daemons and cron jobs. We had most
of the unnecessary processes turned off, which effectively reduce the
variance in measurements. We consider this also an effort to ensure the
repeatability for the experiments, in terms of running-time sample collection.

\todo{Young mentioned exhaustive before, but I now think it should go to the monotonicity section.}
\todo{Rick and Young agreed with Rui about this.}
\shorten{
We first applied these scenarios to DBMS A through E with a few runs of 100 queries. 
As a result, DBMSes A and B showed the repeatability for all queries. 
In particular, they were not affected by which of the above scenarios 
was run or how many times the scenario was run. 
On the contrary, the other DBMSes C, D and E failed to retain repeatability. 
They were sensitive to queries, runs and scenarios.

As a last check, we also took advantage of CLI (Command-Line Interface) on 
the unrepeatable DBMSes in order to see directly the 1M plan without passing 
through the scenarios. 
DBMS C showed the same plan at a cardinality for a query at all times. 
However, the other two DBMSes D and E produced different plans for the same 
query at times.

The empirical observation about repeatability can be summarized as follows. 
Some heuristic algorithms may be exploited for the optimizers of DBMS 
D and E, which may affect the enumeration and decision of query plan. 
(To figure out this, we made several attempts to look for {\tt heuristic} keyword 
from the source code of DBMS D and E optimizers. 
However, it was not clear of where it was used.) 
DBMS C has relatively plenty of operators compared with other DBMSes, so 
its compatible operators may show up at different times, considering that plans 
built by such operators would have the same execution cost. 
In the meantime, DBMS A and B were repeatable.

Based on the observation from all the DBMSes, we eventually ensure 
repeatability (and monotonicity) by using {\tt exhaustive} scenario.
In {\tt exhaustive}, multiple query executions are made at each cardinality, 
varying the max cardinality to the min one; namely, for a query we get a plan 
and execute the query multiple times, saying 10, upon decreasing the cardinality. 
Instead of the incremental 10K-row-deletion, 
changing the cardinality is done by copying from the max table as many rows as needed 
so that all pages are in full with the rows. 
This setup allows us to check if any of the multiple executions produces 
different plans at the same cardinality (and see if the query execution 
times are monotonically decreasing as cardinality decreases). 
Each DBMS ran {\tt exhaustive} with a run of 10 queries, and it had 
no case such that different plans were detected at the same cardinality 
while a query got executed repeatedly.
}

### 2017-08-08 subopt-tods17/OLDtex/summary.tex
subopt tods submission

\section{Summary and Future Work}\label{sec:future}
{\todo{Rick has the modify token}
\todo{Talk about where more work is needed, based on the results}
This paper studies an important
component of a DBMS, the query optimizer. This component is an amazingly
sophisticated piece of code, but is still not well understood after decades
of research and development.

This paper makes the following contributions in
an attempt to gain a new understanding of this component.
\begin{itemize}
\item Introduces a new methodological
perspective, that of {\em empirical generalization}.
\item Uses that perspective to study {\em query suboptimality}, a problem
  not previously investigated.
\item Proposes a {\em predictive model} that attempts a causal explanation of
  suboptimality. 
\item Utilizes a novel {\em research infrastructure}, \azdb, to express, run, and
  analyze experiments.
\item Tests several hypotheses deductively derived from the causal model; in
  each case, the null hypothesis was rejected.
\item Provides a {\em space of hypotheses} that in concert through further
  investigation can refine and support (or not) this model or suggest a
  better model.
\item Thereby provides a path toward scientific progress in the
  understanding of a key enabling technology.
\end{itemize}

We studied a particular
phenomenon, that of the optimizer selecting a wrong plan, indicated by the existence of a query plan that performs more efficiently
than the DBMS' chosen plan, for the same query. From the engineering
perspective, it is of critical importance to understand the prevalence
of suboptimality and its origin(s). Our theory is that suboptimality
is caused in part by the inherent complexity of these system and the
concomitant unanticipated interactions between various rules in the
optimizer.

Through a series of experiments managed by a novel Laboratory
Information Management System termed \azdb, we uncovered compelling evidence 
(a) that suboptimality occurs, (b)~that  suboptimality
correlates with an aspect of query complexity, that of the number of
correlation names used in the query, (c)~that suboptimality correlates
with the cardinality of the effective plan space 
(CEPS)---another (query-specific) measure of optimizer complexity, and (d)~that
query complexity negatively correlates with query complexity and with optimizer
complexity. All of these findings were predicted by our theory.

\comment{
We presented a taxonomy of flutter types and found that the three DBMSes
we studied differed greatly in the types of flutter they produced. We found
that, as expected, flutter was correlated to the number of correlation names 
in the FROM clause. We also found a correlation between the number of 
self-joins and flutter prevalence, and a negative correlation between 
self-joins on the variable table and flutter prevalence. On both, our theory is silent.

Hence, every interaction predicted by our theory was borne out by this experiment, 
thus providing support for our theory.

We also uncovered the surprising result that for many queries, a majority of
the ones we considered, the optimizer picked the wrong query, even when the
cardinality estimates were completely accurate and even for our simplistic
queries. (Note, however, that these results are for only one DBMS.)

One potential engineering ramification of this new understanding is that, 
in cases where the DBMS isn't as sure about the cardinalities of the underlying 
relations or the speed of the disk (if such relations migrated frequently~\cite{reiss03}), the optimizer
shouldn't try as hard to wring out the ``perfect plan,'' Perhaps optimizers
should instead explicitly take uncertainty into account. Indeed, others have
started to argue that uncertainties in the query planning process should be
acknowledged and exploited~\cite{Babcock05}. But those are engineering
considerations; our focus in this work is on the science and on
understanding at a fundamental level these complex systems. Good engineering
should build on solid scientific results.
}
Note that our theory doesn't apply to just one algorithm or to one DBMS.
Rather, it is quite broad, applying to any DBMS with a cost-based
optimizer. We plan to augment \azdb\ to enable other DBMSes to be studied; 
our theory makes specific predictions on what we may observe with these other systems.

\comment{These results also raise several unsettling concerns. These particularly simple
queries used only a few plans within any DBMS. Other more complex queries
will certainly use more plans. These results imply that as a DBMS becomes
more sophisticated, by adding operators, it also become more unstable, as
the prevalence of flutter grows. Perhaps there is a fixed level of
complexity beyond which a cost-based optimizer will be inexorably unstable.}

There are several directions we wish to take this work. 
\c2j{}{Our study raise interesting questions about self-joins, with correlations 
in opposite directions observed. Might this be DBMS-specific? Or is there 
an overarching theory to explain this phenomenon?
}
These investigations are based on very simple SPJ (select-project-join)
queries.  We wish to look at {\em schema complexity}:
the influence of foreign keys and indexes, {\em query complexity}: complex
predicates, subqueries, and aggregates, and {\em instance complexity}:
skewed data.
We also want to manipulate DBMSes internally (at least for
those that are open-source), turning on and off the rules and observing
suboptimality.
%\todo{Rick: DeWitt: anecdotal, broken: first scientific evidence}
Might there be other unanticipated interactions,
that are not known about
simply because they haven't been looked for?

A second concern is the prevalence of sub-optimal plans uncovered in this
work. 
We plan to examine sub-optimality within the other
DBMSes. We need to look at the three complexities listed above also within
this investigation scenario. Does sub-optimality increase with complexity?
Is there a complexity whereby sub-optimality is the rule?

Our predictive model postulates two causal factors to suboptimality:
optimizer complexity and the cost model, the latter specifically, non-linearity of
the cost formul\ae\ and any cost model inaccuracies. In this paper we
considered portions of both factors, but showed only that optimizer
complexity definitely plays a role. It is important to also evaluate the
contribution of non-linearity and cost model inaccuracies. For some DBMSes,
the result of {\tt EXPLAIN PLAN} includes estimated cost of the plan, or even of
each stage of the plan. For these systems, that cost (estimate) is a
dependent variable, one that can be observed. If multiple plans include the
same operator, it may be possible to assemble, from outside the DBMS, an
approximation of the cost formula for that operator, and thus directly
observe whether it is non-linear and/or inaccurate, and then to understand
the impact of that dependent variable on suboptimality. In doing so, one
could test additional hypotheses derived from the causal model introduced in
this paper, ascertain the relative contributions of these two central
factors of the model to suboptimality, and perhaps also identify other
general factors that contribute to suboptimality across DBMSes. The result
would be a truly general, predictive, causal model of this vital component
of a DBMS.

### 2017-08-08 subopt-tods17/OLDtex/timing.tex
subopt tods submission

\subsubsection{Single Execution Time Measurement}\label{sec:qe_time}

\begin{figure}[th]\centering
\includegraphics[width=0.50\textwidth]{figures/time_measurement.eps}
\caption{Processes influencing query execution\label{fig:timing}}
\end{figure}

In this section we discuss how we measure query execution time. 
As illustrated in Figure~\ref{fig:timing}, 
we first get all running processes before executing 
a query $Q$. 
Then timer for $Q$ gets started, and it is stopped after $Q$ gets 
executed. 
The execution time for $Q$ is measured by the time difference, and 
we finally finish single query execution by examining running 
processes right after the execution.

Now let us examine how other processes may influence query execution. 
Assume that all the processes below are not relevant to DBMS processes.
Before timing $Q$, processes $P$1, $P$2 and $P$3 are recorded into 
a map $M$1. 
Another map $M$2, on the contrary, keeps processes $P$3 and $P$5 
captured after the timing. 
For each recorded process, the maps $M$1 and $M$2 keep 
$<$process id (pid), \{process name, minimum/major faults, 
cpu/system times\}$>$. 
The cpu and system times are considered for the subtraction. 
The process information can be achieved by iterating 
{\tt /proc/\{pid\}/stat} associated with each process.

By comparing $M$1 and $M$2, we easily know that $P$3 and $P$5 affected 
$Q$'s execution; thus, the time spent on $P$3 and $P$5 are subtracted from 
the measured query execution time. 
$P$1 and $P$2 that do not appear in $M$2 must have stopped 
before or during the query execution, respectively. 
We cannot determine whether the query execution is affected by 
these stopped processes. 
For instance, $P$2 is explicitly involved with the execution whereas 
$P$1 not. 
What we can know is that regarding the execution, there were stopped 
processes, such as $P$1 and $P$2, as a result of comparison with 
$M$1 and $M$2. 
This execution, therefore, will be thrown out due to the stopped ones.

Note that $P$4 is a {\em phantom} process, since it does appear 
in neither $M$1 nor $M$2. 
The existence of phantom process can be known by comparing 
the number of processes from {\tt /proc/stat} before and after execution.
Like stopped processes, we throw out any execution with a phantom process, 
considering measurement accuracy.

\begin{figure}[t]
\begin{center}
\begin{algorithmic}
{\bf Algorithm} timeSingleQueryExecution($i$, $q$, \\
						\hspace{44.0mm}$exPlan$, $card$):
\STATE $plan$ $\leftarrow$ get(Prepared)QueryPlan($q$)
\IF{$plan$ $\neq$ $exPlan$}
	\STATE Error: two different plans are observed \\ 
	at the same cardinality. \\
\ENDIF
\STATE Flush disk drive/OS/DBMS caches. \\
\STATE $M1$ $\leftarrow$ $extractAllRunningProcs$() \\
\STATE $startTime$ $\leftarrow$ $getCurrTimeMills$() \\
\STATE $executeQuery$($q$) \\ 
\STATE $endTime$ $\leftarrow$ $getCurrTimeMills$() \\	
\STATE $execTime$ $\leftarrow$ $endTime$ - $startTime$ \\	
\STATE $M2$ $\leftarrow$ $extractAllRunningProcs$() \\
\STATE $procDiff$ $\leftarrow$ $diff$($M1$, $M2$) \\	
\STATE RecordQueryExecution($i$, $q$, $card$, \\ 
			\hspace{34.0mm}$execTime$, $procDiff$) \\
\end{algorithmic}
\caption{Timing single query excution\label{alg:timing}}
\end{center}
\end{figure}

### 2017-08-08 subopt-tods17/cover-letter.tex
subopt tods submission

--- 2017-07-26 subopt-tods16/cover-letter.tex
@@
 \begin{flushright}
-\hfill National Institute of Supercomputing and Networking, 
+\hfill National Institute of Supercomputing and Networking
 \end{flushright}
@@
 \begin{flushright}
-\hfill Korea Institute of Science and Technology Information (KISTI), 
+\hfill Korea Institute of Science and Technology Information (KISTI)
 \end{flushright}
@@
 \begin{flushright}
-\hfill Republic of Korea, 34141, 
+\hfill Republic of Korea, 34141
 \end{flushright}
@@
 \begin{flushright}
-\hfill Email: yksuh@kisti.re.kr
+\hfill Email: {\tt yksuh@kisti.re.kr}
 \end{flushright} 
@@
 
-\opening{Dear Prof. Christian S. Jensen:}
+\opening{Dear Prof.~Jensen:}
 
@@
 
-\item No version of this manuscript has been previously published nor be currently under consideration for publication elsewhere. 
-We have anonymized the current submission as directed by the double-blind
-review guidelines on the TODS web site. Note that Section 9 (on pages 36--37) of
-this manuscript summarizes the contributions in an attempt to draw 
-a causal explanation of suboptimality and to uncover a fundamental limit 
-on the number of operators that a DBMS may be able to support.
+\item No version of this manuscript has been previously published nor is
+  currently under consideration for publication elsewhere.
+
+\item We have anonymized the current submission as directed by the double-blind
+review guidelines on the TODS web site. 
+%Note that Section 9 (on pages 36--37) of
+%this manuscript summarizes the contributions in an attempt to draw 
+%a causal explanation of suboptimality and to uncover a fundamental limit 
+%on the number of operators that a DBMS may be able to support.
 

### 2017-08-08 subopt-tods17/paper.tex
subopt tods submission

\documentclass[prodmode,acmtods]{acmsmall}
\pdfminorversion=6

%Metadata Information
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmYear{2017}
\acmMonth{8}

\usepackage{url}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphics}
% \usepackage[total={6.5in,9in}, top=1.25in,left=0.9in]{geometry}
%\usepackage[text={6.5in,9in},centerpage,includefoot]{geometry}
% \usepackage{setspace}
\usepackage{color}
\usepackage{url}
\usepackage{balance}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{subfigure}
\usepackage{nicefrac}

\definecolor{grey}{RGB}{200,200,200}
\newcommand{\hilite}[1]{\colorbox{grey}{#1}}
\newcommand{\hilitey}[1]{\colorbox{yellow}{#1}}
\newcommand{\hiliting}[1]{\colorbox{grey}{#1}}
\long\def\todo#1{\hilitey{{\bf TODO:} {\em #1}}}
\long\def\shorten#1{}

%For last edit: revise and resubmit for SIGMOD'14
\long\def\shortenF#1#2{}%First is letter indicating what has been removed

%For double blind, keep the second argument. To turn off doubleblind, keep
%the first
\long\def\doubleblind#1#2{#2}
%To remove discontinuity, keep the second argument. To include discontinuity,
%keep the first argument
\long\def\discont#1#2{#2}
\def\azdb{\doubleblind{\hbox{\sc AZDBLab}}{\hbox{\sc DBLab}}}
\def\QatC{Q{@}C}
\long\def\comment#1{}
%c2j: Conference to Journal: first parameter is conference, second is journal
% For Conferences: c2j#1#2{#1}
% For Journals: c2j#1#2{2}
\long\def\c2j#1#2{#2}

\begin{document}

\title{%\vspace*{-0.75in}
A Causal Model of \hbox{DBMS} Suboptimality\\
}
\markboth{\ }{A Causal Model of \hbox{DBMS} Suboptimality}

% possible, but not really needed or used for PVLDB:
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as\textit{Author's Guide to Preparing ACM GIG Proceedings Using \LaTeX$2_\epsilon$\ and BibTeX} at \texttt{www.acm.org/eaddress.htm}}}

% ****************** AUTHORS **************************************

% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

%bug is vldb.cls gives an additional authors section for greater than 3...
\doubleblind{\numberofauthors{3}}{}
 %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.

\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
\doubleblind{% 1st. author
Richard T.~Snodgrass\\
       \affaddr{Dept. of Computer Science}\\
       \affaddr{University of Arizona}\\
%       \affaddr{Tucson, AZ}\\
       \email{\small\tt rts@cs.arizona.edu}
\alignauthor
% 2nd. author
\alignauthor
Sabah Currim\\
       \affaddr{Alumni Office}\\
       \affaddr{University of Arizona}\\
%       \affaddr{Tucson, AZ}\\
%       %\vspace*{-0.05in}
       \email{\small\tt scurrim@email.arizona.edu}
% 3rd. author
\alignauthor
Young-Kyoon Suh\\
       \affaddr{Dept. of Computer Science}\\
       \affaddr{University of Arizona}\\
%       \affaddr{Tucson, AZ}\\
       \email{\small\tt yksuh@cs.arizona.edu}
}{}}

\begin{abstract}

The query optimization phase within a database management system (\hbox{DBMS}) ostensibly finds the fastest query
execution plan from a potentially large set of enumerated plans, all of
which correctly compute the specified query. Occasionally the cost-based optimizer
selects a slower plan, for a variety of reasons. We introduce the notion of
{\em empirical suboptimality} of a query plan chosen by the DBMS,
indicated by the existence of a query plan that performs more efficiently
than the chosen plan, for the same query. From an engineering
perspective, it is of critical importance to understand the prevalence
of suboptimality and its causal factors.

We propose a novel structural causal model to explicate the relationship
between various factors in query optimization and empirical
suboptimality. Our model associates suboptimality with the factors of
complexity of the schema, data, query, and optimizer and concomitant
interactions among the components of the optimizer. This model induces a
number of specific hypotheses that were subsequently tested on multiple
\hbox{DBMSes}.

Through a series of experiments that examine the plans for thousands of
queries run on one hundred thousand query/cardinality combinations on four
popular DBMSes, we observe that the dependent construct of empirical
suboptimality prevalence correlates positively with (a)~\hbox{DBMS}
complexity, (b)~schema complex, (c)~data complexity, and (d)~query
complexity, providing empirical support for this model. These
\discont{ten}{nine} factors explain in concert over half of the
variance of suboptimality, across four disparate DBMSes. Thus, it is the
common aspects of these DBMSes that predict
suboptimality, {\em not} the particulars embedded in the inordinate complexity of
each of these DBMSes.

An implication of this causal model is that as query evaluation operators
are added to a DBMS, the prevalence of slower queries will grow. Through a
novel experiment that examines the plans on the afore-mentioned
query/cardinality combinations, we present evidence for a previously-unknown
upper bound on the number of operators a DBMS may be able to support before
performance suffers.  We
show that this upper bound may have already been reached by one or more
extant~DBMSes.

This paper thus provides a new methodology to study mature query optimizers, that
of empirical generalization, proposes a novel causal model for empirical query
suboptimality, and demonstrates an upper bound that may have already
been encountered, with implications for fundamental improvements of those
cost-based query optimizers.
\end{abstract}

% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures,
%performance measures]
\terms{}
\keywords{}
\acmformat{\doubleblind{Richard T.~Snodgrass, Sabah Currim, and Young-Kyoon
    Suh, 2017, A Causal Model of \hbox{DBMS} Suboptimality}{A Causal Model of
    \hbox{DBMS} Suboptimality}}

\doubleblind{\begin{bottomstuff}
This work was supported in part by NSF grants \hbox{IIS-0639106},
IIS-0415101, and EIA-0080123. 
%
%Author's addresses: G. Zhou, Computer Science Department,
\end{bottomstuff}}{}
\maketitle

\section{Introduction}\label{sec:intro}
Database management systems (DBMSes) underlie information systems and hence optimizing their performance is of
critical importance. A relational DBMS's cost-based query optimizer plays an
important role, ostensibly finding a fast query execution plan from a
potentially large set of possible plans, all of which correctly compute the
submitted query, using a cost model that references properties of the
underlying relations. But
what if the optimizer {\em doesn't}: what if it selects a slow plan?

This paper provides a thorough investigation into \hbox{DBMS}
suboptimality: when the \hbox{DBMS} chooses a slower plan over a faster
plan for a given query. We systematically examine the factors influencing
the number of suboptimal queries. There could be multiple causes of the
suboptimality. One possible cause could be some peculiarity within the tens
of thousands of lines of code of that query optimizer. Another possible
cause could be the query's complexity. Prior research in other domains shows
that increasing complexity negatively influences
performance~\cite{campbell88,moody98}. A third possible reason could be some
fundamental limitation {\em within the general architecture of cost-based
  optimization} that will always render a number of queries suboptimal.

To better understand the impact of different factors on suboptimality of
query performance and the relationship between operators, especially in a
dynamic environment, an experimental approach is needed.  Based on existing
research and general knowledge of \hbox{DBMSes}, we propose an innovative
predictive model of suboptimality in query evaluation. From this point on,
by ``suboptimality'' we mean specifically ``empirical suboptimality'',
indicated by the presence of another plan that can be empirically determined
to be faster than the plan that was chosen by the query optimizer. Indeed,
as we will elaborate in Section~\ref{sec:suboptimality}, where we provide a
specific operationalization of this notion, empirical suboptimality is an
{\em ordinal variable}, indicating the degree to which plans across a
range of table cardinalities are slower than other (empirically-identified)
plans. This contrasts with the conventional notion of query plan
suboptimality, which is a binary {\em categorical variable} associated with
a particular cardinality: a plan is either optimal (the best-performing
plan) or it is suboptimal.

We employ an experimental methodology on a collection of \hbox{DBMSes} as
subjects to test our hypotheses with respect to factors influencing
suboptimality, utilizing empirical data collected over a \hbox{cumulative}
16,000 hours (over two years) of query executions.  Our research falls
within creative development of new evaluation methods and metrics for
artifacts, which were identified as important design-science
contributions~\cite{hevner04}.

%\pagebreak
The key contributions of this paper are as follows.
\begin{itemize} 
\item We use an innovative {\em methodology} that treats \hbox{DBMSes} as experimental subjects.

\item We find that for a surprisingly large portion of queries, the plan
  chosen  by the query optimizer is not the best plan, for some cardinality
  of the underlying tables.

\item We propose a {\em predictive model} for \hbox{DBMSes} to better
  understand the factors causing suboptimality.

\item We test the six quite specific hypotheses that arise from that
  model across a wide range of DBMSes, queries, and data, showing
  through correlational and regression analyses that four
  hypotheses are strongly supported, one is weakly in the
  other direction, and
  the sixth is partially supported, lending credibility to our particular model.

\item Through an innovative
  analysis, we track the net cumulative benefit of a succession of operators
  added to the DBMS, and show (a)~that there is a {\em limit} to the number of
  relational operators that a DBMS can accommodate before slower plans start
  to dominate and (b)~that one or more extant relational DBMSes may have
  already reached that limit.

\item The predictive model and these experimental results suggest several specific
  engineering~directions.

\item The model and analyses presented here taken as a whole imply that a
  new approach to query optimization, fundamentally
  different from the cost-based approach utilized over the last forty years, may be required
  to get past the fundamental limitation uncovered in this research.
\end{itemize}

This paper takes a scientifically rigorous approach to an
area previously dominated by the engineering perspective, that of database
query optimization.  Our goal is to \hbox{understand} cost-based query optimizers
as a {\em general} class of computational artifacts and to come up with
insights and ultimately with predictive theories about how such optimizers,
again, as a general class, behave.  These theories can be used to further
improve \hbox{DBMSes} through engineering \hbox{efforts} that benefit from the fundamental
understanding that the scientific perspective can provide.

One might ask, shouldn't the task of optimizing queries be left to DBAs?  In
databases, and especially in data warehouses, the number of users writing
and running queries has been growing exponentially. This growth is aided, in
part, by the drag and drop query tools provided by the different
systems. For example, subject areas allow business users to write queries
without knowing anything about the underlying database structure. This,
coupled with constantly growing data presents new challenges for tuning. For
example, \doubleblind{the data warehouse deployed at the University of
  Arizona}{a large data warehouse we are familiar with} runs about
30,000 queries on a daily basis. 
Also, tuning a subject area or tables for one group of queries can
negatively impact the performance of other queries. Query optimization experts often take hours to tune and test existing canned
queries; the amount expended on this one system for manual query optimization
approaches \$100K/year. Therefore, we argue that it is important to
understand how existing query optimizers can be further improved, and
indeed, whether  fundamental limitation inherent in these optimizers exist
and whether such limitations are indeed already being encountered.

We focus here on the effectiveness of query
optimization. The query optimization phase within a \hbox{DBMS} ostensibly
finds a fast query execution plan, drawn from a potentially large set of enumerated plans, all of
which correctly compute the specified query. (The term ``query optimizer'' is thus aspirational rather
than realistic: the goal of the optimizer isn't really to find the best
plan, but rather to avoid the worst plans.) However, this intermediate goal
might not always be realized in practice, for several possible reasons. One is
that it may not be practical to enumerate all possible plans, and so the
fastest plans, or even the fast plans, may not be considered. Also, because
the optimizer chooses a plan based on its estimated query execution time, if
that estimate is of, the optimizer may instead choose a different, less
efficient plan. From both a scientific perspective and an engineering
perspective, it is of critical importance to understand the phenomenon of
suboptimality.

We study these aspects \c2j{}{across \hbox{DBMSes}} to identify the
underlying causes. We have developed a predictive causal model that
identifies five constructs that may play a role in suboptimality. The ultimate goal is to {\em
  understand} a component within a \hbox{DBMS}, its cost-based optimizer, through the
articulation and empirical testing of a general scientific theory.

In Section~\ref{sec:related} we briefly summarize the vast amount of related
work in query optimization to establish the technical basis for our study.
The following section introduces the methodology we will follow, that of
{\em empirical generalization}~\cite{cohenbook}. We present in
Sections~\ref{sec:model} and~\ref{sec:operationalization} a predictive,
causal model of suboptimality and state six specific hypotheses derived
from that model. We then test these hypotheses across a number of queries,
data, and cardinalities, whose results provide strong support for the
validity of the proposed model.  These are the first scientific results that
we are aware of that apply {\em across} multiple \hbox{DBMSes}, rather than on a
single, specific \hbox{DBMS} or on a specific algorithm. In
Section~\ref{sec:diminishing} presents a follow-on analysis that
uncovers a previously-unknown fundamental limit to the number of operators that a
cost-based optimizer can support, and shows that this limit may have already
been reached by one or more extant DBMSes.
We explore in Section~\ref{sec:engineering} implications of the model for
research in engineering more efficient \hbox{DBMSes}.

\section{Motivation}\label{sec:motivation}

Consider a simple select-project-join (SPJ) query, with a few attributes in
the SELECT clause, a few tables referenced in the FROM clause,
and a few equality predicates in the WHERE clause. This query might be an
excerpt from a more complex query, with the tables being intermediate
results.

%%%% query info: runid: 440, dbms: pgsql, querynum:13, exp: op-1M-100q-6
%\vspace*{-2ex}
\noindent
\hspace{3ex}{\small\begin{verbatim}
        SELECT t0.id1, t0.id2, t2.id4, t1.id1 
        FROM ft_HT3 t2, ft_HT2 t1, ft_HT1 t0 
        WHERE t2.id4=t1.id1 AND t2.id1=t0.id1
\end{verbatim}
}

%\vspace*{-2ex}
\noindent
The optimizer generates
different plans for this query as the cardinality of the {\tt
  ft\_HT1} table varies, an experiment that we will elaborate later in depth.

The upper graph in Figure~\ref{fig:query769} represents the plans chosen by
a common DBMS as the cardinality of {\tt FT\_HT1} decreases from 2M tuples
to 10K tuples in units of 10K tuples. The \hbox{x-axis} depicts the estimated
cardinality and the y-axis  a plan chosen for an interval of
cardinalities. So Plan~P0 was chosen for 2M tuples, switching to Plan~P1 at
1,830,000 tuples, back to Plan~P0 at 1,640,000 tuples, and so on, through
the plan
sequence P0, P1, P0, P1, P2, P0, P2, and finally P3 at 40,000 tuples.

\begin{figure*}[t]
\centering
%originally 30pc
%\epsfig{figure=figures/plan769.pdf,width=36pc}
\includegraphics[width=30pc]{figures/subopt_flutter.pdf}
%\includegraphics[scale=0.42]{figures/subopt_flutter.pdf}
\caption{An Example of Suboptimality and Fluttering\label{fig:query769}}
\end{figure*}

The lower graph in Figure~\ref{fig:query769} indicates the query times
executed at adjacent cardinalities, termed the
``query-at-cardinality'' (Q@C) time, when the plan changed. For some transitions, the Q@C time at
the larger cardinality was also larger, as expected. But for other
transitions, emphasized in red ovals, the Q@C time at the larger cardinality
was {\em smaller}, such as the transition from plan P1 at 1,650,000 to P0 at 1,640,000
tuples: the plan at the higher cardinality actually took {\em less} time
than the plan at the lower cardinality. Such {\em change pairs}, where the time at the
  higher cardinality is shorter, identify suboptimal plans. For the change pair at 720,000
tuples, P0 required 2.35sec whereas P1 at a larger cardinality required
2.41sec (as is common, the plan at the higher cardinality takes more time). This query exhibits seven plan change pairs, two of which are suboptimal.

This query also illustrates an interesting phenomenon, in which the
optimizer returns to an {\em earlier} plan. Sometimes the
optimizer starts oscillating between two plans, sometimes even switching
back and forth when the cardinality estimate changes by a small
percentage. The example query showed returning to P0 twice and to P1 and to
P2 each once.
We call this phenomenon, in which
the query optimizer returns to a previous plan,
``query optimizer flutter'', or simply ``flutter''.

We have found through our experiments that flutter and suboptimality are all
around us: {\em every} \hbox{DBMS} that we have examined, including two
open source \hbox{DBMSes} and two proprietary
\hbox{DBMSes}, covering much of the installed base worldwide, exhibit these
phenomena, even for very simple queries. In the Confirmatory
Experiment described in detail in Section~\ref{sec:experiments}, we started
with 6,967 query instances (a query run on a specific \hbox{DBMS}) after \doubleblind{our}{an} extensive query measurement
protocol~\cite{TTPv1,TTPv2} applied its extensive sanity checks. While about
20\%, or 1,491, of these query instances contained only one
plan, a few of the other query instances switched plans at almost every change in cardinality (we varied the
cardinality in increments of 10K tuples, a total of 200 cardinalities): see
Figure~\ref{fig:planchanges}. Slightly over half, or 3,933 query instances, exhibited
suboptimality somewhere along those 200 cardinalities; a few had
many changes to a plan that was in fact suboptimal, as indicated in
Figure~\ref{fig:suboptplanchanges}, across all four \hbox{DBMSes} considered.

\begin{figure}[t]\centering
\includegraphics[width=0.60\textwidth]{figures/org_ncpq.pdf}
\caption{Cumulative percentage of queries \hbox{exhibiting} the \hbox{indicated}
  number of plan changes\label{fig:planchanges}}
\end{figure}

\begin{figure}[t]\centering
\includegraphics[width=0.60\textwidth]{figures/zi_ncpq.pdf}
\caption{Cumulative percentage of queries exhibiting the \hbox{indicated}
  number of changes to a {\em suboptimal} plan\label{fig:suboptplanchanges}}
\vspace*{-2ex}
\end{figure}

One oft-stated observation is that the role of query optimization is not to
get the {\em best} plan, but rather to get a plan that is acceptably
good. (Thus, the very term ``query optimizer'' is aspirational rather than
accurate.) Figure~\ref{fig:suboptcumulative} shows
the cumulative distribution of the relative amount of suboptimality (where
an $x$-value of 100 denotes that the query ran 100\% slower than the optimal
query, that is, twice as long). The good news is that 2,738 query instances,
or 67\% of the suboptimal queries, exhibited only a small degree of
suboptimality: less than 30\%. The challenge is that over fifth of all
queries (1,355) exhibited a significant amount of suboptimality ($\geq
30\%$). The relative slowdown can thus be quite large for some queries.

We started with 7,640 query instances, each of which is a particular query running on a
particular DBMS, cf.~Experiment 7 of Table~\ref{tab:run_stat}
in Section~\ref{sec:experiments}. After our protocol, we were left with
  6,967 query instances, of which 5,475 had at least one change pair, so those are
  the ones we consider further. Of those, 1,382 had {\em no} suboptimality, so
4,093 (75\%) had some suboptimality.

1,951 (36\% of the query instances with a change pair) have the higher cardinality
running at least 20\% faster than the lower cardinality. That means that
2,142 query instances (slightly over half of the suboptimal queries) were barely
suboptimal ($<$20\% slower) and about a third (1,355) were considerably
suboptimal (where the higher cardinality ran at least 30\% faster than the
lower cardinality).

\begin{figure}\centering
\includegraphics[width=0.60\textwidth]{figures/rel_diff.pdf}
\caption{Cumulative percentage of queries exhibiting the \hbox{indicated}
  percentage relative suboptimality\label{fig:suboptcumulative}}
\end{figure}

We emphasize four important points.
\begin{itemize}
\item First, we used a sophisticated query
measurement methodology that reduces the measurement variance, so that the
query plans we identify as suboptimal definitely are so.
\item Second, these
results are over four \hbox{DBMSes}, and thus, such phenomena are not dependent on
a particular implementation of cost-based optimization. Rather, they seem to
be common to {\em any} cost-based optimizer, independent of the specific
cardinality \hbox{estimation} or plan costing or plan enumeration algorithm or
implementation.
\item Third, suboptimal plans are {\em not} the result of poor
coding or of inadequate algorithms. We view query optimization in modern
\hbox{DBMSes} as an engineering marvel, especially given the complexity of the SQL
language and the requirements and expectations of \hbox{DBMS} users, who often
demand that important queries simply not get slower with a new release of
the \hbox{DBMS}. Rather, the prevalence of suboptimality observed here is a
reflection of the complexity of the task of query optimization.
\item Fourth, we wanted to understand whether a fundamental
limitation exists in the prevalent approach to query optimization utilized
by DBMSes generally, and certainly by the four DBMSes that we studies.
\end{itemize}

This is why we needed a new methodology. We want to understand cost-based
optimization deeply. This means that we need to go beyond the examination of
a single query optimizer, as is done in almost every paper (with a few
important exceptions~\cite{harish07,Haritsa10,Leis15}) over the forty-year history of
query optimization research, to study multiple instances of that
optimization architecture, in an effort to achieve generalizable results.

Thus, in this paper, we articulate a predictive causal model for how and in
what circumstances suboptimality arises and provide compelling evidence via
\hbox{hypothesis} testing that this model accurately characterizes the
\hbox{behavior} of query optimizers in general. This is what is meant by
``empirical generalization''~\cite{cohenbook} and why it is needed to answer
such questions. In the long history of research in database query
optimization, or even of databases in general, our model and its hypothesis
tests are the first predictive results that we are aware of that apply {\em
  across} \hbox{DBMSes}, rather than on a single, specific \hbox{DBMS} or on
a specific algorithm. Our goal is to make statements
that hold across cost-based optimizers in general, thereby moving up the
$y$-axis of Figure~\ref{fig:empirical}.
Such a DBMS-agnostic, though paradigmatic, causal
model can then \hbox{provide} guidance to the community about where fundamental
research is needed and to the \hbox{DBMS} engineers about where to focus
their efforts.

\begin{figure*}[t]
\centering
\includegraphics[width=0.60\textwidth]{figures/understanding.pdf}%0.40
\caption{Empirical Generalization\label{fig:empirical}}
\end{figure*}

\section{Related Work}\label{sec:related}

\shorten{SQL has emerged as the {\em de facto} and {\em de jure} standard relational
database query and update language. It emerged as an ANSI and ISO standard
in 1986--7, with a major extension and refinement as SQL-92~\cite{Melton93} 
and an even larger extension as SQL:1999~\cite{Melton03} and refinement as
SQL:2008~\cite{SQL2008} and SQL:2011~\cite{Melton11}.  
}
There has been extensive work in query optimization over the last 40
years~\cite{Ioannidis96,Jarke84}, during which a particular quite effective
paradigm had taken hold, in both open-source and proprietary DBMSes. In this
over-arching paradigm, query optimization and evaluation
proceeds in several general steps~\cite{Ramakrishnan03}. First, the SQL query is 
translated into alternative query evaluation plans based on the 
relational algebra via {\em query enumeration}. The cost
of each plan is then estimated and the plan with the lowest estimated cost
is chosen. These steps comprise query optimization, specifically {\em
cost-based query optimization}~\cite{Selinger}. The selected query plan is
then evaluated by the query execution engine which implements a set of
physical operators, often several for each logical operator such as
join~\cite{Graefe93}.

An influential survey~\cite{Chaudhuri98} identifies
the major themes that have pursued in the hundreds of articles
published on this general topic. Chaudhuri reviews the many techniques and
approaches that have been developed to represent the query plans, to
enumerate equivalent query plans, to handle some of the more complex lexical
constructs of SQL, to statistically summarize the base data, and to compute the
cost of evaluation plans. He also mentions some of the theoretical work
(which is much less prevalent) to understand the computational complexity of
these algorithms. Most of this research may be classified as adopting an
engineering perspective: how can we architect a query optimizer ``where 
(1)~the search space includes plans that have {\em low cost} (2)~the costing
technique is {\em accurate} (3)~the enumeration algorithm is {\em
  efficient}. Each of these three tasks is nontrivial and that is why
building a good optimizer is an enormous undertaking.'' \cite[page~35]{Chaudhuri98}

To determine the best query access plan, the cost model estimates the
execution time of each plan. There is a vibrant literature on this
\hbox{subject~\cite{Ioannidis03,Mannino88}}, including proposals for
histograms, sampling, and parametric methods. Again, most of these papers
are engineering studies, providing new techniques that improve on the
state-of-the-art through increased accuracy or performance. There have also
been a few mathematical results, such as ``the task of estimating distinct
values is {\em provably} error prone, i.e., for any estimation scheme, there
exists a database where the error is significant''~\cite{Chaudhuri98}.

An optimizer for a language like SQL must contend with a huge search space of
complex queries. Its first objective must be {\em correctness}: that the
resulting query evaluation plan produce the correct result for the
query. This objective must be ensured both by the initial SQL-to-relational
algebra translator and by the subsequent query enumerator. A secondary but clearly very important objective is {\em efficiency}; after
all, that is the raison d'\^etre for this phase. As is well known and has been
alluded to already, the name
for this phase is an exaggeration, as existing optimizers do not produce
provably optimal plans. That said, the
query optimizers of prominent \hbox{DBMSes} generally do a superb job of producing
the best query evaluation plan for most queries. This performance is the
result of a fruitful collaboration between the research community and
developers.

Early investigation of plan suboptimality resulted in approaches such as
dynamic query-reoptimization~\cite{Avnur,Bellamkonda13,kabra98,Li07}, which exploit more
accurate runtime statistics that appear while a query is being executed, to
steer in-flight plan reoptimization. The very presence of such a radical
change to the normal optimize-execute sequence indicates that plan
suboptimality was of interest to some researchers.

However, even with great effort over decades, optimizers as a general class
are still poorly understood. As has been observed, ``query optimization has
acquired the dubious reputation of being something of a black
art''~\cite{Babcock05}. DeWitt has gone farther, stating that ``query
optimizers [do] a terrible job of producing reliable, good plans [for
  complex queries] without a lot of hand
tuning''~\cite[page~59]{winslett02}. And as we will see, suboptimality may
occur in sophisticated query optimizers even when considering only simple queries.

While this paper does not provide direct solutions to address
suboptimality, we envision that by following up on the implications of our proposed predictive model
for suboptimality, engineering practice, such as dynamic reoptimization just
mentioned, may benefit. We elaborate on this subject in
Sections~\ref{sec:root} and \ref{sec:engineering}, where we discuss the engineering implications
of our causal model.

\section{A Model of Suboptimality}\label{sec:model}

The purpose of query optimization is to generate optimal plans.  So
why would suboptimality occur in the first place? Query optimizers are
highly complex, comprised of tens or hundreds of thousands of lines of code. There
are several reasons for this complexity. First, an optimizer must
contend with the richness of the SQL language, whose definition
requires about 2000 pages~\cite{SQL2008}, with a multiple of
linguistic features. Second, an optimizer must contend with the
\hbox{richness} of the physical operators available to it. \hbox{DBMSes} have a
range of algorithms available to evaluate each of many algebraic
operators. Third, the optimizer must contend with an exponential
number of query evaluation plans. Kabra and DeWitt~\cite{kabra98}
identify several other sources of complexity: inaccurate statistics on
the underlying tables and insufficient information about the runtime
system: ``amount of available resources (especially memory), the load
on the system, and the values of host language
variables.''~(page~106). They also mention user-defined data types,
methods, and operators allowed by the newer object-relational
systems~\cite{Melton03}. Thus, the task of optimization is very
complex, with the result that the optimizers themselves consist of a
collection of ``components'', that is, the rules or heuristics that it
uses during optimization, with each of these components being itself
complex.

We wish to understand the causal factors of suboptimality, through a
predictive model that explicitly states the relationships between these
causal factors. We test this model through experiments over tens\shorten{14,700} of thousand
of queries and hundreds of thousands of query executions\shorten{2.4M}, showing that there
is strong support for this model. We then extract engineering implications
from the model, suggestions for the most productive places to look to reduce
suboptimality and thus to improve existing query optimizers.

\begin{figure*}[tb]
\centering
%originally 30pc
%\epsfig{figure=figures/model.pdf,width=30pc}
\includegraphics[width=30pc]{\discont{figures/subopt_model.pdf}{figures/subopt_model_wo_disc_op.pdf}}
\caption{Predictive Model of Suboptimality\label{fig:model}}
\vspace*{3ex}
\end{figure*}

Here we examine the hypothesized influence that each independent
variable will have on the one dependent variable, query Suboptimality (with
some of the influences mediated by one of the constructs). In
the next section we will operationalize these variables, explaining how
each is controlled or measured.

\subsection{Constructs in the Model}

The model concerns five general
constructs that we \hbox{hypothesize} will play a role in suboptimality: optimizer
complexity, schema complexity, query complexity, data complexity, and plan space complexity.
Two constructs include several specific variables that contribute to that
construct as a whole. Our model distills many of the widely-held assumptions
about query optimization. Our contribution is the specific \hbox{structure} of
the model and the specific operationalization of the
\hbox{factors} included in the model.

Our theory, which we will investigate in some detail and which is the
basis for our predictive model, is that suboptimality is due in part
to the complexity of the optimizer and concomitant 
interactions between the different components used within the 
optimizer with various sources of complexity in producing a good (hopefully
optimal) plan.  We argue that with the proliferation of concerns and
variability that an optimizer must contend with, it is extremely
difficult to ensure that for an arbitrary query that the optimizer
will {\em always} pick the right plan. There are many sources of complexity
that may challenge one or more components of the optimizer; our theory
implicates specific sources as contributing to observed
suboptimal plans.

The model depicted in Figure~\ref{fig:model} suggests specific factors that may
impact the prevalence of suboptimality. This model has one dependent
variable, {\em Suboptimality} (as noted in Section~\ref{sec:intro}, specifically, {\em empirical query suboptimality}), on the far right. We observe this \hbox{dependent
}variable in our experiments by determining whether each particular query,
run on a particular \hbox{DBMS} and using a particular schema,
is suboptimal at any cardinality of the input table. In Section~\ref{sec:operationalization}, we
delve into the details of how we operationalize this and the other variables
we now examine. The last two sections summarize our contributions and
suggest three fundamental directions that are suggested by this study.

\subsection{Variables in the Model}
The model has several independent
variables which influence Suboptimality.

For the construct of {\em optimizer
complexity} we have one independent variable, ``\hbox{Number} of operators
available (in the \hbox{DBMS})''. We can manipulate this variable by our choice of
\hbox{DBMS}: each \hbox{DBMS} has a set of operators available to its query
evaluator and available to the query optimizer to use with in query plans.

We have one variable for the construct of {\em schema complexity}:
``Presence of secondary indexes''. If true, that means that every
non-primary-key attribute, for the table for which we vary the cardinality
(termed the {\em variable table}) as well as for the other three tables
(termed the {\em fixed
tables}), has a secondary index associated with it. The rationale is that the
Presence of secondary indexes expands the number of possible plans: each
predicate in the query can often be mapped to one or more operators
utilizing that index.

For the construct of {\em query complexity} we have identified four
variables: ``Number of correlation names'', ``Presence of aggregate'',
``Presence of primary key attribute'', and ``Presence of subquery''. For each independent
variable, we have interventional control in our experiments, in that we can
manipulate the values of these variables through the construction of the
actual query to be optimized. The first is the number of correlation names
defined in the FROM clauses. The second is whether a single aggregate operator
({\tt Sum}) appears in the SELECT clause. The third
is whether a primary key attribute appears in at least one predicate in the
WHERE clause. The fourth is whether a single subquery appears in the
WHERE clause; that subquery evaluates to a single value that is
equality-compared with an attribute, in the where clause. The rationale is
that each factor may expand the number of plans or otherwise render the search for
the optimal plan more complex.

We have one variable for the construct of {\em data complexity}: that of
``Presence of skewed data''.  Skew is generally defined as how values are
distributed with a column of a table. We refine this to ``how many duplicate
values are present,'' with zero skew meaning that there are no duplicate
values present and 100\% skew implies that there is but one value in the
entire column. The Presence of skew complicates query time estimation, which
in turn complicates the search for the optimal plan.

In the middle of the figure is the construct of {\em plan space
complexity}. Given a particular query, its
complexity will impact the total number of candidate plans considered
by the optimizer. This variable is not directly
observable, again, especially for proprietary \hbox{systems}. However, we {\em
can} measure the number of plans actually generated by the optimizer
when presented with different cardinalities of the underlying tables.
We term this set of plans the ``effective plan
space'' and term the number of such plans, the cardinality of the
effective plan space, or ``{\em CEPS}''. Similarly, \discont{we cannot directly observe
the cost model utilized by the optimizer for the operators it considers, but
can classify the cost model of each operator as continuous (smoothly varying
with cardinality of its input(s) and without jumps) or discontinuous
(sometimes producing observable jumps in the predicted
cost), and can thus count the ``Number of operators with
discontinuity''. (Such operators have also been termed
``nonlinear''~\cite{Hulgen03}, though we focus on a more specific aspect of
discontinuity.) Finally,}{} the variable ``Number of repeats'' {\em is}
directly measurable as the number
of times a plan is reused across the cardinality of the variable table. (We
discussed plan reuse and the related phenomenon of {\em flutter} in
Section~\ref{sec:motivation}; Figure~\ref{fig:query769} provided a concrete
example in which plan P0 was encountered three separate times across the 200
cardinalities.) For
example, if the sequence of plans for a query as the cardinality increases
is $A ~ B ~ B ~ B ~ C ~ C ~ A ~ B ~ B~ B ~ C ~ B ~ C ~ C ~ C ~ B$, then the
``Number of repeats'' will be six: removing sequential duplicates results in $A ~ B ~ C
~ A ~ B ~ C ~ B ~ C ~ B$, with the last six distinct plans being duplicates
of the first three. Note that CEPS plus the number of repeats gives you the
number of plans in the sequence, again, after removing sequential
duplicates, nine plans in this particular case. Thus, we associate with the latent construct
of plan space complexity \discont{three}{two} measurable (that is, indirect) variables. Our
  model stipulates that the construct, and thus the \discont{three}{two} associated with
  this construct, intervene between the constructs of optimizer complexity,
  schema complexity and query complexity and the construct of Suboptimality.

Plan space complexity is an {\em intervening} construct, in that it is dependent
on some of the constructs on its left but is observable and thus exerts influence on the dependent
variable on its right. We can observe these variables within experiments and
indirectly influence their value but
cannot directly intervene to specify their value. For example, we can influence the CEPS through
manipulating the values for the independent variables of optimizer and query
complexity, but cannot directly set a value for CEPS within an
experiment.

\subsection{Hypotheses in the Model}\label{sec:hypotheses}
Given these six constructs and \discont{eleven}{ten} specific variables depicted in
Figure~\ref{fig:model}, let's now examine the causal relationships between
these variables, each depicted as a directed arrow between variables (a line
originating or ending at a construct is interpreted as originating or ending at all
variables in that construct). The causal model predicts specific hypotheses  between
the the constructs (in particular, between their associated variables).

One causal factor of this model is the optimizer complexity. We hypothesize
that optimizer complexity has influence over suboptimality indirectly, via
plan space complexity. We hypothesize that an optimizer with a larger number
of available operators will generate more plans and hence increase plan
space complexity.

\vspace{0.6em}\noindent
{\bf Hypothesis 1}: Number of operators available will be positively
correlated with (a)~Number of repeats\discont{, (b)~CEPS, and (c)~Number of operators with
  discontinuity}{ and (b)~CEPS}.

\vspace{0.6em}We now turn to query complexity, a construct associated with four
independent variables. As with the optimizer complexity construct, we
include in our model an
indirect effect through plan space complexity.

A higher value of each of these specific variables implies a more complex
query.  We expect a strong relationship between Number of correlation names
to CEPS (because it is well-known that the number of potential join
combinations is exponential to the number of correlation names)\discont{,}{ and} to Number of
repeats (Number of repeats could partially track CEPS)\discont{, and Number of
operators with discontinuity (for the same reason)}{}.

We also expect a positive relationship between Presence of aggregate (as
that will definitely add at least one operator to the query plan), Presence
of primary key attribute, and
Presence of subquery.

\vspace{0.6em}\noindent
{\bf Hypothesis 2}: Number of correlation names will be
strongly correlated with (a)~Number of repeats\discont{,}{ and} (b)~CEPS\discont{, and (c)~Number of
operators with discontinuity}{}. Presence of aggregate will be positively
correlated those those \discont{three}{two} variables (correlations \discont{d--f}{c--d}) and similarly
with Presence of primary key attribute
(correlations \discont{g--i}{e--f}) and Presence of subquery (correlations \discont{j--l}{g--h}).

\vspace{0.6em}
We feel that skewed data (in the ``data complexity''
construct) will not impact plan space complexity. Rather, our expectation is
that it will negatively impact accuracy of plan cost estimation, and thus
increase suboptimality directly.

\vspace{0.6em}\noindent
{\bf Hypothesis 3}: Presence of skewed data will be negatively correlated with Suboptimality.

\vspace{0.6em}
Let's now turn to the plan space complexity construct. We hypothesize a positive
correlation between the two variables associated with this construct. As the
Number of plans considered by the optimizer increases, so should CEPS\discont{, which
could increase the Number of operators with discontinuity that is observed}.

The model also predicts \discont{two correlations}{a correlation} within plan space
complexity, between 
CEPS and \discont{Number of operators with discontinuity variables and with
}Number of repeats. \discont{

We hypothesize that greater plan space complexity should make it more
difficult to optimize the query.  It is important to note that the
occurrences of suboptimality is not necessarily dependent on the Presence
of an operator with discontinuity. However, we predict that when
discontinuous plan operators appear in candidate plans, they may introduce
complexity to query optimization, especially at the cardinality where
discontinuity is possible. This is because the performance of such operators
is sensitive to the input size in a rather complex way. Any inaccuracy in
the estimation of plan statistics may lead the optimizer to select a suboptimal
plan.

Similarly, a}{A} plan chosen by the optimizer once will be reconsidered at a
later cardinality and perhaps chosen, following
the past experience of selecting and using that plan among many other plans.
Even if many different plans are already used, 
for the same reason the optimizer may revisit a pool of 
the previously used plans and choose one of them than to explore other new plans, 
given the complexity of plan cost estimation. 
Then the larger CEPS, the more times the optimizer repeats using the same plans. 
We thus hypothesize that the Number of repeats has a positive correlation with CEPS.

\vspace{0.6em}\noindent
{\bf Hypothesis 4}: Plan space complexity (CEPS) and \discont{(a)~Number of operators
with discontinuity and (b) }Number of repeats will be positively correlated.

\vspace{0.6em}
The schema complexity construct consists of the variable ``Presence of
secondary indexes''. Such indexes provide opportunities for the optimizer to
consider more candidate plans that use these indexes, due to the additional
query evaluation operators now applicable. Those additional plans enable the optimizer to
possibly do a better job, while also adding complexity to the optimization
process.  The quality of query optimization ``very much depends on how much
the query engine relies on [cardinality] estimates and on how complex the
physical database design is, i.e., the number of indexes
available.''~\cite{Leis15}.

We hypothesize that this factor has a more complex role in the
model, serving as a {\em moderator} of two relationships introduced
above. We hypothesize that the overall effect of the secondary indexes 
is to increase the strength of the relationship between query
complexity and to plan space complexity. (Note that as the Presence of primary
key is required for secondary indexes, we don't include that former
independent variable here.)

\vspace{0.6em}\noindent
{\bf Hypothesis 5}: Presence of secondary indexes will strengthen
the correlations between the Query Complexity construct, specifically, the
variables Number of
correlation names, Presence of aggregate, and Presence of subquery, and
the Plan Space Complexity construct, specifically, variables Number of
repeats\discont{,}{ and} CEPS\discont{, and Number
of operators with discontinuity (correlations a--f and \todo{Rick: check:}j--l listed above in
Hypothesis 2).}{ (correlations a--f listed above in Hypothesis 2).}

\vspace{0.6em}
Finally, we hypothesize that queries with a high plan space complexity
present challenges to the query optimizer, and thus increase the chance that
the query optimizer picks a suboptimal plan. \discont{This is a fairly direct
connection for CEPS and Number of operators with
discontinuity. For}{Specifically, for} the
Number of repeats variable, a query exhibiting a high number of repeats has
more opportunities for suboptimality at some cardinality, just because the
number of repeats is an indication that the query optimizer is struggling.

\vspace{0.6em}\noindent
{\bf Hypothesis 6}: Suboptimality will be positively correlated with Plan
space complexity, that is, (a)~Number of repeats\discont{, (b)~CEPS, and (c)~Number
of operators with discontinuity}{ and (b)~CEPS}.

\vspace{0.6em}
We have just {\em described} how
suboptimality might arise, through a theory and its elaborated causal model,
which implies six specific hypotheses (some with multiple components).
We now need to move to {\em prediction}.
How might we test such a model?

The first step to test this model is to {\em operationalize} each
variable. In the next section we describe explicitly how each variable is
defined. For the independent variables, we must be able to intervene, that is,
set their values before the experimental test commences. For the
dependent variables, we need to be able to measure their values during each experiment.

\section{Variable Operationalization}\label{sec:operationalization}

In this section we specify how we operationalized each of the seven independent
variables in the model.  Recall that each independent variable is a property
of a \hbox{DBMS} (Number of operators available), of the schema (Presence
  of secondary indexes), of a query (Number of correlation names,
Presence of aggregate, Presence of primary key attribute, and
Presence of subquery), of the data (Presence of skewed data), and of
the plan space (Number of repeats\discont{, CEPS, and Number of operators
  with discontinuity}{ and CEPS}). There is 
one dependent variable of our model (Suboptimality).

It is important to note that manipulation must be done {\em
  outside} the \hbox{DBMS}, as we do not have access to the
internal code of proprietary DBMSes. Hence, we do not know ({\em cannot} know) all the plans that were
considered, nor the details of how the plans were selected. But such access
is not needed; indeed, to be able to study a phenomenon across many \hbox{DBMSes},
such access is not feasible. But by designing the experiment to examine
the plans that each \hbox{DBMS} actually produces, and thus to examine phenomena
that can be externally visible, we can obtain valuable insights into general
classes of computational artifacts.

\subsection{Optimizer Complexity}
By ``Number of operators available'' we mean the \hbox{number} of operators
available in the \hbox{DBMS} for selection, projection, join and aggregate
functions (that is, potentially relevant for our queries). Our
experiments intervene on this variable by selecting a particular \hbox{DBMS} on
which to evaluate each query. Across the available \hbox{DBMSes}, as the Number of
operators available increases, the complexity of the optimizer increases
because it has to choose between more operators.

The {\tt EXPLAIN PLAN} facility specifies the operators(s) \hbox{employed} in that
plan. For each \hbox{DBMS}, we collect the unique operators from the plans
and count the number of these distinct operators, each used by for at least one
query at at least one cardinality by that \hbox{DBMS}. The Number of operators used
by a \hbox{DBMS} ranged from 8 to 53 across the queries and data sets
used in the Exhaustive, Exhaustive with Keys, and Exploratory Experiments,
to be discussed in Section~\ref{sec:experiments}.

\subsection{Schema Complexity}
Presence of secondary indexes is easy to operationalize. We generate two
databases, one without any secondary indexes and one with a key specified
for each table on the non-key (that is, other than the first) attribute.

\subsection{Query Complexity}\label{sec:querycomplexity}
Query complexity is also relatively easy to operationalize. We randomly generate queries,
such as the example presented in Section~\ref{sec:motivation}. Each query
is a select-project-join-aggregate query, with a few attributes in the
SELECT clause, a few tables \hbox{referenced} in the FROM clause, a few
equality predicates in the WHERE clause, and zero or one aggregate
functions in the SELECT clause. As such, some of the complexities
mentioned by Kabra and \hbox{DeWitt~\cite{kabra98},} such as
user-defined data types, methods, and operators, are not
considered. Concerning Presence of primary key attribute, if this independent
variable was set to 1, we ensured that there was at least one primary key
attribute in one of the comparisons in the WHERE condition.

The queries were generated
by a simple algorithm\shorten{, the pseudo-code for which is shown in
Figure~\ref{alg:querygen}}. 
\shorten{The query generator has four main components, which are {\tt
  generateSelectClause()}, {\tt generateFromClause()}, {\tt
  generateWhereClause()}, {\tt generateGroupByClause()}. , and {\tt buildAggregateStatement()}.}
The SELECT clause will contain from one to four
attributes, hence, an average of 2.5 attributes. The Number of
correlation names
in the FROM clause varied from one to four, with duplication of tables
allowed (duplicate table names within a FROM clause implies a self-join).
In the queries that were generated, from one to four tables were
mentioned in the FROM clause. Somewhat fewer tables were mentioned than the
Number of correlation names, as the Presence of self-joins reduces the
number of unique tables referenced by the queries.

For the query in Section~\ref{sec:motivation}, the Number of correlation
names is 3, the Presence of aggregate
is false (0), the Presence of primary key attribute is false (0),
and the Presence of subquery is false (0).

%\pagebreak
\shorten{In {\tt generateWhereClause()}, in which all the joins will appear,
  we}We ensure that Cartesian products are
eliminated\c2j{.}{\shorten{({\tt cartesianPossible="false"})}.}  We do this by
connecting the correlation names that appear in the FROM statement via
equi-joins on random attributes. The comparisons are all
equality operators. To ensure that the queries are as simple as possible, we
do not include any additional predicates in the WHERE clause.  \c2j{}{This
  is realized by setting the attributes {\tt maxIsAbsolute} to {\tt true}
  and {\tt complexUsePercentage} to {\tt 100}.  Basically, ``complex''
  predicates eliminate the Cartesian product, and by setting complex
  predicates as ``absolute'', no additional predicates are included except
  for those which are necessary for eliminating Cartesian product.} Also
for simplicity, we include neither disjunctions nor negations.

\shorten{In the \todo{Correct name: generateGroupByClause}, where the aggregate
functions and
grouping occurs, the presence or absence of a group by clause is
recorded. We expect the
complexity of the query to increase when the first aggregate function
is introduced. (We don't consider in this model whether
additional aggregate functions in the same SELECT clause add
complexity}

The value of Presence of subquery was 0 or 1, with 0 indicating no
subquery. For each query needing a subquery, we picked a separate generated query and
rendered it as a subquery to replace an attribute in the WHERE clause of the
original generated query. The following query is an example where this
variable had a value of 1.

\begin{verbatim}
SELECT t0.id2, SUM(t1.id3)
FROM ft_HT2 t0, ft_HT2 t1
WHERE (t0.id2=t1.id1)
GROUP BY t0.id2
\end{verbatim}

We then generated another simple select-project (SP) 
query at random concerning the variable table 
and simply replaced one side (in this case, the right side) with the generated entire 
query as a subquery, to produce this final query.

\begin{verbatim}
SELECT t0.id2, SUM(t1.id3)
FROM ft_HT2 t0, ft_HT2 t1
WHERE (t0.id2 IN (SELECT t2.id3 FROM ft_HT1 t2)) 
GROUP BY t0.id2
\end{verbatim}
We thus have a maximum of only one level of nesting.

\subsection{Data Complexity}\label{sec:datacomplexity}
This independent construct has one variable: Presence of skewed data.

Skew has a very specific definition in statistics, 
involving elongating the left or right tail of a distribution, 
thereby moving the mean left or right of the median 
(in a symmetric distribution the mean = median = mode).

But we start with a distribution without a tail: 
the uniform distribution: the values from 1 to 2 million (2M). 
We consider this to be a skew of 0 (no skew). 
At the other end of the spectrum is one in which all the values are identical, 
or a skew of 1.0.

We define the skew as ``the reciprocal of the Number of distinct
values,'' so $0 < \hbox{\em skew} \leq 1$.
For 2M distinct values, the skew would be $\nicefrac{1}{2M}= 0.000005$, which is practically 0. 
For exactly one distinct value, the skew would be 1.0. For two distinct values, 
the skew would be 0.5. For ten distinct values, the skew would be 0.1.

We can generate the table of 2M rows by generating values sequentially from
1 to the Number of distinct values.  This creates a ``span of values.'' We
repeat this for the second span if necessary, and on and on, until we have
2M values in all.

When varying the cardinality, we remove 10K values from the variable table
and then copy those tuples to a new table to ensure that every page is as
full as possible (that is, 100\% load factor).  This gives us a table of
1.99M tuples. (We then get a query plan for this table.)  We repeat this
removal process until the final cardinality reaches 10K.

The way we effect the 10K removal is as follows. 
The key idea is to remove individual spans until we've deleted 10K
values. Since we don't touch the remaining spans, the Number of distinct
values does not change, and so the skew remains constant.

We use two values of skew: $\nicefrac{1}{2M}$ (termed {\em tiny}) and
$\nicefrac{1}{10K}$ (termed {\em small}).
\shorten{Let's first illustrate with a skew factor of 1.0. 
This translates to exactly 1 value for the entire table, or 2M spans.
To shorten, we remove 10K spans, equal to 10K values. 
This removal keeps the skew at 1.0, due to the unique values in the
remaining spans.
At the end, 10K spans, each with a single value, remain.

What about a skew factor of 0.1? 
This skew factor translates to 10 distinct values. 
So we generate 200K spans, each with ten values. 
To shorten, we delete 1,000 spans, which is equivalent to removing 10K values. 
As before, this does not change the skew.
At the end, the table will have 1,000 spans with 10K values, retaining a
skew factor of 0.1.

At the other end of the spectrum, consider a skew factor of
1/2M.  This skew factor, which is almost 0, gets translated to 2M
distinct values.  So}For the former, we generate a single span of 2M values.  To shorten the
table, it makes no sense to remove this single span in its entirety. But we
can remove 10K values from this single span.  Note that this changes the
skew to 1/1.99M, then eventually to 1/10K, which is still very close to zero
(the skew has changed from .0000005 to .00001).
For the latter, we generate 2,000 spans each of 10K tuples, and drop a span
to reduce to 1.99M tuples, repeating.
\shorten{We use five skewness values: 0.000005 (which we term only slightly
misleadingly as ``no skew''), 0.001, 0.1, 0.5, and 1.0,
for this independent variable.}

\subsection{Plan Space Complexity}\label{sec:discontinuity}

This explanatory construct includes \discont{three}{two} variables.

As discussed in Section~\ref{sec:model}, the ``cardinality of the effective plan
space'' (CEPS) is the number of plans selected as optimal for that query
being evaluated on one of the 200 cardinalities for the variable table. It
is ``effective'' because it was chosen, as opposed to the
plans that were considered but not chosen. (Recall that for proprietary
\hbox{DBMSes}, we do not have access to such plans.) Note that we count only
distinct plans. As we saw in Figure~\ref{fig:query769}, fluttering queries
return to a previous plan. This particular query has a CEPS of 4.

The Number of repeats is just the number of times a plan associated with a
smaller cardinality is repeated, after removing sequential
duplicates.

\discont{We also wanted to evaluate the contribution of cost model non-linearity to
suboptimality. We found that it is possible to assemble, from outside the
\hbox{DBMS}, an approximation of the cost formula for that operator, and thus
directly observe whether it is non-linear. Specifically, the result of SQL's
{\tt EXPLAIN PLAN} (a result that is particular to each \hbox{DBMS}) includes the
{\em estimated cost} of each stage of the plan. That cost (estimate) is a
dependent variable, one that can be observed.

To operationalize the ``Number of operators with discontinuity'' we
classify each \hbox{DBMS} operator as either continuous or discontinuous, to be
defined shortly.  We then examine the queries
to identify the plan change pairs, which provide the {\em effective plan
  space}, a set of plans chosen for one or
more cardinalities of the variable table. For each distinct query plan in the
effective plan space, we
count the number of discontinuous plan operators that appear in that plan. Some of
these plans may contribute no such operators, some may contribute one such
operator, and some may contribute several such operators. We then sum the
counts of the distinct plans in the effective plan space, yielding an
integer as its operationalization
for each query.  This count per query varied from 0 to 85 for the
queries we tested in the confirmatory analysis.

To classify each operator as continuous or discontinuous, we use the data
from the ``Exhaustive'' experiments to be described in detail in Section~\ref{sec:experiments}. We
used a small subset of queries to be used later in testing the model, to
attempt to observe the same operators as encountered in that
study. We
collected the query plan at each possible cardinalities (200 in all) and
looked for ``jumps'', that is, when the cost model for an operator in the
plan exhibited discontinuity. Jumps are determined from the estimated cost
extracted from each plan operator. We provide an example of a jump below. If
a jump is observed, we classify that particular plan operator as
discontinuous.

A {\em jump} is a pair of close cardinalities (in our case, separated by 10K
tuples) in which the query optimizer's cost model is {\em discontinuous},
that is, does not smoothly increase from the lower cardinality to the upper
cardinality. To identify jumps in the first step, we examine the slopes
between each pair of adjacent cardinalities (again, separated by 10K
tuples). We expect that the slope (that is, the first derivative) is
well-behaved and thus does not change much for adjacent cardinalities. 
A jump thus indicates a large deviation in the second derivative of the 
cost model across the two cardinalities.

The Exhaustive experiments gather the cost model for each operator in each
plan for each cardinality (in our case, for cardinalities ranging from 10K
to 2M in steps of 10K, or 200 cardinalities in all that were examined).  Figure~\ref{fig:discontinuity}
presents for a single query\shorten{({\em query4}, exhaustive-10q-1,
  oracle)}, the {\em cost} of each of the hash-join operators utilized in
each plan at a particular cardinality. Each distinct point in the graph
depicts the computed cost of an individual hash-join operator over an input
table of the indicated cardinality (showing only a small portion of the 200
cardinalities).  This figure is typical
(we generally observed jumps at low cardinalities for these particular queries and
relation sizes).  As shown by this figure, the hash-join operator
represented by `+' has two discontinuous jumps, both appearing below
cardinality 150K. The other hash-join operator represented by $\times$ has one
discontinuous jump.  Between the jumps, the behavior is linear. It is our
guess that
such a jump is caused by the transition between one pass of a disk-based hashing
technique to two passes (and indeed for one of the operators, perhaps the one
lower in the operator tree, the transition to three passes, hence, the two
jumps). That said, all
that matters for our predictive model is that operators that experience
such discontinuity in their cost models present opportunities for
suboptimality.

%\begin{figure}[t]\centering
\begin{figure}[t]
\centering
%\includegraphics[width=0.70\textwidth]{figures/discontinuity_400K.pdf}%0.40
\includegraphics[width=0.60\textwidth]{figures/discontinuity_500K.pdf}%0.40
\caption{An Example of Discontinuous Plan \hbox{Operators (Hash-Join)}}
\label{fig:discontinuity}
\end{figure}

\shorten{It is interesting that jumps occurred only at smaller cardinalities. This
might depend on the strategy of how the \hbox{DBMS} utilizes the main memory
buffer.  Assume that the \hbox{DBMS} attempts to use as little memory as possible.
One possibility is that when the cardinality is low, memory allocation is
done in small chunks. Hence as the cardinality grows, more passes are
required. However, when tables get larger, bigger
chunks of memory would be allocated, requiring the operator to take longer
to fill up the available memory. We again emphasize that our focus is on predicting
suboptimality. Perhaps memory allocation has a role in operator
discontinuity, which our model predicts to have a role in query
suboptimality.
}
To identify such jumps, we examine the second derivative (the change in the
first derivative, that is, the change in the slope). A jump will be indicated by a larger than normal second
derivative. By identifying a sudden change in the second derivative, we can
effectively spot the cardinality at which an operator in a plan becomes
discontinuous.

Formally, we compute the slope ({\em S}) of the {\em estimated} cost
(of the cost model, formalized as $C_{\hbox{\small\em card}}$,
where {\em card} is the input
cardinality) between each pair of adjacent cardinalities as
$S_{\hbox{\small\em card}} = (C_{\hbox{\small\em card} + 10K} -
C_{\hbox{\small\em card}}) / 10K$~.
This computes a series of slope values.
We then compute the standard deviation of the slope
values.  For example, examining Figure~\ref{fig:discontinuity}, the slopes
are small except at three places, one for the green operator and two for the
red operator. By identifying the slope values that are greater than one
standard deviation over the average value, the discontinuous operators can
be identified: A ``discontinuous operator'' is one for which a jump is
observed in that operator in one of the plans for at least one of the input
queries.
}{}

\subsection{Empirical Suboptimality}\label{sec:suboptimality}
We now consider the one dependent variable at the core of this investigation.
How might suboptimality be observed? We have developed a system, \azdb, that
allows us to perform experiments to study this phenomenon of suboptimality.
\hbox{\azdb} submits queries to the \hbox{DBMS}, while varying
the cardinality of one of the tables,
requesting in each case the evaluation plan chosen by the \hbox{DBMS}.
This is done using the
{\tt EXPLAIN} SQL facility available in modern \hbox{DBMSes}. (The Picasso system
also used this facility to visualize the plan space chosen by a \hbox{DBMS}
optimizer~\cite{harish07,Haritsa10}.) We can then compare
the performance (execution time) of various plans for the query, to identify those
situations when a suboptimal plan was chosen, when in fact there was a
different plan that was semantically equivalent to the chosen plan (that is,
yielded the identical result) but which ran faster.

We modify the cardinality to produce multiple execution plans for a given
query. For one of the \hbox{DBMSes}, we can modify the stored table
statistics directly. For the other \hbox{DBMSes}, we had to do so
indirectly, by varying the size of the table and running the optimizer on
tables of different size. As we vary the cardinality, we collect the plans
that the optimizer felt were appropriate for that query at the various cardinalities.

Our definition of suboptimality assumes that
actual execution time for any query plan is {\em monotonically non-decreasing}, that is,
unchanged or increasing as the cardinality increases.  The intuitive justification is that at the higher
cardinality, the plan {\em has} to do more work, in terms of CPU time and/or
I/O time, to process the greater number of tuples.\discont{ (For an operator having a discontinuous cost model, as the
anticipated input cardinality grows, there will be {\em jumps}, in which the
predicted cost is temporarily much greater. Section~\ref{sec:discontinuity}
explains further how we specifically
operationalize this property. Also note that we don't consider SQL
operators such as {\tt EXCEPT} that are not monotonic.)}{}
We formalize this property as follows.

\vspace{1em}\noindent
{\bf Definition:} {\em Strict Monotonicity}: given a query $Q$ and an actual
cardinality $a$,

\quad\quad\quad\quad\quad$\forall p \in \hbox{\em plans} (p) \;\forall c > a \;( \hbox{\em time} (p,
c) \geq \hbox{\em time} (p,a) )$

\vspace{1em}\noindent
where {\em plans}($p$) is the set of plans generated by the optimizer and
{\em time}($p$, $c$) is the execution time of plan $p$ on the data set with
the varying table at cardinality $c$.\hfill$\blacksquare$

\vspace{1em}
\noindent
Note that the comparison is with the same plan $p$, occurring at higher
cardinalities.

To test this assumption, we ran an experiment that we term
``Monotonicity''. This experiment considered 60 queries, chosen from the
pool of queries
generated for testing suboptimality, and timed them for cardinalities from 10K
to 2M tuples, in steps of 10K tuples (hence, we used 200 cardinalities), for
each \hbox{DBMS} (for one \hbox{DBMS} that was very slow, we started with
30K tuples, as will be
discussed in Section~\ref{sec:experiments}). We varied the cardinality of the variable table by starting with
the maximum size, running the queries, then deleting 10K tuples and repeating.
We performed an ``{\tt ANALYZE TABLE}'' function to force
the \hbox{DBMS} to update the table's statistics to be accurate before actually
evaluating the query (we did this for all experiments).
We expected that as the cardinality decreased,
the run time would also monotonically decrease. However, due to the variance in query
time measurement observed even when the cardinality was identical, we encountered
spurious violations.

Assuming a normal distribution for our time measurements, 95\% of 
the distribution falls within two $\sigma$ of the mean.
Therefore, to statistically
infer with a 95\% confidence interval that a violation occurred, 
we relaxed our definition of monotonicity to the following.

\vspace{1em}\noindent
{\bf Definition:} {\em Non-Strict Monotonicity}: given a query $Q$, an actual
cardinality $C$, and the standard deviation of the query executions for
cardinality C as $\sigma$, $Q$ is non-strict monotonic if $\forall p \in
\hbox{\em plans} (p) \;\forall c^{\prime} > C ((\hbox{\em time} (p, c^{\prime}) + \sigma_{c^{\prime}}) \geq
(\hbox{\em time} (p,C) - \sigma_{C}))~.~~\hfill\blacksquare$

\vspace{1em}
That said, the suspicious monotonicity violations raised the fundamental question of whether
we were using a sufficiently reliable timing method. The answer was {\em no}, as
the experiment used a (popular but) naive measurement technique 
based on \hbox{end-to-end} timing of which the results are 
typically perturbed by system noise potentially incurred during the measurement. 
We realized that the accuracy of this timing data was
insufficient for this monotonicity test.
We then switched to a new protocol, TTPv1~\cite{TTPv1}, and later TTPv2~\cite{TTPv2},
which fortunately is able to obtain quite accurate and precise timing.

Specifically, as we will see in Section~\ref{sec:experiments}, 
using the protocol we observed only 3,347 violations (0.74\%) of non-strict monotonicity, 
for the largest experiment (Experiment~7: Confirmatory) across all
the \hbox{DBMSes} we studied. 
Hence, this result justified our conclusion that the \hbox{DBMSes} under
study are indeed monotonic.

We can now turn to empirical suboptimality. Recall that the monotonicity test examines
two adjacent {\QatC}s for which the {\em same plan} is observed. To detect
suboptimality, we look for adjacent {\QatC}s with {\em different plans}, the
``change pairs'' mentioned earlier. We look for such change pairs where the computed
query time at the {\em upper} cardinality is {\em smaller} than the computed
query time at the {\em lower} cardinality. Say the lower cardinality used
Plan {\em A} and the upper cardinality exhibited Plan {\em B}. Had the \hbox{DBMS}
query optimizer selected Plan {\em B} for the lower cardinality, the query time would have been smaller than that for
Plan {\em A}, which follows directly from the
monotonicity assumption. The conclusion is that for the lower cardinality, the
optimizer picked the less efficient plan, and thus, this query exhibits
\hbox{suboptimality}. Note that since this approach cannot consider plans
that were never chosen, it very likely misses some suboptimal plans (for
which there was a better plan not seen), and thus produces a
conservative estimate of suboptimality.

Our definition of suboptimality compares the computed run times 
and standard deviations at the cardinality just before the change
pair (designated as $n-1$) and at the change pair (that is, $n$).

The query is said to be suboptimal if
%\hspace*{3em}
$\hbox{\em time}_{n-1} - 0.5 \cdot \hbox{\em stddev}_{n-1} \geq \hbox{\em
time}_n + 0.5 \cdot \hbox{\em stddev}_n$~.\\
In requiring at least one standard deviation of difference, versus just saying a Q@C pair is suboptimal if the time at higher cardinality is faster than the the time of plan at lower cardinality, we minimize spurious or minor differences in plan times. 
For each Q@C pair, Suboptimality is coded as four levels (0--3), based on the distance in 
standard deviations, up to three standard deviations. We chose this cutoff
because 99.7\% of Q@C pairs are less than or equal to three standard
deviations.
\shorten{\begin{itemize}
\item Not suboptimal, i.e., 0, when
$$\hbox{\em time}_{n-1} - 0.5 \cdot \hbox{\em
stddev}_{n-1} < \hbox{\em time}_n + 0.5 \cdot \hbox{\em stddev}_n $$
\item Low suboptimality, i.e., 1, when
$$\hbox{\em time}_{n-1} - 0.5 \cdot \hbox{\em
stddev}_{n-1} \geq \hbox{\em time}_n + 0.5 \cdot \hbox{\em stddev}_n$$
and
$$\hbox{\em time}_{n-1} 1 1.0 \cdot \hbox{\em stddev}_{n-1} < \hbox{\em time}_n +
1.0 \cdot \hbox{\em stddev}_n$$
\item Moderate suboptimality, i.e., 2, when
$$\hbox{\em time}_{n-1} - 1.0
\cdot \hbox{\em stddev}_{n-1} \geq \hbox{\em time}_n + 1.0 \cdot \hbox{\em
stddev}_n$$
and
$$\hbox{\em time}_{n-1} - 1.5 \cdot \hbox{\em stddev}_{n-1} < \hbox{\em time}_n +
1.5 \cdot \hbox{\em stddev}_n$$
\item High suboptimality, i.e., 3, when
$$\hbox{\em time}_{n-1} - 1.5
\cdot \hbox{\em stddev}_{n-1} \geq \hbox{\em time}_n + 1.5 \cdot\hbox{\em stddev}_n$$
\end{itemize}

\noindent}
We then sum this value over the Q@C pairs with different plans
(the change pairs) to arrive at a single integer for the query instance,
thus determining the {\em empirical suboptimality of that query instance}.

Note that while \azdb\ examines the plan at every cardinality, it only has
to actually execute the query at change
pairs.  Since many fewer {\QatC}s were involved, we could try many more
queries than the \hbox{Exhaustive} \hbox{Experiment}. In the Exploratory Experiment to be
described in Section~\ref{sec:experiments}, the value of Suboptimality ranged
from 0 (no suboptimality) to 133, with the majority between 0 and 9. A full 56\% of
the queries were suboptimal.  \hbox{Because} the occurrence of large values of this
measure was so rare, we did a log transformation: $\log_{10} (1 + \mbox{\em
  subopt})$ in the confirmatory analysis.

\section{Testing the Causal Model}
In this section we elaborate on how to test our model and discuss the test
results. Specifically, we describe in detail the environmental configurations,
data sets, and various experiments that we utilized. We then provide
descriptive statistics of the experiments and present the results of
correlational and regression analyses.

\c2j{}{\subsection{Experimental Setup}\label{sec:setup}}
\shorten{As mentioned, we used \azdb\ to run the experiments. This infrastructure,
now comprising around 44K lines of code, handles all the details of
generating the sample data, generating the actual queries, collecting the
various plans generated by the optimizer, running the queries, and storing
the results in a ``labshelf'', a set of tables managed by
a separate \hbox{DBMS}. This infrastructure also performs the many steps
required to ensure repeatability.

We have assembled a hardware lab of dedicated
machines, one \c2j{for each DBMS}{each for the three \hbox{DBMSes}} and one to run the \hbox{DBMS} used
to store the lab shelves. Having dedicated hardware allows us to worry
less about other processes running on the machines that could dirty the
results (these machines are {\em only} for experiments), and allows us to
run extensive experiments involving days or weeks of computation. We have
used this software and hardware lab to perform a number of experiments
examining the phenomenon of suboptimality.
}
\shorten{{\azdb} utilizes experiments defined in XML. It supports {\em experimental
  scenarios}, which are short (perhaps 100 lines of code) Java subclasses
that provide details of how to run kinds of experiments.}

The measurements were collected using Tucson Timing
Protocol\linebreak Version~1 (TTPv1)~\cite{TTPv1} and Version~2 (TTPv2)~\cite{TTPv2} on a suite of five machines, each an
\hbox{Intel} Core i7-870 Lynnfield 2.93GHz quad-core processor on a LGA 1156 95W
motherboard with 4GB of DDR3 1333 dual-channel memory and Western Digital Caviar
Black 1TB 7200rpm SATA hard drive, running Red Hat \hbox{Enterprise} Linux
Server release 5.8 (Tikanga) for TTPv1 and release 6.4 (Santiago) for TTPv2, with a kernel of 2.6.32-358.18.1.  The protocol
provided calculated query evaluation time, including computation and I/O time,
in msec. Both protocols were utilized exactly as specified.

No run violated the experiment-wide sanity checks. For the largest
experiment, the Confirmatory Experiment (number 7) discussed below, approximately
10.5\% of the query executions (QEs) and 5.1\%
of the queries-at-cardinality (Q@Cs) were dropped due to query execution
and Q@C sanity checks. As a result, excessive variation in calculated query
time was observed in only 0.003\% of the Q@Cs. A total of 0.43\% of Q@C adjacent
pairs violated relaxed monotonicity and 0.74\% of the Q@Cs violated strict
monotonicity, which is acceptable.

\doubleblind{In fact, over the course of the research described here, we
  found that currently available techniques for measuring query time did not offer sufficient query time
  precision, and so we developed TTPv1~\cite{TTPv1} to provide that needed
  precision. As our research in suboptimality proceeded, we realized that
  to detect suboptimality for two adjacent cardinalities
  separated by only 10,000 rows (just a few database pages), or even in the case of
  MySQL, 300 rows, we needed even finer resolution. This provided the
  impetus for the massive effort to develop TTPv2~\cite{TTPv2}, which did
  indeed provide the needed precision.}{We
  thank the developers of this protocol and of the software that enabled the
  running of many, many queries for providing us this software.}

In these experiments, we  installed each disk directly
on the machine that also runs the DBMS, ensured a cold cache (disk drive,
disk controller, O/S, and DBMS buffer), and discarded any sequence of query
executions that appear to be the result of query result caching. We also
ensured within-run plan repeatability.

In the following, we describe in detail the data used by our experiments and
the experimental scenarios we defined. More details on both
can be found in Appendix~\ref{sec:app}.

\subsection{Data sets}\label{sec:datasets}
We generate our experiment data set randomly in each of the experiments. However, we
use seeds to control the random data generator so that it can produce
repeatable data as required.

Our experiment data set consists of relational tables. There are two types of
tables. The first is a ``fixed table'' that, once created and
populated, will never be modified in the future.  In contrast, the second
type is a ``variable table''. We alter the cardinality\c2j{ of
  this table}{, physically, of such tables} as the experiments are being performed. 
\c2j{We used a}{We generated four configurations for the data sets. The first configuration
is a small} data set with four tables, each with four integer-typed
attributes. The fixed tables were populated with one million rows and the
variable table with two million rows (sixty thousand rows for one DBMS). Thus, the size of the tables is roughly 16--32Mbytes, with tiny skew (cf.~Section~\ref{sec:datacomplexity}). We also produced versions of the data set
with (i)~small skew, (ii)~primary keys (of the first attribute), and (iii)~primary
keys and secondary indexes, for all tables, as detailed in Appendix~\ref{sec:appdatasets}.

\subsection{The Experiments}\label{sec:experiments}

We are interested in predicting the suboptimal behavior of \hbox{DBMSes} through our
model. We selected four relational
\hbox{DBMSes}, some open source and some proprietary, that are representative of
the relational \hbox{DBMS} market. Each was used in its stock configuration.

In each experiment, we varied the
cardinality from 2M (maximum) to 10K (minimum), in increments of 10K. For
the one \hbox{DBMS} that was slower than the others and was timing out for
the majority of the queries when run between 10K and 2M, we reduced the
size of the tables and varied the cardinality from 60K (maximum) to 300 
(minimum), in increments of 300.

Utilizing JDBC to manipulate independent variables from outside the
  \hbox{DBMS} allows us to empirically generalize by moving up the $y$ axis of
  Figure~\ref{fig:empirical}, from one system to several systems and then to
  a general theory. We don't reveal the identity of the \hbox{DBMSes} we studied, for two reasons. First, commercial \hbox{DBMSes}
include in their user agreements requirements not to release 
performance data. This is detrimental to science, but we have no choice but
to live with that restriction. However, in some sense the specific \hbox{DBMS}
doesn't matter, as we are studying phenomena about cost-based optimizers
{\em in general}, and so are interested in making statements that apply across
the experimental subjects in our study.

\begin{table}[t]
\tbl{Experiments 1--7: Detailed Run Statistics\label{tab:run_stat}}
{%
\resizebox{140mm}{!}
{
\begin{tabular}{c|c|c|c|c|c|c|c}
& {\em Experiment}& {\em Protocol} & {\em Cumulative} & {\em Number of}&{\em
    Number} &{\em Number}& {\em Number of}\\
& & & {\em Hours} & {\em Query Instances}&{\em of Q@Cs} &{\em of QEs}&{\em Retained QEs}\\
\hline
1 & Monotonicity 		& --- & 38 & 60 & 12,000 & 12,000 & 12,000\\
2 & Exhaustive		& TTPv1 & 1,672 & 160 & 32,000 & 320,000 & 244,787\\
3\shorten{4} & Exhaustive with Keys 	& --- & 28 & 200 & 40,000 & --- & ---\\
4\shorten{5} & Initial Exploratory 	& TTPv1 & 560 & 780 & 8,842 & 88,420 & 68,891\\%40512+26367+2012
5\shorten{3} & Refined Exhaustive 	& TTPv2 & 1,544 & 160 & 32,000 & 320,000 & 319,980\\
6 & Exploratory 		& TTPv2 & 1,663 & 1,200 & 12,560 & 125,600 & 114,377\\
7 & Confirmatory 		& TTPv2 & 11,375 & 7,640 & 99,558 & 995,580 & 890,631\\
\multicolumn{3}{c|}{\em Total}	& 16,880 & 10,200 & 236,960 & 1,861,600 & 1,650,666\\
\end{tabular}
}
}
\end{table}

We performed seven separate experiments, each looking at a different
aspect. \hbox{Table~\ref{tab:run_stat}} exhibits the statistics of running a number of 
queries in our experiments. All but the last column list the number of query
instances, etc., gathered by the protocol for each experiment. The last
column gives the number of query executions retained by the protocol. One of
the primary goals of TTPv2 was (i)~to reduce the number of query executions discarded due to
phantom processes that were indirectly detected and (ii)~to collect relevant measures 
directly associated with I/O, such as BlockIO Delay time~\cite{TTPv2}. 
Experiments 5, 6, and~7 benefited from that protocol. Other details about
the experiments can be found in Appendix~\ref{sec:otherdetails}.

The queries came from 16 query sets, summarized below with more details in
\hbox{Appendix~\ref{sec:querysets}}. It is important to emphasize that while the {\em queries} all came
from the same query pool and while the data sets were also shared by the
experiments (see the Appendix for details), the {\em query executions} for
the six experiments, except Experiment~3 with no timing and hence no QEs, 
are disjoint. As a side comment, we mention that for all four \hbox{DBMSes}, the query
plans generated by a \hbox{DBMS} for a particular \QatC\ of a particular query
varied between the experiments, but not between the QEs
of that Q@C, by virtue of the way \doubleblind{we designed the
measurement protocol}{the measurement protocol was designed}.

The first experiment, termed ``Monotonicity'', was described in
Section~\ref{sec:suboptimality}. This experiment ran quickly, as it
only involved 12,000 QEs. That experiment helped us realize that we needed
to be much more sophisticated in our approach to timing queries.

\doubleblind{After we developed our six-step TTPv1 protocol,}{When TTPv1
    became available,} we performed our
second experiment, termed \hbox{``Exhaustive'',} which more accurately tested the 
monotonicity assumption. 
This experiment involved 160 query instances, 32,000 \QatC s (thus the name:
we timed each query ten times at {\em all} cardinalities, generating 200
\QatC s for each query instance), and thus
320,000 QEs (ten QEs for each \QatC ), requiring 1,672 cumulative hours.  The (very small: 1.6\%)
percentage % (16312+8215)/(995321+587546) = 1.6%
of strict monotonicity violations observed was consistent with the remaining
variance of the query time measurement, concluding that none of
the \hbox{DBMSes} violated monotonicity. This also provides a validation of
our definition of suboptimality, which requires monotonicity.

\discont{We then used the plans generated from the Exhaustive Experiment to classify operators as
continuous or discontinuous, as discussed in detail in
Section~\ref{sec:discontinuity}. Again we note that this experiment used a subset of
the set of {\em queries} (but {\em not} query executions) from the query pool,
to ensure that we see the same operators as encountered in that
study.
}{}
We also ran an experiment (termed ``Exhaustive with Keys'') on the 
\hbox{Exhaustive} query set but using the data set with primary keys.\discont{ However, in this
case we did not actually execute
the queries (thus this third experiment was very fast), but rather just
examined the plans that were returned from the \hbox{DBMS} to identify
additional discontinuous operators (five more were found).}{}

The fourth experiment was an ``Initial Exploratory'' 
analysis of a prior version of the causal model.
That earlier model had fewer independent
variables yet also had more complex relationships. Specifically, the model
did not include the independent query complexity variables of Presence of
secondary indexes, Presence of subquery, nor the schema complexity
independent variable of Presence of secondary indexes, nor the plan space
complexity independent variable of Number of repeats. The prior model also
had Presence of primary key attribute in the schema complexity construct rather than
the query complexity construct. In reformulating that independent variable,
we were able to remove a complex mediating moderator between the plan
space complexity and suboptimality constructs.

%% The following shorten corresponds to TTPv1.
In the experiment we ran a representative sample of queries and data sets: (a)~600 queries of one
\hbox{DBMS}, (b)~120 queries of another \hbox{DBMS}, (c)~10 queries
from each other two \hbox{DBMSes}, all without primary keys defined, and (d)~10 queries
from each of the four \hbox{DBMSes} on the primary key data set, for exploration
across all combinations, a total of 780 query instances.  (Again, consult
Appendix~\ref{sec:querysets} for details on these query sets.)
We ran the \hbox{DBMSes} only on the
change pairs, unlike the Exhaustive experiments. The protocol retained about 78\% of the QEs.

As a large number of QEs (roughly a quarter) were discarded by TTPv1 in
these experiments, and as these discarded QEs were correlated with query
time (and thus perhaps with suboptimality),
\doubleblind{we sought to develop a better measurement protocol, which was
  the impetus for TTPv2~\cite{TTPv2}}{we switched to a better measurement
  protocol that had recently become available: TTPv2~\cite{TTPv2}}.
%Our subsequent experiments indeed benefited from the improved protocol, producing 
%highly lossless ($\sim$5\% Q@C drop) but very clean ($\sim$1\% violation) data.
The fifth experiment then was to rerun the Exhaustive experiment  (we term this simply
``Refined Exhaustive'') using that protocol. The main purpose of this experiment 
was to reexamine monotonicity within in the enhanced
protocol.

We observed that the monotonicity violation rate decreased
from 1.54\% to 1.02\% for strict monotonicity and from 0.61\% to 0.60\% for
non-strict monotonicity, thereby reaffirming our monotonicity assumption
(the first benefit). Moreover, the number of retained QEs went up
significantly (compare 244K to 320K retained QEs, a second and the primary benefit of TTPv2), 
while the number of hours required actually went down somewhat (a third benefit).
These benefits reassured us of the quality of data 
for subsequent experiments via the improved protocol.\discont{ Incidentally, no new
discontinuous operators were identified in this experiment.}{}

%% The following corresponds to TTPv2.
We subsequently reran the exploratory analysis using TTPv2, this time on a larger sample of queries and data sets:
(a)~200 queries of four \hbox{DBMSes} without primary keys defined and
(b)~100 queries from each of the four \hbox{DBMSes} on the primary key data set,
for exploration across all combinations, a total of 1,200 query instances. % (= 300 $\times$ 4 (DBMSes))
We termed this sixth experiment ``Exploratory''. This experiment 
retained 12,100 {\QatC}s (3.7\% were dropped), concerning 1,123 query
instances (6.4\% were dropped). Only 443 strict monotonicity violations (0.78\%) and 301 relaxed
violation (0.54\%) were observed. This exploratory analysis allowed us to refine
the operationalizations.

The seventh and final experiment was used for confirmatory analysis of the model and thus was
called ``Confirmatory''. This was the most time-consuming of
the experiments. Here we
used (a)~800 queries on the data set without primary keys defined and (b)~510 of
those queries that had joins on the primary key attributes, on the data set
with primary keys, 
(c)~100 queries on the data set with tiny skew and primary keys,  
(d)~100 queries with a subquery on the data set without primary keys,  
(e)~100 queries on the data set with primary keys and secondary indexes defined, 
(f)~100 queries with a subquery on the data set with primary keys defined,  
(e)~100 queries with a subquery on the data set with primary keys and secondary indexes defined, and 
(g)~100 queries on the data set with primary keys and secondary indexes defined, 
for a total of 1,910 queries and a total of 7,640 query instances (over the
four \hbox{DBMSes}). (Again, see Appendix~\ref{sec:querysets} for details.)
We ran the \hbox{DBMSes} at the 99,558 {\QatC}s that were observed,
roughly 13 per query. The protocol accurately timed
both sides of adjacent {\QatC}s (53,547 change pairs in all).

In this Confirmatory Experiment, we observed 3,347 strict monotonicity
violations (0.74\%), and 1,966 relaxed monotonicity violations (0.43\%) (out
of a total of 452,684 Q@C pairs having identical plans and query instance), which provides
further confidence that monotonicity also applies to operators found in
queries over data in various contexts (as described above) and that our
operationalization of Suboptimality is a valid one.

In the remainder of this section, we focus using the measured independent
and dependent variables in the Confirmatory experiment
to test the predictions that arise out of our causal model in Figure~\ref{fig:model}.

\subsection{Descriptive Statistics}
Several initial conclusions can be drawn from this confirmatory experiment, which was the
culmination of several years of programming effort and about 30 months of
experimental runs, summarized in Table~\ref{tab:run_stat}.

To illustrate the descriptive statistics with the Confirmatory
experiment, we started with 7,640 query instances. These
resulted in 99,558 Q@Cs, each with ten query executions (QEs). Some of these
QEs violated one or more of the (many) sanity checks specified within the
TTPv2 protocol, leaving a remaining 890,631
retained QEs. A few of the query instances then violated one or more of the
Q@C sanity checks in the protocols (657 in all), with 6,983 query instances retained. Additionally, 16 query instances were dropped {\em after} the
protocol because they were missing data for the only change pair in the
query.

We were then left with 6,967 queries, of which we make three general
observations. 
First, perhaps surprisingly, 
more than half (3,933 queries out of those that emerged) exhibited suboptimality
somewhere in the range of cardinality of the varying table. 
{\em Every} \hbox{DBMS} exhibits suboptimality.

Secondly, most (3,370) of those suboptimal queries (86\%) had the maximum value of
Suboptimality  (i.e., level 3) for at least one change pair.
\shorten{a big number for a query could be many pairs, each at a small
  distance, or a few pairs, each at level 3. How many queries had at least
  one pair with subopt>=3? Answer: 670 in 5.20, 36 in 5.3, 531 in 6.0. So,
  total of 1237; query used: select count(*) FROM (select distinct experimentid, experimentname, dbms, runid,
querynum from CF_Analysis_SuboptDefn_Ver1 where subopt>=3);
query used 7_10: select count(*) FROM (select distinct experimentid, experimentname, dbms, runid,
querynum from Cnfm_Analysis_SuboptDefn where subopt_sd>=3);}
The phenomenon of query suboptimality that we observed is likely to be a
fundamental aspect of either the 
\hbox{algorithm} (cost-based optimization) or the
creator of the algorithm (human information processing). Our model includes
both effects.

%\pagebreak
Third, concerning the causal factors of Suboptimality in our model,
\begin{itemize}
\item the cardinality of the effective plan space (CEPS) ranged from 1 to 24
plans across the cardinality range,
\discont{\item a total of 23 operators were classified as discontinuous, out of 26 in all,
\item the Number of operators with discontinuity observed in plans for a query
averaged 9.4, with 85 being the maximum,}{} and
\item the Number of repeats ranged from 0 to 108, with the mean being 4.99.
\end{itemize}

\subsection{Correlational Analysis}\label{sec:corr}
We tested Hypotheses 1--6 using the strength and significance of
correlations of variables involved. These hypotheses predict
\discont{21}{14} main effects
and \discont{9}{6} moderating effects. Table~\ref{tab:hyptest} lists the hypotheses
followed by the correlation observed when testing each hypothesis.  (``NS''
denotes not significant at the $p < 0.05$ level, the accepted standard for
significance. For each dependent variable, we used the Bonferroni correction
to control the familywise error rate. ``---'' denotes no prediction arising
from the model.)  As can be seen, most of the main effects (17) arising from
the causal model are supported and significant.  \discont{The four exceptions are
Hypotheses 1c, 2g, 2j, and 3, highlighted in italics in the table.}{The three exceptions are
Hypotheses 2e, 2g, and 3, highlighted in italics in the table.}

\begin{table}[t]
\tbl{Testing Hypotheses 1--7: Correlations on the Confirmatory Study\label{tab:hyptest}}{%
%\begin{center}
%\resizebox{150mm}{!}{
\begin{tabular}{c|c|c|c\discont{|c}{}}
{\em Variable}& {\em Suboptimality}&{\em Repeats}&{\em CEPS}\discont{ & {\em Discontinuity}}{}\\
\hline
\hline
Operators in \hbox{DBMS} & ---&  {\bf H1a:} 0.35 & {\bf H1b:} 0.27\discont{& {\bf
  H1c:} -{\em 0.09}}{}\\
\hline
Correlation names & --- & {\bf H2a:} 0.17 & {\bf H2b:} 0.46 \discont{&{\bf H2c:} 0.34}{}\\
\hline
Presence of aggregate & --- & \discont{{\bf H2d:} 0.04 & {\bf H2e:} 0.10 & {\bf H2f:} 0.29}{{\bf H2c:} 0.04 & {\bf H2d:} 0.10}\\
\hline
Presence of primary key attribute& --- & \discont{{\bf H2g:} {\em NS} & {\bf H2h:} 0.15 & {\bf H2i:} 0.14}{{\bf H2e:} {\em NS} & {\bf H2f:} 0.15}\\
\hline
Presence of subquery & --- & \discont{{\bf H2j:} {\em NS} & {\bf H2k:} 0.26 & {\bf H2l:} 0.20}{{\bf H2g:} {\em NS} & {\bf H2h:} 0.26}\\
\hline
Presence of skewed data & {\bf H3:} {\em -0.05} & --- & ---  \discont{& ---}{}\\
\hline
Number of repeats& {\bf H6a:} 0.62 & --- & ---  \discont{& ---}{}\\
\hline
CEPS &  {\bf H6b:} 0.61 & {\bf H4b:} 0.45 &  --- \discont{&{\bf H4a:} 0.72}{} \\
\discont{\hline
Number of operators with discontinuity & {\bf H6c:} 0.47 & --- & --- & ---\\}{}
\end{tabular}
}
%\end{center}
%}
\end{table}

\discont{Hypothesis 1c was not supported because the
correlation between Number of operators available and Number of operators
with discontinuity was negative, while we predicted positive
correlation. The strength of the negative correlation was small:
\hbox{-0.09}.

In looking into this further, we noticed that the number of operators deemed
discontinuous by our operationalization in Section~\ref{sec:discontinuity}
was quite high: 26 out of 29 operators in all. Note that the
operationalization utilizes the {\em estimated cost} of each instance of a
particular operator within an observed plan. (The reason is that we could
not directly measure {\em per-operator run times}: we were only able to measure
{\em per-plan run times}, that is, for each Q@C.) The estimated cost reported by
the DBMS sometimes exhibited a lower estimate for the same operator within the
exact same plan at a slightly higher cardinality, which is
counter-intuitive. These findings
indicated to us that perhaps our \hbox{operationalization} of ``discontinuous
operator'' incorrectly classified some operators, due to the vagaries in the
estimated operator costs, and thus this negative correlation is also in doubt.

Hypotheses ~2g and 2j, both involving Number of repeats, were not significant.}{}

Hypothesis 3 was
not supported because the correlation between Presence of skewed data (recall from
Section~\ref{sec:datacomplexity} that absence of skew is operationalized as
{\em tiny} skew and presence of skew is operationalized as {\em small} skew)
and Suboptimality was in the \hbox{opposite} direction. Specifically, we observed a negative correlation,
indicating that as the skew increases from tiny to small, Suboptimality
decreases, though only slightly:~\hbox{-0.05}.

In Table~\ref{tab:hyptest5}, we examine the interaction strength of
Hypothesis 5, between the Query Complexity and Plan Complexity constructs, when
the moderator (Presence of secondary indexes) has a value of absent (``Not
SI'') and when it has a value of present
(``SI''). As before, ``{\em NS}'' denotes no significant at the $p < 0.05$
level; three of these interactions become not significant when secondary
indexes are added. (Recall that it is not possible to compute the
interaction strength for Presence of primary key attribute in the presence
of secondary indexes.)

\begin{table}[t]
\tbl{Testing Hypothesis 5: Interaction Strength on the Confirmatory Study\label{tab:hyptest5}}{%
%\begin{center}
%\resizebox{85mm}{!}{
\begin{tabular}{c|c|c|c|c\discont{|c|c}{}}
{\em Variable}& \multicolumn{2}{|c|}{{\em Repeats}}&\multicolumn{2}{|c}{{\em CEPS}} \discont{& \multicolumn{2}{|c}{{\em Discontinuity}}}{}\\
%\hline
{\em }&{Not SI\em }&{SI\em }&{Not SI\em }&{SI\em }\discont{&{Not SI\em }&{SI\em }}{}\\
%&{\em }&{\em }&{\em }&{\em }\\
\hline
\hline
Correlation names & 0.17 & 0.15 &0.49 &0.39 \discont{& 0.36 & 0.27}{} \\
\hline
Presence of aggregate  & 0.04 & {\em NS} & 0.12 & {\em NS} \discont{& 0.31 & 0.22}{} \\
\hline
Presence of subquery  & 0.03 & {\em NS} & 0.18 & 0.17  \discont{& 0.14 & 0.13}{}\\
\end{tabular}
}
%\end{center}
%}
\end{table}

This table shows that the strength of all \discont{nine}{six} applicable and significant
interactions goes {\em down} just a little (on average by 0.04), which is
opposite to our prediction. Hence, this negative moderation seems small but
robust. One possibility is that rather than the added schema complexity of
indexes increasing suboptimality, secondary indexes may have {\em reduced} the complexity of query
optimization, since the role of secondary indexes within query optimization
is well understood, thereby reducing suboptimality.

In summary, most (\discont{19 out of 21}{12 out of 14}) of the main effect hypotheses were
significant at the $p < 0.05$ level. \discont{Two of those hypotheses were}{One of those hypotheses was} (weakly)
in the opposite direction. The hypothesized moderation (Hypothesis 5) of
Presence of secondary indexes was small but in the opposite direction. Thus,
most of the hypotheses in our model were strongly supported.

\subsection{Regression Analysis}\label{sec:regression}
One use of regression is as a further test of Hypothesis 5, which predicts a (positive)
moderation in the presence of query complexity, specifically the Number of
correlation names, Presence of aggregate, and Presence of subquery
independent variables. Thus our causal model predicts that the interaction
strength should increase when secondary indexes appear in the relational
schema. (The {\em direction} of moderation, whether positive or negative, is not involved in this test; rather,
an increase in interaction strength only supports that there is a moderating
effect, in either direction.) When
secondary indexes were not specified, our model explains 
7.23\% of the variance of Number of repeats\discont{,}{ and} 40.9\% of the variance of CEPS\discont{, and 35.7\% of the variance
of discontinuity}{}. When
secondary indexes were defined on the underlying tables, our model explains
7.23\% of the variance of Number of repeats\discont{,}{ and} 41.1\% of the variance of CEPS\discont{, and
38.9\% of the variance of discontinuity}. These findings are all consistent with our
hypothesis predicting a moderation effect of secondary indexes.

A second use of regression is to compute the amount of variance explained
for the intervening and dependent variables, to indicate how much of the
underlying phenomena are explained holistically by the model.
\shorten{We did regressions on the causal variables for CEPS, Number of repeats and for Number of
operators with discontinuity. Our model explained 41.1\% of the
variance for CEPS, 7.23\% of the variance for Number of repeats, and 38.9\% of the variance for
discontinuity. \shorten{need path analysis}

} So we also ran a regression over the independent variables of the model that
predict suboptimality over the data from the Confirmatory experiment.

Note that our experiment was over four highly complex DBMSes, each with
hundreds of thousands to millions of lines of source code, none shared.
Thus, one might expect the amount of variance explained to be low, with a
considerable amount of variance to be {\em within} the DBMS. What we are getting
at with our causal model is the impact of {\em shared} aspects that arise
from supporting (a)~a common data language (the relational model), (b)~a common
language (SQL), (c)~a common query evaluation approach (operators in the
relational algebra), and (d)~a common query rewriting approach (cost-based query
optimization). Given these basic commonalities, each DBMS supports a
certain number of operators, each is impacted by certain characteristics of
the schema, query, and data, and each generates a certain number of candidate
plans.

Our
model explained \discont{53.6\%}{52.2\%} of the variance of the Suboptimality dependent
variable. That means that over half of the variance of Suboptimality is
explained by our operationalizations of Optimizer Complexity, Schema
Complexity, Query Complexity, Data Complexity, and Plan Space Complexity.

To state this a different way, all other possible causes for Suboptimality
will in concert have less predictive power than the constructs and specific
variables included in the causal model introduced here.

And somewhat extraordinarily, it is the common aspects listed above that predict
Suboptimality, {\em not} the particulars embedded in the inordinate complexity of
each of these DBMSes.

\shorten{
In our experiments, we noticed that some of the \hbox{DBMSes} might at
different times pick a different plan for the same query at the same
cardinality. We addressed this in our experiment by running the optimizer
(with a JDBC {\tt PreparedStatement}) and then executing the query ten times
in fast succession. Each time we did an {\tt EXPLAIN PLAN}  to ensure that
the originally-chosen plan was indeed used (it was).

That said, at the next cardinality, the optimizer went into action again,
and so could have chosen a {\em different} plan for that new cardinality.
Our supposition is that such a {\em heuristic} optimizer introduces
some variance in the cardinality dependent variable.

To test this, we ran regressions on just the \hbox{DBMS} that was demonstrably {\em
  not} heuristic, for the common case of primary keys. Interestingly, only
the Presence of aggregate and CEPS were significant at the 0.05 level, yet
our model explains fully \% of the Suboptimality variance (and \% of the
CEPS variance and \% of the discontinuity variance). This might imply
that heuristic optimization itself introduces perhaps 10--20% of the variance.

The size of the regression coefficients also provides useful
information. From the regressions run on the data, for the
statistically significant causal factors, the highest coefficients, all in
predicting the Suboptimality dependent variable, were CEPS at , number
of tables involved at , number of discontinuous variables at ,
number of operators available in \hbox{DBMS} at . Note though that number of correlation
variables is highly correlated with CEPS; the regression will apportion the variance of the dependent
variable among these factors.
}

\subsection{Summary of Model Testing}
In our experimental design, we started with a structural causal model that
encapsulates our theory for how cost-based query optimizers might select a
suboptimal plan for a query at a cardinality. This model implies the six
specific hypotheses listed in Section~\ref{sec:model}. We then performed
a series of experiments,
\begin{itemize}
\item to refine our operationalizations: specifically \discont{discontinuity,
via the Exhaustive \hbox{Experiment}, and} Number of operators available, via the
Exhaustive with Keys \hbox{Experiment},
\item to test fundamental assumptions, specifically
monotonicity, via the Monotonicity and Refined Exhaustive \hbox{Experiments}, and
\item to test and make minor
refinements to our model: the Initial Exploratory and \hbox{Exploratory}
Experiments.
\end{itemize}
During this exploration, \doubleblind{once we completed the TTPv2 protocol,}{when the TTPv2 protocol became available,} we reran
\hbox{experiments} to avail ourselves of the increased precision due to a much
higher percentage of retained QEs and thus Q@Cs and query instances.

Throughout these six experiments, we were cognizant of the
possibility of Type 1 errors: false positives that lead one to believe a
relationship exists when it doesn't.
To control for such errors, we then performed the final Confirmatory 
Experiment on a completely different data set consisting of many more 
query instances, 7,640 in all, running on four \hbox{DBMSes} that each
utilize cost-based query optimization, to test
our refined model. Statistical inference is only possible in confirmatory
analysis, where the model and hypotheses are selected a priori.

Now that the causal model has been found to be supported by the confirmatory
\hbox{analysis}, we turn to possible implications of this model.

\subsection{Identifying Root Causes of Suboptimality}\label{sec:root}

Our goal in this paper has been to understand cost-based query optimizers
as a \hbox{\em general} class of computational artifacts and to articulate and test
a predictive model \hbox{characterizing} how such optimizers,
again, as a general class, behave.  This model can be used to further
improve \hbox{DBMSes} through engineering \hbox{efforts} that benefit from the fundamental
understanding that the scientific perspective can provide.

Our model includes a number of causal factors of suboptimality. Of these
factors, the regression \hbox{coefficient} that was highest was for Number
of repeats (0.46, normalized). The next \hbox{highest}
regression factor was CEPS (0.27, normalized). These two observations imply that the
number of plans being considered is a major determinate of
suboptimality, indicating that choosing among many plans is hard, despite
many decades of research and development.

The next most influential factor is \discont{the Number of operators with discontinuity
(0.176, normalized),
\hbox{implicating} the cost model. The final factor is}{} Presence of skewed
data (-0.057, normalized), a small effect that has the salutary effect of
decreasing suboptimality.

These factors implicate two broad root causes of suboptimality across \hbox{DBMSes}:
(i)~the plan search process. and (ii)~the cost model.

\section{Diminishing Returns?}\label{sec:diminishing}
One oft-used way to increase DBMS query performance is for the developer to
add to the DBMS source code another physical operator that can be then used
in query evaluation plans~\cite{Graefe93}.

For example, nested loop join was probably the
first implementation of the relational join algebraic operator. B-tree index join
was only possible when such indexes were added (quite early in the history
of modern DBMSes). Adding hashing to a DBMS enables a hash join operator;
adding hash indexes enables (perhaps several physical variants of)
hash-index join.

Each
subsequent generation of the \hbox{DBMS} thus supports an ever-expanding
collection of physical operators, with the current incarnation the most recent
within a series of \hbox{DBMS} {\em generations}. (The study discussed in
this paper above was over a recent generation of each DBMS.)

With this new physical operator
available, the set of possible plans is expanded, and for certain
combinations of query and data, there could be a plan that uses that new
operator that is faster than any plan that can be expressed without that
operator. In any case, the performance of the best plan for each query/data
combination won't get slower, because all prior plans are still
available. That is why the number of physical operators grows over the
releases of a DBMS.

However, this ideal behavior might not always be seen in practice, for
several possible reasons. One is that it may not be practical to enumerate
all possible plans, and so the fastest plan may not even be
considered. Indeed, query optimization is \hbox{NP-complete}, even when using only
one physical join operator~\cite{Ibaraki84}. Also, because the optimizer is
choosing a plan based on its estimated query execution time,
the optimizer ight not choose the plan with the fastest
actual execution time. Thus, if the new plan, with that additional operator,
is estimated to be faster than the current chosen plan, the new plan may be
chosen even though its actual execution time may be slower than another
candidate plan.

This raises a central question: {\em does an additional operator made
  available in a release of a DBMS to speed up some queries actually help or
  hurt the overall performance of that DBMS?}

\subsection{A Gedanken Experiment}\label{sec:gedanken}
Consider a Gedanken experiment over DBMS generations. The experiment
considers the plan selected for each query at each possible cardinality,
that is, each Q@C, for each \hbox{DBMS} generation that adds an
  operator. In early generations, there will be few plan changes for a
given query as the cardinality varies, 
simply \hbox{because} there are few operators available. For
later generations, some Q@Cs will be associated with different plans
enabled by the new operators that were added.

In many (hopefully most) cases, a new plan selected by a generation is
more efficient than the plan selected by the immediate previous
generation. After all, that is the very reason the new operator was
added to that subsequent generation.  However, sometimes that \hbox{DBMS} generation's query optimizer selects a slower plan.
Indeed, the query optimizer also evolves and improves with each new
generation, in part to minimize the chance of selecting a slower plan.

Our predictive model in Figure~\ref{fig:model} suggests that Suboptimality
is causally impacted by the Plan Space Complexity construct, specifically, the
constituent variables of CEPS (cardinality of effective plan
space)\discont{,}{ and} Number
of repeats\discont{, and 
Number of operators with discontinuity}. This Plan Space \hbox{Complexity} construct
is naturally connected to the Number of operators available. Specifically,
our Confirmatory Experiment showed that the one variable of Number of
operators available accounted for 9.21\% of the variance of Suboptimality.

A possible implication of our validated causal model
is that as the evolution of generations of the \hbox{DBMS} adds more and more
operators, suboptimality will increase.

It is important to emphasize that our causal analysis to this point has been
across multiple \hbox{DBMSes}, each with a different set of
operators and even number of operators. Let's reprise from Section~\ref{sec:intro} on
page~\pageref{sec:intro}: our goal thus far has been ``to \hbox{understand}
cost-based query optimizers
as a {\em general} class of computational artifacts and to come up with
insights and ultimately with predictive theories about how such optimizers,
again, as a general class, behave.''

The present discussion has a quite different focus. Here we're talking about
the query optimizer across multiple DBMS
generations, each generation adding one or more operators. (That said, as
we'll show shortly, we can still study this phenomenon of increasing
suboptimality across DBMSes; this remains a scientific question.)

As a refinement, let's assume each subsequent
generation adds a single operator. Starting with a fixed set of queries,
for each DBMS, we run each generation on each Q@C and then sum up the query
times to compute a {\em per-generation (total) time} for that DBMS. We can
also sum over the four DBMSes, to see how this  {\em overall per-generation
  time} (a single number for each generation number) varies with generation.

The underlying question then becomes, does an additional operator made
available in a subsequent generation of the \hbox{DBMS} actually help or
hurt? More fully, is the predicted increase in suboptimality originating
from that added operator, a causal effect discovered and validated by the process
of {\em science}, specifically the application of \hbox{empirical} generalization illustrated
in Figure~\ref{fig:empirical}, compensated for by the increased performance
afforded by that operator, a benefit realized by the application of {\em
  engineering}, the articulation, elaboration, and implementation of new
storage structures and query evaluation algorithms?

Fortunately, the engineering has already been done: current DBMSes have at
their disposal multiple query operators, which have been perfected over the
decades. Adopting the engineering perspective would predict that the overall per-generation
time would monotonically decrease as the generations add query evaluation
operators.

Interestingly, adopting the scientific perspective reaches a different conclusion. Our
causal model in Figure~\ref{fig:model} predicts that as the number of
operators increases, the intervening measures of plan space complexity increase,
which therefore also increase empirical suboptimality. As the generations of the
\hbox{DBMS} contain successively greater numbers of operators, suboptimality
will also increase with generation number.  Given that each new operator
will improve a {\em shrinking}
subset of Q@Cs for any given query, while slower plans can emerge over an
{\em expanding} subset of Q@Cs, our causal model implies that performance
will improve with each successive generation, but then
start to level off as optimizer missteps become more prevalent.
The causal model predicts a point where the
\hbox{increase} in execution time due to further missteps obviate the decrease
enabled by the new operator. This seems to be a fundamental limitation
inherent in cost-based query optimization, due to the inherent inaccuracy of
the cost model and the difficulty of enumerating all plans for queries over
many table.

Put in simpler terms, engineering considerations alone imply that DBMSes can
continue to improve, whereas also adopting the scientific perspective predicts
that eventually each DBMS will hit a wall, beyond which improvement is not possible.

\vspace{1ex}
{\em Does such a limit exist, and if so, how close are modern \hbox{DBMSes} to that limit?}

\subsection{Simulating \hbox{DBMS} Generations}
While we do not have access to prior generations of our \hbox{DBMSes} (which is why
the previous discussion was in the form of a Gedanken experiment), we can
use the data already collected on the
current version available for each \hbox{DBMS}, to {\em simulate} the prior
generations, each successively having one additional operator, and thus a
larger set of realizable query plans. (Note that we use the most recent query
optimizer in all of the simulated generations; only the available operators
will vary across generations.)

In the next section, we will explain how we characterize the generations of
each DBMS. For now assume that each DBMS is associated with a series of
generations, with each generation having one more operator. So generation 1
of the DBMS has just one operator, generation 2 has two operators, etc.

Our data consists of the 8,840 query instances and their 112,118 Q@Cs from
the exploratory and confirmatory experiments (Experiments 6 and 7) of
the previous study. Specifically, we focus on the adjacent (that is,
separated by the minimum cardinality, either 10,000 or 300 rows) Q@Cs within
change pairs 
(let's refer to these as the {\em lower} Q@C and the {\em upper} Q@C,
and thus we also have a {\em lower plan} and an {\em upper plan}) for the
same query at the {\em lower} and {\em upper} cardinalities, each with an actual execution time. All Q@Cs for that query and
for that DBMS having cardinalities between the upper Q@C of one change pair and the
lower Q@C of the next (higher) change pair are associated with that same plan,
guaranteed by the process in which we chose those Q@Cs for
actually timing (recall from Section~\ref{sec:motivation} that we start from
the highest cardinality, that of 2M tuples, looking for changes in the
plan).

For this generational experiment, we associate with each Q@C a generation
(a positive integer) that
is the earliest generation of the DBMS that contains all the operators in the
plan associated with that Q@C. We also only consider change pairs where the lower
plan has a generation distinct from that of the upper plan.  Say the upper
generation is earlier than that of the lower plan, as illustrated in
Figure~\ref{fig:negback}. In this figure, as we scan from right to left in
decreasing cardinality, we first encounter Plan A from generation 2 at the
highest cardinality of 2M rows, then later at 1320K rows (this is the
measured Q@C that has Plan A that is closest to the measurement at 910K
rows), then a change pair with Plan A at 910K rows and Plan B at the
adjacent 900K rows, then later Plan A again at 30K rows. (We show only a few
Q@Cs in this example; there are other change pairs that are not relevant to
our discussion.)

\begin{figure*}[t]
%\centering
%originally 30pc
\includegraphics[width=30pc]{figures/negback.pdf}
\caption{An Example of a Negative Backward Delta\label{fig:negback}}
\end{figure*}

In this case, the thinking goes that, had we been
in the \hbox{DBMS} generation 2, the optimizer would have selected Plan A
throughout, because Plan B simply wasn't possible (as it involves an
operator not present in generation 2). Then when generation 3 was created by
adding an operator, the optimizer chose Plan B for 900K. The reason Plan B
was chosen is that it was faster than Plan A at that cardinality, shown in
the figure by extrapolating the run time down from 910K to 900K (we will
revisit this extrapolation shortly).

In this particular case, as just mentioned, Plan B is more appropriate (faster) than Plan A,
but that is not the only possibility. Sometimes the later generation with
more operators available chooses a plan that is slower
than the plan chosen by the earlier generation, due to the suboptimality
we've observed. (We'll examine an example shortly.)

The question then becomes, does a plan change by a subsequent generation
(\hbox{enabled} by the additional operator in that generation) represent a win (runs faster)
or a loss (runs slower)? More broadly, do the Q@Cs
in the aggregate enabled by each successive generation continue to overcome
the increasing burden of suboptimality?

\subsection{Characterizing \hbox{DBMS} Generations}
To which generation do we assign each \hbox{DBMS} operator?

Note that we don't actually know the specific order in which the operators
were added to each \hbox{DBMS}. But even if we did, that order was somewhat
arbitrary, with a host of considerations going into those decisions over the
years. Given that we are using the same optimizer for each such defined
generation, we'll adopt a more systematic ordering of the operators. We
start by gathering all single-operator plans, all two-operator plans,
and so forth, and order the generation by the prevalence of their
\hbox{appearance} in these query plans.

Specifically, for each \hbox{DBMS} we designate the first
generation to contain the single operator that maximizes the number of plans
at change pairs, that is, maximizing the number of Q@Cs, containing just that
operator. So for example, all the plans generated by one DBMS that have exactly
one operator involve just the Full Table Scan operator.  So no choice was
needed: we designate the
first generation as just having that one operator. Any plan with just that
operator can be constructed by generation 1, as well as by any
subsequent generation (as each generation includes that initial operator).

We then examine the plans containing exactly two operators.
Using DBMS A again, there are two such: one with the Full Table Scan and
Full Table Scan with Join operators (888 Q@Cs) and
one with the Full Table Scan and Ref operators (112 Q@Cs). Full
Table Scan with Join is thus the operator added by Generation 2, given its
prevalence of Q@Cs. Each
subsequent generation adds that operator that maximizes the number of plans
that operator will eventual enable. So Generation 3 adds the Eq\_Ref operator,
as that operator enables 633 plans eventually. Generation 4 adds the Ref
operator and Generation 5 adds the Index operator.

The generations thus can be characterized from the Q@Cs we encountered: five distinct combinations of two
operators, covering 1086 Q@Cs, six combinations of three operators (1126 Q@Cs),
two combinations of four operators (104 Q@Cs), and exactly one
combination of all five operators (6 Q@Cs).\shorten{ Thus for MySQL our data implies
a simulation of five distinct generations, each adding an operator.

\begin{align*}
g_1 &= \{ \hbox{\em full table scan} \}\\
g_2 &= \{ \hbox{\em full table scan}, \hbox{\em scan with join} \}\\
g_3 &= \{ \hbox{\em full table scan}, \hbox{\em scan with join}, \hbox{\em eq\_ref} \}\\
g_4 &= \{ \hbox{\em full table scan}, \hbox{\em scan with join}, \hbox{\em
  eq\_ref}, \hbox{\em ref} \}\\
g_5 &= \{ \hbox{\em full table scan}, \hbox{\em scan with join}, \hbox{\em
  eq\_ref}, \hbox{\em ref}, \hbox{\em index} \}
\end{align*}}
For the four \hbox{DBMSes} in our study, the number of generations ranged from five
to thirteen.

\subsection{Number of Change Pairs Per Generation}\label{sec:CPQ}
As an illustrative example of how we can look at DBMS query optimizer
performance through the lens of change pairs, let's consider  the {\em
  maximum number of change pairs per query}, or {\em maximum CPQ}, from the
perspective of
\hbox{DBMS} generations.

For each query, we count the number of Q@Cs at each
generation, so for instance a query might have five Q@Cs at generation 1, seven Q@Cs at
generation 2, and 17 Q@Cs at generation 5. We then then take the maximum CPQ
over the queries (so perhaps 17 is the maximum for generation 5 over all
queries). Our hypothesis is that as the generation increases, the maximum
query flutter will increase, which then influences the maximum CPQ. The results are shown in Table~\ref{tab:maxcpq}.

\begin{table}[h]
\tbl{Max Change Pairs Per Query, Per Generation\label{tab:maxcpq}}{%
%\begin{center}
\begin{tabular}{c|cr}
{\em Generation}&{\em Maximum Change Pairs}\\
& {\em Per Query}\\
\hline
1 & 1\\
2 & 66\\
3 & 115\\
4 & 136\\
5 & 132\\
6 & 141\\
7 & 42\\
8 & 172\\
9 & 2\\
10 & 12\\
11 & 55\\
12 & 29\\
13 & 1\\
{\em Cumulative} & 172\\
\end{tabular}
%\end{center}
}
\end{table}

What we notice is that the maximum CPQ does increase with generation,
up through generation 8, after which it falls off dramatically. In
retrospect, this fall-off makes sense. There can be only 200 Q@Cs for a
query, because we look for query plan changes only every 10,000 (300 in one
DBMS) tuples down from
a maximum of 2M (60,000) tuples. As the generation increases, there is less ``room''
for more plans (as some fraction of the previously generated plans will
still be quite good or even optimal), and so less opportunity for flutter
which will show up as a large maximum CPQ. And indeed, at generation 8,
there is a query that has an astonishingly high number of change pairs,
172, all waffling between plans within that generation.
This analysis indicates that there is something concerning change pairs that seems to get
critical around generation 8. We'll return to this in Section~\ref{sec:hitthewall}.

\subsection{Using Change Pairs}\label{sec:usingCP}
We now consider how change pairs can be used to evaluate the
effectiveness of different generations of a \hbox{DBMS}.

Define $\hbox{\em ops}(p)$ for a given plan $p$ to be the set of operators
present in that plan, with some operators perhaps repeated in that plan. 
A plan $p$ is {\em applicable} to a DBMS generation $g$ (denoted by a set of
operators) if $\hbox{\em ops}(p) \subseteq g$. By
definition, if a generation is applicable to a given plan, it is
applicable to all subsequent generations, with one being the 
{\em earliest applicable generation}, or {\em mingen}.
\shorten{
\shorten{We couldn't make this one work...
\todo{Young:replace in the above para with an actual SQL example query that was associated
  with two MySQL generations in a Q@C pair. use the generation numbers from
  the next para}

\todo{Rick: I found one example but the cardinality information 
was different than what's stated in the current prose...}

%%runid = 1097, querynum = 46, max_card = 60000, upper_card = 29100, lower_ard = 28800, upper_gen = 4, lower_gen = 5
%%max_cqt: 230, upper_cqt: 110, lower_cqt: 120, lower_extrapolated_cqt: 110
%% experimentname: op-pk-30K-100q-idx, pk: true, sec_idx: true
%% plan 0: ALL:Full Table Scan,ref
%% plan 1: eq_ref,index,ref
%% plan 0 at upper_card, plan 1 at lower_card
\noindent
\hspace{3ex}{\small\begin{verbatim}
        SELECT t3.id2, t3.id4, t0.id3, t2.id1 
        FROM ft_HT3 t2, ft_HT1 t1, ft_HT2 t3, ft_HT1 t0  
        WHERE  (t2.id2=t1.id2 AND t1.id2=t3.id1 AND t3.id1=t0.id4)
\end{verbatim}}
}
}
For each change pair, containing adjacent Q@Cs, we have either one or two
generations to consider. We focus on change pairs with two generations, and start with
those for which $\hbox{\em 
mingen}(\hbox{\em lower}) > \hbox{\em mingen}(\hbox{\em
  upper})$. An example is shown in Figure~\ref{fig:posback}. In this case,
the set of operators in Plan A for the Q@C at cardinality 900K requires at
least generation 3, whereas the set of operators for Plan B for the Q@C at
cardinality 910K requires generation 2. (Note that the generations in
this example
are consecutive, but that is not required. Sometimes the generations within a
change pair are quite different, such as generation 2 at the lower plan of a
pair but generation 5 at the upper plan.)

\begin{figure*}[t]
%\centering
%originally 30pc
\includegraphics[width=30pc]{figures/posback.pdf}
\caption{An Example of a Positive Backward Delta\label{fig:posback}}
\end{figure*}

To examine the wisdom of picking Plan A for 900K (whose run time was measured as
the higher point pair shown, the one above the dashed line), we find the closest Q@C
also having Plan A. There are two illustrated here, at cardinalities 30K and
2000K, respectively, with the closer one at 30K. (If there is no such Q@C,
we can't do the extrapolation, and simply remove that change pair from
further consideration.)

We then extrapolate the measured time for the closest plan of the lower
generation to the cardinality of the adjacent higher generation. In this
particular case, we use the measured query time of Plan
B at cardinalities 30K and 910K to extrapolate an estimated query time at 900K.
This realizes an estimate of how fast the query using Plan B would have run
on the database having the variable table with a cardinality of 900K. In
this particular case, Plan B looks like it would have run {\em faster} (in
fact, was faster even at 910K in this particular example).

\discont{As illustrated in Figure~\ref{fig:discontinuity}, plans can be
discontinuous, an example being external merge sort. For that operator}{Note
  that for some operators, such as external merge sort}, as
the cardinality increases, an additonal pass may at some point be needed,
but the query time will be roughly linear for a set number of passes.
Returning to the situation in Figure~\ref{fig:posback}, the presence of a change in the number of passes, say between 30K and 910K,
will have the effect of flattening the slope slightly for Plan B, such that
this extrapolation will slightly underestimate the run time of
that plan at 900K, and thus overestimate the penalty of going with
Plan A rather than Plan B. That said,
given that we are extrapolating from the actual run time at a very
close cardinality, our extrapolated estimate should be similarly close.

We term this situation a {\em backward extrapolation}, because we are extrapolating
backward from a cardinality of 910K to one of 900K, within the change pair.

With this extrapolation, we can compute, for each change pair, the {\em relative delta}, defined as the
measured time of Plan A (the chosen one) at the lower cardinality (here, 900K) minus the
extrapolated time of Plan B at that same cardinality, divided by the
larger of the measured and extrapolated time of the plan at the higher
generation (here, Plan A). The relative delta is thus scaled by original
measured time at that granularity, and thus has a maximum possible value of 1.

In this case, the relative delta is positive (the measured time for the plan
associated with the higher generation is greater than the extrapolated
time), which indicates that the optimizer chose a slower plan: here,
Plan B should have been chosen.

On the other hand, a {\em negative} relative delta (where the measured
time for the plan associated with the higher generation is lower than the
extrapolated time, such as that illustrated in Figure~\ref{fig:negback})
implies that the additional operator(s) available to the minimally
applicable generation of the upper plan were indeed beneficial, in that that
plan was faster. (In such cases, we again divide by the larger
value, so the minimum possible value is~-1.)

Summarizing, this analysis computes, for a {\em change pair} (a pair of
adjacent Q@Cs for a specific query running on a specific DBMS), a {\em
  relative delta} for the earliest applicable generation. A
positive relative delta (extrapolated in either the forward or backward
direction) reflects that a slower
plan was chosen by the \hbox{DBMS} generation, perhaps due to the greater number of
operators available. A negative relative delta (forward or backward) reflects
that the chosen plan was faster at that cardinality than the one
chosen at the adjacent cardinality.

There are ten orthogonal possibilities for each change pair (that
is, a pair of adjacent Q@Cs), a total of 66,792 change
pairs from the Exploratory and Confirmatory Experiments. (There is also the
case of a query instance containing a lone Q@C, meaning that only one plan was
chosen across all 200 Q@Cs. This occurred for 2126 out of the 8840
query instances, which we don't consider further.)

\begin{enumerate}
\item The pair of Q@Cs share the same generation: $\hbox{\em mingen}(\hbox{\em lower}) =
\hbox{\em mingen}(\hbox{\em upper})$  (62,903 change pairs, or 94.2\% of the
total), which we don't consider further.

\item The extrapolation yielded a computed query
time that was negative, which we also drop (10 change pairs, 0.01\%).

\item The extrapolation was from above and indicated
no suboptimality, termed a {\em negative backward relative
  delta}, as
  exemplified in Figure~\ref{fig:negback}, examined earlier (583 change pairs,
  0.9\%).

\item The extrapolation, termed a {\em positive backward
    relative delta} and exemplified in Figure~\ref{fig:posback}, indicated
  a suboptimal plan (492 change pairs, 0.7\%).

\item The extrapolation was a {\em negative forward relative delta},
  indicating no suboptimality (795 change pairs, 1.2\%).

\item A extrapolation was from below (consider Figure~\ref{fig:negback} but with
  Plan A from generation 5; we would then need to extrapolate from the
  closest Plan B, which is {\em at a smaller cardinality}), in a {\em forward
    direction}, to compute a {\em positive forward relative delta},
  indicating a suboptimal plan (1218 change pairs, 1.8\%).

\item The pair had an upper plan that was newer but no forward extrapolation
  was possible (554 change pairs, 0.8\%).

\item The pair had a lower plan that was newer but no backward extrapolation
  was possible (224 change pairs, 0.3\%).

\item The pair had a relative delta of -1 (2 change pairs, 0.002\%).

\item The pair had a relative delta of 1 (11 change pairs, 0.01\%).

\end{enumerate}
%1 of rel del: 2, -1 of rel del: 21 (neg backward:3, pos backward:18)
From these ten possibilities, we thus retain (3) and (5), which indicate a
faster plan at the later generation (1378 pairs: not suboptimal), and (4) and (6), which indicate a
slower plan at the later generation (1710 pairs: suboptimal).

\subsection{Realizing the Gedanken Experiment}\label{sec:realizing}
We now have the components in place for performing an experiment that
parallels the Gedanken experiment described in Section~\ref{sec:gedanken}.

The analysis in the previous section is for {\em a single pair of adjacent Q@Cs for a single query
running on a specific \hbox{DBMS}},
providing a {\em relative delta} for the minimally \hbox{applicable} generation. A
positive relative delta indicates that the later generation chose a
suboptimal plan; a negative relative delta indicates the later generation
did not. The relative delta is a percentage difference, and so is not \hbox{affected}
by the absolute magnitude of the run time nor by the cardinality in
question. Indeed, because it is a percentage difference, the relative delta
is not affected by the query nor even which \hbox{DBMS} is involved. (DBMSes
vary greatly in the estimated time of individual queries; using the actual
query time would artificially give more weight to change pairs from the
slowest DBMS.) We associate
this relative delta with the later generation, for it is that generation
which had the choice between the two plans for that query.

Consider how the {\em average} relative delta, computed across queries and
DBMSes, of query plans associated with that individual generation, might
behave across successive generations.
The average relative delta for an individual generation is a characterization of the
aggregate impact of the query plans associated with that generation, providing a quantitative estimate of the benefit
of adding that operator. (We use average so that each point is not impacted
by the number of change pairs over which that point is computed. Note that
this approach weights the queries equally. This makes sense for our queries,
summarized in Section~\ref{sec:experiments}; those queries are quite similar.)

What does our causal model in Figure~\ref{fig:model}, supported by the
correlational and regression analyses of the Confirmatory Experiment in
Sections~\ref{sec:corr} and \ref{sec:regression}, say about this? That
causal model asserts that as the number of operators increases in
subsequent generations, the intervening measures in plan space complexity increase, which
therefore impacts suboptimality, also in a positive direction.

Our Gedanken experiment in
Section~\ref{sec:gedanken} takes this behavior and predicts that
as the generations contain successively greater numbers of
operators, optimizer missteps will increase and possibly dominate. What is
the correspondence with the realizable experiment we are now considering?

If we plot the performance of the DBMS on the {\em y}-axis for a
workload consisting of a set of queries over a 
prescribed data set, for a sequence of DBMS generations arranged on the
{\em x}-axis, the average relative delta is in some way a characterization
of the {\em slope} of this relationship. A negative average relative delta
implies that the indicated generation is doing a good job, with less
suboptimality, and so the total workload execution time will go down and
performance will go up; a
positive average relative delta implies that the suboptimal decisions are
dominating, indicated by the total workload execution time going up for that
generation, and thus performance going down.

We expect that the average relative delta for the first few
generations will reflect new operators that improve some
plans, a natural result of the \hbox{efforts} of \hbox{DBMS} developers to
increase performance over successive generations of their \hbox{DBMS}.

That said, all is not rosy in this picture. Each new operator is applicable
to a successively smaller portion of the queries, and perhaps over a
successively smaller portion of the cardinality space.  As already noted,
our causal model predicts that the prevalence of suboptimality will
increase as operators are added.  It seems that even
with DBMS implementers doing smart things, these two considerations predict
that the average relative delta 
will  increase, as suboptimality (a positive relative delta) becomes
more prevalent. This analysis suggests then that the performance curve will
level off and then start dropping off.

So the question comes down to this:  {\em Is there an empirically-determined
point where the decrease due to suboptimality obviates the increase enabled
by the new operator: a DBMS generation where the performance actually drops,
as predicted? How close might modern \hbox{DBMSes} be to that limit?}

\subsection{Trends Across DBMS Generations}\label{sec:trends}
We partition the relevant four sets of change pairs discussed in the
Section~\ref{sec:usingCP} into two groups. The first group consists of those
change pairs with a {\em positive}
relative delta (either forward or backward), indicating that the query
optimizer selected a slower plan at the later generation (corresponding to a
higher generation number), thereby denoting an (empirically) {\em suboptimal
  decision} by the query optimizer at that later generation. The second
group consists of those change pairs 
with a {\em negative} relative delta (either forward or backward) relative
delta, indicating that the query optimizer selected a faster plan at the
later generation, which was thus an (empirically) {\em non-suboptimal
  decision} at that later generation. Note that we can't state
unequivocally that the plan will a negative relative delta is optimal (in the
original, absolute, sense of that term) because there may be a yet another
plan involving the operators 
within that generation that is even faster. That said, in the other case, a positive relative
delta reliably asserts that the plan at this Q@C is \hbox{demonstrably}
{\em empirically suboptimal}, and thus also is suboptimal in the absolute case.

\begin{table}[t]
\tbl{Beneficial Change Pairs\label{tab:non-subopt}}{%
%\begin{center}
\begin{tabular}{c|c|c|c}
{\em Generation}&{\em Number of} &{\em Average}&{\em Cumulative}\\
& {\em Change Pairs} & {\em Relative Delta} & {\em Relative Delta}\\
\hline
1 & --- & --- & 0.00\\
2 & --- & --- & 0.00\\
3 & --- & --- & 0.00\\
4 & 24 & -0.28 & -0.28\\
5 & 282 &-0.118 & -0.130\\
6 & 309 &-0.274 & -0.203\\
7 & 336 & -0.317 & -0.243\\
8 & 43 & -0.20 & -0.241\\
9 & 4 & -0.1 & -0.241\\
10 & 17 &-0.29 & -0.241\\
11 & 357 & -0.178 & -0.225\\
12 & 6 & -0.2 & -0.225\\
13 & 0 & --- & -0.225\\
{\em Cumulative}  &1378 & ---& -0.225\\
\end{tabular}
%\end{center}
}
\end{table}

\begin{table}[t]
\tbl{Deleterious Change Pairs\label{tab:subopt}}{%
%\begin{center}
\begin{tabular}{c|c|c|c}
{\em Generation}&{\em Number of} &{\em Average}&{\em Cumulative}\\
& {\em Change Pairs} & {\em Relative Delta} & {\em Relative Delta}\\
\hline
1 & --- & ---   & ---  \\
2 & --- & ---   & ---  \\
3 & --- & ---   & ---  \\
4 & 2   & 0.2   & 0.2  \\
5 & 211 & 0.076 & 0.077\\
6 & 86  & 0.16  & 0.102\\
7 & 263 & 0.303 & 0.196\\
8 & 273 & 0.303 & 0.231\\
9  & 3  & 0.04  & 0.230\\
10 & 168& 0.319 & 0.245\\
11 & 639& 0.237 & 0.242\\
12 & 53 & 0.35  & 0.245\\
13 & 12 & 0.40  & 0.246\\
{\em Cumulative}&1710&---&0.246\\
\end{tabular}
%\end{center}
}
\end{table}

We first examine those change pairs for which the query optimizer made a
good \hbox{decision}: the non-suboptimal change pairs, summarized in
Table~\ref{tab:non-subopt} and designated as {\em beneficial}.
This data aggregates the results over the
four \hbox{DBMSes}, which had generations ranging from five to thirteen. Overall, only a small number of change pairs, 1378, or about 2\% of the
total, satisfied the requirements listed in Section~\ref{sec:usingCP}, including having a different
generation number for the lower and upper plans. In fact, there were no such
change pairs for the first three generations nor for the last
generation. The second column states the number of change pairs added by
that generation and the third column the average relative delta across just those change
pairs. The last column states the {\em cumulative relative delta} for that
generation, defined as
the sum of the average relative delta for those change pairs associated
with plans having only operators in its generation, divided by the number of such
change pairs. It thus
includes the change pairs associated with all previous generations,
again, as those plans all include operators made available by that generation.

The averages jump around quite a bit, especially for the first few
generations. (One of the reasons is that the number of change pairs
associated with an individual generation varies a
lot.) \hbox{Focusing} on the last column, we see that the cumulative
relative delta starts off at a low of
-0.28 (recall that a negative value is good, as it indicates the
additional operator was effective at lowering the query time) and slowly
increases to -0.225, or more negative numbers trending to less negative
numbers, indicating decreasing benefits of optimization.

This general behavior matches our prediction arising from the structural
causal model: it gets harder for the DBMS query optimizer to squeeze out
performance gains as operators are added to the \hbox{DBMS} over successive
generations.

We now examine those change pairs for which the query optimizer made a poor
decision, in that we can conclude that there was a better plan (the one
right next to it in the change pair), designated as {\em deleterious}. Table~\ref{tab:subopt} provides the
same information across the \hbox{DBMS} generations for the suboptimal
change pairs: those for which the relative deltas are positive, indicating a
decision that increased the query time. While there are still only a small
number of such change pairs, there are more deleterious than beneficial
pairs, with a relatively greater number showing up at more recent
generations. (Half of the beneficial change pairs occurred before generation
8, but half of the deleterious change pairs showed up only in generation 10
or higher.) More strikingly, the cumulative relative delta has the opposite
behavior to the \hbox{non-suboptimal} change pairs (which showed decreasing
benefits): it {\em increases} over the generations (indicating greater
suboptimality), starting at 0.077 and ending at 0.246, a value {\em higher}
than that of the non-suboptimal change pairs. Both trends were as predicted.

\subsection{Have Modern DBMSes Hit The Wall?}\label{sec:hitthewall}
Table~\ref{tab:performance} brings the beneficial and deleterious change
pairs together, stating the total number of change pairs associated with
each generation and the {\em average net relative delta}, again, just for
change pairs associated with that generation. Note that in the third column,
the average net relative delta (per generation), the values transition from
negative (indicating better performance), to quite positive (worse
performance); recall that the relative delta is bounded by -1 and 1.

\begin{table}[t]
\tbl{Assembled Change Pairs and Relative ``Performance''\label{tab:performance}}{%
%\begin{center}
\begin{tabular}{c|c|c|c}
{\em Generation}&{\em Number of} &{\em Average Net}     &{\em Cumulative Net}\\
            & {\em Change Pairs} & {\em Relative Delta} &{\em Relative Delta}\\
\hline
1           & ---                & ---                  & ---\\
2           & ---                & ---                  & ---\\
3           & ---                & ---                  & ---\\
4           & 26                 & -0.24                    & -0.24\\
5           & 493                & -0.035                    & -0.045\\
6           & 395                & -0.179                    & -0.103\\
7           & 599                & -0.045                    & -0.0800\\
8           & 316                & 0.235                    & -0.0256\\
9           & 7                  & -0.04                    & -0.0257\\
10          & 185                & 0.263                    & 0.00750\\
11          & 996                & 0.088                    & 0.0296\\
12          & 59                 & 0.29                    & 0.0346\\
13          & 12                 & 0.40                    & 0.0360\\
{\em Cumulative}& 3088           & ---                  & 0.0360\\
\end{tabular}
%\end{center}
}
%\todo{Young: could you provide the third column? It's the average just for
%  the change pairs for that generation.}
\end{table}

The final column in Table~\ref{tab:performance} provides the
cumulative net relative delta.  This column thus reflects all the change pairs
with plans that would have been generated by that DBMS generation
(hence, including change pairs at that generation along with those
at previous generations). This value provides an indicator of how the four
\hbox{DBMSes} together would have performed (say, in total execution time) on the
workload of the Exploratory and Confirmatory
Experiments, or rather on the 3088 change pairs selected by the criteria in
Section~\ref{sec:usingCP}. (We emphasize that our experiment estimates with
empirical results just
the first derivative of the performance graph, the cumulative net 
relative delta, and then only for a small
number of Q@Cs, and then only as a relative measure, between -1 and 1. It is
thus only very roughly indicative of the {\em shape} of the performance
graph.)

We see that the cumulative net relative delta starts out \hbox{at -0.24}, indicating
that the new operators are {\em increasing} the performance (by
{\em decreasing} the query time) for the workload. This happy situation
unfortunately starts to deteriorate: after generation 6, with each
successive generation, the performance deteriorates. This is as predicted:
the optimizer is struggling with both more options (plans over the available
operators) to select from and a diminished opportunity to make a significant
improvement.

\shorten{OLD: The final column of Table~\ref{tab:subopt} integrates the
  average relative delta to produce a unit-less
{\em relative performance}, an indicator of how the four DBMSes together
would have performed (say, in total execution time) on the workload of the
Exploratory and Confirmatory
Experiments, or rather on the 1378 change pairs selected by the criteria in
Section~\ref{sec:usingCP}. (We emphasize that our experiment estimates just
the first derivative of the performance graph, the average relative delta, and then only for a small
number of Q@Cs, and then only as a relative measure, between -1 and 1. It is
thus only very roughly indicative of the {\em shape} of the performance graph.)
The relative performance gets better (faster, indicated by a falling
value), but that trend again stops at generation 10, after which the
relative performance actually gets worse (slower, indicated by a rising value).}

\begin{figure}
    \subfigure[Cumulative Relative Deltas]{
        %\includegraphics[width=20pc]{figures/gen_benefit.pdf}
        %\includegraphics[scale=0.56]{figures/gen_benefit.pdf}
        %\includegraphics[width=19pc, height=5.5cm]{figures/gen_benefit.pdf}
        \includegraphics[scale=0.735]{figures/gen_benefit.pdf}
        \label{fig:benefit}
    }
    \hspace{-0.17in}
    \subfigure[Performance over Generations]{
        %\includegraphics[scale=0.6]{overall/machine_pt_re.eps}
        %\includegraphics[width=12pc]{figures/perf_gen.pdf}
        %\includegraphics[scale=0.56]{figures/perf_gen.pdf}
        %\includegraphics[width=14.5pc, height=5.5cm]{figures/perf_gen.pdf}
        \includegraphics[scale=0.735]{figures/perf_gen.pdf}
        \label{fig:performance}
    }
%\todo{Young: Please remove the ``Net Cumulative Relative Delta'' curve in
%  the middle and instead put it as 10(b).}
%\todo{Young: also replace ``Suboptimal'' with
%  ``Deleterious (suboptimal)'' and ``Not suboptimal'' with ``Beneficial (not suboptimal)''}
    \caption{Trend of Relative Performance over Generations}
    \label{fig:machine_comp}
\end{figure}

Figure~\ref{fig:machine_comp} tells the story
graphically. Figure~\ref{fig:benefit} plots the last column of
Table~\ref{tab:non-subopt} as the bottom line (beneficial: non-suboptimal cumulative
relative delta) and the last column of Table~\ref{tab:subopt} as the top
line (deleterious: suboptimal cumulative relative delta).
Figure~\ref{fig:performance} plots the ``Relative Performance'', that is,
the negation of the cumulative net relative delta (as a positive relative
delta denotes a slower query, reflecting decreased performance) appearing as the
last column of Table~\ref{tab:subopt}, as this characterizes the relative
performance across generations. This line has a least squares slope of
-0.024.

\subsection{Summary}
The engineering perspective in Section~\ref{sec:gedanken}
predicts that the overall per-generation run time would ``monotonically decrease
as the generations add query evaluation operators.'' The scientific
perspective, applying our validated causal model, reaches the opposite
conclusion: there will be a point where ``the
\hbox{increase} in execution time due to suboptimality obviates the decrease
enabled by the new operator.'' The results of this study clearly support
the latter (cf.~Figure~\ref{fig:performance}).

This highly aggregated result, extracting 3000-odd change pairs having
the specified properties of interest stated in Section~\ref{sec:usingCP} and
drawn from over \hbox{100,000} Q@Cs implies that these four
particular \hbox{DBMSes}, taken together, might be close to or have already
hit the wall, where adding an operator actually slows down the average
query. And recall the analysis in Section~\ref{sec:CPQ} in which a
related problem cropped up around generation 8.  This is the
first study we are aware of that compares the generational trends of
\hbox{DBMSes} over quite disparate code bases, thus getting at fundamental
trends.

It is important to emphasize that we can't say from this one study
whether any of these DBMSes have actually
transitioned to where, for this class of queries, the cost of the
suboptimal change pairs overwhelm the benefits of the non-suboptimal change
pairs. Presumably, the DBMS vendors have done extensive tests to ensure
that the operator added to each generation did in fact effect a speedup on
the representative workloads that they use in evaluating their optimizer
enhancements. However, we {\em can} say that (a)~the trends observed
strongly point to a decreasing benefit and an increasing cost, as predicted
by the simple arguments made above, and (b)~if current DBMSes haven't yet
reached the point of diminishing returns, that possibility looms in the future.

We reiterate the provisos mentioned earlier. In all of these
experiments, we are looking at quite simple queries, over a quite limited
range of data, with only a small percentage (3\%) of change pairs
identified in our experiments. On the other hand, simpler queries
are thought to be easier to generate good plans than more complex queries
(as Hypothesis 2 in Section~\ref{sec:hypotheses} states). Also, relational
data is generally much less uniform in its (i)~data types (we used only
integer data), (ii)~values (the values in our tables are quite evenly
distributed), (ii)~schemas (the schemas of our tables are identical and
quite simple), and (iii)~range of table cardinalities (our smallest
table is $1/{200}^{th}$ of the largest table). The simplicity of queries, schemas,
and tables in this study should independently, and certainly in concert,
minimize the suboptimality observed. Thus, more complex workloads may
present an even higher degree of empirical suboptimality.

\section{Engineering Implications}\label{sec:engineering}
We studied a particular
phenomenon, suboptimality, when the \hbox{optimizer} selecting a slower
plan. This phenomenon is indicated by the existence of a query plan that performs more efficiently
than the DBMS's chosen plan, for the same query. From the engineering
\hbox{perspective}, it is of critical importance to understand the prevalence
of Suboptimality and its causal factors. The genesis of our predictive model
was a sense that Suboptimality is caused in part by the inherent complexity
of these systems and the concomitant interactions between various rules in the
optimizer.

Through a series of experiments managed by our laboratory instrument
management system, \azdb, carried out across several years, we uncovered
several surprising results that provide systematic clues as to where current optimizers come up
short and how they can be further improved.

\begin{itemize}
\item For many queries, a majority of the ones we
considered, the optimizer picked a slower plan for at least one
cardinality, even when the cardinality estimates were completely accurate
and even for our quite simple queries.

\item A quarter of the queries exhibited significant suboptimality ($\geq 20\%$ of
        the run time) at some cardinality.
\end{itemize}
\noindent
These two results indicate that there is still research needed on
this topic. Fortunately, the causal model can suggest
specifically where that research should be focused.

\begin{itemize}

\item Many queries exhibited {\em query fluttering}, in which the query
  optimizer returned to a previous plan at a higher cardinality.
\item Some queries exhibited significant {\em query thrashing}, with a plan change at
         almost every cardinality. While this phenomenon was first visualized
         by Haritsa et al.~\cite{harish07,Haritsa10} on some complex
         queries, we have shown that it is present even in a surprising
         percentage of simple queries.

\item Furthermore, some queries exhibited many changes
{\em to a suboptimal plan} as the \hbox{cardinality} was varied.
\end{itemize}
\noindent
These particular queries, as well as those of the right-hand side of
Figure~\ref{fig:suboptcumulative} exhibiting a large degree of suboptimality, can be a starting point for identifying the root
cause(s) of query thrashing. The phenomenon can be investigated initially on a
per-DBMS basis. Our methodology could then be
used to test proposed causal mechanisms of query thrashing across \hbox{DBMSes}, to
ascertain the generality of any proposed solutions.

\begin{itemize}
\discont{\item The causal model and our experimental results suggest that more
research is needed to improve the cost model of {\em discontinuous
operators}.}{}

\item \discont{We also show that it}{It} may well be useful to explicitly take {\em \hbox{cardinality} estimate
  uncertainty} into account. (Leis et al.~\doubleblind{independently }{}came
  to the same conclusion: ``The results ... demonstrate that the
  state-of-the-art in cardinality estimation is far from
  perfect.''~\cite{Leis15}).

\item This research indicates that aggregates are less of a problem, so
  that aspect of query optimization is in reasonable shape.

\shorten{REVISE to characterize
  more precisely the relative
advantage or disadvantage of heuristic search\shorten{, and \hbox{(c)~research} is needed to
identify the inflection point where adding an operator decreases overall
\hbox{DBMS} performance}raise several unsettling concerns. These particularly simple
queries used only a few plans within any \hbox{DBMS}. Other more complex queries
will certainly use more plans. These results imply that as a \hbox{DBMS} becomes
more sophisticated, by adding operators, it also become more unstable, as
the prevalence of suboptimality grows. Perhaps there is a fixed level of
complexity beyond which a cost-based optimizer will be inexorably unstable.}
\end{itemize}

\discont{We see that costing of \hbox{discontinuous} operators is a
root cause of query suboptimality, and is thus particularly challenging to a
query optimizer. If the cost model is even a little bit off, the optimizer
might be on the wrong side of the ``jump'', thereby selecting a slower
plan.  That the Presence of discontinuous operators has such a high
regression coefficient provides a quite specific guideline: more research is
needed to improve the accuracy of the cost model for such operators, such as
careful calibration that tunes the cost model with more accurate resource
knowledge, including the global memory capacity available, as well as to
improve the algorithms that allocate those buffer pages to specific
operators.}{}

Concerning the plan search
process, which was another identified root cause of suboptimality,\shorten{one preliminary finding was that a heuristic
search process was implicated in 10--20\% of optimizer variance.} in cases
where the \hbox{DBMS} is not as sure about the cardinalities of the underlying
relations or the speed of the disk (e.g.,~if such relations migrated
frequently to a different disk drive~\cite{reiss03}), 
perhaps the optimizer should\shorten{not try as hard to produce out the ``perfect'' plan. 
It might help if optimizers instead}
explicitly take uncertainty into account. Indeed, others have started to
\hbox{argue} that uncertainties in the query planning process should be
acknowledged and exploited~\cite{Babcock05}.

We mentioned dynamic query optimization in Section~\ref{sec:related}
earlier.  Dynamic query-reoptimization normally requires a significant
amount of information to be recorded during query execution, which can incur
non-negligible overhead on the overall query
performance~\cite{Avnur,kabra98}.  We envision that by utilizing the
proposed predictive model for suboptimality, it may be possible to enhance
reoptimization techniques such that given a particular query, a particular
data distribution, and a specific plan operator, just the important
statistics that affect the operator's performance can be identified and
should be recorded, thereby reducing the overhead of bookkeeping irrelevant
\hbox{information}.

Hence, the methodology introduced in this paper suggests fairly specifically where additional
engineering is needed (\discont{the cost model of discontinuous
operators and }{}accommodating cardinality estimate uncertainty) and is not
needed (costing of aggregation).

The generational study in Section~\ref{sec:diminishing} though implies that
the challenge in daunting. That study validates the implications of the
causal model in Figure~\ref{fig:model}, which correctly predicts the almost
inexorable rise in per-generation assembled relative delta, which implies that the
optimizer will eventually hit the wall where it is no longer improving.

We emphasize that this section of this paper, considering engineering
implications of the underlying causal model, contrasts with the rest of the
paper, whose focus is on the science and on understanding these complex
systems at a fundamental level.  Good engineering should be built on solid
scientific results. This paper focuses on the latter.

\section{Summary}\label{sec:summary}

This paper studies an important
component of a \hbox{DBMS}, the query optimizer. This component is an amazingly
sophisticated piece of code, but is still not well understood after decades
of research and development.

While there has been a wealth of research over the last forty years on
engineering approaches and refinements to improve the performance of query
optimization, this is the first paper to the authors' knowledge to take a
scientific approach (in the sense of {\em empirical generalization}),
towards the important phenomenon of {\em suboptimality}.

This paper makes the following contributions in
an \hbox{attempt} to gain new understanding of this
component.
\begin{itemize}
\item Shows that even for simple queries, over a simple schema and
  relatively small range of table sizes, the prevalence of {\em query
  suboptimality}, {\em query flutter}, and {\em query thrashing}, three
  problems that have not been systematically investigated across multiple \hbox{DBMSes}, is
  high, and thus there is still research needed on this mature topic of
  query optimization.

\item Introduces a new {\em methodological
perspective} that treats \hbox{DBMSes} as experimental subjects within empirical generalization.

\item Proposes {\em operationalizations} of several relevant measures that apply
  even to proprietary \hbox{DBMSes}, as well as an overarching {\em predictive
    model} that attempts a causal explanation of suboptimality, encoding
   some of what is known about query optimization.

\doubleblind{\item Utilizes a novel {\em research infrastructure}, \azdb, to express, run, and
  analyze experiments.}{}

\item Tests {\em six hypotheses} deductively derived from the predictive causal model. A
correlational analysis and a regression analysis
provided {\em strong support} for our model, across \hbox{DBMSes}, thus
confirming what was informally articulated but never
coherently tested.

\item Uncovers compelling {\em evidence}
(i)~that (empirical) suboptimality correlates with four operationalizations of query complexity,
(ii)~that suboptimality correlates with \discont{three}{two} operationalizations of plan space
complexity,
(iii)~that the Presence of secondary indexes is a contributor to plan space
\hbox{complexity,}
(iv)~that schema complexity, as operationalized by the
Presence of secondary indexes, moderates these \discont{three}{two}
interactions (though weakly and in the opposite direction), and
(v)~that the Presence of skewed data may diminish
Suboptimality slightly.

\shorten{\item A separate analysis suggested that heuristic plan selection
  appears to also be a source of suboptimality.}

\item Explains a {\em significant portion} (\discont{53.6\%}{52.2\%}) {\em of the variance of empirical suboptimality} for the kinds of queries we looked at, based on the interactions as well as the factors that we identified in the model: optimizer complexity, query complexity, and plan space complexity.  No other factor, as yet unknown, nor combination of factors,
will themselves predict as much variance as the factors we studied in this
paper. And somewhat extraordinarily, it is these common aspects that predict
suboptimality, not the particulars embedded in the inordinate complexity of
each of these DBMSes.
That said, it is certain that there remain several unknown causal factors;
identifying those factors may also have important engineering
implications.

\item Articulates for the first time a {\em limit} on the
  number of operators a \hbox{DBMS} may be able to support, given that the
  empirical evidence suggests that additional operators speed up a smaller
  and smaller portion of the query/cardinality space while incurring an
  increasing chance of suboptimality over the remaining space, which is
  growing.

\item Applies a novel experiment over pairs of adjacent Q@Cs, providing
  empirical evidence that {\em this limit exists} and may have already been
  reached by one or more our subject DBMSes.

\item Identifies {\em specific directions for engineering interventions}.

\item Provides a {\em path toward scientific progress} in the understanding
  of a key enabling technology.  It is important to emphasize that our model
  doesn't apply to only one implementation of the algorithm or to one
  \hbox{DBMS}.  Rather, it is quite broad, applying to any \hbox{DBMS} with
  a cost-based optimizer.

\end{itemize}
This paper thus suggests a framework of casual model elaboration and
directed engineering efforts.

\section{Future Work}\label{sec:future}
There are at least three fundamental directions that could be taken in
future work: (i)~model testing and refinement, (ii)~following up on engineering
implications drawn from the predictive causal model, and (iii)~applying the empirical
generalization perspective to other aspects of DBMSes.

The model proposed and tested here, shown in Figure~\ref{fig:model}, is
relatively simple, with six constructs and their relationships. Some of the
predicted relationships were not borne out in the confirmatory analysis.

Might
further testing support or reject the hypothesized correlation between Number
of repeats (in the Plan complexity construct) and Presence of primary key
attribute and Presence of subquery (both in the Query complexity construct)? Why did the hypothesized moderation
  by the Presence of secondary indexes (in the Schema complexity construct)
  {\em decrease} the influence of Query complexity on Plan space complexity?

\discont{Why was the direction between Number of operators available and Number of
operators with discontinuity opposite from hypothesized? As
mentioned in Section~\ref{sec:corr}, the operationalization of ``discontinuous
operator'' may incorrect classify some operators. We suggest that a more holistic operationalization be
devised that relies not on individual pairs of adjacent Q@Cs but rather on the estimated time for the operator across all Q@Cs in which it
appears in the identical plan, for each query. The analysis could then be rerun, using that refined
definition, on the confirmatory data (note of course that that would
technically be considered an exploratory analysis, as the operationalization
of one of the operators would have changed), followed by a truly
confirmatory analysis on newly gathered data if appropriate.}{}

Finally, why did the Presence of skewed data (actually, the increase in skew
from tiny to small, in the Data complexity construct)
  {\em decrease} Suboptimality? (As others have noted, ``cardinality
  estimates are usually computed based on simplifying assumptions like
  uniformity and independence. In real-world data sets, these assumptions
  are {\em frequently} wrong, which may lead to sub-optimal and sometimes
  disastrous plans ... the cardinality estimators of the major relational
  database systems produce bad estimates for many realistic queries, in
  particular for multi-join queries.''~\cite{Leis15})

Our causal
model is extensible, in that we can add other factors, as long as their proper
operationalization can be established, and additional causal links. It would be
useful to consider
\begin{itemize}
\item {\em schema complexity:} foreign keys and
uniqueness and other kinds of constraints, 
\item {\em query complexity:} complex predicates, multiple nesting of
  subqueries and different types of subqueries, query operators such as {\tt
    EXCEPT} and {\tt UNION}, other
SQL clauses such as {\tt ORDER BY} and {\tt GROUP BY}, and user-defined data
types, methods, and operators,
\item {\em data complexity:} data that doesn't exhibit uniformity or independence
  data, and
\item {\em plan space complexity:} the actual space of plans considered by
  the optimizer (discussed below).
\end{itemize}

Our Confirmatory experiment used only two values of skew, tiny
($\nicefrac{1}{2M}$) and small ($\nicefrac{1}{10K}$), for the data complexity
construct. It would be useful to extend the study to (much) larger
values. (This could increase the number of duplicates in joins, which may
dramatically increase the query time.) But more to the point, our definition
of skew utilizes a uniform distribution of values, albeit across a narrowing
range as skew increases. It would be useful to look at other distributions,
such as an increasing distribution and a normal distribution, as well as a
very spiky distribution, in part to see whether the positive impact of skew
on suboptimality continues (our intuition remains that with substantial
skew, suboptimality will increase). It would also be useful to run experiments over
entirely different benchmarks, including those based on real-world data and
queries such as the Join Order Benchmark~\cite{Leis15}, appropriately
modified to provide a range of values for the various constructs such as
data skew.

Kabra and
\hbox{DeWitt} have identified another source of complexity: inaccurate
statistics on the underlying tables and insufficient information about the
runtime system: ``amount of available resources (especially memory), the
load on the system, and the values of host language
variables.''~\cite[p.~106]{kabra98}.  Might there be other unanticipated relations, that are unknown simply because they haven't been
looked for?

It may be useful to look into query flutter and thrashing in greater detail, as those phenomena
provide concrete indicators of problematic optimizer behavior. One
possible methodological approach is to utilize SQL optimizer hints, such as
``{\tt \*+ SEMIJOIN}'' in MySQL and ``{\tt enable\_hashjoin(false)}'' in PostgreSQL,
to encourage the optimizer to produce more plans at a given cardinality,
that can then be timed to make more explicit the entire plan space (recall that
CEPS, the cardinality of the effective plan space, is no greater and
probably much smaller than the cardinality of the plan space).

Along with adding such factors to the model, one can watch whether these
other factors impact (at all, and if so, how much) these initial constructs
and relationships. So for example it would be interesting to study how a suboptimal subquery can affect
the suboptimality of the containing query.  For instance, is it true that if
many subqueries are themselves suboptimal, does that causally impact whether
the overall query is suboptimal?

It would also be useful to study other relational DBMSes, as the model
should apply to any such DBMS using a cost-based optimizer.

All of the above activities are in the tradition of science, specifically
empirical generalization (cf.~Figure~\ref{fig:empirical}).

The second fundamental direction takes the validated model and draws
engineering implications. Our model has provided specific directions for implementation interventions:
\discont{refining the cost model of discontinuous operators, }{}improving buffer
allocation algorithms\discont{,}{} and accommodating cardinality estimate uncertainty.
One can further ask, for each causal relationship in the model,
relationships that have been validated in empirical studies, what is it about
the cost-based query optimizer that that interaction is evidenced? And then
one can ask, what changes to that optimizer might ameliorate that
relationship?

It may be that that relationship is {\em baked into the optimizer}. For
example, the study in Section~\ref{sec:diminishing} was predicated on some
basic aspects of adding operators: each successive operator {\em benefits} a smaller region
of query/cardinality (specifically, Q@Cs) and is susceptible to
suboptimality on a larger region of Q@Cs
(cf.~Section~\ref{sec:gedanken}). That seems like a fundamental limitation.

But other relationships may be avoidable, through selection of alternative
algorithms. After all, this study generalizes over only four DBMSes.

Engineering also provides alternative methodologies to studying the model
variables, which we have examined only from the ``outside'' of the DBMS. It
would be useful to manipulate \hbox{DBMSes} internally (at least for those
that are open-source or are available within the DBMS vendor), turning on
and off the rules and observing suboptimality. (An example is a
study of cardinality estimation, cost model, and plan enumeration~\cite{Leis15}.) And it may be possible to
determine whether \hbox{certain} query rewrite techniques are employed within each
\hbox{DBMS}, perhaps introducing other ``\hbox{DBMS} complexity'' factor(s) in our
model.

If extant cost-based query optimizers are hitting a wall,
it might be necessary to fundamentally revisit query optimization,
to come up with an entirely new approach that is less impacted by number
of operators and by CEPS and thus avoids the flutter phenomenon.

The identified limit is inherent in cost-based query
optimization. Engineering approaches can help ameliorate the effect
identified here. One approach might be to time query plans as they run
(effectively introducing an empirical cost model) in order to provide more
accurate cost estimates, though that is itself time-consuming
when there is a large plan space. Another that has been proposed is to
replace multiple physical operators of a logical  operator (in this case,
the join operator, which can have as physical operators nested-loop,
sort-merge, hash, and index join) with a single physical operator (in this
case, g-join) that performs equally or better that the alternative physical
operators, hopefully in all cases~\cite{Graefe12}. Doing so can reduce
optimizer errors when choosing between the previously-available multiple
variants. This approach is aligned with our theoretical and empirical
results: if adding physical operators gets us close to or beyond the limit
of performance improvement, as we have
seen, then removing variants should move the DBMS away from that
limit. However, the g-join has not yet been demonstrated to be applicable in
all situations and for all queries, and in any case, the limit still
remains.

Engineering solutions to {\em eliminate} the identified limit, to allow
continued introduction of new operators, thus requires perfecting the cost
model and making query plan enumeration deterministic, neither of which
seems to be practical, or adopting an entirely new tact that eschews
cost-based query optimization entirely, such as a
learning-based approach or defining a single physical operator for
each logical operator.

A third fundamental direction is to return to science, but examine dependent
measures other than suboptimality. This paper demonstrates that by employing an extensible causal model, many complex factors can be studied
via a systematic, statistically sound, scientific manner to better
understand the causal factors and their relationships.
\doubleblind{Another recent example is Suh's validated causal
predictive model for DBMS thrashing~\cite{Suh17}. }{}What other areas of the
rich field of databases might be amenable to this approach?

The methodology utilized in this paper to introduce a predictive causal model has
suggested the scope of the problem of query suboptimality, a number of
contributing factors, and a collection of specific engineering efforts that
can now be considered. An ultimate goal is a refined causal model that
fully explains how query suboptimality arises in cost-based optimizers,
thereby enabling engineering solutions that address this important issue.

\doubleblind{\section{Acknowledgments}
This research was supported in part by NSF grants \hbox{IIS-0639106},
IIS-0415101, and EIA-0080123 and by NRF grant NRF-2011-0020576. We thank Rui Zhang for his help in initiating
this research. We appreciate helpful discussions and insightful feedback from Melanie
Brucks, Curtis Dyreson, Christian Jensen, David Maier, Thomas Matheson, Arash Termehchy,
Abhijit Saha, and Marianne Winslett. Ricardo Carlos, Preetha Chatterjee, Pallavi
Chilappagari, Jennifer Dempsey, David Gallup, Kevan Holdaway, Matthew Wong
Johnson, Andrey
Kvochko, Siou Lin, Adam Robertson,
Lopamudra Sarangi, Linh Tran, Cheng Yi, and Man Zhang contributed
to the \hbox{\azdb} and Phil Kaslo, Tom Lowry, and John
Luiten helped in constructing and maintaining our experimental instrument\c2j{}{,
  a laboratory of six machines and associated software}.
}{}

%\section{Electronic Appendix}
%The online electronic appendix to this article can be accessed in the ACM
%Digital\linebreak Library. That appendix lists...
\doubleblind{\todo{Rick:should the appendix be in the paper, rather than
    being electronic? Should it be part of the paper proper?}Appendix~\ref{sec:app} provides further details on the experiments.}{}

%\vfill\pagebreak
%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.

\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{paper}
\shorten{NOT USED:\newcommand{\etalchar}[1]{$^{#1}$}shorten{\begin{thebibliography}{99}

\vspace{0.1em}
\end{thebibliography}
}
}
\newpage
\appendix
\section{Details on the Experiments}\label{sec:app}
Table~\ref{tab:run_stat} in Section~\ref{sec:experiments} lists the run statistics of the seven experiments
  used in this \hbox{paper}. In this appendix we provide more detailed information
  on the experiments, walking through the columns in succession of
Table~\ref{tab:run_stat2}, given below.

\subsection{Data Sets}\label{sec:appdatasets}
\begin{table}[t]
\tbl{Experiments 1--7: Detailed Run Statistics\label{tab:run_stat2}}
{%
\resizebox{140mm}{!}
{
\begin{tabular}{c|c|c|c|c|c|c}
& {\em Experiment}& {\em Data Sets} & {\em Lab Shelves} & {\em What was} &{\em Was Query}&{\em Number of}\\
&                 & {\em Used}      &                   & {\em Examined?}&{\em Timed?}   &{\em Retained (Raw)}\\
&&&&&& {\em Q@Cs}\\
\hline
1 & Monotonicity 	& A & 6.0     &all cardinalities&yes& 12,000 (12,000)\\
2 & Exhaustive & A & 5.19 + 6.0&all cardinalities&yes& 27,948 (32,000)\\
3\shorten{4} & Exhaustive with Keys & B& 6.0     &change pairs   &no& 40,000 (40,000) \\
4\shorten{5} & Initial Exploratory & A + B&5.19 + 5.2 + 6.0&change pairs&yes& 8,171 (8,842)\\
5\shorten{3} & Refined Exhaustive & A & 7.1     &all cardinalities&yes & 29,515 (32,000)\\
6 & Exploratory 	& A + B&7.1   &change pairs   &yes& 12,100 (12,560)\\
7 & Confirmatory 	&A + B + C +D&7.1&change pairs&yes& 94,502 (99,558)\\
\multicolumn{2}{c|}{\em Total}&      &   &             &   & 184,236 (196,960)\\
\end{tabular}
}
}
\end{table}

The third column identifies the data set(s) used in each experiment, that is, the
specific tables being queried. There are four data sets, named A, B, C, and
D.

We first discuss the features shared between the four data sets. As
introduced in Section~\ref{sec:motivation}, the queries referenced tables
{\tt ft\_HT1}, {\tt ft\_HT2}, {\tt ft\_HT3}, and {\tt ft\_HT4}. All four
tables contain four columns, each of type integer. The specific values of
the rows for all but the first column depend
on the Presence of skewed data. Section~\ref{sec:datacomplexity} provides
the algorithm for generating the values for different values of skew; this
algorithm is used in the second to fourth columns, which for any row will have
identical values. The first column holds a unique integer starting from 1
and going to 60K or 2M, for use in an optional primary key.

There was one version of the
last three tables, for use with MySQL, with cardinality 60K, and one version
for the rest of the DBMSes, with cardinality 2M.

We generate 200 versions of {\tt ft\_HT1}, termed the {\em variable table}. For MySQL, these version contain
300, 600, 900, 1200, $\ldots$, 59,700, and 60,000 rows; for the rest of the
DBMSes, these versions contain 10,000, 20,000, 30,000, $\ldots$, 1,970,000,
and 2M rows, as introduced in Section~\ref{sec:motivation}.

We now identify the four successive data sets, elaborating on the discussion in Sections~\ref{sec:datasets}--\ref{sec:experiments}. Data Set C is the simplest to
describe: it specifies no primary key, has no duplicate rows, and has no skew
(of course, for any of the four tables). Data Set A differs from Data Set C
only in that there is skew.  As summarized at the end of
Section~\ref{sec:datacomplexity}, we use two values of skew, tiny and small.\shorten{five skewness values: 0
(approximately), 0.001, 0.1, 0.5, and 1.0.}

Data Set B is similar to Data Set A, adding the specification of the first
column as the primary key. And Data Set D is similar to Data Set B, adding
the specification that the other three columns should each be associated
with a secondary index, only for each (one) column.
We see the confirmatory experiment examined a much larger variation of data sets than
the exploratory studies.

\subsection{Other Details}\label{sec:otherdetails}
The next column of Table~\ref{tab:run_stat2} concerns the {\em Lab
  Shelf}. \hbox{\azdb} utilizes the metaphor of a bookshelf of lab
notebooks. Here, each shelf is associated with a version of \hbox{\azdb} itself. For the experiments in this paper, we used at various times over
the last three years lab shelves (that is, program versions) 5.19, 5.2,
6.0, and 7.1. Versions 5.19 and 5.2 were very similar; both implemented
TTPv1.  Version 6.0 also implemented TTPv1, but collected more query
measures that were not relevant for this paper. Version 7.1 implemented
TTPv2.

The \hbox{\azdb\ system also includes support} for {\em experiment
  scenarios}, each of which
is a small amount~ (a few hundred lines) of Java code that actually performs the experiment, such as varying the
cardinality and running different queries on the data. The only difference
in the scenario code across these experiments was in accommodating the
details of the data set (that is, creating secondary indexes and data skew).
\shorten{\todo{Young:was this in the scenario code? It seems that this should be
  outside of that: Rick: Yes, this is part of the scenario
  code. Specifically, if the Presence of secondary indexes is true, or the
  Presence of skewed data is false, then in the scenario code an experiment subject creates the indexes or makes no skew when populating tables. After table population is done, the code studies each query, or detects a change pair and makes query executions at that change pair.}}

The bottom line is that while the lab shelf and experiment scenario varied
somewhat, the only important aspect was the {\em Protocol} column of
Table~\ref{tab:run_stat}.

We now turn to the fifth column, ``what was examined?'' Here there are just two
possibilities, all 200 cardinalities or just the cardinalities at which the
query plan changed. The sixth column, ``what was timed?'', indicates that Exhaustive with Keys,
described in Section~\ref{sec:experiments}, just
collected query plans, not timing any of them.

The last column states how many Q@Cs the experiments measured (each with 10
QEs), termed {\em raw}, and how many Q@Cs were retained after the protocol
(listed in the third column of Table~\ref{tab:run_stat}) dropped query
executions and Q@Cs via its many sanity checks. (Note that we don't list QEs
and retained QEs in Table~\ref{tab:run_stat} for Exhaustive with Keys simply
because we did not use timing data in that experiment.)

\subsection{Query Sets}\label{sec:querysets}
The following sets of queries were used in the seven experiments.

\begin{description}
\item[QSa] 100 queries over the four tables, generated as described in Section~\ref{sec:querycomplexity} (on Data Set A)
\item[QSb] 100 queries (on Data Set A)
\item[QSc] 100 queries (on Data Set A)
\item[QSd] 100 queries (on Data Set A)
\item[QSe] 100 queries (on Data Set A)
\item[QSf] 100 queries (on Data Set A)
\item[QSg] 100 queries (on Data Set A)
\item[QSh] 100 queries (on Data Set A)
\item[QSi] 100 queries (on Data Set A)
\item[QSj] 100 queries (on Data Set A)
\item[QSk] A query set consisting of the 390 queries drawn from {\em QSa}--{\em QSj} (on Data Set B)
\item[QSl] A query set consisting of 110 new queries (on Data Set B)
\item[QSm] A query set consisting of 100 queries without aggregates (on Data Set C)
\item[QSn] A subquery query set consisting of 100 queries, each with a subquery (on Data Sets A, B, and D)
\item[QSo] A subquery query set consisting of 100 queries, each with a subquery (on Data Set D)
\item[QSp] A query set consisting of 100 queries drawn from {\em QSk} (on Data Set D)
\end{description}

Experiment 1 (Monotonicity) used the first 50 queries from QSa
for one DBMS plus the first six queries from QSa and two queries each from
{\em QSd} and {\em QSe} for MySQL, for a total of 60 query instances.

Experiment 2 (Exhaustive) used the first 50 queries from {\em QSa}
for the three other DBMSes plus the ten queries for MySQL from Experiment 1, for
a total of 160 query instances. Experiment 5 (Refined Exhaustive) used the same 160 queries.

Experiment 3 (Exhaustive with Keys) used the first 50 queries from {\em QSa}
for the four DBMSes, for a total of 200 query instances.

Experiment 5 (Initial Exploratory) used the (100) queries from {\em
  QSa}--{\em QSf} (for
one DBMS), the first 20 queries from {\em QSa}--{\em QSf} (for another DBMS), the first
10 (primary key) queries from {\em QSa} (for the four DBMSes), and the first 10
(non-primary key) queries from {\em QSa} (for the other two DBMSes), for a total of 780 query instances.

Experiment 6 (Exploratory) used {\em QSa} and {\em QSb}, plus the first 100 (primary
key) queries from {\em QSk}, for the four DBMSes, for a total of 1200 query
instances.

Experiment 7 (Confirmatory) used {\em QSc}--{\em QSj},
along with {\em QSk} except the (first 100) queries included in Experiment 6,
{\em QSl} for primary key (for two runs, or 220 queries),
{\em QSm} for no data skew,
{\em QSn} for primary key and subquery,
{\em QSn} for subquery,
{\em QSn} and {\em QSo} for primary key and secondary index and subquery,
{\em QSp} for primary key and secondary index,
all across the four DBMSes, for a total of 7,640 query instances.

\shorten{
\newpage
\todo{This is old material}

\input{OldTests.tex}

\input{Repeatability.tex}

\input{Measuring.tex}

\input{ExperimentSetup.tex}

\input{Discontinuous.tex}

\input{Results.tex}
}
\end{document}

% LocalWords:  PVLDB

### 2017-08-08 subopt-tods17/data/conf_data_010817.R

@@ -1,0 +1,100 @@
+setwd("I:\\My Documents\\research\\a causal model\\conf_data")
+
+newVarModel <- read.table("conf010817.txt",
+                   header=TRUE,
+                   sep="\t"         # use "\t" for tab-delimited files
+)
+
+
+
+# combined <- merge(mainModel, newVars, by=c("DBMS", "RUNID", "QUERYNUM"))
+
+# save(combined,file="dataC.Rda")
+
+newVarModel$int_si_op<- newVarModel$sec_idx*newVarModel$opCt
+newVarModel$int_si_nf<- newVarModel$sec_idx*newVarModel$NUMFROM
+newVarModel$int_si_na<- newVarModel$sec_idx*newVarModel$NumAgg
+newVarModel$int_si_pp<- newVarModel$sec_idx*newVarModel$pk_pr
+newVarModel$int_si_sp<- newVarModel$sec_idx*newVarModel$sq_pr
+
+# Transform Subopt 
+newVarModel$log_SUBOPT = log10(1 + newVarModel$SUBOPT)
+
+
+
+# excluding "PK_NoJoin" since not populated for smaller set
+woDBMS <- c("opCt", "NUMFROM", "NumAgg", "pk_pr", "sq_pr", "CEPS", "num_repeats", "NumDiscontOperators", "log_SUBOPT", "sec_idx_pr", "skew")
+# woDBMS <- c("queryId", "querynumber", "opCt", "NUMFROM", "CEPS", "NUMSELFJOIN", "CtDbOp", "SUBOPT", "runid", "NumDiscontOperators", "NumAgg", "PK_NoGroupBy", "PK_TwoSided", "PK_InGroupBy", "PK_InOnesideJoin", "PK_NotInGroupBy", "PK_NotInJoin", "pk_pr", "num_repeats", "sec_idx_pr", "sq_pr")
+newVarModel[woDBMS] 
+
+
+# get means for variables in data frame newVars
+# excluding missing values 
+# sapply(newVars[woDBMS], mean, na.rm=TRUE)
+
+# Set no of digits for output
+options(digits=2)
+
+# Correlations/covariances among numeric variables in 
+# data frame newVars. Use listwise deletion of missing data. 
+cor(newVarModel[woDBMS] , use="pairwise.complete.obs", method="pearson") 
+# cov(newVars[woDBMS], use="all.obs")
+
+# Correlations with significance levels
+# install Hmisc
+install.packages("Hmisc")
+
+library(Hmisc)
+
[…]

### 2017-08-08 subopt-tods17/paper.tex
TODS submission on the subopt

@@
 suboptimality prevalence correlates positively with (a)~\hbox{DBMS}
-complexity, (b)~schema complex, (c)~data complexity, and (d)~query
+complexity, (b)~schema complexity, (c)~data complexity, and (d)~query
 complexity, providing empirical support for this model. These

### 2017-08-11 pt_characterization/rep_check/u_s_time.tex
New draft

@@
 	\subfigure[Standard deviation (absolute)]{
-		\includegraphics[scale=0.6]{u_s_time/overall_std.eps}
+		%\includegraphics[scale=0.6]{u_s_time/overall_std.eps}
+		\includegraphics[scale=0.6]{u_s_time/new_overall_pt_std.eps}
 		\label{fig:overall_std}
@@
 	\subfigure[Standard deviation (absolute) in log scale]{
-		\includegraphics[scale=0.6]{u_s_time/overall_std_log.eps}
+		%\includegraphics[scale=0.6]{u_s_time/overall_std_log.eps}
+		\includegraphics[scale=0.6]{u_s_time/new_overall_pt_std_log.eps}
 		\label{fig:overall_std_log}
@@
 		\label{fig:overall_re}
+	}
+	\subfigure[Coefficient of variation in log scale (relative): divisor=task length]{
+		\includegraphics[scale=0.6]{u_s_time/new_overall_pt_re.eps}
+		\label{fig:overall_re2}
 	}

### 2017-08-11 pt_characterization/rep_check/repet_data1/et_hist_sodb9.r

@@ -12,6 +12,8 @@ xmin <-min(x$METIME)
 xmin
 xmax <-max(x$METIME)
 xmax
+task_len <- 1000
+sd(x$METIME)/task_len
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
@@ -38,6 +40,8 @@ ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
 xmin <-min(x$METIME)
 xmax <-max(x$METIME)
+task_len <- 2000
+sd(x$METIME)/task_len
 plot(h, xaxt="n",freq=TRUE,ylim=c(0,ymax), xlim=c(xmin-1,xmax+1),col="blue", main='ET frequency on INC2', 
 sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
 xlab='ET (ms)', ylab=expression('Frequency'))
@@ -56,6 +60,8 @@ xmin <-min(x$METIME)
 xmin
 xmax <-max(x$METIME)
 xmax
+task_len <- 4000
+sd(x$METIME)/task_len
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
@@ -79,6 +85,8 @@ xmin <-min(x$METIME)
 xmin
 xmax <-max(x$METIME)
 xmax
+task_len <- 8000
+sd(x$METIME)/task_len
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
@@ -103,6 +111,8 @@ xmin <-min(x$METIME)
 xmin
 xmax <-max(x$METIME)
 xmax
+task_len <- 16000
+sd(x$METIME)/task_len
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
@@ -127,6 +137,8 @@ xmin <-min(x$METIME)
 xmin
 xmax <-max(x$METIME)
 xmax
+task_len <- 32000
+sd(x$METIME)/task_len
[…]

### 2017-08-11 pt_characterization/rep_check/repet_data1/pt_hist_sodb9.r

@@ -12,6 +12,8 @@ xmin <-min(x$PRTIME)
 xmin
 xmax <-max(x$PRTIME)
 xmax
+task_len <- 1000
+sd(x$PRTIME)/task_len
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
@@ -34,6 +36,8 @@ xmin <-min(x$PRTIME)
 xmin
 xmax <-max(x$PRTIME)
 xmax
+task_len <- 2000
+sd(x$PRTIME)/task_len
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
@@ -63,6 +67,8 @@ ymax <- ceiling(ymax/100)*100
 #ymax <- 200
 xmin <-min(x$PRTIME)
 xmax <-max(x$PRTIME)
+task_len <- 4000
+sd(x$PRTIME)/task_len
 binsize=1
 plot(h, ylim=c(0,ymax), xaxt="n",freq=TRUE,col="green", main='PT frequency on INC4', 
 sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
@@ -83,6 +89,8 @@ xmin <-min(x$PRTIME)
 xmin
 xmax <-max(x$PRTIME)
 xmax
+task_len <- 8000
+sd(x$PRTIME)/task_len
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
@@ -108,6 +116,8 @@ xmin <-min(x$PRTIME)
 xmin
 xmax <-max(x$PRTIME)
 xmax
+task_len <- 16000
+sd(x$PRTIME)/task_len
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
@@ -132,6 +142,8 @@ xmin <-min(x$PRTIME)
 xmin
 xmax <-max(x$PRTIME)
 xmax
+task_len <- 32000
+sd(x$PRTIME)/task_len
[…]

### 2017-08-11 pt_characterization/rep_check/repet_data2/et_hist_sodb9.r

@@ -325,6 +325,8 @@ xmin <-min(x$METIME)
 xmin <- xmin-2
 xmax <-max(x$METIME)
 xmax <- xmax+2
+task_len <-8192
+sd(x$METIME)/task_len
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
@@ -349,6 +351,8 @@ xmin <-min(x$METIME)
 xmin <- xmin-6
 xmax <-max(x$METIME)
 xmax <- xmax+2
+task_len <-16384
+sd(x$METIME)/task_len 
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$METIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)

### 2017-08-11 pt_characterization/rep_check/repet_data2/pt_hist_sodb9.r

@@ -309,7 +309,7 @@ dev.off()
 
 x = read.csv(file="8192_sec.dat",head=TRUE,sep="\t")
 setEPS()
-postscript("8192_sec_pt_hist.eps")
+postscript("8192_sec_pt_hist2.eps")
 x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
 x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
 #x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
@@ -318,6 +318,8 @@ xmin <-min(x$PRTIME)
 xmin <- xmin-2
 xmax <-max(x$PRTIME)
 xmax <- xmax+1
+task_len <-8192
+sd(x$PRTIME)/task_len 
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 ymax <- max(h$counts)
@@ -331,9 +333,12 @@ dev.off()
 
 x = read.csv(file="16384_sec.dat",head=TRUE,sep="\t")
 setEPS()
+task_len <-16384
 postscript("16384_sec_pt_hist.eps")
 x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
 x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+task_len <-16384
+sd(x$PRTIME)/task_len 
 #x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
 binsize=10
 xmin <-min(x$PRTIME)

### 2017-08-11 pt_characterization/rep_check/u_s_time/pt_hist.r

@@ -290,6 +290,7 @@ axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(m
 dev.off()
 
 #################
+numSamples <- 1000
 x = read.csv(file="3_sec.dat",head=TRUE,sep="\t")
 x <- subset(x, x$ITERNUM <= numSamples)
 setEPS()
@@ -298,6 +299,7 @@ binsize=1
 #x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
 xmin <-min(x$PRTIME)
 xmax <-max(x$PRTIME)
+sd(x$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
 ymax <- max(h$counts)
@@ -317,6 +319,7 @@ binsize=1
 #x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
 xmin <-min(x$PRTIME)
 xmax <-max(x$PRTIME)
+sd(x$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
 ymax <- max(h$counts)
@@ -335,24 +338,7 @@ binsize=1
 #x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
 xmin <-min(x$PRTIME)
 xmax <-max(x$PRTIME)
-nbins <- ceiling((xmax-xmin) / binsize)
-h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+1,1),plot=F)
-ymax <- max(h$counts)
-ymax <- ceiling(ymax/100)*100
-plot(h, xaxt='n', axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green",  main='PT frequency on INC12', 
-sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
-xlab='PT (ms)', ylab=expression('Frequency'))
-axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
-dev.off()
-
-x = read.csv(file="12_sec.dat",head=TRUE,sep="\t")
[…]

### 2017-08-19 pt_characterization/rep_check/u_s_time.tex
New 3D plot

@@
 This section exhibits user and system time histograms on the first run of 
-INC with its task length increasing from 1 second to 1024 seconds 
-and the second run of INC from 2048 seconds to 16384 seconds via EMPv5 with no Step 2.
+INC with its task length increasing from 1 second to 4096 seconds 
+and the second run of INC from 8192 seconds to 16384 seconds via EMPv5 with no Step 2.
 The detailed description of the base data is from Table~\ref{tab:exp_notes1}, 
@@
 \subsection{Summary of Histograms}
+Figure~\ref{fig:hist3d} represents a 3D plot of collecting 
+the selected histograms from the first and second runs: 
+Figures ~\ref{fig:s9_r1_pt_hist1},~\ref{fig:s9_r1_pt_hist2},~\ref{fig:s9_r1_pt_hist3}, and ~\ref{fig:s9_r1_pt_hist4} (associated with INC1$\sim$INC4096 of the first run) and~\ref{fig:inc8192_r2_hist_v5} and~\ref{fig:inc16384_r2_hist_v5} (associated with  INC8192$\sim$INC16384 of the second run).
+The x-, y-, z-axes
+indicate bins (PT range), task length, 
+and relative frequency of each histogram.
 
@@
 	\centering
-	\includegraphics[scale=0.6,angle=270]{u_s_time/3d_plot.eps}\label{fig:3d_plot}
+	%\includegraphics[scale=0.6,angle=270]{u_s_time/3d_plot.eps}\label{fig:3d_plot}
+	\includegraphics[scale=0.6]{u_s_time/new_3d_plot}\label{fig:3d_plot}
 	\caption{3D histograms on INC~\label{fig:hist3d}}
@@
 
-\subsection{Absolute and Relative Variance over Increasing Task Length}
+\subsection{Absolute and Relative Variance over Increasing Task Lengths}
+Figure~\ref{fig:cv_inc} exhibits absolute and relative variances over increasing task lengths. 
+More specifically, Figures~\ref{fig:overall_std} and~\ref{fig:overall_std_log} 
+concern the PT standard deviations on INC1$\sim$INC4096 of the first run and 
+INC8192$\sim$INC16384 on the second run, as well as those from 
+the intermediate task lengths described in Table~\ref{tab:exp_notes4}. 
+The y-axis in Figure~\ref{fig:overall_std_log} is taken in log scale.
+Figures~\ref{fig:overall_re} and~\ref{fig:overall_re2} shows 
+the relative variance of the same data set---{\em coefficient of variation} (= standard deviation 
+/ task length or mean). Both the x and y axes in these two figures are taken in log scale.
 
@@
 	}
-	\subfigure[Coefficient of variation in log scale (relative)]{
+	\subfigure[Coefficient of variation in log scale (relative): divisor=mean]{
 		\includegraphics[scale=0.6]{u_s_time/overall_re.eps}

### 2017-08-19 pt_characterization/rep_check/10k_run/hist.r

@@ -73,11 +73,11 @@ sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""),
 xlab='PT (ms)', ylab=expression('Frequency'))
 axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
 dev.off()
-
+numSamples <- 10000
 x = read.csv(file="16_sec.dat",head=TRUE,sep="\t")
 x <- subset(x, x$ITERNUM <= numSamples)
 setEPS()
-postscript("16_sec_pt_hist_v5.eps")
+postscript("16_sec_pt_hist.eps")
 binsize=1
 xmin <-min(x$PRTIME)
 xmax <-max(x$PRTIME)

### 2017-08-19 pt_characterization/rep_check/10k_run/new_3d_plot.r

@@ -9,7 +9,8 @@ xmin <-min(x1$PRTIME)
 xmax <-max(x1$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h1 = hist(x1$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
-x1_pt <- (h1$mids-1000)/1
+#x1_pt <- (h1$mids-1000)/1
+x1_pt <- (h1$mids-min(h1$mids))
 y1_raw_pt <- 1
 y1_pt <- 0
 z1_pt <- h1$counts/numSamples
@@ -23,7 +24,8 @@ xmin <-min(x2$PRTIME)
 xmax <-max(x2$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h2 = hist(x2$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
-x2_pt <- (h2$mids-2000)/2
+#x2_pt <- (h2$mids-2000)/2
+x2_pt <- (h2$mids-min(h2$mids))
 y2_raw_pt <- 2
 y2_pt <- 1
 z2_pt <- h2$counts/numSamples
@@ -37,7 +39,8 @@ xmin <-min(x4$PRTIME)
 xmax <-max(x4$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h4 = hist(x4$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
-x4_pt <- (h4$mids-4000)/4
+#x4_pt <- (h4$mids-4000)/4
+x4_pt <- (h4$mids-min(h4$mids))
 y4_raw_pt <- 4
 y4_pt <- 2
 z4_pt <- h4$counts/numSamples
@@ -52,7 +55,8 @@ xmin <-min(x8$PRTIME)
 xmax <-max(x8$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h8 = hist(x8$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
-x8_pt <- (h8$mids-8000)/8
+#x8_pt <- (h8$mids-8000)/8
+x8_pt <- (h8$mids-min(h8$mids))
 y8_raw_pt <- 8
 y8_pt <- 3
 z8_pt <- h8$counts/numSamples
[…]

### 2017-08-19 pt_characterization/rep_check/u_s_time/new_3d_plot.r

@@ -1,6 +1,4 @@
 #### An R script for 3D plot
-require(rgl)
-library(plot3D)
 numSamples <- 1000
 ### 1 sec
 x1 = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
@@ -9,7 +7,7 @@ xmin <-min(x1$PRTIME)
 xmax <-max(x1$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h1 = hist(x1$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
-x1_pt <- (h1$mids-1000)/1
+x1_pt <- (h1$mids-min(h1$mids))
 y1_raw_pt <- 1
 y1_pt <- 0
 z1_pt <- h1$counts/numSamples
@@ -23,7 +21,8 @@ xmin <-min(x2$PRTIME)
 xmax <-max(x2$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h2 = hist(x2$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
-x2_pt <- (h2$mids-2000)/2
+#x2_pt <- (h2$mids-2000)/2
+x2_pt <- (h2$mids-min(h2$mids))
 y2_raw_pt <- 2
 y2_pt <- 1
 z2_pt <- h2$counts/numSamples
@@ -37,7 +36,7 @@ xmin <-min(x4$PRTIME)
 xmax <-max(x4$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h4 = hist(x4$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
-x4_pt <- (h4$mids-4000)/4
+x4_pt <- (h4$mids-min(h4$mids))
 y4_raw_pt <- 4
 y4_pt <- 2
 z4_pt <- h4$counts/numSamples
@@ -52,7 +51,7 @@ xmin <-min(x8$PRTIME)
 xmax <-max(x8$PRTIME)
 nbins <- ceiling((xmax-xmin) / binsize)
 h8 = hist(x8$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
-x8_pt <- (h8$mids-8000)/8
+x8_pt <- (h8$mids-min(h8$mids))
 y8_raw_pt <- 8
 y8_pt <- 3
[…]

### 2017-09-08 pt_characterization/rep_check/u_s_time.tex
Updated descriptions
Draft for Sep 8

@@
 The x-, y-, z-axes
-indicate bins (PT range), task length, 
-and relative frequency of each histogram.
+indicate the normalized PT ranging from 0 to 100, task length in log scale, 
+and relative frequency of each histogram, such that for each INC it is scaled from 0 (the shortest) to 100 (the highest).
 
 \begin{figure}[htp!]
-\vspace{-.3in}
+%\vspace{-.3in}
 	\centering
 	%\includegraphics[scale=0.6,angle=270]{u_s_time/3d_plot.eps}\label{fig:3d_plot}
+	%\includegraphics[scale=0.6]{u_s_time/new_3d_plot}\label{fig:3d_plot}
 	\includegraphics[scale=0.6]{u_s_time/new_3d_plot}\label{fig:3d_plot}
@@
 	\subfigure[Coefficient of variation in log scale (relative): divisor=mean]{
-		\includegraphics[scale=0.6]{u_s_time/overall_re.eps}
+		\includegraphics[scale=0.6]{u_s_time/overall_pt_re.eps}
 		\label{fig:overall_re}

### 2017-09-13 pt_characterization/rep_check/1024_run.tex
Updated draft

@@
 
-\clearpage
-\newpage
+%\clearpage
+%\newpage
 
@@
 	\subfigure[ET frequency on INC1024-Run1]{
-		\includegraphics[scale=0.43]{1024_run/1024_sec_et_hist1.eps}
+		\includegraphics[scale=0.28]{1024_run/1024_sec_et_hist1.eps}
 		\label{fig:inc1024_run1_et}
@@
 	\subfigure[ET frequency on INC1024-Run2]{
-		\includegraphics[scale=0.43]{1024_run/1024_sec_et_hist2.eps}
+		\includegraphics[scale=0.28]{1024_run/1024_sec_et_hist2.eps}
 		\label{fig:inc1024_run2_et}
@@
 	\subfigure[ET frequency on INC1024-Run3]{
-		\includegraphics[scale=0.43]{1024_run/1024_sec_et_hist3.eps}
+		\includegraphics[scale=0.28]{1024_run/1024_sec_et_hist3.eps}
 		\label{fig:inc1024_run3_et}
@@
 \end{figure}
-
-\vspace\fill
-\clearpage
 
@@
 	\subfigure[PT frequency on INC1024-Run1]{
-		\includegraphics[scale=0.43]{1024_run/1024_sec_pt_hist1.eps}
+		\includegraphics[scale=0.28]{1024_run/1024_sec_pt_hist1.eps}
 		\label{fig:inc1024_run1_pt}
@@
 	\subfigure[PT frequency on INC1024-Run2]{
-		\includegraphics[scale=0.43]{1024_run/1024_sec_pt_hist2.eps}
+		\includegraphics[scale=0.28]{1024_run/1024_sec_pt_hist2.eps}
 		\label{fig:inc1024_run2_pt}
@@
 	\subfigure[PT frequency on INC1024-Run3]{
-		\includegraphics[scale=0.43]{1024_run/1024_sec_pt_hist3.eps}
+		\includegraphics[scale=0.28]{1024_run/1024_sec_pt_hist3.eps}
 		\label{fig:inc1024_run3_pt}

### 2017-09-13 pt_characterization/rep_check/res1.tex
Updated draft

@@
-\section{Histograms on the First Run~\label{sec:sodb9_r1_hist}} 
+\section{Histograms on the First Run~\label{sec:first_run}} 
 This section exhibits histograms on the first run of 
@@
 
-\section{Experiment Notes}
+%\section{Experiment Notes}
+\paragraph{Experiment Notes:}
 Table~\ref{tab:exp_notes1} provides a short description of our experimental runs, 
@@
 \clearpage
+
+\pagebreak
+\newpage
+
+\subsection{Additional Histograms}
+This section exhibits the histograms of INC with some intermediate task lengths under 256 secs. 
+These histograms are intended to investigate where the crossing and merge of two peaks that are consistently observed up to INC128 
+happened. Table~\ref{tab:exp_notes4} provides a description of the intermediate runs. 
+%and the second run of INC from 8192 seconds to 16384 seconds via EMPv5 with no Step 2.
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{2cm}|p{3cm}|p{3cm}|p{4cm}|p{3.5cm}|} \hline
+Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
+{\tt sodb9} &  INC96$\sim$INC256 & 1000 samples, each & 2017-05-24 $\sim$ 2017-06-06 & 
+Figs.~\ref{fig:ut_hist3}, \ref{fig:inc224_ut_hist}, and~\ref{fig:inc256_ut_hist}\\ \hline
+					&  INC3, 6, 12, 24, 48, 72, 80, 88, 104, 112, and 120 & 1000 samples, each & 2017-06-07 $\sim$ 2017-06-16 & Figs.~\ref{fig:inc3_ut_hist}, ~\ref{fig:inc6_ut_hist}, ~\ref{fig:inc12_ut_hist}, ~\ref{fig:inc24_ut_hist},
+~\ref{fig:inc48_ut_hist}, ~\ref{fig:inc72_ut_hist}, ~\ref{fig:inc80_ut_hist}, ~\ref{fig:inc88_ut_hist}, ~\ref{fig:inc104_ut_hist},~\ref{fig:inc112_ut_hist}, and~\ref{fig:inc120_ut_hist}\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Notes on experiment runs used for histograms\label{tab:exp_notes4}}
+\end{table}
+
+\subsubsection{ET}
+Not available at this point, due to the labshelf server's unavailability for the time being.
+
+\subsubsection{PT}
+%The histograms of INC1 through INC4096 are 
+%the same as those of Figures~\ref{fig:s9_r1_pt_hist1},~\ref{fig:s9_r1_pt_hist2},~\ref{fig:s9_r1_pt_hist3}, and~\ref{fig:s9_r1_pt_hist4}.
+
+\begin{figure}[hp!]
+	\centering
+	\subfigure[PT frequency on INC3]{
+		\includegraphics[scale=0.43]{u_s_time/3_sec_pt_hist.eps}
+		\label{fig:inc3_pt_hist}
+	}
+	\subfigure[PT frequency on INC6]{
+		\includegraphics[scale=0.43]{u_s_time/6_sec_pt_hist.eps}
+		\label{fig:inc6_pt_hist}
+	}
+	\subfigure[PT frequency on INC12]{
+		\includegraphics[scale=0.43]{u_s_time/12_sec_pt_hist.eps}
+		\label{fig:inc12_pt_hist}
+	}
+	\subfigure[PT frequency on INC24]{
+		\includegraphics[scale=0.43]{u_s_time/24_sec_pt_hist.eps}
+		\label{fig:inc24_pt_hist}
+	}
+	\caption{PT Histograms of INC3 ... INC24~\label{fig:new_pt_hist1}}
+\end{figure}
+
+\pagebreak
+\newpage
+
+\begin{figure}[hp!]
+	\centering
+	\subfigure[PT frequency on INC48]{
+		\includegraphics[scale=0.43]{u_s_time/48_sec_pt_hist.eps}
+		\label{fig:inc48_pt_hist}
+	}
+	\subfigure[PT frequency on INC72]{
+		\includegraphics[scale=0.43]{u_s_time/72_sec_pt_hist.eps}
+		\label{fig:inc72_pt_hist}
+	}
+	\subfigure[PT frequency on INC80]{
+		\includegraphics[scale=0.43]{u_s_time/80_sec_pt_hist.eps}
+		\label{fig:inc80_pt_hist}
+	}
+	\subfigure[PT frequency on INC88]{
+		\includegraphics[scale=0.43]{u_s_time/88_sec_pt_hist.eps}
+		\label{fig:inc88_pt_hist}
+	}
+	\caption{PT Histograms of INC48 ... INC88~\label{fig:new_pt_hist2}}
+\end{figure}
+
+\pagebreak
+\newpage
+
+\begin{figure}[hp!]
+	\centering
+	\subfigure[PT frequency on INC96]{
+		\includegraphics[scale=0.43]{u_s_time/96_sec_pt_hist.eps}
+		\label{fig:inc96_pt_hist}
+	}
+	\subfigure[PT frequency on INC104]{
+		\includegraphics[scale=0.43]{u_s_time/104_sec_pt_hist.eps}
+		\label{fig:inc104_pt_hist}
+	}
+	\subfigure[PT frequency on INC112]{
+		\includegraphics[scale=0.43]{u_s_time/112_sec_pt_hist.eps}
+		\label{fig:inc112_pt_hist}
+	}
+	\subfigure[PT frequency on INC120]{
+		\includegraphics[scale=0.43]{u_s_time/120_sec_pt_hist.eps}
+		\label{fig:inc120_pt_hist}
+	}
+	\caption{PT Histograms of INC96 ... INC120~\label{fig:new_pt_hist3}}
+\end{figure}
+
+\pagebreak
+\newpage
+
+\begin{figure}[hp!]
+	\centering
+	\subfigure[PT frequency on INC160]{
+		\includegraphics[scale=0.43]{u_s_time/160_sec_pt_hist.eps}
+		\label{fig:inc160_pt_hist}
+	}
+	\subfigure[PT frequency on INC192]{
+		\includegraphics[scale=0.43]{u_s_time/192_sec_pt_hist.eps}
+		\label{fig:inc192_pt_hist}
+	}
+	\subfigure[PT frequency on INC224]{
+		\includegraphics[scale=0.43]{u_s_time/224_sec_pt_hist.eps}
+		\label{fig:inc224_pt_hist}
+	}
+	\caption{PT Histograms of INC169 ... INC224~\label{fig:new_pt_hist4}}
+\end{figure}
+
+\pagebreak
+\newpage
+
+\subsection{Summary}
+
+\paragraph{Stacked Histograms in a 3D Plot:}
+Figure~\ref{fig:hist3d} represents a 3D plot of collecting the histograms of PT 
+exhibited in Section~\ref{sec:first_run}. 
+Note that the histograms from the task lengths of 8192 and 16384 seconds 
+to be shown in Figures~\ref{fig:inc8192_r2_hist_v5} and~\ref{fig:inc16384_r2_hist_v5}. 
+are brought into Figure~\ref{fig:hist3d} for completeness. The x-, y-, z-axes 
+indicate the normalized PT ranging from 0 to 100, task length in log scale, 
+and relative frequency of each histogram, such that for each INC 
+it is scaled from 0 (the shortest) to 100 (the highest).
+
+\begin{figure}[htp!]
+%\vspace{-.3in}
+	\centering
+	%\includegraphics[scale=0.6,angle=270]{u_s_time/3d_plot.eps}\label{fig:3d_plot}
+	%\includegraphics[scale=0.6]{u_s_time/new_3d_plot}\label{fig:3d_plot}
+	\includegraphics[scale=1]{u_s_time/new_3d_plot}\label{fig:3d_plot}
+	\caption{3D histograms on INC~\label{fig:hist3d}}
+\end{figure}
+
+\clearpage
+\pagebreak
+
+%\subsection{Absolute and Relative Variance over Increasing Task Lengths}
+\paragraph{Absolute and Relative Variance over Increasing Task Lengths:}
+Figure~\ref{fig:cv_inc} exhibits absolute and relative variances over increasing task lengths. 
+More specifically, Figures~\ref{fig:overall_std} and~\ref{fig:overall_std_log} 
+concern the PT standard deviations of all the runs including those two longest runs of INC8192 and INC16384 described in Table~\ref{tab:exp_notes2}. 
+The x-axis is task length, and the y-axis standard deviation; Figure~\ref{fig:overall_std_log} is taken in log scale.
+Figures~\ref{fig:overall_re} and~\ref{fig:overall_re2} shows 
+the relative variance of the same data set---{\em coefficient of variation} (= standard deviation 
+/ task length (or mean)). Both of the x and y axes in these two figures are taken in log scale. 
+We also overlap a linear-square-fit of each case on the same figure.
+
+\begin{figure}[htp!]
+	\centering
+	\subfigure[Standard deviation (absolute)]{
+		%\includegraphics[scale=0.6]{u_s_time/overall_std.eps}
+		\includegraphics[scale=0.6]{u_s_time/new_overall_pt_std.eps}
+		\label{fig:overall_std}
+	}
+	\subfigure[Standard deviation (absolute) in log scale]{
+		%\includegraphics[scale=0.6]{u_s_time/overall_std_log.eps}
+		\includegraphics[scale=0.6]{u_s_time/new_overall_pt_std_log.eps}
+		\label{fig:overall_std_log}
+	}
+	\subfigure[Coefficient of variation in log scale (relative): divisor=mean]{
+		\includegraphics[scale=0.6]{u_s_time/overall_pt_re.eps}
+		\label{fig:overall_re}
+	}
+	\subfigure[Coefficient of variation in log scale (relative): divisor=task length]{
+		\includegraphics[scale=0.6]{u_s_time/new_overall_pt_re.eps}
+		\label{fig:overall_re2}
+	}
+	\caption{Absolute and relative variances~\label{fig:cv_inc}}
+\end{figure}

### 2017-09-13 pt_characterization/rep_check/u_s_time.tex
Updated draft

--- 2017-08-11 pt_characterization/rep_check/u_s_time.tex
@@
-\section{Further Investigation of the Existing Runs and Some Additional Runs~\label{sec:u_s_time_hist}} 
-This section exhibits user and system time histograms on the first run of 
-INC with its task length increasing from 1 second to 1024 seconds 
-and the second run of INC from 2048 seconds to 16384 seconds via EMPv5 with no Step 2.
-The detailed description of the base data is from Table~\ref{tab:exp_notes1}, 
-and for some additional data, the corresponding description is given in Table~\ref{tab:exp_notes4}.
+\subsection{Histograms of User and System Time~\label{sec:u_s_time_hist}} 
+This section exhibits histograms on user and system time of the INC runs described in Table~\ref{tab:exp_notes1}.
 
-\begin{table}[h]
-\begin{center}
-\begin{tabular}{|p{2cm}|p{3cm}|p{3cm}|p{4cm}|p{3.5cm}|} \hline
-Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
-{\tt sodb9} &  INC96$\sim$INC256 & 1000 samples, each & 2017-05-24 $\sim$ 2017-06-06 & 
-Figs.~\ref{fig:ut_hist3}, \ref{fig:inc224_ut_hist}, and~\ref{fig:inc256_ut_hist}\\ \hline
-					&  INC3, 6, 12, 24, 48, 72, 80, 88, 104, 112, and 120 & 1000 samples, each & 2017-06-07 $\sim$ 2017-06-16 & Figs.~\ref{fig:inc3_ut_hist}, ~\ref{fig:inc6_ut_hist}, ~\ref{fig:inc12_ut_hist}, ~\ref{fig:inc24_ut_hist},
-~\ref{fig:inc48_ut_hist}, ~\ref{fig:inc72_ut_hist}, ~\ref{fig:inc80_ut_hist}, ~\ref{fig:inc88_ut_hist}, ~\ref{fig:inc104_ut_hist},~\ref{fig:inc112_ut_hist}, and~\ref{fig:inc120_ut_hist}\\ \hline
-%%%%% INC2048 and INC4096 come from the second run
-\end{tabular}
-\end{center}
-\vspace{-.2in}
-\caption{Notes on experiment runs used for histograms\label{tab:exp_notes4}}
-\end{table}
-
-\subsection{Summary of Histograms}
-
-\begin{figure}[htp!]
-\vspace{-.3in}
-	\centering
-	\includegraphics[scale=0.6,angle=270]{u_s_time/3d_plot.eps}\label{fig:3d_plot}
-	\caption{3D histograms on INC~\label{fig:hist3d}}
-\end{figure}
-
-\clearpage
-\pagebreak
-
-\subsection{Absolute and Relative Variance over Increasing Task Length}
-
-\begin{figure}[htp!]
-	\centering
-	\subfigure[Standard deviation (absolute)]{
-		%\includegraphics[scale=0.6]{u_s_time/overall_std.eps}
-		\includegraphics[scale=0.6]{u_s_time/new_overall_pt_std.eps}
-		\label{fig:overall_std}
-	}
-	\subfigure[Standard deviation (absolute) in log scale]{
-		%\includegraphics[scale=0.6]{u_s_time/overall_std_log.eps}
-		\includegraphics[scale=0.6]{u_s_time/new_overall_pt_std_log.eps}
-		\label{fig:overall_std_log}
-	}
-	\subfigure[Coefficient of variation in log scale (relative)]{
-		\includegraphics[scale=0.6]{u_s_time/overall_re.eps}
-		\label{fig:overall_re}
-	}
-	\subfigure[Coefficient of variation in log scale (relative): divisor=task length]{
-		\includegraphics[scale=0.6]{u_s_time/new_overall_pt_re.eps}
-		\label{fig:overall_re2}
-	}
-	\caption{Absolute and relative variances~\label{fig:cv_inc}}
-\end{figure}
-
-\pagebreak
-\newpage
-
-\subsection{Additional Histograms}
-
-\subsubsection{ET}
-Not available at this point, due to the labshelf server's unavailability for the time being.
-
-\subsubsection{PT}
-The histograms of INC1 through INC4096 are 
-the same as those of Figures~\ref{fig:s9_r1_pt_hist1},~\ref{fig:s9_r1_pt_hist2},~\ref{fig:s9_r1_pt_hist3}, and~\ref{fig:s9_r1_pt_hist4}. 
-The following are the histograms of INC with some intermediate task lengths.
-
-\begin{figure}[hp!]
-	\centering
-	\subfigure[PT frequency on INC3]{
-		\includegraphics[scale=0.43]{u_s_time/3_sec_pt_hist.eps}
-		\label{fig:inc3_pt_hist}
-	}
-	\subfigure[PT frequency on INC6]{
-		\includegraphics[scale=0.43]{u_s_time/6_sec_pt_hist.eps}
-		\label{fig:inc6_pt_hist}
-	}
-	\subfigure[PT frequency on INC12]{
-		\includegraphics[scale=0.43]{u_s_time/12_sec_pt_hist.eps}
-		\label{fig:inc12_pt_hist}
-	}
-	\subfigure[PT frequency on INC24]{
-		\includegraphics[scale=0.43]{u_s_time/24_sec_pt_hist.eps}
-		\label{fig:inc24_pt_hist}
-	}
-	\caption{PT Histograms of INC3 ... INC24~\label{fig:new_pt_hist1}}
-\end{figure}
-
-\pagebreak
-\newpage
-
-\begin{figure}[hp!]
-	\centering
-	\subfigure[PT frequency on INC48]{
-		\includegraphics[scale=0.43]{u_s_time/48_sec_pt_hist.eps}
-		\label{fig:inc48_pt_hist}
-	}
-	\subfigure[PT frequency on INC72]{
-		\includegraphics[scale=0.43]{u_s_time/72_sec_pt_hist.eps}
-		\label{fig:inc72_pt_hist}
-	}
-	\subfigure[PT frequency on INC80]{
-		\includegraphics[scale=0.43]{u_s_time/80_sec_pt_hist.eps}
-		\label{fig:inc80_pt_hist}
-	}
-	\subfigure[PT frequency on INC88]{
-		\includegraphics[scale=0.43]{u_s_time/88_sec_pt_hist.eps}
-		\label{fig:inc88_pt_hist}
-	}
-	\caption{PT Histograms of INC48 ... INC88~\label{fig:new_pt_hist2}}
-\end{figure}
-
-\pagebreak
-\newpage
-
-\begin{figure}[hp!]
-	\centering
-	\subfigure[PT frequency on INC96]{
-		\includegraphics[scale=0.43]{u_s_time/96_sec_pt_hist.eps}
-		\label{fig:inc96_pt_hist}
-	}
-	\subfigure[PT frequency on INC104]{
-		\includegraphics[scale=0.43]{u_s_time/104_sec_pt_hist.eps}
-		\label{fig:inc104_pt_hist}
-	}
-	\subfigure[PT frequency on INC112]{
-		\includegraphics[scale=0.43]{u_s_time/112_sec_pt_hist.eps}
-		\label{fig:inc112_pt_hist}
-	}
-	\subfigure[PT frequency on INC120]{
-		\includegraphics[scale=0.43]{u_s_time/120_sec_pt_hist.eps}
-		\label{fig:inc120_pt_hist}
-	}
-	\caption{PT Histograms of INC96 ... INC120~\label{fig:new_pt_hist3}}
-\end{figure}
-
-\pagebreak
-\newpage
-
-\begin{figure}[hp!]
-	\centering
-	\subfigure[PT frequency on INC160]{
-		\includegraphics[scale=0.43]{u_s_time/160_sec_pt_hist.eps}
-		\label{fig:inc160_pt_hist}
-	}
-	\subfigure[PT frequency on INC192]{
-		\includegraphics[scale=0.43]{u_s_time/192_sec_pt_hist.eps}
-		\label{fig:inc192_pt_hist}
-	}
-	\subfigure[PT frequency on INC224]{
-		\includegraphics[scale=0.43]{u_s_time/224_sec_pt_hist.eps}
-		\label{fig:inc224_pt_hist}
-	}
-	\caption{PT Histograms of INC169 ... INC224~\label{fig:new_pt_hist4}}
-\end{figure}
-
-\pagebreak
-\newpage
-
-\subsection{User Time}
+\subsubsection{User Time}
 
@@
 	}
-	\caption{User Time Histograms of INC104 and INC120~\label{fig:ut_hist5}}
+	\caption{User Time Histograms of INC104 and INC120: for the same 120-sec task length, 
+	we used {\it two different} bin sizes to see the influence of the choice of the bin size on the shape of histogram.~\label{fig:ut_hist5}}
 \end{figure}
@@
 
-\subsection{System Time}
+\subsubsection{System Time}
 
@@
 
+\pagebreak
 \clearpage
-\pagebreak
 
-\subsection{Correlation}
+\subsubsection{Correlational Analysis on Selected Runs}
+In this section we perform correlational analysis 
+between the user and system time of INC2048 and INC4096 runs 
+and selected measures of the daemons captured on the two runs.
 
@@
 
-\subsection{Scatter Plots on Some Significant Correlations}
+%\subsection{Scatter Plots on Some Significant Correlations}
+\paragraph{Scatter Plots on Some Significant Correlations:} 
 The following scatter plots correspond to the correlations bold in Table~\ref{tab:corr_table}.
@@
 \clearpage
+\pagebreak
 
-\subsection{Further Investigation of Some Samples}
+\paragraph{Detailed Measures on Some Significant Samples:} 
+The following tables show the breakdown of 
+measures of processes observed on some distinct samples (leftmost or rightmost) 
+in the selected runs.
 
-\subsubsection{Samples in Fig.~\ref{fig:corr_u_mnf}}
-
-\begin{table}[h]
+\begin{table}[htp!]
 \begin{center}
@@
 
-\begin{table}[h]
+\begin{table}[htp!]
 \begin{center}
@@
 
-\clearpage
-\newpage
-
-\subsubsection{Samples in Fig.~\ref{fig:corr_s_wbytes}}
-
-\begin{table}[h]
+\begin{table}[htp!]
 \begin{center}
@@
 \end{table}
+
+\clearpage
+\pagebreak

### 2017-09-13 pt_characterization/rep_check/u_s_time/compute-two-normals/fit.c

@@ -1,0 +1,503 @@
+/* fit.c, authored by Robert S. Maier <rsm@math.arizona.edu>.
+ * Compile by: 
+ * gcc -O fit.c -o fit -lm
+ */
+
+/* Typical invocations, for constructing a normal fit and a binormal
+   (i.e. mixture-of-two-normals) fit, to the frequency histograms
+   obtained from Young's data:
+
+   tail -n +2 256_sec.dat | awk '{print $2}' | sort -n |  fit
+   tail -n +2 64_sec.dat | awk '{print $2}' | sort -n |  fit -b 6 10 2 2 0.3
+
+   For a normal fit, no parameters are specified on the command line.
+   For a binormal fit, the "-b" option is specified, followed by
+   initializations of the 5 parameters of the binormal distribution:
+   mean1, mean2, sd1, sd2, scale1.  The fitting procedure will
+   optimize these 5 parameters, in an iterative manner that terminates
+   when no further improvement can be made.
+
+   Note that the mean1,mean2 parameters, as supplied on the command
+   line, should be relative to the minimum value in the dataset.  Thus
+   in the example above, `mean1=6' and `mean2=10' mean that the first
+   mode is (initially, before optimization) centered at 6 units above
+   the minimum value, and the second mode is centered at 10 units
+   above the minimum value.  These values were chosen by examining the
+   INC64 histogram.
+
+   In the example, `sd1=2' and `sd2=2' specify the initial width,
+   i.e. standard deviation, of each mode to be 2.  These values were
+   also chosen by examining the INC64 histogram.  `scale1=0.3' was
+   chosen because approximately 30% of the total seems to be in the
+   first mode, rather than the second.  (scale1 is the relative
+   strength of the first mode, and by definition, scale2 equals
+   1.0-scale1.)
+*/
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <errno.h>
+#include <limits.h>
+#include <math.h>
+
[…]

### 2017-09-14 pt_characterization/rep_check/u_s_time/new_3d_plot.r
User time only plot

@@ -3,10 +3,10 @@ numSamples <- 1000
 binsize=1
 ### 1 sec
 x1 = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
-xmin <-min(x1$PRTIME)
-xmax <-max(x1$PRTIME)
+xmin <-min(x1$U_TIME)
+xmax <-max(x1$U_TIME)
 nbins <- ceiling((xmax-xmin) / binsize)
-h1 = hist(x1$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+h1 = hist(x1$U_TIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 x1_pt <- round((h1$breaks-min(h1$breaks))/(max(h1$breaks-min(h1$breaks)))*100,0)
 x1_pt
 z1_pt <- round(h1$counts/max(h1$counts)*100,0)
@@ -14,10 +14,10 @@ z1_pt
 ### 2 sec
 x2 = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
 x2 <- subset(x2, x2$ITERNUM <= numSamples)
-xmin <-min(x2$PRTIME)
-xmax <-max(x2$PRTIME)
+xmin <-min(x2$U_TIME)
+xmax <-max(x2$U_TIME)
 nbins <- ceiling((xmax-xmin) / binsize)
-h2 = hist(x2$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+h2 = hist(x2$U_TIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 x2_pt <- round((h2$breaks-min(h2$breaks))/(max(h2$breaks-min(h2$breaks)))*100,0)
 x2_pt
 z2_pt <- round(h2$counts/max(h2$counts)*100,0)
@@ -25,10 +25,10 @@ z2_pt
 ### 4 sec
 x4 = read.csv(file="4_sec.dat",head=TRUE,sep="\t")
 x4 <- subset(x4, x4$ITERNUM <= numSamples)
-xmin <-min(x4$PRTIME)
-xmax <-max(x4$PRTIME)
+xmin <-min(x4$U_TIME)
+xmax <-max(x4$U_TIME)
 nbins <- ceiling((xmax-xmin) / binsize)
-h4 = hist(x4$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
[…]

### 2017-09-19 mlppi-ksii17/review/review_response.tex
MLPPI revision

\documentclass[10pt,letterpaper]{article}

\usepackage{graphicx}
\usepackage{longtable}
\usepackage{algorithmic}
\usepackage{alltt}
\usepackage{array}
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
%\usepackage[caption=false]{caption}
\usepackage[dvipsnames]{color}
%\usepackage{colortbl}
%\usepackage{enumitem}
%\usepackage{eqparbox}
%\usepackage{fixltx2e}
%\usepackage{float}
%\usepackage{floatflt}
%\usepackage{mdwmath}
%\usepackage{mdwtab}
%\usepackage{multirow}
\usepackage{stfloats}
%\usepackage[font=footnotesize]{subfig}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\usepackage[tight,normalsize,sf,SF]{subfigure}
%\usepackage{subfigure} 
\usepackage{times}
\usepackage{url}
\usepackage{verbatim} 
\usepackage{wrapfig}
\usepackage{xspace}
\usepackage{color}

\definecolor{grey}{RGB}{200,200,200}
\newcommand{\hilite}[1]{\colorbox{yellow}{#1}}
\newcommand{\hiliting}[1]{\colorbox{grey}{#1}}
\long\def\todo#1{\hilite{{\bf TODO:} {\em #1}}}

\setlength{\oddsidemargin}{0in}
\setlength{\textwidth}{6.5in}

% new commands: from file -- to permit re-use
%\input{newcommands.tex}

%Hyphenation problems
\hyphenation{name-space}
\hyphenation{schema-location}
\hyphenation{xml-lint}
\hyphenation{time-stamp}
\hyphenation{time-stamps}
\hyphenation{time-stamped}

\urlstyle{rm}
\graphicspath{{./figures/}}

\definecolor{gray}{rgb}{0.5,0.5,0.5}
%\newcommand{\rev}[1]{\textcolor{blue}{#1}}
%\newcommand{\quo}[1]{\textcolor{gray}{#1}}
%\newcommand{\rev}[1]{\vspace{3mm} \noindent {\em {\color{blue}{#1}}}}

\newenvironment{myindentpar}[1]%
{\begin{list}{}
         {\vspace{10pt}
					\setlength{\leftmargin}{#1}}
          \item[]
}
{\end{list}}
\newcommand{\rev}[1]{\begin{myindentpar}{.25in} {\em {\color{blue}{#1}}}\end{myindentpar}}

\newenvironment{myindentparQ}[1]%
{\begin{list}{}
         {\setlength{\leftmargin}{#1}}
          \item[]
}
{\end{list}}
\newcommand{\quo}[1]{\begin{myindentparQ}{.25in} {{\color{gray}{#1}}}\end{myindentparQ}}

\begin{document}

\title{Responses to Editor-in-Chief \& Reviewer Comments}
\author{}
\maketitle

\section*{Overview}\label{sec:overview}
We appreciate these quite thorough reviews. We appreciate the feedback and suggestions.

\newcounter{RP}

\clearpage
\section*{Editor-in-Chief}\label{sec:ed1}

Thank you for these detailed comments, which we respond to individually below.

\rev{
Comments from Editor:

15-Sep-2017

Dear Dr. Young-Kyoon Suh:

Manuscript ID TIIS-IS-2017-Jul-0844 entitled ``MLPPI Wizard: An Automated Multi-level Partitioning Tool on Analytical Workloads" which you submitted to the KSII Transactions on Internet and Information Systems, has been reviewed.  The comments of the reviewer(s) are included at the bottom of this letter.

I am happy to inform you that the reviewer(s) have recommended publication, but also suggest some revisions to your manuscript.  Therefore, I invite you to respond to the reviewer(s)' comments and revise your manuscript.   Please go through the reviewers' comments carefully and then prepare for the revised paper and authors' response. Your revised paper will not be guaranteed to be accepted for publication in the TIIS journal. The editor and reviewers will again review the revised paper and authors’ response.

}

In this response letter we address each of the reviewers' comments. 
We also incorporate into the revised paper our response to each comment.

\rev{
To upload your revised manuscript, log into {\url{https://mc.manuscriptcentral.com/tiisjournal}} and enter your Author Center, where you will find your manuscript title listed under ``Manuscripts with Decisions."  Under ``Actions," click on ``Create a Revision."  Your manuscript number has been appended to denote a revision.
}

We used this procedure for uploading the revision.

\rev{
You may also click the below link to start the revision process (or continue the process if you have already started your revision) for your manuscript. If you use the below link, you will not be required to login to ScholarOne Manuscripts.

*** PLEASE NOTE: This is a two-step process. After clicking on the link, you will be directed to a webpage to confirm. ***

{\url{https://mc.manuscriptcentral.com/tiisjournal?URL_MASK=963223d1218947b5b5750544c56a74b6}}
}

Thank you for this additional information.

\pagebreak

\rev{
1. Revise your manuscript "by using the MS Word."

2. Please highlight the changes to your manuscript within the document by using bold and colored text "in BLUE" or by using the track changes mode in MS Word.

3. In addition to the revised manuscript, the authors are required to write the authors’ response answering the reviewers’ comments. In order to expedite the processing of the revised manuscript, please be as specific as possible in your response to the reviewer(s)' comments.

4. The authors’ response should be included in "the first page of the revised manuscript."

Once the revised manuscript is prepared, you can upload it and submit it through your Author Center.
}

Yes, we used the MS word program for revising our article. As requested, anything new or revised compared to the original submission is colored in {\color{blue}blue} and marked in \textbf{bold} face in the revised. In addition, corrections made are colored in {\color{red}red} for distinction. 
As required, we provide this written response letter to answer each reviewer's comments.
Finally, we include the letter in the first page of the revised article.

\rev{
[IMPORTANT]:  Your original files are available to you when you upload your revised manuscript.  Please delete any redundant files before completing the submission.

Because we are trying to facilitate timely publication of manuscripts submitted to the KSII Transactions on Internet and Information Systems, your revised manuscript should be uploaded "by 15-Oct-2017."  If it is not possible for you to submit your revision in the deadline, we may have to consider your paper as a new submission.

Once again, thank you for submitting your manuscript to the KSII Transactions on Internet and Information Systems and I look forward to receiving your revision.

Sincerely,

Dr. Mohammad Shojafar

Editor,

mohammad.shojafar@uniroma1.it

KSII Transactions on Internet and Information Systems

WWW.ITIIS.ORG
}

We deleted any redundant files during the uploading process. Today ({\bf\underline{October 8th, 2017 in KST}}), the revised manuscript has been uploaded for further review, 
and thus, there is no issue with the specified deadline of October 15, 2017.

\clearpage
\section*{Reviewer 1}\label{sec:rev1}

% 1
\rev{
[Reviewer(s)' Comments to Author]:
Reviewer: 1

Comments to the Author
Generally Fig1. is not a figure scale and can be considered as a structure rather aha a Fig. 1. technically, the primary figures are the general scale and structure of the method which tends to grasp the authors' attention and must be clarified. 
}

\rev{
The presented Wizard Architecture in Fig. 2 needs to be expanded. 
}

\rev{
The table in Page3 does not have any caption to introduces besides, its location is badly organized.
}

\rev{
The main contribution of the paper is so limited and needs to be extended in Section 1. 
}

\rev{ 
In the experiment environmental setting, the deployment in realistic datasets are missing and must be indicated for the reproduction of the future readers.  According to the bar-shaped results, the WIZARD does not provide optimal results compared to the state-of the arts. They must clearly be expressed how this cases can be used in the real-time scenarios. 
}

\rev{
Also, Authors should give a more clear definition of the "application type" used in the paper. For example, what information of applications are used and how to obtain this information. 
}

\rev{
Last but not least backs to the background which is limited and some hot-related works are missing that are listed in i)"Using imperialist competition algorithm for independent task scheduling in grid computing" and ii) "A hybrid metaheuristic algorithm for Job scheduling on computational grids" that addressed the analytical workload exploitation in the real case studies that must be added in the background to make your work holistic for the future readers.
}

\clearpage
\section*{Reviewer 2}\label{sec:rev2}

% 2
\rev{
Reviewing: 2

Comments to the Author
The paper could be published after addressing following comments:
1. Abstract could be better write, in this reviewer point of view, the current abstract is not comprehensive.
}

\rev{
2. The manuscript could be better organize as well. For instance, the related works section could be placed after introduction section no before conclusion.
}

\rev{
3. Providing a comparison table could be useful for readers.
}

\rev{
4. Conclusion is too long.
}

\rev{
5. Introduction section need more references.
}

\end{document}

### 2017-09-21 mlppi-ksii17/review/review_response.tex
paper revised

@@
 }
+The table is now denoted as Table~1 with its corresponding caption provided. We also placed it at the top of page~3, which looks the best location for that.
 
@@
 In the experiment environmental setting, the deployment in realistic datasets are missing and must be indicated for the reproduction of the future readers.  
+}
+
+\rev{
 According to the bar-shaped results, the WIZARD does not provide optimal results compared to the state-of the arts.
@@
 }
+We now have some mentions of the above two works compared to our work. Also, we added these two in the references of the revised manuscript.
 
@@
 }
+A great suggestion. We have relocated the related work right after the introduction in the revision. Note that there are some references added in the related work for completeness.
 

### 2017-09-21 pt_characterization/rep_check/res1.tex
User time only plot
Draft for Sep 15
binormal fit

@@
 
+\pagebreak
+
 \subsubsection{PT}
@@
 
+\pagebreak
+\clearpage
+Figure~\ref{fig:hist3d_u} represents a 3D plot of collecting the histograms of user time only
+exhibited in Section~\ref{sec:first_run}. 
+\begin{figure}[htp!]
+%\vspace{-.3in}
+	\centering
+	\includegraphics[scale=1]{u_s_time/3dplot_utime_only}\label{fig:3d_plot_u}
+	\caption{3D histograms on INC: u time only~\label{fig:hist3d_u}}
+\end{figure}
+
+\clearpage
+\pagebreak
+
+\paragraph{Fitting the Histograms in Bi-normal Distributions:}
+We ran Rob's code to obtain a bi-normal fit for each INC's run. 
+Table~\ref{tab:binormal_fit} exhibits the two means (modes), standard deviations, and scales for each of the INC runs.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|l|l|l|l|l|l|l|} \hline
+INC        & Mean 1 & Mean 2 & Std 1 & Std 2 & Scale 1 & Scale 2\\ \hline
+INC1      & 1.66  & 4.73 & 1.03 & 0.54 & 0.104788 & 0.895212\\ \hline
+INC2      & 2.33  & 7.07 & 2.38 & 0.54 & 0.159807 & 0.840193\\ \hline
+INC4      & 2.17  & 7.89 & 0.99 & 0.86 & 0.209242 & 0.790758\\ \hline
+INC8      & 3.51  & 9.15 & 1.10 & 1.08 & 0.388976 & 0.611024\\ \hline
+INC16    & 4.04  & 9.53 & 0.98 & 1.28 & 0.828952 & 0.171048\\ \hline
+INC32    & 2.76  & 8.26 & 1.28 & 1.12 & 0.725687 & 0.274313\\ \hline
+INC64    & 6.36  & 12.02 & 1.64 & 1.62 &  0.347095 & 0.652905\\ \hline
+INC128  & 4.21  & 8.64 & 1.82 & 3.40 & 0.410859 & 0.589141\\ \hline
+INC256  & 6.08  & 14.61  & 3.59 & 3.75 & 0.124312 & 0.875688\\ \hline
+INC512  & 1.60  & 12.43   & 11.51 & 4.47 & 0.075758 & 0.924242\\ \hline
+INC1024 & 1.13  & 18.88  & 14.99 & 6.01 & 0.055255 & 0.944745\\ \hline
+INC2048 & 29.93 & 52.92 & 8.53 & 6.59 & 1.001873 & -0.00187\\ \hline
+INC4096 & 39.57 & 98.00 & 12.41 & 8.19 & 0.939756 & 0.060244\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Two means in mode computed for each INC~\label{tab:binormal_fit}}
+\end{table}
+
 \clearpage

### 2017-09-22 pt_characterization/rep_check/res1.tex
09/22 draft

@@
 
-\paragraph{Fitting the Histograms in Bi-normal Distributions:}
+\paragraph{Fitting Histograms in Bi-Normal Distributions:}
 We ran Rob's code to obtain a bi-normal fit for each INC's run. 
@@
 \end{table}
+
+\pagebreak
+\clearpage
+
+\begin{figure}[htp!]
+	\centering
+	\subfigure[Mean]{
+		\includegraphics[scale=0.6]{u_s_time/binormal_fit_mean.eps}
+		\label{fig:binormal_mean}
+	}
+	\subfigure[Standard Deviation]{
+		\includegraphics[scale=0.6]{u_s_time/binormal_fit_std.eps}
+		\label{fig:binormal_std}
+	}
+	\subfigure[Scale]{
+		\includegraphics[scale=0.6]{u_s_time/binormal_fit_scale.eps}
+		\label{fig:binormal_scale}
+	}
+	\caption{Plots of two means, standard deviations, and scales for each INC 
+	over increasing task length: based on Table~\ref{tab:binormal_fit}\label{fig:bnormal_Fit}}
+\end{figure}
 

### 2017-09-24 pt_characterization/rep_check/res1.tex
PT draft modified

@@
 
-\subsection{PT}
+\subsection{PT~\label{sec:1st_pt}}
 
@@
 Figure~\ref{fig:hist3d} represents a 3D plot of collecting the histograms of PT 
-exhibited in Section~\ref{sec:first_run}. 
+exhibited in Section~\ref{sec:1st_pt}. 
 Note that the histograms from the task lengths of 8192 and 16384 seconds 
@@
 	\includegraphics[scale=1]{u_s_time/3dplot_utime_only}\label{fig:3d_plot_u}
-	\caption{3D histograms on INC: u time only~\label{fig:hist3d_u}}
+	\caption{3D histograms on INC: user time only~\label{fig:hist3d_u}}
 \end{figure}

### 2017-09-28 mlppi-ksii17/review/review_response.tex
KSII updates

--- 2017-09-19 mlppi-ksii17/review/review_response.tex
@@
 
+We totally agree with you. The old Fig. 1 is not at a real figure. We feel that it had better be put into a table for further reference, which is now {\color{blue}\underline{Tab. 1}}.
+
 \rev{
@@
 
+We significantly expanded the architecture, which is now {\color{blue}\underline{Fig. 1}}, with a specified description of each phase. In addition, 
+the figure now includes (1) the query workload example on the top and (2) the output solution 
+produced by the wizard on the bottom, for better clarification.
+
 \rev{
@@
 }
+Thanks for the comments. To address your concern, we put in tabular form not only that table with no caption but also 
+some other provided examples. We then positioned those new tables on the top of each page in the introduction in the revised.
 
@@
 
+We took out some contributions mentioned in Section 2 (now Section 3) and then incorporated them into the contribution paragraph (on page~5 
+in the revision), which is now substantially extended.
+
 \rev{ 
 In the experiment environmental setting, the deployment in realistic datasets are missing and must be indicated for the reproduction of the future readers.  
+}
+
+\rev{
 According to the bar-shaped results, the WIZARD does not provide optimal results compared to the state-of the arts.
@@
 }
+OK, we now have some mentions of the above two works compared to our work. Also, we added these two in the references in the revision.
 
@@
 }
+A great suggestion.
+We have relocated the related work section right after the introduction in the revised manuscript.
+Note that there are also some  references newly added in that section for completeness.
 
@@
 }
+We've added some references in the contribution paragraph in the introduction on page 5.
 

### 2017-09-29 mlppi-ksii17/review/review_response.tex
Revised article

@@
 
-We significantly expanded the architecture, which is now {\color{blue}\underline{Fig. 1}}, with a specified description of each phase. In addition, 
-the figure now includes (1) the query workload example on the top and (2) the output solution 
-produced by the wizard on the bottom, for better clarification.
+We significantly expanded the architecture, which is now {\color{blue}\underline{Fig.
+1}} on page~7, with a specified description of each phase.
+In addition, the figure now includes (1) the query workload text exemplified in the paper (on the top) and (2) the output solution
+produced by the wizard (on the bottom), for better clarification.
 
@@
 }
-Thanks for the comments. To address your concern, we put in tabular form not only that table with no caption but also 
-some other provided examples. We then positioned those new tables on the top of each page in the introduction in the revised.
+Thanks for the comments. To address your concern, we put in tabular form the table with its caption on page~2 (now, {\color{blue}Tab.~2}). 
+Note that we tabularized other examples for better exposition in the introduction.
 
@@
 } 
-
 We took out some contributions mentioned in Section 2 (now Section 3) and then incorporated them into the contribution paragraph (on page~5 
@@
 
+%Even though our query workloads (similar to {\em q1} in Tab.~3) are synthetic, the queries were generated from a realistic query template, 
+%which is from customer cases like reports and form templates.
+The queries are of type joining a fact table and a dimension table in a popular benchmark, representing a real analytical workload.
+In addition, the proposed wizard allows for general expressions for partitioning definition as addressed in the solution paragraph in the introduction.
+Hence, the results on our workloads do not lose their validity on a real environment.
+Please refer to the third and subsequent paragraphs in Section~5.1 (on page 16).
+The synthetic queries used for our experiments were generated from a realistic query template, which is from customer cases like reports and form templates.
+The queries are of type joining a fact table and a dimension table in a popular benchmark, representing a real analytical workload.
+In addition, the proposed wizard allows for general expressions for partitioning definition as addressed in the solution paragraph in the introduction.
+Hence, the results on our workloads do not lose their validity on a real environment.
+For more details, please refer to the Solution paragraph in the introduction and the third and subsequent paragraphs in Section~5.1.
+
 \rev{
@@
 
+Your confusion comes from Figure 5 in the original submission.
+The proposed wizard obviously outperforms the compared approaches: no partitioning (NO PPI) and a human expert (EXP).
+For better clarification, we removed that figure and instead provided a comparison table.
+Again, our query workloads are from the real benchmark, SSB, and our algorithm is still applicable to any other type of analytical workload, since the search space is driven by the query predicates extracted from the workload.
+
 \rev{
-Also, Authors should give a more clear definition of the "application type" used in the paper. For example, what information of applications are used and how to obtain this information. 
+Also, Authors should give a more clear definition of the ``application type" used in the paper. For example, what information of applications are used and how to obtain this information. 
 }
+
+We understand your concern.
+The application type of the proposed wizard is a \underline{physical database design tool} (fundamentally different from several existing works [4--6].
+Also note that any complex workload, giving many different partitioning options, can benefit from the automated wizard to find almost an optimal solution.
+For more details, please refer to the first, third, and fourth bullets in the Contribution paragraphs in the Introduction.
 
@@
 }
+
 OK, we now have some mentions of the above two works compared to our work. Also, we added these two in the references in the revision.
@@
 
+Yes, we do now provide Tab. 9 for better clarity on the comparison.
+
 \rev{

### 2017-09-30 pt_characterization/rep_check/res1.tex
PT characterizing

@@
 \paragraph{Stacked Histograms in a 3D Plot:}
-Figure~\ref{fig:hist3d} represents a 3D plot of collecting the histograms of PT 
-exhibited in Section~\ref{sec:1st_pt}. 
-Note that the histograms from the task lengths of 8192 and 16384 seconds 
-to be shown in Figures~\ref{fig:inc8192_r2_hist_v5} and~\ref{fig:inc16384_r2_hist_v5}. 
-are brought into Figure~\ref{fig:hist3d} for completeness. The x-, y-, z-axes 
-indicate the normalized PT ranging from 0 to 100, task length in log scale, 
-and relative frequency of each histogram, such that for each INC 
-it is scaled from 0 (the shortest) to 100 (the highest).
+Figure~\ref{fig:hist3d} represents a 3D plot of collecting the histograms of PT in Section~\ref{sec:1st_pt}. 
+Note that we additionally include the histograms from the task lengths of 8192 and 16384 seconds, which will be shown in Figures~\ref{fig:inc8192_r2_hist_v5} and~\ref{fig:inc16384_r2_hist_v5} (on page~47).
+See the legend in the right hand side of the figure, for more details about specific task length information.
+%To obtain the better shape, we remove from those histograms outliers identified via the second step of EMPv5.
+The x-axis corresponds to the normalized PT relative to the minimum, ranging 0 to 100, the y-axis to task length in log scale, and the z-axis to the normalized frequency relative to the highest bar for each histogram.
+
+There are several interesting things to notice.
+First, INC32 has some empty bins in the raw histogram in Figure ~\ref{fig:inc32_r1_et_hist_v5}, thereby zero value on the x-axis for that histogram in brown in Figure~\ref{fig:hist3d}.
 
@@
 	\includegraphics[scale=1]{u_s_time/new_3d_plot}\label{fig:3d_plot}
-	\caption{3D histograms on INC~\label{fig:hist3d}}
+	\caption{3D histograms of PT on doubly-increasing INC~\label{fig:hist3d}}
 \end{figure}
@@
 
-\paragraph{Fitting Histograms in Bi-Normal Distributions:}
-We ran Rob's code to obtain a bi-normal fit for each INC's run. 
-Table~\ref{tab:binormal_fit} exhibits the two means (modes), standard deviations, and scales for each of the INC runs.
+\paragraph{Fitting PT Histograms in Bi-Normal Distributions:}
+We ran Rob's code to obtain a bi-normal fit for each INC's PT histogram 
+in Section~\ref{sec:first_run}. 
+Table~\ref{tab:binormal_fit} exhibits the two means (modes), standard deviations, and scales for the PT histogram associated with each INC.
 

### 2017-09-30 pt_characterization/rep_check/final_data/combined.r

@@ -1,0 +1,48 @@
+x1_1 = read.csv(file="repet_data1/1_sec.dat",head=TRUE,sep="\t")
+x1_2 = read.csv(file="repet_data2/1_sec.dat",head=TRUE,sep="\t")
+x1_ <- c(x1_1$PRTIME, x1_2$PRTIME)
+x1_ <- cbind(x1_)
+x1_n <- seq.int(nrow(x1_))
+x1_ <- cbind(x1_n,x1_)
+colnames(x1_) <- c("NUM", "PRTIME")
+write.table(x1_, "final_data/1_sec.dat", sep="\t",row.names = F)
+x1_up = mean(x1_[,2]) + 2*sd(x1_[,2])
+x1_dn = mean(x1_[,2]) - 2*sd(x1_[,2])
+x1_ = subset(x1_, x1_[,2] >= x1_dn & x1_[,2] <= x1_up)
+xmax <- max(x1_[,2])
+xmin <- min(x1_[,2])
+binsize <- 1
+nbins <- ceiling((xmax-xmin) / binsize)
+h1_ = hist(x1_[,2], right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+h1 = h1_
+x1_pt <- round((h1_$breaks-min(h1_$breaks))/(max(h1_$breaks-min(h1_$breaks)))*100,0)
+x1_pt
+z1_pt <- round(h1_$counts/max(h1_$counts)*100,0)
+z1_pt
+
+
+
+## 2 sec
+x2_1 = read.csv(file="repet_data1/2_sec.dat",head=TRUE,sep="\t")
+x2_2 = read.csv(file="repet_data2/2_sec.dat",head=TRUE,sep="\t")
+x2 <- c(x2_1$PRTIME, x2_2$PRTIME)
+x2 <- cbind(x2)
+x2_n <- seq.int(nrow(x2))
+x2 <- cbind(x2_n,x2)
+colnames(x2) <- c("NUM", "PRTIME")
+write.table(x2, "final_data/2_sec.dat", sep="\t",row.names = F)
+x2_up = mean(x2[,2]) + 2*sd(mean(x2[,2]))
+x2_dn = mean(x2[,2]) + 2*sd(mean(x2[,2]))
+xmax <- max(x2[,2])
+xmin <- min(x2[,2])
+binsize <- 1
+nbins <- ceiling((xmax-xmin) / binsize)
[…]

### 2017-09-30 pt_characterization/rep_check/final_data/combined_3d_hist.r

@@ -1,0 +1,55 @@
+library(plotly)
+INC1 <- list(
+  x = c(0,0,0,0,0,0,0,0), 
+  y = c(0,25,25,50,50,75,75,100),
+  z = c(5,5,43,43,100,100,9,9),
+  line = list(
+    color = "Purpule", 
+    width = 10
+  ), 
+  mode = "lines", 
+  type = "scatter3d"
+)
+INC2 <- list(
+  x = c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1), 
+  y = c(0,10,10,20,20,30,30,40,40,50,50,60,60,70,70,80,80,90,90,100),
+  z = c(0,5,5,5,4,4,4,4,3,3,6,6,15,15,100,100,23,23,1,1),
+  line = list(
+    color = "Purpule", 
+    width = 10
+  ), 
+  mode = "lines", 
+  type = "scatter3d"
+)
+#data <- list(INC1, INC2, INC4, INC8, INC16, INC32, INC64,INC128,INC256,INC512,INC1024,INC2048,INC4096)
+data <- list(INC1)
+layout <- list(
+  autosize = FALSE, 
+  height = 600, 
+  title = "3d Stacked Histogram", 
+  width = 600
+)
+p <- plot_ly()
+p <- add_trace(p, x=INC1$x, y=INC1$y, z=INC1$z, line=INC1$line, mode=INC1$mode, type=INC1$type, name = 'INC1')
+#p <- add_trace(p, x=INC2$x, y=INC2$y, z=INC2$z, line=INC2$line, mode=INC2$mode, type=INC2$type, name = 'INC2')
+#p <- add_trace(p, x=INC4$x, y=INC4$y, z=INC4$z, line=INC4$line, mode=INC4$mode, type=INC4$type, name = 'INC4')
+#p <- add_trace(p, x=INC8$x, y=INC8$y, z=INC8$z, line=INC8$line, mode=INC8$mode, type=INC8$type, name = 'INC8')
+#p <- add_trace(p, x=INC16$x, y=INC16$y, z=INC16$z, line=INC16$line, mode=INC16$mode, type=INC16$type, name = 'INC16')
+#p <- add_trace(p, x=INC32$x, y=INC32$y, z=INC32$z, line=INC32$line, mode=INC32$mode, type=INC32$type, name = 'INC32')
[…]

### 2017-09-30 pt_characterization/rep_check/u_s_time/scratch.r

@@ -1,3 +1,12 @@
+#### FOR OUTPUT
+all_data <- cbind(all_inc,all_ss,ny_raw,nz_raw,ny,nz)
+colnames(all_data) <- c("TASK_LEN (in s)","SAMPLE_SIZE","RAW_BIN (in ms)","RAW_FREQ","NOR_PT","REL_FREQ")
+write.table(all_data, "all_freq_data.txt", sep="\t",row.names = F)
+
+x1 = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
+
+
+
 nx <- c(x1_pt,x2_pt,x4_pt,x8_pt)
 ny <- c(y1_pt,y2_pt,y4_pt,y8_pt)
 nz <- c(z1_pt,z2_pt,z4_pt,z8_pt)

### 2017-09-30 mlppi-ksii17/review/review_response.tex
MLPPI paper

@@
 Please refer to the third and subsequent paragraphs in Section~5.1 (on page 16).
-The synthetic queries used for our experiments were generated from a realistic query template, which is from customer cases like reports and form templates.
+%The synthetic queries used for our experiments were generated from a realistic query template, which is from customer cases like reports and form templates.
 The queries are of type joining a fact table and a dimension table in a popular benchmark, representing a real analytical workload.
-In addition, the proposed wizard allows for general expressions for partitioning definition as addressed in the solution paragraph in the introduction.
+%In addition, the proposed wizard allows for general expressions for partitioning definition as addressed in the solution paragraph in the introduction.
 Hence, the results on our workloads do not lose their validity on a real environment.
 For more details, please refer to the Solution paragraph in the introduction and the third and subsequent paragraphs in Section~5.1.
+Please see the paragraph of `Workload Generation \& Deployment' in Section~5.1: more specifically, the third---fifth
+paragraphs in that section, where we address the use of synthetic queries in general from to represent a realistic scenario and have a mention of reproducibility of our work.
 
@@
 
-We understand your concern.
 The application type of the proposed wizard is a \underline{physical database design tool} (fundamentally different from several existing works [4--6].
@@
 
+OK, we rewrote the abstract.
+
 \rev{
@@
 We have relocated the related work section right after the introduction in the revised manuscript.
-Note that there are also some  references newly added in that section for completeness.
+Note also that there are some new references mentioned in the last paragraph of the related work.
 
@@
 
+We shortened the Conclusion section. Please check page~20.
+
 \rev{

### 2017-10-06 pt_characterization/rep_check/res1.tex
MLPPI paper
updates on the draft

@@
 See the legend in the right hand side of the figure, for more details about specific task length information.
-%To obtain the better shape, we remove from those histograms outliers identified via the second step of EMPv5.
+(To obtain the better shape, we could remove from those histograms the outliers that are identified via the second step of EMPv5.)
 The x-axis corresponds to the normalized PT relative to the minimum, ranging 0 to 100, the y-axis to task length in log scale, and the z-axis to the normalized frequency relative to the highest bar for each histogram.
 
-There are several interesting things to notice.
-First, INC32 has some empty bins in the raw histogram in Figure ~\ref{fig:inc32_r1_et_hist_v5}, thereby zero value on the x-axis for that histogram in brown in Figure~\ref{fig:hist3d}.
+One thing to notice is that there are some runs having a few empty bins 
+in their respective histogram. For instance, INC32 has such an empty bin as shown in Figure~\ref{fig:inc32_r1_et_hist_v5}. 
+Same with INC256 in Figure~\ref{fig:inc256_r1_et_hist_v5}. For such an INC run, the z-axis associated with its histogram has the zero value on the x-axis in Figure~\ref{fig:hist3d}.
 
@@
 in Section~\ref{sec:first_run}. 
-Table~\ref{tab:binormal_fit} exhibits the two means (modes), standard deviations, and scales for the PT histogram associated with each INC.
+Table~\ref{tab:binormal_fit} exhibits the total range of bins, the two means (modes), standard deviations, scales (weights) 
+for the PT histogram, and figure numbers associated with each INC.
 
@@
 \begin{center}
-\begin{tabular}{|l|l|l|l|l|l|l|} \hline
-INC        & Mean 1 & Mean 2 & Std 1 & Std 2 & Scale 1 & Scale 2\\ \hline
-INC1      & 1.66  & 4.73 & 1.03 & 0.54 & 0.104788 & 0.895212\\ \hline
-INC2      & 2.33  & 7.07 & 2.38 & 0.54 & 0.159807 & 0.840193\\ \hline
-INC4      & 2.17  & 7.89 & 0.99 & 0.86 & 0.209242 & 0.790758\\ \hline
-INC8      & 3.51  & 9.15 & 1.10 & 1.08 & 0.388976 & 0.611024\\ \hline
-INC16    & 4.04  & 9.53 & 0.98 & 1.28 & 0.828952 & 0.171048\\ \hline
-INC32    & 2.76  & 8.26 & 1.28 & 1.12 & 0.725687 & 0.274313\\ \hline
-INC64    & 6.36  & 12.02 & 1.64 & 1.62 &  0.347095 & 0.652905\\ \hline
-INC128  & 4.21  & 8.64 & 1.82 & 3.40 & 0.410859 & 0.589141\\ \hline
-INC256  & 6.08  & 14.61  & 3.59 & 3.75 & 0.124312 & 0.875688\\ \hline
-INC512  & 1.60  & 12.43   & 11.51 & 4.47 & 0.075758 & 0.924242\\ \hline
-INC1024 & 1.13  & 18.88  & 14.99 & 6.01 & 0.055255 & 0.944745\\ \hline
-INC2048 & 29.93 & 52.92 & 8.53 & 6.59 & 1.001873 & -0.00187\\ \hline
-INC4096 & 39.57 & 98.00 & 12.41 & 8.19 & 0.939756 & 0.060244\\ \hline
+\begin{tabular}{|l|l|l|l|l|l|l|l|l|} \hline
+INC        & Total Range & Mean 1 & Mean 2 & Std 1 & Std 2 & Scale 1 & Scale 2 & Figures \\ \hline
+INC1      & 6 & 1.66  & 4.73 & 1.03 & 0.54 & 0.10 & 0.90  & Fig.~\ref{fig:inc1_r1_hist_v5}\\ \hline
+INC2      & 9 & 2.33  & 7.07 & 2.38 & 0.54 & 0.16 & 0.84  & Fig.~\ref{fig:inc2_r1_hist_v5} \\ \hline
+INC4      & 9 & 2.17  & 7.89 & 0.99 & 0.86 & 0.21 & 0.79  & Fig.~\ref{fig:inc4_r1_hist_v5} \\ \hline
+INC8      & 9 & 3.51  & 9.15 & 1.10 & 1.08 & 0.39 & 0.61  & Fig.~\ref{fig:inc8_r1_hist_v5}\\ \hline
+INC16    & 10 & 4.04  & 9.53 & 0.98 & 1.28 & 0.83 & 0.17  & Fig.~\ref{fig:inc16_r1_hist_v5}\\ \hline
+INC32    & 22 & 2.76  & 8.26 & 1.28 & 1.12 & 0.73 & 0.27  & Fig.~\ref{fig:inc32_r1_hist_v5}\\ \hline
+INC64    & 14 & 6.36  & 12.02 & 1.64 & 1.62 &  0.35 & 0.65  & Fig.~\ref{fig:inc64_r1_hist_v5}\\ \hline
+INC128  & 16 & 4.21  & 8.64 & 1.82 & 3.40 & 0.41 & 0.59  & Fig.~\ref{fig:inc128_r1_hist_v5}\\ \hline
+INC256  & 38 & 6.08  & 14.61  & 3.59 & 3.75 & 0.12 & 0.88  & Fig.~\ref{fig:inc256_r1_hist_v5}\\ \hline
+INC512  & 20 & 1.60  & 12.43   & 11.51 & 4.47 & 0.08 & 0.92  & Fig.~\ref{fig:inc512_r1_hist_v5}\\ \hline
+INC1024 & 40 & 1.13  & 18.88  & 14.99 & 6.01 & 0.06 & 0.94 & Fig.~\ref{fig:inc1024_r1_hist_v5}\\ \hline
+INC2048 & 65 & 29.93 & 52.92 & 8.53 & 6.59 & 1.00 & -0.00 & Fig.~\ref{fig:inc2048_r1_hist_v5}\\ \hline
+INC4096 & 65 & 39.57 & 98.00 & 12.41 & 8.19 & 0.94 & 0.06  & Fig.~\ref{fig:inc4096_r1_hist_v5}\\ \hline
 \end{tabular}

### 2017-10-12 mlppi-ksii17/review/review_response.tex
Penultimate version for MLPPI

@@
 \clearpage
-\section*{Editor-in-Chief}\label{sec:ed1}
+\section*{Editor}\label{sec:ed1}
 
@@
 
-In this response letter we address each of the reviewers' comments. 
-We also incorporate into the revised paper our response to each comment.
+In this response letter we address each of the reviewers' comments with care. 
+We have incorporated into the revision our response to each comment.
 
@@
 1. Revise your manuscript "by using the MS Word."
+}
 
+Yes, we used the MS word program for revising our article.
+
+\rev{
 2. Please highlight the changes to your manuscript within the document by using bold and colored text "in BLUE" or by using the track changes mode in MS Word.
+}
 
+%OK, anything new or revised compared to the original submission is colored in {\color{blue}blue} and marked in \textbf{bold} face in the revised.
+In addition, corrections made are colored in {\color{red}red} for distinction from the new (or revised) contents.
+OK, the contents in response to the reviewer's comments are colored in {\color{blue}\textbf{blue}} with a bold face.
+In addition, corrections made are colored in {\color{red}red} for distinction from the new (or revised) contents.
+
+\rev{
 3.
@@
 In order to expedite the processing of the revised manuscript, please be as specific as possible in your response to the reviewer(s)' comments.
+}
 
+This written response letter satisfies your requirement. It includes the respective response to each of the two reviewer's comments.
+
+\rev{
 4. The authors’ response should be included in "the first page of the revised manuscript."
+}
 
+Our submission includes this letter in the front page of the revised article.
+
+\rev{
 Once the revised manuscript is prepared, you can upload it and submit it through your Author Center.
@@
 
-Yes, we used the MS word program for revising our article.
-As requested, anything new or revised compared to the original submission is colored in {\color{blue}blue} and marked in \textbf{bold} face in the revised.
-In addition, corrections made are colored in {\color{red}red} for distinction.
-As required, we provide this written response letter to answer each reviewer's comments.
-Finally, we include the letter in the first page of the revised article.
+\rev{
+[IMPORTANT]:  Your original files are available to you when you upload your revised manuscript.  Please delete any redundant files before completing the submission.
+}
+
+We deleted any redundant files during the uploading process.
 
 \rev{
-[IMPORTANT]:  Your original files are available to you when you upload your revised manuscript.  Please delete any redundant files before completing the submission.
+Because we are trying to facilitate timely publication of manuscripts submitted to the KSII Transactions on Internet and Information Systems, your revised manuscript should be uploaded "by 15-Oct-2017."  If it is not possible for you to submit your revision in the deadline, we may have to consider your paper as a new submission.
+}
 
-Because we are trying to facilitate timely publication of manuscripts submitted to the KSII Transactions on Internet and Information Systems, your revised manuscript should be uploaded "by 15-Oct-2017."  If it is not possible for you to submit your revision in the deadline, we may have to consider your paper as a new submission.
+Today ({\bf\underline{October 13th, 2017 in KST}}), prior to the specified deadline of October 15, 2017, 
+the revised manuscript has just been uploaded for further review.
 
+\rev{
 Once again, thank you for submitting your manuscript to the KSII Transactions on Internet and Information Systems and I look forward to receiving your revision.
@@
 }
-
-We deleted any redundant files during the uploading process. Today ({\bf\underline{October 8th, 2017 in KST}}), the revised manuscript has been uploaded for further review, 
-and thus, there is no issue with the specified deadline of October 15, 2017.
 
@@
 
-We totally agree with you. The old Fig. 1 is not at a real figure. We feel that it had better be put into a table for further reference, which is now {\color{blue}\underline{Tab. 1}}.
+We totally agree with you. The old Fig. 1 is not at a real figure. We feel that it had better be put into a table for further reference, which is now {\color{blue}\underline{Tab. 1}} on page~3.
 
@@
 1}} on page~7, with a specified description of each phase.
-In addition, the figure now includes (1) the query workload text exemplified in the paper (on the top) and (2) the output solution
+In addition, the figure shows (1) the input query workload text, which is used throughout the paper (on the top), and (2) the output solution
 produced by the wizard (on the bottom), for better clarification.
@@
 }
+
 Thanks for the comments. To address your concern, we put in tabular form the table with its caption on page~2 (now, {\color{blue}Tab.~2}). 
@@
 }
+
 We took out some contributions mentioned in Section 2 (now Section 3) and then incorporated them into the contribution paragraph (on page~5 
@@
 For more details, please refer to the Solution paragraph in the introduction and the third and subsequent paragraphs in Section~5.1.
-Please see the paragraph of `Workload Generation \& Deployment' in Section~5.1: more specifically, the third---fifth
-paragraphs in that section, where we address the use of synthetic queries in general from to represent a realistic scenario and have a mention of reproducibility of our work.
+%Please see the paragraph of `Workload Generation \& Deployment' in Section~5.1: more specifically, the third to fifth
+%paragraphs in that section, where we address the use of synthetic queries in general from to represent a realistic scenario and have a mention of reproducibility of our work.
+Please see Section~5.1 (more specifically, the third to fifth paragraphs in that section), where we address the use of synthetic queries in general from to represent a realistic scenario.
+We do have a mention of reproducibility of our work (as described in the the second-to-last paragraph).
 
@@
 
-Your confusion comes from Figure 5 in the original submission.
+%Your confusion comes from Figure 5 in the original submission.
 The proposed wizard obviously outperforms the compared approaches: no partitioning (NO PPI) and a human expert (EXP).
@@
 Again, our query workloads are from the real benchmark, SSB, and our algorithm is still applicable to any other type of analytical workload, since the search space is driven by the query predicates extracted from the workload.
+After reading your comment we realized that {\color{blue}Fig.~5}, which is now removed, was confusing.
+We replaced it by a comparison table (now {\color{blue}\underline{Tab.~9}} on page 19) that shows that the proposed wizard outperforms the compared approaches.
+
+\newpage
 
@@
 Also note that any complex workload, giving many different partitioning options, can benefit from the automated wizard to find almost an optimal solution.
-For more details, please refer to the first, third, and fourth bullets in the Contribution paragraphs in the Introduction.
+For more details, please refer to the first, third, and fourth bullets in the Contribution paragraphs in the Introduction section.
 
@@
 
-OK, we now have some mentions of the above two works compared to our work. Also, we added these two in the references in the revision.
+OK, we now have some mentions of the above two works compared to our work. Of course, we have added these two in the references ([27, 28]) in the revision.
 
@@
 The paper could be published after addressing following comments:
+}
+
+Yes, we addressed each of your comments as follows.
+
+\rev{
 1. Abstract could be better write, in this reviewer point of view, the current abstract is not comprehensive.
@@
 
-OK, we rewrote the abstract.
+OK. We totally rewrote the abstract to reflect this comment.
 
@@
 }
+
 A great suggestion.
-We have relocated the related work section right after the introduction in the revised manuscript.
-Note also that there are some new references mentioned in the last paragraph of the related work.
+We have relocated the related work section right after the introduction in the revision.
+Note that the related work section is extended with some new references that are added to respond to another reviewer's comment.
 
@@
 
-Yes, we do now provide Tab. 9 for better clarity on the comparison.
+Yes, we now provide Tab. 9 for better clarity on the comparison.
 
@@
 
-We shortened the Conclusion section. Please check page~20.
+OK, we shortened the Conclusion section, which now sounds more concise.
 
@@
 }
-We've added some references in the contribution paragraph in the introduction on page 5.
+
+We've added some references in the Contribution paragraph in the introduction on page 5.
 

### 2017-10-12 mlppi-ksii17/review/cover-letter.tex
Penultimate version for MLPPI

--- 2017-03-27 mlppi-comsis17/cover-letter.tex
@@
 \begin{flushright}
-\hfill Supercomputing R\&D Center, 
+\hfill School of Computer Science and Engineering
 \end{flushright}
@@
 \begin{flushright}
-\hfill Korea Institute of Science and Technology Information (KISTI), 
+\hfill Kyungpook National University 
 \end{flushright}
@@
 \begin{flushright}
-\hfill Republic of Korea, 34141, 
+\hfill Daegu, 41566 Republic of Korea, 
 \end{flushright}
@@
 \begin{flushright}
-\hfill Tel: +82-42-869-0725
+\hfill Tel: +82-53-950-6372
 \end{flushright}
@@
 \begin{flushright}
-\hfill Email: yksuh@kisti.re.kr
+\hfill Email: yksuh@knu.ac.kr
 \end{flushright} 
@@
 
-\opening{Dear Prof. Mirjana Ivanovi\'c (Editor-in-Chief):}
+\opening{Dear Dr. Mohammad Shojafar (Editor)}
 
-\noindent An earlier version of this submission was published in Proceedings of the 21st ACM \hbox{International} Conference on Information and Knowledge Management (CIKM). 
-This version that I now submit to Computer Science and Information Systems (ComSIS) includes the following substantial extensions.
-
-\begin{itemize}
-
-\item Expanded algorithms: we substantially expand the description of the pre-processing, initial and optimized phase algorithms shortened in the earlier version.
-Furthermore, this submission presents three additional algorithms as follows:
-
-\begin{itemize}
-
-\item Splitting overlapping ranges in Algorithm~2 on page~8,
-\item Checking the allowed field count limit on the split ranges on page~8,
-\item Computing scan cost in Algorithm~3 on page~10, and
-\item Computing query cost in Algorithm~5 on page~13.
-
-\end{itemize}
-
-\item Complexity analysis: we conduct the asymptotic running time analysis on each phase in Section~3 on pages~14-15.
-
-\item Experiments: we provide additional runtime statistics measured from our experiments, as exhibited in Table~1 on page~16. 
-In addition, we compare the recommendation quality of our tool with that of a human expert, as shown in Figure~5 on page~18.
-
-\item More thorough literature review: Section~5 on page~18 discusses in greater detail the similarities and differences between existing tools 
-and ours.
-
-\end{itemize}
+\noindent We are submitting our revised manuscript (ID: TIIS-IS-2017-Jul-0844), entitled ``MLPPI Wizard: An Automated Multi-level Partitioning Tool on Analytical Workloads,'' to KSII Transactions on Internet and Information Systems (TIIS).
+For this revision we carefully addressed each reviewer's comments.
+More details are found in the response letter included in this submission.
+Besides, our submission includes this cover letter and the revised PDF of our paper.
+Thank you for your support.
 

### 2017-10-14 mlppi-ksii17/review/review_response.tex
Submitted Revision

@@
 \section*{Overview}\label{sec:overview}
-We appreciate these quite thorough reviews. We appreciate the feedback and suggestions.
+We appreciate these good reviews. We appreciate the feedback and suggestions.
 
@@
 In this response letter we address each of the reviewers' comments with care. 
-We have incorporated into the revision our response to each comment.
+We have incorporated into the revision our response to each of them.
 
@@
 
-Yes, we used the MS word program for revising our article.
+Yes, we used the MS Word program for revising our article.
 
@@
 
-Our submission includes this letter in the front page of the revised article.
+Yes, we include in our submission this letter that appears on the front page of the revised article.
 
@@
 
-Today ({\bf\underline{October 13th, 2017 in KST}}), prior to the specified deadline of October 15, 2017, 
-the revised manuscript has just been uploaded for further review.
+Today ({\bf\underline{October 14th, 2017 in KST}}), prior to the specified deadline of October 15, 2017, 
+we're submitting this revision for further review. So no violation on the deadline.
 
@@
 
-We took out some contributions mentioned in Section 2 (now Section 3) and then incorporated them into the contribution paragraph (on page~5 
+We took out some contributions mentioned in (the previous) Section 2 (now Section 3) and then incorporated them into the contribution paragraph (on page~5 
 in the revision), which is now substantially extended.
@@
 Please see Section~5.1 (more specifically, the third to fifth paragraphs in that section), where we address the use of synthetic queries in general from to represent a realistic scenario.
-We do have a mention of reproducibility of our work (as described in the the second-to-last paragraph).
+We do have a mention of reproducibility of our work (as described in the second-to-last paragraph).
 
@@
 
-The application type of the proposed wizard is a \underline{physical database design tool} (fundamentally different from several existing works [4--6].
+The application type of the proposed wizard is a \underline{physical database design tool} (fundamentally different from several existing works {\color{red}[4--6]}.
 Also note that any complex workload, giving many different partitioning options, can benefit from the automated wizard to find almost an optimal solution.
-For more details, please refer to the first, third, and fourth bullets in the Contribution paragraphs in the Introduction section.
+For more details, please refer to the first, third, and fifth bullets in the Contribution paragraphs in Section~1 on page~5.
 
@@
 
-OK, we now have some mentions of the above two works compared to our work. Of course, we have added these two in the references ([27, 28]) in the revision.
+OK, we now have some mentions of the above two works compared to our work.
+Accordingly, we have added these two articles in the References ({\color{red}[27, 28]}).
+Please also note that we have moved the Related Work section (now Section~2) (on page~5) right after Section 1, to respond to a concern from another reviewer.
 
@@
 A great suggestion.
-We have relocated the related work section right after the introduction in the revision.
-Note that the related work section is extended with some new references that are added to respond to another reviewer's comment.
+We have moved the Related Work section (now Section~2) upfront (on page 5), right after Section 1.
+Please also note that Section~2 is expanded with some new references ({\color{red}[27, 28]}) to address the concerns from another reviewer.
 
@@
 
-Yes, we now provide Tab. 9 for better clarity on the comparison.
+Yes, we now provide {\color{blue}Tab.~9} for better clarity on the comparison.
 
@@
 
-We've added some references in the Contribution paragraph in the introduction on page 5.
+We've added some references in the Contribution paragraph in Section~1 on page 5.
 

### 2018-02-22 pt_characterization/real_data/paper.tex
new data

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
 \usepackage{auto-pst-pdf}
\usepackage{graphicx}
\usepackage{footnote}
%\usepackage{arydshln}
\usepackage{ifpdf}
\ifpdf
  \usepackage{epstopdf}
\fi
\usepackage{multirow}
\usepackage{epsfig}
\usepackage{float}
\usepackage{url}
\usepackage{color}
\usepackage{subfigure}

%\newcommand\solidrule[1][1cm]{\rule[0.5ex]{#1}{.4pt}}
%\newcommand\dashedrule{\mbox{%
%  \solidrule[2mm]\hspace{2mm}\solidrule[2mm]\hspace{2mm}\solidrule[2mm]}}

%\usepackage{hyperref}

\begin{document}
\title{Characterizing Execution Times on Realistic Programs}

\author{
Database \& Big Data Systems Laboratory,\\
School of Computer Science and Engineering, \\
Kyungpook National University,\\
Young-Kyoon Suh\\
}
\maketitle

\section{Description}
This document characterizes execution times 
measured on several real-world programs with different input sizes. 
To achieve this characterization, we discuss 
various histograms of execution times, measured in program time (PT), of the programs throughout this document. 
In this work we wish to achieve several goals as follows. 
The first goal is to unravel any structure behind the histograms and present insights into how such structure is formed. 
Another goal is to build a statistical distribution (or model) fitting in the histograms. 
From that distribution, we may reach predicting a concrete execution time considering system noise via 
the model on an arbitrary algorithm with a given input on a real execution environment. 
As a note, the execution times were measured along with the EMPv5~\cite{EMP} protocol. 
%In the protocol, we use taskstats C struct to get measures of a captured process. 
%The taskstat's data is delivered via a netlink socket from the kernel space. 
%The receive buffer for the socket is not robust for many observed processes~\cite{Metrology}. 
%Fortunately, there is an average of 95 processes per iteration of a run, 
%which turns out to be fine with the struct. 
%For a much more number of processes, 
%the use of  {\tt /proc/}[pid]{\tt{/stat}} is preferred, 
%as (i) there are equivalent measures available in the {\tt /proc} filesystem, 
%and (ii) there's little constraint on the use as opposed to taskstats.

The following section shows histograms for runs on 
different real-world programs with varying input sizes.

\section{Histograms of the Execution Times on Real-World Programs~\label{sec:real-world}}

In this section we present histogram data 
for two main programs: {\it insertion sort} and {\it matrix multiplication}. 
For the runs of these programs, we varied their input sizes by 2{\small $\times$} 
and measured execution times of the programs over each input size.

\subsection{Insertion sort~\label{sec:sort}} 
This section shows a series of histograms of an insertion sort program. 
The program repeatedly runs 40 times for a given input size. 
The input size for the program 
varies from 100,000 to 3,200,000 integer elements, which are randomly generated. 
Note that each sort program over a specific input size is termed SORT{\it x}: 
for instance, SORT100 indicates the insertion sort program over 100K elements.

Figures~\ref{fig:sort1} and ~\ref{fig:sort2} exhibit 
histograms of the execution times measured on the same insertion sort program as 
the input size grows from 100K to 800K elements by the steps of 2x. 
Note that we used one standard deviation in Figure~\ref{fig:sort400_dist} as
a couple of outliers, which were not eliminated by the original protocol, 
resulted in disturbing the rendering of a clean distribution.

\pagebreak

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on SORT100]{
		\includegraphics[scale=0.43]{sort100_dist.eps}
		\label{fig:sort100_dist}
	}
	\subfigure[PT frequency on SORT200]{
		\includegraphics[scale=0.43]{sort200_dist.eps}
		\label{fig:sort200_dist}
	}
	\subfigure[PT frequency on SORT400]{
		\includegraphics[scale=0.43]{sort400_dist.eps}
		\label{fig:sort400_dist}
	}
	\subfigure[PT frequency on SORT800]{
		\includegraphics[scale=0.43]{sort800_dist.eps}
		\label{fig:sort800_dist}
	}
	\caption{PT Histograms of SORT100 ... SORT800~\label{fig:sort1}}
\end{figure}

\pagebreak

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on SORT1600]{
		\includegraphics[scale=0.43]{sort1600_dist.eps}
		\label{fig:sort1600_dist}
	}
	\subfigure[PT frequency on SORT3200]{
		\includegraphics[scale=0.43]{sort3200_dist.eps}
		\label{fig:sort3200_dist}
	}
	\caption{PT Histograms of SORT1600 and SORT3200~\label{fig:sort2}}
\end{figure}

\subsection{Matrix Multiplication~\label{sec:mm}}

This section shows a series of histograms of 
the execution times of an matrix multiplication program. 
We used the same sample size for each input size of this program: i.e., 40 iterations. 
For simplicity, we used two square matrices for performing their multiplication in the program.  
We also varied the input sizes of each of the two matrices: 
from 1K$\times$1K to 8K$\times$8K integer elements that are also randomly generated. 
Note that each matrix multiplication program for a specific size is called MAT{\it xyyyy}, 
where {\it x} indicates which major, specifically {\em column} vs. {\em row}, is used, and {\it yyyy}, how large a given matrix is. 
For instance, MATC1000 represents a matrix multiplication program 
in column major over two square matrices having 1,000 integer (random) 
elements in a row (and a column).

\subsubsection{Column Major}

Figure~\ref{fig:matc} shows a series of 
histograms of the execution times measured on 
the same matrix multiplication program in column major 
as the input sizes grows from 1000x1000 to 8000x8000 by the steps of 2{\small $\times$}..

\pagebreak

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on MATC1000]{
		\includegraphics[scale=0.43]{matc1000_dist.eps}
		\label{fig:matc100_dist}
	}
	\subfigure[PT frequency on MATC2000]{
		\includegraphics[scale=0.43]{matc2000_dist.eps}
		\label{fig:matc200_dist}
	}
	\subfigure[PT frequency on MATC4000]{
		\includegraphics[scale=0.43]{matc4000_dist.eps}
		\label{fig:matc400_dist}
	}
	\subfigure[PT frequency on MATC8000]{
		\includegraphics[scale=0.43]{matc8000_dist.eps}
		\label{fig:matc800_dist}
	}
	\caption{PT Histograms of MATC1000 ... MATC8000~\label{fig:matc}}
\end{figure}

\subsubsection{Row Major}

Figure~\ref{fig:matc} shows a series of 
histograms of the execution times measured on 
the same matrix multiplication program in row major 
as the input sizes grows from 1K$\times$1K to 8K$\times$8K by the steps of 2{\small $\times$}.

\pagebreak

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on MATR1000]{
		\includegraphics[scale=0.43]{matr1000_dist.eps}
		\label{fig:matr100_dist}
	}
	\subfigure[PT frequency on MATR2000]{
		\includegraphics[scale=0.43]{matr2000_dist.eps}
		\label{fig:matr200_dist}
	}
	\subfigure[PT frequency on MATR4000]{
		\includegraphics[scale=0.43]{matr4000_dist.eps}
		\label{fig:matr400_dist}
	}
	\subfigure[PT frequency on MATR8000]{
		\includegraphics[scale=0.43]{matr8000_dist.eps}
		\label{fig:matr800_dist}
	}
	\caption{PT Histograms of MATR1000 ... MATR8000~\label{fig:matr}}
\end{figure}

\bibliographystyle{abbrv}
\newcommand{\etalchar}[1]{$^{#1}$}
\begin{thebibliography}{99}
\vspace{0.1em}
%\bibitem
%{Avnur:2000:ECA:342009.335420}
%{Sedona}
%Young-Kyoon Suh, ``SEDONA: A Novel Protocol for Identifying Infrequent, Long-running Daemons on a Linux System'', in {\em IEICE Transactions on Information Systems}, Vol. 100D, No. 9, pp.~??--??, 2017.
%\vspace{0.1em}
\bibitem
%{Avnur:2000:ECA:342009.335420}
{EMP}
Young-Kyoon Suh, Richard T. Snodgrass, John Kececioglu, Peter J. Downey, Rob S. Maier, and Cheng Yi, ``EMP: Execution Time Measurement Protocol for Compute-Bound Programs'', in {\em Software: Practice and Experience}, 47(4):559--597, 2017.
\bibitem
%{Avnur:2000:ECA:342009.335420}
{Metrology}
Sabah Currim,Richard T. Snodgrass, Young-Kyoon Suh, and Rui Zhang, ``DBMS Metrology: Measuring Query Time'', in {\em ACM Transactions on Database Systems}, 42(1):3:1--42(+8), 2017.
\end{thebibliography}

\end{document}

### 2018-02-22 pt_characterization/real_data/hist.r

@@ -1,0 +1,163 @@
+### SORT10
+numSamples <- 40
+x = read.csv(file="SORT100.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort100_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 20
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on SORT100 with 40 samples', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+x = read.csv(file="SORT200.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort200_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
[…]

### 2018-02-22 pt_characterization/real_data/matc_hist.r

@@ -1,0 +1,109 @@
+### MATC1000
+numSamples <- 40
+x = read.csv(file="MATC1000.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("matc1000_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 15
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on MATC1000 with 40 samples', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+x = read.csv(file="MATC2000.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("matc2000_dist.eps")
+binsize=1
+x <- subset(x, x$ITERNUM >= 6)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
[…]

### 2018-02-22 pt_characterization/real_data/matr_hist.r

@@ -1,0 +1,109 @@
+### MATR1000
+numSamples <- 40
+x = read.csv(file="MATR1000.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("matr1000_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+#ymax <- ceiling(ymax/100)*100
+ymax <- 20
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on MATR1000 with 40 samples', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+x = read.csv(file="MATR2000.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("matr2000_dist.eps")
+binsize=1
+x <- subset(x, x$ITERNUM >= 6)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
[…]

### 2018-03-08 pt_characterization/real_data/paper.tex
Characterizing draft

@@
 
-\pagebreak
-
 \begin{figure}[h]
 	\centering
+	\subfigure[PT frequency on SORT25]{
+		\includegraphics[scale=0.43]{sort25_dist.eps}
+		\label{fig:sort25_dist}
+	}
+	\subfigure[PT frequency on SORT50]{
+		\includegraphics[scale=0.43]{sort50_dist.eps}
+		\label{fig:sort50_dist}
+	}
 	\subfigure[PT frequency on SORT100]{
@@
 	}
+	\caption{PT Histograms of SORT25 ... SORT200~\label{fig:sort1}}
+\end{figure}
+
+\clearpage
+\pagebreak
+
+\begin{figure}[h]
+	\centering
 	\subfigure[PT frequency on SORT400]{
@@
 	}
-	\caption{PT Histograms of SORT100 ... SORT800~\label{fig:sort1}}
+	\subfigure[PT frequency on SORT1160]{
+		\includegraphics[scale=0.43]{sort1160_dist.eps}
+		\label{fig:sort1160_dist}
+	}
+	\caption{PT Histograms of SORT400 ... SORT1160~\label{fig:sort2}}
 \end{figure}
 
+\clearpage
 \pagebreak
@@
 	\centering
-	\subfigure[PT frequency on SORT1600]{
-		\includegraphics[scale=0.43]{sort1600_dist.eps}
-		\label{fig:sort1600_dist}
+	\subfigure[PT frequency on SORT2320 (Will be updated with the rest.)]{
+		\includegraphics[scale=0.43]{sort2320_dist.eps}
+		\label{fig:sort2320_dist}
 	}
@@
 	}
-	\caption{PT Histograms of SORT1600 and SORT3200~\label{fig:sort2}}
+	\\
+	\subfigure[PT frequency on SORT4640]{
+		%\includegraphics[scale=0.43]{sort4640_dist.eps}
+		\framebox(200,200){Will be available soon.}
+		\label{fig:sort4640_dist}
+	}
+	\subfigure[PT frequency on SORT6400]{
+		\includegraphics[scale=0.43]{sort6400_dist.eps}
+		\label{fig:sort6400_dist}
+	}
+	\caption{PT Histograms of SORT2320 ... SORT6400~\label{fig:sort3}}
 \end{figure}
@@
 
+\pagebreak
+
 \subsubsection{Column Major}
@@
 the same matrix multiplication program in column major 
-as the input sizes grows from 1000x1000 to 8000x8000 by the steps of 2{\small $\times$}..
-
-\pagebreak
+as the input sizes grows.
 
@@
 	\centering
-	\subfigure[PT frequency on MATC1000]{
-		\includegraphics[scale=0.43]{matc1000_dist.eps}
-		\label{fig:matc100_dist}
+	\subfigure[PT frequency on MATC250]{
+		\includegraphics[scale=0.43]{matc250_dist.eps}
+		\label{fig:matc250_dist}
 	}
-	\subfigure[PT frequency on MATC2000]{
-		\includegraphics[scale=0.43]{matc2000_dist.eps}
-		\label{fig:matc200_dist}
+	\subfigure[PT frequency on MATC500]{
+		\includegraphics[scale=0.43]{matc500_dist.eps}
+		\label{fig:matc500_dist}
 	}
-	\subfigure[PT frequency on MATC4000]{
-		\includegraphics[scale=0.43]{matc4000_dist.eps}
-		\label{fig:matc400_dist}
+	\subfigure[PT frequency on MATC1024]{
+		\includegraphics[scale=0.43]{matc1024_dist2.eps}
+		\label{fig:matc1024_dist}
 	}
-	\subfigure[PT frequency on MATC8000]{
-		\includegraphics[scale=0.43]{matc8000_dist.eps}
-		\label{fig:matc800_dist}
+	\subfigure[PT frequency on MATC1440]{
+		\includegraphics[scale=0.43]{matc1440_dist.eps}
+		\label{fig:matc1440_dist}
 	}
-	\caption{PT Histograms of MATC1000 ... MATC8000~\label{fig:matc}}
+	\caption{PT Histograms of MATC250 ... MATC1440~\label{fig:matc}}
 \end{figure}
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on MATC2048]{
+		\includegraphics[scale=0.43]{matc2048_dist2.eps}
+		\label{fig:matc2048_dist}
+	}
+	\subfigure[PT frequency on MATC2880]{
+		\includegraphics[scale=0.43]{matc2880_dist.eps}
+		\label{fig:matc2880_dist}
+	}	
+	\subfigure[PT frequency on MATC4096]{
+		\includegraphics[scale=0.43]{matc4096_dist2.eps}
+		\label{fig:matc4096_dist}
+	}	
+	\subfigure[PT frequency on MATC5760]{
+		\includegraphics[scale=0.43]{matc5760_dist.eps}
+		\label{fig:matc5760_dist}
+	}	
+	\caption{PT Histograms of MATC2048 ... MATC5760~\label{fig:matc2}}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on MATC8192]{
+		\includegraphics[scale=0.43]{matc8192_dist2.eps}
+		\label{fig:matc8192_dist}
+	}
+	\subfigure[PT frequency on MATC11520]{
+		\includegraphics[scale=0.43]{matc11520_dist.eps}
+		\label{fig:matc11520_dist}
+	}
+	\subfigure[PT frequency on MATC16000]{
+		\includegraphics[scale=0.43]{matc16000_dist.eps}
+		\label{fig:matc16000_dist}
+	}
+	\subfigure[PT frequency on MATC23040]{
+		\includegraphics[scale=0.43]{matc23040_dist.eps}
+		\label{fig:matc23040_dist}
+	}
+	\caption{PT Histograms of MATC8192 ... MATC23040~\label{fig:matc3}}
+\end{figure}
+
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATC1000]{
+%		\includegraphics[scale=0.43]{matc1000_dist.eps}
+%		\label{fig:matc100_dist}
+%	}
+%	\subfigure[PT frequency on MATC2000]{
+%		\includegraphics[scale=0.43]{matc2000_dist.eps}
+%		\label{fig:matc200_dist}
+%	}
+%	\subfigure[PT frequency on MATC4000]{
+%		\includegraphics[scale=0.43]{matc4000_dist.eps}
+%		\label{fig:matc400_dist}
+%	}
+%	\subfigure[PT frequency on MATC8000]{
+%		\includegraphics[scale=0.43]{matc8000_dist.eps}
+%		\label{fig:matc800_dist}
+%	}
+%	\caption{PT Histograms of MATC1000 ... MATC8000~\label{fig:matc}}
+%\end{figure}
+
+\clearpage
+\pagebreak
 
@@
 histograms of the execution times measured on 
-the same matrix multiplication program in row major 
-as the input sizes grows from 1K$\times$1K to 8K$\times$8K by the steps of 2{\small $\times$}.
-
-\pagebreak
+the same matrix multiplication program in row major as the input sizes grows.
 
@@
 	\centering
-	\subfigure[PT frequency on MATR1000]{
-		\includegraphics[scale=0.43]{matr1000_dist.eps}
-		\label{fig:matr100_dist}
+	\subfigure[PT frequency on MATR250]{
+		\includegraphics[scale=0.43]{matr250_dist.eps}
+		\label{fig:matr250_dist}
 	}
-	\subfigure[PT frequency on MATR2000]{
-		\includegraphics[scale=0.43]{matr2000_dist.eps}
-		\label{fig:matr200_dist}
+	\subfigure[PT frequency on MATR500]{
+		\includegraphics[scale=0.43]{matr500_dist.eps}
+		\label{fig:matr500_dist}
 	}
-	\subfigure[PT frequency on MATR4000]{
-		\includegraphics[scale=0.43]{matr4000_dist.eps}
-		\label{fig:matr400_dist}
+	\subfigure[PT frequency on MATR1024]{
+		\includegraphics[scale=0.43]{matr1024_dist2.eps}
+		\label{fig:matr1024_dist}
 	}
-	\subfigure[PT frequency on MATR8000]{
-		\includegraphics[scale=0.43]{matr8000_dist.eps}
-		\label{fig:matr800_dist}
+	\subfigure[PT frequency on MATR1440]{
+		\includegraphics[scale=0.43]{matr1440_dist.eps}
+		\label{fig:matr1440_dist}
 	}
-	\caption{PT Histograms of MATR1000 ... MATR8000~\label{fig:matr}}
+	\caption{PT Histograms of MATR250 ... MATR1440~\label{fig:matr}}
 \end{figure}
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on MATR2048]{
+		\includegraphics[scale=0.43]{matr2048_dist2.eps}
+		\label{fig:matr2048_dist}
+	}
+	\subfigure[PT frequency on MATR2880]{
+		\includegraphics[scale=0.43]{matr2880_dist.eps}
+		\label{fig:matr2880_dist}
+	}	
+	\subfigure[PT frequency on MATR4096]{
+		\includegraphics[scale=0.43]{matr4096_dist2.eps}
+		\label{fig:matr4096_dist}
+	}	
+	\subfigure[PT frequency on MATR5760]{
+		\includegraphics[scale=0.43]{matr5760_dist.eps}
+		\label{fig:matr5760_dist}
+	}	
+	\caption{PT Histograms of MATR2048 ... MATR5760~\label{fig:matr2}}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on MATR8192]{
+		\includegraphics[scale=0.43]{matr8192_dist2.eps}
+		\label{fig:matr8192_dist}
+	}
+	\subfigure[PT frequency on MATR11520]{
+		\includegraphics[scale=0.43]{matr11520_dist.eps}
+		\label{fig:matr11520_dist}
+	}	
+	\subfigure[PT frequency on MATR16000]{
+		\includegraphics[scale=0.43]{matr16000_dist.eps}
+		\label{fig:matr16000_dist}
+	}
+	\subfigure[PT frequency on MATR23040]{
+		\includegraphics[scale=0.43]{matr23040_dist.eps}
+		\label{fig:matr23040_dist}
+	}
+	\caption{PT Histograms of MATR8192 ... MATR23040~\label{fig:matr3}}
+\end{figure}
+
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATR1000]{
+%		\includegraphics[scale=0.43]{matr1000_dist.eps}
+%		\label{fig:matr100_dist}
+%	}
+%	\subfigure[PT frequency on MATR2000]{
+%		\includegraphics[scale=0.43]{matr2000_dist.eps}
+%		\label{fig:matr200_dist}
+%	}
+%	\subfigure[PT frequency on MATR4000]{
+%		\includegraphics[scale=0.43]{matr4000_dist.eps}
+%		\label{fig:matr400_dist}
+%	}
+%	\subfigure[PT frequency on MATR8000]{
+%		\includegraphics[scale=0.43]{matr8000_dist.eps}
+%		\label{fig:matr800_dist}
+%	}
+%	\caption{PT Histograms of MATR1000 ... MATR8000~\label{fig:matr}}
+%\end{figure}
+
+\clearpage
+\pagebreak
 

### 2018-03-08 pt_characterization/experiment/char_expr/compile.sh

@@ -1,0 +1,8 @@
+#!/bin/sh
+#sudo gcc -std=c99 proc_monitor.c -o proc_monitor -g -lpthread $(pkg-config --cflags --libs libnl-3.0 libnl-genl-3.0)
+#sudo chmod u+s proc_monitor
+#gcc -o incr_work -O3 incr_work.c
+##gcc -o incr_work.sort -O3 spe_new_exp_sort.c
+gcc -o incr_work.matc -O3 new_exp_matc.c
+gcc -o incr_work.matr -O3 new_exp_matr.c
+mv incr_work.matc incr_work.matr ..

### 2018-03-08 pt_characterization/experiment/char_expr/new_exp_matc.c

@@ -1,0 +1,373 @@
+#include <stdio.h>    /* standard in and output*/
+#include <sys/socket.h> /* for socket() and socket functions*/
+#include <arpa/inet.h>  /* for sockaddr_in and inet_ntoa() */
+#include <stdlib.h>   
+#include <string.h>    
+#include <unistd.h>     /* for close() */
+#include <time.h> 
+#include <limits.h> 
+#define PRINT 0
+#define PORT_NUM 2400
+//#define COEF 1650000000 // 16B
+#define COEF 1460000000 // 16B
+#define NUM_ELE1 20800
+#define NUM_ELE64 174000
+#define NUM_ELE128 241000
+//#define MAT_SZ 10000
+#define COLUMN_MAJOR 1
+#define ROW_MAJOR 2
+//#define COL_COEF 1460000000 
+//#define COL_COEF 1455000000 
+//#define COL_COEF 1450000000 
+//#define ROW_COEF 1457000000 
+#define MAT_COEF 2000000
+
+char c;
+
+/* for sort */
+int* iarr = NULL;
+int* iarr_unsorted = NULL;
+int numCnt = 0;
+int MAX_NUM = 1000000;//1M
+/* for matrix */
+int MAT_SZ = 0;
+int** matrix = NULL;
+
+/* for sort */
+inline void reset(int numCnt){
+	memset(iarr, 0, sizeof(int)*numCnt);
+	memcpy(iarr, iarr_unsorted, sizeof(int)*numCnt);
+}
+
+inline int sort_init(int task_length){
+	/*if(task_length == 1) numCnt = NUM_ELE1;
+	else if(task_length == 64) numCnt = NUM_ELE64;
+	else if(task_length == 128) numCnt = NUM_ELE128;*/
+
+        if(task_length == 100)    numCnt = 100000; // 100k
+        else if(task_length == 200)    numCnt = 200000; // 200k
+        else if(task_length == 400)    numCnt = 400000;// 400k
+        else if(task_length == 800)    numCnt = 800000;// 800k
+
+	// numCnt should be translated from a task length
+	iarr = (int*)malloc(sizeof(int)*numCnt);
+	iarr_unsorted = (int*)malloc(sizeof(int)*numCnt);
+
+	int i;
+	time_t t;
+	/* Intializes random number generator */
+  	srand((unsigned) time(&t));
+	for (i = 0; i < numCnt; i++) {
[…]

### 2018-03-08 pt_characterization/experiment/char_expr/new_exp_matr.c

@@ -1,0 +1,373 @@
+#include <stdio.h>    /* standard in and output*/
+#include <sys/socket.h> /* for socket() and socket functions*/
+#include <arpa/inet.h>  /* for sockaddr_in and inet_ntoa() */
+#include <stdlib.h>   
+#include <string.h>    
+#include <unistd.h>     /* for close() */
+#include <time.h> 
+#include <limits.h> 
+#define PRINT 0
+#define PORT_NUM 2400
+//#define COEF 1650000000 // 16B
+#define COEF 1460000000 // 16B
+#define NUM_ELE1 20800
+#define NUM_ELE64 174000
+#define NUM_ELE128 241000
+//#define MAT_SZ 10000
+#define COLUMN_MAJOR 1
+#define ROW_MAJOR 2
+//#define COL_COEF 1460000000 
+//#define COL_COEF 1455000000 
+//#define COL_COEF 1450000000 
+//#define ROW_COEF 1457000000 
+#define MAT_COEF 2000000
+
+char c;
+
+/* for sort */
+int* iarr = NULL;
+int* iarr_unsorted = NULL;
+int numCnt = 0;
+int MAX_NUM = 1000000;//1M
+/* for matrix */
+int MAT_SZ = 0;
+int** matrix = NULL;
+
+/* for sort */
+inline void reset(int numCnt){
+	memset(iarr, 0, sizeof(int)*numCnt);
+	memcpy(iarr, iarr_unsorted, sizeof(int)*numCnt);
+}
+
+inline int sort_init(int task_length){
+	/*if(task_length == 1) numCnt = NUM_ELE1;
+	else if(task_length == 64) numCnt = NUM_ELE64;
+	else if(task_length == 128) numCnt = NUM_ELE128;*/
+
+        if(task_length == 100)    numCnt = 100000; // 100k
+        else if(task_length == 200)    numCnt = 200000; // 200k
+        else if(task_length == 400)    numCnt = 400000;// 400k
+        else if(task_length == 800)    numCnt = 800000;// 800k
+
+	// numCnt should be translated from a task length
+	iarr = (int*)malloc(sizeof(int)*numCnt);
+	iarr_unsorted = (int*)malloc(sizeof(int)*numCnt);
+
+	int i;
+	time_t t;
+	/* Intializes random number generator */
+  	srand((unsigned) time(&t));
+	for (i = 0; i < numCnt; i++) {
[…]

### 2018-03-08 pt_characterization/experiment/char_expr/spe_new_exp.c

@@ -1,0 +1,363 @@
+#include <stdio.h>    /* standard in and output*/
+#include <sys/socket.h> /* for socket() and socket functions*/
+#include <arpa/inet.h>  /* for sockaddr_in and inet_ntoa() */
+#include <stdlib.h>   
+#include <string.h>    
+#include <unistd.h>     /* for close() */
+#include <time.h> 
+#include <limits.h> 
+#define PRINT 0
+#define PORT_NUM 2400
+//#define COEF 1650000000 // 16B
+#define COEF 1460000000 // 16B
+#define NUM_ELE1 20800
+#define NUM_ELE64 174000
+#define NUM_ELE128 241000
+//#define MAT_SZ 10000
+#define COLUMN_MAJOR 1
+#define ROW_MAJOR 2
+//#define COL_COEF 1460000000 
+//#define COL_COEF 1455000000 
+//#define COL_COEF 1450000000 
+//#define ROW_COEF 1457000000 
+#define MAT_COEF 2000000
+
+char c;
+
+/* for sort */
+int* iarr = NULL;
+int* iarr_unsorted = NULL;
+int numCnt = 0;
+int MAX_NUM = 1000000;//1M
+/* for matrix */
+int MAT_SZ = 0;
+int** matrix = NULL;
+
+/* for sort */
+inline void reset(int numCnt){
+	memset(iarr, 0, sizeof(int)*numCnt);
+	memcpy(iarr, iarr_unsorted, sizeof(int)*numCnt);
+}
+
+inline int sort_init(int task_length){
+	/*if(task_length == 1) numCnt = NUM_ELE1;
+	else if(task_length == 64) numCnt = NUM_ELE64;
+	else if(task_length == 128) numCnt = NUM_ELE128;*/
+
+	if(task_length == 100)    numCnt = 100000; // 100k
+        else if(task_length == 200)    numCnt = 200000; // 200k
+        else if(task_length == 400)    numCnt = 400000;// 400k
+	else if(task_length == 800)    numCnt = 800000;// 800k
+
+	// numCnt should be translated from a task length
+	iarr = (int*)malloc(sizeof(int)*numCnt);
+	iarr_unsorted = (int*)malloc(sizeof(int)*numCnt);
+
+	int i;
+	time_t t;
+	/* Intializes random number generator */
+  	srand((unsigned) time(&t));
+	for (i = 0; i < numCnt; i++) {
[…]

### 2018-03-08 pt_characterization/experiment/char_expr/temp/spe_new_exp_matc.c

@@ -1,0 +1,363 @@
+#include <stdio.h>    /* standard in and output*/
+#include <sys/socket.h> /* for socket() and socket functions*/
+#include <arpa/inet.h>  /* for sockaddr_in and inet_ntoa() */
+#include <stdlib.h>   
+#include <string.h>    
+#include <unistd.h>     /* for close() */
+#include <time.h> 
+#include <limits.h> 
+#define PRINT 0
+#define PORT_NUM 2400
+//#define COEF 1650000000 // 16B
+#define COEF 1460000000 // 16B
+#define NUM_ELE1 20800
+#define NUM_ELE64 174000
+#define NUM_ELE128 241000
+//#define MAT_SZ 10000
+#define COLUMN_MAJOR 1
+#define ROW_MAJOR 2
+//#define COL_COEF 1460000000 
+//#define COL_COEF 1455000000 
+//#define COL_COEF 1450000000 
+//#define ROW_COEF 1457000000 
+#define MAT_COEF 2000000
+
+char c;
+
+/* for sort */
+int* iarr = NULL;
+int* iarr_unsorted = NULL;
+int numCnt = 0;
+int MAX_NUM = 1000000;//1M
+/* for matrix */
+int MAT_SZ = 0;
+int** matrix = NULL;
+
+/* for sort */
+inline void reset(int numCnt){
+	memset(iarr, 0, sizeof(int)*numCnt);
+	memcpy(iarr, iarr_unsorted, sizeof(int)*numCnt);
+}
+
+inline int sort_init(int task_length){
+	/*if(task_length == 1) numCnt = NUM_ELE1;
+	else if(task_length == 64) numCnt = NUM_ELE64;
+	else if(task_length == 128) numCnt = NUM_ELE128;*/
+
+        if(task_length == 100)    numCnt = 100000; // 100k
+        else if(task_length == 200)    numCnt = 200000; // 200k
+        else if(task_length == 400)    numCnt = 400000;// 400k
+        else if(task_length == 800)    numCnt = 800000;// 800k
+
+	// numCnt should be translated from a task length
+	iarr = (int*)malloc(sizeof(int)*numCnt);
+	iarr_unsorted = (int*)malloc(sizeof(int)*numCnt);
+
+	int i;
+	time_t t;
+	/* Intializes random number generator */
+  	srand((unsigned) time(&t));
+	for (i = 0; i < numCnt; i++) {
[…]

### 2018-03-08 pt_characterization/experiment/char_expr/temp/spe_new_exp_matr.c

@@ -1,0 +1,363 @@
+#include <stdio.h>    /* standard in and output*/
+#include <sys/socket.h> /* for socket() and socket functions*/
+#include <arpa/inet.h>  /* for sockaddr_in and inet_ntoa() */
+#include <stdlib.h>   
+#include <string.h>    
+#include <unistd.h>     /* for close() */
+#include <time.h> 
+#include <limits.h> 
+#define PRINT 0
+#define PORT_NUM 2400
+//#define COEF 1650000000 // 16B
+#define COEF 1460000000 // 16B
+#define NUM_ELE1 20800
+#define NUM_ELE64 174000
+#define NUM_ELE128 241000
+//#define MAT_SZ 10000
+#define COLUMN_MAJOR 1
+#define ROW_MAJOR 2
+//#define COL_COEF 1460000000 
+//#define COL_COEF 1455000000 
+//#define COL_COEF 1450000000 
+//#define ROW_COEF 1457000000 
+#define MAT_COEF 2000000
+
+char c;
+
+/* for sort */
+int* iarr = NULL;
+int* iarr_unsorted = NULL;
+int numCnt = 0;
+int MAX_NUM = 1000000;//1M
+/* for matrix */
+int MAT_SZ = 0;
+int** matrix = NULL;
+
+/* for sort */
+inline void reset(int numCnt){
+	memset(iarr, 0, sizeof(int)*numCnt);
+	memcpy(iarr, iarr_unsorted, sizeof(int)*numCnt);
+}
+
+inline int sort_init(int task_length){
+	/*if(task_length == 1) numCnt = NUM_ELE1;
+	else if(task_length == 64) numCnt = NUM_ELE64;
+	else if(task_length == 128) numCnt = NUM_ELE128;*/
+       
+	if(task_length == 100)    numCnt = 100000; // 100k
+        else if(task_length == 200)    numCnt = 200000; // 200k
+        else if(task_length == 400)    numCnt = 400000;// 400k
+	else if(task_length == 800)    numCnt = 800000;// 800k
+
+	// numCnt should be translated from a task length
+	iarr = (int*)malloc(sizeof(int)*numCnt);
+	iarr_unsorted = (int*)malloc(sizeof(int)*numCnt);
+
+	int i;
+	time_t t;
+	/* Intializes random number generator */
+  	srand((unsigned) time(&t));
+	for (i = 0; i < numCnt; i++) {
[…]

### 2018-03-08 pt_characterization/experiment/char_expr/test.c

@@ -1,0 +1,106 @@
+#include <stdio.h>    /* standard in and output*/
+#include <stdlib.h>   
+#include <string.h>    
+#include <unistd.h>     /* for close() */
+#include <time.h> 
+#include <limits.h> 
+#include <sys/time.h>
+#define PRINT 0
+#define PORT_NUM 2400
+#define COEF 1460000000 // 16B
+//#define MAT_SZ 10000
+#define COLUMN_MAJOR 1
+#define ROW_MAJOR 2
+#define MAT_COEF 2000000
+
+/* for sort */
+int* iarr = NULL;
+int* iarr_unsorted = NULL;
+int numCnt = 0;
+int MAX_NUM = 1000000;//1M
+/* for matrix */
+int MAT_SZ = 0;
+int** matrix = NULL;
+
+void mat_init(int task_length);
+void mat_walker(int exp_secs, int type);
+
+int main(int argc, char *argv[]) {
+	int task_length = atoi(argv[1]);
+printf("task length: %d\n", task_length);
+	mat_init(task_length);
+
+struct timeval tim;
+gettimeofday(&tim, NULL);
+double t1=tim.tv_sec*1000+(tim.tv_usec/1000.0);
+clock_t cpu_start = clock();
+	mat_walker(task_length, COLUMN_MAJOR);
+clock_t cpu_end = clock();
+double cpu_temp = (double)(cpu_end-cpu_start) / CLOCKS_PER_SEC;
+gettimeofday(&tim, NULL);
+double t2=tim.tv_sec*1000+(tim.tv_usec/1000.0);
+fprintf(stdout, "Total CPU time on read requests: '%.4f(sec)'\n", cpu_temp);
+fprintf(stdout, "%.6lf msecs elapsed\n", t2-t1);
+
+}
+
+/* for matrix worker */
+inline void mat_init(int task_length){
+	time_t t;
+	/* Intializes random number generator */
+  	srand((unsigned) time(&t));
+	
+	if(task_length == 1024) MAT_SZ = 16000;//1G
+else if(task_length == 250) MAT_SZ = 8000;//1G
+else if(task_length == 500) MAT_SZ = 11314;//1G
+else if(task_length == 1440) MAT_SZ = 17152;//1G
[…]

### 2018-03-08 pt_characterization/real_data/hist.r

@@ -1,5 +1,179 @@
-### SORT10
+### SORT25
 numSamples <- 40
+x = read.csv(file="SORT25.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort25_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 25
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on SORT25 with 40 samples', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### SORT50
+numSamples <- 40
+x = read.csv(file="SORT50.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort50_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
[…]

### 2018-03-08 pt_characterization/real_data/matc_hist.r

@@ -1,9 +1,153 @@
-### MATC1000
+### MATC250
 numSamples <- 40
-x = read.csv(file="MATC1000.dat",head=TRUE,sep="\t")
+x = read.csv(file="MATC250.dat",head=TRUE,sep="\t")
 x <- subset(x, x$ITERNUM <= numSamples)
 setEPS()
-postscript("matc1000_dist.eps")
+postscript("matc250_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 25
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on MATC250 with 40 samples', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### MATC500
+numSamples <- 40
+x = read.csv(file="MATC500.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("matc500_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
[…]

### 2018-03-08 pt_characterization/real_data/matr_hist.r

@@ -1,9 +1,9 @@
 ### MATR1000
 numSamples <- 40
-x = read.csv(file="MATR1000.dat",head=TRUE,sep="\t")
+x = read.csv(file="MATR250.dat",head=TRUE,sep="\t")
 x <- subset(x, x$ITERNUM <= numSamples)
 setEPS()
-postscript("matr1000_dist.eps")
+postscript("matr250_dist.eps")
 binsize=1
 #x <- subset(x, x$ITERNUM != 9394)
 x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
@@ -20,12 +20,148 @@ xmax <-max(x$PRTIME)
 ymax <- max(h$counts)
 #ymax <- ceiling(ymax/100)*100
 ymax <- 20
-plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on MATR1000 with 40 samples', 
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on MATR250 with 40 samples', 
 sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
 xlab='PT (ms)', ylab=expression('Frequency'))
 #axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
+
+x = read.csv(file="MATR500.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("matr500_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
[…]

### 2018-03-08 pt_characterization/real_data/paper.tex
03/08 draft

@@
 \usepackage{color}
+\usepackage{comment}
 \usepackage{subfigure}
@@
 \subsection{Insertion sort~\label{sec:sort}} 
-This section shows a series of histograms of an insertion sort program. 
-The program repeatedly runs 40 times for a given input size. 
+This section shows a series of histograms of an insertion sorting program that 
+sorts the elements of a given array in non-decreasing order. 
+The program repeatedly runs 300 times for a given input size. 
 The input size for the program 
-varies from 100,000 to 3,200,000 integer elements, which are randomly generated. 
+varies from 100,000 to 1,160,000 integer elements, which are randomly generated. 
 Note that each sort program over a specific input size is termed SORT{\it x}: 
@@
 histograms of the execution times measured on the same insertion sort program as 
-the input size grows from 100K to 800K elements by the steps of 2x. 
+the input size grows from 100K to 1,160K elements.% by the steps of 2x. 
 Note that we used one standard deviation in Figure~\ref{fig:sort400_dist} as
@@
 
+\begin{comment}
 \clearpage
@@
 
+\end{comment}
+
 \bibliographystyle{abbrv}

### 2018-03-08 pt_characterization/real_data/hist.r

@@ -56,6 +56,151 @@ xlab='PT (ms)', ylab=expression('Frequency'))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
+### SORT100
+numSamples <- 300
+x = read.csv(file="SORT100.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort100_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 150
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on SORT100 (sample size: 300)', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### SORT200
+numSamples <- 300
+x = read.csv(file="SORT200.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort200_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
[…]

### 2018-03-21 pt_characterization/real_data/paper.tex
Draft for 03/15

--- 2018-02-22 pt_characterization/real_data/paper.tex
@@
 for two main programs: {\it insertion sort} and {\it matrix multiplication}. 
-For the runs of these programs, we varied their input sizes by 2{\small $\times$} 
-and measured execution times of the programs over each input size.
+For the runs of these programs, we varied their input sizes by 2{\small $\times$} while trying 
+some intermediate sizes increased by $\sqrt{2}$ and measured execution times of the programs over each input size.
 
 \subsection{Insertion sort~\label{sec:sort}} 
-This section shows a series of histograms of an insertion sort program. 
-The program repeatedly runs 40 times for a given input size. 
+This section shows a series of histograms of an insertion sorting program that 
+sorts the elements of a given array in non-decreasing order. 
+The program repeatedly runs 300 times for a given input size. 
 The input size for the program 
-varies from 100,000 to 3,200,000 integer elements, which are randomly generated. 
+varies from 100,000 to 1,160,000 integer elements, which are randomly generated. 
 Note that each sort program over a specific input size is termed SORT{\it x}: 
@@
 
-\pagebreak
-
 \begin{figure}[h]
 	\centering
+%	\subfigure[PT frequency on SORT25]{
+%		\includegraphics[scale=0.43]{sort25_dist.eps}
+%		\label{fig:sort25_dist}
+%	}
+%	\subfigure[PT frequency on SORT50]{
+%		\includegraphics[scale=0.43]{sort50_dist.eps}
+%		\label{fig:sort50_dist}
+%	}
 	\subfigure[PT frequency on SORT100]{
@@
 		\label{fig:sort100_dist}
+	}
+	\subfigure[PT frequency on SORT144]{
+		\includegraphics[scale=0.43]{sort144_dist.eps}
+		\label{fig:sort144_dist}
 	}
@@
 	}
+	\subfigure[PT frequency on SORT288]{
+		\includegraphics[scale=0.43]{sort288_dist.eps}
+		\label{fig:sort288_dist}
+	}
+	\caption{PT Histograms of SORT25 ... SORT288~\label{fig:sort1}}
+\end{figure}
+
+\clearpage
+\pagebreak
+
+\begin{figure}[h]
+	\centering
 	\subfigure[PT frequency on SORT400]{
@@
 		\label{fig:sort400_dist}
+	}
+	\subfigure[PT frequency on SORT580]{
+		\includegraphics[scale=0.43]{sort580_dist.eps}
+		\label{fig:sort588_dist}
 	}
@@
 	}
-	\caption{PT Histograms of SORT100 ... SORT800~\label{fig:sort1}}
+	\subfigure[PT frequency on SORT1000]{
+		\includegraphics[scale=0.43]{sort1000_dist.eps}
+		\label{fig:sort1000_dist}
+	}
+	\caption{PT Histograms of SORT400 ... SORT1000~\label{fig:sort2}}
 \end{figure}
-
-\pagebreak
 
@@
 	\centering
-	\subfigure[PT frequency on SORT1600]{
-		\includegraphics[scale=0.43]{sort1600_dist.eps}
-		\label{fig:sort1600_dist}
+	\subfigure[PT frequency on SORT1160]{
+		\includegraphics[scale=0.43]{sort1160_dist.eps}
+		\label{fig:sort1160_dist}
 	}
-	\subfigure[PT frequency on SORT3200]{
-		\includegraphics[scale=0.43]{sort3200_dist.eps}
-		\label{fig:sort3200_dist}
+	\subfigure[PT frequency on SORT2320 (not complete)]{
+		\includegraphics[scale=0.43]{sort2320_dist.eps}
+		\label{fig:sort2320_dist}
 	}
-	\caption{PT Histograms of SORT1600 and SORT3200~\label{fig:sort2}}
+	\caption{PT Histograms of SORT1160 ... SORT2320~\label{fig:sort3}}
 \end{figure}
+
+%\begin{comment}
+
+%\clearpage
+%\pagebreak
+%
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on SORT2320 (Will be updated with the rest.)]{
+%		\includegraphics[scale=0.43]{sort2320_dist.eps}
+%		\label{fig:sort2320_dist}
+%	}
+%	\subfigure[PT frequency on SORT3200]{
+%		\includegraphics[scale=0.43]{sort3200_dist.eps}
+%		\label{fig:sort3200_dist}
+%	}
+%	\\
+%	\subfigure[PT frequency on SORT4640]{
+%		%\includegraphics[scale=0.43]{sort4640_dist.eps}
+%		\framebox(200,200){Will be available soon.}
+%		\label{fig:sort4640_dist}
+%	}
+%	\subfigure[PT frequency on SORT6400]{
+%		\includegraphics[scale=0.43]{sort6400_dist.eps}
+%		\label{fig:sort6400_dist}
+%	}
+%	\caption{PT Histograms of SORT2320 ... SORT6400~\label{fig:sort3}}
+%\end{figure}
+
+\clearpage
+\pagebreak
 
@@
 the same matrix multiplication program in column major 
-as the input sizes grows from 1000x1000 to 8000x8000 by the steps of 2{\small $\times$}..
+as the input sizes grows.
 
-\pagebreak
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATC250]{
+%		\includegraphics[scale=0.43]{matc250_dist.eps}
+%		\label{fig:matc250_dist}
+%	}
+%	\subfigure[PT frequency on MATC500]{
+%		\includegraphics[scale=0.43]{matc500_dist.eps}
+%		\label{fig:matc500_dist}
+%	}
+%	\subfigure[PT frequency on MATC1024]{
+%		\includegraphics[scale=0.43]{matc1024_dist2.eps}
+%		\label{fig:matc1024_dist}
+%	}
+%	\subfigure[PT frequency on MATC1440]{
+%		\includegraphics[scale=0.43]{matc1440_dist.eps}
+%		\label{fig:matc1440_dist}
+%	}
+%	\caption{PT Histograms of MATC250 ... MATC1440~\label{fig:matc}}
+%\end{figure}
+%
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATC2048]{
+%		\includegraphics[scale=0.43]{matc2048_dist2.eps}
+%		\label{fig:matc2048_dist}
+%	}
+%	\subfigure[PT frequency on MATC2880]{
+%		\includegraphics[scale=0.43]{matc2880_dist.eps}
+%		\label{fig:matc2880_dist}
+%	}	
+%	\subfigure[PT frequency on MATC4096]{
+%		\includegraphics[scale=0.43]{matc4096_dist2.eps}
+%		\label{fig:matc4096_dist}
+%	}	
+%	\subfigure[PT frequency on MATC5760]{
+%		\includegraphics[scale=0.43]{matc5760_dist.eps}
+%		\label{fig:matc5760_dist}
+%	}	
+%	\caption{PT Histograms of MATC2048 ... MATC5760~\label{fig:matc2}}
+%\end{figure}
+%
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATC8192]{
+%		\includegraphics[scale=0.43]{matc8192_dist2.eps}
+%		\label{fig:matc8192_dist}
+%	}
+%	\subfigure[PT frequency on MATC11520]{
+%		\includegraphics[scale=0.43]{matc11520_dist.eps}
+%		\label{fig:matc11520_dist}
+%	}
+%	\subfigure[PT frequency on MATC16000]{
+%		\includegraphics[scale=0.43]{matc16000_dist.eps}
+%		\label{fig:matc16000_dist}
+%	}
+%	\subfigure[PT frequency on MATC23040]{
+%		\includegraphics[scale=0.43]{matc23040_dist.eps}
+%		\label{fig:matc23040_dist}
+%	}
+%	\caption{PT Histograms of MATC8192 ... MATC23040~\label{fig:matc3}}
+%\end{figure}
 
@@
 	\centering
-	\subfigure[PT frequency on MATC1000]{
-		\includegraphics[scale=0.43]{matc1000_dist.eps}
-		\label{fig:matc100_dist}
+	\subfigure[PT frequency on MATC1K]{
+		\includegraphics[scale=0.43]{matc1024_dist.eps}
+		\label{fig:matc1k_dist}
 	}
-	\subfigure[PT frequency on MATC2000]{
-		\includegraphics[scale=0.43]{matc2000_dist.eps}
-		\label{fig:matc200_dist}
+	\subfigure[PT frequency on MATC2K]{
+		\includegraphics[scale=0.43]{matc2048_dist.eps}
+		\label{fig:matc2k_dist}
 	}
-	\subfigure[PT frequency on MATC4000]{
-		\includegraphics[scale=0.43]{matc4000_dist.eps}
-		\label{fig:matc400_dist}
+	\subfigure[PT frequency on MATC4K]{
+		\includegraphics[scale=0.43]{matc4096_dist.eps}
+		\label{fig:matc4k_dist}
 	}
-	\subfigure[PT frequency on MATC8000]{
-		\includegraphics[scale=0.43]{matc8000_dist.eps}
-		\label{fig:matc800_dist}
+	\subfigure[PT frequency on MATC8K]{
+		\includegraphics[scale=0.43]{matc8192_dist.eps}
+		\label{fig:matc8k_dist}
 	}
-	\caption{PT Histograms of MATC1000 ... MATC8000~\label{fig:matc}}
+	\caption{PT Histograms of MATC1024 ... MATC8192~\label{fig:matc}}
 \end{figure}
+
+\clearpage
+\pagebreak
 
@@
 histograms of the execution times measured on 
-the same matrix multiplication program in row major 
-as the input sizes grows from 1K$\times$1K to 8K$\times$8K by the steps of 2{\small $\times$}.
+the same matrix multiplication program in row major as the input sizes grows.
 
-\pagebreak
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATR250]{
+%		\includegraphics[scale=0.43]{matr250_dist.eps}
+%		\label{fig:matr250_dist}
+%	}
+%	\subfigure[PT frequency on MATR500]{
+%		\includegraphics[scale=0.43]{matr500_dist.eps}
+%		\label{fig:matr500_dist}
+%	}
+%	\subfigure[PT frequency on MATR1024]{
+%		\includegraphics[scale=0.43]{matr1024_dist2.eps}
+%		\label{fig:matr1024_dist}
+%	}
+%	\subfigure[PT frequency on MATR1440]{
+%		\includegraphics[scale=0.43]{matr1440_dist.eps}
+%		\label{fig:matr1440_dist}
+%	}
+%	\caption{PT Histograms of MATR250 ... MATR1440~\label{fig:matr}}
+%\end{figure}
+
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATR2048]{
+%		\includegraphics[scale=0.43]{matr2048_dist2.eps}
+%		\label{fig:matr2048_dist}
+%	}
+%	\subfigure[PT frequency on MATR2880]{
+%		\includegraphics[scale=0.43]{matr2880_dist.eps}
+%		\label{fig:matr2880_dist}
+%	}	
+%	\subfigure[PT frequency on MATR4096]{
+%		\includegraphics[scale=0.43]{matr4096_dist2.eps}
+%		\label{fig:matr4096_dist}
+%	}	
+%	\subfigure[PT frequency on MATR5760]{
+%		\includegraphics[scale=0.43]{matr5760_dist.eps}
+%		\label{fig:matr5760_dist}
+%	}	
+%	\caption{PT Histograms of MATR2048 ... MATR5760~\label{fig:matr2}}
+%\end{figure}
+%
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATR8192]{
+%		\includegraphics[scale=0.43]{matr8192_dist2.eps}
+%		\label{fig:matr8192_dist}
+%	}
+%	\subfigure[PT frequency on MATR11520]{
+%		\includegraphics[scale=0.43]{matr11520_dist.eps}
+%		\label{fig:matr11520_dist}
+%	}	
+%	\subfigure[PT frequency on MATR16000]{
+%		\includegraphics[scale=0.43]{matr16000_dist.eps}
+%		\label{fig:matr16000_dist}
+%	}
+%	\subfigure[PT frequency on MATR23040]{
+%		\includegraphics[scale=0.43]{matr23040_dist.eps}
+%		\label{fig:matr23040_dist}
+%	}
+%	\caption{PT Histograms of MATR8192 ... MATR23040~\label{fig:matr3}}
+%\end{figure}
 
@@
 	\centering
-	\subfigure[PT frequency on MATR1000]{
-		\includegraphics[scale=0.43]{matr1000_dist.eps}
-		\label{fig:matr100_dist}
+	\subfigure[PT frequency on MATR1K]{
+		\includegraphics[scale=0.43]{matr1024_dist.eps}
+		\label{fig:matr1k_dist}
 	}
-	\subfigure[PT frequency on MATR2000]{
-		\includegraphics[scale=0.43]{matr2000_dist.eps}
-		\label{fig:matr200_dist}
+	\subfigure[PT frequency on MATR2K]{
+		\includegraphics[scale=0.43]{matr2048_dist.eps}
+		\label{fig:matr2k_dist}
 	}
-	\subfigure[PT frequency on MATR4000]{
-		\includegraphics[scale=0.43]{matr4000_dist.eps}
-		\label{fig:matr400_dist}
+	\subfigure[PT frequency on MATR4K]{
+		\includegraphics[scale=0.43]{matr4096_dist.eps}
+		\label{fig:matr4k_dist}
 	}
-	\subfigure[PT frequency on MATR8000]{
-		\includegraphics[scale=0.43]{matr8000_dist.eps}
-		\label{fig:matr800_dist}
+	\subfigure[PT frequency on MATR8K]{
+		\includegraphics[scale=0.43]{matr8192_dist.eps}
+		\label{fig:matr8k_dist}
 	}
-	\caption{PT Histograms of MATR1000 ... MATR8000~\label{fig:matr}}
+	\caption{PT Histograms of MATR1K ... MATR8K~\label{fig:matr}}
 \end{figure}
+
+\clearpage
+\pagebreak
+
+%\end{comment}
 

### 2018-03-21 pt_characterization/real_data/hist.r

@@ -85,6 +85,34 @@ xlab='PT (ms)', ylab=expression('Frequency'))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
+### SORT144
+x = read.csv(file="SORT144.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort144_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 150
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on SORT144 (sample size: 300)', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
 ### SORT200
 numSamples <- 300
 x = read.csv(file="SORT200.dat",head=TRUE,sep="\t")
@@ -114,6 +142,35 @@ xlab='PT (ms)', ylab=expression('Frequency'))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
+### SORT288
+numSamples <- 300
+x = read.csv(file="SORT288.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
[…]

### 2018-03-21 pt_characterization/real_data/matc_hist.r

@@ -57,12 +57,99 @@ axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
 ### MATC1024
-numSamples <- 40
+numSamples <- 300
 x = read.csv(file="MATC1024.dat",head=TRUE,sep="\t")
-x <- subset(x, x$ITERNUM <= numSamples)
+x <- subset(x, x$ITERNUM <= numSamples & x$ITERNUM > 15)
 setEPS()
-postscript("matc1024_dist2.eps")
+postscript("matc1024_dist.eps")
 binsize=1
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 150
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main=paste("PT frequency on MATC1024 (sample size: ",numSamples,")"), 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### MATC2048
+numSamples <- 300
+x = read.csv(file="MATC2048.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples & x$ITERNUM > 5)
+setEPS()
+postscript("matc2048_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
[…]

### 2018-03-21 pt_characterization/real_data/matr_hist.r

@@ -54,12 +54,13 @@ xlab='PT (ms)', ylab=expression('Frequency'))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
+#MATR1024
+numSamples <- 300
 x = read.csv(file="MATR1024.dat",head=TRUE,sep="\t")
-x <- subset(x, x$ITERNUM <= numSamples)
+x <- subset(x, x$ITERNUM <= numSamples & x$ITERNUM > 13)
 setEPS()
-postscript("matr1024_dist2.eps")
+postscript("matr1024_dist.eps")
 binsize=1
-#x <- subset(x, x$ITERNUM != 9394)
 x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
 x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
 x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
@@ -72,21 +73,22 @@ h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
 xmin <-min(x$PRTIME)
 xmax <-max(x$PRTIME)
 ymax <- max(h$counts)
-#ymax <- ceiling(ymax/100)*100
-ymax <- 20
-plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on MATR1024 with 40 samples', 
+ymax <- ceiling(ymax/100)*100
+ymax <- 80
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main=paste("PT frequency on MATR1024 (sample size: ",numSamples,")"), 
 sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
 xlab='PT (ms)', ylab=expression('Frequency'))
 #axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
+# MATR2048
+numSamples <- 300
 x = read.csv(file="MATR2048.dat",head=TRUE,sep="\t")
-x <- subset(x, x$ITERNUM <= numSamples)
+x <- subset(x, x$ITERNUM <= numSamples & x$ITERNUM > 15)
[…]

### 2018-03-29 pt_characterization/real_data/paper.tex
New draft

@@
 
-\subsection{Insertion sort~\label{sec:sort}} 
-This section shows a series of histograms of an insertion sorting program that 
+\subsection{One-to-One Comparison between An Insertion Sort \& A Corresponding INC Programs~\label{sec:sort}} 
+This section compares program time histograms 
+of insertion sort and INC (a nested-for-loop) programs. 
+The insertion sort program 
 sorts the elements of a given array in non-decreasing order.  
 The program repeatedly runs 300 times for a given input size. 
-The input size for the program 
-varies from 100,000 to 1,160,000 integer elements, which are randomly generated. 
+The input size for the program varies from 144K to 344K integer elements, which are randomly generated. 
 Note that each sort program over a specific input size is termed SORT{\it x}: 
 for instance, SORT100 indicates the insertion sort program over 100K elements. 
+An INC program's task length is correspondingly determined by the program time 
+of an SORT program. In Figures~\ref{fig:sort1},~\ref{fig:sort2}, and~\ref{fig:sort3} 
+we perform a match on an SORT program and its corresponding INC program.
 
-Figures~\ref{fig:sort1} and ~\ref{fig:sort2} exhibit 
-histograms of the execution times measured on the same insertion sort program as 
-the input size grows from 100K to 1,160K elements.% by the steps of 2x. 
-Note that we used one standard deviation in Figure~\ref{fig:sort400_dist} as
-a couple of outliers, which were not eliminated by the original protocol, 
-resulted in disturbing the rendering of a clean distribution.
+%
+%Figures~\ref{fig:sort1} and ~\ref{fig:sort2} exhibit 
+%histograms of the execution times measured on the same insertion sort program as 
+%the input size grows from 100K to 1,160K elements.% by the steps of 2x. 
+%Note that we used one standard deviation in Figure~\ref{fig:sort400_dist} as
+%a couple of outliers, which were not eliminated by the original protocol, 
+%resulted in disturbing the rendering of a clean distribution.
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on SORT144]{
+		\includegraphics[scale=0.43]{sort144_dist.eps}
+		\label{fig:sort144_dist}
+	}	
+	\subfigure[PT frequency on INC8.8]{
+		\includegraphics[scale=0.43]{inc8.8_dist.eps}
+		\label{fig:inc8.8_dist}
+	}
+	\subfigure[PT frequency on SORT177]{
+		\includegraphics[scale=0.43]{sort177_dist.eps}
+		\label{fig:sort177_dist}
+	}	
+	\subfigure[PT frequency on INC13]{
+		\includegraphics[scale=0.43]{inc13.0_dist.eps}
+		\label{fig:inc13_dist}
+	}
+	\caption{PT Histograms I~\label{fig:sort1}}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on SORT144]{
+		\includegraphics[scale=0.43]{sort244_dist.eps}
+		\label{fig:sort144_dist}
+	}	
+	\subfigure[PT frequency on INC25.8]{
+		\includegraphics[scale=0.43]{inc25.8_dist.eps}
+		\label{fig:inc25.8_dist}
+	}
+	\subfigure[PT frequency on SORT288]{
+		\includegraphics[scale=0.43]{sort177_dist.eps}
+		\label{fig:sort177_dist}
+	}	
+	\subfigure[PT frequency on INC35]{
+		\includegraphics[scale=0.43]{inc13.0_dist.eps}
+		\label{fig:inc35_dist}
+	}
+	\caption{PT Histogram Comparison II~\label{fig:sort2}}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on SORT344]{
+		\includegraphics[scale=0.43]{sort344_dist.eps}
+		\label{fig:sort344_dist}
+	}	
+	\subfigure[PT frequency on INC51.5]{
+		\includegraphics[scale=0.43]{inc51.5_dist.eps}
+		\label{fig:inc51.5_dist}
+	}
+	\caption{PT Histogram Comparison III~\label{fig:sort3}}
+\end{figure}
+
+\clearpage
+\pagebreak
+
+\section{Additional Insertion Sort~\label{sec:add_new}}
 
@@
 	}
-	\subfigure[PT frequency on SORT144]{
-		\includegraphics[scale=0.43]{sort144_dist.eps}
-		\label{fig:sort144_dist}
-	}
-	\subfigure[PT frequency on SORT200]{
-		\includegraphics[scale=0.43]{sort200_dist.eps}
-		\label{fig:sort200_dist}
-	}
-	\subfigure[PT frequency on SORT288]{
-		\includegraphics[scale=0.43]{sort288_dist.eps}
-		\label{fig:sort288_dist}
-	}
-	\caption{PT Histograms of SORT25 ... SORT288~\label{fig:sort1}}
-\end{figure}
-
-\clearpage
-\pagebreak
-
-\begin{figure}[h]
-	\centering
 	\subfigure[PT frequency on SORT400]{
@@
 	}
+	\caption{PT Histograms of SORT100 ... SORT580~\label{fig:n_sort1}}
+\end{figure}
+
+\clearpage
+\pagebreak
+
+\begin{figure}[h]
+	\centering
 	\subfigure[PT frequency on SORT800]{
@@
 	}
-	\caption{PT Histograms of SORT400 ... SORT1000~\label{fig:sort2}}
-\end{figure}
-
-\begin{figure}[h]
-	\centering
 	\subfigure[PT frequency on SORT1160]{
@@
 	}
-	\caption{PT Histograms of SORT1160 ... SORT2320~\label{fig:sort3}}
+	\caption{PT Histograms of SORT800 ... SORT2320~\label{fig:n_sort2}}
 \end{figure}
+
+%\subsection{Insertion sort~\label{sec:sort}} 
+%This section shows a series of histograms of an insertion sorting program that 
+%sorts the elements of a given array in non-decreasing order. 
+%The program repeatedly runs 300 times for a given input size. 
+%The input size for the program 
+%varies from 100,000 to 1,160,000 integer elements, which are randomly generated. 
+%Note that each sort program over a specific input size is termed SORT{\it x}: 
+%for instance, SORT100 indicates the insertion sort program over 100K elements. 
+%
+%Figures~\ref{fig:sort1} and ~\ref{fig:sort2} exhibit 
+%histograms of the execution times measured on the same insertion sort program as 
+%the input size grows from 100K to 1,160K elements.% by the steps of 2x. 
+%Note that we used one standard deviation in Figure~\ref{fig:sort400_dist} as
+%a couple of outliers, which were not eliminated by the original protocol, 
+%resulted in disturbing the rendering of a clean distribution. 
+%
+%\begin{figure}[h]
+%	\centering
+%%	\subfigure[PT frequency on SORT25]{
+%%		\includegraphics[scale=0.43]{sort25_dist.eps}
+%%		\label{fig:sort25_dist}
+%%	}
+%%	\subfigure[PT frequency on SORT50]{
+%%		\includegraphics[scale=0.43]{sort50_dist.eps}
+%%		\label{fig:sort50_dist}
+%%	}
+%	\subfigure[PT frequency on SORT100]{
+%		\includegraphics[scale=0.43]{sort100_dist.eps}
+%		\label{fig:sort100_dist}
+%	}
+%	\subfigure[PT frequency on SORT144]{
+%		\includegraphics[scale=0.43]{sort144_dist.eps}
+%		\label{fig:sort144_dist}
+%	}
+%	\subfigure[PT frequency on SORT200]{
+%		\includegraphics[scale=0.43]{sort200_dist.eps}
+%		\label{fig:sort200_dist}
+%	}
+%	\subfigure[PT frequency on SORT288]{
+%		\includegraphics[scale=0.43]{sort288_dist.eps}
+%		\label{fig:sort288_dist}
+%	}
+%	\caption{PT Histograms of SORT25 ... SORT288~\label{fig:sort1}}
+%\end{figure}
+%
+%\clearpage
+%\pagebreak
+%
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on SORT400]{
+%		\includegraphics[scale=0.43]{sort400_dist.eps}
+%		\label{fig:sort400_dist}
+%	}
+%	\subfigure[PT frequency on SORT580]{
+%		\includegraphics[scale=0.43]{sort580_dist.eps}
+%		\label{fig:sort588_dist}
+%	}
+%	\subfigure[PT frequency on SORT800]{
+%		\includegraphics[scale=0.43]{sort800_dist.eps}
+%		\label{fig:sort800_dist}
+%	}
+%	\subfigure[PT frequency on SORT1000]{
+%		\includegraphics[scale=0.43]{sort1000_dist.eps}
+%		\label{fig:sort1000_dist}
+%	}
+%	\caption{PT Histograms of SORT400 ... SORT1000~\label{fig:sort2}}
+%\end{figure}
+%
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on SORT1160]{
+%		\includegraphics[scale=0.43]{sort1160_dist.eps}
+%		\label{fig:sort1160_dist}
+%	}
+%	\subfigure[PT frequency on SORT2320 (not complete)]{
+%		\includegraphics[scale=0.43]{sort2320_dist.eps}
+%		\label{fig:sort2320_dist}
+%	}
+%	\caption{PT Histograms of SORT1160 ... SORT2320~\label{fig:sort3}}
+%\end{figure}
 

### 2018-03-29 pt_characterization/real_data/hist.r

@@ -113,6 +113,64 @@ xlab='PT (ms)', ylab=expression('Frequency'))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
+### new SORT144
+numSamples <- 300
+x = read.csv(file="SORT144.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort144_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 150
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on SORT144 (sample size: 300)', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### new SORT177
+numSamples <- 300
+x = read.csv(file="SORT177.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort177_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
[…]

### 2018-03-29 pt_characterization/real_data/inc_hist.r

@@ -1,0 +1,173 @@
+### INC8.8
+numSamples <- 300
+x = read.csv(file="INC8.8.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("inc8.8_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 100
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on INC8.8 (sample size: 300)', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### INC13.0
+numSamples <- 300
+x = read.csv(file="INC13.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("inc13.0_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
[…]

### 2018-04-03 pt_characterization/real_data/paper.tex
intermachine repeatability check

@@
 
-In this section we present histogram data 
-for two main programs: {\it insertion sort} and {\it matrix multiplication}. 
-For the runs of these programs, we varied their input sizes by 2{\small $\times$} while trying 
-some intermediate sizes increased by $\sqrt{2}$ and measured execution times of the programs over each input size.
+In this section we run three kinds of programs: {\it nested for-loop}, {\it insertion sort} and {\it matrix multiplication}.
+For the runs of the latter two programs, we varied their input sizes by 2{\small $\times$} while trying some intermediate sizes increased by $\sqrt{2}$ and measured execution times of the programs over each input size.
+We then exhibit different kinds of histograms in the subsequent sections.
+
+\subsection{Inter-machine Repeatability Check~\label{sec:diff_machine}}
+
+In this section we check if the use of different machines affects the distribution of execution (process) times on the same program. 
+In other words, we examine {\em inter-machine repeatability}. For this examination, we use the nested for loop program 
+with different task lengths: 16 seconds, 13 seconds, and 17.2 seconds, termed {\em INC16}, {\em INC13}, and {\em INC17.2}, respectively.
+
+Figures~\ref{fig:dm_1}, ~\ref{fig:dm_2}, and ~\ref{fig:dm_3} display the process time (PT) histograms on INC16, INC13, and INC17.2 run on our machines (from {\tt sodb8} to {\tt sodb12}), respectively.
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on INC16]{
+		\includegraphics[scale=0.43]{sodb8_INC16_0_dist.eps}
+		\label{fig:s8_inc16_dist}
+	}	
+	\subfigure[PT frequency on INC16]{
+		\includegraphics[scale=0.43]{sodb9_INC16_0_dist.eps}
+		\label{fig:s9_inc16_dist}
+	}
+	\subfigure[PT frequency on INC16]{
+		\includegraphics[scale=0.43]{sodb10_INC16_0_dist.eps}
+		\label{fig:s10_inc16_dist}
+	}	
+	\subfigure[PT frequency on INC16]{
+		\includegraphics[scale=0.43]{sodb12_INC16_0_dist.eps}
+		\label{fig:s12_inc16_dist}
+	}
+	\caption{PT Histograms on INC16~\label{fig:dm_1}}
+\end{figure}
+
+\newpage
+\clearpage
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on INC13]{
+		\includegraphics[scale=0.43]{sodb8_INC13_0_dist.eps}
+		\label{fig:s8_inc13_dist}
+	}	
+	\subfigure[PT frequency on INC13]{
+		\includegraphics[scale=0.43]{sodb9_INC13_0_dist.eps}
+		\label{fig:s9_inc13_dist}
+	}
+	\subfigure[PT frequency on INC13]{
+		\includegraphics[scale=0.43]{sodb10_INC13_0_dist.eps}
+		\label{fig:s10_inc13_dist}
+	}	
+	\subfigure[PT frequency on INC13]{
+		\includegraphics[scale=0.43]{sodb12_INC13_0_dist.eps}
+		\label{fig:s12_inc13_dist}
+	}
+	\caption{PT Histograms on INC13~\label{fig:dm_2}}
+\end{figure}
+
+\newpage
+\clearpage
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on INC17.2]{
+		\includegraphics[scale=0.43]{sodb8_INC17_2_dist.eps}
+		\label{fig:s8_inc17_2_dist}
+	}	
+	\subfigure[PT frequency on INC17.2]{
+		\includegraphics[scale=0.43]{sodb9_INC17_2_dist.eps}
+		\label{fig:s9_inc17_2_dist}
+	}
+	\subfigure[PT frequency on INC17.2]{
+		\includegraphics[scale=0.43]{sodb10_INC17_2_dist.eps}
+		\label{fig:s10_inc17_2_dist}
+	}	
+	\subfigure[PT frequency on INC17.2]{
+		\includegraphics[scale=0.43]{sodb12_INC17_2_dist.eps}
+		\label{fig:s12_inc17_2_dist}
+	}
+	\caption{PT Histograms on INC17.2~\label{fig:dm_3}}
+\end{figure}
+
+\newpage
+\clearpage
 

### 2018-04-03 pt_characterization/real_data/inc_hist.r

@@ -1,3 +1,364 @@
+### sodb8-INC16
+numSamples <- 1000
+x = read.csv(file="sodb8.INC16.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sodb8_INC16_0_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 300 
+title <- paste("PT frequency on sodb8-INC16.0 (sample size: ",numSamples,")")
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main=title, 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### sodb9-INC16
+numSamples <- 1000
+x = read.csv(file="sodb9.INC16.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sodb9_INC16_0_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
[…]

### 2018-04-03 pt_characterization/real_data/test.c

@@ -1,0 +1,12 @@
+#include <stdio.h>
+
+int g_numCnt = 0;
+
+int main(int argc, char* argv[]){
+    g_numCnt++;
+    
+    int l_numCnt2 = 1;
+    printf("numCnt: %d, numCnt2: %d\n", 
+		g_numCnt, l_numCnt2);
+    return 0;
+}

### 2018-04-12 pt_characterization/real_data/paper.tex
New draft

@@
 
-The following section shows histograms for runs on 
-different real-world programs with varying input sizes.
-
-\section{Histograms of the Execution Times on Real-World Programs~\label{sec:real-world}}
-
-In this section we run three kinds of programs: {\it nested for-loop}, {\it insertion sort} and {\it matrix multiplication}.
-For the runs of the latter two programs, we varied their input sizes by 2{\small $\times$} while trying some intermediate sizes increased by $\sqrt{2}$ and measured execution times of the programs over each input size.
-We then exhibit different kinds of histograms in the subsequent sections.
+\section{Preliminary Experiments}
 
@@
 In this section we check if the use of different machines affects the distribution of execution (process) times on the same program. 
-In other words, we examine {\em inter-machine repeatability}. For this examination, we use the nested for loop program 
-with different task lengths: 16 seconds, 13 seconds, and 17.2 seconds, termed {\em INC16}, {\em INC13}, and {\em INC17.2}, respectively.
+In other words, we examine {\em inter-machine repeatability}. For this examination, we use a nested for loop program, here termed INC, 
+with different task lengths: 16 seconds, 13 seconds, and 17.2 seconds, each termed {\em INC16}, {\em INC13}, and {\em INC17.2}, respectively.
 
 Figures~\ref{fig:dm_1}, ~\ref{fig:dm_2}, and ~\ref{fig:dm_3} display the process time (PT) histograms on INC16, INC13, and INC17.2 run on our machines (from {\tt sodb8} to {\tt sodb12}), respectively.
+Based on these figures, we conclude that we should stick to one machine  for consistency, and {\tt sodb9} should be chosen in a sense that the data measured there consistently reveals a distinct binormal pattern.
 
@@
 This section compares program time histograms 
-of insertion sort and INC (a nested-for-loop) programs. 
-The insertion sort program 
-sorts the elements of a given array in non-decreasing order.  
+of an insertion sort and a nested-for-loop programs (termed INC) with different input sizes.
+The insertion sort program sorts the elements of a given array in non-decreasing order.
 The program repeatedly runs 300 times for a given input size.
 The input size for the program varies from 144K to 344K integer elements, which are randomly generated.
-Note that each sort program over a specific input size is termed SORT{\it x}: 
-for instance, SORT100 indicates the insertion sort program over 100K elements. 
-An INC program's task length is correspondingly determined by the program time 
-of an SORT program. In Figures~\ref{fig:sort1},~\ref{fig:sort2}, and~\ref{fig:sort3} 
+Note that each sort program over a specific input size is termed SORT{\it x}: for instance, SORT100 indicates the insertion sort program over 100K elements.
+An INC program's task length is correspondingly determined by the program time of an SORT program.
+In Figures~\ref{fig:sort1},~\ref{fig:sort2}, and~\ref{fig:sort3}
 we perform a match on an SORT program and its corresponding INC program.
@@
 	\centering
-	\subfigure[PT frequency on SORT144]{
+	\subfigure[PT frequency on SORT144 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sort144_dist.eps}
@@
 	}	
-	\subfigure[PT frequency on INC8.8]{
+	\subfigure[PT frequency on INC8.8 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{inc8.8_dist.eps}
@@
 	}
-	\subfigure[PT frequency on SORT177]{
+	\subfigure[PT frequency on SORT177 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sort177_dist.eps}
@@
 	}	
-	\subfigure[PT frequency on INC13]{
+	\subfigure[PT frequency on INC13 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{inc13.0_dist.eps}
@@
 
+\clearpage
+\pagebreak
+
 \begin{figure}[h]
 	\centering
-	\subfigure[PT frequency on SORT144]{
+	\subfigure[PT frequency on SORT144 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sort244_dist.eps}
@@
 	}	
-	\subfigure[PT frequency on INC25.8]{
+	\subfigure[PT frequency on INC25.8 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{inc25.8_dist.eps}
@@
 	}
-	\subfigure[PT frequency on SORT288]{
+	\subfigure[PT frequency on SORT288  on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sort177_dist.eps}
@@
 	}	
-	\subfigure[PT frequency on INC35]{
+	\subfigure[PT frequency on INC35 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{inc13.0_dist.eps}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 	\centering
-	\subfigure[PT frequency on SORT344]{
+	\subfigure[PT frequency on SORT344  on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sort344_dist.eps}
@@
 	}	
-	\subfigure[PT frequency on INC51.5]{
+	\subfigure[PT frequency on INC51.5 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{inc51.5_dist.eps}
@@
 
-\section{Additional Insertion Sort~\label{sec:add_new}}
+\section{Histograms of the Execution Times on Real-World Programs~\label{sec:real-world}} 
+This section shows histograms for runs on 
+different real-world programs with varying input sizes.  
+%We run three kinds of programs: {\it nested for-loop}, {\it insertion sort} and {\it matrix multiplication}.  
+In addition to the insertion sort program, we use another kind of program---{\it matrix multiplication}.
+For these programs, we varied their input sizes by 2{\small $\times$} while trying some intermediate sizes increased by $\sqrt{2}$ and measured execution times of the programs over each input size.
+We then exhibit different kinds of histograms in the subsequent sections.
+
+\clearpage
+\pagebreak
+
+\subsection{Insertion Sort on {\tt sodb9}~\label{sec:new_ins_sort}} 
+This section shows a series of histograms of an insertion sorting program that 
+sorts the elements of a given array in non-decreasing order. 
+The program repeatedly runs 300 times for a given input size. 
+Note that each sort program over a specific input size is termed SORT{\it x}: 
+for instance, SORT100 indicates the insertion sort program over 100K elements. 
+(This experiment was performed on {\tt sodb9}.)
+
+\begin{figure}[htp!]
+	\centering
+	\subfigure[PT frequency on SORT100 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{s9_sort100_dist.eps}
+		\label{fig:s9_sort100_dist}
+	}
+	\subfigure[PT frequency on SORT200 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{s9_sort200_dist.eps}
+		\label{fig:s9_sort200_dist}
+	}
+	\subfigure[PT frequency on SORT400 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{s9_sort400_dist.eps}
+		\label{fig:s9_sort400_dist}
+	}
+	\subfigure[PT frequency on SORT800 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{s9_sort800_dist.eps}
+		\label{fig:s9_sort800_dist}
+	}
+	\caption{PT Histograms of SORT100 ... SORT800 on {\tt sodb9}~\label{fig:new_s9_sort1}}
+\end{figure}
+
+\newpage
+\pagebreak
+
+\section{Appendix}
+
+\subsection{Insertion Sort on {\tt sodb8}~\label{sec:add_new}} 
+This section shows a series of histograms of an insertion sorting program that 
+sorts the elements of a given array in non-decreasing order. 
+The program repeatedly runs 300 times for a given input size. 
+The input size for the program 
+varies from 100,000 to 1,160,000 integer elements, which are randomly generated. 
+Note that each sort program over a specific input size is termed SORT{\it x}: 
+for instance, SORT100 indicates the insertion sort program over 100K elements. 
+This experiment was performed on {\tt sodb8}.
 
@@
 %	}
-	\subfigure[PT frequency on SORT100]{
+	\subfigure[PT frequency on SORT100 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sort100_dist.eps}
@@
 	}
-	\subfigure[PT frequency on SORT400]{
+	\subfigure[PT frequency on SORT400 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sort400_dist.eps}
@@
 	}
-	\subfigure[PT frequency on SORT580]{
+	\subfigure[PT frequency on SORT580 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sort580_dist.eps}
@@
 	\centering
-	\subfigure[PT frequency on SORT800]{
+	\subfigure[PT frequency on SORT800 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sort800_dist.eps}
@@
 	}
-	\subfigure[PT frequency on SORT1000]{
+	\subfigure[PT frequency on SORT1000 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sort1000_dist.eps}
@@
 	}
-	\subfigure[PT frequency on SORT1160]{
+	\subfigure[PT frequency on SORT1160 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sort1160_dist.eps}
@@
 	}
-	\subfigure[PT frequency on SORT2320 (not complete)]{
+	\subfigure[PT frequency on SORT2320 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sort2320_dist.eps}
@@
 the same matrix multiplication program in column major 
-as the input sizes grows.
+as the input sizes grows. The program was run on {\tt sodb10}.
 
@@
 	\centering
-	\subfigure[PT frequency on MATC1K]{
+	\subfigure[PT frequency on MATC1K on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{matc1024_dist.eps}
@@
 	}
-	\subfigure[PT frequency on MATC2K]{
+	\subfigure[PT frequency on MATC2K on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{matc2048_dist.eps}
@@
 	}
-	\subfigure[PT frequency on MATC4K]{
+	\subfigure[PT frequency on MATC4K on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{matc4096_dist.eps}
@@
 	}
-	\subfigure[PT frequency on MATC8K]{
+	\subfigure[PT frequency on MATC8K on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{matc8192_dist.eps}
@@
 the same matrix multiplication program in row major as the input sizes grows. 
+The program was run on {\tt sodb12}.
 
@@
 	\centering
-	\subfigure[PT frequency on MATR1K]{
+	\subfigure[PT frequency on MATR1K on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{matr1024_dist.eps}
@@
 	}
-	\subfigure[PT frequency on MATR2K]{
+	\subfigure[PT frequency on MATR2K on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{matr2048_dist.eps}
@@
 	}
-	\subfigure[PT frequency on MATR4K]{
+	\subfigure[PT frequency on MATR4K on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{matr4096_dist.eps}
@@
 	}
-	\subfigure[PT frequency on MATR8K]{
+	\subfigure[PT frequency on MATR8K on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{matr8192_dist.eps}
@@
 \end{figure}
-
-\clearpage
-\pagebreak
-
-%\end{comment}
 

### 2018-04-12 pt_characterization/real_data/sodb9/sort_hist.r

@@ -1,0 +1,124 @@
+### SORT100
+numSamples <- 300
+x = read.csv(file="SORT100.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("s9_sort100_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 150 
+title <- paste("PT frequency on SORT100 on sodb9 (sample size: ",numSamples,")")
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main=title, 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### SORT200
+numSamples <- 300
+x = read.csv(file="SORT200.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("s9_sort200_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
[…]

### 2018-04-26 pt_characterization/real_data/inc_hist.r
Next version

@@ -120,6 +120,36 @@ dev.off()
 
 ### sodb8-INC13
 numSamples <- 1000
+x = read.csv(file="sodb8.INC13.0_new.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sodb8_INC13_0_new_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 300 
+title <- paste("PT frequency on sodb8-new-INC13.0 (sample size: ",numSamples,")")
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main=title, 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### sodb8-INC13
+numSamples <- 1000
 x = read.csv(file="sodb8.INC13.0.dat",head=TRUE,sep="\t")
 x <- subset(x, x$ITERNUM <= numSamples)
 setEPS()
@@ -139,7 +169,7 @@ xmin <-min(x$PRTIME)
 xmax <-max(x$PRTIME)
 ymax <- max(h$counts)
 ymax <- ceiling(ymax/100)*100
-ymax <- 300 
+ymax <- 500 
[…]

### 2018-05-01 pt_characterization/machine_cons_check/paper.tex
New drafts

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
 \usepackage{auto-pst-pdf}
\usepackage{graphicx}
%\usepackage{arydshln}
\usepackage{ifpdf}
\ifpdf
  \usepackage{epstopdf}
\fi
\usepackage{multirow}
\usepackage{epsfig}
\usepackage{float}
\usepackage{url}
\usepackage{color}
\usepackage{subfigure}

%\newcommand\solidrule[1][1cm]{\rule[0.5ex]{#1}{.4pt}}
%\newcommand\dashedrule{\mbox{%
%  \solidrule[2mm]\hspace{2mm}\solidrule[2mm]\hspace{2mm}\solidrule[2mm]}}

%\usepackage{hyperref}

\begin{document}
\title{Empirical Check of Machine Homogeneity on Timing Protocol}

\author{
Young-Kyoon Suh\\
Data-Intensive Systems Laboratory\\
School of Computer Science and Engineering\\
Kyungpook National University\\
}
\maketitle

\section{Description}
In this work we do a check to see if the result on the same experiment is different 
on different nodes, each with the same machine specification.

\section{Preliminary Experiments}

We provide a short description of our preliminary experimental runs.

\paragraph{Experimental Notes:} In our experiments 
we used four nodes ({\tt sodb8}, {\tt sodb9}, {\tt sodb10}, and {\tt sodb12} ) in the same cluster. 
A short summary of the experiments is exhibited in Table~\ref{tab:exp_notes}.

%% done
\begin{table}[h]
\begin{center}
\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
Machine & Task Length & Description & Time Length\\ \hline
{\tt sodb8} (plugged into \hbox{bottom} {\em left} power strip) & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
{\tt sodb9}  (plugged into \hbox{bottom} {\em left} power strip)  &  INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
{\tt sodb10}  (plugged into \hbox{bottom} {\em left} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
{\tt sodb12}  (plugged into \hbox{bottom} {\em left} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{Detailed description of INC data used for histograms\label{tab:exp_notes}}
\end{table}

\clearpage
\newpage
\paragraph{Measurement Quality Across the SoDB Nodes:} 
Fig.~\ref{fig:machine_comp} shows 
the standard deviation and relative error of 
elapsed time (ET) and process time (PT) on the different SoDB nodes 
as the task length of INC increases.

\begin{figure}[h]
	\centering
	\subfigure[Standard Deviation - ET]{
		\includegraphics[scale=0.6]{overall/machine_et_std.eps}
        \label{fig:et_std}
    }
    \subfigure[Relative Error - ET]{
        \includegraphics[scale=0.6]{overall/machine_et_re.eps}
        \label{fig:et_re}
    }
	\subfigure[Standard Deviation - PT]{
		\includegraphics[scale=0.6]{overall/machine_pt_std.eps}
        \label{fig:pt_std}
    }
    \subfigure[Relative Error - PT]{
        \includegraphics[scale=0.6]{overall/machine_pt_re.eps}
        \label{fig:pt_re}
    }
    \caption{Measurement Quality Comparison among Different SoDB Machines}
    \label{fig:machine_comp}
\end{figure}

%\newpage

Each of the following sections exhibits histograms of ET and PT 
over increasing INC's task lengths on each individual node.

\input{sodb8/res.tex}
\input{sodb9/res.tex}
\input{sodb10/res.tex}
\input{sodb12/res.tex}
\input{post_exp.tex}

\newpage

\section{Conclusion}
My recommendation is to avoid {\tt sodb8} and to use the rest of 
the SoDB nodes for the rest of our experiments. Further investigation is 
needed to examine why the experiment results on {\tt sodb8} are different 
than those of the other nodes.

\end{document}

### 2018-05-01 pt_characterization/machine_cons_check/post_exp.tex
New drafts

\clearpage
\newpage

\section{Post Experiments}

We provide a short description of newer experimental runs.

\paragraph{Experimental Notes:} In the experiments 
we also used four nodes ({\tt sodb8}, {\tt sodb9}, {\tt sodb10}, and {\tt sodb12} ) in the same cluster. 
A short summary of the experiments is shown in Table~\ref{tab:exp_notes2}.

%% done
\begin{table}[h]
\begin{center}
\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
Machine & Task Length & Description & Time Length\\ \hline
{\tt sodb8} (plugged into \hbox{bottom} {\em left} power strip) &  INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
{\tt sodb9}  (plugged into \hbox{bottom} {\em left} power strip)  &  IINC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
{\tt sodb10}  (plugged into \hbox{bottom} {\em left} power strip)  & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
{\tt sodb12}  (plugged into \hbox{bottom} {\em left} power strip)  & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{Detailed description of INC data used for histograms\label{tab:exp_notes2}}
\end{table}

\subsection{Inter-Machine Homogeneity~\label{sec:diff_machine}}

In this section we check if the use of different machines affects the distribution of execution (process) times on the same program. 
In other words, we examine {\em inter-machine repeatability}. For this examination, we use a nested for loop program, here termed INC, 
with different task lengths: 16 seconds, 13 seconds, and 17.2 seconds, each termed {\em INC16}, {\em INC13}, and {\em INC17.2}, respectively.

Figures~\ref{fig:dm_1}, ~\ref{fig:dm_2}, and ~\ref{fig:dm_3} display the process time (PT) histograms on INC16, INC13, and INC17.2 run on our machines (from {\tt sodb8} to {\tt sodb12}), respectively. Based on these figures, we conclude that we should stick to one machine  for consistency, and {\tt sodb9} should be chosen in a sense that the data measured there consistently reveals a distinct binormal pattern.

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on INC16]{
		\includegraphics[scale=0.43]{newer_exp/sodb8_INC16_0_dist.eps}
		\label{fig:s8_inc16_dist}
	}	
	\subfigure[PT frequency on INC16]{
		\includegraphics[scale=0.43]{newer_exp/sodb9_INC16_0_dist.eps}
		\label{fig:s9_inc16_dist}
	}
	\subfigure[PT frequency on INC16]{
		\includegraphics[scale=0.43]{newer_exp/sodb10_INC16_0_dist.eps}
		\label{fig:s10_inc16_dist}
	}	
	\subfigure[PT frequency on INC16]{
		\includegraphics[scale=0.43]{newer_exp/sodb12_INC16_0_dist.eps}
		\label{fig:s12_inc16_dist}
	}
	\caption{PT Histograms on INC16~\label{fig:dm_1}}
\end{figure}

\newpage
\clearpage

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on INC13]{
		\includegraphics[scale=0.43]{newer_exp/sodb8_INC13_0_dist.eps}
		\label{fig:s8_inc13_dist}
	}	
	\subfigure[PT frequency on INC13]{
		\includegraphics[scale=0.43]{newer_exp/sodb9_INC13_0_dist.eps}
		\label{fig:s9_inc13_dist}
	}
	\subfigure[PT frequency on INC13]{
		\includegraphics[scale=0.43]{newer_exp/sodb10_INC13_0_dist.eps}
		\label{fig:s10_inc13_dist}
	}	
	\subfigure[PT frequency on INC13]{
		\includegraphics[scale=0.43]{newer_exp/sodb12_INC13_0_dist.eps}
		\label{fig:s12_inc13_dist}
	}
	\caption{PT Histograms on INC13~\label{fig:dm_2}}
\end{figure}

\newpage
\clearpage

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on INC17.2]{
		\includegraphics[scale=0.43]{newer_exp/sodb8_INC17_2_dist.eps}
		\label{fig:s8_inc17_2_dist}
	}	
	\subfigure[PT frequency on INC17.2]{
		\includegraphics[scale=0.43]{newer_exp/sodb9_INC17_2_dist.eps}
		\label{fig:s9_inc17_2_dist}
	}
	\subfigure[PT frequency on INC17.2]{
		\includegraphics[scale=0.43]{newer_exp/sodb10_INC17_2_dist.eps}
		\label{fig:s10_inc17_2_dist}
	}	
	\subfigure[PT frequency on INC17.2]{
		\includegraphics[scale=0.43]{newer_exp/sodb12_INC17_2_dist.eps}
		\label{fig:s12_inc17_2_dist}
	}
	\caption{PT Histograms on INC17.2~\label{fig:dm_3}}
\end{figure}

\newpage
\clearpage

\subsection{Investigation of Daemons' Influence on Program Time Distribution Regarding Machine Dependence~\label{sec:daemon_impact}}

For the same task length, or INC16, 
I examined in each of the four runs a few iterations at which more than three daemons (except INC and proc monitor processes) that had positive PT were captured.  
From Table~\ref{fig:daemon}, I suspect that PT distribution seems most likely to be affected by two facts: {\it how longer the same daemon ran than usual}, and {\it how many different daemon processes appeared and how long it ran}.

\begin{table}[htp!]
\centering
{
 \begin{tabular}{|p{1.5cm}|p{2cm}|p{12.5cm}|} \hline
Machine Name & Iteration \# & Process Name (id, PT(msec))\\ \hline
{\tt sodb8}  & 174 & {\tt java} ({\tt 2349}, 15), {\tt java} ({\tt 2335}, 2),  {\tt md127\_raid1} ({\tt 457}, 1), {\tt kslowd000} ({\tt 166}, 1), {\tt kslowd001} ({\tt 167}, 1) \\ \cline{2-3}
 					& 278 & {\tt java} ({\tt 2877}, 16), {\tt java} ({\tt 2335}, 2),  {\tt kslowd000} ({\tt 166}, 1), {\tt kslowd001} ({\tt 167}, 1) \\ \cline{2-3}
			  &  486 & {\tt java} ({\tt 3190}, 21), {\tt java} ({\tt 2335}, 2),  {\tt md127\_raid1} ({\tt 457}, 1), {\tt kslowd000} ({\tt 166}, 1)\\ \cline{2-3}
				  &  526 & {\tt md127\_raid1} ({\tt 457}, 1), {\tt kslowd000} ({\tt 166}, 1), {\tt kslowd001} ({\tt 167}, 1), {\tt jbd2/md127-8} ({\tt 470}, 1)\\ \cline{2-3}
				  &  555 & {\tt java} ({\tt 3815}, 8), {\tt java} ({\tt 2335}, 2), {\tt kslowd000} ({\tt 166}, 1), {\tt kslowd001} ({\tt 167}, 1)\\ \hline
{\tt sodb9}  & 105 & {\tt java} ({\tt 6634}, 16), {\tt java} ({\tt 6621}, 2),  {\tt kslowd000} ({\tt 166}, 1), {\tt kslowd001} ({\tt 167}, 1) \\ \cline{2-3}
				  & 175 & {\tt java} ({\tt 6942}, 7), {\tt java} ({\tt 6621}, 2),  {\tt kslowd000} ({\tt 166}, 1), {\tt kslowd001} ({\tt 167}, 1) \\ \cline{2-3}
				  & 245 & {\tt java} ({\tt 7259}, 6), {\tt java} ({\tt 6621}, 2),  {\tt kslowd000} ({\tt 166}, 1), {\tt kslowd001} ({\tt 167}, 1) \\ \cline{2-3}
				  & 280 & {\tt java} ({\tt 7365}, 3), {\tt java} ({\tt 6621}, 2),  {\tt kslowd000} ({\tt 166}, 1), {\tt kslowd001} ({\tt 167}, 1) \\ \cline{2-3}
				  & 350 & {\tt java} ({\tt 7577}, 3), {\tt java} ({\tt 6621}, 2),  {\tt kslowd000} ({\tt 166}, 1), {\tt kslowd001} ({\tt 167}, 1) \\ \cline{2-3}
				  & 420 & {\tt java} ({\tt 7683}, 3), {\tt java} ({\tt 6621}, 2),  {\tt kslowd000} ({\tt 166}, 1), {\tt kslowd001} ({\tt 167}, 1) \\ \cline{2-3}
				  & 490 & {\tt java} ({\tt 7894}, 4), {\tt java} ({\tt 6621}, 2),  {\tt kslowd000} ({\tt 166}, 1), {\tt kslowd001} ({\tt 167}, 1) \\ \hline
{\tt sodb10}  & 105 & {\tt java} ({\tt 28394}, 14), {\tt java} ({\tt 28381}, 2),  {\tt kslowd000} ({\tt 166}, 1), {\tt kslowd001} ({\tt 167}, 1) \\ \cline{2-3}
					 & 314 & {\tt java} ({\tt 28702}, 18), {\tt java} ({\tt 28381}, 2),  {\tt md127\_raid1} ({\tt 455}, 1),  {\tt kslowd001} ({\tt 167}, 1) \\ \hline
 {\tt sodb12}  & 280 &  {\tt kslowd001} ({\tt 167}, 154),  {\tt kslowd000} ({\tt 166}, 150), {\tt java} ({\tt 14820}, 37), {\tt java} ({\tt 14807}, 2) \\ \cline{2-3}
	& 311 &  {\tt kslowd000} ({\tt 166}, 154),  {\tt kslowd001} ({\tt 167}, 153), {\tt java} ({\tt 14807}, 2), {\tt java} ({\tt 15653}, 1), {\tt khugepaged} ({\tt 30}, 1)\\ \cline{2-3}		& 373 &  {\tt kslowd000} ({\tt 166}, 154),  {\tt kslowd001} ({\tt 167}, 152), {\tt java} ({\tt 14807}, 2), {\tt java} ({\tt 15747}, 2)\\ \cline{2-3}
& 435 &  {\tt kslowd000} ({\tt 166}, 154),  {\tt kslowd001} ({\tt 167}, 153), {\tt java} ({\tt 15934}, 6), {\tt java} ({\tt 14807}, 2)\\ \cline{2-3}
& 466 &  {\tt kslowd000} ({\tt 166}, 152), {\tt java} ({\tt 16121}, 6), {\tt java} ({\tt 14807}, 2), {\tt kblockd/0} ({\tt 16}, 1)\\ \cline{2-3}
 \hline
 \end{tabular}
  }
 \caption{Some daemon processes captured across different machines~\label{fig:daemon}}
\end{table}

### 2018-05-01 pt_characterization/machine_cons_check/sodb10/res.tex
New drafts

\newpage

\subsection{{\tt sodb10}~\label{sec:sodb10_hist}} 
This section exhibits histograms on the EMPv5 data obtained on {\tt sodb10}. 
The detailed description of the base data are from Table~\ref{tab:exp_notes}.

\subsubsection{ET}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC1]{
		\includegraphics[scale=0.43]{sodb10/1_sec_et_hist_v5.eps}
		\label{fig:s10_inc1_et_hist_v5}
	}
	\subfigure[ET frequency on INC2]{
		\includegraphics[scale=0.43]{sodb10/2_sec_et_hist_v5.eps}
		\label{fig:s10_inc2_et_hist_v5}
	}
	\subfigure[ET frequency on INC4]{
		\includegraphics[scale=0.43]{sodb10/4_sec_et_hist_v5.eps}
		\label{fig:s10_inc4_et_hist_v5}
	}
	\subfigure[ET frequency on INC8]{
		\includegraphics[scale=0.43]{sodb10/8_sec_et_hist_v5.eps}
		\label{fig:s10_inc8_et_hist_v5}
	}
	\caption{ET Histograms of INC1 ... INC8~\label{fig:s10_et_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC16]{
		\includegraphics[scale=0.43]{sodb10/16_sec_et_hist_v5.eps}
		\label{fig:s10_inc16_et_hist_v5}
	}
	\subfigure[ET frequency on INC32]{
		\includegraphics[scale=0.43]{sodb10/32_sec_et_hist_v5.eps}
		\label{fig:s10_inc32_et_hist_v5}
	}
	\subfigure[ET frequency on INC64]{
		\includegraphics[scale=0.43]{sodb10/64_sec_et_hist_v5.eps}
		\label{fig:s10_inc64_et_hist_v5}
	}
	\subfigure[ET frequency on INC128]{
		\includegraphics[scale=0.43]{sodb10/128_sec_et_hist_v5.eps}
		\label{fig:s10_inc128_et_hist_v5}
	}
	\caption{ET Histograms of INC16 ... INC128~\label{fig:s10_et_hist2}}
\end{figure}

\newpage

\subsubsection{PT}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC1]{
		\includegraphics[scale=0.43]{sodb10/1_sec_pt_hist_v5.eps}
		\label{fig:s10_inc1_hist_v5}
	}
	\subfigure[PT frequency on INC2]{
		\includegraphics[scale=0.43]{sodb10/2_sec_pt_hist_v5.eps}
		\label{fig:s10_inc2_hist_v5}
	}
	\subfigure[PT frequency on INC4]{
		\includegraphics[scale=0.43]{sodb10/4_sec_pt_hist_v5.eps}
		\label{fig:s10_inc4_hist_v5}
	}
	\subfigure[PT frequency on INC8]{
		\includegraphics[scale=0.43]{sodb10/8_sec_pt_hist_v5.eps}
		\label{fig:s10_inc8_hist_v5}
	}
	\caption{PT Histograms of INC1 ... INC8~\label{fig:s10_pt_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC16]{
		\includegraphics[scale=0.43]{sodb10/16_sec_pt_hist_v5.eps}
		\label{fig:s10_inc16_hist_v5}
	}
	\subfigure[PT frequency on INC32]{
		\includegraphics[scale=0.43]{sodb10/32_sec_pt_hist_v5.eps}
		\label{fig:s10_inc32_hist_v5}
	}
	\subfigure[PT frequency on INC64]{
		\includegraphics[scale=0.43]{sodb10/64_sec_pt_hist_v5.eps}
		\label{fig:s10_inc64_hist_v5}
	}
	\subfigure[PT frequency on INC128]{
		\includegraphics[scale=0.43]{sodb10/128_sec_pt_hist_v5.eps}
		\label{fig:s10_inc128_hist_v5}
	}
	\caption{PT Histograms of INC16 ... INC64~\label{fig:s10_pt_hist2}}
\end{figure}

### 2018-05-01 pt_characterization/machine_cons_check/sodb12/res.tex
New drafts

\newpage

\subsection{{\tt sodb12}~\label{sec:sodb12_hist}} 
This section exhibits histograms on the EMPv5 data obtained on {\tt sodb12}. 
The detailed description of the base data are from Table~\ref{tab:exp_notes}.

\subsubsection{ET}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC1]{
		\includegraphics[scale=0.43]{sodb12/1_sec_et_hist_v5.eps}
		\label{fig:s12_inc1_et_hist_v5}
	}
	\subfigure[ET frequency on INC2]{
		\includegraphics[scale=0.43]{sodb12/2_sec_et_hist_v5.eps}
		\label{fig:s12_inc2_et_hist_v5}
	}
	\subfigure[ET frequency on INC4]{
		\includegraphics[scale=0.43]{sodb12/4_sec_et_hist_v5.eps}
		\label{fig:s12_inc4_et_hist_v5}
	}
	\subfigure[ET frequency on INC8]{
		\includegraphics[scale=0.43]{sodb12/8_sec_et_hist_v5.eps}
		\label{fig:s12_inc8_et_hist_v5}
	}
	\caption{ET Histograms of INC1 ... INC8~\label{fig:s12_et_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC16]{
		\includegraphics[scale=0.43]{sodb12/16_sec_et_hist_v5.eps}
		\label{fig:s12_inc16_et_hist_v5}
	}
	\subfigure[ET frequency on INC32]{
		\includegraphics[scale=0.43]{sodb12/32_sec_et_hist_v5.eps}
		\label{fig:s12_inc32_et_hist_v5}
	}
	\subfigure[ET frequency on INC64]{
		\includegraphics[scale=0.43]{sodb12/64_sec_et_hist_v5.eps}
		\label{fig:s12_inc64_et_hist_v5}
	}
	\subfigure[ET frequency on INC128]{
		\includegraphics[scale=0.43]{sodb12/128_sec_et_hist_v5.eps}
		\label{fig:s12_inc128_et_hist_v5}
	}
	\caption{ET Histograms of INC16 ... INC128~\label{fig:s12_et_hist2}}
\end{figure}

\newpage

\subsubsection{PT}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC1]{
		\includegraphics[scale=0.43]{sodb12/1_sec_pt_hist_v5.eps}
		\label{fig:s12_inc1_hist_v5}
	}
	\subfigure[PT frequency on INC2]{
		\includegraphics[scale=0.43]{sodb12/2_sec_pt_hist_v5.eps}
		\label{fig:s12_inc2_hist_v5}
	}
	\subfigure[PT frequency on INC4]{
		\includegraphics[scale=0.43]{sodb12/4_sec_pt_hist_v5.eps}
		\label{fig:s12_inc4_hist_v5}
	}
	\subfigure[PT frequency on INC8]{
		\includegraphics[scale=0.43]{sodb12/8_sec_pt_hist_v5.eps}
		\label{fig:s12_inc8_hist_v5}
	}
	\caption{PT Histograms of INC1 ... INC8~\label{fig:s12_pt_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC16]{
		\includegraphics[scale=0.43]{sodb12/16_sec_pt_hist_v5.eps}
		\label{fig:s12_inc16_hist_v5}
	}
	\subfigure[PT frequency on INC32]{
		\includegraphics[scale=0.43]{sodb12/32_sec_pt_hist_v5.eps}
		\label{fig:s12_inc32_hist_v5}
	}
	\subfigure[PT frequency on INC64]{
		\includegraphics[scale=0.43]{sodb12/64_sec_pt_hist_v5.eps}
		\label{fig:s12_inc64_hist_v5}
	}
	\subfigure[PT frequency on INC128]{
		\includegraphics[scale=0.43]{sodb12/128_sec_pt_hist_v5.eps}
		\label{fig:s12_inc128_hist_v5}
	}
	\caption{PT Histograms of INC16 ... INC64~\label{fig:s12_pt_hist2}}
\end{figure}

### 2018-05-01 pt_characterization/machine_cons_check/sodb8/res.tex
New drafts

\newpage

\subsection{{\tt sodb8}~\label{sec:sodb8_hist}} 
This section exhibits histograms on the EMPv5 data obtained on {\tt sodb8}. 
The detailed description of the base data are from Table~\ref{tab:exp_notes}.

\subsubsection{ET}

\begin{figure}[h]
	\centering
	\subfigure[ET frequency on INC1]{
		\includegraphics[scale=0.43]{sodb8/1_sec_et_hist_v5.eps}
		\label{fig:inc1_et_hist_v5}
	}
	\subfigure[ET frequency on INC2]{
		\includegraphics[scale=0.43]{sodb8/2_sec_et_hist_v5.eps}
		\label{fig:inc2_et_hist_v5}
	}
	\subfigure[ET frequency on INC4]{
		\includegraphics[scale=0.43]{sodb8/4_sec_et_hist_v5.eps}
		\label{fig:inc4_et_hist_v5}
	}
	\subfigure[ET frequency on INC8]{
		\includegraphics[scale=0.43]{sodb8/8_sec_et_hist_v5.eps}
		\label{fig:inc8_et_hist_v5}
	}
	\caption{ET Histograms of INC1 ... INC8~\label{fig:s8_et_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC16]{
		\includegraphics[scale=0.43]{sodb8/16_sec_et_hist_v5.eps}
		\label{fig:inc16_et_hist_v5}
	}
	\subfigure[ET frequency on INC32]{
		\includegraphics[scale=0.43]{sodb8/32_sec_et_hist_v5.eps}
		\label{fig:inc32_et_hist_v5}
	}
	\subfigure[ET frequency on INC64]{
		\includegraphics[scale=0.43]{sodb8/64_sec_et_hist_v5.eps}
		\label{fig:inc64_et_hist_v5}
	}
	\subfigure[ET frequency on INC128]{
		\includegraphics[scale=0.43]{sodb8/128_sec_et_hist_v5.eps}
		\label{fig:inc128_et_hist_v5}
	}
	\caption{ET Histograms of INC16 ... INC128~\label{fig:s8_et_hist2}}
\end{figure}

\newpage

\subsubsection{PT}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC1]{
		\includegraphics[scale=0.43]{sodb8/1_sec_pt_hist_v5.eps}
		\label{fig:inc1_hist_v5}
	}
	\subfigure[PT frequency on INC2]{
		\includegraphics[scale=0.43]{sodb8/2_sec_pt_hist_v5.eps}
		\label{fig:inc2_hist_v5}
	}
	\subfigure[PT frequency on INC4]{
		\includegraphics[scale=0.43]{sodb8/4_sec_pt_hist_v5.eps}
		\label{fig:inc4_hist_v5}
	}
	\subfigure[PT frequency on INC8]{
		\includegraphics[scale=0.43]{sodb8/8_sec_pt_hist_v5.eps}
		\label{fig:inc8_hist_v5}
	}
	\caption{PT Histograms of INC1 ... INC8~\label{fig:s8_pt_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC16]{
		\includegraphics[scale=0.43]{sodb8/16_sec_pt_hist_v5.eps}
		\label{fig:inc16_hist_v5}
	}
	\subfigure[PT frequency on INC32]{
		\includegraphics[scale=0.43]{sodb8/32_sec_pt_hist_v5.eps}
		\label{fig:inc32_hist_v5}
	}
	\subfigure[PT frequency on INC64]{
		\includegraphics[scale=0.43]{sodb8/64_sec_pt_hist_v5.eps}
		\label{fig:inc64_hist_v5}
	}
	\subfigure[PT frequency on INC128]{
		\includegraphics[scale=0.43]{sodb8/128_sec_pt_hist_v5.eps}
		\label{fig:inc128_hist_v5}
	}
	\caption{PT Histograms of INC16 ... INC64~\label{fig:s8_pt_hist2}}
\end{figure}

### 2018-05-01 pt_characterization/machine_cons_check/sodb9/res.tex
New drafts

\newpage

\subsection{{\tt sodb9}~\label{sec:sodb9_hist}} 
This section exhibits histograms on the EMPv5 data obtained on {\tt sodb9}. 
The detailed description of the base data are from Table~\ref{tab:exp_notes}.

\subsubsection{ET}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC1]{
		\includegraphics[scale=0.43]{sodb9/1_sec_et_hist_v5.eps}
		\label{fig:s9_inc1_et_hist_v5}
	}
	\subfigure[ET frequency on INC2]{
		\includegraphics[scale=0.43]{sodb9/2_sec_et_hist_v5.eps}
		\label{fig:s9_inc2_et_hist_v5}
	}
	\subfigure[ET frequency on INC4]{
		\includegraphics[scale=0.43]{sodb9/4_sec_et_hist_v5.eps}
		\label{fig:s9_inc4_et_hist_v5}
	}
	\subfigure[ET frequency on INC8]{
		\includegraphics[scale=0.43]{sodb9/8_sec_et_hist_v5.eps}
		\label{fig:s9_inc8_et_hist_v5}
	}
	\caption{ET Histograms of INC1 ... INC8~\label{fig:s9_et_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[ET frequency on INC16]{
		\includegraphics[scale=0.43]{sodb9/16_sec_et_hist_v5.eps}
		\label{fig:s9_inc16_et_hist_v5}
	}
	\subfigure[ET frequency on INC32]{
		\includegraphics[scale=0.43]{sodb9/32_sec_et_hist_v5.eps}
		\label{fig:s9_inc32_et_hist_v5}
	}
	\subfigure[ET frequency on INC64]{
		\includegraphics[scale=0.43]{sodb9/64_sec_et_hist_v5.eps}
		\label{fig:s9_inc64_et_hist_v5}
	}
	\subfigure[ET frequency on INC128]{
		\includegraphics[scale=0.43]{sodb9/128_sec_et_hist_v5.eps}
		\label{fig:s9_inc128_et_hist_v5}
	}
	\caption{ET Histograms of INC16 ... INC128~\label{fig:s9_et_hist2}}
\end{figure}

\newpage

\subsubsection{PT}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC1]{
		\includegraphics[scale=0.43]{sodb9/1_sec_pt_hist_v5.eps}
		\label{fig:s9_inc1_hist_v5}
	}
	\subfigure[PT frequency on INC2]{
		\includegraphics[scale=0.43]{sodb9/2_sec_pt_hist_v5.eps}
		\label{fig:s9_inc2_hist_v5}
	}
	\subfigure[PT frequency on INC4]{
		\includegraphics[scale=0.43]{sodb9/4_sec_pt_hist_v5.eps}
		\label{fig:s9_inc4_hist_v5}
	}
	\subfigure[PT frequency on INC8]{
		\includegraphics[scale=0.43]{sodb9/8_sec_pt_hist_v5.eps}
		\label{fig:s9_inc8_hist_v5}
	}
	\caption{PT Histograms of INC1 ... INC8~\label{fig:s9_pt_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC16]{
		\includegraphics[scale=0.43]{sodb9/16_sec_pt_hist_v5.eps}
		\label{fig:s9_inc16_hist_v5}
	}
	\subfigure[PT frequency on INC32]{
		\includegraphics[scale=0.43]{sodb9/32_sec_pt_hist_v5.eps}
		\label{fig:s9_inc32_hist_v5}
	}
	\subfigure[PT frequency on INC64]{
		\includegraphics[scale=0.43]{sodb9/64_sec_pt_hist_v5.eps}
		\label{fig:s9_inc64_hist_v5}
	}
	\subfigure[PT frequency on INC128]{
		\includegraphics[scale=0.43]{sodb9/128_sec_pt_hist_v5.eps}
		\label{fig:s9_inc128_hist_v5}
	}
	\caption{PT Histograms of INC16 ... INC64~\label{fig:s9_pt_hist2}}
\end{figure}

### 2018-05-01 pt_characterization/real_data/paper.tex
New drafts

@@
 
-\section{Preliminary Experiments}
-
-\subsection{Inter-machine Repeatability Check~\label{sec:diff_machine}}
-
-In this section we check if the use of different machines affects the distribution of execution (process) times on the same program. 
-In other words, we examine {\em inter-machine repeatability}. For this examination, we use a nested for loop program, here termed INC, 
-with different task lengths: 16 seconds, 13 seconds, and 17.2 seconds, each termed {\em INC16}, {\em INC13}, and {\em INC17.2}, respectively.
-
-Figures~\ref{fig:dm_1}, ~\ref{fig:dm_2}, and ~\ref{fig:dm_3} display the process time (PT) histograms on INC16, INC13, and INC17.2 run on our machines (from {\tt sodb8} to {\tt sodb12}), respectively.
-Based on these figures, we conclude that we should stick to one machine  for consistency, and {\tt sodb9} should be chosen in a sense that the data measured there consistently reveals a distinct binormal pattern.
-
-\begin{figure}[h]
-	\centering
-	\subfigure[PT frequency on INC16]{
-		\includegraphics[scale=0.43]{sodb8_INC16_0_dist.eps}
-		\label{fig:s8_inc16_dist}
-	}	
-	\subfigure[PT frequency on INC16]{
-		\includegraphics[scale=0.43]{sodb9_INC16_0_dist.eps}
-		\label{fig:s9_inc16_dist}
-	}
-	\subfigure[PT frequency on INC16]{
-		\includegraphics[scale=0.43]{sodb10_INC16_0_dist.eps}
-		\label{fig:s10_inc16_dist}
-	}	
-	\subfigure[PT frequency on INC16]{
-		\includegraphics[scale=0.43]{sodb12_INC16_0_dist.eps}
-		\label{fig:s12_inc16_dist}
-	}
-	\caption{PT Histograms on INC16~\label{fig:dm_1}}
-\end{figure}
-
-\newpage
-\clearpage
-
-\begin{figure}[h]
-	\centering
-	\subfigure[PT frequency on INC13]{
-		\includegraphics[scale=0.43]{sodb8_INC13_0_dist.eps}
-		\label{fig:s8_inc13_dist}
-	}	
-	\subfigure[PT frequency on INC13]{
-		\includegraphics[scale=0.43]{sodb9_INC13_0_dist.eps}
-		\label{fig:s9_inc13_dist}
-	}
-	\subfigure[PT frequency on INC13]{
-		\includegraphics[scale=0.43]{sodb10_INC13_0_dist.eps}
-		\label{fig:s10_inc13_dist}
-	}	
-	\subfigure[PT frequency on INC13]{
-		\includegraphics[scale=0.43]{sodb12_INC13_0_dist.eps}
-		\label{fig:s12_inc13_dist}
-	}
-	\caption{PT Histograms on INC13~\label{fig:dm_2}}
-\end{figure}
-
-\newpage
-\clearpage
-
-\begin{figure}[h]
-	\centering
-	\subfigure[PT frequency on INC17.2]{
-		\includegraphics[scale=0.43]{sodb8_INC17_2_dist.eps}
-		\label{fig:s8_inc17_2_dist}
-	}	
-	\subfigure[PT frequency on INC17.2]{
-		\includegraphics[scale=0.43]{sodb9_INC17_2_dist.eps}
-		\label{fig:s9_inc17_2_dist}
-	}
-	\subfigure[PT frequency on INC17.2]{
-		\includegraphics[scale=0.43]{sodb10_INC17_2_dist.eps}
-		\label{fig:s10_inc17_2_dist}
-	}	
-	\subfigure[PT frequency on INC17.2]{
-		\includegraphics[scale=0.43]{sodb12_INC17_2_dist.eps}
-		\label{fig:s12_inc17_2_dist}
-	}
-	\caption{PT Histograms on INC17.2~\label{fig:dm_3}}
-\end{figure}
-
-\newpage
-\clearpage
-
-\subsection{One-to-One Comparison between An Insertion Sort \& A Corresponding INC Programs~\label{sec:sort}} 
-This section compares program time histograms 
-of an insertion sort and a nested-for-loop programs (termed INC) with different input sizes.
-The insertion sort program sorts the elements of a given array in non-decreasing order.
-The program repeatedly runs 300 times for a given input size.
-The input size for the program varies from 144K to 344K integer elements, which are randomly generated.
-Note that each sort program over a specific input size is termed SORT{\it x}: for instance, SORT100 indicates the insertion sort program over 100K elements.
-An INC program's task length is correspondingly determined by the program time of an SORT program.
-In Figures~\ref{fig:sort1},~\ref{fig:sort2}, and~\ref{fig:sort3}
-we perform a match on an SORT program and its corresponding INC program.
-
-%
-%Figures~\ref{fig:sort1} and ~\ref{fig:sort2} exhibit 
-%histograms of the execution times measured on the same insertion sort program as 
-%the input size grows from 100K to 1,160K elements.% by the steps of 2x. 
-%Note that we used one standard deviation in Figure~\ref{fig:sort400_dist} as
-%a couple of outliers, which were not eliminated by the original protocol, 
-%resulted in disturbing the rendering of a clean distribution.
-
-\begin{figure}[h]
-	\centering
-	\subfigure[PT frequency on SORT144 on {\tt sodb8}]{
-		\includegraphics[scale=0.43]{sort144_dist.eps}
-		\label{fig:sort144_dist}
-	}	
-	\subfigure[PT frequency on INC8.8 on {\tt sodb10}]{
-		\includegraphics[scale=0.43]{inc8.8_dist.eps}
-		\label{fig:inc8.8_dist}
-	}
-	\subfigure[PT frequency on SORT177 on {\tt sodb8}]{
-		\includegraphics[scale=0.43]{sort177_dist.eps}
-		\label{fig:sort177_dist}
-	}	
-	\subfigure[PT frequency on INC13 on {\tt sodb10}]{
-		\includegraphics[scale=0.43]{inc13.0_dist.eps}
-		\label{fig:inc13_dist}
-	}
-	\caption{PT Histograms I~\label{fig:sort1}}
-\end{figure}
-
-\clearpage
-\pagebreak
-
-\begin{figure}[h]
-	\centering
-	\subfigure[PT frequency on SORT144 on {\tt sodb8}]{
-		\includegraphics[scale=0.43]{sort244_dist.eps}
-		\label{fig:sort144_dist}
-	}	
-	\subfigure[PT frequency on INC25.8 on {\tt sodb10}]{
-		\includegraphics[scale=0.43]{inc25.8_dist.eps}
-		\label{fig:inc25.8_dist}
-	}
-	\subfigure[PT frequency on SORT288  on {\tt sodb8}]{
-		\includegraphics[scale=0.43]{sort177_dist.eps}
-		\label{fig:sort177_dist}
-	}	
-	\subfigure[PT frequency on INC35 on {\tt sodb10}]{
-		\includegraphics[scale=0.43]{inc13.0_dist.eps}
-		\label{fig:inc35_dist}
-	}
-	\caption{PT Histogram Comparison II~\label{fig:sort2}}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\subfigure[PT frequency on SORT344  on {\tt sodb8}]{
-		\includegraphics[scale=0.43]{sort344_dist.eps}
-		\label{fig:sort344_dist}
-	}	
-	\subfigure[PT frequency on INC51.5 on {\tt sodb10}]{
-		\includegraphics[scale=0.43]{inc51.5_dist.eps}
-		\label{fig:inc51.5_dist}
-	}
-	\caption{PT Histogram Comparison III~\label{fig:sort3}}
-\end{figure}
-
 \clearpage
@@
 \section{Appendix}
+
+\subsection{One-to-One Comparison between An Insertion Sort \& A Corresponding INC Programs~\label{sec:sort}} 
+This section compares program time histograms 
+of an insertion sort and a nested-for-loop programs (termed INC) with different input sizes.
+The insertion sort program sorts the elements of a given array in non-decreasing order.
+The program repeatedly runs 300 times for a given input size.
+The input size for the program varies from 144K to 344K integer elements, which are randomly generated.
+Note that each sort program over a specific input size is termed SORT{\it x}: for instance, SORT100 indicates the insertion sort program over 100K elements.
+An INC program's task length is correspondingly determined by the program time of an SORT program.
+In Figures~\ref{fig:sort1},~\ref{fig:sort2}, and~\ref{fig:sort3}
+we perform a match on an SORT program and its corresponding INC program.
+
+%
+%Figures~\ref{fig:sort1} and ~\ref{fig:sort2} exhibit 
+%histograms of the execution times measured on the same insertion sort program as 
+%the input size grows from 100K to 1,160K elements.% by the steps of 2x. 
+%Note that we used one standard deviation in Figure~\ref{fig:sort400_dist} as
+%a couple of outliers, which were not eliminated by the original protocol, 
+%resulted in disturbing the rendering of a clean distribution.
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on SORT144 on {\tt sodb8}]{
+		\includegraphics[scale=0.43]{sort144_dist.eps}
+		\label{fig:sort144_dist}
+	}	
+	\subfigure[PT frequency on INC8.8 on {\tt sodb10}]{
+		\includegraphics[scale=0.43]{inc8.8_dist.eps}
+		\label{fig:inc8.8_dist}
+	}
+	\subfigure[PT frequency on SORT177 on {\tt sodb8}]{
+		\includegraphics[scale=0.43]{sort177_dist.eps}
+		\label{fig:sort177_dist}
+	}	
+	\subfigure[PT frequency on INC13 on {\tt sodb10}]{
+		\includegraphics[scale=0.43]{inc13.0_dist.eps}
+		\label{fig:inc13_dist}
+	}
+	\caption{PT Histograms I~\label{fig:sort1}}
+\end{figure}
+
+\clearpage
+\pagebreak
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on SORT244 on {\tt sodb8}]{
+		\includegraphics[scale=0.43]{sort244_dist.eps}
+		\label{fig:sort244_dist}
+	}	
+	\subfigure[PT frequency on INC25.8 on {\tt sodb10}]{
+		\includegraphics[scale=0.43]{inc25.8_dist.eps}
+		\label{fig:inc25.8_dist}
+	}
+	\subfigure[PT frequency on SORT288  on {\tt sodb8}]{
+		\includegraphics[scale=0.43]{sort288_dist.eps}
+		\label{fig:sort288_dist}
+	}	
+	\subfigure[PT frequency on INC35 on {\tt sodb10}]{
+		\includegraphics[scale=0.43]{inc35.0_dist.eps}
+		\label{fig:inc35_dist}
+	}
+	\caption{PT Histogram Comparison II~\label{fig:sort2}}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\subfigure[PT frequency on SORT344  on {\tt sodb8}]{
+		\includegraphics[scale=0.43]{sort344_dist.eps}
+		\label{fig:sort344_dist}
+	}	
+	\subfigure[PT frequency on INC51.5 on {\tt sodb10}]{
+		\includegraphics[scale=0.43]{inc51.5_dist.eps}
+		\label{fig:inc51.5_dist}
+	}
+	\caption{PT Histogram Comparison III~\label{fig:sort3}}
+\end{figure}
 

### 2018-05-01 pt_characterization/machine_cons_check/newer_exp/inc_hist.r

@@ -1,0 +1,564 @@
+### sodb8-INC16
+numSamples <- 1000
+x = read.csv(file="sodb8.INC16.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sodb8_INC16_0_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 300 
+title <- paste("PT frequency on sodb8-INC16.0 (sample size: ",numSamples,")")
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main=title, 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### sodb9-INC16
+numSamples <- 1000
+x = read.csv(file="sodb9.INC16.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sodb9_INC16_0_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
[…]

### 2018-05-01 pt_characterization/real_data.new/hist.r

@@ -1,0 +1,711 @@
+### SORT25
+numSamples <- 40
+x = read.csv(file="SORT25.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort25_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 25
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on SORT25 with 40 samples', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### SORT50
+numSamples <- 40
+x = read.csv(file="SORT50.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort50_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
[…]

### 2018-05-01 pt_characterization/real_data.new/matc_hist.r

@@ -1,0 +1,423 @@
+### MATC250
+numSamples <- 40
+x = read.csv(file="MATC250.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("matc250_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 25
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on MATC250 with 40 samples', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### MATC500
+numSamples <- 40
+x = read.csv(file="MATC500.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("matc500_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
[…]

### 2018-05-01 pt_characterization/real_data.new/matmul.c

@@ -1,0 +1,428 @@
+#include <stdio.h>    /* standard in and output*/
+#include <sys/socket.h> /* for socket() and socket functions*/
+#include <arpa/inet.h>  /* for sockaddr_in and inet_ntoa() */
+#include <stdlib.h>   
+#include <string.h>    
+#include <unistd.h>     /* for close() */
+#include <time.h> 
+#include <limits.h> 
+#define PRINT 0
+#define PORT_NUM 2400
+//#define COEF 1650000000 // 16B
+#define COEF 1460000000 // 16B
+#define NUM_ELE1 20800
+#define NUM_ELE64 174000
+#define NUM_ELE128 241000
+//#define MAT_SZ 10000
+#define COLUMN_MAJOR 1
+#define ROW_MAJOR 2
+//#define COL_COEF 1460000000 
+//#define COL_COEF 1455000000 
+//#define COL_COEF 1450000000 
+//#define ROW_COEF 1457000000 
+#define MAT_COEF 2000000
+#define SEED 10000
+
+char c;
+
+/* for sort */
+int* iarr = NULL;
+int* iarr_unsorted = NULL;
+int numCnt = 0;
+int MAX_NUM = 1000;//10K
+/* for matrix */
+int MAT_SZ = 0;
+int** matrixA = NULL;
+int** matrixB = NULL;
+int** matrixC = NULL;
+
+/* for sort */
+inline void reset(int numCnt){
+	memset(iarr, 0, sizeof(int)*numCnt);
+	memcpy(iarr, iarr_unsorted, sizeof(int)*numCnt);
+}
+
+inline int sort_init(int task_length){
+	/*if(task_length == 1) numCnt = NUM_ELE1;
+	else if(task_length == 64) numCnt = NUM_ELE64;
+	else if(task_length == 128) numCnt = NUM_ELE128;*/
+
+        /*if(task_length == 100)    numCnt = 100000; // 100k
+        else if(task_length == 200)    numCnt = 200000; // 200k
+        else if(task_length == 400)    numCnt = 400000;// 400k
+        else if(task_length == 800)    numCnt = 800000;// 800k*/
+	numCnt = task_length;
+
+	// numCnt should be translated from a task length
+	iarr = (int*)malloc(sizeof(int)*numCnt);
+	iarr_unsorted = (int*)malloc(sizeof(int)*numCnt);
+
+	int i;
+	time_t t;
+	/* Intializes random number generator */
[…]

### 2018-05-01 pt_characterization/real_data.new/matr_hist.r

@@ -1,0 +1,381 @@
+### MATR1000
+numSamples <- 40
+x = read.csv(file="MATR250.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("matr250_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+#ymax <- ceiling(ymax/100)*100
+ymax <- 20
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on MATR250 with 40 samples', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+x = read.csv(file="MATR500.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("matr500_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
[…]

### 2018-05-01 pt_characterization/real_data.new/sort_hist.r

@@ -1,0 +1,173 @@
+### SORT144
+numSamples <- 300
+x = read.csv(file="SORT144.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort144_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 150
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main='PT frequency on SORT144 (sample size: 300)', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### new SORT177
+numSamples <- 300
+x = read.csv(file="SORT177.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("sort177_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
[…]

### 2018-05-04 pt_characterization/machine_cons_check/paper.tex
Ricks comment applied

@@
 Machine & Task Length & Description & Time Length\\ \hline
-{\tt sodb8} (plugged into \hbox{bottom} {\em left} power strip) & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
-{\tt sodb9}  (plugged into \hbox{bottom} {\em left} power strip)  &  INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
-{\tt sodb10}  (plugged into \hbox{bottom} {\em left} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
-{\tt sodb12}  (plugged into \hbox{bottom} {\em left} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb8} (plugged into {\em bottom left} power strip) & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb9}  (plugged into {\em the upper left} power strip)  &  INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb10}  (plugged into {\em the upper left} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb12}  (plugged into {\em the upper left} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
 \end{tabular}
@@
 
-\newpage
+%\newpage
 

### 2018-05-04 pt_characterization/machine_cons_check/post_exp.tex
Ricks comment applied

@@
 Machine & Task Length & Description & Time Length\\ \hline
-{\tt sodb8} (plugged into \hbox{bottom} {\em left} power strip) &  INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
-{\tt sodb9}  (plugged into \hbox{bottom} {\em left} power strip)  &  IINC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
-{\tt sodb10}  (plugged into \hbox{bottom} {\em left} power strip)  & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
-{\tt sodb12}  (plugged into \hbox{bottom} {\em left} power strip)  & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
+{\tt sodb8} (plugged into {\em bottom left} power strip) &  INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
+{\tt sodb9}  (plugged into {\em the upper left} power strip) &  IINC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
+{\tt sodb10} (plugged into {\em the upper left} power strip)  & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
+{\tt sodb12} (plugged into {\em the upper left} power strip) & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
 \end{tabular}
@@
 
+\subsection{Intra-Machine Homogeneity~\label{sec:intra_machine}}
+
+A short description of the runs we used for this experiment are exhibited in Table~\ref{tab:exp_notes3}.
+
+%\paragraph{Experimental Notes:} In the experiments 
+%we used only {\tt sodb9}. A short summary of the experiments is shown in Table~\ref{tab:exp_notes3}.
+%\vspace{-.3in}
+%% done
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
+Machine & Task Length & Description & Time Length\\ \hline
+{\tt sodb9}  (plugged into {\em the upper left} power strip) &  INC1$\sim$INC128 & Runs of 1,000 samples & 2018-04-28 $\sim$2018-04-30\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Detailed description of INC data used for histograms\label{tab:exp_notes3}}
+\end{table}
+
+%\newpage
+%\clearpage
+
+\begin{figure}[H]
+	\centering
+	\subfigure[PT frequency on INC1 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{new_INC_data_s9/s9_INC1_dist.eps}
+		\label{fig:s9_inc1_dist}
+	}
+	\subfigure[PT frequency on INC2 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{new_INC_data_s9/s9_INC2_dist.eps}
+		\label{fig:s9_inc2_dist}
+	}
+	\subfigure[PT frequency on INC4 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{new_INC_data_s9/s9_INC4_dist.eps}
+		\label{fig:s9_inc4_dist}
+	}
+	\subfigure[PT frequency on INC8 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{new_INC_data_s9/s9_INC8_dist.eps}
+		\label{fig:s9_inc8_dist}
+	}
+	\caption{PT Histograms on INC1$\sim$INC8\label{fig:new_data1}}
+\end{figure}
+
+\newpage
+\clearpage
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on INC16 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{new_INC_data_s9/s9_INC16_dist.eps}
+		\label{fig:s9_inc16_dist2}
+	}
+	\subfigure[PT frequency on INC32 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{new_INC_data_s9/s9_INC32_dist.eps}
+		\label{fig:s9_inc32_dist}
+	}
+	\subfigure[PT frequency on INC64 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{new_INC_data_s9/s9_INC64_dist.eps}
+		\label{fig:s9_inc64_dist}
+	}
+	\subfigure[PT frequency on INC128 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{new_INC_data_s9/s9_INC128_dist.eps}
+		\label{fig:s9_inc128_dist}
+	}
+	\caption{PT Histograms on INC16$\sim$INC128\label{fig:new_data2}}
+\end{figure}
+
+\newpage
+\clearpage
+
 \subsection{Investigation of Daemons' Influence on Program Time Distribution Regarding Machine Dependence~\label{sec:daemon_impact}}

### 2018-05-04 pt_characterization/machine_cons_check/sodb10/res.tex
Ricks comment applied

@@
 	\centering
-	\subfigure[ET frequency on INC1]{
+	\subfigure[ET frequency on INC1 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/1_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC2]{
+	\subfigure[ET frequency on INC2 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/2_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC4]{
+	\subfigure[ET frequency on INC4 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/4_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC8]{
+	\subfigure[ET frequency on INC8 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/8_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[ET frequency on INC16]{
+	\subfigure[ET frequency on INC16 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/16_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC32]{
+	\subfigure[ET frequency on INC32 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/32_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC64]{
+	\subfigure[ET frequency on INC64 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/64_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC128]{
+	\subfigure[ET frequency on INC128 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/128_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC1]{
+	\subfigure[PT frequency on INC1 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/1_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC2]{
+	\subfigure[PT frequency on INC2 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/2_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC4]{
+	\subfigure[PT frequency on INC4 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/4_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC8]{
+	\subfigure[PT frequency on INC8 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/8_sec_pt_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC16]{
+	\subfigure[PT frequency on INC16 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/16_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC32]{
+	\subfigure[PT frequency on INC32 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/32_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC64]{
+	\subfigure[PT frequency on INC64 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/64_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC128]{
+	\subfigure[PT frequency on INC128 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{sodb10/128_sec_pt_hist_v5.eps}

### 2018-05-04 pt_characterization/machine_cons_check/sodb12/res.tex
Ricks comment applied

@@
 	\centering
-	\subfigure[ET frequency on INC1]{
+	\subfigure[ET frequency on INC1 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/1_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC2]{
+	\subfigure[ET frequency on INC2 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/2_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC4]{
+	\subfigure[ET frequency on INC4 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/4_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC8]{
+	\subfigure[ET frequency on INC8 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/8_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[ET frequency on INC16]{
+	\subfigure[ET frequency on INC16 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/16_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC32]{
+	\subfigure[ET frequency on INC32 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/32_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC64]{
+	\subfigure[ET frequency on INC64 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/64_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC128]{
+	\subfigure[ET frequency on INC128 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/128_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC1]{
+	\subfigure[PT frequency on INC1 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/1_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC2]{
+	\subfigure[PT frequency on INC2 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/2_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC4]{
+	\subfigure[PT frequency on INC4 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/4_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC8]{
+	\subfigure[PT frequency on INC8 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/8_sec_pt_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC16]{
+	\subfigure[PT frequency on INC16 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/16_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC32]{
+	\subfigure[PT frequency on INC32 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/32_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC64]{
+	\subfigure[PT frequency on INC64 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/64_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC128]{
+	\subfigure[PT frequency on INC128 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{sodb12/128_sec_pt_hist_v5.eps}

### 2018-05-04 pt_characterization/machine_cons_check/sodb8/res.tex
Ricks comment applied

@@
 	\centering
-	\subfigure[ET frequency on INC1]{
+	\subfigure[ET frequency on INC1 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/1_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC2]{
+	\subfigure[ET frequency on INC2 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/2_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC4]{
+	\subfigure[ET frequency on INC4 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/4_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC8]{
+	\subfigure[ET frequency on INC8 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/8_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[ET frequency on INC16]{
+	\subfigure[ET frequency on INC16 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/16_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC32]{
+	\subfigure[ET frequency on INC32 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/32_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC64]{
+	\subfigure[ET frequency on INC64 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/64_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC128]{
+	\subfigure[ET frequency on INC128 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/128_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC1]{
+	\subfigure[PT frequency on INC1 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/1_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC2]{
+	\subfigure[PT frequency on INC2 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/2_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC4]{
+	\subfigure[PT frequency on INC4 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/4_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC8]{
+	\subfigure[PT frequency on INC8 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/8_sec_pt_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC16]{
+	\subfigure[PT frequency on INC16 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/16_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC32]{
+	\subfigure[PT frequency on INC32 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/32_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC64]{
+	\subfigure[PT frequency on INC64 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/64_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC128]{
+	\subfigure[PT frequency on INC128 on {\tt sodb8}]{
 		\includegraphics[scale=0.43]{sodb8/128_sec_pt_hist_v5.eps}

### 2018-05-04 pt_characterization/machine_cons_check/sodb9/res.tex
Ricks comment applied

@@
 	\centering
-	\subfigure[ET frequency on INC1]{
+	\subfigure[ET frequency on INC1 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/1_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC2]{
+	\subfigure[ET frequency on INC2 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/2_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC4]{
+	\subfigure[ET frequency on INC4 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/4_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC8]{
+	\subfigure[ET frequency on INC8 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/8_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[ET frequency on INC16]{
+	\subfigure[ET frequency on INC16 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/16_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC32]{
+	\subfigure[ET frequency on INC32 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/32_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC64]{
+	\subfigure[ET frequency on INC64 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/64_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC128]{
+	\subfigure[ET frequency on INC128 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/128_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC1]{
+	\subfigure[PT frequency on INC1 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/1_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC2]{
+	\subfigure[PT frequency on INC2 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/2_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC4]{
+	\subfigure[PT frequency on INC4 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/4_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC8]{
+	\subfigure[PT frequency on INC8 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/8_sec_pt_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC16]{
+	\subfigure[PT frequency on INC16 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/16_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC32]{
+	\subfigure[PT frequency on INC32 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/32_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC64]{
+	\subfigure[PT frequency on INC64 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/64_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC128]{
+	\subfigure[PT frequency on INC128 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{sodb9/128_sec_pt_hist_v5.eps}

### 2018-05-04 pt_characterization/real_data/paper.tex
realistic programs
Ricks comment applied

@@
 
-\clearpage
-\pagebreak
+%\clearpage
+%\pagebreak
 
@@
 We then exhibit different kinds of histograms in the subsequent sections.
+
+\paragraph{Experimental Notes:} In the experiments 
+we also used four nodes ({\tt sodb8}, {\tt sodb9}, {\tt sodb10}, and {\tt sodb12} ) in the same cluster. 
+A short summary of the experiments is shown in Table~\ref{tab:exp_notes}.
+
+%% done
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
+Machine & Task Length & Description & Time Length\\ \hline
+%{\tt sodb8}  (plugged into {\em bottom left} power strip) & SORT100, SORT200, SORT400, SORT800 & Runs of 300 samples & 2018-??-?? $\sim$2018-??-??\\ \hline
+{\tt sodb9}  (plugged into {\em the upper left} power strip) & SORT100, SORT200, SORT400, SORT800 & Runs of 300 samples & 2018-04-09 $\sim$2018-04-12\\ \hline
+{\tt sodb10} (plugged into {\em the upper left} power strip)  & MATC1K, MATC2K, MATC4K, MATC8K & Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
+{\tt sodb12} (plugged into {\em the upper left} power strip) & MATR1K, MATR2K, MATR4K, MATR8K& Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Detailed description of INC data used for histograms\label{tab:exp_notes}}
+\end{table}
 
@@
 for instance, SORT100 indicates the insertion sort program over 100K elements. 
-(This experiment was performed on {\tt sodb9}.)
+This experiment was performed on {\tt sodb9} plugged into {\em the upper left} power strip.
 
@@
 
+\clearpage
+\pagebreak
+
+\subsection{Matrix Multiplication~\label{sec:mm}}
+
+This section shows a series of histograms of 
+the execution times of an matrix multiplication program. 
+We used the same sample size for each input size of this program: i.e., 40 iterations. 
+For simplicity, we used two square matrices for performing their multiplication in the program.  
+We also varied the input sizes of each of the two matrices: 
+from 1K$\times$1K to 8K$\times$8K integer elements that are also randomly generated. 
+Note that each matrix multiplication program for a specific size is called MAT{\it xyyyy}, 
+where {\it x} indicates which major, specifically {\em column} vs. {\em row}, is used, and {\it yyyy}, how large a given matrix is. 
+For instance, MATC1000 represents a matrix multiplication program 
+in column major over two square matrices having 1,000 integer (random) 
+elements in a row (and a column).
+
+\pagebreak
+
+\subsubsection{Column Major}
+
+Figure~\ref{fig:matc} shows a series of 
+histograms of the execution times measured on 
+the same matrix multiplication program in column major 
+as the input sizes grows. The program was run on {\tt sodb10} plugged into {\em the upper left} power strip.
+
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATC250]{
+%		\includegraphics[scale=0.43]{matc250_dist.eps}
+%		\label{fig:matc250_dist}
+%	}
+%	\subfigure[PT frequency on MATC500]{
+%		\includegraphics[scale=0.43]{matc500_dist.eps}
+%		\label{fig:matc500_dist}
+%	}
+%	\subfigure[PT frequency on MATC1024]{
+%		\includegraphics[scale=0.43]{matc1024_dist2.eps}
+%		\label{fig:matc1024_dist}
+%	}
+%	\subfigure[PT frequency on MATC1440]{
+%		\includegraphics[scale=0.43]{matc1440_dist.eps}
+%		\label{fig:matc1440_dist}
+%	}
+%	\caption{PT Histograms of MATC250 ... MATC1440~\label{fig:matc}}
+%\end{figure}
+%
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATC2048]{
+%		\includegraphics[scale=0.43]{matc2048_dist2.eps}
+%		\label{fig:matc2048_dist}
+%	}
+%	\subfigure[PT frequency on MATC2880]{
+%		\includegraphics[scale=0.43]{matc2880_dist.eps}
+%		\label{fig:matc2880_dist}
+%	}	
+%	\subfigure[PT frequency on MATC4096]{
+%		\includegraphics[scale=0.43]{matc4096_dist2.eps}
+%		\label{fig:matc4096_dist}
+%	}	
+%	\subfigure[PT frequency on MATC5760]{
+%		\includegraphics[scale=0.43]{matc5760_dist.eps}
+%		\label{fig:matc5760_dist}
+%	}	
+%	\caption{PT Histograms of MATC2048 ... MATC5760~\label{fig:matc2}}
+%\end{figure}
+%
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATC8192]{
+%		\includegraphics[scale=0.43]{matc8192_dist2.eps}
+%		\label{fig:matc8192_dist}
+%	}
+%	\subfigure[PT frequency on MATC11520]{
+%		\includegraphics[scale=0.43]{matc11520_dist.eps}
+%		\label{fig:matc11520_dist}
+%	}
+%	\subfigure[PT frequency on MATC16000]{
+%		\includegraphics[scale=0.43]{matc16000_dist.eps}
+%		\label{fig:matc16000_dist}
+%	}
+%	\subfigure[PT frequency on MATC23040]{
+%		\includegraphics[scale=0.43]{matc23040_dist.eps}
+%		\label{fig:matc23040_dist}
+%	}
+%	\caption{PT Histograms of MATC8192 ... MATC23040~\label{fig:matc3}}
+%\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on MATC1K on {\tt sodb10}]{
+		\includegraphics[scale=0.43]{matc1024_dist.eps}
+		\label{fig:matc1k_dist}
+	}
+	\subfigure[PT frequency on MATC2K on {\tt sodb10}]{
+		\includegraphics[scale=0.43]{matc2048_dist.eps}
+		\label{fig:matc2k_dist}
+	}
+	\subfigure[PT frequency on MATC4K on {\tt sodb10}]{
+		\includegraphics[scale=0.43]{matc4096_dist.eps}
+		\label{fig:matc4k_dist}
+	}
+	\subfigure[PT frequency on MATC8K on {\tt sodb10}]{
+		\includegraphics[scale=0.43]{matc8192_dist.eps}
+		\label{fig:matc8k_dist}
+	}
+	\caption{PT Histograms of MATC1024 ... MATC8192~\label{fig:matc}}
+\end{figure}
+
+\clearpage
+\pagebreak
+
+\subsubsection{Row Major}
+
+Figure~\ref{fig:matc} shows a series of 
+histograms of the execution times measured on 
+the same matrix multiplication program in row major as the input sizes grows. 
+The program was run on {\tt sodb12} plugged into {\em the upper left} power strip.
+
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATR250]{
+%		\includegraphics[scale=0.43]{matr250_dist.eps}
+%		\label{fig:matr250_dist}
+%	}
+%	\subfigure[PT frequency on MATR500]{
+%		\includegraphics[scale=0.43]{matr500_dist.eps}
+%		\label{fig:matr500_dist}
+%	}
+%	\subfigure[PT frequency on MATR1024]{
+%		\includegraphics[scale=0.43]{matr1024_dist2.eps}
+%		\label{fig:matr1024_dist}
+%	}
+%	\subfigure[PT frequency on MATR1440]{
+%		\includegraphics[scale=0.43]{matr1440_dist.eps}
+%		\label{fig:matr1440_dist}
+%	}
+%	\caption{PT Histograms of MATR250 ... MATR1440~\label{fig:matr}}
+%\end{figure}
+
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATR2048]{
+%		\includegraphics[scale=0.43]{matr2048_dist2.eps}
+%		\label{fig:matr2048_dist}
+%	}
+%	\subfigure[PT frequency on MATR2880]{
+%		\includegraphics[scale=0.43]{matr2880_dist.eps}
+%		\label{fig:matr2880_dist}
+%	}	
+%	\subfigure[PT frequency on MATR4096]{
+%		\includegraphics[scale=0.43]{matr4096_dist2.eps}
+%		\label{fig:matr4096_dist}
+%	}	
+%	\subfigure[PT frequency on MATR5760]{
+%		\includegraphics[scale=0.43]{matr5760_dist.eps}
+%		\label{fig:matr5760_dist}
+%	}	
+%	\caption{PT Histograms of MATR2048 ... MATR5760~\label{fig:matr2}}
+%\end{figure}
+%
+%\begin{figure}[h]
+%	\centering
+%	\subfigure[PT frequency on MATR8192]{
+%		\includegraphics[scale=0.43]{matr8192_dist2.eps}
+%		\label{fig:matr8192_dist}
+%	}
+%	\subfigure[PT frequency on MATR11520]{
+%		\includegraphics[scale=0.43]{matr11520_dist.eps}
+%		\label{fig:matr11520_dist}
+%	}	
+%	\subfigure[PT frequency on MATR16000]{
+%		\includegraphics[scale=0.43]{matr16000_dist.eps}
+%		\label{fig:matr16000_dist}
+%	}
+%	\subfigure[PT frequency on MATR23040]{
+%		\includegraphics[scale=0.43]{matr23040_dist.eps}
+%		\label{fig:matr23040_dist}
+%	}
+%	\caption{PT Histograms of MATR8192 ... MATR23040~\label{fig:matr3}}
+%\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on MATR1K on {\tt sodb12}]{
+		\includegraphics[scale=0.43]{matr1024_dist.eps}
+		\label{fig:matr1k_dist}
+	}
+	\subfigure[PT frequency on MATR2K on {\tt sodb12}]{
+		\includegraphics[scale=0.43]{matr2048_dist.eps}
+		\label{fig:matr2k_dist}
+	}
+	\subfigure[PT frequency on MATR4K on {\tt sodb12}]{
+		\includegraphics[scale=0.43]{matr4096_dist.eps}
+		\label{fig:matr4k_dist}
+	}
+	\subfigure[PT frequency on MATR8K on {\tt sodb12}]{
+		\includegraphics[scale=0.43]{matr8192_dist.eps}
+		\label{fig:matr8k_dist}
+	}
+	\caption{PT Histograms of MATR1K ... MATR8K~\label{fig:matr}}
+\end{figure}
+
 \newpage
@@
 \section{Appendix}
+
+\subsection{Insertion Sort on {\tt sodb8}~\label{sec:add_new}} 
+This section shows a series of histograms of an insertion sorting program that 
+sorts the elements of a given array in non-decreasing order. 
+The program repeatedly runs 300 times for a given input size. 
+The input size for the program 
+varies from 100,000 to 1,160,000 integer elements, which are randomly generated. 
+Note that each sort program over a specific input size is termed SORT{\it x}: 
+for instance, SORT100 indicates the insertion sort program over 100K elements. 
+This experiment was performed on {\tt sodb8} plugged into {\em bottom left} power strip.
+
+\begin{figure}[h]
+	\centering
+%	\subfigure[PT frequency on SORT25]{
+%		\includegraphics[scale=0.43]{sort25_dist.eps}
+%		\label{fig:sort25_dist}
+%	}
+%	\subfigure[PT frequency on SORT50]{
+%		\includegraphics[scale=0.43]{sort50_dist.eps}
+%		\label{fig:sort50_dist}
+%	}
+	\subfigure[PT frequency on SORT100 on {\tt sodb8}]{
+		\includegraphics[scale=0.43]{sort100_dist.eps}
+		\label{fig:sort100_dist}
+	}
+	\subfigure[PT frequency on SORT400 on {\tt sodb8}]{
+		\includegraphics[scale=0.43]{sort400_dist.eps}
+		\label{fig:sort400_dist}
+	}
+	\subfigure[PT frequency on SORT580 on {\tt sodb8}]{
+		\includegraphics[scale=0.43]{sort580_dist.eps}
+		\label{fig:sort588_dist}
+	}
+	\caption{PT Histograms of SORT100 ... SORT580~\label{fig:n_sort1}}
+\end{figure}
+
+\clearpage
+\pagebreak
+
+\begin{figure}[h]
+	\centering
+	\subfigure[PT frequency on SORT800 on {\tt sodb8}]{
+		\includegraphics[scale=0.43]{sort800_dist.eps}
+		\label{fig:sort800_dist}
+	}
+	\subfigure[PT frequency on SORT1000 on {\tt sodb8}]{
+		\includegraphics[scale=0.43]{sort1000_dist.eps}
+		\label{fig:sort1000_dist}
+	}
+	\subfigure[PT frequency on SORT1160 on {\tt sodb8}]{
+		\includegraphics[scale=0.43]{sort1160_dist.eps}
+		\label{fig:sort1160_dist}
+	}
+	\subfigure[PT frequency on SORT2320 on {\tt sodb8}]{
+		\includegraphics[scale=0.43]{sort2320_dist.eps}
+		\label{fig:sort2320_dist}
+	}
+	\caption{PT Histograms of SORT800 ... SORT2320~\label{fig:n_sort2}}
+\end{figure}
+
+\clearpage
+\pagebreak
 
@@
 we perform a match on an SORT program and its corresponding INC program. 
+{\tt sodb8} was plugged into {\em bottom left} power strip 
+while {\tt sodb10} plugged into {\em the upper left} power strip.
 
@@
 	\caption{PT Histogram Comparison III~\label{fig:sort3}}
-\end{figure}
-
-\subsection{Insertion Sort on {\tt sodb8}~\label{sec:add_new}} 
-This section shows a series of histograms of an insertion sorting program that 
-sorts the elements of a given array in non-decreasing order. 
-The program repeatedly runs 300 times for a given input size. 
-The input size for the program 
-varies from 100,000 to 1,160,000 integer elements, which are randomly generated. 
-Note that each sort program over a specific input size is termed SORT{\it x}: 
-for instance, SORT100 indicates the insertion sort program over 100K elements. 
-This experiment was performed on {\tt sodb8}.
-
-\begin{figure}[h]
-	\centering
-%	\subfigure[PT frequency on SORT25]{
-%		\includegraphics[scale=0.43]{sort25_dist.eps}
-%		\label{fig:sort25_dist}
-%	}
-%	\subfigure[PT frequency on SORT50]{
-%		\includegraphics[scale=0.43]{sort50_dist.eps}
-%		\label{fig:sort50_dist}
-%	}
-	\subfigure[PT frequency on SORT100 on {\tt sodb8}]{
-		\includegraphics[scale=0.43]{sort100_dist.eps}
-		\label{fig:sort100_dist}
-	}
-	\subfigure[PT frequency on SORT400 on {\tt sodb8}]{
-		\includegraphics[scale=0.43]{sort400_dist.eps}
-		\label{fig:sort400_dist}
-	}
-	\subfigure[PT frequency on SORT580 on {\tt sodb8}]{
-		\includegraphics[scale=0.43]{sort580_dist.eps}
-		\label{fig:sort588_dist}
-	}
-	\caption{PT Histograms of SORT100 ... SORT580~\label{fig:n_sort1}}
-\end{figure}
-
-\clearpage
-\pagebreak
-
-\begin{figure}[h]
-	\centering
-	\subfigure[PT frequency on SORT800 on {\tt sodb8}]{
-		\includegraphics[scale=0.43]{sort800_dist.eps}
-		\label{fig:sort800_dist}
-	}
-	\subfigure[PT frequency on SORT1000 on {\tt sodb8}]{
-		\includegraphics[scale=0.43]{sort1000_dist.eps}
-		\label{fig:sort1000_dist}
-	}
-	\subfigure[PT frequency on SORT1160 on {\tt sodb8}]{
-		\includegraphics[scale=0.43]{sort1160_dist.eps}
-		\label{fig:sort1160_dist}
-	}
-	\subfigure[PT frequency on SORT2320 on {\tt sodb8}]{
-		\includegraphics[scale=0.43]{sort2320_dist.eps}
-		\label{fig:sort2320_dist}
-	}
-	\caption{PT Histograms of SORT800 ... SORT2320~\label{fig:n_sort2}}
 \end{figure}
@@
 
-\clearpage
-\pagebreak
-
-\subsection{Matrix Multiplication~\label{sec:mm}}
-
-This section shows a series of histograms of 
-the execution times of an matrix multiplication program. 
-We used the same sample size for each input size of this program: i.e., 40 iterations. 
-For simplicity, we used two square matrices for performing their multiplication in the program.  
-We also varied the input sizes of each of the two matrices: 
-from 1K$\times$1K to 8K$\times$8K integer elements that are also randomly generated. 
-Note that each matrix multiplication program for a specific size is called MAT{\it xyyyy}, 
-where {\it x} indicates which major, specifically {\em column} vs. {\em row}, is used, and {\it yyyy}, how large a given matrix is. 
-For instance, MATC1000 represents a matrix multiplication program 
-in column major over two square matrices having 1,000 integer (random) 
-elements in a row (and a column).
-
-\pagebreak
-
-\subsubsection{Column Major}
-
-Figure~\ref{fig:matc} shows a series of 
-histograms of the execution times measured on 
-the same matrix multiplication program in column major 
-as the input sizes grows. The program was run on {\tt sodb10}.
-
-%\begin{figure}[h]
-%	\centering
-%	\subfigure[PT frequency on MATC250]{
-%		\includegraphics[scale=0.43]{matc250_dist.eps}
-%		\label{fig:matc250_dist}
-%	}
-%	\subfigure[PT frequency on MATC500]{
-%		\includegraphics[scale=0.43]{matc500_dist.eps}
-%		\label{fig:matc500_dist}
-%	}
-%	\subfigure[PT frequency on MATC1024]{
-%		\includegraphics[scale=0.43]{matc1024_dist2.eps}
-%		\label{fig:matc1024_dist}
-%	}
-%	\subfigure[PT frequency on MATC1440]{
-%		\includegraphics[scale=0.43]{matc1440_dist.eps}
-%		\label{fig:matc1440_dist}
-%	}
-%	\caption{PT Histograms of MATC250 ... MATC1440~\label{fig:matc}}
-%\end{figure}
-%
-%\begin{figure}[h]
-%	\centering
-%	\subfigure[PT frequency on MATC2048]{
-%		\includegraphics[scale=0.43]{matc2048_dist2.eps}
-%		\label{fig:matc2048_dist}
-%	}
-%	\subfigure[PT frequency on MATC2880]{
-%		\includegraphics[scale=0.43]{matc2880_dist.eps}
-%		\label{fig:matc2880_dist}
-%	}	
-%	\subfigure[PT frequency on MATC4096]{
-%		\includegraphics[scale=0.43]{matc4096_dist2.eps}
-%		\label{fig:matc4096_dist}
-%	}	
-%	\subfigure[PT frequency on MATC5760]{
-%		\includegraphics[scale=0.43]{matc5760_dist.eps}
-%		\label{fig:matc5760_dist}
-%	}	
-%	\caption{PT Histograms of MATC2048 ... MATC5760~\label{fig:matc2}}
-%\end{figure}
-%
-%\begin{figure}[h]
-%	\centering
-%	\subfigure[PT frequency on MATC8192]{
-%		\includegraphics[scale=0.43]{matc8192_dist2.eps}
-%		\label{fig:matc8192_dist}
-%	}
-%	\subfigure[PT frequency on MATC11520]{
-%		\includegraphics[scale=0.43]{matc11520_dist.eps}
-%		\label{fig:matc11520_dist}
-%	}
-%	\subfigure[PT frequency on MATC16000]{
-%		\includegraphics[scale=0.43]{matc16000_dist.eps}
-%		\label{fig:matc16000_dist}
-%	}
-%	\subfigure[PT frequency on MATC23040]{
-%		\includegraphics[scale=0.43]{matc23040_dist.eps}
-%		\label{fig:matc23040_dist}
-%	}
-%	\caption{PT Histograms of MATC8192 ... MATC23040~\label{fig:matc3}}
-%\end{figure}
-
-\begin{figure}[h]
-	\centering
-	\subfigure[PT frequency on MATC1K on {\tt sodb10}]{
-		\includegraphics[scale=0.43]{matc1024_dist.eps}
-		\label{fig:matc1k_dist}
-	}
-	\subfigure[PT frequency on MATC2K on {\tt sodb10}]{
-		\includegraphics[scale=0.43]{matc2048_dist.eps}
-		\label{fig:matc2k_dist}
-	}
-	\subfigure[PT frequency on MATC4K on {\tt sodb10}]{
-		\includegraphics[scale=0.43]{matc4096_dist.eps}
-		\label{fig:matc4k_dist}
-	}
-	\subfigure[PT frequency on MATC8K on {\tt sodb10}]{
-		\includegraphics[scale=0.43]{matc8192_dist.eps}
-		\label{fig:matc8k_dist}
-	}
-	\caption{PT Histograms of MATC1024 ... MATC8192~\label{fig:matc}}
-\end{figure}
-
-\clearpage
-\pagebreak
-
-\subsubsection{Row Major}
-
-Figure~\ref{fig:matc} shows a series of 
-histograms of the execution times measured on 
-the same matrix multiplication program in row major as the input sizes grows. 
-The program was run on {\tt sodb12}.
-
-%\begin{figure}[h]
-%	\centering
-%	\subfigure[PT frequency on MATR250]{
-%		\includegraphics[scale=0.43]{matr250_dist.eps}
-%		\label{fig:matr250_dist}
-%	}
-%	\subfigure[PT frequency on MATR500]{
-%		\includegraphics[scale=0.43]{matr500_dist.eps}
-%		\label{fig:matr500_dist}
-%	}
-%	\subfigure[PT frequency on MATR1024]{
-%		\includegraphics[scale=0.43]{matr1024_dist2.eps}
-%		\label{fig:matr1024_dist}
-%	}
-%	\subfigure[PT frequency on MATR1440]{
-%		\includegraphics[scale=0.43]{matr1440_dist.eps}
-%		\label{fig:matr1440_dist}
-%	}
-%	\caption{PT Histograms of MATR250 ... MATR1440~\label{fig:matr}}
-%\end{figure}
-
-%\begin{figure}[h]
-%	\centering
-%	\subfigure[PT frequency on MATR2048]{
-%		\includegraphics[scale=0.43]{matr2048_dist2.eps}
-%		\label{fig:matr2048_dist}
-%	}
-%	\subfigure[PT frequency on MATR2880]{
-%		\includegraphics[scale=0.43]{matr2880_dist.eps}
-%		\label{fig:matr2880_dist}
-%	}	
-%	\subfigure[PT frequency on MATR4096]{
-%		\includegraphics[scale=0.43]{matr4096_dist2.eps}
-%		\label{fig:matr4096_dist}
-%	}	
-%	\subfigure[PT frequency on MATR5760]{
-%		\includegraphics[scale=0.43]{matr5760_dist.eps}
-%		\label{fig:matr5760_dist}
-%	}	
-%	\caption{PT Histograms of MATR2048 ... MATR5760~\label{fig:matr2}}
-%\end{figure}
-%
-%\begin{figure}[h]
-%	\centering
-%	\subfigure[PT frequency on MATR8192]{
-%		\includegraphics[scale=0.43]{matr8192_dist2.eps}
-%		\label{fig:matr8192_dist}
-%	}
-%	\subfigure[PT frequency on MATR11520]{
-%		\includegraphics[scale=0.43]{matr11520_dist.eps}
-%		\label{fig:matr11520_dist}
-%	}	
-%	\subfigure[PT frequency on MATR16000]{
-%		\includegraphics[scale=0.43]{matr16000_dist.eps}
-%		\label{fig:matr16000_dist}
-%	}
-%	\subfigure[PT frequency on MATR23040]{
-%		\includegraphics[scale=0.43]{matr23040_dist.eps}
-%		\label{fig:matr23040_dist}
-%	}
-%	\caption{PT Histograms of MATR8192 ... MATR23040~\label{fig:matr3}}
-%\end{figure}
-
-\begin{figure}[h]
-	\centering
-	\subfigure[PT frequency on MATR1K on {\tt sodb12}]{
-		\includegraphics[scale=0.43]{matr1024_dist.eps}
-		\label{fig:matr1k_dist}
-	}
-	\subfigure[PT frequency on MATR2K on {\tt sodb12}]{
-		\includegraphics[scale=0.43]{matr2048_dist.eps}
-		\label{fig:matr2k_dist}
-	}
-	\subfigure[PT frequency on MATR4K on {\tt sodb12}]{
-		\includegraphics[scale=0.43]{matr4096_dist.eps}
-		\label{fig:matr4k_dist}
-	}
-	\subfigure[PT frequency on MATR8K on {\tt sodb12}]{
-		\includegraphics[scale=0.43]{matr8192_dist.eps}
-		\label{fig:matr8k_dist}
-	}
-	\caption{PT Histograms of MATR1K ... MATR8K~\label{fig:matr}}
-\end{figure}
-
 \bibliographystyle{abbrv}

### 2018-05-04 pt_characterization/rep_check/1024_run.tex
Ricks comment applied

@@
 Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
-{\tt sodb9} &  INC1024 & 300 samples, each & 2017-04-12 $\sim$ 2017-04-23 & 
+{\tt sodb9} (plugged into {\em the upper left} power strip) &  INC1024 & 300 samples, each & 2017-04-12 $\sim$ 2017-04-23 & 
 Figs.~\ref{fig:s9_inc1024_et} and~\ref{fig:s9_inc1024_pt}\\ \hline
@@
 	\centering
-	\subfigure[ET frequency on INC1024-Run1]{
+	\subfigure[ET frequency on INC1024-Run1 on {\tt sodb9}]{
 		\includegraphics[scale=0.28]{1024_run/1024_sec_et_hist1.eps}
@@
 	}
-	\subfigure[ET frequency on INC1024-Run2]{
+	\subfigure[ET frequency on INC1024-Run2 on {\tt sodb9}]{
 		\includegraphics[scale=0.28]{1024_run/1024_sec_et_hist2.eps}
@@
 	}
-	\subfigure[ET frequency on INC1024-Run3]{
+	\subfigure[ET frequency on INC1024-Run3 on {\tt sodb9}]{
 		\includegraphics[scale=0.28]{1024_run/1024_sec_et_hist3.eps}
@@
 
+\clearpage
+\newpage
+
 \subsection{PT}
@@
 	\centering
-	\subfigure[PT frequency on INC1024-Run1]{
+	\subfigure[PT frequency on INC1024-Run1 on {\tt sodb9}]{
 		\includegraphics[scale=0.28]{1024_run/1024_sec_pt_hist1.eps}
@@
 	}
-	\subfigure[PT frequency on INC1024-Run2]{
+	\subfigure[PT frequency on INC1024-Run2 on {\tt sodb9}]{
 		\includegraphics[scale=0.28]{1024_run/1024_sec_pt_hist2.eps}
@@
 	}
-	\subfigure[PT frequency on INC1024-Run3]{
+	\subfigure[PT frequency on INC1024-Run3 on {\tt sodb9}]{
 		\includegraphics[scale=0.28]{1024_run/1024_sec_pt_hist3.eps}

### 2018-05-04 pt_characterization/rep_check/10k_run.tex
Ricks comment applied

@@
 Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
-{\tt sodb9} &  INC8 & 10000 samples & 2017-03-29 $\sim$ 2017-03-30 & Figs.~\ref{fig:inc8_10k_et_hist_v5} and~\ref{fig:inc8_10k_pt_hist_v5}\\ \hline
-{\tt sodb10} &  INC16 & 10000 samples & 2017-03-29 $\sim$ 2017-03-31 & 
+{\tt sodb9} (plugged into {\em the upper left} power strip) &  INC8 & 10,000 samples & 2017-03-29 $\sim$ 2017-03-30 & Figs.~\ref{fig:inc8_10k_et_hist_v5} and~\ref{fig:inc8_10k_pt_hist_v5}\\ \hline
+{\tt sodb10} (plugged into {\em the upper left} power strip) &  INC16 & 10,000 samples & 2017-03-29 $\sim$ 2017-03-31 & 
 Figs.~\ref{fig:inc16_10k_et_hist_v5} and~\ref{fig:inc16_10k_pt_hist_v5}\\ \hline
@@
 	\centering
-	\subfigure[ET frequency on INC8]{
+	\subfigure[ET frequency on INC8 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{10k_run/8_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC8]{
+	\subfigure[PT frequency on INC8 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{10k_run/8_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC16]{
+	\subfigure[ET frequency on INC16 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{10k_run/16_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC16]{
+	\subfigure[PT frequency on INC16 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{10k_run/16_sec_pt_hist_v5.eps}

### 2018-05-04 pt_characterization/rep_check/res1.tex
Ricks comment applied

@@
 This section exhibits histograms on the first run of 
-INC with its task length increasing from 1 second to 4096 seconds. 
+INC with its task length increasing from 1 second to 4,096 seconds. 
 The detailed description of the base data is from Table~\ref{tab:exp_notes1}.
@@
 \begin{center}
-\begin{tabular}{|p{2cm}|p{3cm}|p{3cm}|p{4cm}|p{3cm}|} \hline
+\begin{tabular}{|p{3cm}|p{2cm}|p{3cm}|p{4cm}|p{3cm}|} \hline
 Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
-{\tt sodb9} &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-02 $\sim$ 2017-03-04 & Figs.~\ref{fig:s9_r1_et_hist1},~\ref{fig:s9_r1_et_hist2},~\ref{fig:s9_r1_pt_hist1}, and~\ref{fig:s9_r1_pt_hist2}\\ \hline
-{\tt sodb9} &  INC128$\sim$INC1024 & 300 samples, each & 2017-03-04 $\sim$ 2017-03-11 & 
+{\tt sodb9} (plugged into {\em the upper left} power strip)  &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-02 $\sim$ 2017-03-04 & Figs.~\ref{fig:s9_r1_et_hist1},~\ref{fig:s9_r1_et_hist2},~\ref{fig:s9_r1_pt_hist1}, and~\ref{fig:s9_r1_pt_hist2}\\ \hline
+{\tt sodb9} (plugged into {\em the upper left} power strip)  &  INC128$\sim$ INC1024 & 300 samples, each & 2017-03-04 $\sim$ 2017-03-11 & 
 Figs.~\ref{fig:s9_r1_et_hist3} and~\ref{fig:s9_r1_pt_hist3}\\ \hline
-{\tt sodb10} & INC2048 & 300 samples & 2017-03-02 $\sim$ 2017-03-09 & Figs.~\ref{fig:inc2048_r1_et_hist_v5} and~\ref{fig:inc2048_r1_hist_v5}\\ \hline
-{\tt sodb12} & INC4096 & 300 samples & 2017-02-13 $\sim$ 2017-02-27 & Figs.~\ref{fig:inc4096_r1_et_hist_v5} and~\ref{fig:inc4096_r1_hist_v5}\\ \hline
+{\tt sodb10} (plugged into {\em the upper left} power strip)  & INC2048 & 300 samples & 2017-03-02 $\sim$ 2017-03-09 & Figs.~\ref{fig:inc2048_r1_et_hist_v5} and~\ref{fig:inc2048_r1_hist_v5}\\ \hline
+{\tt sodb12} (plugged into {\em the upper left} power strip)  & INC4096 & 300 samples & 2017-02-13 $\sim$ 2017-02-27 & Figs.~\ref{fig:inc4096_r1_et_hist_v5} and~\ref{fig:inc4096_r1_hist_v5}\\ \hline
 \end{tabular}
@@
 	\centering
-	\subfigure[ET frequency on INC1]{
+	\subfigure[ET frequency on INC1 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/1_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC2]{
+	\subfigure[ET frequency on INC2 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/2_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC4]{
+	\subfigure[ET frequency on INC4 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/4_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC8]{
+	\subfigure[ET frequency on INC8 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/8_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[ET frequency on INC16]{
+	\subfigure[ET frequency on INC16 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/16_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC32]{
+	\subfigure[ET frequency on INC32 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/32_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC64]{
+	\subfigure[ET frequency on INC64 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/64_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[ET frequency on INC128]{
+	\subfigure[ET frequency on INC128 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/128_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC256]{
+	\subfigure[ET frequency on INC256 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/256_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC512]{
+	\subfigure[ET frequency on INC512 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/512_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC1024]{
+	\subfigure[ET frequency on INC1024 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/1024_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[ET frequency on INC2048]{
+	\subfigure[ET frequency on INC2048  on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{repet_data1/2048_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC4096]{
+	\subfigure[ET frequency on INC4096 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{repet_data1/4096_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC1]{
+	\subfigure[PT frequency on INC1 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/1_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC2]{
+	\subfigure[PT frequency on INC2 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/2_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC4]{
+	\subfigure[PT frequency on INC4 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/4_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC8]{
+	\subfigure[PT frequency on INC8 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/8_sec_pt_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC16]{
+	\subfigure[PT frequency on INC16 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/16_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC32]{
+	\subfigure[PT frequency on INC32 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/32_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC64]{
+	\subfigure[PT frequency on INC64 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/64_sec_pt_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC128]{
+	\subfigure[PT frequency on INC128 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/128_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC256]{
+	\subfigure[PT frequency on INC256 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/256_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC512]{
+	\subfigure[PT frequency on INC512 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/512_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC1024]{
+	\subfigure[PT frequency on INC1024 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data1/1024_sec_pt_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC2048]{
+	\subfigure[PT frequency on INC2048 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{repet_data1/2048_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC4096]{
+	\subfigure[PT frequency on INC4096 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{repet_data1/4096_sec_pt_hist_v5.eps}
@@
 Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
-{\tt sodb9} &  INC96$\sim$INC256 & 1000 samples, each & 2017-05-24 $\sim$ 2017-06-06 & 
-Figs.~\ref{fig:ut_hist3}, \ref{fig:inc224_ut_hist}, and~\ref{fig:inc256_ut_hist}\\ \hline
-					&  INC3, 6, 12, 24, 48, 72, 80, 88, 104, 112, and 120 & 1000 samples, each & 2017-06-07 $\sim$ 2017-06-16 & Figs.~\ref{fig:inc3_ut_hist}, ~\ref{fig:inc6_ut_hist}, ~\ref{fig:inc12_ut_hist}, ~\ref{fig:inc24_ut_hist},
+{\tt sodb9} &  INC96$\sim$INC256 & 1000 samples, each & 2017-05-24 $\sim$ 2017-06-06 & Figs.~\ref{fig:ut_hist3}, \ref{fig:inc224_ut_hist}, and~\ref{fig:inc256_ut_hist}\\ \cline{2-4}
+(plugged into {\em the upper left} power strip)	&  INC3, 6, 12, 24, 48, 72, 80, 88, 104, 112, and 120 & 1000 samples, each & 2017-06-07 $\sim$ 2017-06-16 & Figs.~\ref{fig:inc3_ut_hist}, ~\ref{fig:inc6_ut_hist}, ~\ref{fig:inc12_ut_hist}, ~\ref{fig:inc24_ut_hist},
 ~\ref{fig:inc48_ut_hist}, ~\ref{fig:inc72_ut_hist}, ~\ref{fig:inc80_ut_hist}, ~\ref{fig:inc88_ut_hist}, ~\ref{fig:inc104_ut_hist},~\ref{fig:inc112_ut_hist}, and~\ref{fig:inc120_ut_hist}\\ \hline
@@
 	\centering
-	\subfigure[PT frequency on INC3]{
+	\subfigure[PT frequency on INC3  on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/3_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC6]{
+	\subfigure[PT frequency on INC6 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/6_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC12]{
+	\subfigure[PT frequency on INC12  on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/12_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC24]{
+	\subfigure[PT frequency on INC24 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/24_sec_pt_hist.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC48]{
+	\subfigure[PT frequency on INC48 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/48_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC72]{
+	\subfigure[PT frequency on INC72 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/72_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC80]{
+	\subfigure[PT frequency on INC80 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/80_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC88]{
+	\subfigure[PT frequency on INC88 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/88_sec_pt_hist.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC96]{
+	\subfigure[PT frequency on INC96 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/96_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC104]{
+	\subfigure[PT frequency on INC104 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/104_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC112]{
+	\subfigure[PT frequency on INC112 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/112_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC120]{
+	\subfigure[PT frequency on INC120 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/120_sec_pt_hist.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC160]{
+	\subfigure[PT frequency on INC160 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/160_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC192]{
+	\subfigure[PT frequency on INC192 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/192_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC224]{
+	\subfigure[PT frequency on INC224 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/224_sec_pt_hist.eps}
@@
 	\includegraphics[scale=1]{u_s_time/new_3d_plot}\label{fig:3d_plot}
-	\caption{3D histograms of PT on doubly-increasing INC~\label{fig:hist3d}}
+	\caption{3D histograms of PT on doubly-increasing INC on {\tt sodb9} for up to INC1024, on {\tt sodb10} for INC2048, on {\tt sodb12} for INC4096~\label{fig:hist3d}}
 \end{figure}
@@
 	\includegraphics[scale=1]{u_s_time/3dplot_utime_only}\label{fig:3d_plot_u}
-	\caption{3D histograms on INC: user time only~\label{fig:hist3d_u}}
+	\caption{3D histograms on INC on {\tt sodb9} for up to INC1024, on {\tt sodb10} for INC2048, on {\tt sodb12} for INC4096: user time only~\label{fig:hist3d_u}}
 \end{figure}

### 2018-05-04 pt_characterization/rep_check/res2.tex
Ricks comment applied

@@
 Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
-{\tt sodb9} &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-13 $\sim$ 2017-03-14 & Figs.~\ref{fig:s9_r2_et_hist1},~\ref{fig:s9_r2_et_hist2},~\ref{fig:s9_r2_pt_hist1}, and~\ref{fig:s9_r2_pt_hist2}\\ \hline
+{\tt sodb9} (plugged into {\em the upper left} power strip) &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-13 $\sim$ 2017-03-14 & Figs.~\ref{fig:s9_r2_et_hist1},~\ref{fig:s9_r2_et_hist2},~\ref{fig:s9_r2_pt_hist1}, and~\ref{fig:s9_r2_pt_hist2}\\ \hline
 %{\tt sodb9} &  INC128$\sim$INC1024 & 300 samples, each & 2017-03-14 $\sim$ 2017-03-21 & 
 %Figs.~\ref{fig:s9_r2_et_hist3} and~\ref{fig:s9_r2_pt_hist3}\\ \hline
-{\tt sodb9} &  INC512$\sim$INC1024 & 300 samples, each & 2017-03-17 $\sim$ 2017-03-21 & 
+{\tt sodb9} (plugged into {\em the upper left} power strip) &  INC512$\sim$INC1024 & 300 samples, each & 2017-03-17 $\sim$ 2017-03-21 & 
 Figs.~\ref{fig:s9_r2_et_hist3} and~\ref{fig:s9_r2_pt_hist3}\\ \hline
-{\tt sodb10} & INC2048 & 300 samples & 2017-03-13 $\sim$ 2017-03-20 & Figs.~\ref{fig:inc2048_r2_et_hist_v5} and~\ref{fig:inc2048_r2_hist_v5}\\ \hline
-{\tt sodb12} & INC4096 & 300 samples & 2017-03-02 $\sim$ 2017-03-17 & Figs.~\ref{fig:inc4096_r2_et_hist_v5} and~\ref{fig:inc4096_r2_hist_v5}\\ \hline
-{\tt sodb10} & INC8192& 261 samples & 2017-04-27 $\sim$ 2017-05-21 & Figs.~\ref{fig:inc8192_r2_et_hist_v5} and~\ref{fig:inc8192_r2_hist_v5}\\ \hline
-{\tt sodb12} & INC16384& 130 samples & 2017-04-27 $\sim$ 2017-05-21 & Figs.~\ref{fig:inc16384_r2_et_hist_v5} and~\ref{fig:inc16384_r2_hist_v5}\\ \hline
+{\tt sodb10} (plugged into {\em the upper left} power strip) & INC2048 & 300 samples & 2017-03-13 $\sim$ 2017-03-20 & Figs.~\ref{fig:inc2048_r2_et_hist_v5} and~\ref{fig:inc2048_r2_hist_v5}\\ \hline
+{\tt sodb12} (plugged into {\em the upper left} power strip) & INC4096 & 300 samples & 2017-03-02 $\sim$ 2017-03-17 & Figs.~\ref{fig:inc4096_r2_et_hist_v5} and~\ref{fig:inc4096_r2_hist_v5}\\ \hline
+{\tt sodb10} (plugged into {\em the upper left} power strip) & INC8192& 261 samples & 2017-04-27 $\sim$ 2017-05-21 & Figs.~\ref{fig:inc8192_r2_et_hist_v5} and~\ref{fig:inc8192_r2_hist_v5}\\ \hline
+{\tt sodb12} (plugged into {\em the upper left} power strip) & INC16384& 130 samples & 2017-04-27 $\sim$ 2017-05-21 & Figs.~\ref{fig:inc16384_r2_et_hist_v5} and~\ref{fig:inc16384_r2_hist_v5}\\ \hline
 \end{tabular}
@@
 	\centering
-	\subfigure[ET frequency on INC1]{
+	\subfigure[ET frequency on INC1 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/1_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC2]{
+	\subfigure[ET frequency on INC2 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/2_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC4]{
+	\subfigure[ET frequency on INC4 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/4_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC8]{
+	\subfigure[ET frequency on INC8 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/8_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[ET frequency on INC16]{
+	\subfigure[ET frequency on INC16 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/16_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC32]{
+	\subfigure[ET frequency on INC32 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/32_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC64]{
+	\subfigure[ET frequency on INC64 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/64_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[ET frequency on INC128]{
+	\subfigure[ET frequency on INC128 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/128_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC256]{
+	\subfigure[ET frequency on INC256 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/256_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC512]{
+	\subfigure[ET frequency on INC512 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/512_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC1024]{
+	\subfigure[ET frequency on INC1024 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/1024_sec_et_hist_v5.eps}
@@
 	\centering
-	\subfigure[ET frequency on INC2048]{
+	\subfigure[ET frequency on INC2048 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{repet_data2/2048_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC4096]{
+	\subfigure[ET frequency on INC4096 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{repet_data2/4096_sec_et_hist_v5.eps}
@@
 	}
-	\subfigure[ET frequency on INC8192]{
+	\subfigure[ET frequency on INC8192 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{repet_data2/8192_sec_et_hist.eps}
@@
 	}
-	\subfigure[ET frequency on INC16384]{
+	\subfigure[ET frequency on INC16384 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{repet_data2/16384_sec_et_hist.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC1]{
+	\subfigure[PT frequency on INC1 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/1_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC2]{
+	\subfigure[PT frequency on INC2 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/2_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC4]{
+	\subfigure[PT frequency on INC4 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/4_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC8]{
+	\subfigure[PT frequency on INC8 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/8_sec_pt_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC16]{
+	\subfigure[PT frequency on INC16 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/16_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC32]{
+	\subfigure[PT frequency on INC32 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/32_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC64]{
+	\subfigure[PT frequency on INC64 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/64_sec_pt_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC128]{
+	\subfigure[PT frequency on INC128 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/128_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC256]{
+	\subfigure[PT frequency on INC256 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/256_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC512]{
+	\subfigure[PT frequency on INC512 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/512_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC1024]{
+	\subfigure[PT frequency on INC1024 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{repet_data2/1024_sec_pt_hist_v5.eps}
@@
 	\centering
-	\subfigure[PT frequency on INC2048]{
+	\subfigure[PT frequency on INC2048 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{repet_data2/2048_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC4096]{
+	\subfigure[PT frequency on INC4096 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{repet_data2/4096_sec_pt_hist_v5.eps}
@@
 	}
-	\subfigure[PT frequency on INC8192]{
+	\subfigure[PT frequency on INC8192 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{repet_data2/8192_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC16384]{
+	\subfigure[PT frequency on INC16384 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{repet_data2/16384_sec_pt_hist.eps}

### 2018-05-04 pt_characterization/rep_check/u_s_time.tex
Ricks comment applied

@@
 	\centering
-	\subfigure[User time frequency on INC1]{
+	\subfigure[User time frequency on INC1 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/1_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC2]{
+	\subfigure[User time frequency on INC2 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/2_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC3]{
+	\subfigure[User time frequency on INC3 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/3_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC4]{
+	\subfigure[User time frequency on INC4 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/4_sec_ut_hist.eps}
@@
 	\centering
-	\subfigure[User time frequency on INC6]{
+	\subfigure[User time frequency on INC6 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/6_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC8]{
+	\subfigure[User time frequency on INC8 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/8_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC12]{
+	\subfigure[User time frequency on INC12 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/12_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC16]{
+	\subfigure[User time frequency on INC16 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/16_sec_ut_hist.eps}
@@
 	\centering
-	\subfigure[User time frequency on INC24]{
+	\subfigure[User time frequency on INC24 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/24_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC32]{
+	\subfigure[User time frequency on INC32 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/32_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC48]{
+	\subfigure[User time frequency on INC48 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/48_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC64]{
+	\subfigure[User time frequency on INC64 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/64_sec_ut_hist.eps}
@@
 	\centering
-	\subfigure[User time frequency on INC72]{
+	\subfigure[User time frequency on INC72 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/72_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC80]{
+	\subfigure[User time frequency on INC80 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/80_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC88]{
+	\subfigure[User time frequency on INC88 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/88_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC96]{
+	\subfigure[User time frequency on INC96 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/96_sec_ut_hist.eps}
@@
 	\centering
-	\subfigure[User time frequency on INC104]{
+	\subfigure[User time frequency on INC104 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/104_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC112]{
+	\subfigure[User time frequency on INC112 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/112_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC120 with bin size = 1 ms]{
+	\subfigure[User time frequency on INC120  on {\tt sodb9} with bin size = 1 ms]{
 		\includegraphics[scale=0.43]{u_s_time/120_sec_ut_hist1.eps}
@@
 	}
-	\subfigure[User time frequency on INC120 with bin size = 2 ms]{
+	\subfigure[User time frequency on INC120  on {\tt sodb9} with bin size = 2 ms]{
 		\includegraphics[scale=0.43]{u_s_time/120_sec_ut_hist.eps}
@@
 	\centering
-	\subfigure[User time frequency on INC128]{
+	\subfigure[User time frequency on INC128  on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/128_sec_ut_hist2.eps}
@@
 	}
-	\subfigure[User time frequency on INC160]{
+	\subfigure[User time frequency on INC160 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/160_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC192]{
+	\subfigure[User time frequency on INC192 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/192_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC224]{
+	\subfigure[User time frequency on INC224 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/224_sec_ut_hist.eps}
@@
 	\centering
-	\subfigure[User time frequency on INC256]{
+	\subfigure[User time frequency on INC256 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/256_sec_ut_hist2.eps}
@@
 	}
-	\subfigure[User time frequency on INC512 (with one extreme outlier of 513,259 msec removed)]{
+	\subfigure[User time frequency on INC512 on {\tt sodb9} (with one extreme outlier of 513,259 msec removed)]{
 		\includegraphics[scale=0.43]{u_s_time/512_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC1024]{
+	\subfigure[User time frequency on INC1024 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/1024_sec_ut_hist.eps}
@@
 	\centering
-	\subfigure[User time frequency on INC2048]{
+	\subfigure[User time frequency on INC2048 on {\tt sodb10}]{
 		\includegraphics[scale=0.43]{u_s_time/2048_sec_ut_hist.eps}
@@
 	}
-	\subfigure[User time frequency on INC4096]{
+	\subfigure[User time frequency on INC4096 on {\tt sodb12}]{
 		\includegraphics[scale=0.43]{u_s_time/4096_sec_ut_hist.eps}
@@
 	\centering
-	\subfigure[System time frequency on INC1]{
+	\subfigure[System time frequency on INC1 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/1_sec_st_hist.eps}
@@
 	}
-	\subfigure[System time frequency on INC2]{
+	\subfigure[System time frequency on INC2 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/2_sec_st_hist.eps}
@@
 	}
-	\subfigure[System time frequency on INC4]{
+	\subfigure[System time frequency on INC4 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/4_sec_st_hist.eps}
@@
 	}
-	\subfigure[System time frequency on INC8]{
+	\subfigure[System time frequency on INC8 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/8_sec_st_hist.eps}
@@
 	\centering
-	\subfigure[System time frequency on INC16]{
+	\subfigure[System time frequency on INC16 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/16_sec_st_hist.eps}
@@
 	}
-	\subfigure[System time frequency on INC32]{
+	\subfigure[System time frequency on INC32 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/32_sec_st_hist.eps}
@@
 	}
-	\subfigure[System time frequency on INC64]{
+	\subfigure[System time frequency on INC64 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/64_sec_st_hist.eps}
@@
 	\centering
-	\subfigure[System time frequency on INC128]{
+	\subfigure[System time frequency on INC128 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/128_sec_st_hist.eps}
@@
 	}
-	\subfigure[System time frequency on INC256]{
+	\subfigure[System time frequency on INC256 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/256_sec_st_hist.eps}
@@
 	}
-	\subfigure[System time frequency on INC512]{
+	\subfigure[System time frequency on INC512 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/512_sec_st_hist.eps}
@@
 	}
-	\subfigure[System time frequency on INC1024]{
+	\subfigure[System time frequency on INC1024 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/1024_sec_st_hist.eps}
@@
 	\centering
-	\subfigure[System time frequency on INC2048]{
+	\subfigure[System time frequency on INC2048 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/2048_sec_st_hist.eps}
@@
 	}
-	\subfigure[System time frequency on INC4096]{
+	\subfigure[System time frequency on INC4096 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{u_s_time/4096_sec_st_hist.eps}

### 2018-05-04 pt_characterization/machine_cons_check/new_INC_data_s9/inc_hist.r

@@ -1,0 +1,241 @@
+### sodb9-INC1
+numSamples <- 1000
+x = read.csv(file="INC1.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("s9_INC1_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 800 
+title <- paste("PT frequency of INC1 on sodb9 (sample size: ",numSamples,")")
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main=title, 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+### sodb9-INC2
+numSamples <- 1000
+x = read.csv(file="INC2.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("s9_INC2_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
[…]

### 2018-05-16 pt_characterization/machine_cons_check/post_exp.tex
Added INC13

@@
 \clearpage
+\subsection{Power Strip Influence~\label{sec:sodb8}}
+
+A short description of the runs we used for this experiment are exhibited in Table~\ref{tab:exp_notes4}.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
+Machine & Task Length & Description & Time Length\\ \hline
+{\tt sodb8}  (plugged into {\em bottom left} power strip) &  INC13 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
+{\tt sodb8}  (plugged into {\em the upper left} power strip) &  INC13 & Runs of 1,000 samples & 2018-04-28 $\sim$2018-04-30\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Detailed description of INC data used for histograms\label{tab:exp_notes4}}
+\end{table}
+
+\begin{figure}[H]
+	\centering
+	\subfigure[PT frequency on INC13 on {\tt sodb8}  (plugged into {\em bottom left} power strip) (copied from Fig.~\ref{fig:s8_inc13_dist})]{
+		\includegraphics[scale=0.43]{newer_exp/sodb8_INC13_0_dist.eps}
+		\label{fig:s8_inc13_dist_before}
+	}
+	\subfigure[PT frequency on INC13 on {\tt sodb8}  (plugged into {\em the upper left} power strip)]{
+		\includegraphics[scale=0.43]{new_INC_data_s9/s8_INC13_dist.eps}
+		\label{fig:s8_inc13_dist_after}
+	}
+	\caption{PT Histograms on INC13 on {\tt sodb8} Before and After Power Strip Relocation\label{fig:power_strip}}
+\end{figure}
+
+\newpage
+\clearpage
 

### 2018-05-16 pt_characterization/machine_cons_check/new_INC_data_s9/inc_hist.r

@@ -1,3 +1,33 @@
+### sodb8-INC13
+numSamples <- 1000
+x = read.csv(file="INC13.0.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("s8_INC13_dist.eps")
+binsize=1
+#x <- subset(x, x$ITERNUM != 9394)
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+xmin
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+xmin <-min(x$PRTIME)
+xmax <-max(x$PRTIME)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 400 
+title <- paste("PT frequency of INC13 on sodb8 (sample size: ",numSamples,")")
+plot(h, xaxt='n', freq=TRUE,ylim=c(0,ymax), col="blue", main=title, 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+#axis(side=1,at=seq(min(h$mids),max(h$mids)+binsize,binsize),labels=seq(ceiling(min(h$mids)),ceiling(max(h$mids))+binsize,binsize))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
 ### sodb9-INC1
 numSamples <- 1000
 x = read.csv(file="INC1.0.dat",head=TRUE,sep="\t")

### 2018-05-18 pt_characterization/machine_cons_check/paper.tex
Rick's comments & INC data

@@
 \paragraph{Experimental Notes:} In our experiments 
-we used four nodes ({\tt sodb8}, {\tt sodb9}, {\tt sodb10}, and {\tt sodb12} ) in the same cluster. 
+we used four nodes ({\tt sodb8}, {\tt sodb9}, {\tt sodb10}, and {\tt sodb12}) in the same cluster. 
 A short summary of the experiments is exhibited in Table~\ref{tab:exp_notes}.
@@
 {\tt sodb10}  (plugged into {\em the upper left} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
-{\tt sodb12}  (plugged into {\em the upper left} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb12}  (plugged into {\em the upper right} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
 \end{tabular}

### 2018-05-18 pt_characterization/machine_cons_check/post_exp.tex
Rick's comments & INC data

@@
 \paragraph{Experimental Notes:} In the experiments 
-we also used four nodes ({\tt sodb8}, {\tt sodb9}, {\tt sodb10}, and {\tt sodb12} ) in the same cluster. 
+we also used four nodes ({\tt sodb8}, {\tt sodb9}, {\tt sodb10}, and {\tt sodb12}) in the same cluster. 
 A short summary of the experiments is shown in Table~\ref{tab:exp_notes2}.
@@
 Machine & Task Length & Description & Time Length\\ \hline
-{\tt sodb8} (plugged into {\em bottom left} power strip) &  INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
+{\tt sodb8} (plugged into {\em the upper right} power strip) &  INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
 {\tt sodb9}  (plugged into {\em the upper left} power strip) &  IINC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
 {\tt sodb10} (plugged into {\em the upper left} power strip)  & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
-{\tt sodb12} (plugged into {\em the upper left} power strip) & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
+{\tt sodb12} (plugged into {\em the upper right} power strip) & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
 \end{tabular}

### 2018-05-18 pt_characterization/real_data/paper.tex
Rick's comments & INC data

@@
 {\tt sodb10} (plugged into {\em the upper left} power strip)  & MATC1K, MATC2K, MATC4K, MATC8K & Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
-{\tt sodb12} (plugged into {\em the upper left} power strip) & MATR1K, MATR2K, MATR4K, MATR8K& Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
+{\tt sodb12} (plugged into {\em the upper right} power strip) & MATR1K, MATR2K, MATR4K, MATR8K& Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
 \end{tabular}
@@
 the same matrix multiplication program in row major as the input sizes grows. 
-The program was run on {\tt sodb12} plugged into {\em the upper left} power strip.
+The program was run on {\tt sodb12} plugged into {\em the upper right} power strip.
 
@@
 for instance, SORT100 indicates the insertion sort program over 100K elements. 
-This experiment was performed on {\tt sodb8} plugged into {\em bottom left} power strip.
+This experiment was performed on {\tt sodb8} \underline{plugged into {\em bottom left}} power strip.
 
@@
 we perform a match on an SORT program and its corresponding INC program. 
-{\tt sodb8} was plugged into {\em bottom left} power strip 
-while {\tt sodb10} plugged into {\em the upper left} power strip.
+\underline{{\tt sodb8}} was plugged into \underline{{\em bottom left} power strip} 
+while \underline{{\tt sodb10}} plugged into \underline{{\em the upper left} power strip}.
 

### 2018-05-18 pt_characterization/rep_check/res1.tex
Rick's comments & INC data

@@
 {\tt sodb10} (plugged into {\em the upper left} power strip)  & INC2048 & 300 samples & 2017-03-02 $\sim$ 2017-03-09 & Figs.~\ref{fig:inc2048_r1_et_hist_v5} and~\ref{fig:inc2048_r1_hist_v5}\\ \hline
-{\tt sodb12} (plugged into {\em the upper left} power strip)  & INC4096 & 300 samples & 2017-02-13 $\sim$ 2017-02-27 & Figs.~\ref{fig:inc4096_r1_et_hist_v5} and~\ref{fig:inc4096_r1_hist_v5}\\ \hline
+{\tt sodb12} (plugged into {\em the upper right} power strip)  & INC4096 & 300 samples & 2017-02-13 $\sim$ 2017-02-27 & Figs.~\ref{fig:inc4096_r1_et_hist_v5} and~\ref{fig:inc4096_r1_hist_v5}\\ \hline
 \end{tabular}

### 2018-05-18 pt_characterization/rep_check/res2.tex
Rick's comments & INC data

@@
 {\tt sodb10} (plugged into {\em the upper left} power strip) & INC2048 & 300 samples & 2017-03-13 $\sim$ 2017-03-20 & Figs.~\ref{fig:inc2048_r2_et_hist_v5} and~\ref{fig:inc2048_r2_hist_v5}\\ \hline
-{\tt sodb12} (plugged into {\em the upper left} power strip) & INC4096 & 300 samples & 2017-03-02 $\sim$ 2017-03-17 & Figs.~\ref{fig:inc4096_r2_et_hist_v5} and~\ref{fig:inc4096_r2_hist_v5}\\ \hline
+{\tt sodb12} (plugged into {\em the upper right} power strip) & INC4096 & 300 samples & 2017-03-02 $\sim$ 2017-03-17 & Figs.~\ref{fig:inc4096_r2_et_hist_v5} and~\ref{fig:inc4096_r2_hist_v5}\\ \hline
 {\tt sodb10} (plugged into {\em the upper left} power strip) & INC8192& 261 samples & 2017-04-27 $\sim$ 2017-05-21 & Figs.~\ref{fig:inc8192_r2_et_hist_v5} and~\ref{fig:inc8192_r2_hist_v5}\\ \hline
-{\tt sodb12} (plugged into {\em the upper left} power strip) & INC16384& 130 samples & 2017-04-27 $\sim$ 2017-05-21 & Figs.~\ref{fig:inc16384_r2_et_hist_v5} and~\ref{fig:inc16384_r2_hist_v5}\\ \hline
+{\tt sodb12} (plugged into {\em the upper right} power strip) & INC16384& 130 samples & 2017-04-27 $\sim$ 2017-05-21 & Figs.~\ref{fig:inc16384_r2_et_hist_v5} and~\ref{fig:inc16384_r2_hist_v5}\\ \hline
 \end{tabular}

### 2018-09-14 pt_characterization/machine_cons_check/paper.tex
Rick's changes

@@
 Young-Kyoon Suh\\
-Data-Intensive Systems Laboratory\\
+Data \& Knowledge Engineering Laboratory\\
 School of Computer Science and Engineering\\
@@
 A short summary of the experiments is exhibited in Table~\ref{tab:exp_notes}.
+One thing to mention is that data on INC17.2 with old power supply does not exist. 
+Note however that we do have data with INC17.2 with new power supply as described in Table~\ref{tab:exp_notes2}.
 
@@
 Machine & Task Length & Description & Time Length\\ \hline
-{\tt sodb8} (plugged into {\em bottom left} power strip) & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
-{\tt sodb9}  (plugged into {\em the upper left} power strip)  &  INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
-{\tt sodb10}  (plugged into {\em the upper left} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
-{\tt sodb12}  (plugged into {\em the upper right} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb8} (plugged into the {\em bottom left} power strip) & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb9}  (plugged into the {\em upper left} power strip)  &  INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb10}  (plugged into the {\em upper left} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb12}  (plugged into the {\em upper right} power strip)  & INC1$\sim$INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
 \end{tabular}

### 2018-09-14 pt_characterization/machine_cons_check/post_exp.tex
Rick's changes

@@
 Machine & Task Length & Description & Time Length\\ \hline
-{\tt sodb8} (plugged into {\em the upper right} power strip) &  INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
-{\tt sodb9}  (plugged into {\em the upper left} power strip) &  IINC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
-{\tt sodb10} (plugged into {\em the upper left} power strip)  & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
-{\tt sodb12} (plugged into {\em the upper right} power strip) & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
+{\tt sodb8} (plugged into the {\em upper right} power strip) &  INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
+{\tt sodb9}  (plugged into the {\em upper left} power strip) &  IINC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
+{\tt sodb10} (plugged into the {\em upper left} power strip)  & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
+{\tt sodb12} (plugged into the {\em upper right} power strip) & INC16, INC13, and INC17.2 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
 \end{tabular}
@@
 Machine & Task Length & Description & Time Length\\ \hline
-{\tt sodb9}  (plugged into {\em the upper left} power strip) &  INC1$\sim$INC128 & Runs of 1,000 samples & 2018-04-28 $\sim$2018-04-30\\ \hline
+{\tt sodb9}  (plugged into the {\em upper left} power strip) &  INC1$\sim$INC128 & Runs of 1,000 samples & 2018-04-28 $\sim$2018-04-30\\ \hline
 \end{tabular}
@@
 Machine & Task Length & Description & Time Length\\ \hline
-{\tt sodb8}  (plugged into {\em bottom left} power strip) &  INC13 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
-{\tt sodb8}  (plugged into {\em the upper left} power strip) &  INC13 & Runs of 1,000 samples & 2018-04-28 $\sim$2018-04-30\\ \hline
+{\tt sodb8}  (plugged into the {\em bottom left} power strip) &  INC13 & Runs of 1,000 samples & 2018-03-17 $\sim$2018-03-18\\ \hline
+{\tt sodb8}  (plugged into the {\em upper right} power strip) &  INC13 & Runs of 1,000 samples & 2018-04-28 $\sim$2018-04-30\\ \hline
 \end{tabular}
@@
 	\centering
-	\subfigure[PT frequency on INC13 on {\tt sodb8}  (plugged into {\em bottom left} power strip) (copied from Fig.~\ref{fig:s8_inc13_dist})]{
+	\subfigure[PT frequency on INC13 on {\tt sodb8}  (plugged into the {\em bottom left} power strip) (copied from Fig.~\ref{fig:s8_inc13_dist})]{
 		\includegraphics[scale=0.43]{newer_exp/sodb8_INC13_0_dist.eps}
@@
 	}
-	\subfigure[PT frequency on INC13 on {\tt sodb8}  (plugged into {\em the upper left} power strip)]{
+	\subfigure[PT frequency on INC13 on {\tt sodb8}  (plugged into the {\em upper right} power strip)]{
 		\includegraphics[scale=0.43]{new_INC_data_s9/s8_INC13_dist.eps}

### 2018-09-14 pt_characterization/real_data/paper.tex
Rick's changes

@@
 \author{
-Database \& Big Data Systems Laboratory,\\
+Data \& Knowledge Engineering Laboratory,\\
 School of Computer Science and Engineering, \\

### 2018-09-14 pt_characterization/real_data/paper.tex
More changes

--- 2018-05-18 pt_characterization/real_data/paper.tex
@@
 \author{
-Database \& Big Data Systems Laboratory,\\
+Young-Kyoon Suh,\\
+Data \& Knowledge Engineering Laboratory,\\
 School of Computer Science and Engineering, \\
-Kyungpook National University,\\
-Young-Kyoon Suh\\
+Kyungpook National University\\
 }
@@
 for instance, SORT100 indicates the insertion sort program over 100K elements. 
-This experiment was performed on {\tt sodb8} \underline{plugged into {\em bottom left}} power strip.
+%This experiment was performed on {\tt sodb8} \underline{plugged into {\em bottom left}} power strip.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
+Machine & Task Length & Description & Time Length\\ \hline
+{\tt sodb8}  (plugged into the {\em bottom left} power strip) & SORT100 ... SORT2320 & Runs of 300 samples & 2018-03-02 $\sim$2018-03-10\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Detailed description of INC data used for the subsequent histograms\label{tab:exp_notes1}}
+\end{table}
 
@@
 we perform a match on an SORT program and its corresponding INC program. 
-\underline{{\tt sodb8}} was plugged into \underline{{\em bottom left} power strip} 
-while \underline{{\tt sodb10}} plugged into \underline{{\em the upper left} power strip}.
+%\underline{{\tt sodb8}} was plugged into \underline{{\em bottom left} power strip} 
+%while \underline{{\tt sodb10}} plugged into \underline{{\em the upper left} power strip}.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
+Machine & Task Length & Description & Time Length\\ \hline
+{\tt sodb8}  (plugged into the {\em bottom left} power strip) & SORT144 ... SORT344 & Runs of 300 samples & 2018-03-09 $\sim$2018-03-17\\ \hline
+{\tt sodb10}  (plugged into the {\em upper left} power strip) & INC8.8 ... INC51.5 & Runs of 300 samples & 2018-03-18 $\sim$2018-03-18\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Detailed description of INC data used for the subsequent histograms\label{tab:exp_notes2}}
+\end{table}
 

### 2018-09-19 pt_characterization/rep_check/res2.tex
TODOs done

@@
 \end{table}
+
+\clearpage
+\newpage
+
+\subsubsection{Analysis on Some Outliers Found from INC16 and INC64 on {\tt sodb9}}
+In this section we investigate what happened about some outlying samples on INC16 and INC64 
+shown in Figures~\ref{fig:inc16_r2_hist_v5} and~\ref{fig:inc64_r2_hist_v5}.
+
+%Regarding INC16, we looked into detailed information of processes that were 
+%captured at some executions corresponding to the two bins of 16,041 (frequency = 3) 
+%and 16,050 (frequency = 1) msecs in PT. 
+Regarding INC16, we looked into detailed information of processes that were 
+captured at one execution corresponding to the bin of 16,050 (frequency = 1) msec in PT. 
+For the purpose of comparison, we also explored processes captured 
+at the second-highest bin of 16,041 msec in PT.
+
+Table~\ref{tab:inc16_daemons} shows the overall information of the captured processes at the two bins. 
+As indicated by the table, there were 
+some infrequent daemon processes such as {\tt sshd}, {\tt grep}, {\tt md0\_raid1}, {\tt cifsd}, running for some time 
+when the INC16 program was run.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|l|l|l|l|} \hline
+Iteration number &  Process ID & Process Name  & PT (in msec)\\ \hline
+77 (, 209, or 569) & 10185 & {\tt incr\_work} (=INC16) & 16,041 \\ \hline
+& 28525 & {\tt proc\_monitor} & 196 \\ \hline
+& 167 & {\tt kslowd001} &  1 \\ \hline
+& 166 & {\tt kslowd000} &  1 \\ \hline \hline 
+703 & 10185 & {\tt incr\_work} (=INC16) & 16,050 \\ \hline
+& 28525 & {\tt proc\_monitor} & 196 \\ \hline
+& 12322 & {\tt sshd} & 14 \\ \hline
+& 12313 & {\tt sshd}  & 13 \\ \hline
+& 12324 & {\tt grep}  &  6 \\ \hline
+& 472 & {\tt md0\_raid1} &  2 \\ \hline
+& 12314 & {\tt sshd}  &  2 \\ \hline
+& 12315 & {\tt grep}  &  2 \\ \hline
+& 12319 & {\tt grep}  &  1 \\ \hline
+& 167 & {\tt kslowd001} &  1 \\ \hline
+& 12323 & {\tt sshd}  &  1 \\ \hline
+& 166 & {\tt kslowd000} &  1 \\ \hline
+& 12328 & {\tt grep}  &  1 \\ \hline
+& 1925 & {\tt cifsd} &  1 \\ \hline 
+\hline
+\end{tabular}
+\end{center}
+\caption{Daemons observed at the second-rightmost and rightmost bins in Figure~\ref{fig:inc16_r2_hist_v5}~\label{tab:inc16_daemons}}
+\end{table}
+
+Regarding INC64, we looked into detailed information of processes that were 
+captured at one execution corresponding to the bin of 64,151 (frequency = 1) msec in PT. 
+For the purpose of comparison, we also explored processes captured 
+at the second-highest bin of 16,041 msec in PT.
+
+Table~\ref{tab:inc64_daemons} shows the overall information of the captured processes at the two bins. 
+As indicated by the table, there were also 
+the same kind of infrequent daemon processes such as {\tt sshd}, {\tt grep}, {\tt md0\_raid1} as well as 
+{\tt java} running for some time when the INC64 program was run. 
+%In particular, we may not able to avoid hosting java for the longer INCs 
+%as it appears by all means java ran due to any reason including GC 
+%during such INCs' execution for some time. 
+It is conjectured that for a longer INC program, it is more frequent to have {\tt java} with some positive time in PT.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|l|l|l|l|} \hline
+Iteration number &  Process ID & Process Name  & PT (in msec)\\ \hline
+795 (, 125, or 236) & 16256 & {\tt incr\_work} (=INC64) & 16,142 \\ \hline
+	& 28525 & {\tt proc\_monitor} & 198 \\ \hline
+	& 167 & {\tt kslowd001} &  4 \\ \hline
+	& 166 & {\tt kslowd000} &  3 \\ \hline 
+(125) & (472) & ({\tt md0\_raid1}) & (2) \\ \hline
+    & 16246 & {\tt java} &  2 \\ \hline 
+\hline 
+759 & 16256 & {\tt incr\_work} (=INC64) & 64,151 \\ \hline
+& 28525 & {\tt proc\_monitor} & 198 \\ \hline
+& 19320 & {\tt sshd} & 13 \\ \hline
+& 19329 & {\tt sshd}  & 12 \\ \hline
+& 19331 & {\tt grep}  &  5 \\ \hline
+& 19322 & {\tt grep}  &  5 \\ \hline
+& 167 & {\tt kslowd001} & 3\\ \hline
+& 166 & {\tt kslowd000} &  3 \\ \hline
+& 16246 & {\tt java} &  2 \\ \hline
+& 2067 & {\tt sshd}  &  2 \\ \hline
+& 19335 & {\tt grep}  &  1 \\ \hline
+& 19330 & {\tt sshd}  &  1 \\ \hline
+& 472 & {\tt md0\_raid1} &  1 \\ \hline
+& 19326 & {\tt grep}  &  1 \\ \hline
+\hline
+\end{tabular}
+\end{center}
+\caption{Daemons observed at the second-rightmost and rightmost bins in Figure~\ref{fig:inc64_r2_hist_v5}~\label{tab:inc64_daemons}}
+\end{table}
+
+\clearpage
+\newpage

### 2018-09-27 pt_characterization/machine_cons_check/post_exp.tex
ing...

@@
 I examined in each of the four runs a few iterations at which more than three daemons (except INC and proc monitor processes) that had positive PT were captured.  
-From Table~\ref{fig:daemon}, I suspect that PT distribution seems most likely to be affected by two facts: {\it how longer the same daemon ran than usual}, and {\it how many different daemon processes appeared and how long it ran}.
+From Table~\ref{tab:daemon}, I suspect that PT distribution seems most likely to be affected by two facts: {\it how longer the same daemon ran than usual}, and {\it how many different daemon processes appeared and how long it ran}.
 
@@
   }
- \caption{Some daemon processes captured across different machines~\label{fig:daemon}}
+ \caption{Some daemon processes captured across different machines~\label{tab:daemon}}
 \end{table}
+
+Similarly, we looked at the data from INC16 on {\tt sodb12}.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
+Machine & Task Length & Description & Time Length\\ \hline
+{\tt sodb12}  (plugged into the {\em upper right} power strip) &  INC128 & Runs of 1,000 samples & 2017-02-09$\sim$2017-02-10\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Detailed description of INC128 data\label{tab:exp_notes5}}
+\end{table}
+
+Table~\ref{tab:daemon2} shows a list of daemon processes observed at some iterations for which more than three daemons revealed positive PT while the INC16 program was run.
+\begin{table}[htp!]
+\centering
+{
+ \begin{tabular}{|p{1.5cm}|p{2cm}|p{12.5cm}|} \hline
+Machine Name & Iteration \# & Process Name (id, PT(msec))\\ \hline
+ {\tt sodb12}  & 33 & {\tt flush-9:0} ({\tt 25187}, 1), {\tt java} ({\tt 32627}, 3), {\tt java} ({\tt 304}, 1), {\tt md0\_raid1} ({\tt 484}, 1) \\ \cline{2-3}
+					  & 138  & {\tt flush-9:0} ({\tt 25187}, 1), {\tt java} ({\tt 734}, 1), {\tt java} ({\tt 32627}, 3), {\tt md0\_raid1} ({\tt 484}, 1)  \\ \cline{2-3}
+					  & 181 & {\tt flush-9:0} ({\tt 25187}, 1), {\tt java} ({\tt 908}, 1), {\tt java} ({\tt 32627}, 3), {\tt jbd2/md0-8} ({\tt 497}, 1) \\ \cline{2-3}
+				     & 228 &  {\tt flush-9:0} ({\tt 25187}, 1), {\tt java} ({\tt 1099}, 1), {\tt java} ({\tt 32627}, 3), {\tt md0\_raid1} ({\tt 484}, 1)\\ \cline{2-3}
+ \hline
+ \end{tabular}
+  }
+ \caption{Some daemon processes captured across different machines~\label{tab:daemon2}}
+\end{table}

### 2018-09-27 pt_characterization/machine_cons_check/post_exp.tex
Rick's concerns being handled

--- 2018-09-14 pt_characterization/machine_cons_check/post_exp.tex
@@
   }
- \caption{Some daemon processes captured across different machines~\label{fig:daemon}}
+ \caption{Some daemon processes captured across different machines~\label{tab:daemon}}
 \end{table}
+
+Similarly, we looked at the data from INC128 on {\tt sodb12}.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
+Machine & Task Length & Description & Time Length\\ \hline
+{\tt sodb12}  (plugged into the {\em upper right} power strip) &  INC128 & Runs of 1,000 samples & 2017-02-09$\sim$2017-02-10\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Detailed description of the INC128 data\label{tab:exp_notes5}}
+\end{table}
+
+Table~\ref{tab:daemon2} shows a list of daemon processes observed at some iterations for which more than three daemons revealed positive PT while the INC128 program was run.
+\begin{table}[htp!]
+\centering
+{
+ \begin{tabular}{|p{1.5cm}|p{2cm}|p{12.5cm}|} \hline
+Machine Name & Iteration \# & Process Name (id, PT(msec))\\ \hline
+ {\tt sodb12}  & 33 & {\tt flush-9:0} ({\tt 25187}, 1), {\tt java} ({\tt 304}, 1), {\tt java} ({\tt 32627}, 3), {\tt md0\_raid1} ({\tt 484}, 1) \\ \cline{2-3}
+					  & 138  & {\tt flush-9:0} ({\tt 25187}, 1), {\tt java} ({\tt 734}, 1), {\tt java} ({\tt 32627}, 3), {\tt md0\_raid1} ({\tt 484}, 1)  \\ \cline{2-3}
+					  & 181 & {\tt flush-9:0} ({\tt 25187}, 1), {\tt java} ({\tt 908}, 1), {\tt java} ({\tt 32627}, 3), {\tt jbd2/md0-8} ({\tt 497}, 1) \\ \cline{2-3}
+				     & 228 &  {\tt flush-9:0} ({\tt 25187}, 1), {\tt java} ({\tt 1099}, 1), {\tt java} ({\tt 32627}, 3), {\tt md0\_raid1} ({\tt 484}, 1)\\ \cline{2-3}
+ \hline
+ \end{tabular}
+  }
+ \caption{Some daemon processes captured at some iterations of INC128~\label{tab:daemon2}}
+\end{table}
+
+Table~\ref{tab:exp_notes6} gives a background description of what environment 
+was set for running the MATR8K program.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
+Machine & Task Length & Description & Time Length\\ \hline
+{\tt sodb12} (plugged into {\em the upper right} power strip) & MATR8K& Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Detailed description of the MATR8K data\label{tab:exp_notes6}}
+\end{table}
+
+Table~\ref{tab:daemon3} shows a list of daemon processes 
+observed at some iterations for which a number of daemons revealed positive PT while the MATR8K program was run.
+\begin{table}[htp!]
+\centering
+{ %{\tt bash} ({\tt 17403}, 1), {\tt bash} ({\tt 17400}, 7),  
+ \begin{tabular}{|p{1.5cm}|p{2cm}|p{12.5cm}|} \hline
+Machine Name & Iteration \# & Process Name (id, PT(msec))\\ \hline
+ {\tt sodb12}  & 
+ 36 & ({\tt MATR8K} ({\tt 17244}, 2183274),{\tt proc\_monitor} ({\tt 14166}, 200)), 
+ {\tt kslowd001} ({\tt 167}, 16959),  {\tt kslowd000} ({\tt 166}, 16707), 
+ {\tt md0\_raid1 } ({\tt 482}, 19),  {\tt sshd} ({\tt 17398}, 18), {\tt bash}  ({\tt 17400}, 7),  
+ {\tt flush-9:0} ({\tt 718}, 3), {\tt java} ({\tt 17234}, 2), {\tt kblockd/0} ({\tt 16}, 2), 
+ {\tt grep} ({\tt 17406}, 1), {\tt sshd} ({\tt 17399}, 1), 
+ {\tt events/0} ({\tt 7}, 1), {\tt bash} ({\tt 17403}, 1))\\ \cline{2-3} 
+&  119 & ({\tt MATR8K} ({\tt 17244}, 2183217),{\tt proc\_monitor} ({\tt 14166}, 204))
+ {\tt kslowd001} ({\tt 167}, 16784),  {\tt kslowd000} ({\tt 166}, 16839), 
+ {\tt md0\_raid1 } ({\tt 482}, 16),  {\tt  jbd2/md0-8} ({\tt 495}, 13), 
+ {\tt sshd} ({\tt 17773}, 7), {\tt sshd} ({\tt 17771}, 7), 
+ {\tt kblockd/0} ({\tt 16}, 4),  {\tt events/0} ({\tt 7}, 3), 
+ {\tt flush-9:0} ({\tt 718}, 3), {\tt java} ({\tt 17234}, 2), 
+ {\tt sshd} ({\tt 17772}, 2),  {\tt sshd} ({\tt 2148}, 1), 
+ {\tt java} ({\tt 17768}, 2), {\tt sshd} ({\tt 17774}, 1)\\ \cline{2-3}
+ \hline
+ \end{tabular}
+  }
+ \caption{Some daemon processes captured at some iterations of MATR8K~\label{tab:daemon3}}
+\end{table}

### 2018-09-28 pt_characterization/machine_cons_check/daemon_analysis/paper.tex
Rick's request

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
 \usepackage{auto-pst-pdf}
\usepackage{graphicx}
%\usepackage{arydshln}
\usepackage{ifpdf}
\ifpdf
  \usepackage{epstopdf}
\fi
\usepackage{multirow}
\usepackage{epsfig}
\usepackage{float}
\usepackage{url}
\usepackage{color}
\usepackage{subfigure}

%\newcommand\solidrule[1][1cm]{\rule[0.5ex]{#1}{.4pt}}
%\newcommand\dashedrule{\mbox{%
%  \solidrule[2mm]\hspace{2mm}\solidrule[2mm]\hspace{2mm}\solidrule[2mm]}}

%\usepackage{hyperref}

\begin{document}
\title{Daemon Analysis across Different Machines}

\author{
Young-Kyoon Suh\\
Data \& Knowledge Engineering Laboratory\\
School of Computer Science and Engineering\\
Kyungpook National University\\
}
\maketitle

\section{Investigation on Some Daemons' Appearance Across Machines}

\begin{table}[htp!]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 INC1 &  1,000 & kslowd000 & sodb8 & 1 & 1 & 1 & 19 \\ \hline
 &   &  & sodb9 & 1 & 1 & 1 & 15 \\ \hline
 &   &  & sodb10 & 1 & 1 & 1 & 19 \\ \hline 
 \hline
  &  & kslowd001 & sodb8 & 1 & 1 & 1 & 15 \\ \hline
 &   &  & sodb9 & 1 & 1 & 1 & 18 \\ \hline
 &   &  & sodb10 & 1 & 1 & 1 & 15 \\ \hline 
 %&   &  & sodb12 & 1 & 1 & 1 & 19 \\ \hline 
%  \hline
 \end{tabular}
  }
 \caption{Daemon statistics on the INC1 data across machines~\label{tab:inc1}}
\end{table}

\begin{table}[htp!]
\centering
{\small
 \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
 Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
 MAT8K & 300 & grep & sodb12 & 1 & 1 & 1 & 4 \\ \hline
 \hline
  &  & kslowd000  & sodb12 & 16540 & 17033 & 16780.43 & 300 \\ \hline
 &  & kslowd000	& sodb10 & 102 & 107 & 103.33 & 300 \\ \hline 
 \hline
  &  & kslowd001  & sodb12 & 16541 & 17000 & 16777.18 & 300 \\ \hline
  &  & kslowd001	& sodb10 & 102 & 108 & 103.27 & 300 \\ \hline 
 \hline
  &  & sshd  & sodb12 & 1 & 20 & 5.91 & 43 \\ \hline
 &  & sshd & sodb10 & 1 & 8  &  4.17 & 42 \\ \hline 
%   \hline
 \end{tabular}
  }
 \caption{Daemon statistics on the MAT8K data across machines~\label{tab:mat8K}}
\end{table}

\end{document}

### 2018-09-29 pt_characterization/machine_cons_check/daemon_analysis/paper.tex
Daemon analysis done

@@
 
-\begin{table}[htp!]
-\centering
-{\small
- \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
- Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
- INC1 &  1,000 & kslowd000 & sodb8 & 1 & 1 & 1 & 19 \\ \hline
- &   &  & sodb9 & 1 & 1 & 1 & 15 \\ \hline
- &   &  & sodb10 & 1 & 1 & 1 & 19 \\ \hline 
- \hline
-  &  & kslowd001 & sodb8 & 1 & 1 & 1 & 15 \\ \hline
- &   &  & sodb9 & 1 & 1 & 1 & 18 \\ \hline
- &   &  & sodb10 & 1 & 1 & 1 & 15 \\ \hline 
- %&   &  & sodb12 & 1 & 1 & 1 & 19 \\ \hline 
-%  \hline
+\subsection{INC1, INC16, and INC128}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
+Machine & Task Length & Description & Time Length\\ \hline
+{\tt sodb8} (plugged into the {\em bottom left} power strip) & INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb9}  (plugged into the {\em upper left} power strip)  & INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb10}  (plugged into the {\em upper left} power strip)  & INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb12}  (plugged into the {\em upper right} power strip)  & INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
 \end{tabular}
-  }
- \caption{Daemon statistics on the INC1 data across machines~\label{tab:inc1}}
+\end{center}
+\vspace{-.2in}
+\caption{Detailed description of the INC1, INC16, INC128 data used for histograms\label{tab:exp_notesINCs}}
 \end{table}
@@
  Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
- MAT8K & 300 & grep & sodb12 & 1 & 1 & 1 & 4 \\ \hline
+ INC1 &  1,000 & {\tt kslowd000} & {\tt sodb8} & 1 & 1 & 1 & 19 \\ \hline
+ &   &  & {\tt sodb9} & 1 & 1 & 1 & 15 \\ \hline
+ &   &  & {\tt sodb10} & 1 & 1 & 1 & 19 \\ \hline 
  \hline
-  &  & kslowd000  & sodb12 & 16540 & 17033 & 16780.43 & 300 \\ \hline
- &  & kslowd000	& sodb10 & 102 & 107 & 103.33 & 300 \\ \hline 
- \hline
-  &  & kslowd001  & sodb12 & 16541 & 17000 & 16777.18 & 300 \\ \hline
-  &  & kslowd001	& sodb10 & 102 & 108 & 103.27 & 300 \\ \hline 
- \hline
-  &  & sshd  & sodb12 & 1 & 20 & 5.91 & 43 \\ \hline
- &  & sshd & sodb10 & 1 & 8  &  4.17 & 42 \\ \hline 
+  &  & {\tt kslowd001} & {\tt sodb8} & 1 & 1 & 1 & 15 \\ \hline
+ &   &  & {\tt sodb9} & 1 & 1 & 1 & 18 \\ \hline
+ &   &  & {\tt sodb10} & 1 & 1 & 1 & 15 \\ \hline 
+ %&   &  & {\tt sodb12} & 1 & 1 & 1 & 19 \\ \hline 
 %  \hline
@@
   }
- \caption{Daemon statistics on the MAT8K data across machines~\label{tab:mat8K}}
+ \caption{\textbf{The statistics of the daemons} on the INC1 data across machines~\label{tab:inc1}}
+\end{table}
+
+\begin{table}[htp!]
+\centering
+{\small
+ \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
+ Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
+ INC16 &  1,000 & {\tt kslowd000} & {\tt sodb8} & 1 & 1 & 1 & 243 \\ \hline
+ &   &  & {\tt sodb9} & 1 & 2 & 1.01 & 242 \\ \hline
+ &   &  & {\tt sodb10} & 1 & 1 & 1 & 241 \\ \hline 
+ \hline
+  &  & {\tt kslowd001} & {\tt sodb8} & 1 & 1 & 1 & 243 \\ \hline
+ &   &  & {\tt sodb9} & 1 & 2 & 1 & 241 \\ \hline
+ &   &  & {\tt sodb10} & 1 & 1 & 1 & 243 \\ \hline 
+ %&   &  & {\tt sodb12} & 1 & 1 & 1 & 19 \\ \hline 
+%  \hline
+ \end{tabular}
+  }
+ \caption{The statistics of the daemons on the INC16 data across machines~\label{tab:inc16}}
+\end{table}
+
+\begin{table}[H]
+\centering
+{\small
+ \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
+ Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
+ INC128 &  300 & {\tt kslowd000} & {\tt sodb8} & 6 & 8 & 6.43 & 300 \\ \hline
+ &   &  & {\tt sodb9} & 6 & 7 & 6.42 & 300 \\ \hline
+ &   &  & {\tt sodb10} & 6 & 7 & 6.42 & 300 \\ \hline 
+ \hline
+  &  & {\tt kslowd001} & {\tt sodb8} & 6 & 7 & 6.43 & 300 \\ \hline
+ &   &  & {\tt sodb9} & 6 & 8 & 6.42 & 300 \\ \hline
+ &   &  & {\tt sodb10} & 6 & 8 & 6.43 & 300 \\ \hline 
+ %&   &  & {\tt sodb12} & 1 & 1 & 1 & 19 \\ \hline 
+%  \hline
+ \end{tabular}
+  }
+ \caption{The statistics of the daemons on the INC128 data across machines~\label{tab:inc128}}
+\end{table}
+
+\clearpage
+\newpage
+\subsection{INC1024}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
+Machine & Task Length & Description & Time Length\\ \hline
+{\tt sodb9} (plugged into {\em the upper left} power strip) & INC1024 & 300 samples, each & 2017-03-17 $\sim$ 2017-03-21  \\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Detailed description of the INC1024 data \label{tab:exp_notes1024}}
+\end{table}
+
+\begin{table}[htp!]
+\centering
+{\small
+ \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
+ Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
+ INC1024 &  300 & {\tt grep} & {\tt sodb9} & 1 & 7 & 3.23 & 13 \\ \hline
+ \hline
+  &   & {\tt kslowd000} &  & 51 & 53 & 51.37 & 300 \\ \hline
+ \hline
+  &  & {\tt kslowd001} &  & 51 & 53 & 51.35 & 300 \\ \hline
+ \hline
+  &  & {\tt sshd} & & 1 & 18 & 7.55 & 20 \\ \hline
+ \end{tabular}
+  }
+ \caption{The statistics of the daemons on the INC1024 data across machines~\label{tab:inc1024}}
+\end{table}
+
+\clearpage
+\newpage
+\subsection{INC16384}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{4cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
+Machine & Task Length & Description & Time Length\\ \hline
+{\tt sodb12} (plugged into {\em the upper right} power strip) & INC16384& 130 samples & 2017-04-27 $\sim$ 2017-05-21\\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Detailed description of the INC16384 data \label{tab:exp_notes16384}}
+\end{table}
+
+\begin{table}[htp!]
+\centering
+{\small
+ \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
+ Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
+ INC16384 & 130 & {\tt grep} & {\tt sodb12} & 1 & 9 & 3.21& 105 \\ \hline
+ \hline
+  &  & {\tt sshd} & {\tt sodb12} & 1 & 20 & 8.11 & 140 \\ \hline
+ \end{tabular}
+  }
+ \caption{The statistics of the daemons on the INC16384 data across machines~\label{tab:inc16384}}
+\end{table}
+
+\clearpage
+\newpage
+\subsection{SORT100, SORT800, MAT1K, and MAT8K}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{4cm}|p{4cm}|p{4cm}|p{4cm}|} \hline
+Machine & Task Length & Description & Time Length\\ \hline
+{\tt sodb9}  (plugged into {\em the upper left} power strip) & SORT100, SORT800 & Runs of 300 samples & 2018-04-09 $\sim$2018-04-12\\ \hline
+%{\tt sodb10} (plugged into {\em the upper left} power strip)  & MATC1K, MATC8K & Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
+{\tt sodb10} (plugged into {\em the upper left} power strip)  & MAT1K, MAT8K & Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
+%{\tt sodb12} (plugged into {\em the upper right} power strip) & MATR1K, MATR8K & Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
+{\tt sodb12} (plugged into {\em the upper right} power strip) & MAT1K, MAT8K & Runs of 300 samples & 2018-03-02 $\sim$2018-03-12 \\ \hline
+\end{tabular}
+\end{center}
+\vspace{-.2in}
+\caption{Detailed description of the SORT, MAT data\label{tab:exp_notes_sort_mat}}
+\end{table}
+
+\begin{table}[htp!]
+\centering
+{\small
+ \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
+ Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
+ SORT100 & 300 & {\tt kslowd000}  & {\tt sodb9} & 1 & 1 & 1 & 64 \\ \hline
+  &  & {\tt kslowd001}  & 	& 1 & 2 & 1.02 & 63  \\ \hline
+  &  & {\tt sshd}  & & 2 & 8 & 5 & 4 \\ \hline
+ \end{tabular}
+  }
+ \caption{The statistics of the daemons on the SORT100 data across machines~\label{tab:sort100}}
+\end{table}
+
+\begin{table}[htp!]
+\centering
+{\small
+ \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
+ Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
+ SORT800 & 300 & {\tt kslowd000}  & {\tt sodb9} & 13 & 15 & 13.78 & 300 \\ \hline
+  &  & {\tt kslowd001}  & 	& 13 & 15 & 13.78 & 300  \\ \hline
+  &  & {\tt sshd}  & & 1 & 9 & 4.8 & 5 \\ \hline
+ \end{tabular}
+  }
+ \caption{The statistics of the daemons on the SORT800 data across machines~\label{tab:sort800}}
+\end{table}
+
+\begin{table}[htp!]
+\centering
+{\small
+ \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
+ Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
+ MAT1K & 300 & {\tt kslowd000}  & {\tt sodb12} & 38 & 161 &  150.85 & 52 \\ \hline
+ 			 &  	    & {\tt kslowd000}	 & {\tt sodb10} & 1 & 1 & 1 & 50 \\ \hline 
+ \hline
+  &  & {\tt kslowd001}  & {\tt sodb12} & 146 & 161 & 154.28 & 50 \\ \hline
+  &  & {\tt kslowd001}	& {\tt sodb10} & 1 & 1 & 1 & 52 \\ \hline 
+ \hline
+  &  & {\tt sshd}  & {\tt sodb12} & 1 & 7 & 4.25 & 4 \\ \hline
+ &  & {\tt sshd} & {\tt sodb10} & 1 & 7  &  4 & 4 \\ \hline 
+%   \hline
+ \end{tabular}
+  }
+ \caption{The statistics of the daemons on the MAT1K data across machines~\label{tab:mat1K}}
+\end{table}
+
+\begin{table}[htp!]
+\centering
+{\small
+ \begin{tabular}{|l|l|l|l|l||l|l|l|} \hline
+ Task & Total & Process & Machine & Min (ms) & Max (ms) & Avg (ms) & \# of Iters\\ \hline
+ MAT8K & 300 & {\tt grep} & {\tt sodb12} & 1 & 1 & 1 & 4 \\ \hline
+ \hline
+  &  & {\tt kslowd000}  & {\tt sodb12} & 16540 & 17033 & 16780.43 & 300 \\ \hline
+ &  & {\tt kslowd000}	& {\tt sodb10} & 102 & 107 & 103.33 & 300 \\ \hline 
+ \hline
+  &  & {\tt kslowd001}  & {\tt sodb12} & 16541 & 17000 & 16777.18 & 300 \\ \hline
+  &  & {\tt kslowd001}	& {\tt sodb10} & 102 & 108 & 103.27 & 300 \\ \hline 
+ \hline
+  &  & {\tt sshd}  & {\tt sodb12} & 1 & 20 & 5.91 & 43 \\ \hline
+ &  & {\tt sshd} & {\tt sodb10} & 1 & 8  &  4.17 & 42 \\ \hline 
+%   \hline
+ \end{tabular}
+  }
+ \caption{The statistics of the daemons on the MAT8K data across machines~\label{tab:mat8K}}
 \end{table}

### 2018-10-19 pt_characterization/machine_cons_check/daemon_analysis/paper.tex
Rob's docs

@@
 Machine & Task Length & Description & Time Length\\ \hline
-{\tt sodb8} (plugged into the {\em bottom left} power strip) & INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
-{\tt sodb9}  (plugged into the {\em upper left} power strip)  & INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
-{\tt sodb10}  (plugged into the {\em upper left} power strip)  & INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
-{\tt sodb12}  (plugged into the {\em upper right} power strip)  & INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb8} (plugged into the {\em bottom left} power strip) & INC1, INC16, INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb9}  (plugged into the {\em upper left} power strip)  & NC1, INC16, INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb10}  (plugged into the {\em upper left} power strip)  & NC1, INC16, INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
+{\tt sodb12}  (plugged into the {\em upper right} power strip)  & NC1, INC16, INC128 & Runs of 300 samples & 2017-02-09 $\sim$ 2017-02-10\\ \hline
 \end{tabular}
@@
   }
- \caption{\textbf{The statistics of the daemons} on the INC1 data across machines~\label{tab:inc1}}
+ \caption{The statistics of the daemons on the INC1 data across machines~\label{tab:inc1}}
 \end{table}
@@
 
-\clearpage
-\newpage
+%\clearpage
+%\newpage
 \subsection{INC1024}
@@
 
-\clearpage
-\newpage
+%\clearpage
+%\newpage
 \subsection{INC16384}

### 2018-12-12 pt_characterization/protocol/intermediate_data/paper.tex
New experiments

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
 \usepackage{auto-pst-pdf}
\usepackage{graphicx}
%\usepackage{arydshln}
\usepackage{ifpdf}
\ifpdf
  \usepackage{epstopdf}
\fi
\usepackage{multirow}
\usepackage{epsfig}
\usepackage{float}
\usepackage{url}
\usepackage{color}
\usepackage{subfigure}

%\newcommand\solidrule[1][1cm]{\rule[0.5ex]{#1}{.4pt}}
%\newcommand\dashedrule{\mbox{%
%  \solidrule[2mm]\hspace{2mm}\solidrule[2mm]\hspace{2mm}\solidrule[2mm]}}

%\usepackage{hyperref}

\begin{document}
\title{A Study of INC Runs in Sequence}

\author{
Young-Kyoon Suh\\
}
\maketitle

\section{Description}
In this document we discuss characteristics of INC runs in sequence.

There are two big subjects included in the document.
Specifically, the first subject relates running a series of INCs in a batch, pausing the run, and then resuming 
the run, rather than repeating the same INC multiple times. 
That is, INC1, ..., INC16384 are consecutively run as a batch, and 
then the entire batch run sleeps for an hour, and then 
another batch run of INC1, ..., INC16384 begins, stops, and resumes, and so forth.

The second subject concerns breaking up 
a whole run of 1000 samples into 200 samples, thereby yielding five different histograms.

\subsection{Experiment Notes}
Table~\ref{tab:exp_notes} provides a short description of our experimental runs, 
on which the following histograms are based.

%% done
\begin{table}[h]
\begin{center}
\begin{tabular}{|p{2cm}|p{3cm}|p{6cm}|p{4cm}|} \hline
Machine & Task Length (sec) & Description & Experiment Period\\ \hline
{\tt sodb9} &  INC16 & Each run with 1,000 samples & 2017-03-02 $\sim$ 2017-03-07\\ \hline
{\tt sodb9} &  INC1~INC16384 & One run in which each INC has 30 samples & 2018-11-17 $\sim$ 2018-12-02\\ \hline
\end{tabular}
\end{center}
\vspace{-.2in}
\caption{Notes on experiment runs used for histograms\label{tab:exp_notes}}
\end{table}

\newpage
\section{INC Run in Sequence}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC1 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{1_sec_pt_hist.eps}
		\label{fig:inc1}
	}
	\subfigure[PT frequency on INC2 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{2_sec_pt_hist.eps}
		\label{fig:inc2}
	}
	\subfigure[PT frequency on INC4 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{4_sec_pt_hist.eps}
		\label{fig:inc4}
	}
	\subfigure[PT frequency on INC8 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{8_sec_pt_hist.eps}
		\label{fig:inc8}
	}
	\caption{PT Histograms of INC1 ... INC8~\label{fig:s9_r2_pt_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC16 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{16_sec_pt_hist.eps}
		\label{fig:inc16}
	}
	\subfigure[PT frequency on INC32 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{32_sec_pt_hist.eps}
		\label{fig:inc32}
	}
	\subfigure[PT frequency on INC64 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{64_sec_pt_hist.eps}
		\label{fig:inc64}
	}
	\caption{PT Histograms of INC16 ... INC64\label{fig:s9_r2_pt_hist2}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC128 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{128_sec_pt_hist.eps}
		\label{fig:inc128}
	}
	\subfigure[PT frequency on INC256 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{256_sec_pt_hist.eps}
		\label{fig:inc256}
	}
	\subfigure[PT frequency on INC512 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{512_sec_pt_hist.eps}
		\label{fig:inc512}
	}
	\subfigure[PT frequency on INC1024 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{1024_sec_pt_hist.eps}
		\label{fig:inc1024}
	}
	\caption{PT Histograms of INC256 ... INC1024~\label{fig:s9_r2_pt_hist3}}
\end{figure}

\begin{figure}[t]
	\centering
	\subfigure[PT frequency on INC2048 on {\tt sodb10}]{
		\includegraphics[scale=0.43]{2048_sec_pt_hist.eps}
		\label{fig:inc2048}
	}
	\subfigure[PT frequency on INC4096 on {\tt sodb12}]{
		\includegraphics[scale=0.43]{4096_sec_pt_hist.eps}
		\label{fig:inc4096}
	}
	\subfigure[PT frequency on INC8192 on {\tt sodb10}]{
		\includegraphics[scale=0.43]{8192_sec_pt_hist.eps}
		\label{fig:inc8192}
	}
	\subfigure[PT frequency on INC16384 on {\tt sodb12}]{
		\includegraphics[scale=0.43]{16384_sec_pt_hist.eps}
		\label{fig:inc16384}
	}
	\caption{PT Histograms of INC2048 ... INC16384~\label{fig:s9_r2_pt_hist4}}
\end{figure}

\newpage
\section{Decomposition of INC16 Samples}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC16 with the first 200 samples on {\tt sodb9}]{
		\includegraphics[scale=0.43]{16_sec_pt_hist0_1.eps}
		\label{fig:inc16-1}
	}
	\subfigure[PT frequency on INC16 with the next 200 samples on {\tt sodb9}]{
		\includegraphics[scale=0.43]{16_sec_pt_hist0_2.eps}
		\label{fig:inc16-2}
	}
	\subfigure[PT frequency on INC16 with the next 200 samples on {\tt sodb9}]{
		\includegraphics[scale=0.43]{16_sec_pt_hist0_3.eps}
		\label{fig:inc16-3}
	}
	\caption{PT Histograms of INC16-1 ~ INC16-3~\label{fig:inc16_hist1}}
\end{figure}

\begin{figure}[hp!]
	\centering
	\subfigure[PT frequency on INC16 with the next 200 samples on {\tt sodb9}]{
		\includegraphics[scale=0.43]{16_sec_pt_hist0_4.eps}
		\label{fig:inc16-4}
	}
	\subfigure[PT frequency on INC16 with the next 200 samples on {\tt sodb9}]{
		\includegraphics[scale=0.43]{16_sec_pt_hist0_5.eps}
		\label{fig:inc16-5}
	}
	\caption{PT Histograms of INC16-4 ~ INC16-5\label{fig:inc16_hist2}}
\end{figure}

\end{document}

### 2018-12-12 pt_characterization/protocol/intermediate_data/pt_hist_sodb9.r

@@ -1,0 +1,867 @@
+numSamples <- 30
+x = read.csv(file="1_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("1_sec_pt_hist.eps")
+binsize=1
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmin
+xmax <-max(x$PRTIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <- 15
+plot(h, xaxt="n",axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green", main='PT frequency on INC1', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+numSamples <- 30
+x = read.csv(file="2_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("2_sec_pt_hist.eps")
+binsize=1
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmin
+xmax <-max(x$PRTIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <-25
[…]

### 2018-12-15 pt_characterization/protocol/intermediate_data/paper.tex
Doc for new runs

@@
 	\centering
-	\subfigure[PT frequency on INC2048 on {\tt sodb10}]{
+	\subfigure[PT frequency on INC2048 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{2048_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC4096 on {\tt sodb12}]{
+	\subfigure[PT frequency on INC4096 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{4096_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC8192 on {\tt sodb10}]{
+	\subfigure[PT frequency on INC8192 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{8192_sec_pt_hist.eps}
@@
 	}
-	\subfigure[PT frequency on INC16384 on {\tt sodb12}]{
+	\subfigure[PT frequency on INC16384 on {\tt sodb9}]{
 		\includegraphics[scale=0.43]{16384_sec_pt_hist.eps}
@@
 
-\newpage
+\clearpage
 \section{Decomposition of INC16 Samples}
 
-\begin{figure}[hp!]
+\begin{figure}[htp!]
 	\centering
@@
 	}
-	\caption{PT Histograms of INC16-1 ~ INC16-3~\label{fig:inc16_hist1}}
-\end{figure}
-
-\begin{figure}[hp!]
-	\centering
 	\subfigure[PT frequency on INC16 with the next 200 samples on {\tt sodb9}]{
@@
 	}
+	\caption{PT Histograms of INC16-1$~$INC16-4~\label{fig:inc16_hist1}}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
 	\subfigure[PT frequency on INC16 with the next 200 samples on {\tt sodb9}]{
@@
 	}
-	\caption{PT Histograms of INC16-4 ~ INC16-5\label{fig:inc16_hist2}}
+	\caption{PT Histograms of INC16-5\label{fig:inc16_hist2}}
 \end{figure}

### 2018-12-15 pt_characterization/protocol/intermediate_data/pt_hist_sodb9.r

@@ -118,6 +118,246 @@ xlab='PT (ms)', ylab=expression('Frequency'))
 axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
 dev.off()
 
+numSamples <- 30
+x = read.csv(file="32_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("32_sec_pt_hist.eps")
+binsize=1
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmin
+xmax <-max(x$PRTIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <-10
+plot(h, xaxt="n",axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green", main='PT frequency on INC32', 
+sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), 
+xlab='PT (ms)', ylab=expression('Frequency'))
+axis(side=1,at=h$mids,labels=seq(xmin,xmax,1))
+dev.off()
+
+numSamples <- 30
+x = read.csv(file="64_sec.dat",head=TRUE,sep="\t")
+x <- subset(x, x$ITERNUM <= numSamples)
+setEPS()
+postscript("64_sec_pt_hist.eps")
+binsize=1
+x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+xmin <-min(x$PRTIME)
+xmin
+xmax <-max(x$PRTIME)
+xmax
+nbins <- ceiling((xmax-xmin) / binsize)
+h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F)
+ymax <- max(h$counts)
+ymax <- ceiling(ymax/100)*100
+ymax <-6
[…]

### 2018-12-15 pt_characterization/protocol/intermediate_data/paper.tex
Slightly modified

@@
 \section{Description}
-In this document we discuss characteristics of INC runs in sequence.
+In this document we study the characteristics of INC runs that are executed in sequence.
 
-There are two big subjects included in the document.
-Specifically, the first subject relates running a series of INCs in a batch, pausing the run, and then resuming 
-the run, rather than repeating the same INC multiple times. 
-That is, INC1, ..., INC16384 are consecutively run as a batch, and 
-then the entire batch run sleeps for an hour, and then 
-another batch run of INC1, ..., INC16384 begins, stops, and resumes, and so forth.
+Broadly speaking, there are two {\em subjects} included in the document.
+The first subject, which is discussed in Section 1, involves {\it i}) running as a {\em batch} a series of INCs with increasing task length and {\it ii}) pausing the batch run.
+This entire batch run including the pause continues until the specified number of iterations, say 30, is specified.
+More specifically, INC1, ..., INC16384 are consecutively run as a batch, and
+then the entire batch run sleeps for an hour, and then another batch run begins and stops, and again the new batch run resumes.
+This type of batch run is distinct from repeatedly running the same INC with a single task length.
 
-The second subject concerns breaking up 
-a whole run of 1000 samples into 200 samples, thereby yielding five different histograms.
+The second subject, which is discussed in Section 2, iconcerns dividing 
+a single run of 1000 samples on INC16 into five consecutive subruns of 200 samples.  
+For each subrun, we examine their respective PT (program time) histograms 
+and see if the histograms look similar to each other.
 
@@
 {\tt sodb9} &  INC16 & Each run with 1,000 samples & 2017-03-02 $\sim$ 2017-03-07\\ \hline
-{\tt sodb9} &  INC1~INC16384 & One run in which each INC has 30 samples & 2018-11-17 $\sim$ 2018-12-02\\ \hline
+{\tt sodb9} &  INC1$\sim$INC16384 & One run in which each INC has 30 samples & 2018-11-17 $\sim$ 2018-12-02\\ \hline
 \end{tabular}
@@
 	}
-	\caption{PT Histograms of INC16-1$~$INC16-4~\label{fig:inc16_hist1}}
+	\caption{PT Histograms of INC16-1$\sim$INC16-4~\label{fig:inc16_hist1}}
 \end{figure}

### 2019-04-04 pt_characterization/protocol/intermediate_data/res2-backup.tex
Up-to-date version adding some intermediate runs
Rick's requests

--- 2018-09-19 pt_characterization/rep_check/res2.tex
@@
 This section exhibits histograms on the second run of 
-INC with its task length increasing from 1 second to 4096 seconds, via EMPv5 without Step 2. 
+INC with its task length increasing from 1 second to 4096 seconds, via EMPv5 without Step 2, except 
+some intermediate runs applied by EMPv4. 
 The detailed description of the base data is from Table~\ref{tab:exp_notes2}.
@@
 \begin{center}
-\begin{tabular}{|p{2cm}|p{3cm}|p{3cm}|p{4cm}|p{3.5cm}|} \hline
+\begin{tabular}{|p{2cm}|p{3cm}|p{3cm}|p{4cm}|p{4cm}|} \hline
 Machine & Task Length (sec) & Description & Experiment Period & Relevant \linebreak Histograms\\ \hline
-{\tt sodb9} (plugged into {\em the upper left} power strip) &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-13 $\sim$ 2017-03-14 & Figs.~\ref{fig:s9_r2_et_hist1},~\ref{fig:s9_r2_et_hist2},~\ref{fig:s9_r2_pt_hist1}, and~\ref{fig:s9_r2_pt_hist2}\\ \hline
-%{\tt sodb9} &  INC128$\sim$INC1024 & 300 samples, each & 2017-03-14 $\sim$ 2017-03-21 & 
-%Figs.~\ref{fig:s9_r2_et_hist3} and~\ref{fig:s9_r2_pt_hist3}\\ \hline
+{\tt sodb9} (plugged into {\em the upper left} power strip) &  INC1$\sim$INC64 & 1,000 samples, each & 2017-03-13 $\sim$ 2017-03-14 & Figs.~\ref{fig:s9_r2_et_hist1},~\ref{fig:s9_r2_et_hist2},~\ref{fig:s9_r2_pt_hist1},~\ref{fig:s9_r2_pt_hist2},~\ref{fig:s9_r2_pt_hist2-2},~\ref{fig:s9_r2_pt_hist3}\\ \hline
+
+{\tt sodb9} (plugged into {\em the upper left} power strip) &  INC19 and INC20 & 1,000 samples, each, applied by EMPv4 & 2018-12-03 $\sim$ 2018-12-03 & Fig.~\ref{fig:s9_r2_pt_hist2}\\ \hline
+
+{\tt sodb9} (plugged into {\em the upper left} power strip) &  INC21.125, INC60, and INC62 & 1000 samples, each, applied by EMPv4 & 2019-03-04 $\sim$ 2019-03-14 & Figs.~\ref{fig:s9_r2_pt_hist2} and~\ref{fig:s9_r2_pt_hist2-2}\\ \hline
+
+{\tt sodb9} (plugged into {\em the upper left} power strip) &  INC128$\sim$INC1024 & 300 samples, each & 2017-03-14 $\sim$ 2017-03-21 & 
+Figs.~\ref{fig:s9_r2_et_hist3}, ~\ref{fig:s9_r2_pt_hist3}, and~\ref{fig:inc1024_r2_hist_v5}\\ \hline
+
+{\tt sodb9} (plugged into {\em the upper left} power strip) &  INC224 & 1,000 samples, applied by EMPv4 & 2017-05-31 & Fig.~\ref{fig:inc224_v4}\\ \hline
+
 {\tt sodb9} (plugged into {\em the upper left} power strip) &  INC512$\sim$INC1024 & 300 samples, each & 2017-03-17 $\sim$ 2017-03-21 & 
-Figs.~\ref{fig:s9_r2_et_hist3} and~\ref{fig:s9_r2_pt_hist3}\\ \hline
+Figs.~\ref{fig:s9_r2_et_hist3},~\ref{fig:s9_r2_pt_hist3}, and~\ref{fig:inc1024_r2_hist_v5}\\ \hline
 {\tt sodb10} (plugged into {\em the upper left} power strip) & INC2048 & 300 samples & 2017-03-13 $\sim$ 2017-03-20 & Figs.~\ref{fig:inc2048_r2_et_hist_v5} and~\ref{fig:inc2048_r2_hist_v5}\\ \hline
@@
 
+%\begin{figure}[hp!]
+%	\centering
+%	\subfigure[PT frequency on INC16 on {\tt sodb9}]{
+%		\includegraphics[scale=0.43]{repet_data2/16_sec_pt_hist_v5.eps}
+%		\label{fig:inc16_r2_hist_v5}
+%	}
+%	\subfigure[PT frequency on INC32 on {\tt sodb9}]{
+%		\includegraphics[scale=0.43]{repet_data2/32_sec_pt_hist_v5.eps}
+%		\label{fig:inc32_r2_hist_v5}
+%	}
+%	\subfigure[PT frequency on INC64 on {\tt sodb9}]{
+%		\includegraphics[scale=0.43]{repet_data2/64_sec_pt_hist_v5.eps}
+%		\label{fig:inc64_r2_hist_v5}
+%	}
+%	\caption{PT Histograms of INC16 ... INC64\label{fig:s9_r2_pt_hist2}}
+%\end{figure}
+
 \begin{figure}[hp!]
@@
 	}
+	\subfigure[PT frequency on INC19 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data2/19_sec_pt_hist.eps}
+		\label{fig:inc19_hist_v4}
+	}
+	\subfigure[PT frequency on INC20 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data2/20_sec_pt_hist.eps}
+		\label{fig:inc20_hist_v4}
+	}
+	\subfigure[PT frequency on INC21.125 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data2/21_125_sec_pt_hist.eps}
+		\label{fig:inc21_125_hist_v4}
+	}
+	\caption{PT Histograms of INC16 ... INC32\label{fig:s9_r2_pt_hist2}}
+\end{figure}
+
+\begin{figure}[hp!]
+	\centering
 	\subfigure[PT frequency on INC32 on {\tt sodb9}]{
@@
 		\label{fig:inc32_r2_hist_v5}
+	}
+	\subfigure[PT frequency on INC60 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data2/60_0_sec_pt_hist.eps}
+		\label{fig:inc60_hist_v4}
+	}
+	\subfigure[PT frequency on INC62 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data2/62_0_sec_pt_hist.eps}
+		\label{fig:inc62_hist_v4}
 	}
@@
 	}
-	\caption{PT Histograms of INC16 ... INC64\label{fig:s9_r2_pt_hist2}}
+	\caption{PT Histograms of INC32 ... INC64\label{fig:s9_r2_pt_hist2-2}}
 \end{figure}
@@
 		\label{fig:inc128_r2_hist_v5}
+	}
+	\subfigure[PT frequency on INC224 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data2/224_sec_pt_hist.eps}
+		\label{fig:inc224_v4}
 	}
@@
 	}
+	\caption{PT Histograms of INC128 ... INC512~\label{fig:s9_r2_pt_hist3}}
+\end{figure}
+
+\begin{figure}[t]
+	\centering
 	\subfigure[PT frequency on INC1024 on {\tt sodb9}]{
@@
 	}
-	\caption{PT Histograms of INC256 ... INC1024~\label{fig:s9_r2_pt_hist3}}
-\end{figure}
-
-\begin{figure}[t]
-	\centering
 	\subfigure[PT frequency on INC2048 on {\tt sodb10}]{
@@
 	}
+	\caption{PT Histograms of INC1024 ... IINC4096~\label{fig:s9_r2_pt_hist4}}
+\end{figure}
+
+\begin{figure}[t]
+	\centering
 	\subfigure[PT frequency on INC8192 on {\tt sodb10}]{
@@
 	}
-	\caption{PT Histograms of INC2048 ... INC16384~\label{fig:s9_r2_pt_hist4}}
+	\caption{PT Histograms of INC8192 ... INC16384~\label{fig:s9_r2_pt_hist5}}
 \end{figure}

### 2019-04-04 pt_characterization/rep_check/res1.tex
Up-to-date version adding some intermediate runs
Rick's requests

--- 2018-05-04 pt_characterization/rep_check/res1.tex
@@
 {\tt sodb9} (plugged into {\em the upper left} power strip)  &  INC1$\sim$INC64 & 1000 samples, each & 2017-03-02 $\sim$ 2017-03-04 & Figs.~\ref{fig:s9_r1_et_hist1},~\ref{fig:s9_r1_et_hist2},~\ref{fig:s9_r1_pt_hist1}, and~\ref{fig:s9_r1_pt_hist2}\\ \hline
+
+{\tt sodb9} (plugged into {\em the upper left} power strip) &  INC19, INC20, INC21.125, INC60, INC62, INC224 & 1000 samples, each, applied by EMPv4 & 2017-05-31, 2018-12-03, 2019-03-04 $\sim$ 2019-03-14 & Figs.~\ref{fig:s9_r1_pt_hist2},~\ref{fig:s9_r1_pt_hist2-2}, and~\ref{fig:inc224_v4}\\ \hline
+
 {\tt sodb9} (plugged into {\em the upper left} power strip)  &  INC128$\sim$ INC1024 & 300 samples, each & 2017-03-04 $\sim$ 2017-03-11 & 
-Figs.~\ref{fig:s9_r1_et_hist3} and~\ref{fig:s9_r1_pt_hist3}\\ \hline
+Figs.~\ref{fig:s9_r1_et_hist3},~\ref{fig:s9_r1_pt_hist3}, and~\ref{fig:s9_r1_pt_hist4}\\ \hline
+
 {\tt sodb10} (plugged into {\em the upper left} power strip)  & INC2048 & 300 samples & 2017-03-02 $\sim$ 2017-03-09 & Figs.~\ref{fig:inc2048_r1_et_hist_v5} and~\ref{fig:inc2048_r1_hist_v5}\\ \hline
-{\tt sodb12} (plugged into {\em the upper left} power strip)  & INC4096 & 300 samples & 2017-02-13 $\sim$ 2017-02-27 & Figs.~\ref{fig:inc4096_r1_et_hist_v5} and~\ref{fig:inc4096_r1_hist_v5}\\ \hline
+{\tt sodb12} (plugged into {\em the upper right} power strip)  & INC4096 & 300 samples & 2017-02-13 $\sim$ 2017-02-27 & Figs.~\ref{fig:inc4096_r1_et_hist_v5} and~\ref{fig:inc4096_r1_hist_v5}\\ \hline
 \end{tabular}
@@
 	}
+	\subfigure[PT frequency on INC19 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data1/19_sec_pt_hist.eps}
+		\label{fig:inc19_hist_v4}
+	}
+	\subfigure[PT frequency on INC20 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data1/20_sec_pt_hist.eps}
+		\label{fig:inc20_hist_v4}
+	}
+	\subfigure[PT frequency on INC21.125 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data1/21_125_sec_pt_hist.eps}
+		\label{fig:inc21_125_hist_v4}
+	}
+	\caption{PT Histograms of INC16 ... INC21.125\label{fig:s9_r1_pt_hist2}}
+\end{figure}
+
+\begin{figure}[hp!]
+	\centering
 	\subfigure[PT frequency on INC32 on {\tt sodb9}]{
@@
 		\label{fig:inc32_r1_hist_v5}
+	}
+	\subfigure[PT frequency on INC60 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data1/60_0_sec_pt_hist.eps}
+		\label{fig:inc60_hist_v4}
+	}
+	\subfigure[PT frequency on INC62 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data1/62_0_sec_pt_hist.eps}
+		\label{fig:inc62_hist_v4}
 	}
@@
 	}
-	\caption{PT Histograms of INC16 ... INC64\label{fig:s9_r1_pt_hist2}}
+	\caption{PT Histograms of INC32 ... INC64\label{fig:s9_r1_pt_hist2-2}}
 \end{figure}
@@
 		\label{fig:inc128_r1_hist_v5}
+	}
+	\subfigure[PT frequency on INC224 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data1/224_sec_pt_hist.eps}
+		\label{fig:inc224_v4}
 	}
@@
 	}
-	\subfigure[PT frequency on INC1024 on {\tt sodb9}]{
-		\includegraphics[scale=0.43]{repet_data1/1024_sec_pt_hist_v5.eps}
-		\label{fig:inc1024_r1_hist_v5}
-	}
 	\caption{PT Histograms of INC256 ... INC1024~\label{fig:s9_r1_pt_hist3}}
@@
 	\centering
+	\subfigure[PT frequency on INC1024 on {\tt sodb9}]{
+		\includegraphics[scale=0.43]{repet_data1/1024_sec_pt_hist_v5.eps}
+		\label{fig:inc1024_r1_hist_v5}
+	}
 	\subfigure[PT frequency on INC2048 on {\tt sodb10}]{
@@
 	}
-	\caption{PT Histograms of INC2048 and INC4096~\label{fig:s9_r1_pt_hist4}}
+	\caption{PT Histograms of INC1024 ... INC4096~\label{fig:s9_r1_pt_hist4}}
 \end{figure}

### 2019-04-04 pt_characterization/protocol/intermediate_data/fp_incr_work.c

@@ -1,0 +1,197 @@
+#include <stdio.h>    /* standard in and output*/
+#include <sys/socket.h> /* for socket() and socket functions*/
+#include <arpa/inet.h>  /* for sockaddr_in and inet_ntoa() */
+#include <stdlib.h>   
+#include <string.h>    
+#include <unistd.h>     /* for close() */
+#include <time.h> 
+#include <limits.h> 
+
+#define PORT_NUM 2400
+#define COEF 1460000000 // 16B
+#define MAX_INCR 420000000
+
+char c;
+
+#if 1 
+inline void incr_work(float exp_secs) {
+	unsigned long long i, j, x;
+	//printf("Increment work of %d seconds starts!\n", exp_secs);
+	//max = 2147483647;
+	//max = 2000000000; 
+	double fn = (long)COEF*(float)exp_secs;
+	//printf("max iterations: %llu\n", n);
+	for (i = 1; i <= (unsigned long long)fn; i++) {
+		x = 0;
+		for (j = 0; j < UINT_MAX; j++) {
+			x++;
+		}
+	}
+	//sleep(exp_secs);
+	//printf("Done!\n");
+}
+#else
+inline void incr_work(int exp_secs) {
+        unsigned long long i, j,k, x;
+        for (i = 1; i <= (long)exp_secs; i++) {
+                x = 0;
+                for (j = 0; j < MAX_INCR; j++) {
+                        x++;
+                }
+        }
+}
+#endif
+
+void reply(char c, int connected){
+//	printf("Response signal has been sent! : %d on %c\n\n", success, c);
+	if (send(connected, "1", 1, 0) == -1) {
+		perror("send() failed");
+	}
+//	printf("Response signal has been sent! : %d on %c\n\n", success, c);
+//	success = 0;
+	close (connected);
+}
+
+int main(int argc, char *argv[]) {
+	float exp_run_time = 0;
+	int total_size = sizeof(int);
+
+	int sock, connected, bytes_received, true = 1;
+	struct sockaddr_in server_addr, client_addr;
+	int sin_size;
+	if ((sock = socket(AF_INET, SOCK_STREAM, 0)) == -1) {
+//		if ((sock = socket(AF_LOCAL, SOCK_STREAM, 0)) == -1) {
+		perror("Socket");
+		exit(1);
+	}
+printf("command socket : %d\n", sock);
+	
[…]

### 2019-04-04 pt_characterization/protocol/intermediate_data/pt_hist_sodb9.r

@@ -1,409 +1,465 @@
-###### INC11.125
-numSamples <- 1000 
-task_len <- 11 
-fraction <- 125
-dataFileName <- paste("INC",task_len,".",fraction,".dat",sep="")
-x = read.csv(file=dataFileName,head=TRUE,sep="\t")
-x <- subset(x, x$ITERNUM <= numSamples)
-setEPS() 
-histFileName <- paste(task_len,"_",fraction,"_sec_pt_hist.eps",sep="")
-postscript(histFileName)
-binsize=1
-x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
-x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
-#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
-xmin <-min(x$PRTIME)
-xmin
-xmax <-max(x$PRTIME)
-xmax
-nbins <- ceiling((xmax-xmin) / binsize)
-h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F) 
-tl <- h$counts  
-c <- 0
-for (v in tl) {
-    c <- c+1
-    tl[c:1] <- task_len+fraction/1000
+###### INC224
+numSamples <- 1000
+task_len <- 224
+fileName <- paste("inc",task_len,"_hist_bin_count.txt",sep="") 
+for (task_len in c(224)){
+	fraction <- 0
+	INC_label <- paste(task_len,sep="")
+	dataFileName <- paste(task_len,"_sec.dat",sep="")
+	histFileName <- paste(task_len,"_sec_pt_hist.eps",sep="")
+	x = read.csv(file=dataFileName,head=TRUE,sep="\t")
+	x <- subset(x, x$ITERNUM <= numSamples)
+	setEPS() 
+	postscript(histFileName)
+	binsize=1
+	x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+	x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+	#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+	xmin <-min(x$PRTIME)
+	xmin
+	xmax <-max(x$PRTIME)
+	xmax
+	nbins <- ceiling((xmax-xmin) / binsize)
[…]

### 2019-08-14 pt_characterization/protocol/interleaved/paper.tex
Directory reorganized

--- 2018-12-15 pt_characterization/protocol/intermediate_data/paper.tex
@@
 Broadly speaking, there are two {\em subjects} included in the document.
-The first subject, which is discussed in Section 1, involves {\it i}) running as a {\em batch} a series of INCs with increasing task length and {\it ii}) pausing the batch run.
+The first subject, which is discussed in Section~2, involves an experiment of {\it i}) running as a {\em batch} a series of INCs with increasing task length and {\it ii}) pausing the batch run.
 This entire batch run including the pause continues until the specified number of iterations, say 30, is specified.
@@
 
-The second subject, which is discussed in Section 2, iconcerns dividing 
-a single run of 1000 samples on INC16 into five consecutive subruns of 200 samples.  
+The second subject, which is discussed in Section~3, concerns dividing 
+the 1000 samples of an INC16 run into five consecutive subruns of 200 samples.  
 For each subrun, we examine their respective PT (program time) histograms 
-and see if the histograms look similar to each other.
+and check to see if the histograms look similar to each other.
 

### 2019-08-14 pt_characterization/protocol/intermediate_data/pt_hist_sodb9.r

@@ -1,3 +1,120 @@
+###### INC118-123
+numSamples <- 1000
+task_len <- 118
+fileName <- paste("inc",task_len,"-",task_len+2,"_hist_bin_count.txt",sep="") 
+fraction <- 0 
+count <- -1
+for (task_len in c(118,119,120,121,122,123)){
+	count <- count+1
+	fraction <- 0
+	dataFileName <- paste("INC",task_len,".",fraction,".dat",sep="")
+	INC_label <- paste(task_len,".",fraction,sep="")
+	histFileName <- paste(task_len,"_sec_pt_hist.eps",sep="")
+	print(dataFileName)
+	x = read.csv(file=dataFileName,head=TRUE,sep="\t") 
+	x <- subset(x, x$ITERNUM <= numSamples)
+	setEPS() 
+	postscript(histFileName)
+	binsize=1
+	x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
+	x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
+	#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
+	xmin <-min(x$PRTIME)
+	xmin
+	xmax <-max(x$PRTIME)
+	xmax
+	nbins <- ceiling((xmax-xmin) / binsize)
+	h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F) 
+	ymax <- max(h$counts)
+	ymax <- ceiling(ymax/100)*100 
+	#ymax <- 600 
+	plot(h, xaxt="n",axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green", 
+	main=paste("PT frequency on INC",INC_label,sep=""), 
+	sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), xlab='PT (ms)', ylab=expression('Frequency'))
+	axis(side=1,at=h$mids,labels=seq(min(h$breaks),max(h$breaks)-1,binsize))
+	dev.off()
+	tl <- h$counts  
+	c <- 0
+	for (v in tl) {
+	    c <- c+1 
+	    tl[c:1] <- task_len
+	}
+	y <- cbind(tl,floor(h$mids),h$counts) 
+	y
+	if(task_len == 118){
+		write.table(y, file=fileName, row.names = FALSE, col.names = c('INC Number', 'Bin Number', 'Counts'), sep = "\t")
+	}else{
[…]

### 2019-08-14 pt_characterization/protocol/intermediate_data/recent_data/pt_hist_sodb9.r

@@ -1,236 +1,0 @@
-###### INC8192
-numSamples <- 300
-task_len <- 8192
-fileName <- paste("inc",task_len,"_hist_bin_count.txt",sep="") 
-fraction <- 0 
-dataFileName <- paste("INC",task_len,".",fraction,".dat",sep="")
-INC_label <- paste(task_len,".",fraction,sep="")
-histFileName <- paste(task_len,"_sec_pt_hist.eps",sep="")
-print(dataFileName)
-x = read.csv(file=dataFileName,head=TRUE,sep="\t") 
-x <- subset(x, x$ITERNUM <= numSamples)
-setEPS() 
-postscript(histFileName)
-binsize=5
-x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
-x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
-#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
-xmin <-min(x$PRTIME)
-xmin
-xmax <-max(x$PRTIME)
-xmax
-nbins <- ceiling((xmax-xmin) / binsize)
-h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F) 
-ymax <- max(h$counts)
-ymax <- ceiling(ymax/100)*100 
-ymax <- 40 
-plot(h, xaxt="n",axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green", 
-main=paste("PT frequency on INC",INC_label,sep=""), 
-sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), xlab='PT (ms)', ylab=expression('Frequency'))
-axis(side=1,at=h$mids,labels=seq(min(h$breaks),max(h$breaks)-1,binsize))
-dev.off()
-tl <- h$counts  
-c <- 0
-for (v in tl) {
-    c <- c+1 
-    tl[c:1] <- task_len
-}
-y <- cbind(tl,floor(h$mids),h$counts) 
-y
-if(task_len == 8192){
-	write.table(y, file=fileName, row.names = FALSE, col.names = c('INC Number', 'Bin Number', 'Counts'), sep = "\t")
-}else{
-	write.table(y, file=fileName, append = TRUE, row.names = FALSE, col.names = FALSE, sep = "\t")
-}
-
-
-### INC72,160,192,224
-numSamples <- 1000 
-start <- 72
[…]

### 2019-08-14 pt_characterization/protocol/intermediate_data/rerun_incs/daemon_stat.r

@@ -1,75 +1,0 @@
-###### INC80 
-task_len <- "old"
-dataFileName <- paste("inc80_",task_len,"_daemon_stat.txt",sep="")
-x = read.csv(file=dataFileName,head=TRUE,sep="\t")
-xmin <- min(x$INC_PT) 
-xmin
-xmax<- max(x$INC_PT) 
-xmax
-ymin <- min(x$KSD0_PT) 
-ymax <- max(x$KSD0_PT) 
-
-task_len <- "new"
-dataFileName <- paste("inc80_",task_len,"_daemon_stat.txt",sep="")
-x = read.csv(file=dataFileName,head=TRUE,sep="\t")
-xmin <- min(x$INC_PT) 
-xmin
-xmax<- max(x$INC_PT) 
-xmax
-ymin <- min(x$KSD0_PT) 
-ymax <- max(x$KSD0_PT) 
-#plot(x$INC_PT, x$KSD0_PT,xlim=c(xmin,xmax),ylim=c(0,ymax), xlab='INC PT (ms)', ylab='kslowd000 PT (ms)')
-plot(x$INC_PT, x$KSD0_PT, xaxt='n')  
-axis(side=1,at=seq(xmin,xmax,1),labels=seq(xmin,xmax,1))
-axis(2)
-box()
-
-dev.off()
-
-
-
-
-
-numSamples <- 1000
-task_len <- 80
-fileName <- paste("inc",task_len,"_hist_bin_count.txt",sep="") 
-fraction <- 0 
-dataFileName <- paste("INC",task_len,".",fraction,".dat",sep="")
-INC_label <- paste(task_len,".",fraction,sep="")
-histFileName <- paste(task_len,"_sec_pt_hist.eps",sep="")
-print(dataFileName)
-x = read.csv(file=dataFileName,head=TRUE,sep="\t") 
-x <- subset(x, x$ITERNUM <= numSamples)
-setEPS() 
-postscript(histFileName)
-binsize=1
-x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
-x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
-#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
-xmin <-min(x$PRTIME)
-xmin
-xmax <-max(x$PRTIME)
-xmax
-nbins <- ceiling((xmax-xmin) / binsize)
-h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F) 
[…]

### 2019-08-14 pt_characterization/protocol/intermediate_data/rerun_incs/pt_hist_sodb9.r

@@ -1,2236 +1,0 @@
-###### INC1-64
-numSamples <- 1000
-fileName <- paste("inc1-64_hist_bin_count.txt",sep="") 
-fraction <- 0 
-for (task_len in c(1,2,4,8,16,32,64)){
-	fraction <- 0
-	dataFileName <- paste("INC",task_len,".",fraction,".dat",sep="")
-	INC_label <- paste(task_len,".",fraction,sep="")
-	histFileName <- paste(task_len,"_sec_pt_hist.eps",sep="")
-	print(dataFileName)
-	x = read.csv(file=dataFileName,head=TRUE,sep="\t") 
-	x <- subset(x, x$ITERNUM <= numSamples)
-	setEPS() 
-	postscript(histFileName)
-	binsize=1
-	x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
-	x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
-	#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
-	xmin <-min(x$PRTIME)
-	xmin
-	xmax <-max(x$PRTIME)
-	xmax
-	nbins <- ceiling((xmax-xmin) / binsize)
-	h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F) 
-	ymax <- max(h$counts)
-	ymax <- ceiling(ymax/100)*100 
-	#ymax <- 600 
-	plot(h, xaxt="n",axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green", 
-	main=paste("PT frequency on INC",INC_label,sep=""), 
-	sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), xlab='PT (ms)', ylab=expression('Frequency'))
-	axis(side=1,at=h$mids,labels=seq(min(h$breaks),max(h$breaks)-1,binsize))
-	dev.off()
-	tl <- h$counts  
-	c <- 0
-	for (v in tl) {
-	    c <- c+1 
-	    tl[c:1] <- task_len
-	}
-	y <- cbind(tl,floor(h$mids),h$counts) 
-	y
-	if(task_len == 1){
-		write.table(y, file=fileName, row.names = FALSE, col.names = c('INC Number', 'Bin Number', 'Counts'), sep = "\t")
-	}else{
[…]

### 2024-09-13 pt_characterization/paper2.tex
data cleansing

\documentclass[10pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{epsfig}
\usepackage{float}
\usepackage{subfigure}
\usepackage{algorithm}
\usepackage{algorithmic}
\def\ert{\mbox{{\it expRunTime}}}
\def\sertput{\mbox{{\it sendExpRunTimeToPUT}}}
\def\rpe{\mbox{{\it recordPUTExecution}}}
\def\procDiff{\mbox{{\it procDiff}}}
\def\getProcDiff{\mbox{{\it getProcDiff}}}
\def\bfps{\mbox{{\it beforeProcs}}}
\def\afps{\mbox{{\it afterProcs}}}
\def\gpi{\mbox{{\it getProcInfo}}}
\def\ect{\mbox{{\it elapsedClockTime}}}
\def\pt{\mbox{{\it ProgramTime}}}
\def\cpt{\mbox{{\it computeProgramTime}}}
\long\def\shorten#1{}
\begin{document}
\title{Characterization of Program Execution Time Distribution}

\author{
Young-Kyoon Suh\\
Department of Computer Science \\
University of Arizona \\
\small\tt \{yksuh\}@cs.arizona.edu
}
\maketitle

\section{Observed Phenomena}
In this section, we discuss interesting phenomena observed from our measurements on 
a program-under-test (PUT) with a given task length. 
PUT is a simple compute-only program that runs a nested-for loop for a given amount of time (task length) and terminates.

%For instance, PUT8 denotes PUT taking 8 seconds as its task length. 
%In this report, we use two types of program execution time: elapsed time (ET) and program time (PT). 
%ET is computed as the difference of the wall-clock times measured before and after PUT execution.
%PT is calculated as the sum of the user time and system time of PUT.

\subsection{Different Bands in Elapsed Time}
\label{sec:bands}

We introduce an interesting phenomenon regarding the elapsed time (ET) of PUT.
ET is computed as the difference of the wall-clock times measured before and after PUT execution.

\paragraph{Presumption} Little structure in ET

\paragraph{Description} Figure~\ref{fig:et_plot} shows the distribution of ET on PUT16 and PUT32. 
In this report, PUT$x$ denotes the PUT with $x$-second task length.
We observe from these data that some or many of ET records were concentrated within a certain interval, 
called {\em band} (or layer). The used data were obtained by EMPv2.

\begin{figure}[hp!]
	\centering
	\subfigure[Per-execution ET on PUT16]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/16_sec_et_all.eps}
		\label{fig:16_sec_ect}
	}
	\subfigure[Per-execution ET on PUT32]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/32_sec_et_all.eps}
		\label{fig:32_sec_ect}
	}
	\caption{Per-execution ET on PUT16 and PUT32~\label{fig:et_plot}}
\end{figure}

\paragraph{Analysis} As illustrated in Figure~\ref{fig:16_sec_ect}, 
there was the compact band formed around 16,650 msecs in the $y$-axis corresponding to most of the PUT executions. 
Another {\it weak} band around 16,900 msecs was formed as well. 
Furthermore, Figure~\ref{fig:32_sec_ect} demonstrates this similar occurrence observed in PUT32. 
Most of the $y$-axis values were clustered around 33,250 msecs. 
There were also other higher bands consisting of around 33,400 or 33,500 msecs. 
For each experiment we indeed were able to see an intensively-clustered band (formed by around the median ET measurements) 
and a few bands weakly formed above and below the main band. 
%In particular, Y-axis values belonging to a higher band than the main band seem periodic.

\paragraph{Hypothesis} ``{\it A high band seems formed 
due to daemon processes that periodically came up and ran for a moment, thus affecting 
ET at certain executions.}''

\paragraph{Test} 
When we looked at the data in PUT32, a raw ET value ({\it e.g}, 33,500 msecs) 
exceeding its median appeared at the 547th, 573th, 600th, and 626th. 
The same ET value was seen at the 731th, 758th, 916th, 942th, 968th, and 
995th. 
In other words, every 26 or 27 executions ET was higher than the median.
In PUT16, the measured high ET value was also observed every 51 executions. 
Based on these investigations, 
relatively more executed processes were included in the high ET measurements. 
Considering there were no other processes except PUT, 
we infer that some daemon processes must have contributed to forming such a high band. 
This hypothesis, therefore, is supported.

\paragraph{Status} This cause is understood.

\clearpage
\newpage

\subsection{Periodic Dips in PT}\label{sec:pd_pt}

We now turn our discussion to several phenomena regarding PT. 
PT is calculated as the sum of the user time and system time of PUT.

\paragraph{Presumption} PT is totally flat.

\paragraph{Description} 
In the measured data by EMPv2, 
we observed periodically-appearing dips. 
Figure~\ref{fig:pt_all} exhibits the distribution of PT in the experiments 
where the assigned task lengths were one, two, and four seconds.

\begin{figure}[H]
	\centering
	\subfigure[Per-execution PT on PUT1]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/1_sec_pt_all.eps}
		\label{fig:1_sec_pt_all}
	}
	\subfigure[Per-execution PT on PUT2]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/2_sec_pt_all.eps}
		\label{fig:2_sec_pt_all}
	}
	\subfigure[Per-execution PT on PUT4]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/4_sec_pt_all.eps}
		\label{fig:4_sec_pt_all}
	}
	\caption{Per-execution PT on PUT1, PUT2, and PUT4~\label{fig:pt_all}}
\end{figure}

\paragraph{Analysis}
At first, the calculated PT values were plotted around the median. 
At some point, however, the PT values dropped to the minimum, 
and then the values bounced back up to the median. 
This pattern repeated after a certain number of executions; 
namely, this dip seemed to occur every around $x$ seconds after the first appearance. 
Specifically, the dip first appeared after the 200th execution (about 200 seconds) and 
reappeared every approximately 400 seconds, as shown in Figure~\ref{fig:1_sec_pt_all}.
In PUT2, these dips first emerged around the 200th execution, 
and they reoccurred after around 800 seconds, as illustrated in Figure~\ref{fig:2_sec_pt_all}.
In Figure~\ref{fig:4_sec_pt_all}, 
the first dip was observed at around the 100th execution and repeated after approximately 800 seconds. 
Note that the $x$-axis values represents execution number, rather than actual time.
The same execution number across different tasks do not mean that the respective actual times are same. 
The 200th execution in PUT2 and PUT4 could happen around 400 seconds and 800 seconds, respectively, after the tasks start.
%Although weak, the dips seem to occur every around 400 seconds even in the eight-second task.

\paragraph{Hypothesis I} ``{\it These dips are caused by adjustments by NTP (Network Time Protocol)~\cite{Mills} used to 
synchronize its local system time to a server in the network.}''

This hypothesis is based on the assumption that PT will be faster whenever the system time is corrected 
by the NTP daemon ({\tt ntpd}).

\paragraph{Test} To confirm this, we ran some additional experiments in which {\tt ntpd} was disabled.
In these experiments, EMPv1 was used.

The experiment results are shown in Figure~\ref{fig:ntpd_off_pt}. 
Unfortunately, the dips did not go away. 
Conversely, overall dips became thinner and appeared more frequently. 
For instance, the dips shown in Figure~\ref{fig:ntp_off_1_sec_pt_all} were weaker and more frequent, 
compared with the ones in Figure~\ref{fig:1_sec_pt_all}. 
The same trend was seen in Figures~\ref{fig:ntp_off_2_sec_pt_all} and~\ref{fig:ntp_off_4_sec_pt_all} although less strong and somewhat seemingly random. 
In addition, the overall distribution of PT looked more dispersed, weakening the compactness of the bands.

\begin{figure}[H]
	\centering
	\subfigure[Per-execution PT on PUT1]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-off-turbo-on/1_sec_pt_all.eps}
		\label{fig:ntp_off_1_sec_pt_all}
	}
	\subfigure[Per-execution PT on PUT2]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-off-turbo-on/2_sec_pt_all.eps}
		\label{fig:ntp_off_2_sec_pt_all}
	}
	\subfigure[Per-execution PT on PUT4]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-off-turbo-on/4_sec_pt_all.eps}
		\label{fig:ntp_off_4_sec_pt_all}
	}
	\caption{Per-execution PT on PUT1, PUT2, and PUT4 when {\tt ntpd} was turned off~\label{fig:ntpd_off_pt}}
\end{figure}

Turning off {\tt ntpd} did not help us to figure out the cause of the dips. 
Our hypothesis, thus, is not supported.

\paragraph{Hypothesis II} ``{\it Disabling the Turbo mode set in a processor will make the dips go away.}''

Some processors support ``turbo mode'' and scale up the frequency and voltage at times. 
This hypothesis studies the impact of the frequency scaling caused by the turbo mode.

%Turbo mode dynamically increases the processor's frequency as needed by taking advantage of thermal and power headroom. 
%Speedstep is meant to reduce power usage/heat based on the load of the processor. (slow down an idle processor)
%Turbo boost increase processor speed based on load and temperature. So it will find the so called sweet spot between temperature and performance up to a certain clock speed.

\paragraph{Test} 
We found that the processor type of our experiment machine is Intel Core {\it i}7 CPU 870 @ 2.93GHz. 
This processor enabled by default the turbo mode through the feature named ``Intel$\textregistered$ Turbo Boost Technology''. 
In addition, the turbo mode can be influenced by another feature of the processor, 
called ``Enhanced Intel SpeedStep$\textregistered$ Technology'', which can reduce power usage/heat based on the load of the processor. 
To see if the Turbo mode was the cause of the dips, we disabled both features. The frequency scaling was stopped.

Figure~\ref{fig:turbo_off_pt} shows the distribution of PT on PUT1, PUT2, and PUT4. 
In these data, EMPv2 was used. 
Compared with Figure~\ref{fig:pt_all}, overall the degree of dips became much more discrete and weaker. 
In addition, the standard variations of PT on the PUT tasks reduced to 
1.65/2.78/5.75 from 2.39/4.33/7.52, respectively. 
This implies that turning off the turbo mode helped us reduce the variation of PT. 
As a result, most executions yielded flat PT measurements.
Nevertheless, the dips were still observed.

\begin{figure}[H]
	\centering
	\subfigure[Per-execution PT on PUT1]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-off/1_sec_pt_all.eps}
		\label{fig:tm_1_sec_pt_all}
	}
	\subfigure[Per-execution PT on PUT2]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-off/2_sec_pt_all.eps}
		\label{fig:tm_2_sec_pt_all}
	}
	\subfigure[Per-execution PT on PUT4]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-off/4_sec_pt_all.eps}
		\label{fig:tm_4_sec_pt_all}
	}
	\caption{Per-execution PT on PUT1, PUT2, PUT4, and PUT8 when the turbo mode was turned off~\label{fig:turbo_off_pt}}
\end{figure}

%As shown in Figures~\ref{fig:turbo_off_pt}, overall PT was flat although the dots look scattered due to resolution . 
%Specifically, the standard variations of PT in the 1/2/4-second tasks reduced to 1.65/2.78/5.75 from 2.39/4.33/7.52, respectively.

%Figure~\ref{fig:turbo_off_ect} shows the distribution of ET in the one-/two-/four-second tasks. 
%Interestingly, the 
%
%\begin{figure}[htp!]
%	\centering
%	\subfigure[Per-execution ET on the 1-second task]{
%		\includegraphics[scale=0.43]{figures/turbo_mode/1_sec_ect_all.eps}
%		\label{fig:tm_1_sec_ect_all}
%	}
%	\subfigure[Per-execution ET on the 2-second task]{
%		\includegraphics[scale=0.43]{figures/turbo_mode/2_sec_ect_all.eps}
%		\label{fig:tm_2_sec_ect_all}
%	}
%	\subfigure[Per-execution ET on the 4-second task]{
%		\includegraphics[scale=0.43]{figures/turbo_mode/4_sec_ect_all.eps}
%		\label{fig:tm_4_sec_ect_all}
%	}
%%	\subfigure[Per-execution PT on the 8-second task]{
%%		\includegraphics[scale=0.43]{figures/ntp-off/8_sec_pt_all.eps}
%%		\label{fig:ntp_off_8_sec_pt_all}
%%	}
%%	\caption{Per-execution PT on the 1/2/4/8-second task when {\tt ntpd} was turned off\label{fig:ntpd_off_pt}}
%	\caption{Per-execution ET on the 1/2/4-second task when Turbo mode was turned off\label{fig:turbo_off_ect}}
%\end{figure}
%
%They all look flat except a couple of executions.

\paragraph{Status} This cause is yet understood.

In subsequent sections, we show the experiment results after the Turbo mode was off.

\newpage

\subsection{Synchronized ET and PT}

\paragraph{Presumption} Totally correlated

\paragraph{Description} 
Figure~\ref{fig:sync_time} shows ET and PT measurements on PUT2 and PUT4. 
The common phenomenon is that most of the measured ET and PT data looked synchronized. 
The phenomenon was observed from the EMPv3 data.

\begin{figure}[H]
	\centering
	\subfigure[Per-execution ET on PUT2]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-off/2_sec_et_all.eps}
		\label{fig:2_sec_ect}
	}
	\subfigure[Per-execution PT on PUT2]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-off/2_sec_pt_all.eps}
		\label{fig:2_sec_pt_all2}
	}
	\subfigure[Per-execution ET on PUT4]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-off/4_sec_et_all.eps}
		\label{fig:4_sec_ect}
	}
	\subfigure[Per-execution PT on PUT4]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-off/4_sec_pt_all.eps}
		\label{fig:4_sec_pt_all2}
	}
	\caption{Per-execution ET/PT on PUT2 and PUT4~\label{fig:sync_time}}
\end{figure}

\paragraph{Analysis} In PUT2, if we overlap Figures~\ref{fig:2_sec_ect} and \ref{fig:2_sec_pt_all2}, 
we know that most executions have exactly the same ET and PT. 
This also holds true for PUT4, as shown in Figures~\ref{fig:4_sec_ect} and \ref{fig:4_sec_pt_all2}. 
The synchronization does not take place in some executions, where some other daemon processes had indeed run together.

%\paragraph{Hypothesis} ``{\it When the number of daemon processes is very low, ET and PT will be synchronized.}''. 
%
%\paragraph{Test} 
%For this test, we used EMPv2. 
%Figure~\ref{fig:daemon} shows the number of daemon processes observed when measuring the execution time of PUT2 and PUT4.
%In Figures~\ref{fig:2_sec_low_daemon} and \ref{fig:4_sec_low_daemon}, 
%we found that when fewer than three daemon processes (low) was present, 
%ET and PT measurements were exactly the same. 
%Figures~\ref{fig:2_sec_high_daemon} and \ref{fig:4_sec_high_daemon} exhibit the measurements with 
%more than three daemon processes (high). 
%For these measurements, both did not get synchronized; 
%namely, the measured ET values were greater than the measured PT values. 
%This hypothesis, therefore, is supported. 
%
%\begin{figure}[hp!]
%	\centering
%	\subfigure[Low number of daemon processes]{
%		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/2_sec_num_low_daemons.eps}
%		\label{fig:2_sec_low_daemon}
%	}
%	\subfigure[High number of daemon processes]{
%		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/2_sec_num_high_daemons.eps}
%		\label{fig:2_sec_high_daemon}
%	}
%	\subfigure[Low number of daemon processes]{
%		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/4_sec_num_low_daemons.eps}
%		\label{fig:4_sec_low_daemon}
%	}
%	\subfigure[High number of daemon processes]{
%		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/4_sec_num_high_daemons.eps}
%		\label{fig:4_sec_high_daemon}
%	}
%	\caption{Number of daemon processes per execution\label{fig:daemon}}
%\end{figure}
%
%\paragraph{Status} We understood this.

\newpage

\subsection{Outliers in PT}

\paragraph{Presumption} Large outlier in PT should not occur.

\paragraph{Description} 
This phenomenon indicates that a couple of PT measurements showed exceedingly high peaks.

\paragraph{Analysis} 
%In Figure~\ref{fig:pt_all}, we've already seen several outliers across different measurements.
%There is one outlier observed on PUT8, as shown in Figure~\ref{fig:1_sec_pt_all}, 
%and there are also several outliers seen in Figures~\ref{fig:1_sec_pt_all} and \ref{fig:4_sec_pt_all}. 
%Furthermore, we can see these abnormal PT in heavier tasks.

Figure~\ref{fig:peak_pt} exhibits the PT measurements on PUT16, PUT32, and PUT64. 
These data were obtained by EMPv3. 
There were three outliers observed at around the 90th and the 820th executions, as shown in Figure~\ref{fig:16_sec_pt_all}.
Likewise, a couple of outliers were also observed from the measurements on PUT32 and in PUT64, 
as illustrated in Figures~\ref{fig:32_sec_pt_all} and \ref{fig:64_sec_pt_all}.

\begin{figure}[H]
\centering
	\subfigure[Per-execution PT on PUT16]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-off/16_sec_pt_all.eps}
		\label{fig:16_sec_pt_all}
	}
	\subfigure[Per-execution PT on PUT32]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-off/32_sec_pt_all.eps}
		\label{fig:32_sec_pt_all}
	}
	\subfigure[Per-execution PT on PUT64]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-off/64_sec_pt_all.eps}
		\label{fig:64_sec_pt_all}
	}
	\caption{Per-execution PT on PUT8, PUT16, PUT32, and PUT64~\label{fig:peak_pt}}
\end{figure}

There are three common things that we can draw regarding this phenomenon. 
First of all, literally all the peaks are far above the main band. 
Second, every experiment reveals a few peaks. 
Lastly, their occurrence does not have any periodic cycle. In other words, 
the outliers are scattered all over the places. 
It looks irregular and appears randomly at a specific execution.

\paragraph{Hypothesis} ``{\it The outliers could be correlated with processes 
scheduled to run at some point during our experiment.}''\\

%\paragraph{Hypothesis} ``{\it Considering that there could be processes that can be scheduled 
%during the experiment, these peaks could be correlated to this.}''
%
We looked at the data revealing the peaks in every experiment. 
Interestingly, {\em CPU\_COUNT}, representing \# of context swaps, is positively 
correlated with peak. PT grows linearly with the increase of \# of context switches. 
The correlation coefficient of PT and the context switches is 0.67, which is quite high. 
(The correlation between PT and \# of run processes is also positive.) 
Hence, this hypothesis is supported.

\paragraph{Status} We have understood this.

\newpage

\subsection{Periodic Peaks in Frequency of PT}

\paragraph{Presumption} No peak except the median PT

\paragraph{Description} 
Figure~\ref{fig:pt_hist} shows the frequency of PT in the experiments 
whose respective expected running time is 4 and 8 seconds. 
This phenomenon indicates that a few peaks in frequency of PT are regularly seen. 
The phenomenon was observed from the EMPv2 data.

\begin{figure}[H]
	\centering
	\subfigure[PT frequency on PUT4]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/4_sec_pt_hist.eps}
		\label{fig:4_sec_pt_hist}
	}
	\subfigure[PT frequency on PUT8]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/8_sec_pt_hist.eps}
		\label{fig:8_sec_pt_hist}
	}
	\caption{PT frequency on PUT4 and PUT8~\label{fig:pt_hist}}
\end{figure}

\paragraph{Analysis} In Figure~\ref{fig:4_sec_pt_hist}, we could see two distinct peaks 
formed around 4,138 msecs and 4,142 msecs before the highest peak around 4,155 msecs. 
This pattern was also observed in Figure~\ref{fig:8_sec_pt_hist}, 
where there were about three peaks shown at 8,275 msecs, 8,292 msecs, and 8,300 msecs 
before the highest one around 8,318 msecs. 
The peaks appeared to repeat every around 8 msecs. 
This periodicty seems formed by the same PT values on PUT4 and PUT8. 
In particular, these PT values were below the median. 
The low PT values were calculated when PT periodically went down to the minimum, 
as discussed in Section~\ref{sec:pd_pt}. 
As a consequence, periodic peaks in frequency were observed, as illustrated in Figure~\ref{fig:pt_hist}.

\paragraph{Hypothesis} ``{\it Heavier PUT will also have similar periodic peaks in frequency due to their low PT values.}''

\paragraph{Test} 
Such similar peaks in frequency periodically 
appeared in the measurements on PUT16 and PUT32, as depicted in Figure~\ref{fig:extra_pt_hist}. 
Again, EMPv2 was used for this test.

\begin{figure}[H]
	\centering
	\subfigure[PT frequency on PUT16]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/16_sec_pt_hist.eps}
		\label{fig:16_sec_pt_hist}
	}
	\subfigure[PT frequency on PUT32]{
		\includegraphics[scale=0.43]{figures/sodb8-ntp-on-turbo-on/32_sec_pt_hist.eps}
		\label{fig:32_sec_pt_hist}
	}
	\caption{PT frequency on PUT16 and PUT32~\label{fig:extra_pt_hist}}
\end{figure}

Therefore, this hypothesis is supported.

\paragraph{Status} The cause is understood.

\newpage

\subsection{Increasing Minimum and Maximum Difference in PT}
%Figure~\ref{fig:mm_pt_dist} shows the distance between the minimum and maximum PT 
%over increasing amount of work.

\paragraph{Presumption} Almost zero or very small constant

\paragraph{Description} 
This phenomenon indicates that the difference between the maximum and minimum of the measured PT values 
grows linearly over increasing task length, as shown in Figure~\ref{fig:mm_pt_dist}. 
Note that $xy$ axes are in log scale. 
This phenomenon was observed from the EMPv2 data.

% 60 =  1.778 => 0.058
% 89 = 1.9493 => 0.043
% 106 = 2.02  => 0.025
% 121 = 2.08  => 0.0146
% 195 = 2.29  => 0.0117
% 770 = 2.89  => 0.023
% 1038 = 3.01 => 0.0156

\begin{figure}[H]
\centering
\includegraphics[scale=0.8]{figures/sodb8-ntp-on-turbo-on/overall_pt_diff.eps}
\caption{Increasing difference over increasing task length~\label{fig:mm_pt_dist}}
\end{figure}

\paragraph{Analysis} As task length increases, the PT difference increases. 
Considering that the $y$-axis values are plotted in log scale, the difference grows linearly. 
It appears to be synchronized with the degree of increasing task length. 
Our expectation on PT was flat, and thus, the increasing difference was strange.

\paragraph{Hypothesis} ?

\paragraph{Test} ?

%\paragraph{Hypothesis} ``{\it Disabling the Turbo mode will show the flat PT difference.}''
%
%\paragraph{Test} Figure~\ref{fig:off_turbo_pt_dist} shows the PT difference when the Turbo mode was off.
%
%\begin{figure}[hp!]
%\centering
%\includegraphics[scale=0.9]{figures/turbo_mode/overall_pt_diff.eps}
%\caption{PT difference over increasing expected running time under the disabled Turbo mode\label{fig:off_turbo_pt_dist}}
%\end{figure}
%
%Unfortunately, overall the PT difference still grows over increasing expected running time even 
%after the mode was not enabled. It went down at the four-second task, compared to that of the two-second task. 
%However, the increase is still shown.

%\paragraph{Hypothesis} ``{\it Even if we increase the expected running time of PUT,
%the difference in PT will be constant.}''
%
%Based on our expectation that PT should be flat, we anticipated that 
%the minimum and maximum PT through 1000 executions did not differ much, regardless of how long PUT took. 
%However, we observed that the difference got bigger over increasing the degree of the task. 
%For instance, the difference started with 60 msecs in the one-second task. 
%As the expected running times increase, the difference also grows somewhat linearly. 
%The difference reaches up to around 1040 msecs in the sixty-four second work. 
%Thus, the hypothesis is not supported. 
%
\paragraph{Status} Not understood.
%
%\subsection{Miscellaneous}
%
%\begin{enumerate}
%\item Note that the X-axis in th
%\end{enumerate}
%
%
\vspace{-0.01in}

\newpage

\section{Exploring a Proper Distribution for Measured Data}

Figures~\ref{fig:extra_pt_hist1},~\ref{fig:extra_pt_hist2},~\ref{fig:extra_pt_hist3},~\ref{fig:extra_pt_hist4},~\ref{fig:extra_pt_hist5},~\ref{fig:extra_pt_hist6}, 
and~\ref{fig:extra_pt_hist7} represent the histograms of ET and PT on the 1 $\sim$ 4096 second tasks. 
Nnormal and beta distributions are fitted on the histograms as well.
For these histograms, we used EMPv5, which identified and removed outliers in its measurements.

\begin{figure}[H]
	\centering
	\subfigure[ET frequency on PUT1]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/1_sec_et_hist.eps}
		\label{fig:1_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT1]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/1_sec_pt_hist.eps}
		\label{fig:1_sec_pt_hist1}
	}
	\subfigure[ET frequency on PUT2]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/2_sec_et_hist.eps}
		\label{fig:2_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT2]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/2_sec_pt_hist.eps}
		\label{fig:2_sec_pt_hist1}
	}
	\caption{Normal and beta distributions fitting onto the histograms of the measurements of PUT1 and PUT2~\label{fig:extra_pt_hist1}}
\end{figure}

\begin{figure}[H]
	\centering
	\subfigure[ET frequency on PUT4]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/4_sec_et_hist.eps}
		\label{fig:4_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT4]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/4_sec_pt_hist.eps}
		\label{fig:4_sec_pt_hist1}
	}
	\subfigure[ET frequency on PUT8]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/8_sec_et_hist.eps}
		\label{fig:8_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT8]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/8_sec_pt_hist.eps}
		\label{fig:8_sec_pt_hist1}
	}
	\caption{Normal and beta distributions fitting onto the histograms of the measurements of PUT4 and PUT8~\label{fig:extra_pt_hist2}}
\end{figure}

\begin{figure}[H]
	\centering
	\subfigure[ET frequency on PUT16]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/16_sec_et_hist.eps}
		\label{fig:16_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT16]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/16_sec_pt_hist.eps}
		\label{fig:16_sec_pt_hist1}
	}
	\subfigure[ET frequency on PUT32]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/32_sec_et_hist.eps}
		\label{fig:32_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT32]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/32_sec_pt_hist.eps}
		\label{fig:32_sec_pt_hist1}
	}
	\caption{Normal and beta distributions fitting onto the histograms of the measurements of PUT16 and PUT32~\label{fig:extra_pt_hist3}}
\end{figure}

\begin{figure}
	\centering
	\subfigure[ET frequency on PUT64]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/64_sec_et_hist.eps}
		\label{fig:64_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT64]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/64_sec_pt_hist.eps}
		\label{fig:64_sec_pt_hist1}
	}
	\subfigure[ET frequency on PUT128]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/128_sec_et_hist.eps}
		\label{fig:128_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT128]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/128_sec_pt_hist.eps}
		\label{fig:128_sec_pt_hist1}
	}
	\caption{Normal and beta distributions fitting onto the histograms of the measurements of PUT64 and PUT128~\label{fig:extra_pt_hist4}}
\end{figure}

\begin{figure}[H]
	\centering
	\subfigure[ET frequency on PUT256]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/256_sec_et_hist.eps}
		\label{fig:256_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT256]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/256_sec_pt_hist.eps}
		\label{fig:256_sec_pt_hist1}
	}
	\subfigure[ET frequency on PUT512]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/512_sec_et_hist.eps}
		\label{fig:512_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT512]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/512_sec_pt_hist.eps}
		\label{fig:512_sec_pt_hist1}
	}
	\caption{Normal and beta distributions fitting onto the histograms of the measurements of PUT256 and PUT512~\label{fig:extra_pt_hist5}}
\end{figure}

\begin{figure}[H]
	\centering
	\subfigure[ET frequency on PUT1024]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/1024_sec_et_hist.eps}
		\label{fig:1024_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT1024]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/1024_sec_pt_hist.eps}
		\label{fig:1024_sec_pt_hist1}
	}
	\subfigure[ET frequency on PUT2048]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/2048_sec_et_hist.eps}
		\label{fig:2048_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT2048]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/2048_sec_pt_hist.eps}
		\label{fig:2048_sec_pt_hist1}
	}
	\caption{Normal and beta distributions fitting onto the histograms of the measurements of PUT1024 and PUT2048~\label{fig:extra_pt_hist6}}
\end{figure}

\begin{figure}[H]
	\centering
		\subfigure[ET frequency on PUT4096]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/4096_sec_et_hist.eps}
		\label{fig:4096_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT4096]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/4096_sec_pt_hist.eps}
		\label{fig:4096_sec_pt_hist1}
	}
	\caption{Normal and beta distributions fitting onto the histograms of the measurements of PUT4096~\label{fig:extra_pt_hist7}}
\end{figure}

\begin{figure}[H]
	\centering
		\subfigure[ET frequency on PUT8192]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/8192_sec_et_hist.eps}
		\label{fig:8192_sec_et_hist1}
	}
	\subfigure[PT frequency on PUT8192]{
		\includegraphics[scale=0.43]{figures/sodb12-ntp-on-turbo-off/8192_sec_pt_hist.eps}
		\label{fig:8192_sec_pt_hist1}
	}
	\caption{Normal and beta distributions fitting onto the histograms of the measurements of PUT8192~\label{fig:extra_pt_hist8}}
\end{figure}

\bibliographystyle{abbrv}
\newcommand{\etalchar}[1]{$^{#1}$}
\begin{thebibliography}{99}
\bibitem{Mills} David L. Mills, ``Internet Time Synchronization: The Network Time Protocol,'' in {\it IEEE Trans. on Comm.}, Vol. 39, No. 10, pp.~1482--1493, October, 1991.
\end{thebibliography}
\end{document}

### 2024-09-13 pt_characterization/protocol/intermediate_data/rerun_incs/pt_hist_sodb9_v2.r

@@ -1,48 +1,0 @@
-###### INC1-64
-fileName <- paste("inc72-224_hist_bin_count.txt",sep="") 
-fraction <- 0 
-for (task_len in c(72,160,192,224)){
-	numSamples <- 1000
-	if(task_len == 
-
-	fraction <- 0
-	dataFileName <- paste("INC",task_len,".",fraction,".dat",sep="")
-	INC_label <- paste(task_len,".",fraction,sep="")
-	histFileName <- paste(task_len,"_sec_pt_hist.eps",sep="")
-	print(dataFileName)
-	x = read.csv(file=dataFileName,head=TRUE,sep="\t") 
-	x <- subset(x, x$ITERNUM <= numSamples)
-	setEPS() 
-	postscript(histFileName)
-	binsize=1
-	x_up = mean(x$PRTIME) + 2*sd(x$PRTIME)
-	x_dn = mean(x$PRTIME) - 2*sd(x$PRTIME)
-	#x = subset(x, x$PRTIME >= x_dn & x$PRTIME <= x_up)
-	xmin <-min(x$PRTIME)
-	xmin
-	xmax <-max(x$PRTIME)
-	xmax
-	nbins <- ceiling((xmax-xmin) / binsize)
-	h = hist(x$PRTIME, right=F, breaks=seq(xmin,xmax+binsize,binsize),plot=F) 
-	ymax <- max(h$counts)
-	ymax <- ceiling(ymax/100)*100 
-	#ymax <- 600 
-	plot(h, xaxt="n",axes = TRUE,freq=TRUE,ylim=c(0,ymax), col="green", 
-	main=paste("PT frequency on INC",INC_label,sep=""), 
-	sub=paste("(n=",nrow(x),", bin_size=",binsize,"ms)",sep=""), xlab='PT (ms)', ylab=expression('Frequency'))
-	axis(side=1,at=h$mids,labels=seq(min(h$breaks),max(h$breaks)-1,binsize))
-	dev.off()
-	tl <- h$counts  
-	c <- 0
-	for (v in tl) {
-	    c <- c+1 
-	    tl[c:1] <- task_len
-	}
-	y <- cbind(tl,floor(h$mids),h$counts) 
-	y
-	if(task_len == 1){
-		write.table(y, file=fileName, row.names = FALSE, col.names = c('INC Number', 'Bin Number', 'Counts'), sep = "\t")
-	}else{
[…]
